### 2020-07-28 tex/gitlinks.tex
better spot param; added tex

\newcommand{\codelink}[1]{\href{https://github.com/user/repo/blob/fe12fa16a5cc76c1ba72d97241fb9545da144edf/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/user/repo/blob/fe12fa16a5cc76c1ba72d97241fb9545da144edf/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/user/repo/blob/fe12fa16a5cc76c1ba72d97241fb9545da144edf/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/user/repo/_build}{#1}}

### 2020-07-28 tex/ms.tex
better spot param; added tex

% !TeX root = ./ms.tex
\documentclass[modern]{aastex62}

% Load the corTeX style definitions
\input{cortex}

% Load custom style
\input{style}

% Bibliography
\bibliographystyle{aasjournal}

\usepackage{etoolbox}
\makeatletter % we need to patch \env@cases that has @ in its name
\patchcmd{\env@cases}{\quad}{\qquad\qquad}{}{}
\makeatother

\usepackage{enumitem}

% Begin!
\begin{document}

% Title
\title{%
    \textbf{
        Interpretable Gaussian Processes for Stellar Light Curves
    }
}

% Author list
\author[0000-0002-0296-3826]{Rodrigo Luger}\altaffiliation{Flatiron Fellow}
\email{rluger@flatironinstitute.org}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY}
\affil{Virtual~Planetary~Laboratory, University~of~Washington, Seattle, WA}
%

\keywords{methods: analytic}

% \begin{abstract}
%     Abstract here.
%     %
%     \href{https://github.com/rodluger/starry_process}{\color{linkcolor}\faGithub}
% \end{abstract}

%
\section{The Spot Expansion}
\label{sec:spot}
%
We adopt the following expression for the spherical harmonic coefficient
of degree $l$ and order $m$ in the expansion of a spot at
$\theta = \varphi = 0$:
%
\begin{align}
    \label{eq:ylm0}
    y_{lm}(\delta, r, \theta = 0, \varphi = 0) =
    \begin{cases}
        1 - \dfrac{\delta c r}{2 (1 + c r)}
         & l = m = 0    \\[2em]
        -\dfrac{\delta c r \left( 2 + c r \right)}
        {2 \sqrt{2l + 1} (1 + c r)^{l + 1}}
         & l > 0, m = 0 \\[2em]
        0
         & m \ne 0
    \end{cases}
\end{align}
%
where $\delta \in [-\infty, 1]$ is the fractional decrease in the brightness
at the center of the spot and $r \in [0, 1]$ is a normalized spot radius.
The quantity $c$ is a normalization constant for the radius (see below).

The expression in Equation~(\ref{eq:ylm0}) is convenient because it satisfies
three important properties:
%
\begin{enumerate}[itemsep=2pt,parsep=1pt,label=\textbf{\arabic*}]
    \item The surface intensity monotonically increases away from the spot center
    \item The surface intensity at the spot center is $1 - \delta$
    \item The surface intensity at the antipode of the spot center is unity
\end{enumerate}
%
These properties may be demonstrated by considering the expression for the
surface intensity at a given point $(\theta, \varphi)$:
%
\begin{align}
    I(\theta, \varphi)
     & =
    \sum\limits_{l=0}^\infty \sum\limits_{m=-l}^l
    y_{lm} Y_{lm}(\theta, \varphi) \\
     & = \sum\limits_{l=0}^\infty
    y_{l0} \sqrt{2l + 1} P_l(\cos\theta)
\end{align}
%
where $P_l$ is the Legendre polynomial of degree $l$ and we have implicitly
assumed a normalization such that the integral of our expansion over
the unit sphere is $4\pi$.
Combining this with Equation~(\ref{eq:ylm0}) and rearranging, we may write
%
\begin{align}
    \label{eq:Igen}
    I(\theta, \varphi) =
    1 + \dfrac{\delta c r}{2}
    -
    \dfrac{\delta c r \left( 2 + c r \right)}{2 (1 + c r)}
    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + c r}\right)^l P_l(\cos\theta)
    \quad.
\end{align}
%
The summation in Equation~(\ref{eq:Igen}) has a closed-form expression in
terms of the generating function of the Legendre polynomials:
%
\begin{align}
    \label{eq:gen}
    \sum\limits_{l=0}^\infty t^l P_l(\cos\theta) = \frac{1}{\sqrt{1 + t^2 - 2 t \cos\theta}}
    \quad,
\end{align}
%
so we may express the intensity in the fairly simple form
%
\begin{align}
    \label{eq:Ifinal}
    I(\theta, \varphi) = A - \frac{B}{\sqrt{C - \cos\theta}}
    \quad,
\end{align}
%
where
%
\begin{align}
    A & = 1 + \dfrac{\delta c r}{2}                      \\
    B & = \delta c r (2 + cr) \sqrt{\dfrac{1}{8 + 8 cr}} \\
    C & = \dfrac{1 + (1 + c r)^2}{2 + 2 c r}
\end{align}
%
are positive constants.

Differentiating Equation~(\ref{eq:Ifinal}) with respect to $\theta$, we have
%
\begin{align}
    \label{eq:Ideriv}
    \dfrac{\mathrm{d}I(\theta, \varphi)}{\mathrm{d}\theta} & =
    -\frac{B\sin\theta}{2(C - \cos\theta)^\frac{3}{2}}
    \quad,
\end{align}
%
which is zero only for $\theta = 0$ (for which $I(\theta, \varphi)$ is
minimized) and $\theta = \pi$ (for which it is maximized). The intensity
therefore increases monotonically from the spot center to the antipode,
as stated in \textbf{1}. The value at the minimum is
%
\begin{align}
    I_{\mathrm{min}} & = A - \dfrac{B}{\sqrt{C - 1}} \\
                     & = 1 - \delta
    \quad,
\end{align}
%
as stated in \textbf{2}, and the value at the maximum is
%
\begin{align}
    I_{\mathrm{max}} & = A - \dfrac{B}{\sqrt{C + 1}} \\
                     & = 1
    \quad,
\end{align}
%
as stated in \textbf{3}.

% At the spot center, $\theta = 0$, so $\cos\theta = 1$. Since $P_l(1) = 1$ for
% all $l$, we have
% %
% \begin{align}
%     I(0)
%      & =
%     \sum\limits_{l=0}^\infty y_{l0} \sqrt{2l + 1} \\
%      & =
%     1 - \dfrac{\delta c r}{2 (1 + c r)}
%     -
%     \dfrac{\delta c r (2 + c r)}{2 (1 + c r)}
%     \sum\limits_{l=1}^\infty
%     \dfrac{1}
%     {(1 + c r)^{l}}                               \\
%      & =
%     1 -
%     \delta
%     \left[
%         \dfrac{c r}{2 (1 + c r)}
%         - \dfrac{2 + c r}{2(1 + c r)}
%         \right]                                   \\
%      & = 1 - \delta
%     \quad,
% \end{align}
% %
% as stated in \textbf{2}, where we made use of the fact that
% %
% \begin{align}
%     \sum\limits_{l=1}^\infty
%     \dfrac{1}
%     {(1 + x)^{l}} = \frac{1}{x}
%     \quad.
% \end{align}
% %
% Finally, at the antipode, $\theta = \pi$,
% $\cos\theta = -1$, and $P_l(\cos\theta) = (-1)^l$, so
% %
% \begin{align}
%     I(\pi)
%      & =
%     \sum\limits_{l=0}^\infty y_{l0} (-1)^l \sqrt{2l + 1} \\
%      & =
%     1 - \dfrac{\delta c r}{2 (1 + c r)}
%     -
%     \dfrac{\delta c r (2 + c r)}{2 (1 + c r)}
%     \sum\limits_{l=1}^\infty
%     \dfrac{(-1)^l}
%     {(1 + c r)^{l}}                                      \\
%      & =
%     1 - \dfrac{\delta c r}{2 (1 + c r)}
%     +
%     \dfrac{\delta c r}{2 (1 + c r)}                      \\
%      & = 1
%     \quad,
% \end{align}
% as stated in \textbf{3}, where we made use of the fact that
% %
% \begin{align}
%     \sum\limits_{l=1}^\infty
%     \dfrac{(-1)^l}
%     {(1 + x)^{l}} = -\frac{1}{2 + x}
%     \quad.
% \end{align}
%
%\bibliography

\end{document}

### 2020-07-28 tests/test_spot.py

@@ -8,15 +8,15 @@ from scipy.stats import lognorm
 from tqdm import tqdm
 
 
-def test_first_moment(ydeg=5, mu_s=0.1, nu_s=0.01, mu_a=-3.0, nu_a=1.0):
+def test_first_moment(ydeg=5, mu_r=0.1, nu_r=0.01, mu_d=-0.25, nu_d=1.0):
 
     np.random.seed(0)
-    nsamples = 10000000
-    atol = 1e-4
+    nsamples = 1000000
+    atol = 2e-4
 
     # Get analytic integral
     S = SpotIntegral(ydeg)
-    S.set_params(mu_s, nu_s, mu_a, nu_a)
+    S.set_params(mu_r, nu_r, mu_d, nu_d)
     mu = S.first_moment()
 
     # Integrate numerically
@@ -24,30 +24,29 @@ def test_first_moment(ydeg=5, mu_s=0.1, nu_s=0.01, mu_a=-3.0, nu_a=1.0):
     y = np.zeros((N, nsamples))
 
     # Draw the spot size
-    alpha_s, beta_s = get_alpha_beta(mu_s, nu_s)
-    s = Beta.rvs(alpha_s, beta_s, size=nsamples)
+    alpha_r, beta_r = get_alpha_beta(mu_r, nu_r)
+    r = Beta.rvs(alpha_r, beta_r, size=nsamples)
 
     # Draw the spot amplitude
-    a = lognorm.rvs(s=np.sqrt(nu_a), scale=np.exp(mu_a), size=nsamples)
+    delta = 1 - lognorm.rvs(s=np.sqrt(nu_d), scale=np.exp(mu_d), size=nsamples)
 
     # Compute the spot expansions
-    for l in tqdm(range(1, ydeg + 1)):
-        y[l * (l + 1)] = -a * (1 + s) ** (-(l ** 2))
-    mu_num = np.mean(y, axis=1).reshape(-1)
+    s = spot(ydeg, r, delta, c=1.0)
+    mu_num = np.mean(s, axis=0)
 
     # Compare
     assert np.allclose(mu, mu_num, atol=atol)
 
 
-def test_second_moment(ydeg=5, mu_s=0.1, nu_s=0.01, mu_a=-3.0, nu_a=1.0):
+def test_second_moment(ydeg=5, mu_r=0.1, nu_r=0.01, mu_d=-0.25, nu_d=1.0):
 
     np.random.seed(0)
-    nsamples = 10000000
[…]

### 2020-07-28 tex/gen_links.py

@@ -1,0 +1,27 @@
+from __future__ import print_function
+import subprocess
+import os
+
+# Generate the github links
+hash = subprocess.check_output(["git", "rev-parse", "HEAD"]).decode("utf-8")[:-1]
+slug = "rodluger/starry_process"
+with open("gitlinks.tex", "w") as f:
+    print(
+        r"\newcommand{\codelink}[1]{\href{https://github.com/%s/blob/%s/tex/figures/#1.py}{\codeicon}\,\,}"
+        % (slug, hash),
+        file=f,
+    )
+    print(
+        r"\newcommand{\animlink}[1]{\href{https://github.com/%s/blob/%s/tex/figures/#1.gif}{\animicon}\,\,}"
+        % (slug, hash),
+        file=f,
+    )
+    print(
+        r"\newcommand{\prooflink}[1]{\href{https://github.com/%s/blob/%s/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}"
+        % (slug, hash),
+        file=f,
+    )
+    print(
+        r"\newcommand{\cilink}[1]{\href{https://dev.azure.com/%s/_build}{#1}}" % (slug),
+        file=f,
+    )

### 2020-07-29 tex/ms.tex
more text

@@
 %
+\subsection{Preliminaries}
+%
 We adopt the following expression for the spherical harmonic coefficient
@@
     \label{eq:ylm0}
-    y_{lm}(\delta, r, \theta = 0, \varphi = 0) =
+    y_{lm}(r, \delta) =
     \begin{cases}
@@
 %
-where $\delta \in [-\infty, 1]$ is the fractional decrease in the brightness
-at the center of the spot and $r \in [0, 1]$ is a normalized spot radius.
-The quantity $c$ is a normalization constant for the radius (see below).
-
+where $\delta \in [-\infty, 1]$ is the spot contrast (the fractional
+decrease in the brightness
+at the center of the spot) and $r \in (0, 1]$ is the normalized spot radius.
+The quantity $c \in (0, \infty)$ is a normalization constant for the radius
+(more details below).
 The expression in Equation~(\ref{eq:ylm0}) is convenient because it satisfies
-three important properties:
+four important properties:
 %
@@
     \item The surface intensity at the antipode of the spot center is unity
+    \item The size of the spot increases monotonically with $r$
 \end{enumerate}
@@
 \begin{align}
+    \begin{split}
         I(\theta, \varphi)
@@
         y_{l0} \sqrt{2l + 1} P_l(\cos\theta)
+    \end{split}
 \end{align}
@@
     \label{eq:Ifinal}
-    I(\theta, \varphi) = A - \frac{B}{\sqrt{C - \cos\theta}}
+    I(\theta, \varphi) & = A - \frac{B}{\sqrt{C - \cos\theta}}
     \quad,
-\end{align}
     %
-where
+    \\
+    \intertext{where}
     %
-\begin{align}
+    \begin{split}
         A & = 1 + \dfrac{\delta c r}{2}                      \\
@@
         C & = \dfrac{1 + (1 + c r)^2}{2 + 2 c r}
+    \end{split}
 \end{align}
@@
 \begin{align}
+    \begin{split}
         I_{\mathrm{min}} & = A - \dfrac{B}{\sqrt{C - 1}} \\
@@
         \quad,
+    \end{split}
 \end{align}
@@
 \begin{align}
+    \begin{split}
         I_{\mathrm{max}} & = A - \dfrac{B}{\sqrt{C + 1}} \\
@@
 as stated in \textbf{3}.
+Finally, to show \textbf{4}, let us compute the half width at half minimum
+$\Delta\theta$ of the intensity profile:
+%
+\begin{align}
+    1 - \left(A - \dfrac{B}{\sqrt{C - \cos{\Delta\theta}}}\right) =
+    \dfrac{1}{2}\delta
+\end{align}
+%
+Solving for $\Delta\theta$ yields
+%
+\begin{align}
+    \Delta\theta =
+    \cos^{-1} \left[ \dfrac{2 + 3 c r (2 + c r)}{2 (1 + c r)^3} \right]
+    \quad.
+\end{align}
+%
+Differentiation with respect to $r$ yields
+%
+\begin{align}
+    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}r} =
+    \frac{3 c^2 r \left(2 + c r\right)}{\left(1 + c r\right)^4
+        \sqrt{\frac{c^2 r^2 \left(2 + c r\right)^2 \left(1 + 2 c r\right)
+                \left(3 + 2 c r\right)}{\left(1 + c r\right)^6}}}
+    \quad,
+\end{align}
+%
+which is positive definite for all $c, r > 0$.
 
-% At the spot center, $\theta = 0$, so $\cos\theta = 1$. Since $P_l(1) = 1$ for
-% all $l$, we have
-% %
-% \begin{align}
-%     I(0)
-%      & =
-%     \sum\limits_{l=0}^\infty y_{l0} \sqrt{2l + 1} \\
-%      & =
-%     1 - \dfrac{\delta c r}{2 (1 + c r)}
-%     -
-%     \dfrac{\delta c r (2 + c r)}{2 (1 + c r)}
-%     \sum\limits_{l=1}^\infty
-%     \dfrac{1}
-%     {(1 + c r)^{l}}                               \\
-%      & =
-%     1 -
-%     \delta
-%     \left[
-%         \dfrac{c r}{2 (1 + c r)}
-%         - \dfrac{2 + c r}{2(1 + c r)}
-%         \right]                                   \\
-%      & = 1 - \delta
-%     \quad,
-% \end{align}
-% %
-% as stated in \textbf{2}, where we made use of the fact that
-% %
-% \begin{align}
-%     \sum\limits_{l=1}^\infty
-%     \dfrac{1}
-%     {(1 + x)^{l}} = \frac{1}{x}
-%     \quad.
-% \end{align}
-% %
-% Finally, at the antipode, $\theta = \pi$,
-% $\cos\theta = -1$, and $P_l(\cos\theta) = (-1)^l$, so
-% %
-% \begin{align}
-%     I(\pi)
-%      & =
-%     \sum\limits_{l=0}^\infty y_{l0} (-1)^l \sqrt{2l + 1} \\
-%      & =
-%     1 - \dfrac{\delta c r}{2 (1 + c r)}
-%     -
-%     \dfrac{\delta c r (2 + c r)}{2 (1 + c r)}
-%     \sum\limits_{l=1}^\infty
-%     \dfrac{(-1)^l}
-%     {(1 + c r)^{l}}                                      \\
-%      & =
-%     1 - \dfrac{\delta c r}{2 (1 + c r)}
-%     +
-%     \dfrac{\delta c r}{2 (1 + c r)}                      \\
-%      & = 1
-%     \quad,
-% \end{align}
-% as stated in \textbf{3}, where we made use of the fact that
-% %
-% \begin{align}
-%     \sum\limits_{l=1}^\infty
-%     \dfrac{(-1)^l}
-%     {(1 + x)^{l}} = -\frac{1}{2 + x}
-%     \quad.
-% \end{align}
+\subsection{Distributions over $r$ and $\delta$}
 %
+We model the normalized spot radius $r$ as a random variable drawn from
+a Beta distribution, whose PDF is given by
+%
+\begin{align}
+    p \big(r \, \big| \, \alpha, \beta \big)
+     & =
+    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+    r^{\alpha - 1}
+    (1 - r)^{\beta - 1}
+    \quad,
+\end{align}
+%
+where $\Gamma$ is the Gamma function and $\alpha$ and $\beta$ are
+hyperparameters characterizing the shape of the distribution.
+%
+Since we require the brghtness $b = 1 - \delta$ at the center of the spot to
+be non-negative, it is convenient to parametrize the distribution of $b$
+instead of $\delta$. We model $b$ as a random variable drawn from a log-normal
+distribution:
+%
+\begin{align}
+    p \big(b \, \big| \, \mu, \nu \big)
+     & =
+    \frac{1}{b\sqrt{2\pi\nu}}
+    \exp\left[
+        -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
+        \right]
+    \quad,
+\end{align}
+%
+where $\mu$ and $\nu$ are the mean and variance of the log-normal, respectively.
+The joint distribution over $r$ and $b$ is therefore separable,
+by construction:
+%
+\begin{align}
+    p \big( r, b \, \big| \, \alpha, \beta, \mu, \nu \big)
+     & =
+    p \big(r \, \big| \, \alpha, \beta \big)
+    \,
+    p \big(b \, \big| \, \mu, \nu \big)
+    \quad.
+\end{align}
+%
+
+\subsection{First moment}
+%
+The first moment of the distribution of $y_{lm}$ over $r$ and $b$ is
+%
+\begin{align}
+    \label{eq:Erb}
+    \begin{split}
+        \mathrm{E} \big[ y_{lm} \big] &=
+        \int\limits_0^\infty \int\limits_0^1
+        y_{lm} (r, 1 - b)
+        p \big(r, 1 - b \, \big| \, \alpha, \beta, \mu, \nu \big)
+        \mathrm{d} r
+        \,
+        \mathrm{d} b
+        \\
+        &=
+        \int\limits_0^\infty
+        \mathrm{E}_r \big[ y_{lm} \big]
+        p \big(b \, \big| \, \mu, \nu \big)
+        \mathrm{d} b
+        \quad,
+    \end{split}
+    %
+    \\
+    \intertext{where}
+    \label{eq:Er}
+    %
+    \mathrm{E}_r \big[ y_{lm} \big]
+     & =
+    \int\limits_0^1
+    y_{lm} (r, 1 - b)
+    p \big(r \, \big| \, \alpha, \beta \big)
+    \mathrm{d} r
+    \quad.
+\end{align}
+%
+Expanding Equation~(\ref{eq:Er}), we obtain
+%
+\begin{align}
+    \mathrm{E}_r \big[ y_{lm} \big]
+     & =
+    \begin{dcases}
+        \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+        \int\limits_0^1
+        \left(
+        1 - \dfrac{(1 - b) c r}{2 (1 + c r)}
+        \right)
+        r^{\alpha - 1}
+        (1 - r)^{\beta - 1}
+        \mathrm{d} r
+         &
+        \qquad
+        l = m = 0    \\[2em]
+        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+        \int\limits_0^1
+        \dfrac{(1 - b) c r \left( 2 + c r \right)r^{\alpha - 1}
+        (1 - r)^{\beta - 1}}
+        {2 \sqrt{2l + 1} (1 + c r)^{l + 1}}
+        \mathrm{d} r
+         &
+        \qquad
+        l > 0, m = 0 \\[2em]
+        0
+         &
+        \qquad m \ne 0
+        \quad.
+    \end{dcases}
+\end{align}
+%
+Thanks to the dependence of the spherical harmonic coefficients on only
+powers of $r$ and $(1 + cr)$, the integrals above may be expressed in closed
+form in terms of the hypergeometric function ${_2F_1}$:
+%
+\begin{align}
+    \mathrm{E}_r \big[ y_{lm} \big]
+     & =
+    \begin{dcases}
+        1 -
+        \dfrac{(1 - b) \alpha c}{2(\alpha + \beta)} F_l
+         &
+        \qquad
+        l = m = 0    \\[2em]
+        -\dfrac{(1 - b) \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
+        \left[
+            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
+            \right]
+         &
+        \qquad
+        l > 0, m = 0 \\[2em]
+        0
+         &
+        \qquad m \ne 0
+    \end{dcases}
+\end{align}
+%
+where we define
+%
+\begin{align}
+    \begin{split}
+        F_l & \equiv {_2F_1}(l + 1, \alpha + 1; \alpha + \beta + 1; -c)
+        \\[1em]
+        G_l & \equiv {_2F_1}(l + 1, \alpha + 2; \alpha + \beta + 2; -c)
+        \quad.
+    \end{split}
+\end{align}
+%
+Note that for a given value of $\alpha$, $\beta$, and $c$, we need only compute
+$F_0$, $F_1$, $G_0$, and $G_1$ directly, since the remaining terms may be
+obtained recursively:
+%
+\begin{align}
+    \begin{split}
+        F_l &=
+        \dfrac{\alpha + \beta + 1 - l}{2l}F_{l - 2}
+        - \dfrac{2\alpha + \beta + 2 - 3l}{2l}F_{l - 1}
+        \\[1em]
+        G_l &=
+        \dfrac{\alpha + \beta + 2 - l}{2l}G_{l - 2}
+        - \dfrac{2\alpha + \beta + 4 - 3l}{2l}G_{l - 1}
+        \quad.
+    \end{split}
+\end{align}
+%
+We may now evaluate the integral in Equation~(\ref{eq:Erb}):
+%
+\begin{align}
+    \mathrm{E} \big[ y_{lm} \big] & =
+    \begin{dcases}
+        \frac{1}{b\sqrt{2\pi\nu}}
+        \int\limits_0^\infty
+        \left(
+        1 -
+        \dfrac{(1 - b) \alpha c}{2(\alpha + \beta)} F_l
+        \right)
+        \\
+        \qquad\qquad\times
+        \exp\left[
+            -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
+            \right]
+        \mathrm{d} b
+         &
+        \qquad\qquad
+        l = m = 0
+        \\
+        -
+        \frac{1}{b\sqrt{2\pi\nu}}
+        \int\limits_0^\infty
+        \dfrac{(1 - b) \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
+        \\
+        \qquad\qquad\times
+        \left[
+            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
+            \right]
+        \\
+        \qquad\qquad\times
+        \exp\left[
+            -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
+            \right]
+        \mathrm{d} b
+         &
+        \qquad\qquad
+        l > 0, m = 0 \\[2em]
+        0
+         &
+        \qquad\qquad m \ne 0
+    \end{dcases}
+\end{align}
+%
+Most of the terms in the integrands above are constants, so these
+integrals may also be expressed in closed form:
+%
+\begin{align}
+    \mathrm{E} \big[ y_{lm} \big]
+     & =
+    \begin{dcases}
+        1 -
+        \dfrac{\gamma \alpha c}{2(\alpha + \beta)} F_l
+         &
+        \qquad
+        l = m = 0    \\[2em]
+        -\dfrac{\gamma \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
+        \left[
+            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
+            \right]
+         &
+        \qquad
+        l > 0, m = 0 \\[2em]
+        0
+         &
+        \qquad m \ne 0
+    \end{dcases}
+\end{align}
+%
+where we define
+%
+\begin{align}
+    \gamma \equiv 1 - \exp\left[ \mu + \frac{1}{2}\nu\right]
+    \quad.
+\end{align}
+%
+
+\subsection{Second moment}
+%
+The second moment of the distribution of $y_{lm}$ over $r$ and $b$ is
+%
+\begin{align}
+    \label{eq:Erb2}
+    \begin{split}
+        \mathrm{E} \big[ y_{lm} y_{l'm'} \big] &=
+        \int\limits_0^\infty \int\limits_0^1
+        y_{lm} (r, 1 - b)
+        y_{l'm'} (r, 1 - b)
+        p \big(r, 1 - b \, \big| \, \alpha, \beta, \mu, \nu \big)
+        \mathrm{d} r
+        \,
+        \mathrm{d} b
+        \\
+        &=
+        \int\limits_0^\infty
+        \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]
+        p \big(b \, \big| \, \mu, \nu \big)
+        \mathrm{d} b
+        \quad,
+    \end{split}
+    %
+    \\
+    \intertext{where}
+    \label{eq:Er2}
+    %
+    \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]
+     & =
+    \int\limits_0^1
+    y_{lm} (r, 1 - b)
+    y_{l'm'} (r, 1 - b)
+    p \big(r \, \big| \, \alpha, \beta \big)
+    \mathrm{d} r
+    \quad.
+\end{align}
+%
+Expanding Equation~(\ref{eq:Er2}), we obtain
+%
+\begin{align}
+    \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]
+     & =
+    \begin{dcases}
+        \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+        \int\limits_0^1
+        \left(
+        1 - \dfrac{(1 - b) c r}{2 (1 + c r)}
+        \right)^2
+        r^{\alpha - 1}
+        (1 - r)^{\beta - 1}
+        \mathrm{d} r
+         &
+        \quad\quad
+        \parbox[t]{\linewidth}{$l = l',              \\m = m' = 0$}
+        \\[2em]
+        %
+        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+        \int\limits_0^1
+        \left(1 - \dfrac{(1 - b) c r}{2 (1 + c r)}\right)
+        \\
+        \qquad\qquad\quad\quad\times
+        \dfrac{(1 - b) c r \left( 2 + c r \right) r^{\alpha - 1}
+            (1 - r)^{\beta - 1}}
+        {2 \sqrt{2l + 1} (1 + c r)^{l + 1}}
+        \mathrm{d} r
+         &
+        \quad\quad
+        \parbox[t]{\linewidth}{\vspace{-3em}$l > 0,  \\l' = 0, \\m = m' = 0$}
+        \\[2em]
+        %
+        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+        \int\limits_0^1
+        \left(1 - \dfrac{(1 - b) c r}{2 (1 + c r)}\right)
+        \\
+        \qquad\qquad\quad\quad\times
+        \dfrac{(1 - b) c r \left( 2 + c r \right) r^{\alpha - 1}
+            (1 - r)^{\beta - 1}}
+        {2 \sqrt{2l' + 1} (1 + c r)^{l' + 1}}
+        \mathrm{d} r
+         &
+        \quad\quad
+        \parbox[t]{\linewidth}{\vspace{-3em}$l = 0,  \\l' > 0, \\m = m' = 0$}
+        \\[2em]
+        %
+        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+        \int\limits_0^1
+        \dfrac{(1 - b)^2 c^2 r^2 \left( 2 + c r \right)^2 r^{\alpha - 1}
+            (1 - r)^{\beta - 1}}
+        {4 \sqrt{(2l + 1)(2l' + 1)} (1 + c r)^{l + l' + 2}}
+        \mathrm{d} r
+         &
+        \quad\quad
+        \parbox[t]{\linewidth}{\vspace{-1.5em}$l > 0 \\l' > 0 \\m = m' = 0$}
+        \\[2em]
+        %
+        0
+         &
+        \quad\quad m, m' \ne 0
+        \quad.
+    \end{dcases}
+\end{align}
+
 %\bibliography

### 2020-07-29 tex/ms.tex
second moment of rb dist
typos; finished mom2

@@
 \begin{align}
+    \label{eq:lognormal}
     p \big(b \, \big| \, \mu, \nu \big)
@@
         1 -
-        \dfrac{(1 - b) \alpha c}{2(\alpha + \beta)} F_l
+        \dfrac{(1 - b) \alpha c}{2(\alpha + \beta)} G_0^1
          &
@@
         \left[
-            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
+            G_l^1 + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l^2
             \right]
@@
 \begin{align}
-    \begin{split}
-        F_l & \equiv {_2F_1}(l + 1, \alpha + 1; \alpha + \beta + 1; -c)
-        \\[1em]
-        G_l & \equiv {_2F_1}(l + 1, \alpha + 2; \alpha + \beta + 2; -c)
+    G_j^k & \equiv {_2F_1}(j + 1, \alpha + k; \alpha + \beta + k; -c)
     \quad.
-    \end{split}
 \end{align}
@@
 Note that for a given value of $\alpha$, $\beta$, and $c$, we need only compute
-$F_0$, $F_1$, $G_0$, and $G_1$ directly, since the remaining terms may be
+$G_0^1$, $G_1^1$, $G_0^2$, and $G_1^2$ directly, since the remaining terms may be
 obtained recursively:
@@
 \begin{align}
-    \begin{split}
-        F_l &=
-        \dfrac{\alpha + \beta + 1 - l}{2l}F_{l - 2}
-        - \dfrac{2\alpha + \beta + 2 - 3l}{2l}F_{l - 1}
-        \\[1em]
-        G_l &=
-        \dfrac{\alpha + \beta + 2 - l}{2l}G_{l - 2}
-        - \dfrac{2\alpha + \beta + 4 - 3l}{2l}G_{l - 1}
+    G_l^k & =
+    \dfrac{\alpha + \beta + k - l}{2l}G_{l - 2}^k
+    - \dfrac{2\alpha + \beta + 2k - 3l}{2l}G_{l - 1}^k
     \quad.
-    \end{split}
 \end{align}
 %
-We may now evaluate the integral in Equation~(\ref{eq:Erb}):
-%
-\begin{align}
-    \mathrm{E} \big[ y_{lm} \big] & =
-    \begin{dcases}
-        \frac{1}{b\sqrt{2\pi\nu}}
-        \int\limits_0^\infty
-        \left(
-        1 -
-        \dfrac{(1 - b) \alpha c}{2(\alpha + \beta)} F_l
-        \right)
-        \\
-        \qquad\qquad\times
-        \exp\left[
-            -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
-            \right]
-        \mathrm{d} b
-         &
-        \qquad\qquad
-        l = m = 0
-        \\
-        -
-        \frac{1}{b\sqrt{2\pi\nu}}
-        \int\limits_0^\infty
-        \dfrac{(1 - b) \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
-        \\
-        \qquad\qquad\times
-        \left[
-            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
-            \right]
-        \\
-        \qquad\qquad\times
-        \exp\left[
-            -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
-            \right]
-        \mathrm{d} b
-         &
-        \qquad\qquad
-        l > 0, m = 0 \\[2em]
-        0
-         &
-        \qquad\qquad m \ne 0
-    \end{dcases}
-\end{align}
-%
-Most of the terms in the integrands above are constants, so these
-integrals may also be expressed in closed form:
+We may now evaluate the integral in Equation~(\ref{eq:Erb}), with
+PDF given by Equation~(\ref{eq:lognormal}).
+Most of the terms in the integrands are constants, so the
+integrals are straightforward to evaluate and may also
+be expressed in closed form:
 %
@@
         1 -
-        \dfrac{\gamma \alpha c}{2(\alpha + \beta)} F_l
+        \dfrac{\gamma_1 \alpha c}{2(\alpha + \beta)} G_0^1
          &
@@
         l = m = 0    \\[2em]
-        -\dfrac{\gamma \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
+        -\dfrac{\gamma_1 \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
         \left[
-            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
+            G_l^1 + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l^2
             \right]
@@
 \begin{align}
-    \gamma \equiv 1 - \exp\left[ \mu + \frac{1}{2}\nu\right]
+    \gamma_1 \equiv 1 - \exp\left[ \mu + \frac{1}{2}\nu\right]
     \quad.
@@
         %
-        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+        \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
         \int\limits_0^1
@@
         \quad\quad
-        \parbox[t]{\linewidth}{\vspace{-1.5em}$l > 0 \\l' > 0 \\m = m' = 0$}
+        \parbox[t]{\linewidth}{\vspace{-1.5em}$l > 0, \\l' > 0, \\m = m' = 0$}
         \\[2em]
@@
 \end{align}
+%
+These integrals again reduce to closed form expressions:
+%
+\begin{align}
+    \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]
+     & =
+    \begin{dcases}
+        1 + \frac{\left(\alpha +1\right) \alpha
+            \left(1-b\right)^2 c^2}{4 \left(c+1\right)
+            \left(\alpha + \beta \right)}
+        \\[0.75em]
+        \quad\quad\quad\quad\quad
+        -\frac{\alpha  \left(1-b\right) c}{\alpha +\beta }
+        \\[0.75em]
+        \quad\quad\quad\quad\quad
+        \times
+        \bigg[G_0^1
+            +
+            \frac{\left(\alpha +1\right) \left(1-b\right)
+                c \left(\alpha +\beta +\alpha  c+c\right)}{4 \left(c+1\right)
+                \left(\alpha +\beta +1\right)} G_0^2\bigg]
+         &
+        \quad\quad
+        \parbox[t]{\linewidth}{\vspace{-4.75em}$l = l', \\m = m' = 0$}
+        \\[2em]
+        %
+        -\frac{\alpha (1-b) c}{2 \sqrt{2 l+1} (\alpha +\beta )}
+        \\[0.75em]
+        \quad\quad\quad\quad\quad
+        \times
+        \bigg[
+            2G_l^1
+            + \dfrac{c(\alpha + 1)}{\alpha + \beta + 1} G_l^2
+            \\[0.75em]
+        \qquad
+        -\dfrac{c(1-b)(\alpha + 1)}{\alpha + \beta + 1} G_{l+1}^2
+        \\[0.75em]
+        \qquad
+        -\dfrac{c^2(1-b)(\alpha + 1)(\alpha + 2)}{2(\alpha + \beta + 1)(\alpha + \beta + 2)} G_{l+1}^3
+        \bigg]
+         &
+        \quad\quad
+        \parbox[t]{\linewidth}{\vspace{-7.5em}$l > 0,   \\l' = 0, \\m = m' = 0$}
+        \\[2em]
+        %
+        \textrm{same as previous case, but with $l \rightarrow l'$}
+         &
+        \quad\quad
+        \parbox[t]{\linewidth}{\vspace{-2em}$l = 0,     \\l' > 0, \\m = m' = 0$}
+        \\[2em]
+        %
+        \frac{\alpha  (\alpha +1) (1-b)^2 c^2}{(\alpha +\beta )
+            (\alpha +\beta +1) \sqrt{(2 l+1) (2 p+1)}}
+        \\[0.75em]
+        \quad\quad\quad\quad\quad
+        \times
+        \bigg[
+        G_{l+l'+1}^2
+        +
+        \frac{c(\alpha + 2)}{\alpha + \beta + 2}G_{l+l'+1}^3
+        \\[0.75em]
+        \qquad
+        +
+        \frac{c^2(\alpha + 2)(\alpha + 3)}{4(\alpha + \beta + 2)(\alpha + \beta + 3)}G_{l+l'+1}^4
+        \bigg]
+         &
+        \quad\quad
+        \parbox[t]{\linewidth}{\vspace{-5.25em}$l > 0,  \\l' > 0, \\m = m' = 0$}
+        \\[2em]
+        %
+        0
+         &
+        \quad\quad m, m' \ne 0
+        \quad.
+    \end{dcases}
+\end{align}
+%
+The $G_j^k$ functions may again be computed recursively, although we now also
+require the direct evaluation of $G_0^3$, $G_1^3$, $G_0^4$, and $G_1^4$ as
+our lower boundary conditions.
+
+The final step is to evaluate the integral in Equation~(\ref{eq:Er2}), given
+the PDF in Equation~(\ref{eq:lognormal}):
+%
+\begin{align}
+    \mathrm{E} \big[ y_{lm} y_{l'm'} \big]
+     & =
+    \begin{dcases}
+        1 + \frac{\left(\alpha +1\right) \alpha
+            \gamma_2 c^2}{4 \left(c+1\right)
+            \left(\alpha + \beta \right)}
+        \\[0.75em]
+        \quad\quad\quad\quad\quad
+        -\frac{\alpha  \gamma_1 c}{\alpha +\beta }
+        \\[0.75em]
+        \quad\quad\quad\quad\quad
+        \times
+        \bigg[G_0^1
+            +
+            \frac{\left(\alpha +1\right) \gamma_2
+                c \left(\alpha +\beta +\alpha  c+c\right)}{4 \gamma_1 \left(c+1\right)
+                \left(\alpha +\beta +1\right)} G_0^2\bigg]
+         &
+        \quad\quad
+        \parbox[t]{\linewidth}{\vspace{-4.75em}$l = l', \\m = m' = 0$}
+        \\[2em]
+        %
+        -\frac{\alpha \gamma_1 c}{2 \sqrt{2 l+1} (\alpha +\beta )}
+        \\[0.75em]
+        \quad\quad\quad\quad\quad
+        \times
+        \bigg[
+            2G_l^1
+            + \dfrac{c(\alpha + 1)}{\alpha + \beta + 1} G_l^2
+            \\[0.75em]
+        \qquad
+        -\dfrac{c\gamma_2(\alpha + 1)}{\gamma_1(\alpha + \beta + 1)} G_{l+1}^2
+        \\[0.75em]
+        \qquad
+        -\dfrac{c^2\gamma_2(\alpha + 1)(\alpha + 2)}{2\gamma_1(\alpha + \beta + 1)(\alpha + \beta + 2)} G_{l+1}^3
+        \bigg]
+         &
+        \quad\quad
+        \parbox[t]{\linewidth}{\vspace{-7.5em}$l > 0,   \\l' = 0, \\m = m' = 0$}
+        \\[2em]
+        %
+        \textrm{same as previous case, but with $l \rightarrow l'$}
+         &
+        \quad\quad
+        \parbox[t]{\linewidth}{\vspace{-2em}$l = 0,     \\l' > 0, \\m = m' = 0$}
+        \\[2em]
+        %
+        \frac{\alpha  (\alpha +1) \gamma_2 c^2}{(\alpha +\beta )
+            (\alpha +\beta +1) \sqrt{(2 l+1) (2 p+1)}}
+        \\[0.75em]
+        \quad\quad\quad\quad\quad
+        \times
+        \bigg[
+        G_{l+l'+1}^2
+        +
+        \frac{c(\alpha + 2)}{\alpha + \beta + 2}G_{l+l'+1}^3
+        \\[0.75em]
+        \qquad
+        +
+        \frac{c^2(\alpha + 2)(\alpha + 3)}{4(\alpha + \beta + 2)(\alpha + \beta + 3)}G_{l+l'+1}^4
+        \bigg]
+         &
+        \quad\quad
+        \parbox[t]{\linewidth}{\vspace{-5.25em}$l > 0,  \\l' > 0, \\m = m' = 0$}
+        \\[2em]
+        %
+        0
+         &
+        \quad\quad m, m' \ne 0
+        \quad,
+    \end{dcases}
+\end{align}
+%
+where we define
+%
+\begin{align}
+    \gamma_2 \equiv 1
+    - 2\exp\bigg[ \mu + \frac{1}{2}\nu\bigg]
+    + \exp\bigg[ 2\mu + 2\nu\bigg]
+    \quad.
+\end{align}
 

### 2020-07-30 tex/ms.tex
coded up second moment

@@
         \quad\quad
-        \parbox[t]{\linewidth}{$l = l',               \\m = m' = 0$}
+        \parbox[t]{\linewidth}{$l = l' = 0,           \\m = m' = 0$}
         \\[2em]
@@
         %
-        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
-        \int\limits_0^1
-        \left(1 - \dfrac{(1 - b) c r}{2 (1 + c r)}\right)
-        \\
-        \qquad\qquad\quad\quad\times
-        \dfrac{(1 - b) c r \left( 2 + c r \right) r^{\alpha - 1}
-            (1 - r)^{\beta - 1}}
-        {2 \sqrt{2l' + 1} (1 + c r)^{l' + 1}}
-        \mathrm{d} r
-         &
-        \quad\quad
-        \parbox[t]{\linewidth}{\vspace{-3em}$l = 0,   \\l' > 0, \\m = m' = 0$}
-        \\[2em]
-        %
         \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
@@
         \quad\quad m, m' \ne 0
-        \quad.
+        \quad,
     \end{dcases}
 \end{align}
+%
+where, by symmetry, the expression for $l' > 0, \, l = 0, \, m = m' = 0$ is the same as
+in the second case, provided we make the substitution $l \rightarrow l'$.
 %
@@
         \quad\quad
-        \parbox[t]{\linewidth}{\vspace{-4.75em}$l = l', \\m = m' = 0$}
+        \parbox[t]{\linewidth}{\vspace{-4.75em}$l = l' = 0, \\m = m' = 0$}
         \\[2em]
@@
         \\[2em]
-        %
-        \textrm{same as previous case, but with $l \rightarrow l'$}
-         &
-        \quad\quad
-        \parbox[t]{\linewidth}{\vspace{-2em}$l = 0,     \\l' > 0, \\m = m' = 0$}
-        \\[2em]
-        %
         \frac{\alpha  (\alpha +1) (1-b)^2 c^2}{(\alpha +\beta )
-            (\alpha +\beta +1) \sqrt{(2 l+1) (2 p+1)}}
+            (\alpha +\beta +1) \sqrt{(2 l+1) (2 l'+1)}}
         \\[0.75em]
@@
         %
-        \textrm{same as previous case, but with $l \rightarrow l'$}
-         &
-        \quad\quad
-        \parbox[t]{\linewidth}{\vspace{-2em}$l = 0,     \\l' > 0, \\m = m' = 0$}
-        \\[2em]
-        %
         \frac{\alpha  (\alpha +1) \gamma_2 c^2}{(\alpha +\beta )
-            (\alpha +\beta +1) \sqrt{(2 l+1) (2 p+1)}}
+            (\alpha +\beta +1) \sqrt{(2 l+1) (2 l'+1)}}
         \\[0.75em]
@@
     + \exp\bigg[ 2\mu + 2\nu\bigg]
-    \quad.
+    \quad,
 \end{align}
+%
+and the expression for $l' > 0, \, l = 0, \, m = m' = 0$ is again the same as in
+the second case, provided we make the substitution $l \rightarrow l'$.
 

### 2020-07-30 starry_gp/spot.py
fixes

@@ -2,39 +2,6 @@ from .transform import eigen, get_alpha_beta
 import numpy as np
 from scipy.special import hyp2f1
 
-# DEBUG
-
-from scipy.stats import beta as Beta
-from scipy.stats import lognorm
-
-
-def second_integral(ydeg, mu_r=0.1, nu_r=0.01, mu_b=-3.0, nu_b=1.0, c=1.0, K=1000000):
-    # Draw
-    alpha, beta = get_alpha_beta(mu_r, nu_r)
-    r = Beta.rvs(alpha, beta, size=K)
-    delta = 1 - lognorm.rvs(s=np.sqrt(nu_b), scale=np.exp(mu_b), size=K)
-
-    # Transform
-    s = np.zeros((K, ydeg + 1))
-    s[:, 0] = 1 - 0.5 * delta * c * r / (1 + c * r)
-    for l in range(1, ydeg + 1):
-        s[:, l] = (
-            -0.5
-            * delta
-            * c
-            * r
-            * (2 + c * r)
-            / (np.sqrt(2 * l + 1) * (1 + c * r) ** (l + 1))
-        )
-    l = np.arange(ydeg + 1)
-    S = np.zeros((K, (ydeg + 1) * (ydeg + 1)))
-    S[:, l * (l + 1)] = s
-
-    # Compute second moment
-    mu = np.mean(S, axis=0)
-    cov = np.cov(S.T)
-    return cov + mu.reshape(-1, 1) @ mu.reshape(1, -1)
-
 
 class SpotIntegral(object):
     """Marginalizes over the spot size and contrast distribution.
@@ -70,30 +37,18 @@ class SpotIntegral(object):
         l = np.arange(ydeg + 1)
         self.i = l * (l + 1)
         self.ij = np.ix_(self.i, self.i)
+        self.E1 = np.zeros(self.N)
+        self.E2 = np.zeros((self.N, self.N))
         self.set_params()
 
     def set_params(self, mu_r=0.1, nu_r=0.01, mu_b=-3.0, nu_b=1.0):
+
+        # Get the Beta params
         alpha, beta = get_alpha_beta(mu_r, nu_r)
-        g1 = 1 - np.exp(mu_b + 0.5 * nu_b)
-        g2 = 1 - 2 * np.exp(mu_b + 0.5 * nu_b) + np.exp(2 * mu_b + 2 * nu_b)
-
-        # Integral normalization
[…]

### 2020-07-30 tests/test_gp.py

@@ -1,8 +1,10 @@
 from starry_gp.gp import YlmGP
-from numerical import get_spot_function
+from starry_gp.transform import get_alpha_beta
+from numerical import spot
 import numpy as np
 from tqdm import tqdm
 from scipy.stats import beta as Beta
+from scipy.stats import lognorm as LogNormal
 
 
 def test_moments():
@@ -10,35 +12,43 @@ def test_moments():
     # Settings
     ydeg = 5
     atol = 5e-2
-    mu_beta = 0.5
-    nu_beta = 0.01
-    mu_lns = np.log(0.1)
-    sig_lns = 0.1
-    mu_lna = np.log(0.2)
-    sig_lna = 0.1
+    mu_lat = 0.9
+    nu_lat = 0.1
+    mu_r = 0.05
+    nu_r = 0.1
+    mu_b = np.log(0.3)
+    nu_b = 0.1
     np.random.seed(0)
     nsamples = 100000
 
     # Integrate analytically
     gp = YlmGP(ydeg)
-    gp.set_params(mu_beta, nu_beta, mu_lns, sig_lns, mu_lna, sig_lna)
+    gp.set_params(mu_lat, nu_lat, mu_r, nu_r, mu_b, nu_b)
     mu = gp.mu
     cov = gp.cov
 
     # Integrate numerically
-    spot = get_spot_function(ydeg)
     y = np.zeros((nsamples, (ydeg + 1) ** 2))
-    alpha = mu_beta * (1 / nu_beta - 1)
-    beta = (1 - mu_beta) * (1 / nu_beta - 1)
+    alpha_lat, beta_lat = get_alpha_beta(mu_lat, nu_lat)
+    alpha_r, beta_r = get_alpha_beta(mu_r, nu_r)
     for n in tqdm(range(nsamples)):
-        lnsigma = mu_lns + sig_lns * np.random.randn()
-        sigma = np.exp(lnsigma)
-        lnamp = mu_lna + sig_lna * np.random.randn()
-        amp = 1 / (1 + np.exp(-lnamp))
-        lat = np.arccos(Beta.rvs(alpha, beta)) * 180.0 / np.pi
+
+        # Draw the spot size
+        r = Beta.rvs(alpha_r, beta_r, size=nsamples)
+
+        # Draw the spot amplitude
+        delta = 1 - LogNormal.rvs(s=np.sqrt(nu_b), scale=np.exp(mu_b), size=nsamples)
+
+        # Draw the latitude
[…]

### 2020-07-30 tests/test_values.py

@@ -1,75 +1,0 @@
-from starry_gp.gp import YlmGP
-import starry
-import numpy as np
-import matplotlib.pyplot as plt
-
-plt.style.use("default")
-
-
-def test_infs():
-    gp = YlmGP(10)
-
-    fig, ax = plt.subplots(2)
-
-    gp.set_params(0.81, 0.0007, -2.9, 0.30, -2.50, 0.50)
-    ax[0].plot(np.diag(gp.cov))
-
-    ax[1].plot(np.diag(gp.S.Q))
-
-    gp.set_params(0.81, 0.0007, -2.9, 0.40, -2.50, 0.50)
-    ax[0].plot(np.diag(gp.cov))
-
-    ax[1].plot(np.diag(gp.S.Q))
-
-    gp.set_params(0.81, 0.0007, -2.9, 0.45, -2.50, 0.50)
-    ax[0].plot(np.diag(gp.cov))
-
-    ax[1].plot(np.diag(gp.S.Q))
-
-    ax[0].set_yscale("log")
-    ax[1].set_yscale("log")
-
-    fig, ax = plt.subplots(2)
-    gp.set_params(0.81, 0.0007, -2.9, 0.40, -2.50, 0.50)
-    ax[0].imshow(gp.S.coeffs @ gp.S.Q)
-    gp.set_params(0.81, 0.0007, -2.9, 0.45, -2.50, 0.50)
-    ax[1].imshow(gp.S.coeffs @ gp.S.Q)
-
-    # DEBUG
-    s = np.linspace(0.35, 0.45, 100)
-    v = np.zeros_like(s)
-    for i in range(len(s)):
-        gp.set_params(0.81, 0.0007, -2.9, s[i], -2.50, 0.50)
-        v[i] = gp.S.U[-1, 4]
-    plt.figure()
-    plt.plot(s, v)
-    plt.show()
-
-
-if __name__ == "__main__":
-    # test_infs()
-
-    intensity = -0.1
-    sigma = 0.6
-    map = starry.Map(ydeg=10, lazy=False)
-    map.add_spot(intensity=intensity, sigma=sigma)
-
-    # Compute the intensity along the equator
-    # Remove the baseline intensity and normalize it
-    lon = np.linspace(-180, 180, 1000)
-    baseline = 1.0 / np.pi
-    I = -(map.intensity(lon=lon) - baseline) / (map.intensity(lon=0) - baseline)
-
-    # Compute the intensity of a normalized gaussian
[…]

### 2020-07-30 tex/ms.tex
fixes
more general r param

--- 2020-07-29 tex/ms.tex
@@
     \begin{cases}
-        1 - \dfrac{\delta c r}{2 (1 + c r)}
+        1 - \dfrac{\delta r'}{2 (1 + r')}
          & l = m = 0    \\[2em]
-        -\dfrac{\delta c r \left( 2 + c r \right)}
-        {2 \sqrt{2l + 1} (1 + c r)^{l + 1}}
+        -\dfrac{\delta r' \left( 2 + r' \right)}
+        {2 \sqrt{2l + 1} (1 + r')^{l + 1}}
          & l > 0, m = 0 \\[2em]
@@
 %
-where $\delta \in [-\infty, 1]$ is the spot contrast (the fractional
+where
+%
+\begin{align}
+    \label{eq:rprime}
+    r' \equiv c_0 + c_1 r
+\end{align}
+%
+and $c_0, c_1 > 0$ are scaling constants (defined in more detail below).
+In the expressions above,
+$\delta \in (-\infty, 1]$ is the spot contrast (the fractional
 decrease in the brightness
-at the center of the spot) and $r \in (0, 1]$ is the normalized spot radius.
-The quantity $c \in (0, \infty)$ is a normalization constant for the radius
-(more details below).
+at the center of the spot) and $r \in [0, 1]$ is the normalized spot radius.
 The expression in Equation~(\ref{eq:ylm0}) is convenient because it satisfies
-four important properties:
+five important properties:
 %
 \begin{enumerate}[itemsep=2pt,parsep=1pt,label=\textbf{\arabic*}]
+    \item The surface intensity is azimuthally symmetric about the spot center
     \item The surface intensity monotonically increases away from the spot center
@@
 the unit sphere is $4\pi$.
-Combining this with Equation~(\ref{eq:ylm0}) and rearranging, we may write
+Note that the expression above is independent of the azimuth $\varphi$,
+a consequence of the fact that all harmonics with $m \ne 0$ are zero; this
+expansion is therefore azimuthally symmetric about the spot center, as
+stated in \textbf{1}.
+%
+Combining the above expression with Equation~(\ref{eq:ylm0}) and
+rearranging, we may write
 %
@@
     I(\theta, \varphi) =
-    1 + \dfrac{\delta c r}{2}
+    I(\theta) =
+    1 + \dfrac{\delta r'}{2}
     -
-    \dfrac{\delta c r \left( 2 + c r \right)}{2 (1 + c r)}
-    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + c r}\right)^l P_l(\cos\theta)
+    \dfrac{\delta r' \left( 2 + r' \right)}{2 (1 + r')}
+    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + r'}\right)^l P_l(\cos\theta)
     \quad.
@@
     \label{eq:Ifinal}
-    I(\theta, \varphi) & = A - \frac{B}{\sqrt{C - \cos\theta}}
+    I(\theta) & = A - \frac{B}{\sqrt{C - \cos\theta}}
     \quad,
@@
     \begin{split}
-        A & = 1 + \dfrac{\delta c r}{2}                      \\
-        B & = \delta c r (2 + cr) \sqrt{\dfrac{1}{8 + 8 cr}} \\
-        C & = \dfrac{1 + (1 + c r)^2}{2 + 2 c r}
+        A & = 1 + \dfrac{\delta r'}{2}                      \\
+        B & = \delta r' (2 + r') \sqrt{\dfrac{1}{8 + 8 r'}} \\
+        C & = \dfrac{1 + (1 + r')^2}{2 + 2 r'}
     \end{split}
@@
     \label{eq:Ideriv}
-    \dfrac{\mathrm{d}I(\theta, \varphi)}{\mathrm{d}\theta} & =
+    \dfrac{\mathrm{d}I(\theta)}{\mathrm{d}\theta} & =
     -\frac{B\sin\theta}{2(C - \cos\theta)^\frac{3}{2}}
@@
 %
-which is zero only for $\theta = 0$ (for which $I(\theta, \varphi)$ is
+which is zero only for $\theta = 0$ (for which $I(\theta)$ is
 minimized) and $\theta = \pi$ (for which it is maximized). The intensity
 therefore increases monotonically from the spot center to the antipode,
-as stated in \textbf{1}. The value at the minimum is
+as stated in \textbf{2}. The value at the minimum is
 %
@@
 %
-as stated in \textbf{2}, and the value at the maximum is
+as stated in \textbf{3}, and the value at the maximum is
 %
@@
 %
-as stated in \textbf{3}.
-Finally, to show \textbf{4}, let us compute the half width at half minimum
+as stated in \textbf{4}.
+Finally, to show \textbf{5}, let us compute the half width at half minimum
 $\Delta\theta$ of the intensity profile:
@@
     \Delta\theta =
-    \cos^{-1} \left[ \dfrac{2 + 3 c r (2 + c r)}{2 (1 + c r)^3} \right]
+    \cos^{-1} \left[ \dfrac{2 + 3 r' (2 + r')}{2 (1 + r')^3} \right]
     \quad.
@@
 %
-Differentiation with respect to $r$ yields
+Differentiation with respect to $r'$ yields
 %
 \begin{align}
-    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}r} =
-    \frac{3 c^2 r \left(2 + c r\right)}{\left(1 + c r\right)^4
-        \sqrt{\frac{c^2 r^2 \left(2 + c r\right)^2 \left(1 + 2 c r\right)
-                \left(3 + 2 c r\right)}{\left(1 + c r\right)^6}}}
+    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}r'} =
+    \frac{3 r' \left(2 + r'\right)}{\left(1 + r'\right)^4
+        \sqrt{\frac{r'^2 \left(2 + r'\right)^2 \left(1 + 2 r'\right)
+                \left(3 + 2 r'\right)}{\left(1 + r'\right)^6}}}
     \quad,
@@
 %
-which is positive definite for all $c, r > 0$.
+which is positive definite for all $r' > 0$.
 
@@
 %
-Since we require the brghtness $b = 1 - \delta$ at the center of the spot to
-be non-negative, it is convenient to parametrize the distribution of $b$
-instead of $\delta$. We model $b$ as a random variable drawn from a log-normal
+Since we require the brightness
+%
+\begin{align}
+    b \equiv 1 - \delta
+\end{align}
+%
+at the center of the spot to
+be non-negative, it is convenient to treat $b \in [0, \infty)$ as our
+random variable instead of $\delta$. We model $b$ as a random variable
+drawn from a log-normal
 distribution:
@@
      & =
+    \resizebox{.75\hsize}{!}{$
             \begin{dcases}
@@
                 \left(
-        1 - \dfrac{(1 - b) c r}{2 (1 + c r)}
+                1 - \dfrac{(1 - b) (c_0 + c_1 r)}{2 (1 + c_0 + c_1 r)}
                 \right)
@@
                 \int\limits_0^1
-        \dfrac{(1 - b) c r \left( 2 + c r \right)r^{\alpha - 1}
+                \dfrac{(1 - b) (c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)r^{\alpha - 1}
                 (1 - r)^{\beta - 1}}
-        {2 \sqrt{2l + 1} (1 + c r)^{l + 1}}
+                {2 \sqrt{2l + 1} (1 + c_0 + c_1 r)^{l + 1}}
                 \mathrm{d} r
@@
             \end{dcases}
+        $}
 \end{align}
@@
 Thanks to the dependence of the spherical harmonic coefficients on only
-powers of $r$ and $(1 + cr)$, the integrals above may be expressed in closed
+powers of $r$ and $(1 + c_0 + c_1 r)$, the integrals above may be expressed in closed
 form in terms of the hypergeometric function ${_2F_1}$:
@@
      & =
+    \resizebox{.75\hsize}{!}{$
             \begin{dcases}
                 1 -
-        \dfrac{(1 - b) \alpha c}{2(\alpha + \beta)} F_l
+                \dfrac{(1 - b)}{2(1 + c_0)}
+                \bigg[c_0H_0^0 + c_1 H_0^1\bigg]
                  &
@@
                 l = m = 0    \\[2em]
-        -\dfrac{(1 - b) \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
-        \left[
-            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
-            \right]
+                -\dfrac{(1 - b)}{2\sqrt{2l + 1} (1 + c_0)^{l+1}}
+                \bigg[
+                    c_0(2+c_0)H_l^0
+                    +
+                    2c_1(1 + c_0)H_l^1
+                    +
+                    c_1^2 H_l^2
+                    \bigg]
                  &
@@
 \begin{align}
-    \begin{split}
-        F_l & \equiv {_2F_1}(l + 1, \alpha + 1; \alpha + \beta + 1; -c)
-        \\[1em]
-        G_l & \equiv {_2F_1}(l + 1, \alpha + 2; \alpha + \beta + 2; -c)
+    H_j^k & \equiv \left(\prod_{n=0}^{k-1} \lambda_n\right) G_j^k
     \quad.
-    \end{split}
+\end{align}
+%
+with
+%
+\begin{align}
+    \lambda_n & \equiv \dfrac{\alpha + n}{\alpha + \beta + n}
+\end{align}
+%
+and
+%
+\begin{align}
+    G_j^k & \equiv {_2F_1}\left(j + 1, \alpha + k; \alpha + \beta + k; -\dfrac{c_1}{1 + c_0}\right)
+    \quad.
 \end{align}
@@
 Note that for a given value of $\alpha$, $\beta$, and $c$, we need only compute
-$F_0$, $F_1$, $G_0$, and $G_1$ directly, since the remaining terms may be
+$G_0^0$, $G_1^0$, $G_0^1$, and $G_1^1$ directly, since the remaining terms may be
 obtained recursively:
@@
 \begin{align}
-    \begin{split}
-        F_l &=
-        \dfrac{\alpha + \beta + 1 - l}{2l}F_{l - 2}
-        - \dfrac{2\alpha + \beta + 2 - 3l}{2l}F_{l - 1}
-        \\[1em]
-        G_l &=
-        \dfrac{\alpha + \beta + 2 - l}{2l}G_{l - 2}
-        - \dfrac{2\alpha + \beta + 4 - 3l}{2l}G_{l - 1}
+    \label{eq:Grec}
+    G_j^k & =
+    \dfrac{\alpha + \beta + k - j}{2j}G_{j - 2}^k
+    - \dfrac{2\alpha + \beta + 2k - 3j}{2j}G_{j - 1}^k
     \quad.
-    \end{split}
 \end{align}
 %
-We may now evaluate the integral in Equation~(\ref{eq:Erb}):
+\textbf{TODO: Recursion for k > 1.\\}
 %
-\begin{align}
-    \mathrm{E} \big[ y_{lm} \big] & =
-    \begin{dcases}
-        \frac{1}{b\sqrt{2\pi\nu}}
-        \int\limits_0^\infty
-        \left(
-        1 -
-        \dfrac{(1 - b) \alpha c}{2(\alpha + \beta)} F_l
-        \right)
-        \\
-        \qquad\qquad\times
-        \exp\left[
-            -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
-            \right]
-        \mathrm{d} b
-         &
-        \qquad\qquad
-        l = m = 0
-        \\
-        -
-        \frac{1}{b\sqrt{2\pi\nu}}
-        \int\limits_0^\infty
-        \dfrac{(1 - b) \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
-        \\
-        \qquad\qquad\times
-        \left[
-            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
-            \right]
-        \\
-        \qquad\qquad\times
-        \exp\left[
-            -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
-            \right]
-        \mathrm{d} b
-         &
-        \qquad\qquad
-        l > 0, m = 0 \\[2em]
-        0
-         &
-        \qquad\qquad m \ne 0
-    \end{dcases}
-\end{align}
-%
-Most of the terms in the integrands above are constants, so these
-integrals may also be expressed in closed form:
+We may now evaluate the integral in Equation~(\ref{eq:Erb}), with
+PDF given by Equation~(\ref{eq:lognormal}).
+Most of the terms in the integrands are constants, so the
+integrals are straightforward to evaluate and may also
+be expressed in closed form:
 %
@@
      & =
+    \resizebox{.75\hsize}{!}{$
             \begin{dcases}
                 1 -
-        \dfrac{\gamma \alpha c}{2(\alpha + \beta)} F_l
+                \dfrac{\gamma_1}{2(1 + c_0)}
+                \bigg[c_0 H_0^0 + c_1 H_0^1\bigg]
                  &
@@
                 l = m = 0    \\[2em]
-        -\dfrac{\gamma \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
-        \left[
-            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
-            \right]
+                -\dfrac{\gamma_1}{2\sqrt{2l + 1} (1 + c_0)^{l+1}}
+                \bigg[
+                    c_0(2+c_0)H_l^0
+                    +
+                    2c_1(1 + c_0)H_l^1
+                    +
+                    c_1^2 H_l^2
+                    \bigg]
                  &
@@
      & =
+    \resizebox{1.5\hsize}{!}{$
             \begin{dcases}
@@
                 \left(
-        1 - \dfrac{(1 - b) c r}{2 (1 + c r)}
+                1 - \dfrac{(1-b) (c_0 + c_1 r)}{2 (1 + c_0 + c_1 r)}
                 \right)^2
+                \\
+                \qquad\qquad\quad\quad\quad\quad\quad
+                \times
+                \,\,
                 r^{\alpha - 1}
@@
                  &
-        \quad\quad
-        \parbox[t]{\linewidth}{$l = l',              \\m = m' = 0$}
+                \qquad
+                \parbox[t]{\linewidth}{\vspace{-2.5em}$l = l' = 0, \\m = m' = 0$}
                 \\[2em]
+                %
+                %
                 %
@@
                 \int\limits_0^1
-        \left(1 - \dfrac{(1 - b) c r}{2 (1 + c r)}\right)
+                \left(
+                1 - \dfrac{(1-b) (c_0 + c_1 r)}{2 (1 + c_0 + c_1 r)}
+                \right)
                 \\
-        \qquad\qquad\quad\quad\times
-        \dfrac{(1 - b) c r \left( 2 + c r \right) r^{\alpha - 1}
-            (1 - r)^{\beta - 1}}
-        {2 \sqrt{2l + 1} (1 + c r)^{l + 1}}
+                \qquad\qquad\quad\quad\quad\quad\quad\quad\quad\quad\quad
+                \times
+                %
+                \left(
+                \dfrac{(1-b) (c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)}
+                    {2 \sqrt{2l + 1} (1 + c_0 + c_1 r)^{l + 1}}
+                \right)
+                %
+                \\[0.75em]
+                \qquad\qquad\quad\quad\quad\quad\quad\quad\quad\quad\quad
+                \times
+                \,\,
+                r^{\alpha - 1}
+                (1 - r)^{\beta - 1}
                 \mathrm{d} r
                  &
-        \quad\quad
-        \parbox[t]{\linewidth}{\vspace{-3em}$l > 0,  \\l' = 0, \\m = m' = 0$}
+                \qquad
+                \parbox[t]{\linewidth}{\vspace{-5em}$l > 0,        \\l' = 0, \\m = m' = 0$}
                 \\[2em]
+                %
+                %
                 %
@@
                 \int\limits_0^1
-        \left(1 - \dfrac{(1 - b) c r}{2 (1 + c r)}\right)
+                \left(
+                \dfrac{(1-b) (c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)}
+                    {2 \sqrt{2l + 1} (1 + c_0 + c_1 r)^{l + 1}}
+                \right)
                 \\
-        \qquad\qquad\quad\quad\times
-        \dfrac{(1 - b) c r \left( 2 + c r \right) r^{\alpha - 1}
-            (1 - r)^{\beta - 1}}
-        {2 \sqrt{2l' + 1} (1 + c r)^{l' + 1}}
+                \qquad\qquad\quad\quad\quad\quad\quad\quad\quad\quad\quad
+                \times
+                %
+                \left(
+                \dfrac{(1-b) (c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)}
+                    {2 \sqrt{2l' + 1} (1 + c_0 + c_1 r)^{l' + 1}}
+                \right)
+                %
+                \\[0.75em]
+                \qquad\qquad\quad\quad\quad\quad\quad\quad\quad\quad\quad
+                \times
+                \,\,
+                r^{\alpha - 1}
+                (1 - r)^{\beta - 1}
                 \mathrm{d} r
                  &
-        \quad\quad
-        \parbox[t]{\linewidth}{\vspace{-3em}$l = 0,  \\l' > 0, \\m = m' = 0$}
+                \qquad
+                \parbox[t]{\linewidth}{\vspace{-5em}$l > 0,        \\l' > 0, \\m = m' = 0$}
                 \\[2em]
                 %
-        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
-        \int\limits_0^1
-        \dfrac{(1 - b)^2 c^2 r^2 \left( 2 + c r \right)^2 r^{\alpha - 1}
-            (1 - r)^{\beta - 1}}
-        {4 \sqrt{(2l + 1)(2l' + 1)} (1 + c r)^{l + l' + 2}}
-        \mathrm{d} r
-         &
-        \quad\quad
-        \parbox[t]{\linewidth}{\vspace{-1.5em}$l > 0 \\l' > 0 \\m = m' = 0$}
-        \\[2em]
+                %
                 %
@@
                  &
-        \quad\quad m, m' \ne 0
+                \qquad m, m' \ne 0
+                \quad,
+            \end{dcases}
+        $}
+    \raisetag{14.5em}
+\end{align}
+%
+where, by symmetry, the expression for $l' > 0, \, l = 0, \, m = m' = 0$ is the same as
+in the second case, provided we make the substitution $l \rightarrow l'$.
+%
+These integrals again reduce to closed form expressions:
+%
+\begin{align}
+    \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]
+     & =
+    \resizebox{1.4\hsize}{!}{$
+            \begin{dcases}
+                1 + \dfrac{1-b}{1+c_0}
+                \bigg[
+                    -c_0 H_0^0
+                    -
+                    c_1 H_0^1
+                    \\[0.75em]
+                \qquad\qquad\enspace\enspace\quad\quad
+                +
+                \dfrac{1-b}{2(1+c_0)}
+                \bigg(
+                \dfrac{c_0^2}{2}H_1^0
+                +
+                c_0 c_1 H_1^1
+                +
+                \dfrac{c_1^2}{2} H_1^2
+                \bigg)
+                \bigg]
+                 &
+                \qquad
+                \parbox[t]{\linewidth}{\vspace{-3.5em}$l = l' = 0, \\m = m' = 0$}
+                \\[2em]
+                %
+                %
+                %
+                \dfrac{1-b}{2\sqrt{2l + 1}(1+c_0)^{l+1}}
+                \\[0.75em]
+                \enspace\quad
+                \times
+                \bigg[
+                    c_0(2 + c_0) H_l^0 + 2 c_1(1 + c_0) H_l^1
+                    + c_1^2 H_l^2
+                    \\[0.75em]
+                \qquad
+                - \dfrac{1 - b}{2(1 + c_0)} \bigg(
+                c_0^2(2 + c_0) H_{l+1}^0
+                + c_0 c_1 (4 + 3 c_0) H_{l+1}^1
+                \\[0.75em]
+                \qquad
+                + (2 + 3 c_0) c_1^2 H_{l + 1}^2
+                + c_1^3 H_{l + 1}^3
+                \bigg)
+                \bigg]
+                 &
+                \qquad
+                \parbox[t]{\linewidth}{\vspace{-7em}$l > 0,        \\l' = 0, \\m = m' = 0$}
+                \\[2em]
+                %
+                %
+                %
+                \dfrac{(1 - b)^2}{4\sqrt{(2l + 1)(2l' + 1)}(1+c_0)^{l+l'+2}}
+                \\[0.75em]
+                \enspace\quad
+                \times
+                \bigg[
+                    c_0^2(2+c_0)^2 H_{l + l' + 1}^0
+                    + 4 c_0 (1 + c_0) (2 + c_0) c_1 H_{l + l' + 1}^1
+                    \\[0.75em]
+                \qquad
+                + 2 \big(2 + 3 c_0 (2 + c_0)\big)c_1^2 H_{l + l' + 1}^2
+                + 4(1 + c_0)c_1^3 H_{l + l' + 1}^3
+                \\[0.75em]
+                \qquad
+                + c_1^4 H_{l + l' + 1}^4
+                \bigg]
+                 &
+                \qquad
+                \parbox[t]{\linewidth}{\vspace{-5em}$l > 0,        \\l' > 0, \\m = m' = 0$}
                 \quad.
             \end{dcases}
+        $}
+    \raisetag{15em}
 \end{align}
+%
+The final step is to evaluate the integral in Equation~(\ref{eq:Er2}), given
+the PDF in Equation~(\ref{eq:lognormal}):
+%
+\begin{align}
+    \mathrm{E} \big[ y_{lm} y_{l'm'} \big]
+     & =
+    ?
+\end{align}
+%
+where we define
+%
+\begin{align}
+    \gamma_2 \equiv 1
+    - 2\exp\bigg[ \mu + \frac{1}{2}\nu\bigg]
+    + \exp\bigg[ 2\mu + 2\nu\bigg]
+    \quad,
+\end{align}
+%
+and the expression for $l' > 0, \, l = 0, \, m = m' = 0$ is again the same as in
+the second case, provided we make the substitution $l \rightarrow l'$.
 

### 2020-07-31 tex/ms.tex
added recursion
added figs

@@
 %
-\subsection{Preliminaries}
+\subsection{Functional form}
 %
@@
 \begin{align}
+    \label{eq:hwhm}
     \Delta\theta =
@@
 which is positive definite for all $r' > 0$.
+
+\subsection{Behavior at finite degree}
+%
+The properties of the spot expansion described in the previous section
+(in particular its monotonicity and values at the spot center and antipode)
+are inherited from the generating function of the Legendre polynomials
+(Equation~\ref{eq:gen}). These properties are therefore only rigorously
+true when the expansion is taken to spherical
+harmonic degree $l_{\mathrm{max}} = \infty$. For truncated expansions,
+a value of $r'$ in Equation~(\ref{eq:ylm0}) that is too small will lead
+to both ringing and a value at the spot center that is in general larger
+than $1 - \delta$.
+
+\begin{figure}[th!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/spot_expansion.pdf}
+        \oscaption{spot_expansion}{%
+            Polar intensity profiles for spots with different normalized
+            radii $r$ in the range $(0, 1]$, computed at spherical harmonic
+            degree $l_{\mathrm{max}} = 20$. Dotted black curves show profiles
+            corresponding to $r < 0$ (corresponding to $r' < c_0$
+            in Equation~\ref{eq:rprime}), for which the expansion does
+            not converge, leading to ringing and a smaller spot contrast.
+            \label{fig:spot_expansion}
+        }
+    \end{centering}
+\end{figure}
+
+To circumvent these issues, we impose a minimum radius
+$r'_{\mathrm{min}}$, whose value is
+a function of the degree of the expansion. Since the normalized spot radius
+$r$ is defined in the range $[0, 1]$ (Equation~\ref{eq:rprime}), this
+minimum value is equal to $c_0$. We compute $c_0 = c_0(l_{\mathrm{max}})$
+by numerically finding the
+value of $r'$ below which the fractional error in the intensity at the
+center of the spot exceeds some threshold, which we take to be 0.01.
+Since the width of the spot $\Delta\theta$ approaches $\pi$ as
+$r' \rightarrow \infty$ (Equation~\ref{eq:hwhm}), we also impose a
+maximum radius $r'_{\mathrm{max}}$, which we define to be the radius
+corresponding to
+$\Delta\theta = \nicefrac{5\pi}{12} = 75^\circ$. This choice is rather
+arbitrary, but motivated by the fact that below this value, $\Delta\theta$
+scales approximately logarithmically with $r'$ (see Figure~\ref{fig:hwhm}).
+
+\begin{figure}[th!]
+    \begin{centering}
+        \includegraphics[width=0.4615\linewidth]{figures/hwhm.pdf}
+        \includegraphics[width=0.475\linewidth]{figures/minr.pdf}
+        \oscaption{minr}{%
+            Foo bar.
+            \codelink{hwhm}
+            \label{fig:hwhm}
+        }
+    \end{centering}
+\end{figure}
 
@@
     G_j^k & =
-    \dfrac{\alpha + \beta + k - j}{2j}G_{j - 2}^k
-    - \dfrac{2\alpha + \beta + 2k - 3j}{2j}G_{j - 1}^k
+    \bigg[
+        \dfrac{(\alpha + \beta + k - j)(1 + c_0)}{j(1 + c_0 + c_1)}
+        \bigg] G_{j - 2}^k
+    \nonumber
+    \\[0.75em]
+          &
+    + \hspace{1.5pt}
+    \bigg[
+        -\dfrac{(\alpha + \beta - 2j + k)(1 + c_0) + (\alpha - j + k)c_1}
+        {j(1 + c_0 + c_1)}
+        \bigg]
+    G_{j - 1}^k
+    \\[1.5em]
+    G_j^k & =
+    \bigg[
+        \frac{(\alpha + \beta + k - 2)(1 + c_0)}
+        {\lambda_{k - 1}(\alpha + \beta - j + k - 2)c_1}
+        \bigg]
+    G_{j}^{k - 2}
+    \nonumber
+    \\[0.75em]
+          & + \hspace{1.5pt}
+    \bigg[
+        -\frac{\alpha - (\alpha + k)c_1 + \beta + k + (\alpha + \beta + k - 2)c_0 + (j + 2)c_1 - 2}
+        {\lambda_{k-1} (\alpha + \beta - j + k - 2)c_1}
+        \bigg]
+    G_{j}^{k - 1}
     \quad.
 \end{align}
-%
-\textbf{TODO: Recursion for k > 1.\\}
 %
@@
                 \qquad
-                \parbox[t]{\linewidth}{\vspace{-5em}$l > 0,        \\l' > 0, \\m = m' = 0$}
+                \parbox[t]{\linewidth}{\vspace{-6em}$l > 0,        \\l' > 0, \\m = m' = 0$}
+                \\[2em]
+                %
+                %
+                %
+                0
+                 &
+                \qquad m, m' \ne 0
                 \quad.
@@
      & =
-    ?
+    \resizebox{1.4\hsize}{!}{$
+            \begin{dcases}
+                1 + \dfrac{\gamma_1}{1+c_0}
+                \bigg[
+                    -c_0 H_0^0
+                    -
+                    c_1 H_0^1
+                    \\[0.75em]
+                \qquad\qquad\enspace\enspace\quad\quad
+                +
+                \dfrac{\gamma_2}{2\gamma_1(1+c_0)}
+                \bigg(
+                \dfrac{c_0^2}{2}H_1^0
+                +
+                c_0 c_1 H_1^1
+                +
+                \dfrac{c_1^2}{2} H_1^2
+                \bigg)
+                \bigg]
+                 &
+                \qquad
+                \parbox[t]{\linewidth}{\vspace{-3.5em}$l = l' = 0, \\m = m' = 0$}
+                \\[2em]
+                %
+                %
+                %
+                \dfrac{\gamma_1}{2\sqrt{2l + 1}(1+c_0)^{l+1}}
+                \\[0.75em]
+                \enspace\quad
+                \times
+                \bigg[
+                    c_0(2 + c_0) H_l^0 + 2 c_1(1 + c_0) H_l^1
+                    + c_1^2 H_l^2
+                    \\[0.75em]
+                \qquad
+                - \dfrac{\gamma_2}{2\gamma_1(1 + c_0)} \bigg(
+                c_0^2(2 + c_0) H_{l+1}^0
+                + c_0 c_1 (4 + 3 c_0) H_{l+1}^1
+                \\[0.75em]
+                \qquad
+                + (2 + 3 c_0) c_1^2 H_{l + 1}^2
+                + c_1^3 H_{l + 1}^3
+                \bigg)
+                \bigg]
+                 &
+                \qquad
+                \parbox[t]{\linewidth}{\vspace{-7em}$l > 0,        \\l' = 0, \\m = m' = 0$}
+                \\[2em]
+                %
+                %
+                %
+                \dfrac{\gamma_2}{4\sqrt{(2l + 1)(2l' + 1)}(1+c_0)^{l+l'+2}}
+                \\[0.75em]
+                \enspace\quad
+                \times
+                \bigg[
+                    c_0^2(2+c_0)^2 H_{l + l' + 1}^0
+                    + 4 c_0 (1 + c_0) (2 + c_0) c_1 H_{l + l' + 1}^1
+                    \\[0.75em]
+                \qquad
+                + 2 \big(2 + 3 c_0 (2 + c_0)\big)c_1^2 H_{l + l' + 1}^2
+                + 4(1 + c_0)c_1^3 H_{l + l' + 1}^3
+                \\[0.75em]
+                \qquad
+                + c_1^4 H_{l + l' + 1}^4
+                \bigg]
+                 &
+                \qquad
+                \parbox[t]{\linewidth}{\vspace{-6em}$l > 0,        \\l' > 0, \\m = m' = 0$}
+                \\[2em]
+                %
+                %
+                %
+                0
+                 &
+                \qquad m, m' \ne 0
+                \quad.
+            \end{dcases}
+        $}
 \end{align}

### 2020-07-31 tex/figures/spot_expansion.py

@@ -1,0 +1,145 @@
+import matplotlib.pyplot as plt
+from matplotlib import colors
+import numpy as np
+import starry
+from scipy.optimize import minimize
+
+starry.config.quiet = True
+
+
+def hwhm(rprime):
+    """
+    Return the half-width at half minimum as a function of r'.
+    
+    """
+    return (
+        np.arccos((2 + 3 * rprime * (2 + rprime)) / (2 * (1 + rprime) ** 3))
+        * 180
+        / np.pi
+    )
+
+
+def peak_error(ydeg, rprime):
+    """
+    Returns the error in the intensity at the spot center.
+    
+    """
+    delta = 1.0
+    I = 1 - 0.5 * delta * rprime / (1 + rprime)
+    for l in range(1, ydeg + 1):
+        I -= 0.5 * delta * rprime * (2 + rprime) / (1 + rprime) ** (l + 1)
+    return np.abs(I)
+
+
+def min_rprime(ydeg, tol=1e-2):
+    """
+    Returns the smallest value of r' such that the error on the peak
+    intensity is less than `tol`.
+
+    """
+    f = lambda rprime: (peak_error(ydeg, rprime) - tol) ** 2
+    res = minimize(f, 0.25)
+    return res.x
+
+
+def max_rprime(hmwhm_max=75):
+    """
+    Returns the value of r' corresponding to `hwhm_max`.
+    
+    """
+    f = lambda rprime: (hwhm(rprime) - hmwhm_max) ** 2
+    res = minimize(f, 10)
+    return res.x
+
+
+def get_c(ydeg, tol=1e-2, hwhm_max=75):
+    """
+    Returns the c_0, c_1 coefficients for the radius.
+
+    """
+    rmin = min_rprime(ydeg, tol=tol)
+    rmax = max_rprime(hmwhm_max=hwhm_max)
+    c0 = rmin
+    c1 = rmax - rmin
+    return c0, c1
+
+
+ydeg = 20
+map = starry.Map(ydeg, lazy=False)
+
+# Normalized radius in (0, 1]
+r = np.logspace(-2, 1, 15)
+
+# Convert to actual radius used in the expansion, r'
+c0, c1 = get_c(ydeg)
+rprime = c0 + c1 * r
+
+# Add a few values below the minimum to show the ringing
+rprime = np.append([c0 / 8, c0 / 4, c0 / 2], rprime)
+
+# Spot contrast
+delta = 1
+
+# Longitude array
+lon = np.linspace(-180, 180, 1000)
+
[…]

### 2020-08-04 tex/ms.tex
better fig
fixed figures

@@
 
-\begin{figure}[th!]
+\begin{figure}[p!]
     \begin{centering}
@@
 
-\begin{figure}[th!]
+\begin{figure}[p!]
     \begin{centering}
-        \includegraphics[width=0.4615\linewidth]{figures/hwhm.pdf}
-        \includegraphics[width=0.475\linewidth]{figures/minr.pdf}
-        \oscaption{minr}{%
-            Foo bar.
-            \codelink{hwhm}
+        \includegraphics[width=\linewidth]{figures/hwhm.pdf}
+        \oscaption{hwhm}{%
+            \emph{Left.} The width $\Delta\theta$
+            (Equation~\ref{eq:hwhm}) of the
+            spot as a function of the radius parameter $r'$ (bottom axis)
+            and the normalized radius $r$ (top axis), computed for
+            $l_{\mathrm{max}} = 20$. The green dots indicate the
+            minimum value of $r'$ for which the fractional error in the
+            spot intensity due
+            to the truncated expansion is less than one percent
+            and the value of $r'$ corresponding to
+            $\Delta\theta = 75^\circ$, respectively. These points correspond to
+            the minimum and maximum spot sizes we consider at
+            $l_{\mathrm{max}} = 20$
+            (blue shading). Note that
+            in this regime, the width of the spot scales approximately
+            logarithmically with $r'$.
+            %
+            \emph{Right.} The spot width $\Delta\theta$ below which the
+            fractional error in the intensity exceeds various thresholds
+            as a function of maximum spherical harmonic degree
+            $l_{\mathrm{max}}$. The value of $\Delta\theta$ for a tolerance
+            of one percent at $l_{\mathrm{max}} = 20$ is indicated as the
+            green dot for reference.
             \label{fig:hwhm}
@@
 %
+\vspace{1in}
+%
 \begin{align}
@@
         $}
-    \raisetag{15em}
+    \raisetag{16.5em}
 \end{align}
 %
-The final step is to evaluate the integral in Equation~(\ref{eq:Er2}), given
+
+\clearpage
+
+%
+\noindent The final step is to evaluate the integral in Equation~(\ref{eq:Er2}), given
 the PDF in Equation~(\ref{eq:lognormal}):
@@
         $}
+    \raisetag{16.5em}
 \end{align}

### 2020-08-04 tex/ms.tex
better english

@@
 %
-\subsection{Functional form}
+\subsection{Preliminaries}
+%
+\subsubsection{Functional form}
+\label{sec:spot-function}
 %
@@
 The expression in Equation~(\ref{eq:ylm0}) is convenient because it satisfies
-five important properties:
+six important properties:
 %
@@
 %
-These properties may be demonstrated by considering the expression for the
-surface intensity at a given point $(\theta, \varphi)$:
+These properties may be demonstrated by considering the
+expression for the surface intensity at a given point $(\theta, \varphi)$:
 %
@@
 
-\subsection{Behavior at finite degree}
+\subsubsection{Behavior at finite degree}
 %
@@
 \end{figure}
-
-To circumvent these issues, we impose a minimum radius
-$r'_{\mathrm{min}}$, whose value is
-a function of the degree of the expansion. Since the normalized spot radius
-$r$ is defined in the range $[0, 1]$ (Equation~\ref{eq:rprime}), this
-minimum value is equal to $c_0$. We compute $c_0 = c_0(l_{\mathrm{max}})$
-by numerically finding the
-value of $r'$ below which the fractional error in the intensity at the
-center of the spot exceeds some threshold, which we take to be 0.01.
-Since the width of the spot $\Delta\theta$ approaches $\pi$ as
-$r' \rightarrow \infty$ (Equation~\ref{eq:hwhm}), we also impose a
-maximum radius $r'_{\mathrm{max}}$, which we define to be the radius
-corresponding to
-$\Delta\theta = \nicefrac{5\pi}{12} = 75^\circ$. This choice is rather
-arbitrary, but motivated by the fact that below this value, $\Delta\theta$
-scales approximately logarithmically with $r'$ (see Figure~\ref{fig:hwhm}).
 
@@
 
-\subsection{Distributions over $r$ and $\delta$}
+To circumvent these issues, we impose a minimum radius
+$r'_{\mathrm{min}}$, whose value is
+a function of the degree of the expansion. Since the normalized spot radius
+$r$ is defined in the range $[0, 1]$ (Equation~\ref{eq:rprime}), this
+minimum value is equal to $c_0$. We compute $c_0 = c_0(l_{\mathrm{max}})$
+by numerically finding the
+value of $r'$ below which the fractional error in the intensity at the
+center of the spot exceeds some threshold, which we take to be 0.01.
+Since the width of the spot $\Delta\theta$ approaches $\pi$ as
+$r' \rightarrow \infty$ (Equation~\ref{eq:hwhm}), we also impose a
+maximum radius $r'_{\mathrm{max}}$, which we define to be the radius
+corresponding to
+$\Delta\theta = \nicefrac{5\pi}{12} = 75^\circ$. This choice is rather
+arbitrary, but motivated by the fact that below this value, $\Delta\theta$
+scales approximately logarithmically with $r'$ (see Figure~\ref{fig:hwhm}).
+
+\subsection{Probability distribution}
 %
-We model the normalized spot radius $r$ as a random variable drawn from
-a Beta distribution, whose PDF is given by
+Our goal now is to characterize the distribution of $y_{lm}$ values as a
+function of certain statistics of the spots: specifically, the distribution of
+spot radii and intensities. As we will see, if we model these with a Beta and
+a log-normal distribution, respectively, we may analytically compute the
+first two moments of the distribution of $y_{lm}$ conditioned on the parameters
+of these two prior distributions.
+
+\subsubsection{Distributions over $r$ and $\delta$}
+%
+Since we require the normalized spot radius $r$ to be bounded
+on the interval $[0, 1]$, we place a Beta prior on this quantity, whose
+probability density function (PDF) is
 %
@@
 %
-Since we require the brightness
+And since we require the brightness
 %
@@
 be non-negative, it is convenient to treat $b \in [0, \infty)$ as our
-random variable instead of $\delta$. We model $b$ as a random variable
-drawn from a log-normal
-distribution:
+random variable instead of $\delta$. We therefore place a log-normal
+prior on $b$, whose PDF is
 %
@@
 %
+In the next two sections, we derive closed form expressions for the first
+and second moments of the distribution
+%
+$p \big( y_{lm} \, \big| \, \alpha, \beta, \mu, \nu \big)$.
 
-\subsection{First moment}
+\subsubsection{First moment}
 %
@@
 
-\subsection{Second moment}
+\subsubsection{Second moment}
 %

### 2020-08-10 tex/gitlinks.tex
setup azure

@@
-\newcommand{\codelink}[1]{\href{https://github.com/user/repo/blob/fe12fa16a5cc76c1ba72d97241fb9545da144edf/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/user/repo/blob/fe12fa16a5cc76c1ba72d97241fb9545da144edf/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/user/repo/blob/fe12fa16a5cc76c1ba72d97241fb9545da144edf/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}
-\newcommand{\cilink}[1]{\href{https://dev.azure.com/user/repo/_build}{#1}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/e5ab316138ff452eef773701f33097cf0c53c3a9/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/e5ab316138ff452eef773701f33097cf0c53c3a9/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/e5ab316138ff452eef773701f33097cf0c53c3a9/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}
+\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-08-10 tex/ms.tex
a work in progress

@@
 
-%
-\section{The Spot Expansion}
+\section{Introduction}
+\label{sec:intro}
+\xxx{Talk about starry, gps, and whatnot.}
+
+\section{Spherical Harmonics}
+\label{sec:ylm}
+%
+\xxx{Introduce the spherical harmonics.}
+
+The real spherical harmonics are indexed by their degree $l \in [0, \infty]$
+and order $m \in [-l, l]$. It is convenient to collect the spherical harmonic coefficients of
+a given expansion into a vector $\mathbf{y}$ indexed by a single
+integer $n$, where
+%
+\begin{align}
+    \label{eq:n}
+    n = l^2 + l + m
+\end{align}
+%
+and, conversely,
+%
+\begin{align}
+    \label{eq:lm}
+    \begin{split}
+        l & = \floor{\sqrt{n}}
+        \\
+        m & = n - l^2 - l
+        \quad.
+    \end{split}
+\end{align}
+
+\section{Variational Inference}
+\label{sec:variational}
+%
+\xxx{Explain what we're doing and the connection between our GP and VI.}
+
+The mean $\pmb{\mu}$ and covariance $\pmb{\Sigma}$ of the flux GP are given by
+%
+\begin{align}
+    \label{eq:mean}
+    \pmb{\mu}(\pmb{\theta})
+     & = \mathrm{E} \Big[ \mathbf{f}(\pmb{\theta}) \, \Big| \, \pmb{\theta} \Big]
+    \\
+    \label{eq:cov}
+    \pmb{\Sigma}(\pmb{\theta})
+     & = \mathrm{E} \Big[ \mathbf{f}(\pmb{\theta}) \, \mathbf{f}^\top(\pmb{\theta}) \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}^2(\pmb{\theta})
+\end{align}
+%
+where $\mathrm{E}[\cdots]$ denotes the expectation value,
+$\mathbf{f}$ is a random vector-valued variable corresponding to
+a set of flux measurements, $\pmb{\theta}$ is a vector
+of hyperparameters on which the GP depends, and the squaring operation in
+Equation~(\ref{eq:cov}) is performed element-wise.
+The vector $\pmb{\theta}$ includes physical properties of the star such
+as its inclination $i$ and rotational period $P$ as well as parameters
+describing the shape of the probability density function (PDF) governing
+the distribution of features on the surface.
+%
+Since the flux
+$\mathbf{f}$ is linear in
+the spherical harmonic coefficient vector $\mathbf{y}$,
+%
+\begin{align}
+    \label{eq:fAy}
+    \mathbf{f}(\pmb{\theta}) = \mathbf{A}(\pmb{\theta}) \, \mathbf{y}
+    \quad,
+\end{align}
+%
+where $\mathbf{A}$ is the \starry design matrix \citep{Luger2019},
+we may write the mean and covariance as
+%
+\begin{align}
+    \pmb{\mu}(\pmb{\theta})
+     & = \mathbf{A}(\pmb{\theta}) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta})
+    \\
+    \pmb{\Sigma}(\pmb{\theta})
+     & = \mathbf{A}(\pmb{\theta}) \, \pmb{\Sigma}_{\mathbf{y}} \, \mathbf{A}^\top(\pmb{\theta})
+    \quad,
+\end{align}
+%
+where
+%
+\begin{align}
+    \label{eq:mean_y}
+    \pmb{\mu}_{\mathbf{y}}(\pmb{\theta})
+     & = \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta} \Big]
+    \\
+    \label{eq:cov_y}
+    \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta})
+     & = \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}_{\mathbf{y}}^2(\pmb{\theta})
+\end{align}
+%
+are the mean and covariance of the spherical harmonic GP.
+The expectation values in Equations~(\ref{eq:mean_y}) and (\ref{eq:cov_y})
+above are given by the integrals
+%
+\begin{align}
+    \label{eq:exp_y}
+    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta} \Big]
+     & =
+    \int \mathbf{y}(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta})\mathrm{d}\mathbf{x}
+    \\
+    \label{eq:exp_yy}
+    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big]
+     & =
+    \int \mathbf{y}(\mathbf{x} ) \mathbf{y}^\top(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta})\mathrm{d}\mathbf{x}
+    \quad,
+\end{align}
+%
+where $\mathbf{x}$ is a random vector-valued variable corresponding to a particular
+distribution of features on the surface and
+$p(\mathbf{x} \, \big| \, \pmb{\theta})$ is its PDF.
+
+The goal of this paper is to show that for suitable choices of $\mathbf{y}(\mathbf{x})$
+and $p(\mathbf{x} \, \big| \, \pmb{\theta})$, the integrals in the expressions
+above have closed form solutions that may be evaluated quickly.
+%
+As we are specifically interested in modeling the effect of star spots
+on stellar light curves, we let
+%
+\begin{align}
+    \mathbf{x} = \big\{ \varphi, \lambda, \delta, r \big\}
+\end{align}
+%
+and
+%
+\begin{align}
+    \label{eq:RRs}
+    \mathbf{y}(\mathbf{x}) =
+    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda)
+    \,
+    \mathbf{R}_{\hat{\mathbf{x}}}(\phi)
+    \,
+    \mathbf{s}(\delta, r)
+    \quad,
+\end{align}
+%
+where $\lambda$ is the longitude of a spot, $\phi$ is its latitude, $\delta$
+is its contrast, and $r$ is its radius. The vector function $\mathbf{s}(\delta, r)$
+returns the spherical harmonic expansion of a circular spot at $\lambda = \phi = 0$,
+$\mathbf{R}_{\hat{\mathbf{x}}}(\phi)$ is the Wigner matrix that rotates the
+expansion about $\hat{\mathbf{x}}$ such that the spot is centered at a
+latitude $\phi$, and $\mathbf{R}_{\hat{\mathbf{y}}}(\lambda)$ is the Wigner
+matrix that then rotates the
+expansion about $\hat{\mathbf{y}}$ such that the spot is centered at a
+longitude $\lambda$; these three functions are detailed in the sections below.
+%
+Equation~(\ref{eq:RRs}) thus provides a way of converting a random variable
+$\mathbf{x}$ describing the size, contrast, and position of a spot to the
+corresponding representation in terms of spherical harmonics.
+%
+Note, importantly, that we are not interested in any specific value of
+$\mathbf{y}$; rather, we would like to know its expectation value under
+the probability distribution governing the different spot properties $\mathbf{x}$,
+i.e., $p(\mathbf{x} \, \big| \, \pmb{\theta})$.
+%
+For simplicity, we assume that $p(\mathbf{x} \, \big| \, \pmb{\theta})$
+is separable in each of the four spot properties:
+%
+\begin{align}
+    p(\mathbf{x} \, \big| \, \pmb{\theta})
+    =
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi})\,
+    p(\delta \, \big| \, \pmb{\theta}_{\delta}) \,
+    p(r \, \big| \, \pmb{\theta}_{r})
+    \quad,
+\end{align}
+%
+where
+%
+\begin{align}
+    \pmb{\theta} = \big\{
+    \pmb{\theta}_{\lambda}, \,
+    \pmb{\theta}_{\phi}, \,
+    \pmb{\theta}_{\delta}, \,
+    \pmb{\theta}_{r} \big\}
+    \quad.
+\end{align}
+%
+This allows us to rewrite the expectation integrals (\ref{eq:exp_y})
+and (\ref{eq:exp_yy}) as
+%
+\begin{align}
+    \label{eq:exp_y_sep}
+    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta} \Big]
+     & =
+    \mathbf{e_3}(\pmb{\theta})
+\end{align}
+\begin{align}
+    \label{eq:exp_yy_sep}
+    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big]
+     & =
+    \mathbf{E_3}(\pmb{\theta})
+\end{align}
+%
+where we define the first moment integrals
+%
+\begin{align}
+    \mathbf{e_1}(\pmb{\theta}_r, \pmb{\theta}_\delta)
+     & =
+    \iint
+    \mathbf{s}(\delta, r) \,
+    p(r \, \big| \, \pmb{\theta}_{r}) \,
+    \mathrm{d}r \,
+    p(\delta \, \big| \, \pmb{\theta}_{\delta}) \,
+    \mathrm{d}\delta
+    %
+    \\[1em]
+    %
+    \mathbf{e_2}(\pmb{\theta}_\phi, \mathbf{S})
+     & =
+    \int
+    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
+    \mathbf{e_1} \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    %
+    \\[1em]
+    %
+    \mathbf{e_3}(\pmb{\theta}_\lambda, \mathbf{P})
+     & =
+    \int
+    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{e_2} \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+\end{align}
+%
+and the second moment integrals
+%
+\begin{align}
+    \mathbf{E_1}(\pmb{\theta}_r, \pmb{\theta}_\delta)
+     & =
+    \iint
+    \mathbf{s}(\delta, r) \, \mathbf{s}^\top(\delta, r) \,
+    p(r \, \big| \, \pmb{\theta}_{r}) \,
+    \mathrm{d}r \,
+    p(\delta \, \big| \, \pmb{\theta}_{\delta})
+    \mathrm{d}\delta
+    %
+    \\[1em]
+    %
+    \mathbf{E_2}(\pmb{\theta}_\phi, \mathbf{S})
+     & =
+    \int
+    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
+    \mathbf{E_1} \,
+    \mathbf{E_1}^\top \,
+    \mathbf{R}_{\hat{\mathbf{x}}}^\top(\phi) \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi})
+    \mathrm{d}\phi
+    %
+    \\[1em]
+    %
+    \mathbf{E_3}(\pmb{\theta}_\lambda, \mathbf{P})
+     & =
+    \int
+    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{E_2} \,
+    \mathbf{E_2}^\top \,
+    \mathbf{R}_{\hat{\mathbf{y}}}^\top(\lambda) \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda})
+    \mathrm{d}\phi
+    \quad.
+\end{align}
+%
+We devote the next three sections to the computation of these
+integrals.
+
+%
+\section{The Spot Integrals: \lowercase{$\mathbf{e_1}$} and $\mathbf{E_1}$}
 \label{sec:spot}
@@
 %
-We adopt the following expression for the spherical harmonic coefficient
-of degree $l$ and order $m$ in the expansion of a spot at
+We adopt the following expression for the $n^{\mathrm{th}}$ term in $\mathbf{s}$
+in the expansion of a spot at
 $\theta = \varphi = 0$:
@@
 \begin{align}
-    \label{eq:ylm0}
-    y_{lm}(r, \delta) =
+    \label{eq:s}
+    s_{n}(\delta, r) =
     \begin{cases}
@@
 at the center of the spot) and $r \in [0, 1]$ is the normalized spot radius.
-The expression in Equation~(\ref{eq:ylm0}) is convenient because it satisfies
+The expression in Equation~(\ref{eq:s}) is convenient because it satisfies
 six important properties:
@@
         & =
-        \sum\limits_{l=0}^\infty \sum\limits_{m=-l}^l
-        y_{lm} Y_{lm}(\theta, \varphi) \\
-        & = \sum\limits_{l=0}^\infty
-        y_{l0} \sqrt{2l + 1} P_l(\cos\theta)
+        \sum\limits_{n=0}^\infty
+        s_{n} Y_{n}(\theta, \varphi) \\
+        & =
+        \sum\limits_{n=0}^\infty
+        \begin{cases}
+            s_{n} \sqrt{2l + 1} P_l(\cos\theta)
+              & m = 0              \\[2em]
+            0 & \mathrm{otherwise}
+        \end{cases}
     \end{split}
@@
 %
-where $P_l$ is the Legendre polynomial of degree $l$ and we have implicitly
+where $Y_n$ is the spherical harmonic of degree $l(n)$ and order $m(n)$
+(see Equation~\ref{eq:lm}),
+$P_l$ is the Legendre polynomial of degree $l$ and we have implicitly
 assumed a normalization such that the integral of our expansion over
@@
 %
-Combining the above expression with Equation~(\ref{eq:ylm0}) and
+Combining the above expression with Equation~(\ref{eq:s}) and
 rearranging, we may write
@@
 harmonic degree $l_{\mathrm{max}} = \infty$. For truncated expansions,
-a value of $r'$ in Equation~(\ref{eq:ylm0}) that is too small will lead
+a value of $r'$ in Equation~(\ref{eq:s}) that is too small will lead
 to both ringing and a value at the spot center that is in general larger
@@
 %
-Our goal now is to characterize the distribution of $y_{lm}$ values as a
+Our goal now is to characterize the distribution of $s_{n}$ values as a
 function of certain statistics of the spots: specifically, the distribution of
@@
 a log-normal distribution, respectively, we may analytically compute the
-first two moments of the distribution of $y_{lm}$ conditioned on the parameters
+first two moments of the distribution of $s_{n}$ conditioned on the parameters
 of these two prior distributions.
@@
 %
-$p \big( y_{lm} \, \big| \, \alpha, \beta, \mu, \nu \big)$.
+$p \big( s_{n} \, \big| \, \alpha, \beta, \mu, \nu \big)$.
 
@@
 %
-The first moment of the distribution of $y_{lm}$ over $r$ and $b$ is
+The first moment of the distribution of $s_{n}$ over $r$ and $b$ is
 %
@@
     \begin{split}
-        \mathrm{E} \big[ y_{lm} \big] &=
+        \mathrm{E} \big[ s_{n} \big] &=
         \int\limits_0^\infty \int\limits_0^1
-        y_{lm} (r, 1 - b)
+        s_{n} (r, 1 - b)
         p \big(r, 1 - b \, \big| \, \alpha, \beta, \mu, \nu \big)
@@
         \int\limits_0^\infty
-        \mathrm{E}_r \big[ y_{lm} \big]
+        \mathrm{E}_r \big[ s_{n} \big]
         p \big(b \, \big| \, \mu, \nu \big)
@@
     %
-    \mathrm{E}_r \big[ y_{lm} \big]
+    \mathrm{E}_r \big[ s_{n} \big]
      & =
     \int\limits_0^1
-    y_{lm} (r, 1 - b)
+    s_{n} (r, 1 - b)
     p \big(r \, \big| \, \alpha, \beta \big)
@@
 \begin{align}
-    \mathrm{E}_r \big[ y_{lm} \big]
+    \mathrm{E}_r \big[ s_{n} \big]
      & =
@@
 \begin{align}
-    \mathrm{E} \big[ y_{lm} \big]
+    \mathrm{E} \big[ s_{n} \big]
      & =
@@
 %
-The second moment of the distribution of $y_{lm}$ over $r$ and $b$ is
+The second moment of the distribution of $s_{n}$ over $r$ and $b$ is
 %
@@
     \begin{split}
-        \mathrm{E} \big[ y_{lm} y_{l'm'} \big] &=
+        \mathrm{E} \big[ s_{n} s_{n'} \big] &=
         \int\limits_0^\infty \int\limits_0^1
-        y_{lm} (r, 1 - b)
-        y_{l'm'} (r, 1 - b)
+        s_{n} (r, 1 - b)
+        s_{n'} (r, 1 - b)
         p \big(r, 1 - b \, \big| \, \alpha, \beta, \mu, \nu \big)
@@
         \int\limits_0^\infty
-        \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]
+        \mathrm{E}_r \big[ s_{n} s_{n'} \big]
         p \big(b \, \big| \, \mu, \nu \big)
@@
     %
-    \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]
+    \mathrm{E}_r \big[ s_{n} s_{n'} \big]
      & =
     \int\limits_0^1
-    y_{lm} (r, 1 - b)
-    y_{l'm'} (r, 1 - b)
+    s_{n} (r, 1 - b)
+    s_{n'} (r, 1 - b)
     p \big(r \, \big| \, \alpha, \beta \big)
@@
 \begin{align}
-    \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]
+    \mathrm{E}_r \big[ s_{n} s_{n'} \big]
      & =
@@
 %
-where, by symmetry, the expression for $l' > 0, \, l = 0, \, m = m' = 0$ is the same as
+where $l' = l'(n')$ and $m' = m'(n')$ are again given by Equation~(\ref{eq:lm})
+and, by symmetry, the expression for $l' > 0, \, l = 0, \, m = m' = 0$ is the same as
 in the second case, provided we make the substitution $l \rightarrow l'$.
@@
 \begin{align}
-    \mathrm{E} \big[ y_{lm} y_{l'm'} \big]
+    \mathrm{E} \big[ s_{n} s_{n'} \big]
      & =
@@
 
-%\bibliography
+\bibliography{bib}
 

### 2020-08-11 tex/ms.tex
typo
better english

@@
 
-\section{Variational Inference}
-\label{sec:variational}
-%
-\xxx{Explain what we're doing and the connection between our GP and VI.}
-
-The mean $\pmb{\mu}$ and covariance $\pmb{\Sigma}$ of the flux GP are given by
+\section{Gaussian Process}
+\label{sec:gp}
+%
+
+% The vector $\pmb{\theta}$ includes physical properties of the star such
+% as its inclination $i$ and rotational period $P$ as well as parameters
+% describing the shape of the probability density function (PDF) governing
+% the distribution of features on the surface.
+
+Let
+$\mathbf{f} = \left( f_0 \, f_1 \, \cdots \,  f_K \right)^\top$
+denote a vector of $K$ flux measurements at times
+$\left( t_0 \,  t_1 \,  \cdots \, t_K \right)^\top$.
+Conditioned on certain physical properties of the star, $\pmb{\theta}$,
+we wish to compute the mean $\pmb{\mu}(\pmb{\theta})$ and
+covariance $\pmb{\Sigma}(\pmb{\theta})$
+of $\mathbf{f}$, which together fully specify our GP in flux.
+%
+As with any random variable, the mean and covariance may be computed from
+the expectation value of $\mathbf{f}$ and
+$\mathbf{f}\,\mathbf{f}^\top$, respectively:
 %
@@
     \pmb{\mu}(\pmb{\theta})
-     & = \mathrm{E} \Big[ \mathbf{f}(\pmb{\theta}) \, \Big| \, \pmb{\theta} \Big]
+     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \pmb{\theta} \Big]
     \\
@@
     \pmb{\Sigma}(\pmb{\theta})
-     & = \mathrm{E} \Big[ \mathbf{f}(\pmb{\theta}) \, \mathbf{f}^\top(\pmb{\theta}) \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}^2(\pmb{\theta})
-\end{align}
-%
-where $\mathrm{E}[\cdots]$ denotes the expectation value,
-$\mathbf{f}$ is a random vector-valued variable corresponding to
-a set of flux measurements, $\pmb{\theta}$ is a vector
-of hyperparameters on which the GP depends, and the squaring operation in
-Equation~(\ref{eq:cov}) is performed element-wise.
-The vector $\pmb{\theta}$ includes physical properties of the star such
-as its inclination $i$ and rotational period $P$ as well as parameters
-describing the shape of the probability density function (PDF) governing
-the distribution of features on the surface.
-%
-Since the flux
-$\mathbf{f}$ is linear in
-the spherical harmonic coefficient vector $\mathbf{y}$,
+     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}^2(\pmb{\theta})
+\end{align}
+%
+where the squaring operation in Equation~(\ref{eq:cov}) is
+performed element-wise.
+%
+In \citet{Luger2019} we showed that $\mathbf{f}$ may be computed from a
+linear operation on the vector of spherical harmonic coefficients
+describing the surface, $\mathbf{y}$:
 %
@@
     \label{eq:fAy}
-    \mathbf{f}(\pmb{\theta}) = \mathbf{A}(\pmb{\theta}) \, \mathbf{y}
-    \quad,
-\end{align}
-%
-where $\mathbf{A}$ is the \starry design matrix \citep{Luger2019},
-we may write the mean and covariance as
+    \mathbf{f} = \mathbf{A} \, \mathbf{y}
+    \quad,
+\end{align}
+%
+where $\mathbf{A}$ is the \starry design matrix, which is implicitly
+a function of $\pmb{\theta}$ (as it depends on the stellar inclination
+and rotation period, for example).
+%
+Given Equation~(\ref{eq:fAy}),
+we may write the mean and covariance of our flux GP as
 %
@@
 %
-are the mean and covariance of the spherical harmonic GP.
-The expectation values in Equations~(\ref{eq:mean_y}) and (\ref{eq:cov_y})
-above are given by the integrals
+are the mean and covariance of the GP in the spherical harmonics basis.
+The bulk of the math in this paper is devoted to computing
+the expectations in the expressions above, which
+are given by the integrals
 %
@@
 where $\mathbf{x}$ is a random vector-valued variable corresponding to a particular
-distribution of features on the surface and
-$p(\mathbf{x} \, \big| \, \pmb{\theta})$ is its PDF.
-
-The goal of this paper is to show that for suitable choices of $\mathbf{y}(\mathbf{x})$
+distribution of features on the surface  (i.e., the size and location of star spots)
+and $p(\mathbf{x} \, \big| \, \pmb{\theta})$ is its probability density
+function (PDF).
+
+In the sections that follow we will show that for suitable choices of $\mathbf{y}(\mathbf{x})$
 and $p(\mathbf{x} \, \big| \, \pmb{\theta})$, the integrals in the expressions
@@
 \begin{align}
-    \mathbf{x} = \big\{ \varphi, \lambda, \delta, r \big\}
+    \mathbf{x} = \left( \lambda \,\, \phi \,\, \xi \,\, r \right)^\top
 \end{align}
@@
     \,
-    \mathbf{s}(\delta, r)
-    \quad,
-\end{align}
-%
-where $\lambda$ is the longitude of a spot, $\phi$ is its latitude, $\delta$
-is its contrast, and $r$ is its radius. The vector function $\mathbf{s}(\delta, r)$
-returns the spherical harmonic expansion of a circular spot at $\lambda = \phi = 0$,
+    \xi
+    \,
+    \mathbf{s}(r)
+    \quad,
+\end{align}
+%
+where $\lambda$ is the longitude of a spot, $\phi$ is its latitude, $\xi$
+is its contrast, and $r$ is its radius. The vector function $\mathbf{s}(r)$
+returns the spherical harmonic expansion of a negative unit brightness
+circular spot at $\lambda = \phi = 0$,
 $\mathbf{R}_{\hat{\mathbf{x}}}(\phi)$ is the Wigner matrix that rotates the
@@
 Equation~(\ref{eq:RRs}) thus provides a way of converting a random variable
-$\mathbf{x}$ describing the size, contrast, and position of a spot to the
+$\mathbf{x}$ describing the size, brightness, and position of a spot to the
 corresponding representation in terms of spherical harmonics.
@@
     p(\phi \, \big| \, \pmb{\theta}_{\phi})\,
-    p(\delta \, \big| \, \pmb{\theta}_{\delta}) \,
+    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
     p(r \, \big| \, \pmb{\theta}_{r})
@@
 \begin{align}
-    \pmb{\theta} = \big\{
-    \pmb{\theta}_{\lambda}, \,
-    \pmb{\theta}_{\phi}, \,
-    \pmb{\theta}_{\delta}, \,
-    \pmb{\theta}_{r} \big\}
+    \pmb{\theta} = \left(
+    \pmb{\theta}_{\lambda} \, \,
+    \pmb{\theta}_{\phi} \, \,
+    \pmb{\theta}_{\xi} \, \,
+    \pmb{\theta}_{r} \right)^\top
     \quad.
@@
      & =
-    \mathbf{e_3}(\pmb{\theta})
-\end{align}
-\begin{align}
+    \mathbf{e_4}(\pmb{\theta})
+    \\[1em]
     \label{eq:exp_yy_sep}
@@
      & =
-    \mathbf{E_3}(\pmb{\theta})
+    \mathbf{E_4}(\pmb{\theta})
 \end{align}
@@
 \begin{align}
-    \mathbf{e_1}(\pmb{\theta}_r, \pmb{\theta}_\delta)
-     & =
-    \iint
-    \mathbf{s}(\delta, r) \,
+    \label{eq:e1}
+    \mathbf{e_1}(\pmb{\theta}_r)
+     & \equiv
+    \int
+    \mathbf{s}(r) \,
     p(r \, \big| \, \pmb{\theta}_{r}) \,
-    \mathrm{d}r \,
-    p(\delta \, \big| \, \pmb{\theta}_{\delta}) \,
-    \mathrm{d}\delta
+    \mathrm{d}r
     %
     \\[1em]
-    %
-    \mathbf{e_2}(\pmb{\theta}_\phi, \mathbf{S})
-     & =
+    \label{eq:e2}
+    \mathbf{e_2}(\pmb{\theta}_b, \mathbf{e_1})
+     & \equiv
+    \int
+    \xi \,
+    \mathbf{e_1} \,
+    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
+    \mathrm{d}\xi
+    %
+    \\[1em]
+    %
+    \label{eq:e3}
+    \mathbf{e_3}(\pmb{\theta}_\phi, \mathbf{e_2})
+     & \equiv
     \int
     \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{e_1} \,
+    \mathbf{e_2} \,
     p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
@@
     %
-    \mathbf{e_3}(\pmb{\theta}_\lambda, \mathbf{P})
-     & =
+    \label{eq:e4}
+    \mathbf{e_4}(\pmb{\theta}_\lambda, \mathbf{e_3})
+     & \equiv
     \int
     \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
-    \mathbf{e_2} \,
+    \mathbf{e_3} \,
     p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
@@
 \begin{align}
-    \mathbf{E_1}(\pmb{\theta}_r, \pmb{\theta}_\delta)
-     & =
-    \iint
-    \mathbf{s}(\delta, r) \, \mathbf{s}^\top(\delta, r) \,
+    \label{eq:E1}
+    \mathbf{E_1}(\pmb{\theta}_r)
+     & \equiv
+    \int
+    \mathbf{s}(r) \, \mathbf{s}^\top(r) \,
     p(r \, \big| \, \pmb{\theta}_{r}) \,
-    \mathrm{d}r \,
-    p(\delta \, \big| \, \pmb{\theta}_{\delta})
-    \mathrm{d}\delta
+    \mathrm{d}r
     %
@@
     %
-    \mathbf{E_2}(\pmb{\theta}_\phi, \mathbf{S})
-     & =
+    \label{eq:E2}
+    \mathbf{E_2}(\pmb{\theta}_b, \mathbf{E_1})
+     & \equiv
+    \int
+    \xi^2 \,
+    \mathbf{E_1} \,
+    \mathbf{E_1}^\top \,
+    p(\xi \, \big| \, \pmb{\theta}_b)
+    \mathrm{d}\xi
+    %
+    \\[1em]
+    %
+    \label{eq:E3}
+    \mathbf{E_3}(\pmb{\theta}_\phi, \mathbf{E_2})
+     & \equiv
     \int
     \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{E_1} \,
-    \mathbf{E_1}^\top \,
+    \mathbf{E_2} \,
+    \mathbf{E_2}^\top \,
     \mathbf{R}_{\hat{\mathbf{x}}}^\top(\phi) \,
@@
     %
-    \mathbf{E_3}(\pmb{\theta}_\lambda, \mathbf{P})
-     & =
+    \label{eq:E4}
+    \mathbf{E_4}(\pmb{\theta}_\lambda, \mathbf{E_3})
+     & \equiv
     \int
     \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
-    \mathbf{E_2} \,
-    \mathbf{E_2}^\top \,
+    \mathbf{E_3} \,
+    \mathbf{E_3}^\top \,
     \mathbf{R}_{\hat{\mathbf{y}}}^\top(\lambda) \,
@@
 %
-We devote the next three sections to the computation of these
+We devote the next four sections to the computation of these eight
 integrals.
@@
 %
-\section{The Spot Integrals: \lowercase{$\mathbf{e_1}$} and $\mathbf{E_1}$}
-\label{sec:spot}
-%
-\subsection{Preliminaries}
-%
-\subsubsection{Functional form}
-\label{sec:spot-function}
+\section{The Size Integrals: \lowercase{$\mathbf{e_1}$} and $\mathbf{E_1}$}
+\label{sec:size}
+%
+\subsection{Functional form}
+\label{sec:size-function}
 %
 We adopt the following expression for the $n^{\mathrm{th}}$ term in $\mathbf{s}$
-in the expansion of a spot at
-$\theta = \varphi = 0$:
+in the expansion of a dark, unit contrast circular spot at
+$\lambda = \phi = 0$:
 %
@@
     \label{eq:s}
-    s_{n}(\delta, r) =
+    s_{n}(r) =
     \begin{cases}
-        1 - \dfrac{\delta r'}{2 (1 + r')}
-         & l = m = 0    \\[2em]
-        -\dfrac{\delta r' \left( 2 + r' \right)}
+        \dfrac{r' \delta_{l0}}{2}
+        -\dfrac{r' \left( 2 + r' \right)}
         {2 \sqrt{2l + 1} (1 + r')^{l + 1}}
-         & l > 0, m = 0 \\[2em]
+         & m = 0
+        \\
         0
-         & m \ne 0
+         & \mathrm{otherwise}
     \end{cases}
@@
 where
+$l = l(n), m = m(n)$ (Equation~\ref{eq:lm}) and
+$\delta$ is the Kronecker delta. The parameter $r'$ is defined as
 %
@@
     r' \equiv c_0 + c_1 r
-\end{align}
-%
-and $c_0, c_1 > 0$ are scaling constants (defined in more detail below).
-In the expressions above,
-$\delta \in (-\infty, 1]$ is the spot contrast (the fractional
-decrease in the brightness
-at the center of the spot) and $r \in [0, 1]$ is the normalized spot radius.
+    \quad,
+\end{align}
+%
+where $c_0, c_1 > 0$ are scaling constants (defined in more detail below)
+and $r \in [0, 1]$ is the normalized spot radius.
 The expression in Equation~(\ref{eq:s}) is convenient because it satisfies
-six important properties:
+five important properties:
 %
@@
     \item The surface intensity monotonically increases away from the spot center
-    \item The surface intensity at the spot center is $1 - \delta$
-    \item The surface intensity at the antipode of the spot center is unity
+    \item The surface intensity at the spot center is $-1$
+    \item The surface intensity at the antipode of the spot center is zero
     \item The size of the spot increases monotonically with $r$
@@
 These properties may be demonstrated by considering the
-expression for the surface intensity at a given point $(\theta, \varphi)$:
+expression for the surface intensity at a given polar angle $\theta$ and
+azimuth $\varphi$:
 %
@@
         \sum\limits_{n=0}^\infty
-        \begin{cases}
-            s_{n} \sqrt{2l + 1} P_l(\cos\theta)
-              & m = 0              \\[2em]
-            0 & \mathrm{otherwise}
-        \end{cases}
+        s_{n} \sqrt{2l + 1} \delta_{m0} P_l(\cos\theta)
     \end{split}
@@
 (see Equation~\ref{eq:lm}),
-$P_l$ is the Legendre polynomial of degree $l$ and we have implicitly
+$P_l$ is the Legendre polynomial of degree $l$,
+and we have implicitly
 assumed a normalization such that the integral of our expansion over
@@
     I(\theta) =
-    1 + \dfrac{\delta r'}{2}
+    \dfrac{r'}{2}
     -
-    \dfrac{\delta r' \left( 2 + r' \right)}{2 (1 + r')}
+    \dfrac{r' \left( 2 + r' \right)}{2 (1 + r')}
     \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + r'}\right)^l P_l(\cos\theta)
@@
 %
-so we may express the intensity in the fairly simple form
+so we may express the intensity as a function of polar angle
+in the fairly simple form
 %
@@
     \begin{split}
-        A & = 1 + \dfrac{\delta r'}{2}                      \\
-        B & = \delta r' (2 + r') \sqrt{\dfrac{1}{8 + 8 r'}} \\
+        A & = \dfrac{r'}{2}                      \\
+        B & = r' (2 + r') \sqrt{\dfrac{1}{8 + 8 r'}} \\
         C & = \dfrac{1 + (1 + r')^2}{2 + 2 r'}
@@
         I_{\mathrm{min}} & = A - \dfrac{B}{\sqrt{C - 1}} \\
-        & = 1 - \delta
+        & = -1
         \quad,
@@
         I_{\mathrm{max}} & = A - \dfrac{B}{\sqrt{C + 1}} \\
-        & = 1
+        & = 0
         \quad,
@@
 \begin{align}
-    1 - \left(A - \dfrac{B}{\sqrt{C - \cos{\Delta\theta}}}\right) =
-    \dfrac{1}{2}\delta
+    A - \dfrac{B}{\sqrt{C - \cos{\Delta\theta}}} =
+    -\dfrac{1}{2}
 \end{align}
@@
 which is positive definite for all $r' > 0$.
-
-\subsubsection{Behavior at finite degree}
-%
-The properties of the spot expansion described in the previous section
-(in particular its monotonicity and values at the spot center and antipode)
-are inherited from the generating function of the Legendre polynomials
-(Equation~\ref{eq:gen}). These properties are therefore only rigorously
-true when the expansion is taken to spherical
-harmonic degree $l_{\mathrm{max}} = \infty$. For truncated expansions,
-a value of $r'$ in Equation~(\ref{eq:s}) that is too small will lead
-to both ringing and a value at the spot center that is in general larger
-than $1 - \delta$.
-
-\begin{figure}[p!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/spot_expansion.pdf}
-        \oscaption{spot_expansion}{%
-            Polar intensity profiles for spots with different normalized
-            radii $r$ in the range $(0, 1]$, computed at spherical harmonic
-            degree $l_{\mathrm{max}} = 20$. Dotted black curves show profiles
-            corresponding to $r < 0$ (corresponding to $r' < c_0$
-            in Equation~\ref{eq:rprime}), for which the expansion does
-            not converge, leading to ringing and a smaller spot contrast.
-            \label{fig:spot_expansion}
-        }
-    \end{centering}
-\end{figure}
 
@@
 
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/spot_expansion.pdf}
+        \oscaption{spot_expansion}{%
+            Polar intensity profiles for spots with different normalized
+            radii $r$ in the range $(0, 1]$, computed at spherical harmonic
+            degree $l_{\mathrm{max}} = 20$. Dotted black curves show profiles
+            corresponding to $r < 0$ (corresponding to $r' < c_0$
+            in Equation~\ref{eq:rprime}), for which the expansion does
+            not converge, leading to ringing and a smaller spot contrast.
+            \label{fig:spot_expansion}
+        }
+    \end{centering}
+\end{figure}
+
+It is important to note that the
+properties of the spot expansion described above
+(in particular its monotonicity and values at the spot center and antipode)
+are inherited from the generating function of the Legendre polynomials
+(Equation~\ref{eq:gen}). These properties are therefore only rigorously
+true when the expansion is taken to spherical
+harmonic degree $l_{\mathrm{max}} = \infty$. For truncated expansions,
+a value of $r'$ in Equation~(\ref{eq:s}) that is too small will lead
+to both ringing and a value at the spot center that is in general larger
+than $-1$.
+
 To circumvent these issues, we impose a minimum radius
@@
 center of the spot exceeds some threshold, which we take to be 0.01.
-Since the width of the spot $\Delta\theta$ approaches $\pi$ as
+Since the half width of the spot $\Delta\theta$ approaches $\nicefrac{\pi}{2}$ as
 $r' \rightarrow \infty$ (Equation~\ref{eq:hwhm}), we also impose a
@@
 arbitrary, but motivated by the fact that below this value, $\Delta\theta$
-scales approximately logarithmically with $r'$ (see Figure~\ref{fig:hwhm}).
-
-\subsection{Probability distribution}
-%
-Our goal now is to characterize the distribution of $s_{n}$ values as a
-function of certain statistics of the spots: specifically, the distribution of
-spot radii and intensities. As we will see, if we model these with a Beta and
-a log-normal distribution, respectively, we may analytically compute the
-first two moments of the distribution of $s_{n}$ conditioned on the parameters
-of these two prior distributions.
-
-\subsubsection{Distributions over $r$ and $\delta$}
-%
-Since we require the normalized spot radius $r$ to be bounded
-on the interval $[0, 1]$, we place a Beta prior on this quantity, whose
-probability density function (PDF) is
-%
-\begin{align}
-    p \big(r \, \big| \, \alpha, \beta \big)
+scales approximately logarithmically with $r'$.
+
+The left panel of Figure~\ref{fig:hwhm} shows this scaling. The minimum
+and maximum values of $r'$ for $l_{\mathrm{max}} = 20$ are indicated
+for reference; these correspond to $r = 0$ and $r = 1$, respectively
+(top axis). The right panel of the figure shows the minimum spot half width
+$\Delta\theta$ as a function of the spherical harmonic degree of the
+expansion for different error thresholds: ten percent (dashed), one
+percent (solid), and one part per thousand (dotted).
+For $l_{\mathrm{max}} = 20$, the minimum spot half width is about 17$^\circ$.
+To model spots smaller than this, one must go to larger $l_{\mathrm{max}}$;
+note, however, how slowly $\Delta\theta$ decreases above
+$l_{\mathrm{max}} = 20$.
+
+The intensity as a function of polar angle away from the spot center
+is shown in Figure~\ref{fig:spot_expansion} for a range of $r$ values.
+Three spots with $r < 0$ are also shown; note the presence of
+ringing and the fact that the intensity at the spot center is significantly
+larger than $-1$.
+
+\subsection{PDF}
+\label{sec:size-pdf}
+%
+Our goal now is to characterize the distribution of $\mathbf{s}$ as a
+function of parameters $\pmb{\theta}_r$ describing the distribution of
+spot radii. Since $r \in [0, 1]$, we choose to model $r$ as a random
+Beta-distributed variable; as we will see, this choice will allow us to
+analytically compute the first two moments of the distribution of
+$\mathbf{s}$ conditioned on $\pmb{\theta}_r$.
+
+%
+The Beta distribution in $r$ has hyperparameters
+%
+\begin{align}
+    \pmb{\theta}_r = \left(
+    \alpha \, \, \, \,
+    \beta \right)^\top
+\end{align}
+%
+and PDF given by
+%
+\begin{align}
+    \label{eq:pdf_r}
+    p \big(r \, \big| \, \pmb{\theta}_r \big)
      & =
@@
 %
-where $\Gamma$ is the Gamma function and $\alpha$ and $\beta$ are
-hyperparameters characterizing the shape of the distribution.
-%
-And since we require the brightness
-%
-\begin{align}
-    b \equiv 1 - \delta
-\end{align}
-%
-at the center of the spot to
-be non-negative, it is convenient to treat $b \in [0, \infty)$ as our
-random variable instead of $\delta$. We therefore place a log-normal
-prior on $b$, whose PDF is
-%
-\begin{align}
-    \label{eq:lognormal}
-    p \big(b \, \big| \, \mu, \nu \big)
-     & =
-    \frac{1}{b\sqrt{2\pi\nu}}
-    \exp\left[
-        -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
-        \right]
-    \quad,
-\end{align}
-%
-where $\mu$ and $\nu$ are the mean and variance of the log-normal, respectively.
-The joint distribution over $r$ and $b$ is therefore separable,
-by construction:
-%
-\begin{align}
-    p \big( r, b \, \big| \, \alpha, \beta, \mu, \nu \big)
-     & =
-    p \big(r \, \big| \, \alpha, \beta \big)
-    \,
-    p \big(b \, \big| \, \mu, \nu \big)
-    \quad.
-\end{align}
-%
-In the next two sections, we derive closed form expressions for the first
-and second moments of the distribution
-%
-$p \big( s_{n} \, \big| \, \alpha, \beta, \mu, \nu \big)$.
-
-\subsubsection{First moment}
-%
-The first moment of the distribution of $s_{n}$ over $r$ and $b$ is
-%
-\begin{align}
-    \label{eq:Erb}
-    \begin{split}
-        \mathrm{E} \big[ s_{n} \big] &=
-        \int\limits_0^\infty \int\limits_0^1
-        s_{n} (r, 1 - b)
-        p \big(r, 1 - b \, \big| \, \alpha, \beta, \mu, \nu \big)
-        \mathrm{d} r
-        \,
-        \mathrm{d} b
-        \\
-        &=
-        \int\limits_0^\infty
-        \mathrm{E}_r \big[ s_{n} \big]
-        p \big(b \, \big| \, \mu, \nu \big)
-        \mathrm{d} b
-        \quad,
-    \end{split}
-    %
-    \\
-    \intertext{where}
-    \label{eq:Er}
-    %
-    \mathrm{E}_r \big[ s_{n} \big]
-     & =
-    \int\limits_0^1
-    s_{n} (r, 1 - b)
-    p \big(r \, \big| \, \alpha, \beta \big)
-    \mathrm{d} r
-    \quad.
-\end{align}
-%
-Expanding Equation~(\ref{eq:Er}), we obtain
-%
-\begin{align}
-    \mathrm{E}_r \big[ s_{n} \big]
+with support in $0 \leq r \leq 1$,
+where $\Gamma$ is the Gamma function.
+
+\subsection{First moment}
+\label{sec:size-mom1}
+%
+The first moment of $\mathbf{s}$ is given by Equation~(\ref{eq:e1}).
+We may use Equations~(\ref{eq:s}) and (\ref{eq:pdf_r}) to write the
+$n^{\mathrm{th}}$ term of $\mathbf{{e_1}}(\pmb{\theta}_r)$ as
+%
+\begin{align}
+    {e_1}_n(\alpha, \beta)
      & =
@@
             \begin{dcases}
-                \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+                -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
                 \int\limits_0^1
-                \left(
-                1 - \dfrac{(1 - b) (c_0 + c_1 r)}{2 (1 + c_0 + c_1 r)}
-                \right)
+                \dfrac{
+                    (c_0 + c_1 r)
                     r^{\alpha - 1}
                     (1 - r)^{\beta - 1}
+                }{2 (1 + c_0 + c_1 r)}
                 \mathrm{d} r
@@
                 \int\limits_0^1
-                \dfrac{(1 - b) (c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)r^{\alpha - 1}
+                \dfrac{(c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)r^{\alpha - 1}
                 (1 - r)^{\beta - 1}}
@@
 \begin{align}
-    \mathrm{E}_r \big[ s_{n} \big]
+    {e_1}_n(\alpha, \beta)
      & =
@@
             \begin{dcases}
-                1 -
-                \dfrac{(1 - b)}{2(1 + c_0)}
+                -
+                \dfrac{1}{2(1 + c_0)}
                 \bigg[c_0H_0^0 + c_1 H_0^1\bigg]
@@
                 l = m = 0    \\[2em]
-                -\dfrac{(1 - b)}{2\sqrt{2l + 1} (1 + c_0)^{l+1}}
+                -\dfrac{1}{2\sqrt{2l + 1} (1 + c_0)^{l+1}}
                 \bigg[
@@
 %
-Note that for a given value of $\alpha$, $\beta$, and $c$, we need only compute
+Note that for fixed $\alpha$, $\beta$, $c_0$, and $c_1$, we need only compute
 $G_0^0$, $G_1^0$, $G_0^1$, and $G_1^1$ directly, since the remaining terms may be
@@
 %
-We may now evaluate the integral in Equation~(\ref{eq:Erb}), with
-PDF given by Equation~(\ref{eq:lognormal}).
-Most of the terms in the integrands are constants, so the
-integrals are straightforward to evaluate and may also
-be expressed in closed form:
-%
-\begin{align}
-    \mathrm{E} \big[ s_{n} \big]
-     & =
-    \resizebox{.75\hsize}{!}{$
-            \begin{dcases}
-                1 -
-                \dfrac{\gamma_1}{2(1 + c_0)}
-                \bigg[c_0 H_0^0 + c_1 H_0^1\bigg]
-                 &
-                \qquad
-                l = m = 0    \\[2em]
-                -\dfrac{\gamma_1}{2\sqrt{2l + 1} (1 + c_0)^{l+1}}
-                \bigg[
-                    c_0(2+c_0)H_l^0
-                    +
-                    2c_1(1 + c_0)H_l^1
-                    +
-                    c_1^2 H_l^2
-                    \bigg]
-                 &
-                \qquad
-                l > 0, m = 0 \\[2em]
-                0
-                 &
-                \qquad m \ne 0
-            \end{dcases}
-        $}
-\end{align}
-%
-where we define
-%
-\begin{align}
-    \gamma_1 \equiv 1 - \exp\left[ \mu + \frac{1}{2}\nu\right]
-    \quad.
-\end{align}
-%
-
-\subsubsection{Second moment}
-%
-The second moment of the distribution of $s_{n}$ over $r$ and $b$ is
-%
-\begin{align}
-    \label{eq:Erb2}
-    \begin{split}
-        \mathrm{E} \big[ s_{n} s_{n'} \big] &=
-        \int\limits_0^\infty \int\limits_0^1
-        s_{n} (r, 1 - b)
-        s_{n'} (r, 1 - b)
-        p \big(r, 1 - b \, \big| \, \alpha, \beta, \mu, \nu \big)
-        \mathrm{d} r
-        \,
-        \mathrm{d} b
-        \\
-        &=
-        \int\limits_0^\infty
-        \mathrm{E}_r \big[ s_{n} s_{n'} \big]
-        p \big(b \, \big| \, \mu, \nu \big)
-        \mathrm{d} b
-        \quad,
-    \end{split}
-    %
-    \\
-    \intertext{where}
-    \label{eq:Er2}
-    %
-    \mathrm{E}_r \big[ s_{n} s_{n'} \big]
-     & =
-    \int\limits_0^1
-    s_{n} (r, 1 - b)
-    s_{n'} (r, 1 - b)
-    p \big(r \, \big| \, \alpha, \beta \big)
-    \mathrm{d} r
-    \quad.
-\end{align}
-%
-Expanding Equation~(\ref{eq:Er2}), we obtain
-%
-\begin{align}
-    \mathrm{E}_r \big[ s_{n} s_{n'} \big]
-     & =
+
+%
+\vspace{1em}
+%
+
+\subsection{Second moment}
+\label{sec:size-mom2}
+%
+The second moment of $\mathbf{s}$ is given by Equation~(\ref{eq:E1}).
+We may again use Equations~(\ref{eq:s}) and (\ref{eq:pdf_r}) to write the
+term at index $(n, n')$ of $\mathbf{{E_1}}(\pmb{\theta}_r)$ as
+%
+\begin{align}
+    {E_1}_{n,n'} (\alpha, \beta) & =
     \resizebox{1.5\hsize}{!}{$
@@
                 \left(
-                1 - \dfrac{(1-b) (c_0 + c_1 r)}{2 (1 + c_0 + c_1 r)}
+                \dfrac{c_0 + c_1 r}{2 (1 + c_0 + c_1 r)}
                 \right)^2
@@
                 %
-                -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+                \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
                 \int\limits_0^1
                 \left(
-                1 - \dfrac{(1-b) (c_0 + c_1 r)}{2 (1 + c_0 + c_1 r)}
+                \dfrac{(c_0 + c_1 r)}{2 (1 + c_0 + c_1 r)}
                 \right)
                 \\
-                \qquad\qquad\quad\quad\quad\quad\quad\quad\quad\quad\quad
+                \qquad\qquad\quad\quad\quad\quad\quad
                 \times
@@
                 \left(
-                \dfrac{(1-b) (c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)}
+                \dfrac{(c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)}
                     {2 \sqrt{2l + 1} (1 + c_0 + c_1 r)^{l + 1}}
@@
                 \\[0.75em]
-                \qquad\qquad\quad\quad\quad\quad\quad\quad\quad\quad\quad
+                \qquad\qquad\quad\quad\quad\quad\quad
                 \times
@@
                 %
-                -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+                \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
                 \int\limits_0^1
                 \left(
-                \dfrac{(1-b) (c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)}
+                \dfrac{(c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)}
                     {2 \sqrt{2l + 1} (1 + c_0 + c_1 r)^{l + 1}}
@@
                 \\
-                \qquad\qquad\quad\quad\quad\quad\quad\quad\quad\quad\quad
+                \qquad\qquad\quad\quad\quad\quad\quad
                 \times
@@
                 \left(
-                \dfrac{(1-b) (c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)}
+                \dfrac{(c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)}
                     {2 \sqrt{2l' + 1} (1 + c_0 + c_1 r)^{l' + 1}}
@@
         $}
-    \raisetag{14.5em}
+    \raisetag{15.5em}
 \end{align}
@@
 %
-\vspace{1in}
-%
-\begin{align}
-    \mathrm{E}_r \big[ s_{n} s_{n'} \big]
-     & =
+\begin{align}
+    {E_1}_{n,n'} (\alpha, \beta) & =
     \resizebox{1.4\hsize}{!}{$
             \begin{dcases}
-                1 + \dfrac{1-b}{1+c_0}
-                \bigg[
-                    -c_0 H_0^0
-                    -
-                    c_1 H_0^1
-                    \\[0.75em]
-                \qquad\qquad\enspace\enspace\quad\quad
+                \dfrac{1}{4(1+c_0)^2}
+                \bigg(
+                c_0^2 H_1^0
                 +
-                \dfrac{1-b}{2(1+c_0)}
-                \bigg(
-                \dfrac{c_0^2}{2}H_1^0
+                2 c_0 c_1 H_1^1
                 +
-                c_0 c_1 H_1^1
-                +
-                \dfrac{c_1^2}{2} H_1^2
+                c_1^2 H_1^2
                 \bigg)
-                \bigg]
-                 &
-                \qquad
-                \parbox[t]{\linewidth}{\vspace{-3.5em}$l = l' = 0, \\m = m' = 0$}
+                 &
+                \qquad
+                \parbox[t]{\linewidth}{\vspace{-1.5em}$l = l' = 0, \\m = m' = 0$}
                 \\[2em]
@@
                 %
-                \dfrac{1-b}{2\sqrt{2l + 1}(1+c_0)^{l+1}}
+                \dfrac{1}{4\sqrt{2l + 1}(1+c_0)^{l+2}}
                 \\[0.75em]
@@
                 \times
-                \bigg[
-                    c_0(2 + c_0) H_l^0 + 2 c_1(1 + c_0) H_l^1
-                    + c_1^2 H_l^2
-                    \\[0.75em]
-                \qquad
-                - \dfrac{1 - b}{2(1 + c_0)} \bigg(
+                \bigg(
                 c_0^2(2 + c_0) H_{l+1}^0
@@
                 \bigg)
-                \bigg]
-                 &
-                \qquad
-                \parbox[t]{\linewidth}{\vspace{-7em}$l > 0,        \\l' = 0, \\m = m' = 0$}
+                 &
+                \qquad
+                \parbox[t]{\linewidth}{\vspace{-5.5em}$l > 0,      \\l' = 0, \\m = m' = 0$}
                 \\[2em]
@@
                 %
-                \dfrac{(1 - b)^2}{4\sqrt{(2l + 1)(2l' + 1)}(1+c_0)^{l+l'+2}}
+                \dfrac{1}{4\sqrt{(2l + 1)(2l' + 1)}(1+c_0)^{l+l'+2}}
                 \\[0.75em]
@@
         $}
-    \raisetag{16.5em}
-\end{align}
-%
-
-\clearpage
-
-%
-\noindent The final step is to evaluate the integral in Equation~(\ref{eq:Er2}), given
-the PDF in Equation~(\ref{eq:lognormal}):
-%
-\begin{align}
-    \mathrm{E} \big[ s_{n} s_{n'} \big]
-     & =
-    \resizebox{1.4\hsize}{!}{$
-            \begin{dcases}
-                1 + \dfrac{\gamma_1}{1+c_0}
-                \bigg[
-                    -c_0 H_0^0
-                    -
-                    c_1 H_0^1
-                    \\[0.75em]
-                \qquad\qquad\enspace\enspace\quad\quad
-                +
-                \dfrac{\gamma_2}{2\gamma_1(1+c_0)}
-                \bigg(
-                \dfrac{c_0^2}{2}H_1^0
-                +
-                c_0 c_1 H_1^1
-                +
-                \dfrac{c_1^2}{2} H_1^2
-                \bigg)
-                \bigg]
-                 &
-                \qquad
-                \parbox[t]{\linewidth}{\vspace{-3.5em}$l = l' = 0, \\m = m' = 0$}
-                \\[2em]
-                %
-                %
-                %
-                \dfrac{\gamma_1}{2\sqrt{2l + 1}(1+c_0)^{l+1}}
-                \\[0.75em]
-                \enspace\quad
-                \times
-                \bigg[
-                    c_0(2 + c_0) H_l^0 + 2 c_1(1 + c_0) H_l^1
-                    + c_1^2 H_l^2
-                    \\[0.75em]
-                \qquad
-                - \dfrac{\gamma_2}{2\gamma_1(1 + c_0)} \bigg(
-                c_0^2(2 + c_0) H_{l+1}^0
-                + c_0 c_1 (4 + 3 c_0) H_{l+1}^1
-                \\[0.75em]
-                \qquad
-                + (2 + 3 c_0) c_1^2 H_{l + 1}^2
-                + c_1^3 H_{l + 1}^3
-                \bigg)
-                \bigg]
-                 &
-                \qquad
-                \parbox[t]{\linewidth}{\vspace{-7em}$l > 0,        \\l' = 0, \\m = m' = 0$}
-                \\[2em]
-                %
-                %
-                %
-                \dfrac{\gamma_2}{4\sqrt{(2l + 1)(2l' + 1)}(1+c_0)^{l+l'+2}}
-                \\[0.75em]
-                \enspace\quad
-                \times
-                \bigg[
-                    c_0^2(2+c_0)^2 H_{l + l' + 1}^0
-                    + 4 c_0 (1 + c_0) (2 + c_0) c_1 H_{l + l' + 1}^1
-                    \\[0.75em]
-                \qquad
-                + 2 \big(2 + 3 c_0 (2 + c_0)\big)c_1^2 H_{l + l' + 1}^2
-                + 4(1 + c_0)c_1^3 H_{l + l' + 1}^3
-                \\[0.75em]
-                \qquad
-                + c_1^4 H_{l + l' + 1}^4
-                \bigg]
-                 &
-                \qquad
-                \parbox[t]{\linewidth}{\vspace{-6em}$l > 0,        \\l' > 0, \\m = m' = 0$}
-                \\[2em]
-                %
-                %
-                %
-                0
-                 &
-                \qquad m, m' \ne 0
-                \quad.
-            \end{dcases}
-        $}
-    \raisetag{16.5em}
-\end{align}
-%
-where we define
-%
-\begin{align}
-    \gamma_2 \equiv 1
-    - 2\exp\bigg[ \mu + \frac{1}{2}\nu\bigg]
+    \raisetag{13.5em}
+\end{align}
+%
+
+\section{The Contrast Integrals: \lowercase{$\mathbf{e_2}$} and $\mathbf{E_2}$}
+\label{sec:contrast}
+%
+\subsection{Functional form}
+\label{sec:contrast-function}
+Since the spot function $\mathbf{s}$ (Equation~\ref{eq:s}) is normalized
+to $-1$, we may simply scale it by the contrast $\xi$ to obtain a spot
+whose peak intensity is $-\xi$.
+Assuming the baseline (unspotted) stellar intensity is unity, $\xi$
+is the fractional decrease in the brightness relative to the stellar
+baseline: i.e., the spot contrast.
+%
+
+\subsection{PDF}
+\label{sec:contrast-pdf}
+We require our
+spot contrast to be no larger than $1$ (corresponding to zero intensity
+at the center of the spot when the baseline is unity) to enforce physical
+intensities
+everywhere. We have thus far implicitly assumed the spot is dark, but
+in principle bright spots may also exist
+(such as the plages observed on the Sun), corresponding to contrasts
+$\xi < 0$. We therefore require a PDF with support over $(-\infty, 1)$.
+Defining
+%
+\begin{align}
+    b(\xi) \equiv 1 - \xi
+\end{align}
+%
+to be the actual brightnesss of a unit-intensity stellar surface at
+the center of the spot,
+we can obtain the required support by modeling
+$b \in (0, \infty)$ as a log-normal
+random variable:
+%
+\begin{align}
+    \label{eq:lognormal}
+    p \big(b \, \big| \, \mu, \nu \big)
+     & =
+    \frac{1}{b\sqrt{2\pi\nu}}
+    \exp\left[
+        -\dfrac{\big(\ln b - \mu\big)^2}{2\nu}
+        \right]
+    \quad.
+\end{align}
+%
+where $\mu$ and $\nu$ are the mean and variance in $b$, respectively.
+The corresponding PDF in terms of $\xi$ is
+%
+\begin{align}
+    \label{eq:lognormal}
+    p \big(\xi \, \big| \, \pmb{\theta}_\xi \big)
+     & =
+    \frac{1}{(1 - \xi)\sqrt{2\pi\nu}}
+    \exp\left[
+        -\dfrac{\big(\ln (1 - \xi) - \mu\big)^2}{2\nu}
+        \right]
+    \quad.
+\end{align}
+%
+where the hyperparameters are
+%
+\begin{align}
+    \pmb{\theta}_\xi = \left(
+    \mu \, \, \, \,
+    \nu \right)^\top
+    \quad.
+\end{align}
+%
+
+\subsection{First moment}
+\label{sec:contrast-mom1}
+The first moment of the spot contrast distribution is given by
+Equation~(\ref{eq:e2}), which may easily be solved analytically:
+%
+\begin{align}
+    \mathbf{e_2}(\pmb{\theta}_b, \mathbf{e_1})
+     & =
+    \int_{-\infty}^1
+    \xi \,
+    \mathbf{e_1} \,
+    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
+    \mathrm{d}\xi
+    \nonumber
+    \\
+     & =
+    1 - \exp\left[ \mu + \frac{1}{2}\nu\right] \mathbf{e_1}
+    \quad.
+\end{align}
+%
+
+\subsection{Second moment}
+\label{sec:contrast-mom2}
+The second moment of the spot contrast distribution is given by
+Equation~(\ref{eq:E2}), which also evaluates to a simple closed form:
+%
+\begin{align}
+    \mathbf{E_2}(\pmb{\theta}_b, \mathbf{E_1})
+     & =
+    \int_{-\infty}^1
+    \xi^2 \,
+    \mathbf{E_1} \,
+    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
+    \mathrm{d}\xi
+    \nonumber
+    \\
+     & =
+    1 - 2\exp\bigg[ \mu + \frac{1}{2}\nu\bigg]
     + \exp\bigg[ 2\mu + 2\nu\bigg]
-    \quad,
-\end{align}
-%
-and the expression for $l' > 0, \, l = 0, \, m = m' = 0$ is again the same as in
-the second case, provided we make the substitution $l \rightarrow l'$.
+    \quad.
+\end{align}
 
 \bibliography{bib}
-
 \end{document}

### 2020-08-11 tex/gitlinks.tex
tests, text, and bugfixes

--- 2020-07-28 tex/gitlinks.tex
@@
-\newcommand{\codelink}[1]{\href{https://github.com/user/repo/blob/fe12fa16a5cc76c1ba72d97241fb9545da144edf/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/user/repo/blob/fe12fa16a5cc76c1ba72d97241fb9545da144edf/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/user/repo/blob/fe12fa16a5cc76c1ba72d97241fb9545da144edf/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}
-\newcommand{\cilink}[1]{\href{https://dev.azure.com/user/repo/_build}{#1}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_gp/blob/eb1218eef62c9812177173e82fdc6d095fc1e432/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_gp/blob/eb1218eef62c9812177173e82fdc6d095fc1e432/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_gp/blob/eb1218eef62c9812177173e82fdc6d095fc1e432/#1}{\raisebox{-0.1em}{\prooficon}}}
+\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_gp/_build}{#1}}

### 2020-08-11 starry_gp/transform.py

@@ -2,6 +2,7 @@ from .wigner import R
 import numpy as np
 import scipy
 from scipy.linalg import eigh
+from scipy.optimize import minimize
 from packaging import version
 
 # Kewyord to `eigh` changed in 1.5.0
@@ -9,6 +10,9 @@ if version.parse(scipy.__version__) < version.parse("1.5.0"):
     eigvals = "eigvals"
 else:
     eigvals = "subset_by_index"
+
+
+__all__ = ["eigen", "get_c0_c1", "get_alpha_beta", "TransformIntegral"]
 
 
 def eigen(Q, n=None):
@@ -21,6 +25,63 @@ def eigen(Q, n=None):
     w, U = eigh(Q, **{eigvals: (N - n, N - 1)})
     U = U @ np.diag(np.sqrt(np.maximum(0, w)))
     return U[:, ::-1]
+
+
+def hwhm(rprime):
+    """
+    Return the half-width at half minimum as a function of r'.
+    
+    """
+    return (
+        np.arccos((2 + 3 * rprime * (2 + rprime)) / (2 * (1 + rprime) ** 3))
+        * 180
+        / np.pi
+    )
+
+
+def peak_error(ydeg, rprime):
+    """
+    Returns the error in the intensity at the spot center.
+    
+    """
+    xi = 1.0
+    I = 1 - 0.5 * xi * rprime / (1 + rprime)
+    for l in range(1, ydeg + 1):
+        I -= 0.5 * xi * rprime * (2 + rprime) / (1 + rprime) ** (l + 1)
+    return np.abs(I)
+
+
+def min_rprime(ydeg, tol=1e-2):
+    """
+    Returns the smallest value of r' such that the error on the peak
+    intensity is less than `tol`.
+
+    """
+    f = lambda rprime: (peak_error(ydeg, rprime) - tol) ** 2
+    res = minimize(f, 0.25)
+    return res.x
+
+
+def max_rprime(hmwhm_max=75):
+    """
+    Returns the value of r' corresponding to `hwhm_max`.
+    
+    """
+    f = lambda rprime: (hwhm(rprime) - hmwhm_max) ** 2
+    res = minimize(f, 10)
+    return res.x
+
+
+def get_c0_c1(ydeg, tol=1e-2, hwhm_max=75):
+    """
+    Returns the c_0, c_1 coefficients for the radius transformation.
+
+    """
[…]

### 2020-08-11 tests/test_hypgeo.py

@@ -1,0 +1,81 @@
+import numpy as np
+from scipy.special import hyp2f1
+from starry_gp.transform import get_c0_c1
+
+
+def get_G_numerical(ydeg=10, alpha=5.0, beta=3.3):
+    """
+    Compute the G_j^k matrix numerically.
+    
+    """
+    c0, c1 = get_c0_c1(ydeg)
+    z = -c1 / (1 + c0)
+    G = np.empty((2 * ydeg + 2, 5))
+    for j in range(2 * ydeg + 2):
+        for k in range(5):
+            G[j, k] = hyp2f1(j + 1, alpha + k, alpha + beta + k, z)
+    return G
+
+
+def get_G(ydeg=10, alpha=5.0, beta=3.3):
+    """
+    Compute the G_j^k matrix recursively.
+
+    """
+    # Parameters
+    c0, c1 = get_c0_c1(ydeg)
+    z = -c1 / (1 + c0)
+    ab = alpha + beta
+    lam = np.array(
+        [
+            alpha / ab,
+            (alpha + 1) / (ab + 1),
+            (alpha + 2) / (ab + 2),
+            (alpha + 3) / (ab + 3),
+        ]
+    )
+
+    # Hypergeometric sequence
+    G = np.empty((2 * ydeg + 2, 5))
+
+    # Compute the first four terms explicitly
+    G[0, 0] = hyp2f1(1, alpha, ab, z)
+    G[0, 1] = hyp2f1(1, alpha + 1, ab + 1, z)
+    G[1, 0] = hyp2f1(2, alpha, ab, z)
+    G[1, 1] = hyp2f1(2, alpha + 1, ab + 1, z)
+
+    # Recurse upward in k
+    for j in range(2):
+        for k in range(2, 5):
+            A = (ab + k - 2) * (1 + c0)
+            B = (
+                alpha
+                - (alpha + k) * c1
+                + beta
+                + k
+                + (ab + k - 2) * c0
+                + (j + 2) * c1
+                - 2
+            )
+            C = lam[k - 1] * (ab - j + k - 2) * c1
+            G[j, k] = (A * G[j, k - 2] - B * G[j, k - 1]) / C
+
+    # Now recurse upward in j
+    for j in range(2, 2 * ydeg + 2):
+        for k in range(5):
+            A = (ab + k - j) * (1 + c0)
[…]

### 2020-08-12 tex/gitlinks.tex
unit tests for equations

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_gp/blob/eb1218eef62c9812177173e82fdc6d095fc1e432/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_gp/blob/eb1218eef62c9812177173e82fdc6d095fc1e432/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_gp/blob/eb1218eef62c9812177173e82fdc6d095fc1e432/#1}{\raisebox{-0.1em}{\prooficon}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_gp/blob/a67c59d30e9067923e402b0af7029bd5f7dac000/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_gp/blob/a67c59d30e9067923e402b0af7029bd5f7dac000/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_gp/blob/a67c59d30e9067923e402b0af7029bd5f7dac000/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_gp/_build}{#1}}

### 2020-08-12 tex/tests/run.py

@@ -1,0 +1,17 @@
+"""
+Runs a single test. Called from the Makefile.
+
+"""
+
+import pytest
+import sys
+
+assert len(sys.argv) == 2
+test_file = sys.argv[1]
+log_file = test_file.replace(".py", ".tex")
+with open(log_file, "w") as f:
+    result = pytest.main([test_file]).value
+    if result == 0:
+        print(r"\testpassicon", file=f)
+    else:
+        print(r"\testfailicon", file=f)

### 2020-08-12 tex/tests/test_hypgeo.py

@@ -1,3 +1,7 @@
+"""
+Test the recurrence relations for the hypergeometric function 2F1.
+
+"""
 import numpy as np
 from scipy.special import hyp2f1
 from starry_gp.transform import get_c0_c1
@@ -25,13 +29,12 @@ def get_G(ydeg=10, alpha=5.0, beta=3.3):
     # Parameters
     c0, c1 = get_c0_c1(ydeg)
     z = -c1 / (1 + c0)
-    ab = alpha + beta
     lam = np.array(
         [
-            alpha / ab,
-            (alpha + 1) / (ab + 1),
-            (alpha + 2) / (ab + 2),
-            (alpha + 3) / (ab + 3),
+            alpha / (alpha + beta),
+            (alpha + 1) / (alpha + beta + 1),
+            (alpha + 2) / (alpha + beta + 2),
+            (alpha + 3) / (alpha + beta + 3),
         ]
     )
 
@@ -39,34 +42,30 @@ def get_G(ydeg=10, alpha=5.0, beta=3.3):
     G = np.empty((2 * ydeg + 2, 5))
 
     # Compute the first four terms explicitly
-    G[0, 0] = hyp2f1(1, alpha, ab, z)
-    G[0, 1] = hyp2f1(1, alpha + 1, ab + 1, z)
-    G[1, 0] = hyp2f1(2, alpha, ab, z)
-    G[1, 1] = hyp2f1(2, alpha + 1, ab + 1, z)
+    G[0, 0] = hyp2f1(1, alpha, alpha + beta, z)
+    G[0, 1] = hyp2f1(1, alpha + 1, alpha + beta + 1, z)
+    G[1, 0] = hyp2f1(2, alpha, alpha + beta, z)
+    G[1, 1] = hyp2f1(2, alpha + 1, alpha + beta + 1, z)
 
     # Recurse upward in k
     for j in range(2):
         for k in range(2, 5):
-            A = (ab + k - 2) * (1 + c0)
-            B = (
-                alpha
-                - (alpha + k) * c1
-                + beta
-                + k
-                + (ab + k - 2) * c0
-                + (j + 2) * c1
-                - 2
+            A = ((alpha + beta + k - 2) * (1 + c0)) / (
+                (alpha + beta - j + k - 2) * lam[k - 1] * c1
             )
-            C = lam[k - 1] * (ab - j + k - 2) * c1
[…]

### 2020-08-13 tex/gitlinks.tex
contrast integral

--- 2020-08-11 tex/gitlinks.tex
@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_gp/blob/eb1218eef62c9812177173e82fdc6d095fc1e432/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_gp/blob/eb1218eef62c9812177173e82fdc6d095fc1e432/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_gp/blob/eb1218eef62c9812177173e82fdc6d095fc1e432/#1}{\raisebox{-0.1em}{\prooficon}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_gp/blob/770b39c492f27102f9a66dc6061f06370f0306e9/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_gp/blob/770b39c492f27102f9a66dc6061f06370f0306e9/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_gp/blob/770b39c492f27102f9a66dc6061f06370f0306e9/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_gp/_build}{#1}}

### 2020-08-13 starry_gp/latitude.py

@@ -6,11 +6,13 @@ import numpy as np
 class LatitudeIntegral(TransformIntegral):
     def __init__(self, ydeg):
         self._c_op = LatitudeGP(ydeg)
-        super().__init__(ydeg, cos_alpha=0, sin_alpha=1, cos_gamma=0, sin_gamma=-1)
+        super().__init__(
+            ydeg, cos_alpha=0, sin_alpha=1, cos_gamma=0, sin_gamma=-1
+        )
         self.set_params()
 
-    def _compute_basis_integrals(self, mu_lat=0.5, nu_lat=0.01):
-        alpha, beta = get_alpha_beta(mu_lat, nu_lat)
+    def _compute_basis_integrals(self, mu=0.5, nu=0.01):
+        alpha, beta = get_alpha_beta(mu, nu)
         self._c_op.compute(alpha, beta)
         self.q = self._c_op.q
         self.Q = self._c_op.Q

### 2020-08-13 tests/test_contrast.py

@@ -1,0 +1,40 @@
+from starry_gp.contrast import ContrastIntegral
+import numpy as np
+from scipy.stats import lognorm as LogNormal
+
+
+def test_contrast(ydeg=5, mu=-0.1, nu=0.01):
+
+    # Settings
+    np.random.seed(0)
+    nsamples = int(1e8)
+    atol = 2.0e-5
+
+    # Random input moments
+    np.random.seed(0)
+    N = (ydeg + 1) ** 2
+    s = np.random.randn(N)
+    eigS = np.random.randn(N, N) / N
+    S = eigS @ eigS.T
+
+    # Get analytic integral
+    I = ContrastIntegral(ydeg)
+    I.set_params(mu, nu)
+    e = I.first_moment(s)
+    eigE = I.second_moment(eigS)
+    E = eigE @ eigE.T
+
+    # Integrate numerically
+    y = np.zeros((N, nsamples))
+
+    # Draw the spot contrast
+    b = LogNormal.rvs(scale=np.exp(mu), s=np.sqrt(nu), size=nsamples)
+    xi = 1 - b
+
+    # Compute the moments
+    e_num = np.mean(xi) * s
+    E_num = np.mean(xi ** 2) * S
+
+    # Compare
+    assert np.allclose(e, e_num, atol=atol), "error in first moment"
+    assert np.allclose(E, E_num, atol=atol), "error in second moment"

### 2020-08-18 tex/ms.tex
tests, text, and bugfixes
unit tests for equations
contrast integral
changing spot param slightly

@@
 \begin{align}
-    \mathbf{x} = \left( \lambda \,\, \phi \,\, \xi \,\, r \right)^\top
+    \mathbf{x} = \left( \xi \,\, \lambda \,\, \phi \,\, \rho \right)^\top
 \end{align}
@@
     \mathbf{y}(\mathbf{x}) =
+    \xi
+    \,
     \mathbf{R}_{\hat{\mathbf{y}}}(\lambda)
@@
     \,
-    \xi
-    \,
-    \mathbf{s}(r)
-    \quad,
-\end{align}
-%
-where $\lambda$ is the longitude of a spot, $\phi$ is its latitude, $\xi$
-is its contrast, and $r$ is its radius. The vector function $\mathbf{s}(r)$
+    \mathbf{s}(\rho)
+    \quad,
+\end{align}
+%
+where $\xi$ is the contrast of a spot,
+$\lambda$ is its longitude, $\phi$ is its latitude,
+and $\rho$ is its radius.
+The vector function $\mathbf{s}(\rho)$
 returns the spherical harmonic expansion of a negative unit brightness
@@
     =
+    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
     p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
     p(\phi \, \big| \, \pmb{\theta}_{\phi})\,
-    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
-    p(r \, \big| \, \pmb{\theta}_{r})
+    p(\rho \, \big| \, \pmb{\theta}_{\rho})
     \quad,
@@
     \pmb{\theta} = \left(
+    \pmb{\theta}_{\xi} \, \,
     \pmb{\theta}_{\lambda} \, \,
     \pmb{\theta}_{\phi} \, \,
-    \pmb{\theta}_{\xi} \, \,
-    \pmb{\theta}_{r} \right)^\top
+    \pmb{\theta}_{\rho} \right)^\top
     \quad.
@@
     \label{eq:e1}
-    \mathbf{e_1}(\pmb{\theta}_r)
+    \mathbf{e_1}(\pmb{\theta}_\rho)
      & \equiv
     \int
-    \mathbf{s}(r) \,
-    p(r \, \big| \, \pmb{\theta}_{r}) \,
-    \mathrm{d}r
+    \mathbf{s}(\rho) \,
+    p(\rho \, \big| \, \pmb{\theta}_{\rho}) \,
+    \mathrm{d}\rho
     %
     \\[1em]
+    %
     \label{eq:e2}
-    \mathbf{e_2}(\pmb{\theta}_b, \mathbf{e_1})
+    \mathbf{e_2}(\pmb{\theta}_\phi, \mathbf{e_1})
+     & \equiv
+    \int
+    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
+    \mathbf{e_1} \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    %
+    \\[1em]
+    %
+    \label{eq:e3}
+    \mathbf{e_3}(\pmb{\theta}_\lambda, \mathbf{e_2})
+     & \equiv
+    \int
+    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{e_2} \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    \\[1em]
+    \label{eq:e4}
+    \mathbf{e_4}(\pmb{\theta}_\xi, \mathbf{e_3})
      & \equiv
@@
     \xi \,
-    \mathbf{e_1} \,
+    \mathbf{e_3} \,
     p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
@@
     %
+\end{align}
+%
+and the second moment integrals
+%
+\begin{align}
+    \label{eq:E1}
+    \mathbf{E_1}(\pmb{\theta}_\rho)
+     & \equiv
+    \int
+    \mathbf{s}(\rho) \, \mathbf{s}^\top(\rho) \,
+    p(\rho \, \big| \, \pmb{\theta}_{\rho}) \,
+    \mathrm{d}\rho
+    %
     \\[1em]
     %
-    \label{eq:e3}
-    \mathbf{e_3}(\pmb{\theta}_\phi, \mathbf{e_2})
+    \label{eq:E2}
+    \mathbf{E_2}(\pmb{\theta}_\phi, \mathbf{E_1})
      & \equiv
@@
     \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{e_2} \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathbf{E_1} \,
+    \mathbf{E_1}^\top \,
+    \mathbf{R}_{\hat{\mathbf{x}}}^\top(\phi) \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi})
     \mathrm{d}\phi
@@
     %
-    \label{eq:e4}
-    \mathbf{e_4}(\pmb{\theta}_\lambda, \mathbf{e_3})
+    \label{eq:E3}
+    \mathbf{E_3}(\pmb{\theta}_\lambda, \mathbf{E_2})
      & \equiv
@@
     \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
-    \mathbf{e_3} \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
-\end{align}
-%
-and the second moment integrals
-%
-\begin{align}
-    \label{eq:E1}
-    \mathbf{E_1}(\pmb{\theta}_r)
-     & \equiv
-    \int
-    \mathbf{s}(r) \, \mathbf{s}^\top(r) \,
-    p(r \, \big| \, \pmb{\theta}_{r}) \,
-    \mathrm{d}r
-    %
+    \mathbf{E_2} \,
+    \mathbf{E_2}^\top \,
+    \mathbf{R}_{\hat{\mathbf{y}}}^\top(\lambda) \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda})
+    \mathrm{d}\phi
     \\[1em]
     %
-    \label{eq:E2}
-    \mathbf{E_2}(\pmb{\theta}_b, \mathbf{E_1})
+    \label{eq:E4}
+    \mathbf{E_4}(\pmb{\theta}_\xi, \mathbf{E_3})
      & \equiv
@@
     \xi^2 \,
-    \mathbf{E_1} \,
-    \mathbf{E_1}^\top \,
-    p(\xi \, \big| \, \pmb{\theta}_b)
-    \mathrm{d}\xi
-    %
-    \\[1em]
-    %
-    \label{eq:E3}
-    \mathbf{E_3}(\pmb{\theta}_\phi, \mathbf{E_2})
-     & \equiv
-    \int
-    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{E_2} \,
-    \mathbf{E_2}^\top \,
-    \mathbf{R}_{\hat{\mathbf{x}}}^\top(\phi) \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi})
-    \mathrm{d}\phi
-    %
-    \\[1em]
-    %
-    \label{eq:E4}
-    \mathbf{E_4}(\pmb{\theta}_\lambda, \mathbf{E_3})
-     & \equiv
-    \int
-    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
     \mathbf{E_3} \,
     \mathbf{E_3}^\top \,
-    \mathbf{R}_{\hat{\mathbf{y}}}^\top(\lambda) \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda})
-    \mathrm{d}\phi
+    p(\xi \, \big| \, \pmb{\theta}_\xi)
+    \mathrm{d}\xi
+    %
     \quad.
@@
 \subsection{Functional form}
+%
+
+\subsubsection{$l_{\mathrm{max}} = \infty$}
 \label{sec:size-function}
@@
 \begin{align}
-    \label{eq:s}
-    s_{n}(r) =
+    \label{eq:sinf}
+    s_{n}^\infty(\rho) =
     \begin{cases}
-        \dfrac{r' \delta_{l0}}{2}
-        -\dfrac{r' \left( 2 + r' \right)}
-        {2 \sqrt{2l + 1} (1 + r')^{l + 1}}
-         & m = 0
-        \\
+        -\dfrac{\rho}{2(1 + \rho)}
+         & l = m = 0
+        \\[1em]
+        -\dfrac{\rho \left( 2 + \rho \right)}
+        {2 \sqrt{2l + 1} (1 + \rho)^{l + 1}}
+         & l > 0, \, m = 0
+        \\[1em]
         0
@@
 where
-$l = l(n), m = m(n)$ (Equation~\ref{eq:lm}) and
-$\delta$ is the Kronecker delta. The parameter $r'$ is defined as
-%
-\begin{align}
-    \label{eq:rprime}
-    r' \equiv c_0 + c_1 r
-    \quad,
-\end{align}
-%
-where $c_0, c_1 > 0$ are scaling constants (defined in more detail below)
-and $r \in [0, 1]$ is the normalized spot radius.
-The expression in Equation~(\ref{eq:s}) is convenient because it satisfies
+$l = l(n), m = m(n)$ (Equation~\ref{eq:lm}),
+$\delta$ is the Kronecker delta,
+$\rho \in [0, \infty)$ is the spot radius,
+and the $\infty$ superscript denotes the fact that this expression
+is valid in the limit $l_{\mathrm{max}} \rightarrow \infty$
+(more on this below).
+Equation~(\ref{eq:sinf}) is convenient because it satisfies
 five important properties:
@@
     \item The surface intensity at the antipode of the spot center is zero
-    \item The size of the spot increases monotonically with $r$
+    \item The size of the spot increases monotonically with $\rho$
 \end{enumerate}
@@
         \sum\limits_{n=0}^\infty
-        s_{n} Y_{n}(\theta, \varphi) \\
+        s_{n}^\infty Y_{n}(\theta, \varphi) \\
         & =
         \sum\limits_{n=0}^\infty
-        s_{n} \sqrt{2l + 1} \delta_{m0} P_l(\cos\theta)
+        s_{n}^\infty \sqrt{2l + 1} \delta_{m0} P_l(\cos\theta)
     \end{split}
@@
 %
-Combining the above expression with Equation~(\ref{eq:s}) and
+Combining the above expression with Equation~(\ref{eq:sinf}) and
 rearranging, we may write
@@
     I(\theta) =
-    \dfrac{r'}{2}
+    \dfrac{\rho}{2}
     -
-    \dfrac{r' \left( 2 + r' \right)}{2 (1 + r')}
-    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + r'}\right)^l P_l(\cos\theta)
+    \dfrac{\rho \left( 2 + \rho \right)}{2 (1 + \rho)}
+    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + \rho}\right)^l P_l(\cos\theta)
     \quad.
@@
     \begin{split}
-        A & = \dfrac{r'}{2}                      \\
-        B & = r' (2 + r') \sqrt{\dfrac{1}{8 + 8 r'}} \\
-        C & = \dfrac{1 + (1 + r')^2}{2 + 2 r'}
+        A & = \dfrac{\rho}{2}                      \\
+        B & = \rho (2 + \rho) \sqrt{\dfrac{1}{8 + 8 \rho}} \\
+        C & = \dfrac{1 + (1 + \rho)^2}{2 + 2 \rho}
     \end{split}
@@
     \Delta\theta =
-    \cos^{-1} \left[ \dfrac{2 + 3 r' (2 + r')}{2 (1 + r')^3} \right]
-    \quad.
-\end{align}
-%
-Differentiation with respect to $r'$ yields
-%
-\begin{align}
-    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}r'} =
-    \frac{3 r' \left(2 + r'\right)}{\left(1 + r'\right)^4
-        \sqrt{\frac{r'^2 \left(2 + r'\right)^2 \left(1 + 2 r'\right)
-                \left(3 + 2 r'\right)}{\left(1 + r'\right)^6}}}
-    \quad,
-\end{align}
-%
-which is positive definite for all $r' > 0$.
-
-\begin{figure}[p!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/hwhm.pdf}
-        \oscaption{hwhm}{%
-            \emph{Left.} The width $\Delta\theta$
-            (Equation~\ref{eq:hwhm}) of the
-            spot as a function of the radius parameter $r'$ (bottom axis)
-            and the normalized radius $r$ (top axis), computed for
-            $l_{\mathrm{max}} = 20$. The green dots indicate the
-            minimum value of $r'$ for which the fractional error in the
-            spot intensity due
-            to the truncated expansion is less than one percent
-            and the value of $r'$ corresponding to
-            $\Delta\theta = 75^\circ$, respectively. These points correspond to
-            the minimum and maximum spot sizes we consider at
-            $l_{\mathrm{max}} = 20$
-            (blue shading). Note that
-            in this regime, the width of the spot scales approximately
-            logarithmically with $r'$.
-            %
-            \emph{Right.} The spot width $\Delta\theta$ below which the
-            fractional error in the intensity exceeds various thresholds
-            as a function of maximum spherical harmonic degree
-            $l_{\mathrm{max}}$. The value of $\Delta\theta$ for a tolerance
-            of one percent at $l_{\mathrm{max}} = 20$ is indicated as the
-            green dot for reference.
-            \label{fig:hwhm}
-        }
-    \end{centering}
-\end{figure}
+    \cos^{-1} \left[ \dfrac{2 + 3 \rho (2 + \rho)}{2 (1 + \rho)^3} \right]
+    \quad.
+\end{align}
+%
+Differentiation with respect to $r$ yields
+%
+\begin{align}
+    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}\rho} =
+    \frac{3}{\left(1 + \rho\right)
+        \sqrt{\left(1 + 2 \rho\right)
+            \left(3 + 2 \rho\right)}}
+    \quad,
+\end{align}
+%
+which is positive definite for all $\rho > 0$.
 
@@
         \oscaption{spot_expansion}{%
+            \emph{Top.}
             Polar intensity profiles for spots with different normalized
-            radii $r$ in the range $(0, 1]$, computed at spherical harmonic
-            degree $l_{\mathrm{max}} = 20$. Dotted black curves show profiles
-            corresponding to $r < 0$ (corresponding to $r' < c_0$
-            in Equation~\ref{eq:rprime}), for which the expansion does
-            not converge, leading to ringing and a smaller spot contrast.
+            radii $\rho$ in the range $(0, 1]$, computed at spherical harmonic
+            degree $l_{\mathrm{max}} = \LMAX$.
+            \emph{Bottom.}
+            The width $\Delta\theta$
+            (Equation~\ref{eq:hwhm}) of the
+            spot as a function of the radius parameter $r$ (left panel)
+            and the normalized radius $\rho$ (right panel), computed for
+            $l_{\mathrm{max}} = \LMAX$.
             \label{fig:spot_expansion}
@@
 
+\subsubsection{Finite $l_{\mathrm{max}}$}
+%
 It is important to note that the
@@
 true when the expansion is taken to spherical
-harmonic degree $l_{\mathrm{max}} = \infty$. For truncated expansions,
-a value of $r'$ in Equation~(\ref{eq:s}) that is too small will lead
-to both ringing and a value at the spot center that is in general larger
-than $-1$.
-
-To circumvent these issues, we impose a minimum radius
-$r'_{\mathrm{min}}$, whose value is
-a function of the degree of the expansion. Since the normalized spot radius
-$r$ is defined in the range $[0, 1]$ (Equation~\ref{eq:rprime}), this
-minimum value is equal to $c_0$. We compute $c_0 = c_0(l_{\mathrm{max}})$
-by numerically finding the
-value of $r'$ below which the fractional error in the intensity at the
-center of the spot exceeds some threshold, which we take to be 0.01.
-Since the half width of the spot $\Delta\theta$ approaches $\nicefrac{\pi}{2}$ as
-$r' \rightarrow \infty$ (Equation~\ref{eq:hwhm}), we also impose a
-maximum radius $r'_{\mathrm{max}}$, which we define to be the radius
-corresponding to
-$\Delta\theta = \nicefrac{5\pi}{12} = 75^\circ$. This choice is rather
-arbitrary, but motivated by the fact that below this value, $\Delta\theta$
-scales approximately logarithmically with $r'$.
-
-The left panel of Figure~\ref{fig:hwhm} shows this scaling. The minimum
-and maximum values of $r'$ for $l_{\mathrm{max}} = 20$ are indicated
-for reference; these correspond to $r = 0$ and $r = 1$, respectively
-(top axis). The right panel of the figure shows the minimum spot half width
-$\Delta\theta$ as a function of the spherical harmonic degree of the
-expansion for different error thresholds: ten percent (dashed), one
-percent (solid), and one part per thousand (dotted).
-For $l_{\mathrm{max}} = 20$, the minimum spot half width is about 17$^\circ$.
-To model spots smaller than this, one must go to larger $l_{\mathrm{max}}$;
-note, however, how slowly $\Delta\theta$ decreases above
-$l_{\mathrm{max}} = 20$.
-
-The intensity as a function of polar angle away from the spot center
-is shown in Figure~\ref{fig:spot_expansion} for a range of $r$ values.
-Three spots with $r < 0$ are also shown; note the presence of
-ringing and the fact that the intensity at the spot center is significantly
-larger than $-1$.
+harmonic degree $l_{\mathrm{max}} = \infty$. Expansions truncated to finite
+$l_{\mathrm{max}}$ will in general experience ringing and a spot amplitude
+whose (absolute) value is in general smaller than unity.
+
+To mitigate these issues, we will do two things: we will impose a minimum
+radius $r_{\mathrm{min}}$ to prevent highly oscillatory behavior within
+the domain of our function; and we will scale the spot expansion
+by an overall amplitude factor so that
+the peak spot intensity is as close to (negative) unity as possible
+everywhere within the domain.
+We find that this can be accomplished by expressing the
+$n^{\mathrm{th}}$ term in the spot expansion $\mathbf{s}$ as
+%
+\begin{align}
+    s_n(\rho) = s_n^\infty\left(r\right) \left(1 + c_2 r^{-1} \left(1 + r\right) ^ {-c_3}\right)
+\end{align}
+%
+where
+%
+\begin{align}
+    r \equiv c_0 + c_1\rho
+\end{align}
+%
+is a rescaled radius and $c_0, c_1, c_2, c_3$ are constants at a given
+value of $l_{\mathrm{max}}$.
+%
+Plugging this into Equation~(\ref{eq:sinf}), we have
+%
+\begin{align}
+    \label{eq:s}
+    s_{n}(\rho) =
+    \begin{cases}
+        -\dfrac{r}{2(1 + r)} - \dfrac{c_2}{2(1 + r)^{c_3 + 1}}
+         & l = m = 0
+        \\[1em]
+        -\dfrac{r \left( 2 + r \right)}
+        {2 \sqrt{2l + 1} (1 + r)^{l + 1}}
+        - \dfrac{c_2\left( 2 + r \right)}
+        {2 \sqrt{2l + 1} (1 + r)^{c_3 + l + 1}}
+         & l > 0, \, m = 0
+        \\[1em]
+        0
+         & \mathrm{otherwise}
+    \end{cases}
+\end{align}
+%
+Note that in the limit $l_{\mathrm{max}} \rightarrow \infty$, we may set
+$c_0 = 0$, $c_1 = 1$, and $c_2 = 0$ to recover the original parametrization.
+At finite $l_{\mathrm{max}}$, however, we solve for the $c_n$ numerically,
+by requiring the following conditions:
+
+\xxx{TODO...}
 
@@
 The first moment of $\mathbf{s}$ is given by Equation~(\ref{eq:e1}).
-We may use Equations~(\ref{eq:s}) and (\ref{eq:pdf_r}) to write the
+We may use Equations~(\ref{eq:sinf}) and (\ref{eq:pdf_r}) to write the
 $n^{\mathrm{th}}$ term of $\mathbf{{e_1}}(\pmb{\theta}_r)$ as
@@
 %
-\begin{align}
+\begin{proof}{test_hypgeo}
     \label{eq:Grec}
+    \begin{split}
         G_j^k & =
@@
             \bigg] G_{j - 2}^k
-    \nonumber
         \\[0.75em]
@@
         \bigg[
-        -\dfrac{(\alpha + \beta - 2j + k)(1 + c_0) + (\alpha - j + k)c_1}
+            1 - \dfrac{(\alpha + \beta + k - j)(1 + c_0) + (\alpha + k)c_1}
             {j(1 + c_0 + c_1)}
@@
             \frac{(\alpha + \beta + k - 2)(1 + c_0)}
-        {\lambda_{k - 1}(\alpha + \beta - j + k - 2)c_1}
+            {(\alpha + \beta - j + k - 2)\lambda_{k - 1}c_1}
             \bigg]
         G_{j}^{k - 2}
-    \nonumber
         \\[0.75em]
@@
         \bigg[
-        -\frac{\alpha - (\alpha + k)c_1 + \beta + k + (\alpha + \beta + k - 2)c_0 + (j + 2)c_1 - 2}
-        {\lambda_{k-1} (\alpha + \beta - j + k - 2)c_1}
+            \frac{1}{\lambda_{k - 1}}
+            -\frac{(\alpha + \beta + k - 2)(1 + c_0) + \beta c_1}
+            {(\alpha + \beta - j + k - 2)\lambda_{k-1}c_1}
             \bigg]
@@
         \quad.
-\end{align}
+    \end{split}
+    \raisetag{7.5em}
+\end{proof}
 %
@@
 The second moment of $\mathbf{s}$ is given by Equation~(\ref{eq:E1}).
-We may again use Equations~(\ref{eq:s}) and (\ref{eq:pdf_r}) to write the
+We may again use Equations~(\ref{eq:sinf}) and (\ref{eq:pdf_r}) to write the
 term at index $(n, n')$ of $\mathbf{{E_1}}(\pmb{\theta}_r)$ as
@@
 
-\section{The Contrast Integrals: \lowercase{$\mathbf{e_2}$} and $\mathbf{E_2}$}
+\section{The Latitude Integrals: \lowercase{$\mathbf{e_2}$} and $\mathbf{E_2}$}
+\label{sec:latitude}
+\xxx{TODO.}
+
+\section{The Longitude Integrals: \lowercase{$\mathbf{e_3}$} and $\mathbf{E_3}$}
+\label{sec:longitude}
+\xxx{TODO.}
+
+\section{The Contrast Integrals: \lowercase{$\mathbf{e_4}$} and $\mathbf{E_4}$}
 \label{sec:contrast}
@@
 \label{sec:contrast-function}
-Since the spot function $\mathbf{s}$ (Equation~\ref{eq:s}) is normalized
+Since the spot function $\mathbf{s}$ (Equation~\ref{eq:sinf}) is normalized
 to $-1$, we may simply scale it by the contrast $\xi$ to obtain a spot
@@
 The first moment of the spot contrast distribution is given by
-Equation~(\ref{eq:e2}), which may easily be solved analytically:
-%
-\begin{align}
-    \mathbf{e_2}(\pmb{\theta}_b, \mathbf{e_1})
+Equation~(\ref{eq:e4}), which may easily be solved analytically:
+%
+\begin{align}
+    \mathbf{e_4}(\pmb{\theta}_\xi, \mathbf{e_3})
      & =
@@
      & =
-    1 - \exp\left[ \mu + \frac{1}{2}\nu\right] \mathbf{e_1}
+    \left(1 - \exp\left[ \mu + \frac{1}{2}\nu\right]\right) \mathbf{e_3}
     \quad.
@@
 The second moment of the spot contrast distribution is given by
-Equation~(\ref{eq:E2}), which also evaluates to a simple closed form:
-%
-\begin{align}
-    \mathbf{E_2}(\pmb{\theta}_b, \mathbf{E_1})
+Equation~(\ref{eq:E4}), which also evaluates to a simple closed form:
+%
+\begin{align}
+    \mathbf{E_4}(\pmb{\theta}_\xi, \mathbf{E_3})
      & =
@@
     \xi^2 \,
-    \mathbf{E_1} \,
+    \mathbf{E_3} \,
+    \mathbf{E_3}^\top \,
     p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
@@
      & =
-    1 - 2\exp\bigg[ \mu + \frac{1}{2}\nu\bigg]
-    + \exp\bigg[ 2\mu + 2\nu\bigg]
+    \left(1 - 2\exp\bigg[ \mu + \frac{1}{2}\nu\bigg]
+    + \exp\bigg[ 2\mu + 2\nu\bigg]\right)
+    \mathbf{E_3} \,
+    \mathbf{E_3}^\top
     \quad.

### 2020-08-18 starry_gp/integrals.py

@@ -62,11 +62,31 @@ class MomentIntegral(object):
 
 
 class WignerIntegral(MomentIntegral):
+    def __init__(self, parent=None, ydeg=None, **kwargs):
+
+        # Stability fix
+        if ydeg is None:
+            if parent is not None:
+                ydeg = parent.ydeg
+            else:
+                assert False, "please provide `ydeg`."
+        if ydeg > 15:
+            # This will slow things down quite a bit,
+            # but is necessary since otherwise things go
+            # unstable in the eigendecomposition
+            self.neig = (ydeg + 1) ** 2
+            self.driver = "ev"
+        else:
+            self.neig = 2 * ydeg + 1
+            self.driver = None
+
+        super().__init__(parent=parent, ydeg=ydeg, **kwargs)
+
     def _compute_basis_integrals(self, *args, **kwargs):
         raise NotImplementedError("Must be subclassed.")
 
     def _compute_U(self):
-        self.U = eigen(self.Q, self.n)
+        self.U = eigen(self.Q, self.neig, driver=self.driver)
 
     def _compute_t(self):
         self.t = [np.zeros((self.n, self.n)) for l in range(self.ydeg + 1)]
@@ -75,7 +95,7 @@ class WignerIntegral(MomentIntegral):
 
     def _compute_T(self):
         self.T = [
-            np.zeros((self.n, self.n, self.n)) for l in range(self.ydeg + 1)
+            np.zeros((self.n, self.neig, self.n)) for l in range(self.ydeg + 1)
         ]
         for l in range(self.ydeg + 1):
             i = slice(l ** 2, (l + 1) ** 2)
@@ -95,7 +115,7 @@ class WignerIntegral(MomentIntegral):
         return mu
 
     def _second_moment(self, eigE):
-        sqrtC = np.zeros((self.N, self.n, eigE.shape[-1]))
+        sqrtC = np.zeros((self.N, self.neig, eigE.shape[-1]))
         for l in range(self.ydeg + 1):
             i = slice(l ** 2, (l + 1) ** 2)
             sqrtC[i] = self.T[l] @ eigE[i]

### 2020-08-18 starry_gp/transforms.py

@@ -7,21 +7,28 @@ from packaging import version
 # Kewyord to `eigh` changed in 1.5.0
 if version.parse(scipy.__version__) < version.parse("1.5.0"):
     eigvals = "eigvals"
+    driver_allowed = False
 else:
     eigvals = "subset_by_index"
+    driver_allowed = True
 
 
 __all__ = ["eigen", "get_c0_c1", "get_alpha_beta"]
 
 
-def eigen(Q, n=None):
+def eigen(Q, n=None, driver=None):
     """
     
     """
     N = Q.shape[0]
-    if n is None:
-        n = N
-    w, U = eigh(Q, **{eigvals: (N - n, N - 1)})
+    if n is None or n == N:
+        if driver_allowed:
+            kwargs = {"driver": driver}
+        else:
+            kwargs = {}
+    else:
+        kwargs = {eigvals: (N - n, N - 1)}
+    w, U = eigh(Q, **kwargs)
     U = U @ np.diag(np.sqrt(np.maximum(0, w)))
     return U[:, ::-1]
 

### 2020-08-18 tex/figures/spot_expansion.py

@@ -1,145 +1,303 @@
 import matplotlib.pyplot as plt
-from matplotlib import colors
+from matplotlib import colors, patches
 import numpy as np
 import starry
 from scipy.optimize import minimize
+from tqdm import tqdm
 
+
+def hwhm(r):
+    """
+    Return the theoretical half-width at half minimum as a function of r.
+    
+    """
+    return np.arccos((2 + 3 * r * (2 + r)) / (2 * (1 + r) ** 3)) * 180 / np.pi
+
+
+def r_max(hwhm_max=60):
+    """
+    Returns the value of r corresponding to `hwhm_max`.
+    
+    """
+    f = lambda r: (hwhm(r) - hwhm_max) ** 2
+    res = minimize(f, 2.0)
+    return res.x[0]
+
+
+def corr(r, c2, c3):
+    """Intensity correction function."""
+    return 1 + c2 * r ** -1 * (1 + r) ** -c3
+
+
+def get_c(ydeg, hwhm_max=60, hwhm_min=15, npts=500):
+    """
+    Return the coefficients for the radius transformation.
+
+    """
+    c = np.zeros(4)
+
+    # Minimum r: we need to optimize numerically
+    loss = lambda p: (hwhm_empirical(ydeg, p[0]) - hwhm_min) ** 2
+    res = minimize(loss, 0.1526)
+    rmin = res.x[0]
+    c[0] = rmin
+
+    # Maximum r (easy)
+    rmax = r_max(hwhm_max=hwhm_max)
+    c[1] = rmax - rmin
+
+    # Now compute the coefficients of the intensity
+    # correction, c[2] and c[3].
+
+    # Array over which to compute the loss
+    r = np.linspace(rmin, rmax, npts)
+
+    # Get the actual (absolute value of the) intensity at the peak
+    l = np.arange(ydeg + 1).reshape(1, -1)
+    term = np.sum((1 + r.reshape(-1, 1)) ** -l, axis=-1)
+    I = -0.5 * r * (1 - (2 + r) / (1 + r) * term)
+
+    # This is the factor by which we need to normalize the function
+    norm = 1.0 / I
+
+    # Find the coefficients of the fit (least squares)
[…]

### 2020-08-25 tex/gitlinks.tex
rename

--- 2020-08-11 tex/gitlinks.tex
@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_gp/blob/eb1218eef62c9812177173e82fdc6d095fc1e432/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_gp/blob/eb1218eef62c9812177173e82fdc6d095fc1e432/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_gp/blob/eb1218eef62c9812177173e82fdc6d095fc1e432/#1}{\raisebox{-0.1em}{\prooficon}}}
-\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_gp/_build}{#1}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/770b39c492f27102f9a66dc6061f06370f0306e9/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/770b39c492f27102f9a66dc6061f06370f0306e9/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/770b39c492f27102f9a66dc6061f06370f0306e9/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-09 tex/first_draft/ms.tex
update tex

--- 2020-08-18 tex/ms.tex
@@
 %
+\xxx{TODO: Explain why we need the corrected, scaled radius.}
 
@@
     \label{eq:sinf}
-    s_{n}^\infty(\rho) =
+    s_{n}'(\rho') =
+    -\dfrac{1}{2\sqrt{2l + 1}}
     \begin{cases}
-        -\dfrac{\rho}{2(1 + \rho)}
+        \dfrac{\rho'}{1 + \rho'}
          & l = m = 0
         \\[1em]
-        -\dfrac{\rho \left( 2 + \rho \right)}
-        {2 \sqrt{2l + 1} (1 + \rho)^{l + 1}}
+        \dfrac{\rho' \left( 2 + \rho' \right)}
+        {(1 + \rho')^{l + 1}}
          & l > 0, \, m = 0
@@
 $\delta$ is the Kronecker delta,
-$\rho \in [0, \infty)$ is the spot radius,
-and the $\infty$ superscript denotes the fact that this expression
-is valid in the limit $l_{\mathrm{max}} \rightarrow \infty$
+$\rho' \in [0, \infty)$ is the spot radius,
+and the prime superscript on both $s_{n}'$ and $\rho'$ denotes
+the fact that this expression
+is valid only in the limit $l_{\mathrm{max}} \rightarrow \infty$
 (more on this below).
@@
     \item The surface intensity at the antipode of the spot center is zero
-    \item The size of the spot increases monotonically with $\rho$
+    \item The size of the spot increases monotonically with $\rho'$
 \end{enumerate}
@@
         \sum\limits_{n=0}^\infty
-        s_{n}^\infty Y_{n}(\theta, \varphi) \\
+        s_{n}' Y_{n}(\theta, \varphi) \\
         & =
         \sum\limits_{n=0}^\infty
-        s_{n}^\infty \sqrt{2l + 1} \delta_{m0} P_l(\cos\theta)
+        s_{n}' \sqrt{2l + 1} \delta_{m0} P_l(\cos\theta)
     \end{split}
@@
     I(\theta) =
-    \dfrac{\rho}{2}
+    \dfrac{\rho'}{2}
     -
-    \dfrac{\rho \left( 2 + \rho \right)}{2 (1 + \rho)}
-    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + \rho}\right)^l P_l(\cos\theta)
+    \dfrac{\rho' \left( 2 + \rho' \right)}{2 (1 + \rho')}
+    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + \rho'}\right)^l P_l(\cos\theta)
     \quad.
@@
     \begin{split}
-        A & = \dfrac{\rho}{2}                      \\
-        B & = \rho (2 + \rho) \sqrt{\dfrac{1}{8 + 8 \rho}} \\
-        C & = \dfrac{1 + (1 + \rho)^2}{2 + 2 \rho}
+        A & = \dfrac{\rho'}{2}                      \\
+        B & = \rho' (2 + \rho') \sqrt{\dfrac{1}{8 + 8 \rho'}} \\
+        C & = \dfrac{1 + (1 + \rho')^2}{2 + 2 \rho'}
     \end{split}
@@
     \label{eq:hwhm}
-    \Delta\theta =
-    \cos^{-1} \left[ \dfrac{2 + 3 \rho (2 + \rho)}{2 (1 + \rho)^3} \right]
+    \Delta\theta(\rho') =
+    \cos^{-1} \left[ \dfrac{2 + 3 \rho' (2 + \rho')}{2 (1 + \rho')^3} \right]
     \quad.
@@
 \begin{align}
-    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}\rho} =
-    \frac{3}{\left(1 + \rho\right)
-        \sqrt{\left(1 + 2 \rho\right)
-            \left(3 + 2 \rho\right)}}
-    \quad,
-\end{align}
-%
-which is positive definite for all $\rho > 0$.
+    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}\rho'} =
+    \frac{3}{\left(1 + \rho'\right)
+        \sqrt{\left(1 + 2 \rho'\right)
+            \left(3 + 2 \rho'\right)}}
+    \quad,
+\end{align}
+%
+which is positive definite for all $\rho' > 0$.
+
+For future reference, Equation~(\ref{eq:hwhm}) may be inverted to obtain the
+radius $\rho'$ corresponding to a particular value of the half width
+at half minimum:
+%
+\begin{align}
+    \label{eq:hwhm}
+    \rho'(\Delta\theta) =
+    \left(\dfrac{1 +
+        \cos\left(\dfrac{2\Delta\theta}{3}\right) +
+        \sqrt{3}\sin\left(\dfrac{2\Delta\theta}{3}\right)}
+    {2 \cos\Delta\theta}\right) - 1
+    \quad.
+\end{align}
+%
 
@@
 %
-It is important to note that the
+The
 properties of the spot expansion described above
@@
 \begin{align}
-    s_n(\rho) = s_n^\infty\left(r\right) \left(1 + c_2 r^{-1} \left(1 + r\right) ^ {-c_3}\right)
+    s_n(\rho) = s_n'\left( \rho' \right) \left(1 + c_2 \left(1 - \rho\right) ^ {c_3}\right)
 \end{align}
@@
 \begin{align}
-    r \equiv c_0 + c_1\rho
-\end{align}
-%
-is a rescaled radius and $c_0, c_1, c_2, c_3$ are constants at a given
+    \begin{split}
+        \rho' &= \rho'\left(\rho\right) \\ &\equiv c_0 + c_1\rho
+        \quad.
+    \end{split}
+\end{align}
+%
+and $c_0, c_1, c_2, c_3$ are constants at a given
 value of $l_{\mathrm{max}}$.
@@
     s_{n}(\rho) =
+    -\dfrac{1 + c_2 \left(1 - \rho\right) ^ {c_3}}{2\sqrt{2l + 1}}
     \begin{cases}
-        -\dfrac{r}{2(1 + r)} - \dfrac{c_2}{2(1 + r)^{c_3 + 1}}
+        \dfrac{\rho'}{1 + \rho'}
          & l = m = 0
         \\[1em]
-        -\dfrac{r \left( 2 + r \right)}
-        {2 \sqrt{2l + 1} (1 + r)^{l + 1}}
-        - \dfrac{c_2\left( 2 + r \right)}
-        {2 \sqrt{2l + 1} (1 + r)^{c_3 + l + 1}}
+        \dfrac{ \rho' \left( 2 + \rho' \right)}
+        {\left(1 + \rho'\right)^{l + 1}}
          & l > 0, \, m = 0
@@
     \end{cases}
+    \quad.
 \end{align}

### 2020-11-09 tex/ms.tex
about to switch to new size param
update tex

@@
 \label{sec:intro}
-\xxx{Talk about starry, gps, and whatnot.}
+\xxx{Talk about starry, gps, etc.}
 
@@
     \pmb{\Sigma}(\pmb{\theta})
-     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}^2(\pmb{\theta})
+     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}(\pmb{\theta}) \pmb{\mu}^\top(\pmb{\theta})
+    \quad.
 \end{align}
-%
-where the squaring operation in Equation~(\ref{eq:cov}) is
-performed element-wise.
 %
@@
     \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta})
-     & = \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}_{\mathbf{y}}^2(\pmb{\theta})
+     & = \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}) \pmb{\mu}_{\mathbf{y}}^\top(\pmb{\theta})
 \end{align}
@@
 \begin{align}
-    \mathbf{x} = \left( \xi \,\, \lambda \,\, \phi \,\, \rho \right)^\top
+    \mathbf{x} = \left( c \,\, \lambda \,\, \phi \,\, r \right)^\top
 \end{align}
@@
     \mathbf{y}(\mathbf{x}) =
-    \xi
+    c
     \,
@@
     \,
-    \mathbf{s}(\rho)
+    \mathbf{s}(r)
     \quad,
@@
 %
-where $\xi$ is the contrast of a spot,
+where $c$ is the contrast of a spot,
 $\lambda$ is its longitude, $\phi$ is its latitude,
-and $\rho$ is its radius.
-The vector function $\mathbf{s}(\rho)$
+and $r$ is its radius.
+The vector function $\mathbf{s}(r)$
 returns the spherical harmonic expansion of a negative unit brightness
@@
     =
-    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
+    p(c \, \big| \, \pmb{\theta}_{c}) \,
     p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
     p(\phi \, \big| \, \pmb{\theta}_{\phi})\,
-    p(\rho \, \big| \, \pmb{\theta}_{\rho})
+    p(r \, \big| \, \pmb{\theta}_{r})
     \quad,
@@
     \pmb{\theta} = \left(
-    \pmb{\theta}_{\xi} \, \,
+    \pmb{\theta}_{c} \, \,
     \pmb{\theta}_{\lambda} \, \,
     \pmb{\theta}_{\phi} \, \,
-    \pmb{\theta}_{\rho} \right)^\top
+    \pmb{\theta}_{r} \right)^\top
     \quad.
@@
     \label{eq:e1}
-    \mathbf{e_1}(\pmb{\theta}_\rho)
+    \mathbf{e_1}(\pmb{\theta}_r)
      & \equiv
     \int
-    \mathbf{s}(\rho) \,
-    p(\rho \, \big| \, \pmb{\theta}_{\rho}) \,
-    \mathrm{d}\rho
+    \mathbf{s}(r) \,
+    p(r \, \big| \, \pmb{\theta}_{r}) \,
+    \mathrm{d}r
     %
@@
     \label{eq:e4}
-    \mathbf{e_4}(\pmb{\theta}_\xi, \mathbf{e_3})
+    \mathbf{e_4}(\pmb{\theta}_c, \mathbf{e_3})
      & \equiv
     \int
-    \xi \,
+    c \,
     \mathbf{e_3} \,
-    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
-    \mathrm{d}\xi
+    p(c \, \big| \, \pmb{\theta}_{c}) \,
+    \mathrm{d}c
     %
@@
     \label{eq:E1}
-    \mathbf{E_1}(\pmb{\theta}_\rho)
+    \mathbf{E_1}(\pmb{\theta}_r)
      & \equiv
     \int
-    \mathbf{s}(\rho) \, \mathbf{s}^\top(\rho) \,
-    p(\rho \, \big| \, \pmb{\theta}_{\rho}) \,
-    \mathrm{d}\rho
+    \mathbf{s}(r) \, \mathbf{s}^\top(r) \,
+    p(r \, \big| \, \pmb{\theta}_{r}) \,
+    \mathrm{d}r
     %
@@
     \label{eq:E4}
-    \mathbf{E_4}(\pmb{\theta}_\xi, \mathbf{E_3})
+    \mathbf{E_4}(\pmb{\theta}_c, \mathbf{E_3})
      & \equiv
     \int
-    \xi^2 \,
+    c^2 \,
     \mathbf{E_3} \,
     \mathbf{E_3}^\top \,
-    p(\xi \, \big| \, \pmb{\theta}_\xi)
-    \mathrm{d}\xi
+    p(c \, \big| \, \pmb{\theta}_c)
+    \mathrm{d}c
     %
@@
 
-%
-\section{The Size Integrals: \lowercase{$\mathbf{e_1}$} and $\mathbf{E_1}$}
-\label{sec:size}
-%
-\subsection{Functional form}
-%
-
-\subsubsection{$l_{\mathrm{max}} = \infty$}
-\label{sec:size-function}
-%
-We adopt the following expression for the $n^{\mathrm{th}}$ term in $\mathbf{s}$
-in the expansion of a dark, unit contrast circular spot at
-$\lambda = \phi = 0$:
-%
-\begin{align}
-    \label{eq:sinf}
-    s_{n}^\infty(\rho) =
-    \begin{cases}
-        -\dfrac{\rho}{2(1 + \rho)}
-         & l = m = 0
-        \\[1em]
-        -\dfrac{\rho \left( 2 + \rho \right)}
-        {2 \sqrt{2l + 1} (1 + \rho)^{l + 1}}
-         & l > 0, \, m = 0
-        \\[1em]
-        0
-         & \mathrm{otherwise}
-    \end{cases}
-\end{align}
-%
-where
-$l = l(n), m = m(n)$ (Equation~\ref{eq:lm}),
-$\delta$ is the Kronecker delta,
-$\rho \in [0, \infty)$ is the spot radius,
-and the $\infty$ superscript denotes the fact that this expression
-is valid in the limit $l_{\mathrm{max}} \rightarrow \infty$
-(more on this below).
-Equation~(\ref{eq:sinf}) is convenient because it satisfies
-five important properties:
-%
-\begin{enumerate}[itemsep=2pt,parsep=1pt,label=\textbf{\arabic*}]
-    \item The surface intensity is azimuthally symmetric about the spot center
-    \item The surface intensity monotonically increases away from the spot center
-    \item The surface intensity at the spot center is $-1$
-    \item The surface intensity at the antipode of the spot center is zero
-    \item The size of the spot increases monotonically with $\rho$
-\end{enumerate}
-%
-These properties may be demonstrated by considering the
-expression for the surface intensity at a given polar angle $\theta$ and
-azimuth $\varphi$:
-%
-\begin{align}
-    \begin{split}
-        I(\theta, \varphi)
-        & =
-        \sum\limits_{n=0}^\infty
-        s_{n}^\infty Y_{n}(\theta, \varphi) \\
-        & =
-        \sum\limits_{n=0}^\infty
-        s_{n}^\infty \sqrt{2l + 1} \delta_{m0} P_l(\cos\theta)
-    \end{split}
-\end{align}
-%
-where $Y_n$ is the spherical harmonic of degree $l(n)$ and order $m(n)$
-(see Equation~\ref{eq:lm}),
-$P_l$ is the Legendre polynomial of degree $l$,
-and we have implicitly
-assumed a normalization such that the integral of our expansion over
-the unit sphere is $4\pi$.
-Note that the expression above is independent of the azimuth $\varphi$,
-a consequence of the fact that all harmonics with $m \ne 0$ are zero; this
-expansion is therefore azimuthally symmetric about the spot center, as
-stated in \textbf{1}.
-%
-Combining the above expression with Equation~(\ref{eq:sinf}) and
-rearranging, we may write
-%
-\begin{align}
-    \label{eq:Igen}
-    I(\theta, \varphi) =
-    I(\theta) =
-    \dfrac{\rho}{2}
-    -
-    \dfrac{\rho \left( 2 + \rho \right)}{2 (1 + \rho)}
-    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + \rho}\right)^l P_l(\cos\theta)
-    \quad.
-\end{align}
-%
-The summation in Equation~(\ref{eq:Igen}) has a closed-form expression in
-terms of the generating function of the Legendre polynomials:
-%
-\begin{align}
-    \label{eq:gen}
-    \sum\limits_{l=0}^\infty t^l P_l(\cos\theta) = \frac{1}{\sqrt{1 + t^2 - 2 t \cos\theta}}
-    \quad,
-\end{align}
-%
-so we may express the intensity as a function of polar angle
-in the fairly simple form
-%
-\begin{align}
-    \label{eq:Ifinal}
-    I(\theta) & = A - \frac{B}{\sqrt{C - \cos\theta}}
-    \quad,
-    %
-    \\
-    \intertext{where}
-    %
-    \begin{split}
-        A & = \dfrac{\rho}{2}                      \\
-        B & = \rho (2 + \rho) \sqrt{\dfrac{1}{8 + 8 \rho}} \\
-        C & = \dfrac{1 + (1 + \rho)^2}{2 + 2 \rho}
-    \end{split}
-\end{align}
-%
-are positive constants.
-
-Differentiating Equation~(\ref{eq:Ifinal}) with respect to $\theta$, we have
-%
-\begin{align}
-    \label{eq:Ideriv}
-    \dfrac{\mathrm{d}I(\theta)}{\mathrm{d}\theta} & =
-    -\frac{B\sin\theta}{2(C - \cos\theta)^\frac{3}{2}}
-    \quad,
-\end{align}
-%
-which is zero only for $\theta = 0$ (for which $I(\theta)$ is
-minimized) and $\theta = \pi$ (for which it is maximized). The intensity
-therefore increases monotonically from the spot center to the antipode,
-as stated in \textbf{2}. The value at the minimum is
-%
-\begin{align}
-    \begin{split}
-        I_{\mathrm{min}} & = A - \dfrac{B}{\sqrt{C - 1}} \\
-        & = -1
-        \quad,
-    \end{split}
-\end{align}
-%
-as stated in \textbf{3}, and the value at the maximum is
-%
-\begin{align}
-    \begin{split}
-        I_{\mathrm{max}} & = A - \dfrac{B}{\sqrt{C + 1}} \\
-        & = 0
-        \quad,
-    \end{split}
-\end{align}
-%
-as stated in \textbf{4}.
-Finally, to show \textbf{5}, let us compute the half width at half minimum
-$\Delta\theta$ of the intensity profile:
-%
-\begin{align}
-    A - \dfrac{B}{\sqrt{C - \cos{\Delta\theta}}} =
-    -\dfrac{1}{2}
-\end{align}
-%
-Solving for $\Delta\theta$ yields
-%
-\begin{align}
-    \label{eq:hwhm}
-    \Delta\theta =
-    \cos^{-1} \left[ \dfrac{2 + 3 \rho (2 + \rho)}{2 (1 + \rho)^3} \right]
-    \quad.
-\end{align}
-%
-Differentiation with respect to $r$ yields
-%
-\begin{align}
-    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}\rho} =
-    \frac{3}{\left(1 + \rho\right)
-        \sqrt{\left(1 + 2 \rho\right)
-            \left(3 + 2 \rho\right)}}
-    \quad,
-\end{align}
-%
-which is positive definite for all $\rho > 0$.
-
-\begin{figure}[p!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/spot_expansion.pdf}
-        \oscaption{spot_expansion}{%
-            \emph{Top.}
-            Polar intensity profiles for spots with different normalized
-            radii $\rho$ in the range $(0, 1]$, computed at spherical harmonic
-            degree $l_{\mathrm{max}} = \LMAX$.
-            \emph{Bottom.}
-            The width $\Delta\theta$
-            (Equation~\ref{eq:hwhm}) of the
-            spot as a function of the radius parameter $r$ (left panel)
-            and the normalized radius $\rho$ (right panel), computed for
-            $l_{\mathrm{max}} = \LMAX$.
-            \label{fig:spot_expansion}
-        }
-    \end{centering}
-\end{figure}
-
-\subsubsection{Finite $l_{\mathrm{max}}$}
-%
-It is important to note that the
-properties of the spot expansion described above
-(in particular its monotonicity and values at the spot center and antipode)
-are inherited from the generating function of the Legendre polynomials
-(Equation~\ref{eq:gen}). These properties are therefore only rigorously
-true when the expansion is taken to spherical
-harmonic degree $l_{\mathrm{max}} = \infty$. Expansions truncated to finite
-$l_{\mathrm{max}}$ will in general experience ringing and a spot amplitude
-whose (absolute) value is in general smaller than unity.
-
-To mitigate these issues, we will do two things: we will impose a minimum
-radius $r_{\mathrm{min}}$ to prevent highly oscillatory behavior within
-the domain of our function; and we will scale the spot expansion
-by an overall amplitude factor so that
-the peak spot intensity is as close to (negative) unity as possible
-everywhere within the domain.
-We find that this can be accomplished by expressing the
-$n^{\mathrm{th}}$ term in the spot expansion $\mathbf{s}$ as
-%
-\begin{align}
-    s_n(\rho) = s_n^\infty\left(r\right) \left(1 + c_2 r^{-1} \left(1 + r\right) ^ {-c_3}\right)
-\end{align}
-%
-where
-%
-\begin{align}
-    r \equiv c_0 + c_1\rho
-\end{align}
-%
-is a rescaled radius and $c_0, c_1, c_2, c_3$ are constants at a given
-value of $l_{\mathrm{max}}$.
-%
-Plugging this into Equation~(\ref{eq:sinf}), we have
-%
-\begin{align}
-    \label{eq:s}
-    s_{n}(\rho) =
-    \begin{cases}
-        -\dfrac{r}{2(1 + r)} - \dfrac{c_2}{2(1 + r)^{c_3 + 1}}
-         & l = m = 0
-        \\[1em]
-        -\dfrac{r \left( 2 + r \right)}
-        {2 \sqrt{2l + 1} (1 + r)^{l + 1}}
-        - \dfrac{c_2\left( 2 + r \right)}
-        {2 \sqrt{2l + 1} (1 + r)^{c_3 + l + 1}}
-         & l > 0, \, m = 0
-        \\[1em]
-        0
-         & \mathrm{otherwise}
-    \end{cases}
-\end{align}
-%
-Note that in the limit $l_{\mathrm{max}} \rightarrow \infty$, we may set
-$c_0 = 0$, $c_1 = 1$, and $c_2 = 0$ to recover the original parametrization.
-At finite $l_{\mathrm{max}}$, however, we solve for the $c_n$ numerically,
-by requiring the following conditions:
-
-\xxx{TODO...}
-
-\subsection{PDF}
-\label{sec:size-pdf}
-%
-Our goal now is to characterize the distribution of $\mathbf{s}$ as a
-function of parameters $\pmb{\theta}_r$ describing the distribution of
-spot radii. Since $r \in [0, 1]$, we choose to model $r$ as a random
-Beta-distributed variable; as we will see, this choice will allow us to
-analytically compute the first two moments of the distribution of
-$\mathbf{s}$ conditioned on $\pmb{\theta}_r$.
-
-%
-The Beta distribution in $r$ has hyperparameters
-%
-\begin{align}
-    \pmb{\theta}_r = \left(
-    \alpha \, \, \, \,
-    \beta \right)^\top
-\end{align}
-%
-and PDF given by
-%
-\begin{align}
-    \label{eq:pdf_r}
-    p \big(r \, \big| \, \pmb{\theta}_r \big)
-     & =
-    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
-    r^{\alpha - 1}
-    (1 - r)^{\beta - 1}
-    \quad,
-\end{align}
-%
-with support in $0 \leq r \leq 1$,
-where $\Gamma$ is the Gamma function.
-
-\subsection{First moment}
-\label{sec:size-mom1}
-%
-The first moment of $\mathbf{s}$ is given by Equation~(\ref{eq:e1}).
-We may use Equations~(\ref{eq:sinf}) and (\ref{eq:pdf_r}) to write the
-$n^{\mathrm{th}}$ term of $\mathbf{{e_1}}(\pmb{\theta}_r)$ as
-%
-\begin{align}
-    {e_1}_n(\alpha, \beta)
-     & =
-    \resizebox{.75\hsize}{!}{$
-            \begin{dcases}
-                -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
-                \int\limits_0^1
-                \dfrac{
-                    (c_0 + c_1 r)
-                    r^{\alpha - 1}
-                    (1 - r)^{\beta - 1}
-                }{2 (1 + c_0 + c_1 r)}
-                \mathrm{d} r
-                 &
-                \qquad
-                l = m = 0    \\[2em]
-                -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
-                \int\limits_0^1
-                \dfrac{(c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)r^{\alpha - 1}
-                (1 - r)^{\beta - 1}}
-                {2 \sqrt{2l + 1} (1 + c_0 + c_1 r)^{l + 1}}
-                \mathrm{d} r
-                 &
-                \qquad
-                l > 0, m = 0 \\[2em]
-                0
-                 &
-                \qquad m \ne 0
-                \quad.
-            \end{dcases}
-        $}
-\end{align}
-%
-Thanks to the dependence of the spherical harmonic coefficients on only
-powers of $r$ and $(1 + c_0 + c_1 r)$, the integrals above may be expressed in closed
-form in terms of the hypergeometric function ${_2F_1}$:
-%
-\begin{align}
-    {e_1}_n(\alpha, \beta)
-     & =
-    \resizebox{.75\hsize}{!}{$
-            \begin{dcases}
-                -
-                \dfrac{1}{2(1 + c_0)}
-                \bigg[c_0H_0^0 + c_1 H_0^1\bigg]
-                 &
-                \qquad
-                l = m = 0    \\[2em]
-                -\dfrac{1}{2\sqrt{2l + 1} (1 + c_0)^{l+1}}
-                \bigg[
-                    c_0(2+c_0)H_l^0
-                    +
-                    2c_1(1 + c_0)H_l^1
-                    +
-                    c_1^2 H_l^2
-                    \bigg]
-                 &
-                \qquad
-                l > 0, m = 0 \\[2em]
-                0
-                 &
-                \qquad m \ne 0
-            \end{dcases}
-        $}
-\end{align}
-%
-where we define
-%
-\begin{align}
-    H_j^k & \equiv \left(\prod_{n=0}^{k-1} \lambda_n\right) G_j^k
-    \quad.
-\end{align}
-%
-with
-%
-\begin{align}
-    \lambda_n & \equiv \dfrac{\alpha + n}{\alpha + \beta + n}
-\end{align}
-%
-and
-%
-\begin{align}
-    G_j^k & \equiv {_2F_1}\left(j + 1, \alpha + k; \alpha + \beta + k; -\dfrac{c_1}{1 + c_0}\right)
-    \quad.
-\end{align}
-%
-Note that for fixed $\alpha$, $\beta$, $c_0$, and $c_1$, we need only compute
-$G_0^0$, $G_1^0$, $G_0^1$, and $G_1^1$ directly, since the remaining terms may be
-obtained recursively:
-%
-\begin{proof}{test_hypgeo}
-    \label{eq:Grec}
-    \begin{split}
-        G_j^k & =
-        \bigg[
-            \dfrac{(\alpha + \beta + k - j)(1 + c_0)}{j(1 + c_0 + c_1)}
-            \bigg] G_{j - 2}^k
-        \\[0.75em]
-        &
-        + \hspace{1.5pt}
-        \bigg[
-            1 - \dfrac{(\alpha + \beta + k - j)(1 + c_0) + (\alpha + k)c_1}
-            {j(1 + c_0 + c_1)}
-            \bigg]
-        G_{j - 1}^k
-        \\[1.5em]
-        G_j^k & =
-        \bigg[
-            \frac{(\alpha + \beta + k - 2)(1 + c_0)}
-            {(\alpha + \beta - j + k - 2)\lambda_{k - 1}c_1}
-            \bigg]
-        G_{j}^{k - 2}
-        \\[0.75em]
-        & + \hspace{1.5pt}
-        \bigg[
-            \frac{1}{\lambda_{k - 1}}
-            -\frac{(\alpha + \beta + k - 2)(1 + c_0) + \beta c_1}
-            {(\alpha + \beta - j + k - 2)\lambda_{k-1}c_1}
-            \bigg]
-        G_{j}^{k - 1}
-        \quad.
-    \end{split}
-    \raisetag{7.5em}
-\end{proof}
-%
-
-%
-\vspace{1em}
-%
-
-\subsection{Second moment}
-\label{sec:size-mom2}
-%
-The second moment of $\mathbf{s}$ is given by Equation~(\ref{eq:E1}).
-We may again use Equations~(\ref{eq:sinf}) and (\ref{eq:pdf_r}) to write the
-term at index $(n, n')$ of $\mathbf{{E_1}}(\pmb{\theta}_r)$ as
-%
-\begin{align}
-    {E_1}_{n,n'} (\alpha, \beta) & =
-    \resizebox{1.5\hsize}{!}{$
-            \begin{dcases}
-                \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
-                \int\limits_0^1
-                \left(
-                \dfrac{c_0 + c_1 r}{2 (1 + c_0 + c_1 r)}
-                \right)^2
-                \\
-                \qquad\qquad\quad\quad\quad\quad\quad
-                \times
-                \,\,
-                r^{\alpha - 1}
-                (1 - r)^{\beta - 1}
-                \mathrm{d} r
-                 &
-                \qquad
-                \parbox[t]{\linewidth}{\vspace{-2.5em}$l = l' = 0, \\m = m' = 0$}
-                \\[2em]
-                %
-                %
-                %
-                \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
-                \int\limits_0^1
-                \left(
-                \dfrac{(c_0 + c_1 r)}{2 (1 + c_0 + c_1 r)}
-                \right)
-                \\
-                \qquad\qquad\quad\quad\quad\quad\quad
-                \times
-                %
-                \left(
-                \dfrac{(c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)}
-                    {2 \sqrt{2l + 1} (1 + c_0 + c_1 r)^{l + 1}}
-                \right)
-                %
-                \\[0.75em]
-                \qquad\qquad\quad\quad\quad\quad\quad
-                \times
-                \,\,
-                r^{\alpha - 1}
-                (1 - r)^{\beta - 1}
-                \mathrm{d} r
-                 &
-                \qquad
-                \parbox[t]{\linewidth}{\vspace{-5em}$l > 0,        \\l' = 0, \\m = m' = 0$}
-                \\[2em]
-                %
-                %
-                %
-                \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
-                \int\limits_0^1
-                \left(
-                \dfrac{(c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)}
-                    {2 \sqrt{2l + 1} (1 + c_0 + c_1 r)^{l + 1}}
-                \right)
-                \\
-                \qquad\qquad\quad\quad\quad\quad\quad
-                \times
-                %
-                \left(
-                \dfrac{(c_0 + c_1 r) \left( 2 + c_0 + c_1 r \right)}
-                    {2 \sqrt{2l' + 1} (1 + c_0 + c_1 r)^{l' + 1}}
-                \right)
-                %
-                \\[0.75em]
-                \qquad\qquad\quad\quad\quad\quad\quad
-                \times
-                \,\,
-                r^{\alpha - 1}
-                (1 - r)^{\beta - 1}
-                \mathrm{d} r
-                 &
-                \qquad
-                \parbox[t]{\linewidth}{\vspace{-5em}$l > 0,        \\l' > 0, \\m = m' = 0$}
-                \\[2em]
-                %
-                %
-                %
-                0
-                 &
-                \qquad m, m' \ne 0
-                \quad,
-            \end{dcases}
-        $}
-    \raisetag{15.5em}
-\end{align}
-%
-where $l' = l'(n')$ and $m' = m'(n')$ are again given by Equation~(\ref{eq:lm})
-and, by symmetry, the expression for $l' > 0, \, l = 0, \, m = m' = 0$ is the same as
-in the second case, provided we make the substitution $l \rightarrow l'$.
-%
-These integrals again reduce to closed form expressions:
-%
-\begin{align}
-    {E_1}_{n,n'} (\alpha, \beta) & =
-    \resizebox{1.4\hsize}{!}{$
-            \begin{dcases}
-                \dfrac{1}{4(1+c_0)^2}
-                \bigg(
-                c_0^2 H_1^0
-                +
-                2 c_0 c_1 H_1^1
-                +
-                c_1^2 H_1^2
-                \bigg)
-                 &
-                \qquad
-                \parbox[t]{\linewidth}{\vspace{-1.5em}$l = l' = 0, \\m = m' = 0$}
-                \\[2em]
-                %
-                %
-                %
-                \dfrac{1}{4\sqrt{2l + 1}(1+c_0)^{l+2}}
-                \\[0.75em]
-                \enspace\quad
-                \times
-                \bigg(
-                c_0^2(2 + c_0) H_{l+1}^0
-                + c_0 c_1 (4 + 3 c_0) H_{l+1}^1
-                \\[0.75em]
-                \qquad
-                + (2 + 3 c_0) c_1^2 H_{l + 1}^2
-                + c_1^3 H_{l + 1}^3
-                \bigg)
-                 &
-                \qquad
-                \parbox[t]{\linewidth}{\vspace{-5.5em}$l > 0,      \\l' = 0, \\m = m' = 0$}
-                \\[2em]
-                %
-                %
-                %
-                \dfrac{1}{4\sqrt{(2l + 1)(2l' + 1)}(1+c_0)^{l+l'+2}}
-                \\[0.75em]
-                \enspace\quad
-                \times
-                \bigg[
-                    c_0^2(2+c_0)^2 H_{l + l' + 1}^0
-                    + 4 c_0 (1 + c_0) (2 + c_0) c_1 H_{l + l' + 1}^1
-                    \\[0.75em]
-                \qquad
-                + 2 \big(2 + 3 c_0 (2 + c_0)\big)c_1^2 H_{l + l' + 1}^2
-                + 4(1 + c_0)c_1^3 H_{l + l' + 1}^3
-                \\[0.75em]
-                \qquad
-                + c_1^4 H_{l + l' + 1}^4
-                \bigg]
-                 &
-                \qquad
-                \parbox[t]{\linewidth}{\vspace{-6em}$l > 0,        \\l' > 0, \\m = m' = 0$}
-                \\[2em]
-                %
-                %
-                %
-                0
-                 &
-                \qquad m, m' \ne 0
-                \quad.
-            \end{dcases}
-        $}
-    \raisetag{13.5em}
-\end{align}
-%
-
-\section{The Latitude Integrals: \lowercase{$\mathbf{e_2}$} and $\mathbf{E_2}$}
-\label{sec:latitude}
-\xxx{TODO.}
-
-\section{The Longitude Integrals: \lowercase{$\mathbf{e_3}$} and $\mathbf{E_3}$}
-\label{sec:longitude}
-\xxx{TODO.}
-
-\section{The Contrast Integrals: \lowercase{$\mathbf{e_4}$} and $\mathbf{E_4}$}
-\label{sec:contrast}
-%
-\subsection{Functional form}
-\label{sec:contrast-function}
-Since the spot function $\mathbf{s}$ (Equation~\ref{eq:sinf}) is normalized
-to $-1$, we may simply scale it by the contrast $\xi$ to obtain a spot
-whose peak intensity is $-\xi$.
-Assuming the baseline (unspotted) stellar intensity is unity, $\xi$
-is the fractional decrease in the brightness relative to the stellar
-baseline: i.e., the spot contrast.
-%
-
-\subsection{PDF}
-\label{sec:contrast-pdf}
-We require our
-spot contrast to be no larger than $1$ (corresponding to zero intensity
-at the center of the spot when the baseline is unity) to enforce physical
-intensities
-everywhere. We have thus far implicitly assumed the spot is dark, but
-in principle bright spots may also exist
-(such as the plages observed on the Sun), corresponding to contrasts
-$\xi < 0$. We therefore require a PDF with support over $(-\infty, 1)$.
-Defining
-%
-\begin{align}
-    b(\xi) \equiv 1 - \xi
-\end{align}
-%
-to be the actual brightnesss of a unit-intensity stellar surface at
-the center of the spot,
-we can obtain the required support by modeling
-$b \in (0, \infty)$ as a log-normal
-random variable:
-%
-\begin{align}
-    \label{eq:lognormal}
-    p \big(b \, \big| \, \mu, \nu \big)
-     & =
-    \frac{1}{b\sqrt{2\pi\nu}}
-    \exp\left[
-        -\dfrac{\big(\ln b - \mu\big)^2}{2\nu}
-        \right]
-    \quad.
-\end{align}
-%
-where $\mu$ and $\nu$ are the mean and variance in $b$, respectively.
-The corresponding PDF in terms of $\xi$ is
-%
-\begin{align}
-    \label{eq:lognormal}
-    p \big(\xi \, \big| \, \pmb{\theta}_\xi \big)
-     & =
-    \frac{1}{(1 - \xi)\sqrt{2\pi\nu}}
-    \exp\left[
-        -\dfrac{\big(\ln (1 - \xi) - \mu\big)^2}{2\nu}
-        \right]
-    \quad.
-\end{align}
-%
-where the hyperparameters are
-%
-\begin{align}
-    \pmb{\theta}_\xi = \left(
-    \mu \, \, \, \,
-    \nu \right)^\top
-    \quad.
-\end{align}
-%
-
-\subsection{First moment}
-\label{sec:contrast-mom1}
-The first moment of the spot contrast distribution is given by
-Equation~(\ref{eq:e4}), which may easily be solved analytically:
-%
-\begin{align}
-    \mathbf{e_4}(\pmb{\theta}_\xi, \mathbf{e_3})
-     & =
-    \int_{-\infty}^1
-    \xi \,
-    \mathbf{e_3} \,
-    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
-    \mathrm{d}\xi
-    \nonumber
-    \\
-     & =
-    \left(1 - \exp\left[ \mu + \frac{1}{2}\nu\right]\right) \mathbf{e_3}
-    \quad.
-\end{align}
-%
-
-\subsection{Second moment}
-\label{sec:contrast-mom2}
-The second moment of the spot contrast distribution is given by
-Equation~(\ref{eq:E4}), which also evaluates to a simple closed form:
-%
-\begin{align}
-    \mathbf{E_4}(\pmb{\theta}_\xi, \mathbf{E_3})
-     & =
-    \int_{-\infty}^1
-    \xi^2 \,
-    \mathbf{E_3} \,
-    \mathbf{E_3}^\top \,
-    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
-    \mathrm{d}\xi
-    \nonumber
-    \\
-     & =
-    \left(1 - 2\exp\bigg[ \mu + \frac{1}{2}\nu\bigg]
-    + \exp\bigg[ 2\mu + 2\nu\bigg]\right)
-    \mathbf{E_3} \,
-    \mathbf{E_3}^\top
-    \quad.
-\end{align}
-
 \bibliography{bib}

### 2020-11-10 tex/ms.tex
working on paper

@@
      & =
-    \mathbf{e_4}(\pmb{\theta})
+    \mathbf{e}_c
     \\[1em]
@@
      & =
-    \mathbf{E_4}(\pmb{\theta})
+    \mathbf{E}_c
 \end{align}
@@
     \label{eq:e1}
-    \mathbf{e_1}(\pmb{\theta}_r)
+    \mathbf{e}_r
      & \equiv
@@
     \label{eq:e2}
-    \mathbf{e_2}(\pmb{\theta}_\phi, \mathbf{e_1})
+    \mathbf{e}_\phi
      & \equiv
@@
     \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{e_1} \,
+    \mathbf{e}_r \,
     p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
@@
     \label{eq:e3}
-    \mathbf{e_3}(\pmb{\theta}_\lambda, \mathbf{e_2})
+    \mathbf{e}_\lambda
      & \equiv
@@
     \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
-    \mathbf{e_2} \,
+    \mathbf{e}_\phi \,
     p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
@@
     \label{eq:e4}
-    \mathbf{e_4}(\pmb{\theta}_c, \mathbf{e_3})
+    \mathbf{e}_c
      & \equiv
@@
     c \,
-    \mathbf{e_3} \,
+    \mathbf{e}_\lambda \,
     p(c \, \big| \, \pmb{\theta}_{c}) \,
@@
     \label{eq:E1}
-    \mathbf{E_1}(\pmb{\theta}_r)
+    \mathbf{E}_r
      & \equiv
@@
     \label{eq:E2}
-    \mathbf{E_2}(\pmb{\theta}_\phi, \mathbf{E_1})
+    \mathbf{E}_\phi
      & \equiv
@@
     \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{E_1} \,
-    \mathbf{E_1}^\top \,
+    \mathbf{E}_r \,
+    \mathbf{E}_r^\top \,
     \mathbf{R}_{\hat{\mathbf{x}}}^\top(\phi) \,
@@
     \label{eq:E3}
-    \mathbf{E_3}(\pmb{\theta}_\lambda, \mathbf{E_2})
+    \mathbf{E}_\lambda
      & \equiv
@@
     \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
-    \mathbf{E_2} \,
-    \mathbf{E_2}^\top \,
+    \mathbf{E}_\phi \,
+    \mathbf{E}_\phi^\top \,
     \mathbf{R}_{\hat{\mathbf{y}}}^\top(\lambda) \,
@@
     \label{eq:E4}
-    \mathbf{E_4}(\pmb{\theta}_c, \mathbf{E_3})
+    \mathbf{E}_c
      & \equiv
@@
     c^2 \,
-    \mathbf{E_3} \,
-    \mathbf{E_3}^\top \,
+    \mathbf{E}_\lambda \,
+    \mathbf{E}_\lambda^\top \,
     p(c \, \big| \, \pmb{\theta}_c)
@@
 
+\section{The Size and Contrast Integrals}
+\label{sec:size-contrast}
+
+\section{The Latitude and Longitude Integrals}
+\label{sec:lat-lon}
+%
+Our goal in this section is to compute the integrals...
+
+%
+The Wigner rotation matrix for real spherical harmonics up to degree $l_{\mathrm{max}}$
+may be written as the block-diagonal matrix
+%
+\begin{align}
+    \setstackgap{L}{1.25\baselineskip}
+    \fixTABwidth{T}
+    \mathbf{R} =
+    \parenMatrixstack{
+    \quad\quad \, \mathbf{R}^0 \, \quad\quad
+     &              &              &        &                               \\
+     & \mathbf{R}^1 &              &        &                               \\
+     &              & \mathbf{R}^2 &        &                               \\
+     &              &              & \ddots &                               \\
+     &              &              &        & \mathbf{R}^{l_{\mathrm{max}}}
+    }\quad,
+\end{align}
+%
+where
+%
+\begin{align}
+    \mathbf{R}^l = {\mathbf{U}^l}^{-1} \mathbf{D}^l \mathbf{U}^l
+\end{align}
+%
+is the Wigner rotation matrix for a single spherical harmonic degree,
+%
+\begin{align}
+    \label{eq:U}
+    \setstackgap{L}{1.25\baselineskip}
+    \fixTABwidth{T}
+    \mathbf{U}^l =
+    \frac{1}{\sqrt{2}}
+    \parenMatrixstack{
+        \quad\quad\, \ddots \, \quad\quad\quad\quad
+           &       &        &       &          &    &   &    & \Ddots \\
+           & \imag &        &       &          &    &   & 1  &        \\
+           &       & \imag  &       &          &    & 1 &    &        \\
+           &       &        & \imag &          & 1  &   &    &        \\
+           &       &        &       & \sqrt{2} &    &   &    &        \\
+           &       &        & \imag &          & -1 &   &    &        \\
+           &       & -\imag &       &          &    & 1 &    &        \\
+           & \imag &        &       &          &    &   & -1 &        \\
+    \Ddots &       &        &       &          &    &   &    & \ddots
+    }
+\end{align}
+%
+describes the transformation from complex to real spherical harmonics,
+and $\mathbf{D}$ is the Wigner matrix for complex spherical harmonics,
+whose terms are given by the expression
+%
+\begin{align}
+    D_{m,m'}^l(\alpha, \beta, \gamma) = \mathrm{e}^{-\imag m' \alpha}
+    d_{m,m'}^l(\beta) \mathrm{e}^{-\imag m \gamma}
+\end{align}
+%
+where $\alpha$, $\beta$, and $\gamma$ are the Euler angles describing
+the rotation in the
+$\hat{\mathbf{z}}{-}\hat{\mathbf{y}}{-}\hat{\mathbf{z}}$ convention
+and $\imag$ is the imaginary unit. The terms of the $d$-matrix
+depend on powers of $\sin\left(\nicefrac{\beta}{2}\right)$
+and $\cos\left(\nicefrac{\beta}{2}\right)$
+\citep[c.f. Equation C15 in][]{Luger2019}, but it is convenient to use
+the half-angle formula to express these terms instead as
+%
+\begin{align}
+    d_{m,m'}^l(\beta) =
+    \sum\limits_i c_{m,m',i}^{l}
+    \mathrm{sgn}(\sin\beta)^{i}
+    (1 - \cos\beta)^{\frac{2l - i}{2}}
+    (1 + \cos\beta)^\frac{i}{2}
+\end{align}
+%
+where
+%
+\begin{align}
+    c_{m,m',i}^{l} & =
+    \resizebox{.75\hsize}{!}{$
+            \begin{cases}
+                \dfrac{
+                \left(-1\right)^{\frac{2l + 3m + m' - i}{2}}
+                \sqrt{\left(l - m\right)!
+                    \left(l + m\right)!
+                    \left(l - m'\right)!
+                    \left(l + m'\right)!}
+                }{
+                2^l
+                \left(\frac{i - m - m'}{2}\right)!
+                \left(\frac{i + m + m'}{2}\right)!
+                \left(\frac{2l - i - m + m'}{2}\right)!
+                \left(\frac{2l - i + m - m'}{2}\right)!
+                }
+                %
+                 & m - m' - i \,\, \mathrm{even}
+                %
+                \\[0.5em]
+                0
+                %
+                 & m - m' - i \,\, \mathrm{odd}
+            \end{cases}
+        $}
+\end{align}
+%
+Rotation by
+an angle $\phi$ about the $\hat{\mathbf{x}}$ axis may be accomplished by choosing
+Euler angles
+%
+\begin{align}
+    \alpha & = \frac{\pi}{2} \nonumber \\
+    \beta  & = \phi \nonumber          \\
+    \gamma & = -\frac{\pi}{2}
+    \quad,
+\end{align}
+%
+which makes it clear that all terms in
+$\mathbf{R}_{\hat{\mathbf{x}}}(\phi)$ are equal to (weighted) sums of powers
+of $(1 \pm \cos\phi)$.
+%
+Since our goal is to compute integrals of these terms multiplied by
+a probability density function, it is convenient to model
+$\cos\phi$ as a Beta-distributed variable. As we will see, this
+choice will allow us to
+analytically compute the first two moments of the distribution of
+$\phi$ conditioned on $\pmb{\theta}_\phi$.
+
+%
+The Beta distribution in $\cos\phi$ has hyperparameters
+%
+\begin{align}
+    \pmb{\theta}_\phi = \left(
+    \alpha \, \, \, \,
+    \beta \right)^\top
+\end{align}
+%
+and PDF given by
+%
+\begin{align}
+    \label{eq:cosphi-pdf}
+    p \big(\cos\phi \, \big| \, \pmb{\theta}_\phi \big)
+     & =
+    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+    (\cos\phi)^{\alpha - 1}
+    (1 - \cos\phi)^{\beta - 1}
+    \quad,
+\end{align}
+%
+where $\Gamma$ is the Gamma function. The implied distribution for $\phi$
+may be computed by a straightforward change of variable:
+%
+\begin{align}
+    \label{eq:phi-pdf}
+    p \big(\phi \, \big| \, \pmb{\theta}_\phi \big)
+     & =
+    \dfrac{\Gamma(\alpha + \beta)}{2\Gamma(\alpha)\Gamma(\beta)}
+    \big|
+    \sin\phi
+    \big|
+    (\cos\phi)^{\alpha - 1}
+    (1 - \cos\phi)^{\beta - 1}
+    \quad,
+\end{align}
+%
+for $\phi \in \left[ -\frac{\pi}{2}, \frac{\pi}{2} \right]$.
+
+Since the Wigner matrices are block diagonal, we may evaluate the moments of the
+distribution one spherical harmonic degree at a time. To that end, let us
+write
+%
+\begin{align}
+    \mathbf{e}_\phi
+     & \equiv
+    \int
+    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
+    \mathbf{e}_r \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    \nonumber
+    \\
+    %
+     & =
+    %
+    \left(
+    \mathbf{e}_\phi^0
+    \,\,\,
+    \mathbf{e}_\phi^1
+    \,\,\,
+    \mathbf{e}_\phi^2
+    \,\,\,
+    \cdots
+    \,\,\,
+    \mathbf{e}_\phi^{l_{\mathrm{max}}}
+    \right)^\top
+    \quad,
+\end{align}
+%
+where
+%
+\begin{align}
+    \mathbf{e}_\phi^l
+     & =
+    \int
+    \mathbf{R}^l_{\hat{\mathbf{x}}}(\phi) \,
+    \mathbf{e}^l_r \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    \nonumber \\
+    %
+     & =
+    %
+    {\mathbf{U}^l}^{-1}
+    \pmb{\rho}^l
+    \,
+    \mathbf{U}^l
+    \quad,
+\end{align}
+%
+where we define
+%
+\begin{align}
+    \pmb{\rho}^l
+     & \equiv
+    \int
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi) \,
+    \mathbf{e}^l_r \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    \,
+\end{align}
+%
+with
+%
+\begin{align}
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi)
+     & =
+    \mathbf{D}^l\left(\frac{\pi}{2}, \phi, -\frac{\pi}{2}\right)
+    \quad.
+\end{align}
+%
+We may write the terms of $\pmb{\rho}^l$ as
+%
+\begin{align}
+    \rho^l_m
+     & =
+    \int
+    \sum\limits_{m'}
+    {D^l_{\hat{\mathbf{x}}}}_{m,m'}(\phi) \,
+    {e^l_r}_{m'} \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    %
+    \nonumber \\[0.5em]
+    %
+     & =
+    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+    \sum\limits_{m'}
+    {e^l_r}_{m'}
+    \mathrm{e}^{\frac{\imag \pi}{2}(m - m')}
+    \sum\limits_i c_{m,m',i}^{l}
+    \,
+    q^l_i(\pmb{\theta}_{\phi})
+    \quad,
+\end{align}
+%
+where
+%
+\begin{align}
+    q^l_i(\pmb{\theta}_{\phi})
+     & =
+    \frac{1}{2}
+    \int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}
+    %
+    \mathrm{sgn}(\sin\phi)^{i}
+    \big| \sin\phi \big|
+    (\cos\phi)^{\alpha - 1}
+    (1 - \cos\phi)^{l + \beta - \frac{i}{2} - 1}
+    (1 + \cos\phi)^\frac{i}{2}
+    %
+    \,
+    \mathrm{d}\phi
+    \nonumber \\[0.5em]
+    %
+     & =
+    % %
+    % \begin{cases}
+    %     \displaystyle\int_{0}^{\frac{\pi}{2}}
+    %     %
+    %     (\cos\phi)^{\alpha - 1}
+    %     (1 - \cos\phi)^{l + \beta - \frac{i}{2} - 1}
+    %     (1 + \cos\phi)^\frac{i}{2}
+    %     %
+    %     \,
+    %     \sin\phi
+    %     \,
+    %     \mathrm{d}\phi
+    %     %
+    %     & i \,\, \mathrm{even}
+    %     %
+    %     \\
+    %     %
+    %     0
+    %     %
+    %     & i \,\, \mathrm{odd}
+    % \end{cases}
+    % \nonumber \\[0.5em]
+    % & =
+    % %
+    \begin{cases}
+        \displaystyle\int_{0}^{1}
+        %
+        x^{\alpha - 1}
+        (1 - x)^{l + \beta - \frac{i}{2} - 1 }
+        (1 + x)^\frac{i}{2}
+        %
+        \,
+        \mathrm{d}x
+        %
+         & i \,\, \mathrm{even}
+        %
+        \\
+        %
+        0
+        %
+         & i \,\, \mathrm{odd} \quad,
+    \end{cases}
+\end{align}
+%
+and in the last line we made use of the transformation $x = \cos\phi$.
+%
+The integral in the expression above has a closed-form solution in terms
+of the hypergeometric function $_2F_1$:
+%
+\begin{align}
+    %
+    q^l_i(\pmb{\theta}_{\phi})
+     & =
+    \begin{cases}
+        \dfrac{\Gamma(\alpha)\Gamma(l + \beta - \frac{i}{2})}{\Gamma(l + \alpha + \beta - \frac{i}{2})}
+        \,
+        _2F_1\Big(
+        -\frac{i}{2},
+        \alpha;
+        l + \alpha + \beta - \frac{i}{2};
+        -1
+        \Big)
+        %
+         & i \,\, \mathrm{even}
+        %
+        \\
+        %
+        0
+        %
+         & i \,\, \mathrm{odd} \quad.
+    \end{cases}
+\end{align}
+
 \bibliography{bib}

### 2020-11-10 tex/gitlinks.tex
working on tex

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/770b39c492f27102f9a66dc6061f06370f0306e9/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/770b39c492f27102f9a66dc6061f06370f0306e9/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/770b39c492f27102f9a66dc6061f06370f0306e9/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/77ecd5e5ec33f62d328f46350d17696547e5c0d3/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/77ecd5e5ec33f62d328f46350d17696547e5c0d3/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/77ecd5e5ec33f62d328f46350d17696547e5c0d3/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-10 tex/ms.tex
fix math
working on tex

--- 2020-11-09 tex/ms.tex
@@
 
+\section{The Size and Contrast Integrals}
+\label{sec:size-contrast}
+
+\section{The Latitude and Longitude Integrals}
+\label{sec:lat-lon}
+%
+Our goal in this section is to compute the integrals...
+
+\subsection{Rotation matrices}
+%
+The Wigner rotation matrix for real spherical harmonics up to degree $l_{\mathrm{max}}$
+may be written as the block-diagonal matrix
+%
+\begin{align}
+    \setstackgap{L}{1.25\baselineskip}
+    \fixTABwidth{T}
+    \mathbf{R} =
+    \parenMatrixstack{
+    \quad\quad \, \mathbf{R}^0 \, \quad\quad
+     &              &              &        &                               \\
+     & \mathbf{R}^1 &              &        &                               \\
+     &              & \mathbf{R}^2 &        &                               \\
+     &              &              & \ddots &                               \\
+     &              &              &        & \mathbf{R}^{l_{\mathrm{max}}}
+    }\quad,
+\end{align}
+%
+where
+%
+\begin{align}
+    \mathbf{R}^l = {\mathbf{U}^l}^{-1} \mathbf{D}^l \mathbf{U}^l
+\end{align}
+%
+is the Wigner rotation matrix for a single spherical harmonic degree,
+%
+\begin{align}
+    \label{eq:U}
+    \setstackgap{L}{1.25\baselineskip}
+    \fixTABwidth{T}
+    \mathbf{U}^l =
+    \frac{1}{\sqrt{2}}
+    \parenMatrixstack{
+        \quad\quad\, \ddots \, \quad\quad\quad\quad
+           &       &        &       &          &    &   &    & \Ddots \\
+           & \imag &        &       &          &    &   & 1  &        \\
+           &       & \imag  &       &          &    & 1 &    &        \\
+           &       &        & \imag &          & 1  &   &    &        \\
+           &       &        &       & \sqrt{2} &    &   &    &        \\
+           &       &        & \imag &          & -1 &   &    &        \\
+           &       & -\imag &       &          &    & 1 &    &        \\
+           & \imag &        &       &          &    &   & -1 &        \\
+    \Ddots &       &        &       &          &    &   &    & \ddots
+    }
+\end{align}
+%
+describes the transformation from complex to real spherical harmonics,
+and $\mathbf{D}$ is the Wigner matrix for complex spherical harmonics,
+whose terms are given by the expression
+%
+\begin{align}
+    D_{m,m'}^l(\alpha, \beta, \gamma) = \mathrm{e}^{-\imag m' \alpha}
+    d_{m,m'}^l(\beta) \mathrm{e}^{-\imag m \gamma}
+\end{align}
+%
+where $\alpha$, $\beta$, and $\gamma$ are the Euler angles describing
+the rotation in the
+$\hat{\mathbf{z}}{-}\hat{\mathbf{y}}{-}\hat{\mathbf{z}}$ convention
+and $\imag$ is the imaginary unit. The terms of the $d$-matrix
+depend on powers of $\sin\left(\nicefrac{\beta}{2}\right)$
+and $\cos\left(\nicefrac{\beta}{2}\right)$
+\citep[c.f. Equation C15 in][]{Luger2019}, but it is convenient to use
+the half-angle formula to express these terms instead as
+%
+\begin{proof}{test_dlmmp}
+    d_{m,m'}^l(\beta) =
+    \sum\limits_{i=0}^{2l} c_{m,m',i}^{l}
+    \mathrm{sgn}(\sin\beta)^{i}
+    (1 - \cos\beta)^{\frac{2l - i}{2}}
+    (1 + \cos\beta)^\frac{i}{2}
+\end{proof}
+%
+where
+%
+\begin{proof}{test_dlmmp}
+    c_{m,m',i}^{l} & =
+    \resizebox{.75\hsize}{!}{$
+            \begin{cases}
+                \dfrac{
+                \left(-1\right)^{\frac{2l + 3m + m' - i}{2}}
+                \sqrt{\left(l - m\right)!
+                    \left(l + m\right)!
+                    \left(l - m'\right)!
+                    \left(l + m'\right)!}
+                }{
+                2^l
+                \left(\frac{i - m - m'}{2}\right)!
+                \left(\frac{i + m + m'}{2}\right)!
+                \left(\frac{2l - i - m + m'}{2}\right)!
+                \left(\frac{2l - i + m - m'}{2}\right)!
+                }
+                %
+                 & m - m' - i \,\, \mathrm{even}
+                %
+                \\[0.5em]
+                0
+                %
+                 & m - m' - i \,\, \mathrm{odd}
+            \end{cases}
+        $}
+\end{proof}
+%
+
+\subsection{Latitude}
+%
+The latitude integrals (Equations~\ref{eq:e2} and \ref{eq:E2}) involve
+rotations about $\hat{\mathbf{x}}$, which
+may be accomplished by choosing
+Euler angles $\alpha = \nicefrac{\pi}{2}$, $\beta = \phi$, and
+$\gamma = -\nicefrac{\pi}{2}$, such that
+%
+\begin{align}
+    \mathbf{R}^l_{\hat{\mathbf{x}}}(\phi)
+     & =
+    {\mathbf{U}^l}^{-1} \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi) \mathbf{U}^l
+\end{align}
+%
+with
+\begin{align}
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi)
+     & =
+    \mathbf{D}^l\left(\frac{\pi}{2}, \phi, -\frac{\pi}{2}\right)
+    \quad.
+\end{align}
+%
+From the expressions above, it is clear that all terms in
+$\mathbf{R}_{\hat{\mathbf{x}}}(\phi)$ are equal to (weighted) sums of powers
+of $(1 \pm \cos\phi)$.
+%
+Since our goal is to compute integrals of these terms multiplied by
+a probability density function, it is convenient to model
+$\cos\phi$ as a Beta-distributed variable. As we will see, this
+choice will allow us to
+analytically compute the first two moments of the distribution of
+$\phi$ conditioned on $\pmb{\theta}_\phi$.
+
+%
+The Beta distribution in $\cos\phi$ has hyperparameters
+%
+\begin{align}
+    \pmb{\theta}_\phi = \left(
+    \alpha \, \, \, \,
+    \beta \right)^\top
+\end{align}
+%
+and PDF given by
+%
+\begin{align}
+    \label{eq:cosphi-pdf}
+    p \big(\cos\phi \, \big| \, \pmb{\theta}_\phi \big)
+     & =
+    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+    (\cos\phi)^{\alpha - 1}
+    (1 - \cos\phi)^{\beta - 1}
+    \quad,
+\end{align}
+%
+where $\Gamma$ is the Gamma function. The implied distribution for $\phi$
+may be computed by a straightforward change of variable:
+%
+\begin{proof}{test_beta_transform}
+    \label{eq:phi-pdf}
+    p \big(\phi \, \big| \, \pmb{\theta}_\phi \big)
+    & =
+    \dfrac{\Gamma(\alpha + \beta)}{2\Gamma(\alpha)\Gamma(\beta)}
+    \big|
+    \sin\phi
+    \big|
+    (\cos\phi)^{\alpha - 1}
+    (1 - \cos\phi)^{\beta - 1}
+    \quad,
+\end{proof}
+%
+for $\phi \in \left[ -\frac{\pi}{2}, \frac{\pi}{2} \right]$.
+
+Since the Wigner matrices are block diagonal, we may evaluate the moments of the
+distribution one spherical harmonic degree at a time. To that end, let us
+write
+%
+\begin{proof}{test_ephi}
+    \mathbf{e}_\phi
+    & \equiv
+    \int
+    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
+    \mathbf{e}_r \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    \nonumber
+    \\
+    %
+    & =
+    %
+    \left(
+    \mathbf{e}_\phi^0
+    \,\,\,
+    \mathbf{e}_\phi^1
+    \,\,\,
+    \mathbf{e}_\phi^2
+    \,\,\,
+    \cdots
+    \,\,\,
+    \mathbf{e}_\phi^{l_{\mathrm{max}}}
+    \right)^\top
+    \quad,
+\end{proof}
+%
+and
+%
+\begin{proof}{}
+    \mathbf{E}_\phi
+    & \equiv
+    \int
+    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
+    \mathbf{E}_r \,
+    \mathbf{E}_r^\top \,
+    \mathbf{R}_{\hat{\mathbf{x}}}^\top(\phi) \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    \nonumber
+    \\
+    %
+    & =
+    %
+    \setstackgap{L}{1.25\baselineskip}
+    \fixTABwidth{T}
+    \parenMatrixstack{
+    \mathbf{E}_\phi^{0,0} & \mathbf{E}_\phi^{0,1} & \mathbf{E}_\phi^{0,2} & \cdots & \mathbf{E}_\phi^{0,l_{\mathrm{max}}} \\
+    \mathbf{E}_\phi^{1,0} & \mathbf{E}_\phi^{1,1} & \mathbf{E}_\phi^{1,2} & \cdots & \mathbf{E}_\phi^{1,l_{\mathrm{max}}} \\
+    \mathbf{E}_\phi^{2,0} & \mathbf{E}_\phi^{2,1} & \mathbf{E}_\phi^{2,2} & \cdots & \mathbf{E}_\phi^{2,l_{\mathrm{max}}} \\
+    \vdots & \vdots & \vdots & \ddots & \vdots \\
+    \mathbf{E}_\phi^{l_{\mathrm{max}},0} & \mathbf{E}_\phi^{l_{\mathrm{max}},1} & \mathbf{E}_\phi^{l_{\mathrm{max}},2} & \cdots & \mathbf{E}_\phi^{l_{\mathrm{max}},l_{\mathrm{max}}}
+    }
+    \quad,
+\end{proof}
+%
+where
+%
+\begin{proof}{test_ephi}
+    \mathbf{e}_\phi^l
+    & =
+    \int
+    \mathbf{R}^l_{\hat{\mathbf{x}}}(\phi) \,
+    \mathbf{e}^l_r \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    \nonumber \\
+    %
+    & =
+    %
+    {\mathbf{U}^l}^{-1}
+    \pmb{\rho}^l
+    \quad,
+\end{proof}
+%
+and
+%
+\begin{proof}{}
+    \mathbf{E}_\phi^{l,l'}
+    & =
+    \int
+    \mathbf{R}^l_{\hat{\mathbf{x}}}(\phi) \,
+    \mathbf{e}^l_r \,
+    {\mathbf{R}^{l'}_{\hat{\mathbf{x}}}}^\top(\phi) \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    \nonumber \\
+    %
+    & =
+    %
+    \cdots
+\end{proof}
+%
+\xxx{work in progress...}
+In the expressions above, we define the vectors
+%
+\begin{align}
+    \pmb{\rho}^l
+     & \equiv
+    \int
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi) \,
+    \bar{\mathbf{e}}^l_r \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    \,
+\end{align}
+%
+and
+%
+\begin{align}
+    \bar{\mathbf{e}}^l_r
+     & \equiv
+    \mathbf{U}^l
+    \mathbf{e}^l_r
+    \quad.
+\end{align}
+%
+We may write the terms of $\pmb{\rho}^l$ as
+%
+\begin{align}
+    \rho^l_m
+     & =
+    \int
+    \sum\limits_{m'}
+    {D^l_{\hat{\mathbf{x}}}}_{m,m'}(\phi) \,
+    {\bar{e}^l}{{}_r}_{m'} \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    %
+    \nonumber \\[0.5em]
+    %
+     & =
+    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+    \sum\limits_{m'}
+    {\bar{e}^l}{{}_r}_{m'}
+    \mathrm{e}^{\frac{\imag \pi}{2}(m - m')}
+    \sum\limits_i c_{m,m',i}^{l}
+    \,
+    q^l_i(\pmb{\theta}_{\phi})
+    \quad,
+\end{align}
+%
+where
+%
+\begin{proof}{test_qli}
+    q^l_i(\pmb{\theta}_{\phi})
+    & =
+    \frac{1}{2}
+    \int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}
+    %
+    \mathrm{sgn}(\sin\phi)^{i}
+    \big| \sin\phi \big|
+    (\cos\phi)^{\alpha - 1}
+    (1 - \cos\phi)^{l + \beta - \frac{i}{2} - 1}
+    (1 + \cos\phi)^\frac{i}{2}
+    %
+    \,
+    \mathrm{d}\phi
+    \nonumber \\[0.5em]
+    %
+    & =
+    %
+    \begin{cases}
+        \displaystyle\int_{0}^{1}
+        %
+        x^{\alpha - 1}
+        (1 - x)^{l + \beta - \frac{i}{2} - 1 }
+        (1 + x)^\frac{i}{2}
+        %
+        \,
+        \mathrm{d}x
+        %
+         & i \,\, \mathrm{even}
+        %
+        \\
+        %
+        0
+        %
+         & i \,\, \mathrm{odd} \quad,
+    \end{cases}
+\end{proof}
+%
+and in the last line we made use of the transformation $x = \cos\phi$.
+%
+The integral in the expression above has a closed-form solution in terms
+of the hypergeometric function $_2F_1$:
+%
+\begin{proof}{test_qli}
+    %
+    \label{eq:qli}
+    q^l_i(\pmb{\theta}_{\phi})
+    & =
+    \resizebox{.75\hsize}{!}{$
+            \begin{cases}
+                \dfrac{\Gamma(\alpha)\Gamma(l + \beta - \frac{i}{2})}{\Gamma(l + \alpha + \beta - \frac{i}{2})}
+                \,
+                _2F_1\Big(
+                -\frac{i}{2},
+                \alpha;
+                l + \alpha + \beta - \frac{i}{2};
+                -1
+                \Big)
+                %
+                 & i \,\, \mathrm{even}
+                %
+                \\
+                %
+                0
+                %
+                 & i \,\, \mathrm{odd} \quad.
+            \end{cases}
+        $}
+\end{proof}
+%
+In order to compute the integral $\mathbf{e}_\phi$
+(Equation~\ref{eq:e2}) we must evaluate Equation~(\ref{eq:qli})
+for all $0 \leq l \leq l_{\mathrm{max}}$, $0 \leq i \leq 2l$,
+which can be done efficiently via upward recursion relations for both
+the gamma function and the hypergeometric function.
+
 \bibliography{bib}

### 2020-11-10 tex/tests/test_dlmmp.py

@@ -1,0 +1,76 @@
+import numpy as np
+from scipy.special import gamma as Gamma
+from scipy.special import beta as B
+from scipy.special import hyp2f1
+from sympy.physics.quantum.spin import Rotation
+from sympy import re
+from tqdm import tqdm
+
+
+def factorial(n):
+    """Just the factorial function, no frills."""
+    return Gamma(n + 1)
+
+
+def clmmpi(l, m, mp, i):
+    """The Wigner-d matrix coefficients c^l_{m, m', i}."""
+    if (m - mp - i) % 2 == 0:
+        return (
+            (-1) ** ((2 * l + 3 * m + mp - i) / 2)
+            * np.sqrt(
+                factorial(l - m)
+                * factorial(l + m)
+                * factorial(l - mp)
+                * factorial(l + mp)
+            )
+            / (
+                2 ** l
+                * factorial((i - m - mp) / 2)
+                * factorial((i + m + mp) / 2)
+                * factorial((2 * l - i - m + mp) / 2)
+                * factorial((2 * l - i + m - mp) / 2)
+            )
+        )
+    else:
+        return 0.0
+
+
+def dlmmp(l, m, mp, beta):
+    """The Wigner-d matrix term d^l_{m,m'}(beta)."""
+    d = 0
+    for i in range(2 * l + 1):
+        d += (
+            clmmpi(l, m, mp, i)
+            * np.sign(np.sin(beta)) ** i
+            * (1 - np.cos(beta)) ** ((2 * l - i) / 2)
+            * (1 + np.cos(beta)) ** (i / 2)
+        )
+    return d
+
+
+def dlmmp_sympy(l, m, mp, beta):
+    """
+    The Wigner-d matrix term d^l_{m,m'}(beta), evaluated using `sympy`.
+
+    Note that the transformation in `sympy` is a **passive** transformation,
+    (i.e., it is a rotation of the coordinate system, not of actual points
+    in 3D space), so we flip the sign of the angle of rotation when 
+    comparing to our implementation.
+    """
+    return np.float(re(Rotation.d(l, m, mp, -beta).doit().evalf()))
+
+
+def test_dlmmp(lmax=3, nbeta=3):
+    """
+    Compare our expression to the one in `sympy` up to degree `lmax`
+    for a few different values of `beta`.
+    """
+    np.random.seed(0)
+    for n in tqdm(range(nbeta)):
+        beta = np.random.uniform(-0.5 * np.pi, 0.5 * np.pi)
[…]

### 2020-11-11 tex/ms.tex
more text

@@
 
+\subsubsection{First moment}
+%
 Since the Wigner matrices are block diagonal, we may evaluate the moments of the
 distribution one spherical harmonic degree at a time. To that end, let us
-write
+write the first moment integral as
 %
@@
 %
-and
-%
-\begin{proof}{}
-    \mathbf{E}_\phi
-    & \equiv
-    \int
-    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{E}_r \,
-    \mathbf{E}_r^\top \,
-    \mathbf{R}_{\hat{\mathbf{x}}}^\top(\phi) \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
-    \mathrm{d}\phi
-    \nonumber
-    \\
-    %
-    & =
-    %
-    \setstackgap{L}{1.25\baselineskip}
-    \fixTABwidth{T}
-    \parenMatrixstack{
-    \mathbf{E}_\phi^{0,0} & \mathbf{E}_\phi^{0,1} & \mathbf{E}_\phi^{0,2} & \cdots & \mathbf{E}_\phi^{0,l_{\mathrm{max}}} \\
-    \mathbf{E}_\phi^{1,0} & \mathbf{E}_\phi^{1,1} & \mathbf{E}_\phi^{1,2} & \cdots & \mathbf{E}_\phi^{1,l_{\mathrm{max}}} \\
-    \mathbf{E}_\phi^{2,0} & \mathbf{E}_\phi^{2,1} & \mathbf{E}_\phi^{2,2} & \cdots & \mathbf{E}_\phi^{2,l_{\mathrm{max}}} \\
-    \vdots & \vdots & \vdots & \ddots & \vdots \\
-    \mathbf{E}_\phi^{l_{\mathrm{max}},0} & \mathbf{E}_\phi^{l_{\mathrm{max}},1} & \mathbf{E}_\phi^{l_{\mathrm{max}},2} & \cdots & \mathbf{E}_\phi^{l_{\mathrm{max}},l_{\mathrm{max}}}
-    }
-    \quad,
-\end{proof}
-%
 where
@@
     {\mathbf{U}^l}^{-1}
-    \pmb{\rho}^l
+    \mathbf{p}^l
     \quad,
@@
 %
-and
-%
-\begin{proof}{}
-    \mathbf{E}_\phi^{l,l'}
-    & =
-    \int
-    \mathbf{R}^l_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{e}^l_r \,
-    {\mathbf{R}^{l'}_{\hat{\mathbf{x}}}}^\top(\phi) \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
-    \mathrm{d}\phi
-    \nonumber \\
-    %
-    & =
-    %
-    \cdots
-\end{proof}
-%
-\xxx{work in progress...}
-In the expressions above, we define the vectors
+and we define
 %
 \begin{align}
-    \pmb{\rho}^l
+    \label{eq:pl}
+    \mathbf{p}^l
      & \equiv
@@
     \mathrm{d}\phi
-    \,
-\end{align}
-%
-and
-%
-\begin{align}
+    \\
     \bar{\mathbf{e}}^l_r
@@
 %
-We may write the terms of $\pmb{\rho}^l$ as
+The integral $\mathbf{p}^l$ defined above has a closed-form solution.
+To show this, we write the terms of $\mathbf{p}^l$ as
 %
 \begin{align}
-    \rho^l_m
+    p^l_m
      & =
     \int
-    \sum\limits_{m'}
-    {D^l_{\hat{\mathbf{x}}}}_{m,m'}(\phi) \,
-    {\bar{e}^l}{{}_r}_{m'} \,
+    \sum\limits_{\mu=-l}^l
+    {D^l_{\hat{\mathbf{x}}}}_{m,\mu}(\phi) \,
+    {\bar{e}^l}{{}_r}_{\mu} \,
     p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
@@
     \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
-    \sum\limits_{m'}
-    {\bar{e}^l}{{}_r}_{m'}
-    \mathrm{e}^{\frac{\imag \pi}{2}(m - m')}
-    \sum\limits_i c_{m,m',i}^{l}
+    \sum\limits_{\mu=-l}^l
+    {\bar{e}^l}{{}_r}_{\mu}
+    \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu)}
+    \sum\limits_{i=0}^{2l} c_{m,\mu,i}^{l}
     \,
@@
     q^l_i(\pmb{\theta}_{\phi})
-    & =
+    & \equiv
     \frac{1}{2}
@@
 
+\subsubsection{Second moment}
+%
+Similarly as before, let us write the second moment integral as
+%
+\begin{proof}{}
+    \mathbf{E}_\phi
+    & \equiv
+    \int
+    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
+    \mathbf{E}_r \,
+    \mathbf{E}_r^\top \,
+    \mathbf{R}_{\hat{\mathbf{x}}}^\top(\phi) \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    \nonumber
+    \\
+    %
+    & =
+    %
+    \setstackgap{L}{1.25\baselineskip}
+    \fixTABwidth{T}
+    \parenMatrixstack{
+    \mathbf{E}_\phi^{0,0} & \mathbf{E}_\phi^{0,1} & \mathbf{E}_\phi^{0,2} & \cdots & \mathbf{E}_\phi^{0,l_{\mathrm{max}}} \\
+    \mathbf{E}_\phi^{1,0} & \mathbf{E}_\phi^{1,1} & \mathbf{E}_\phi^{1,2} & \cdots & \mathbf{E}_\phi^{1,l_{\mathrm{max}}} \\
+    \mathbf{E}_\phi^{2,0} & \mathbf{E}_\phi^{2,1} & \mathbf{E}_\phi^{2,2} & \cdots & \mathbf{E}_\phi^{2,l_{\mathrm{max}}} \\
+    \vdots & \vdots & \vdots & \ddots & \vdots \\
+    \mathbf{E}_\phi^{l_{\mathrm{max}},0} & \mathbf{E}_\phi^{l_{\mathrm{max}},1} & \mathbf{E}_\phi^{l_{\mathrm{max}},2} & \cdots & \mathbf{E}_\phi^{l_{\mathrm{max}},l_{\mathrm{max}}}
+    }
+    \quad,
+\end{proof}
+%
+where
+%
+\begin{proof}{}
+    \mathbf{E}_\phi^{l,l'}
+    & =
+    \int
+    \mathbf{R}^l_{\hat{\mathbf{x}}}(\phi) \,
+    \mathbf{E}^l_r \,
+    {\mathbf{E}^l_r}^\top \,
+    {\mathbf{R}^{l'}_{\hat{\mathbf{x}}}}^\top(\phi) \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    \nonumber \\
+    %
+    & =
+    %
+    {\mathbf{U}^l}^{-1}
+    \mathbf{P}^l
+    {{\mathbf{U}^l}^{-1}}^\top
+\end{proof}
+%
+and we define
+%
+\begin{align}
+    \label{eq:Pl}
+    \mathbf{P}^l
+     & \equiv
+    \int
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi) \,
+    \bar{\mathbf{E}}^l_r \,
+    {\mathbf{D}^l_{\hat{\mathbf{x}}}}^\top(\phi) \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    \\
+    \bar{\mathbf{E}}^l_r
+     & \equiv
+    \mathbf{U}^l
+    \mathbf{E}^l_r
+    {\mathbf{E}^l_r}^\top
+    {\mathbf{U}^l}^\top
+    \quad.
+\end{align}
+%
+As before, we may express the solution to the integral $\mathbf{P}^l$ in
+closed form. Let us write the terms of $\mathbf{P}^l$ as
+%
+\begin{align}
+    P^l_{m,m'}
+     & =
+    \int
+    \sum\limits_{\mu=-l}^l
+    \sum\limits_{\nu=-l}^l
+    {D^l_{\hat{\mathbf{x}}}}_{m,\mu}(\phi) \,
+    {\bar{E}^l}{{}_r}_{\mu,\nu} \,
+    {D^l_{\hat{\mathbf{x}}}}_{\nu,m'}(\phi) \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\phi
+    %
+    \nonumber \\[0.5em]
+    %
+     & =
+    \resizebox{.75\hsize}{!}{$
+        \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+        \sum\limits_{\mu=-l}^l
+        \sum\limits_{\nu=-l}^l
+        {\bar{E}^l}{{}_r}_{\mu,\nu}
+        \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu + \nu - m')}
+        \sum\limits_{i=0}^{2l}
+        \sum\limits_{j=0}^{2l}
+        c_{m,\mu,i}^{l}
+        c_{\nu,m',j}^{l}
+        \,
+        Q^l_{i,j}(\pmb{\theta}_{\phi})
+    $}
+    \quad,
+\end{align}
+%
+where, similarly to before,
+%
+\begin{proof}{}
+    Q^l_{i,j}(\pmb{\theta}_{\phi})
+    & \equiv
+    \frac{1}{2}
+    \int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}
+    %
+    \mathrm{sgn}(\sin\phi)^{i+j}
+    \big| \sin\phi \big|
+    (\cos\phi)^{\alpha - 1}
+    (1 - \cos\phi)^{2l + \beta - \frac{i+j}{2} - 1}
+    (1 + \cos\phi)^\frac{i+j}{2}
+    %
+    \,
+    \mathrm{d}\phi
+    \nonumber \\[0.5em]
+    %
+    & =
+    %
+    \begin{cases}
+        \displaystyle\int_{0}^{1}
+        %
+        x^{\alpha - 1}
+        (1 - x)^{2l + \beta - \frac{i+j}{2} - 1 }
+        (1 + x)^\frac{i+j}{2}
+        %
+        \,
+        \mathrm{d}x
+        %
+         & i+j \,\, \mathrm{even}
+        %
+        \\
+        %
+        0
+        %
+         & i+j \,\, \mathrm{odd} \quad.
+    \end{cases}
+\end{proof}
+%
+We may again express this integral in closed form:
+%
+\begin{proof}{}
+    %
+    \label{eq:Qlij}
+    Q^l_{i,j}(\pmb{\theta}_{\phi})
+    & =
+    \resizebox{.75\hsize}{!}{$
+            \begin{cases}
+                \dfrac{\Gamma(\alpha)\Gamma(2l + \beta - \frac{i+j}{2})}{\Gamma(2l + \alpha + \beta - \frac{i+j}{2})}
+                \,
+                _2F_1\Big(
+                -\frac{i+j}{2},
+                \alpha;
+                2l + \alpha + \beta - \frac{i+j}{2};
+                -1
+                \Big)
+                %
+                 & i+j \,\, \mathrm{even}
+                %
+                \\
+                %
+                0
+                %
+                 & i+j \,\, \mathrm{odd} \quad.
+            \end{cases}
+        $}
+\end{proof}
+%
+As before, this integral may be evaluated recursively to efficiently compute all
+of the terms in $\mathbf{E}_\phi$.
+
 \bibliography{bib}

### 2020-11-11 tex/gitlinks.tex
better caching; more text; more proofs

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/77ecd5e5ec33f62d328f46350d17696547e5c0d3/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/77ecd5e5ec33f62d328f46350d17696547e5c0d3/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/77ecd5e5ec33f62d328f46350d17696547e5c0d3/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/183ebfe9e8b50599ac94eec0b58c35cbed8bce63/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/183ebfe9e8b50599ac94eec0b58c35cbed8bce63/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/183ebfe9e8b50599ac94eec0b58c35cbed8bce63/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-11 tex/ms.tex
better caching; more text; more proofs

@@
 The latitude integrals (Equations~\ref{eq:e2} and \ref{eq:E2}) involve
-rotations about $\hat{\mathbf{x}}$, which
+rotations by an angle $\phi$ about $\hat{\mathbf{x}}$, which
 may be accomplished by choosing
@@
 \begin{align}
-    \label{eq:pl}
-    \mathbf{p}^l
+    \label{eq:plphi}
+    \mathbf{p}^l_\phi
      & \equiv
@@
 %
-The integral $\mathbf{p}^l$ defined above has a closed-form solution.
-To show this, we write the terms of $\mathbf{p}^l$ as
+The integral $\mathbf{p}^l_\phi$ defined above has a closed-form solution.
+To show this, we write the terms of $\mathbf{p}^l_\phi$ as
 %
 \begin{align}
-    p^l_m
+    {p^l_\phi}_m
      & =
@@
     \,
-    q^l_i(\pmb{\theta}_{\phi})
+    {q^l_\phi}_i(\pmb{\theta}_{\phi})
     \quad,
@@
 \begin{proof}{test_qli}
-    q^l_i(\pmb{\theta}_{\phi})
+    {q^l_\phi}_i(\pmb{\theta}_{\phi})
     & \equiv
@@
 %
-\begin{proof}{test_qli}
+\begin{proof}{test_qlphii}
     %
-    \label{eq:qli}
-    q^l_i(\pmb{\theta}_{\phi})
+    \label{eq:qlphii}
+    {q^l_\phi}_i(\pmb{\theta}_{\phi})
     & =
@@
 In order to compute the integral $\mathbf{e}_\phi$
-(Equation~\ref{eq:e2}) we must evaluate Equation~(\ref{eq:qli})
+(Equation~\ref{eq:e2}) we must evaluate Equation~(\ref{eq:qlphii})
 for all $0 \leq l \leq l_{\mathrm{max}}$, $0 \leq i \leq 2l$,
@@
     {\mathbf{U}^l}^{-1}
-    \mathbf{P}^l
+    \mathbf{P}^l_\phi
     {{\mathbf{U}^l}^{-1}}^\top
@@
 \begin{align}
-    \label{eq:Pl}
-    \mathbf{P}^l
+    \label{eq:Plphi}
+    \mathbf{P}^l_\phi
      & \equiv
@@
 %
-As before, we may express the solution to the integral $\mathbf{P}^l$ in
-closed form. Let us write the terms of $\mathbf{P}^l$ as
+As before, we may express the solution to the integral $\mathbf{P}^l_\phi$ in
+closed form. Let us write the terms of $\mathbf{P}^l_\phi$ as
 %
 \begin{align}
-    P^l_{m,m'}
+    {P^l_\phi}_{m,m'}
      & =
@@
         \,
-        Q^l_{i,j}(\pmb{\theta}_{\phi})
+        {Q^l_\phi}_{i,j}(\pmb{\theta}_{\phi})
     $}
@@
 \begin{proof}{}
-    Q^l_{i,j}(\pmb{\theta}_{\phi})
+    {Q^l_\phi}_{i,j}(\pmb{\theta}_{\phi})
     & \equiv
@@
     %
-    \label{eq:Qlij}
-    Q^l_{i,j}(\pmb{\theta}_{\phi})
+    \label{eq:Qlphiij}
+    {Q^l_\phi}_{i,j}(\pmb{\theta}_{\phi})
     & =
@@
 
+\subsection{Longitude}
+%
+The longitude integrals (Equations~\ref{eq:e3} and \ref{eq:E3}) involve
+rotations by an angle $\lambda$ about $\hat{\mathbf{y}}$, which
+may be accomplished by choosing
+Euler angles $\alpha = 0$, $\beta = \lambda$, and
+$\gamma = 0$, such that
+%
+\begin{align}
+    \mathbf{R}^l_{\hat{\mathbf{y}}}(\lambda)
+     & =
+    {\mathbf{U}^l}^{-1} \mathbf{D}^l_{\hat{\mathbf{y}}}(\lambda) \mathbf{U}^l
+\end{align}
+%
+with
+\begin{align}
+    \mathbf{D}^l_{\hat{\mathbf{y}}}(\lambda)
+     & =
+    \mathbf{D}^l\left(0, \lambda, 0\right)
+    \quad.
+\end{align}
+%
+Since we expect the longitudinal distribution of features on the surfaces
+of stars to be (on average) isotropic, we will place a uniform prior on
+$\lambda \in [-\pi, \pi)$:
+%
+\begin{align}
+    p(\lambda \big| \pmb{\theta}_\lambda)
+     & =
+    \begin{cases}
+        \frac{1}{2\pi} & -\pi \leq \lambda < \pi
+        \\
+        0              & \mathrm{otherwise} \quad.
+    \end{cases}
+\end{align}
+%
+We therefore have no hyperparameters
+controlling the longitudinal distribution, i.e.,
+%
+\begin{align}
+    \pmb{\theta}_\lambda = \left( \,\,\, \right)
+    \quad.
+\end{align}
+%
+
+\subsubsection{First moment}
+%
+As before, we will solve for the terms of the moment integrals one
+spherical harmonic degree at a time:
+%
+\begin{proof}{test_elam}
+    \mathbf{e}_\lambda
+    & \equiv
+    \int
+    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{e}_\phi \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    \nonumber
+    \\
+    %
+    & =
+    %
+    \left(
+    \mathbf{e}_\lambda^0
+    \,\,\,
+    \mathbf{e}_\lambda^1
+    \,\,\,
+    \mathbf{e}_\lambda^2
+    \,\,\,
+    \cdots
+    \,\,\,
+    \mathbf{e}_\lambda^{l_{\mathrm{max}}}
+    \right)^\top
+    \quad,
+\end{proof}
+%
+where
+%
+\begin{proof}{test_elam}
+    \mathbf{e}_\lambda^l
+    & =
+    \int
+    \mathbf{R}^l_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{e}^l_\phi \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    \nonumber \\
+    %
+    & =
+    %
+    {\mathbf{U}^l}^{-1}
+    \mathbf{p}^l_\lambda
+    \quad,
+\end{proof}
+%
+and we define
+%
+\begin{align}
+    \label{eq:pllam}
+    \mathbf{p}^l_\lambda
+     & \equiv
+    \int
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(\lambda) \,
+    \bar{\mathbf{e}}^l_\phi \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    \\
+    \bar{\mathbf{e}}^l_\phi
+     & \equiv
+    \mathbf{U}^l
+    \mathbf{e}^l_\phi
+    \quad.
+\end{align}
+%
+The integral $\mathbf{p}^l_\lambda$ defined above has a closed-form solution.
+To show this, we write the terms of $\mathbf{p}^l_\lambda$ as
+%
+\begin{align}
+    {p^l_\lambda}_m
+     & =
+    \int
+    \sum\limits_{\mu=-l}^l
+    {D^l_{\hat{\mathbf{y}}}}_{m,\mu}(\lambda) \,
+    {\bar{e}^l}{{}_\phi}_{\mu} \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    %
+    \nonumber \\[0.5em]
+    %
+     & =
+    \sum\limits_{\mu=-l}^l
+    {\bar{e}^l}{{}_\phi}_{\mu}
+    \sum\limits_{i=0}^{2l} c_{m,\mu,i}^{l}
+    \,
+    {q^l_\lambda}_i
+    \quad,
+\end{align}
+%
+where
+%
+\begin{proof}{test_qllam}
+    {q^l_\lambda}_i
+    & \equiv
+    \frac{1}{2\pi}
+    \int_{-\pi}^{\pi}
+    %
+    \mathrm{sgn}(\sin\lambda)^{i}
+    (1 - \cos\lambda)^{l - \frac{i}{2}}
+    (1 + \cos\lambda)^\frac{i}{2}
+    %
+    \,
+    \mathrm{d}\lambda
+    \nonumber \\[0.5em]
+    %
+    & =
+    %
+    \begin{cases}
+        \dfrac{2^l \Gamma \left(\frac{1+i}{2}\right)
+            \Gamma \left(l+\frac{1-i}{2}\right)}{\pi  \Gamma (l+1)}
+        %
+         & i \,\, \mathrm{even}
+        %
+        \\
+        %
+        0
+        %
+         & i \,\, \mathrm{odd} \quad,
+    \end{cases}
+\end{proof}
+%
+whose terms may easily be computed by upward recursion.
+
 \bibliography{bib}

### 2020-11-11 starry_process/__init__.py

@@ -24,6 +24,7 @@ from . import (
     longitude,
     wigner,
     defaults,
+    cache,
 )
 from .sp import StarryProcess
 from .latitude import gauss2beta, beta2gauss

### 2020-11-11 starry_process/cache.py

@@ -1,0 +1,33 @@
+from pathlib import Path
+import numpy as np
+import functools
+
+
+def cache(*cache_args):
+    def decorator(func):
+        @functools.wraps(func)
+        def wrapper(self, *args, clobber=False, **kwargs):
+            try:
+                CACHE_DIR = Path.home() / ".starry_process"
+                CACHE_DIR.mkdir(exist_ok=True)
+                classname = str(self.__class__).split("'")[1]
+                file = CACHE_DIR / (
+                    "{}.{}-".format(classname, func.__name__)
+                    + "-".join(
+                        ["{}".format(getattr(self, arg)) for arg in cache_args]
+                    )
+                    + ".npz"
+                )
+                if (not clobber) and file.exists():
+                    return np.load(str(file))
+                else:
+                    results = func(self, *args, **kwargs)
+                    np.savez(file, **results)
+                    return results
+            except:
+                # caching failed!
+                return func(self, *args, **kwargs)
+
+        return wrapper
+
+    return decorator

### 2020-11-11 tex/tests/test_qllam.py

@@ -1,0 +1,31 @@
+import numpy as np
+from scipy.integrate import quad
+from scipy.special import gamma as Gamma
+
+
+def integrand(phi, l, i):
+    sin = np.sin(phi)
+    cos = np.cos(phi)
+    return (
+        np.sign(sin) ** i * (1 - cos) ** (l - i / 2) * (1 + cos) ** (i / 2)
+    ) / (2 * np.pi)
+
+
+def qli_numerical(l, i):
+    return quad(integrand, -np.pi, np.pi, args=(l, i))[0]
+
+
+def qli_closed_form(l, i):
+    return (
+        2 ** l
+        / np.pi
+        * Gamma(0.5 * (1 + i))
+        * Gamma(l + 0.5 * (1 - i))
+        / Gamma(l + 1)
+    )
+
+
+def test_qli(lmax=10):
+    for l in range(lmax + 1):
+        for i in range(0, 2 * l + 1, 2):
+            assert np.allclose(qli_closed_form(l, i), qli_numerical(l, i),)

### 2020-11-11 tex/gitlinks.tex
bugfix

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/183ebfe9e8b50599ac94eec0b58c35cbed8bce63/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/183ebfe9e8b50599ac94eec0b58c35cbed8bce63/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/183ebfe9e8b50599ac94eec0b58c35cbed8bce63/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/3e02de349ee1df17319a412d32355b7ceaeff3dd/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/3e02de349ee1df17319a412d32355b7ceaeff3dd/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/3e02de349ee1df17319a412d32355b7ceaeff3dd/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-11 tex/ms.tex
bugfix

@@
 
-\section{The Size and Contrast Integrals}
-\label{sec:size-contrast}
+\section{The Size Integral}
+\label{sec:size}
+%
+\xxx{Coming soon.}
 
@@
 %
-Our goal in this section is to compute the integrals...
+Our goal in this section is to compute the first and second moments
+of the latitude ($\mathbf{e}_\phi$, $\mathbf{E}_\phi$)
+and longitude ($\mathbf{e}_\lambda$, $\mathbf{E}_\lambda$) distributions
+(Equations \ref{eq:e2}, \ref{eq:e3}, \ref{eq:E2}, and \ref{eq:E3},
+respectively). These involve integrals over the terms in the Wigner
+rotation matrix for spherical harmonics, which we discuss below.
 
@@
 
+\subsubsection{Second moment}
+%
+We write the second moment integral as
+%
+\begin{proof}{}
+    \mathbf{E}_\lambda
+    & \equiv
+    \int
+    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{E}_\phi \,
+    \mathbf{E}_\phi^\top \,
+    \mathbf{R}_{\hat{\mathbf{y}}}^\top(\lambda) \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    \nonumber
+    \\
+    %
+    & =
+    %
+    \setstackgap{L}{1.25\baselineskip}
+    \fixTABwidth{T}
+    \parenMatrixstack{
+    \mathbf{E}_\lambda^{0,0} & \mathbf{E}_\lambda^{0,1} & \mathbf{E}_\lambda^{0,2} & \cdots & \mathbf{E}_\lambda^{0,l_{\mathrm{max}}} \\
+    \mathbf{E}_\lambda^{1,0} & \mathbf{E}_\lambda^{1,1} & \mathbf{E}_\lambda^{1,2} & \cdots & \mathbf{E}_\lambda^{1,l_{\mathrm{max}}} \\
+    \mathbf{E}_\lambda^{2,0} & \mathbf{E}_\lambda^{2,1} & \mathbf{E}_\lambda^{2,2} & \cdots & \mathbf{E}_\lambda^{2,l_{\mathrm{max}}} \\
+    \vdots & \vdots & \vdots & \ddots & \vdots \\
+    \mathbf{E}_\lambda^{l_{\mathrm{max}},0} & \mathbf{E}_\lambda^{l_{\mathrm{max}},1} & \mathbf{E}_\lambda^{l_{\mathrm{max}},2} & \cdots & \mathbf{E}_\lambda^{l_{\mathrm{max}},l_{\mathrm{max}}}
+    }
+    \quad,
+\end{proof}
+%
+where
+%
+\begin{proof}{}
+    \mathbf{E}_\lambda^{l,l'}
+    & =
+    \int
+    \mathbf{R}^l_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{E}^l_\phi \,
+    {\mathbf{E}^l_\phi}^\top \,
+    {\mathbf{R}^{l'}_{\hat{\mathbf{y}}}}^\top(\lambda) \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    \nonumber \\
+    %
+    & =
+    %
+    {\mathbf{U}^l}^{-1}
+    \mathbf{P}^l_\lambda
+    {{\mathbf{U}^l}^{-1}}^\top
+\end{proof}
+%
+and we define
+%
+\begin{align}
+    \label{eq:Pllam}
+    \mathbf{P}^l_\lambda
+     & \equiv
+    \int
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(\lambda) \,
+    \bar{\mathbf{E}}^l_\phi \,
+    {\mathbf{D}^l_{\hat{\mathbf{x}}}}^\top(\lambda) \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    \\
+    \bar{\mathbf{E}}^l_\phi
+     & \equiv
+    \mathbf{U}^l
+    \mathbf{E}^l_\phi
+    {\mathbf{E}^l_\phi}^\top
+    {\mathbf{U}^l}^\top
+    \quad.
+\end{align}
+%
+We then express the terms of $\mathbf{P}^l_\lambda$ as
+%
+\begin{align}
+    {P^l_\lambda}_{m,m'}
+     & =
+    \int
+    \sum\limits_{\mu=-l}^l
+    \sum\limits_{\nu=-l}^l
+    {D^l_{\hat{\mathbf{x}}}}_{m,\mu}(\lambda) \,
+    {\bar{E}^l}{{}_\phi}_{\mu,\nu} \,
+    {D^l_{\hat{\mathbf{x}}}}_{\nu,m'}(\lambda) \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    %
+    \nonumber \\[0.5em]
+    %
+     & =
+    \sum\limits_{\mu=-l}^l
+    \sum\limits_{\nu=-l}^l
+    {\bar{E}^l}{{}_\phi}_{\mu,\nu}
+    \sum\limits_{i=0}^{2l}
+    \sum\limits_{j=0}^{2l}
+    c_{m,\mu,i}^{l}
+    c_{\nu,m',j}^{l}
+    \,
+    {Q^l_\lambda}_{i,j}
+    \quad,
+\end{align}
+%
+where
+%
+\begin{proof}{}
+    {Q^l_\lambda}_{i,j}
+    & \equiv
+    \frac{1}{2\pi}
+    \int_{-\pi}^{\pi}
+    %
+    \mathrm{sgn}(\sin\lambda)^{i+j}
+    (1 - \cos\lambda)^{2l - \frac{i+j}{2}}
+    (1 + \cos\lambda)^\frac{i+j}{2}
+    %
+    \,
+    \mathrm{d}\lambda
+    \nonumber \\[0.5em]
+    %
+    & =
+    %
+    \begin{cases}
+        \dfrac{2^{2l} \Gamma \left(\frac{1+i+j}{2}\right)
+            \Gamma \left(2l+\frac{1-i-j}{2}\right)}{\pi  \Gamma (2l+1)}
+        %
+         & i+j \,\, \mathrm{even}
+        %
+        \\
+        %
+        0
+        %
+         & i+j \,\, \mathrm{odd} \quad,
+    \end{cases}
+\end{proof}
+%
+whose terms may again be computed by upward recursion.
+
+\section{The Contrast Integral}
+\label{sec:contrast}
+%
+\xxx{Coming soon.}
+
 \bibliography{bib}

### 2020-11-11 tests/test_sum.py

@@ -29,9 +29,9 @@ def test_sample(tol=5):
     )
 
     # Get the 3 lowest local minima
-    grad = np.gradient(Ilat)
+    grad = np.gradient(I)
     idx = (grad[1:] > 0) & (grad[:-1] < 0)
-    k = np.argsort(Ilat[:-1][idx])[:3]
+    k = np.argsort(I[:-1][idx])[:3]
     min_lats = np.sort(lat[:-1][idx][k])
 
     # Now check that these are about (-45, 0, 45)

### 2020-11-11 tex/figures/nullspace.py

@@ -6,124 +6,73 @@ import numpy as np
 from tqdm import tqdm
 
 # Settings
-ninc = 8
+ninc = 50
 ntheta = 300
-ydeg = 6
+ydeg = 14
 L = 1e9
 C = 1
 
 # Compute the posterior shrinkage & fractional rank at each inclination
 map = starry.Map(ydeg, lazy=False)
-map15 = starry.Map(15, lazy=False)
-incs = np.linspace(20.0, 90.0, ninc)
+incs = np.linspace(0.0, 90.0, ninc)[::-1]
 theta = np.linspace(0, 360, ntheta, endpoint=False)
-A = [None for i in range(ninc)]
-A15 = [None for i in range(ninc)]
-S = [None for i in range(ninc + 1)]
-R15 = [None for i in range(ninc + 1)]
+A = []
+S = []
+Sc = []
+R = []
+Rc = []
 flux = np.random.randn(ntheta)
-for i, inc in enumerate(incs):
-    map.inc = inc
-    A[i] = map.design_matrix(theta=theta)
-    cho_C = starry.linalg.solve(
-        design_matrix=A[i], data=flux, C=C, L=L, N=map.Ny,
-    )[1].eval()
-    S[i] = 1 - np.diag(cho_C @ cho_C.T) / L
-    map15.inc = inc
-    A15[i] = map15.design_matrix(theta=theta)
-    R15[i] = np.linalg.matrix_rank(A15[i].T @ A15[i]) / map15.Ny
+for i, inc in tqdm(enumerate(incs), total=len(incs)):
 
-# Posterior shrinkage & fractional rank for all datasets combined
-cho_C = starry.linalg.solve(
-    design_matrix=np.vstack(A), data=np.tile(flux, ninc), C=C, L=L, N=map.Ny,
-)[1].eval()
-S[-1] = 1 - np.diag(cho_C @ cho_C.T) / L
-R15[-1] = np.linalg.matrix_rank(np.vstack(A15).T @ np.vstack(A15)) / map15.Ny
+    # At this inclination
+    map.inc = inc
+    A.append(map.design_matrix(theta=theta))
+    cho_C = starry.linalg.solve(
+        design_matrix=A[-1], data=flux, C=C, L=L, N=map.Ny,
+    )[1].eval()
+    S.append(1 - np.diag(cho_C @ cho_C.T) / L)
+
+    # Cumulative
+    cho_C = starry.linalg.solve(
[…]

### 2020-11-11 tex/gitlinks.tex
paper

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/3e02de349ee1df17319a412d32355b7ceaeff3dd/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/3e02de349ee1df17319a412d32355b7ceaeff3dd/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/3e02de349ee1df17319a412d32355b7ceaeff3dd/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/7c725300f4f9663d61e19d5d1b01bb3daeab4b15/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/7c725300f4f9663d61e19d5d1b01bb3daeab4b15/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/7c725300f4f9663d61e19d5d1b01bb3daeab4b15/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-12 tex/gitlinks.tex
new figure

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/7c725300f4f9663d61e19d5d1b01bb3daeab4b15/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/7c725300f4f9663d61e19d5d1b01bb3daeab4b15/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/7c725300f4f9663d61e19d5d1b01bb3daeab4b15/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/e1b3d9d74fa35b77f657f5abedc00dc277e99e6d/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/e1b3d9d74fa35b77f657f5abedc00dc277e99e6d/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/e1b3d9d74fa35b77f657f5abedc00dc277e99e6d/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-12 tex/ms.tex
paper
new figure

@@
 
+\appendix
+
 \section{Spherical Harmonics}
@@
 %
-\xxx{Introduce the spherical harmonics.}
-
-The real spherical harmonics are indexed by their degree $l \in [0, \infty]$
-and order $m \in [-l, l]$. It is convenient to collect the spherical harmonic coefficients of
+The spherical harmonics $Y_{l,m}$ are a complete set of orthogonal functions on the
+surface of the sphere and, as such, are a convenient basis in which to construct
+our gaussian process. As in \citet{Luger2019}, here we describe the intensity
+on the surface of a star as an expansion in the real spherical harmonics, which is
+fully specified by a set of spherical harmonic coefficients
+$\{ y_{l,m}\}$, where $l \in [0, l_{\mathrm{max}}]$ is known as
+the degree and
+$m \in [-l, l]$ is known as the order, with $l_{\mathrm{max}}$ equal to the highest
+degree in the expansion.
+It is often convenient to collect the spherical harmonic coefficients of
 a given expansion into a vector $\mathbf{y}$ indexed by a single
@@
 \end{align}
-
-\section{Gaussian Process}
+%
+\citet{Luger2019} derived the expression for the total disk-integrated
+flux $\mathbf{f}$ (i.e., the light curve)
+one would observe from a star rotating about a fixed axis
+if the stellar surface intensity is well described by the spherical harmonic
+coefficient vector $\mathbf{y}$:
+%
+\begin{align}
+    \label{eq:fAy}
+    \mathbf{f} = \pmb{\mathcal{A}} \, \mathbf{y}
+    \quad,
+\end{align}
+%
+where $\pmb{\mathcal{A}} = \pmb{\mathcal{A}}(t, i, P)$ is the \starry
+design matrix, a purely linear operator that transforms from the spherical
+harmonic basis to the flux basis; it is implicitly
+a function of the time $t$, the stellar inclination $i$, and the stellar
+rotation period $P$. For reference, the rows of $\pmb{\mathcal{A}}$ are given by
+Equation~(18) in \citet{Luger2019}.
+
+An in-depth treatment of the spherical harmonics is beyond the scope of this
+paper. For more information, including closed-form expressions for $Y_{l,m}$
+and details on their usage in modeling stellar surfaces, the reader is
+encouraged to consult \citet{Luger2019}.
+
+\section{A Gaussian Process for Star Spots}
 \label{sec:gp}
@@
 denote a vector of $K$ flux measurements at times
-$\left( t_0 \,  t_1 \,  \cdots \, t_K \right)^\top$.
+$\mathbf{t} = \left( t_0 \,  t_1 \,  \cdots \, t_K \right)^\top$.
 Conditioned on certain physical properties of the star, $\pmb{\theta}$,
@@
 %
-In \citet{Luger2019} we showed that $\mathbf{f}$ may be computed from a
-linear operation on the vector of spherical harmonic coefficients
-describing the surface, $\mathbf{y}$:
-%
-\begin{align}
-    \label{eq:fAy}
-    \mathbf{f} = \mathbf{A} \, \mathbf{y}
-    \quad,
-\end{align}
-%
-where $\mathbf{A}$ is the \starry design matrix, which is implicitly
-a function of $\pmb{\theta}$ (as it depends on the stellar inclination
-and rotation period, for example).
-%
-Given Equation~(\ref{eq:fAy}),
+Given the linear relationship between flux and spherical harmonic
+coefficients (Equation~\ref{eq:fAy}),
 we may write the mean and covariance of our flux GP as
@@
     \pmb{\mu}(\pmb{\theta})
-     & = \mathbf{A}(\pmb{\theta}) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta})
+     & = \pmb{\mathcal{A}}(\pmb{\theta}) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta})
     \\
     \pmb{\Sigma}(\pmb{\theta})
-     & = \mathbf{A}(\pmb{\theta}) \, \pmb{\Sigma}_{\mathbf{y}} \, \mathbf{A}^\top(\pmb{\theta})
+     & = \pmb{\mathcal{A}}(\pmb{\theta}) \, \pmb{\Sigma}_{\mathbf{y}} \, \pmb{\mathcal{A}}^\top(\pmb{\theta})
     \quad,
@@
 %
-We devote the next four sections to the computation of these eight
+We devote the next several sections to the computation of these eight
 integrals.
@@
 %
-\xxx{Coming soon.}
+Our goal in this section is to compute the first and second moments
+of the size distribution ($\mathbf{e}_\phi$, $\mathbf{E}_\phi$) under
+a suitable spherical harmonic expansion $\mathbf{s}(r)$ of the spot profile
+and a suitable probability distribution function for the spot radius,
+$p(r \, \big| \, \pmb{\theta}_{r})$.
+
+\xxx{...}
+
+\subsection{Spot profile}
+%
+Let the brightness $b$ an angle $\vartheta$ away from the
+center of a spot of negative unit intensity and radius $r$ be
+%
+\begin{align}
+    b(\vartheta; r) & = \frac{1}{1 + \exp\left(\dfrac{r-\vartheta}{s}\right)} - 1
+\end{align}
+%
+for some (constant) shape parameter $s$. In the limit $s \rightarrow 0$, $b$ approaches an
+inverted top-hat function with half-width equal to $r$. For $s > 0$, each half
+of $b$ is a sigmoid with half-width at half-minimum equal to $r$. For definiteness,
+in our implementation of the algorithm we choose $s = 0.0033 \, \mathrm{rad}$,
+which is small compared to features of interest
+but not too small as to create numerical issues when computing
+model gradients.
+
+Our goal is to expand the function above in spherical harmonics. To that end,
+we note that in a frame where the spot is centered on $\hat{\mathbf{z}}$
+(i.e., at polar angle $\vartheta = 0$), the brightness profile is azimuthally
+symmetric, so the only nonzero coefficients in the spherical harmonic
+expansion are those with order $m = 0$. The corresponding spherical harmonics are
+simply proportional to the Legendre polynomials in $\cos\vartheta$, so our task is
+simplified to finding the Legendre polynomial expansion of $b$.
+Define a vector $\pmb{\vartheta}$ of $K$ equally-spaced points
+between $0$ and $\pi$, with coefficients given by
+%
+\begin{align}
+    \vartheta_k = \frac{k\pi}{K-1}
+    \quad.
+\end{align}
+%
+We wish to model the brightness evaluated at each $\vartheta_k$
+as a weighted combination of Legendre polynomials,
+%
+\begin{align}
+    \mathbf{b}(\pmb{\vartheta}; r) = \mathbf{B}(\pmb{\vartheta}) \, \tilde{\mathbf{s}}(r)
+\end{align}
+%
+%
+where $\mathbf{B}(\pmb{\vartheta})$ is a design matrix whose columns are
+the weighted Legendre polynomials,
+%
+\begin{align}
+    B_{k,l}(\vartheta_k) & = \sqrt{2l + 1} \, P_l\left( \cos \vartheta_k \right)
+    \quad,
+\end{align}
+%
+and $\tilde{\mathbf{s}}(r)$ are the coefficients of the expansion.
+These are related to the full vector of spherical harmonic coefficients describing
+the spot, $\mathbf{s}(r)$, by
+%
+\begin{align}
+    s_{l,m}(r) = \tilde{s}_l(r) \delta_{m,0}
+    \quad.
+\end{align}
+%
+To find the coefficients $\tilde{\mathbf{s}}(r)$, we solve the (linear) inverse problem,
+%
+\begin{align}
+    \tilde{\mathbf{s}}(r) & = \mathbf{B}^+(\pmb{\vartheta}) \, \mathbf{b}(\pmb{\vartheta}; r)
+\end{align}
+%
+%
+where
+%
+\begin{align}
+    \mathbf{B}^+(\pmb{\vartheta}) & = \mathbf{S} \Big(\mathbf{B}^\top(\pmb{\vartheta}) \mathbf{B}(\pmb{\vartheta}) +
+    \epsilon \mathbf{I}\Big)^{-1} \mathbf{B}^\top(\pmb{\vartheta})
+\end{align}
+%
+is the smoothed pseudo-inverse of $\mathbf{B}(\pmb{\vartheta})$ with small regularization
+parameter $\epsilon$, $\mathbf{I}$ is the identity matrix, and
+$\mathbf{S}$ is a diagonal smoothing matrix with coefficients
+%
+\begin{align}
+    S_{k,l} = \exp\left[-\frac{l(l + 1)}{2\sigma^2}\right] \delta_{k,l}
+\end{align}
+%
+for smoothing strength $\sigma$. For $\epsilon \rightarrow 0$ and
+$\sigma \rightarrow \infty$,
+$\mathbf{B}^+(\pmb{\vartheta})$ is the exact pseudo-inverse of
+$\mathbf{B}(\pmb{\vartheta})$. However,
+$\epsilon > 0$ is chosen for improved numerical stability and
+$\sigma > 0$ is chosen to mitigate the effect of ringing in the solution.
+In practice, we obtain good results with $\epsilon \approx 10^{-9}$
+and $\sigma \approx 15$.
+
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/spot_profile.pdf}
+        \oscaption{spot_profile}{%
+            Intensity profiles for spots with different
+            radii $r$ computed at spherical harmonic
+            degree $l_{\mathrm{max}} = \LMAX$.
+            \label{fig:spot_profile}
+        }
+    \end{centering}
+\end{figure}
 
@@
 %
-\begin{proof}{test_qli}
+\begin{proof}{test_qlphii}
     {q^l_\phi}_i(\pmb{\theta}_{\phi})

### 2020-11-12 tex/gitlinks.tex
more text

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/e1b3d9d74fa35b77f657f5abedc00dc277e99e6d/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/e1b3d9d74fa35b77f657f5abedc00dc277e99e6d/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/e1b3d9d74fa35b77f657f5abedc00dc277e99e6d/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/0f5cd85041405565c68a93eca39244838420c99d/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/0f5cd85041405565c68a93eca39244838420c99d/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/0f5cd85041405565c68a93eca39244838420c99d/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-12 tex/ms.tex
more text

@@
 
-\xxx{...}
-
 \subsection{Spot profile}
 %
-Let the brightness $b$ an angle $\vartheta$ away from the
+We model the brightness $b$ an angle $\vartheta$ away from the
 center of a spot of negative unit intensity and radius $r$ be
@@
 \begin{align}
-    b(\vartheta; r) & = \frac{1}{1 + \exp\left(\dfrac{r-\vartheta}{s}\right)} - 1
+    b(r; \vartheta) & = \frac{1}{1 + \exp\left(\dfrac{r-\vartheta}{s}\right)} - 1
 \end{align}
@@
 for some (constant) shape parameter $s$. In the limit $s \rightarrow 0$, $b$ approaches an
-inverted top-hat function with half-width equal to $r$. For $s > 0$, each half
-of $b$ is a sigmoid with half-width at half-minimum equal to $r$. For definiteness,
-in our implementation of the algorithm we choose $s = 0.0033 \, \mathrm{rad}$,
-which is small compared to features of interest
-but not too small as to create numerical issues when computing
-model gradients.
-
-Our goal is to expand the function above in spherical harmonics. To that end,
+inverted top-hat function with half-width equal to $r$,
+corresponding to a circular spot of uniform intensity. For $s > 0$, each half
+of $b$ is a sigmoid with half-width at half-minimum equal to $r$.
+In our implementation of the algorithm we choose $s = 0.2^\circ$,
+which is small compared to features of interest but not too small as to
+create numerical issues when computing model gradients (which would be
+undefined at the spot boundary if the spot profile were truly an inverted top-hat).
+
+Our goal now is to expand the function above in spherical harmonics. To that end,
 we note that in a frame where the spot is centered on $\hat{\mathbf{z}}$
@@
 \begin{align}
-    \mathbf{b}(\pmb{\vartheta}; r) = \mathbf{B}(\pmb{\vartheta}) \, \tilde{\mathbf{s}}(r)
-\end{align}
-%
-%
-where $\mathbf{B}(\pmb{\vartheta})$ is a design matrix whose columns are
+    \mathbf{b}(r) = \mathbf{B} \, \tilde{\mathbf{s}}(r)
+\end{align}
+%
+%
+where $\mathbf{B}$ is a design matrix whose columns are
 the weighted Legendre polynomials,
@@
 \begin{align}
-    B_{k,l}(\vartheta_k) & = \sqrt{2l + 1} \, P_l\left( \cos \vartheta_k \right)
+    B_{k,l} & = \sqrt{2l + 1} \, P_l\left( \cos \vartheta_k \right)
     \quad,
@@
     s_{l,m}(r) = \tilde{s}_l(r) \delta_{m,0}
-    \quad.
-\end{align}
-%
+    \quad,
+\end{align}
+%
+where $\delta$ is the Kronecker delta function.
 To find the coefficients $\tilde{\mathbf{s}}(r)$, we solve the (linear) inverse problem,
@@
 \begin{align}
-    \tilde{\mathbf{s}}(r) & = \mathbf{B}^+(\pmb{\vartheta}) \, \mathbf{b}(\pmb{\vartheta}; r)
+    \tilde{\mathbf{s}}(r) & = \mathbf{B}^+ \, \mathbf{b}(r)
 \end{align}
@@
 \begin{align}
-    \mathbf{B}^+(\pmb{\vartheta}) & = \mathbf{S} \Big(\mathbf{B}^\top(\pmb{\vartheta}) \mathbf{B}(\pmb{\vartheta}) +
-    \epsilon \mathbf{I}\Big)^{-1} \mathbf{B}^\top(\pmb{\vartheta})
-\end{align}
-%
-is the smoothed pseudo-inverse of $\mathbf{B}(\pmb{\vartheta})$ with small regularization
+    \mathbf{B}^+ & = \mathbf{S} \Big(\mathbf{B}^\top \mathbf{B} +
+    \epsilon \mathbf{I}\Big)^{-1} \mathbf{B}^\top
+\end{align}
+%
+is the smoothed pseudo-inverse of $\mathbf{B}$ with small regularization
 parameter $\epsilon$, $\mathbf{I}$ is the identity matrix, and
@@
 $\sigma \rightarrow \infty$,
-$\mathbf{B}^+(\pmb{\vartheta})$ is the exact pseudo-inverse of
-$\mathbf{B}(\pmb{\vartheta})$. However,
+$\mathbf{B}^+$ is the exact pseudo-inverse of
+$\mathbf{B}$. However,
 $\epsilon > 0$ is chosen for improved numerical stability and
@@
             degree $l_{\mathrm{max}} = \LMAX$.
+            For $r \gtrsim 10^\circ$, the spherical harmonic expansion
+            captures the spot shape and intensity reasonably well,
+            albeit with some ringing due to the truncated expansion.
             \label{fig:spot_profile}
@@
 \end{figure}
+
+Figure~\ref{fig:spot_profile} shows the intensity profile for spots of
+different radii expanded to spherical harmonic degree $l_{\mathrm{max}} = \LMAX$,
+computed from
+%
+\begin{align}
+    \mathbf{b'}(r) & = \mathbf{B}
+    \mathbf{B}^+ \, \mathbf{b}(r) \quad.
+\end{align}
+%
+The average intensity
+within the spots is close to $-1$ and the half-widths at half-minimum are
+equal to the spot radii, as expected.
+The effect of ringing due to the truncated spherical harmonic expansion is
+evident, although it is strongly suppressed compared to an expansion
+without the smoothing term (i.e., $\sigma = \infty$). However, for
+$r \lesssim 10^\circ$, an expansion to $l_{\mathrm{max}} = \LMAX$ is
+insufficient to correctly model the spot, as can be seen from the
+$r = 5^\circ$ profile (dashed curve). Expansions to higher spherical harmonic
+degree allow one to model spots with radii smaller than $10^\circ$, although
+at increased computational cost and potential numerical stability issues;
+we discuss this point at length in \S\ref{sec:tiny-spots}.
+
+\subsection{Probability density function}
+%
+For simplicity, we will adopt a uniform probability distribution for the
+spot radius, characterized by a mean radius $r_0$ and a half-width $\Delta$:
+%
+\begin{align}
+    p(r \, \big| \, \pmb{\theta}_{r})
+    &=
+    \begin{cases}
+        \frac{1}{2\Delta} & r_0 - \Delta \leq r \leq r_0 + \Delta
+        \\
+        0 & \mathrm{otherwise}
+        \quad,
+    \end{cases}
+\end{align}
+%
+where the hyperparameters of the distribution are
+%
+\begin{align}
+    \pmb{\theta}_r = \left(
+    r_0 \, \, \, \,
+    \Delta \right)^\top
+    \quad.
+\end{align}
+%
+As we argue in the text, in practice it is often difficult to constrain the 
+moments of the radius distribution above the first (the mean). It is 
+therefore useful to also consider the limiting case of the radius distribution
+as $\Delta \rightarrow 0$, in which case the PDF becomes
+%
+\begin{align}
+    p(r \, \big| \, \pmb{\theta}_{r}, \Delta = 0)
+    &=
+    \delta(r - r_0)
+    \quad,
+\end{align}
+%
+where $\delta$ is the delta function.
+%
+
+\subsection{First moment}
+%
+The first moment of the radius distribution is (Equation~\ref{eq:e1})
+%
+\begin{align}
+    \mathbf{e}_r
+     & \equiv
+    \int
+    \mathbf{s}(r) \,
+    p(r \, \big| \, \pmb{\theta}_{r}) \,
+    \mathrm{d}r
+    \nonumber \\
+    &=
+    \frac{1}{2\Delta}
+    \int_{r_0 - \Delta}^{r_0 + \Delta}
+    \mathbf{s}(r) \,
+    \mathrm{d}r
+    \quad.
+\end{align}
+%
+Using the equations from the previous section, its components may be written
+%
+\begin{align}
+    {e_r}_{l,m}
+    &=
+    \frac{\delta_{m,0}}{2\Delta}
+    \int_{r_0 - \Delta}^{r_0 + \Delta}
+    \tilde{s}_{l,m}(r) \,
+    \mathrm{d}r
+    \nonumber \\
+    &=
+    \frac{\delta_{m,0}}{2\Delta}
+    \sum_{k=0}^{K-1} B^+_{l,m,k}
+    \int_{r_0 - \Delta}^{r_0 + \Delta}
+    b_{k}(r)
+    \mathrm{d}r
+    \nonumber \\
+    &=
+    \frac{\delta_{m,0} s}{2\Delta}
+    \sum_{k=0}^{K-1} B^+_{l,m,k}
+    \ln 
+    \left(
+        \frac{
+            1+\exp\left(\frac{r -\Delta -\vartheta_k}{s}\right)
+        }
+        {
+            1+\exp\left(\frac{r + \Delta -\vartheta_k}{s}\right)
+        }
+    \right)
+    \quad.
+\end{align}
+%
+In the limit $\Delta \rightarrow 0$,
+%
+\begin{align}
+    \lim_{\Delta \rightarrow 0}
+    \mathbf{e}_r
+    &=
+    \mathbf{B}^+ \mathbf{b}(r)
+    \quad.
+\end{align}
+%
+
+\subsection{Second moment}
+%
+The second moment of the radius distribution is (Equation~\ref{eq:E1})
+%
+\begin{align}
+    \mathbf{E}_r
+     & \equiv
+    \int
+    \mathbf{s}(r) \,
+    \mathbf{s}^\top(r) \,
+    p(r \, \big| \, \pmb{\theta}_{r}) \,
+    \mathrm{d}r
+    \nonumber \\
+    &=
+    \frac{1}{2\Delta}
+    \int_{r_0 - \Delta}^{r_0 + \Delta}
+    \mathbf{s}(r) \,
+    \mathbf{s}^\top(r) \,
+    \mathrm{d}r
+    \quad.
+\end{align}
+%
+As before, its components may be written
+%
+%
+\begin{align}
+    {E_r}_{l,m,l',m'}
+    &=
+    \frac{\delta_{m,0}\delta_{m',0}}{2\Delta}
+    \int_{r_0 - \Delta}^{r_0 + \Delta}
+    \tilde{s}_{l,m}(r) \,
+    \tilde{s}_{l',m'}(r) \,
+    \mathrm{d}r
+    \nonumber \\
+    &=
+    \frac{\delta_{m,0}\delta_{m',0}}{2\Delta}
+    \sum_{k=0}^{K-1} B^+_{l,m,k}
+    \sum_{k'=0}^{K-1} B^+_{l',m',k'}
+    \int_{r_0 - \Delta}^{r_0 + \Delta}
+    b_{k}(r)
+    b_{k'}(r)
+    \mathrm{d}r
+    \nonumber \\
+    &= 
+    \cdots
+\end{align}
+\xxx{...}
 

### 2020-11-12 tex/figures/spot_profile.py

@@ -1,0 +1,74 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy.special import legendre as P
+
+
+def BBp(ydeg=15, npts=1000, eps=1e-9, sigma=15, **kwargs):
+    """
+    Return the matrix product B . B+. This expands the 
+    spot profile `b` in Legendre polynomials, then projects
+    back onto intensity space.
+
+    """
+    theta = np.linspace(0, np.pi, npts)
+    cost = np.cos(theta)
+    B = np.hstack(
+        [
+            np.sqrt(2 * l + 1) * P(l)(cost).reshape(-1, 1)
+            for l in range(ydeg + 1)
+        ]
+    )
+    BInv = np.linalg.solve(B.T @ B + eps * np.eye(ydeg + 1), B.T)
+    l = np.arange(ydeg + 1)
+    i = l * (l + 1)
+    S = np.exp(-0.5 * i / sigma ** 2)
+    BInv = S[:, None] * BInv
+    return B @ BInv
+
+
+def b(theta, r, s=0.0033 * 180 / np.pi):
+    """
+    The sigmoid spot profile.
+
+    """
+    return 1 / (1 + np.exp((r - theta) / s)) - 1
+
+
+def b_ylm(theta, r, s=0.0033 * 180 / np.pi, **kwargs):
+    """
+    The spherical harmonic expansion of the spot profile,
+    projected back onto intensity space.
+    
+    """
+    return BBp(npts=len(theta), **kwargs) @ b(theta, r, s=s)
+
+
+fig, ax = plt.subplots(1)
+theta = np.linspace(-90, 90, 1000)
+r = [5, 10, 20, 30, 40]
+cmap = plt.get_cmap("plasma")
+color = [cmap(0.75 * (k / (len(r) - 1))) for k in range(len(r))]
+ls = ["--"] + ["-"] * (len(r) - 1)
+alpha = [0.25] + [1.0] * (len(r) - 1)
+label1 = r"$\mathbf{b}(\mathbf{\vartheta})$"
+label2 = r"$\mathbf{B} \mathbf{B}^+ \mathbf{b}(\mathbf{\vartheta})$"
+
+for k in range(len(r)):
+    ax.plot(
+        theta,
+        b_ylm(np.abs(theta), r[k]),
+        ls=ls[k],
+        alpha=alpha[k],
+        color=color[k],
[…]

### 2020-11-13 tex/gitlinks.tex
more text

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/0f5cd85041405565c68a93eca39244838420c99d/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/0f5cd85041405565c68a93eca39244838420c99d/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/0f5cd85041405565c68a93eca39244838420c99d/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/3c024c44d780c51442e91d488e120a8eb4a05da5/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/3c024c44d780c51442e91d488e120a8eb4a05da5/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/3c024c44d780c51442e91d488e120a8eb4a05da5/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-13 tex/ms.tex
more text

@@
 \begin{align}
-    \mathbf{x} = \left( c \,\, \lambda \,\, \phi \,\, r \right)^\top
+    \mathbf{x} = 
+    \left( 
+        c_0 \,\, \cdots \,\, c_{N-1} \,\, 
+        ,
+        \lambda_0 \,\, \cdots \,\, \lambda_{N-1} \,\, 
+        ,
+        \phi_0 \,\, \cdots \,\, \phi_{N-1} \,\, 
+        ,
+        r_0 \,\, \cdots \,\, r_{N-1} \,\, 
+    \right)^\top
 \end{align}
@@
     \mathbf{y}(\mathbf{x}) =
-    c
+    \sum_{n=0}^{N-1}
+    c_n
     \,
-    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda)
+    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda_n)
     \,
-    \mathbf{R}_{\hat{\mathbf{x}}}(\phi)
+    \mathbf{R}_{\hat{\mathbf{x}}}(\phi_n)
     \,
-    \mathbf{s}(r)
-    \quad,
-\end{align}
-%
-where $c$ is the contrast of a spot,
-$\lambda$ is its longitude, $\phi$ is its latitude,
-and $r$ is its radius.
-The vector function $\mathbf{s}(r)$
+    \mathbf{s}(r_n)
+    \quad,
+\end{align}
+%
+where $c_n$ is the contrast of the $n^\mathrm{th}$ spot,
+$\lambda_n$ is its longitude, $\phi_n$ is its latitude,
+$r_n$ is its radius, and $N$ is the total number of spots.
+The vector function $\mathbf{s}(r_n)$
 returns the spherical harmonic expansion of a negative unit brightness
-circular spot at $\lambda = \phi = 0$,
-$\mathbf{R}_{\hat{\mathbf{x}}}(\phi)$ is the Wigner matrix that rotates the
+circular spot of radius $r_n$ at $\lambda = \phi = 0$,
+$\mathbf{R}_{\hat{\mathbf{x}}}(\phi_n)$ is the Wigner matrix that rotates the
 expansion about $\hat{\mathbf{x}}$ such that the spot is centered at a
-latitude $\phi$, and $\mathbf{R}_{\hat{\mathbf{y}}}(\lambda)$ is the Wigner
+latitude $\phi_n$, and $\mathbf{R}_{\hat{\mathbf{y}}}(\lambda_n)$ is the Wigner
 matrix that then rotates the
 expansion about $\hat{\mathbf{y}}$ such that the spot is centered at a
-longitude $\lambda$; these three functions are detailed in the sections below.
+longitude $\lambda_n$; these three functions are detailed in the sections below.
 %
 Equation~(\ref{eq:RRs}) thus provides a way of converting a random variable
-$\mathbf{x}$ describing the size, brightness, and position of a spot to the
+$\mathbf{x}$ describing the size, brightness, and position of spots to the
 corresponding representation in terms of spherical harmonics.
@@
 For simplicity, we assume that $p(\mathbf{x} \, \big| \, \pmb{\theta})$
-is separable in each of the four spot properties:
+is separable in each of the four spot properties, and that all of the spots
+are drawn from the same distribution:
 %
@@
     =
-    p(c \, \big| \, \pmb{\theta}_{c}) \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi})\,
-    p(r \, \big| \, \pmb{\theta}_{r})
+    \prod_{n=0}^{N-1}
+    p(c_n \, \big| \, \pmb{\theta}_{c}) \,
+    p(\lambda_n \, \big| \, \pmb{\theta}_{\lambda}) \,
+    p(\phi_n \, \big| \, \pmb{\theta}_{\phi})\,
+    p(r_n \, \big| \, \pmb{\theta}_{r})
     \quad,
@@
     \pmb{\theta}_{r} \right)^\top
-    \quad.
-\end{align}
+\end{align}
+%
+is the vector of yet-to-be-specified hyperparameters describing the process.
 %
@@
      & =
-    \mathbf{e}_c
+    N \mathbf{e}_c
     \\[1em]
@@
      & =
-    \mathbf{E}_c
+    N \mathbf{E}_c
 \end{align}
@@
     %
-    \quad.
-\end{align}
-%
+    \quad,
+\end{align}
+%
+In Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), we used the fact 
+that both the mean and the variance of the sum of $N$
+independent, identically-distributed random variables are equal to $N$
+times the individual mean and the variance, respectively.
+
 We devote the next several sections to the computation of these eight
@@
 \begin{align}
+    \label{eq:brvartheta}
     b(r; \vartheta) & = \frac{1}{1 + \exp\left(\dfrac{r-\vartheta}{s}\right)} - 1
@@
 \begin{align}
-    \mathbf{b}(r) = \mathbf{B} \, \tilde{\mathbf{s}}(r)
-\end{align}
-%
-%
-where $\mathbf{B}$ is a design matrix whose columns are
+    \mathbf{B} \, \tilde{\mathbf{s}}(r) = \mathbf{b}(r)
+\end{align}
+%
+%
+where $\mathbf{b}(r)$ is computed by evaluating Equation~(\ref{eq:brvartheta})
+at each of the $\vartheta_k$,
+$\mathbf{B}$ is a design matrix whose columns are
 the weighted Legendre polynomials,
@@
 \begin{align}
-    s_{l,m}(r) = \tilde{s}_l(r) \delta_{m,0}
+    s^l_m(r) = \tilde{s}_l(r) \delta_{m,0}
     \quad,
@@
 Figure~\ref{fig:spot_profile} shows the intensity profile for spots of
-different radii expanded to spherical harmonic degree $l_{\mathrm{max}} = \LMAX$,
-computed from
-%
-\begin{align}
-    \mathbf{b'}(r) & = \mathbf{B}
-    \mathbf{B}^+ \, \mathbf{b}(r) \quad.
-\end{align}
+different radii expanded to spherical harmonic degree $l_{\mathrm{max}} = \LMAX$.
+% computed from
+% %
+% \begin{align}
+%     \mathbf{b'}(r) & = \mathbf{B}
+%     \mathbf{B}^+ \, \mathbf{b}(r) \quad.
+% \end{align}
 %
@@
 \begin{align}
-    {e_r}_{l,m}
+    (e_r)^l_m
     &=
@@
     \int_{r_0 - \Delta}^{r_0 + \Delta}
-    \tilde{s}_{l,m}(r) \,
+    \tilde{s}_{l}(r) \,
     \mathrm{d}r
@@
     \frac{\delta_{m,0}}{2\Delta}
-    \sum_{k=0}^{K-1} B^+_{l,m,k}
+    \sum_{k=0}^{K-1} B^+_{l,k}
     \int_{r_0 - \Delta}^{r_0 + \Delta}
@@
     \frac{\delta_{m,0} s}{2\Delta}
-    \sum_{k=0}^{K-1} B^+_{l,m,k}
+    \sum_{k=0}^{K-1} B^+_{l,k}
     \ln 
@@
 \begin{align}
-    {E_r}_{l,m,l',m'}
+    (E_r)^{l,l'}_{m,m'}
     &=
@@
     \int_{r_0 - \Delta}^{r_0 + \Delta}
-    \tilde{s}_{l,m}(r) \,
-    \tilde{s}_{l',m'}(r) \,
+    \tilde{s}^l_{m}(r) \,
+    \tilde{s}^{l'}_{m'}(r) \,
     \mathrm{d}r
@@
     \frac{\delta_{m,0}\delta_{m',0}}{2\Delta}
-    \sum_{k=0}^{K-1} B^+_{l,m,k}
-    \sum_{k'=0}^{K-1} B^+_{l',m',k'}
+    \sum_{k=0}^{K-1} B^+_{l,k}
+    \sum_{k'=0}^{K-1} B^+_{l',k'}
     \int_{r_0 - \Delta}^{r_0 + \Delta}
@@
     &= 
-    \cdots
-\end{align}
-\xxx{...}
+    \frac{\delta_{m,0}\delta_{m',0}s}{2\Delta}
+    \sum_{k=0}^{K-1} B^+_{l,k}
+    \sum_{k'=0}^{K-1} B^+_{l',k'}
+    C_{k,k'}(r)
+    \quad,
+\end{align}
+%
+where
+%
+\begin{align}
+    C_{k,k'}(r)
+    &\equiv
+    \frac{
+        \exp\left(\frac{\vartheta_{k} - \vartheta_{k'}}{s}\right)
+        \ln 
+        \left(
+            \frac{
+                1+\exp\left(\frac{r -\Delta -\vartheta_{k}}{s}\right)
+            }
+            {
+                1+\exp\left(\frac{r + \Delta -\vartheta_{k}}{s}\right)
+            }
+        \right)
+        -
+        \ln 
+        \left(
+            \frac{
+                1+\exp\left(\frac{r -\Delta -\vartheta_{k'}}{s}\right)
+            }
+            {
+                1+\exp\left(\frac{r + \Delta -\vartheta_{k'}}{s}\right)
+            }
+        \right)
+    }{
+        1
+        -
+        \exp\left(\frac{\vartheta_{k} - \vartheta_{k'}}{s}\right)
+    }
+    \quad.
+\end{align}
+%
+In the limit $\Delta \rightarrow 0$,
+%
+\begin{align}
+    \lim_{\Delta \rightarrow 0}
+    \mathbf{E}_r
+    &=
+    \mathbf{B}^+ \mathbf{b}(r) \mathbf{b}^\top(r) {\mathbf{B}^+}^\top
+    \quad.
+\end{align}
+%
 

### 2020-11-13 tex/gitlinks.tex
fixed typos in paper; more text

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/3c024c44d780c51442e91d488e120a8eb4a05da5/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/3c024c44d780c51442e91d488e120a8eb4a05da5/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/3c024c44d780c51442e91d488e120a8eb4a05da5/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/150c31cb799f4eb003031cfb856cd3b53ab5f5fc/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/150c31cb799f4eb003031cfb856cd3b53ab5f5fc/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/150c31cb799f4eb003031cfb856cd3b53ab5f5fc/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-13 tex/ms.tex
fixed typos in paper; more text

@@
     \mathbf{E}_r \,
-    \mathbf{E}_r^\top \,
     \mathbf{R}_{\hat{\mathbf{x}}}^\top(\phi) \,
@@
     \mathbf{E}_\phi \,
-    \mathbf{E}_\phi^\top \,
     \mathbf{R}_{\hat{\mathbf{y}}}^\top(\lambda) \,
@@
     \mathbf{E}_\lambda \,
-    \mathbf{E}_\lambda^\top \,
     p(c \, \big| \, \pmb{\theta}_c)
@@
     p(r \, \big| \, \pmb{\theta}_{r})
-    &=
+     & =
     \begin{cases}
@@
     p(r \, \big| \, \pmb{\theta}_{r}, \Delta = 0)
-    &=
+     & =
     \delta(r - r_0)
@@
     \nonumber \\
-    &=
+     & =
     \frac{1}{2\Delta}
@@
     (e_r)^l_m
-    &=
+     & =
     \frac{\delta_{m,0}}{2\Delta}
@@
     \nonumber \\
-    &=
+     & =
     \frac{\delta_{m,0}}{2\Delta}
@@
     \nonumber \\
-    &=
+     & =
     \frac{\delta_{m,0} s}{2\Delta}
@@
     \mathbf{e}_r
-    &=
+     & =
     \mathbf{B}^+ \mathbf{b}(r)
@@
     (E_r)^{l,l'}_{m,m'}
-    &=
+     & =
     \frac{\delta_{m,0}\delta_{m',0}}{2\Delta}
@@
     \nonumber \\
-    &=
+     & =
     \frac{\delta_{m,0}\delta_{m',0}}{2\Delta}
@@
     \nonumber \\
-    &= 
+     & =
     \frac{\delta_{m,0}\delta_{m',0}s}{2\Delta}
@@
     C_{k,k'}(r)
-    &\equiv
+     & \equiv
     \frac{
@@
     \mathbf{E}_r
-    &=
+     & =
     \mathbf{B}^+ \mathbf{b}(r) \mathbf{b}^\top(r) {\mathbf{B}^+}^\top
@@
     \mathbf{E}^l_r \,
-    {\mathbf{E}^l_r}^\top \,
     {\mathbf{R}^{l'}_{\hat{\mathbf{x}}}}^\top(\phi) \,
@@
     \mathbf{E}^l_r
-    {\mathbf{E}^l_r}^\top
     {\mathbf{U}^l}^\top
@@
     \mathbf{E}^l_\phi \,
-    {\mathbf{E}^l_\phi}^\top \,
     {\mathbf{R}^{l'}_{\hat{\mathbf{y}}}}^\top(\lambda) \,
@@
     \mathbf{E}^l_\phi
-    {\mathbf{E}^l_\phi}^\top
     {\mathbf{U}^l}^\top
@@
 %
-\xxx{Coming soon.}
-
+The final integrals we must take in our computation of
+$\pmb{\mu}_{\mathbf{y}}(\pmb{\theta})$
+amd $\pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta})$ are the integrals over
+the spot contrast distribution, $\mathbf{e}_c$ and $\mathbf{E}_c$.
+These are by far the easiest, since the spot contrast is a scalar
+multiplier of the spherical harmonic coefficient vector, so we can
+pull the terms $\mathbf{e}_\lambda$ and $\mathbf{E}_\lambda$ out
+of the
+integrals in Equations (\ref{eq:e4}) and (\ref{eq:E4}) to write
+%
+\begin{align}
+    \mathbf{e}_c
+     & \equiv
+    \mathbf{e}_\lambda \,
+    \int
+    c \,
+    p(c \, \big| \, \pmb{\theta}_{c}) \,
+    \mathrm{d}c
+    \\
+    \mathbf{E}_c
+     & \equiv
+    \mathbf{E}_\lambda \,
+    \int
+    c^2 \,
+    p(c \, \big| \, \pmb{\theta}_c)
+    \mathrm{d}c
+    \quad.
+\end{align}
+%
+These integrals may be computed analytically for any choice of
+probability density function $p(c \, \big| \, \pmb{\theta}_c)$
+with closed-form moments.
+%
+However, in practice, it is quite difficult to constrain the
+spot contrast from light curves, let alone higher moments of its
+distribution; this is due largely to the fact that the contrast $c$
+is extremely degenerate with the total number of spots $N$
+(see \S\ref{sec:cN-degeneracy}).
+%
+In our imeplementation of the algorithm,
+we therefore choose the simplest possible probability distribution,
+a delta function:
+%
+\begin{align}
+    p(c \, \big| \, \pmb{\theta}_{c}) = \delta(c - c_0)
+\end{align}
+%
+characterized by a single parameter, the contrast of the spots:
+%
+\begin{align}
+    \pmb{\theta}_c = \left( c_0 \right)^\top
+    \quad.
+\end{align}
+%
+The moment integrals are then trivial to evaluate:
+%
+\begin{align}
+    \mathbf{e}_c & = c \, \mathbf{e}_\lambda
+    \\
+    \mathbf{E}_c & = c^2 \mathbf{E}_\lambda
+    \quad.
+\end{align}
+%
 \bibliography{bib}

### 2020-11-13 tex/gitlinks.tex
more text

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/150c31cb799f4eb003031cfb856cd3b53ab5f5fc/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/150c31cb799f4eb003031cfb856cd3b53ab5f5fc/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/150c31cb799f4eb003031cfb856cd3b53ab5f5fc/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/bea234c69e977f48e7acb13bb803f991e89aa4e9/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/bea234c69e977f48e7acb13bb803f991e89aa4e9/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/bea234c69e977f48e7acb13bb803f991e89aa4e9/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-13 tex/ms.tex
more text

@@
 %
-where $\pmb{\mathcal{A}} = \pmb{\mathcal{A}}(t, i, P)$ is the \starry
+where $\pmb{\mathcal{A}}$ is the \starry
 design matrix, a purely linear operator that transforms from the spherical
 harmonic basis to the flux basis; it is implicitly
-a function of the time $t$, the stellar inclination $i$, and the stellar
-rotation period $P$. For reference, the rows of $\pmb{\mathcal{A}}$ are given by
+a function of the stellar inclination $i$, the stellar
+rotation period $P$, the stellar limb darkening coefficients, etc.
+For reference, the rows of $\pmb{\mathcal{A}}$ are given by
 Equation~(18) in \citet{Luger2019}.
@@
 $\mathbf{t} = \left( t_0 \,  t_1 \,  \cdots \, t_K \right)^\top$.
-Conditioned on certain physical properties of the star, $\pmb{\theta}$,
-we wish to compute the mean $\pmb{\mu}(\pmb{\theta})$ and
-covariance $\pmb{\Sigma}(\pmb{\theta})$
+Conditioned on certain physical properties of the
+star
+(such as its inclination, rotation period, etc.), $\pmb{\theta}_\star$,
+and on certain properties of the  star spots
+(such as their sizes, positions, etc.), $\pmb{\theta}_\bullet$,
+we wish to compute the mean $\pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$ and
+covariance $\pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$
 of $\mathbf{f}$, which together fully specify our GP in flux.
@@
     \label{eq:mean}
-    \pmb{\mu}(\pmb{\theta})
-     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \pmb{\theta} \Big]
+    \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
+     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \pmb{\theta}_\star, \pmb{\theta}_\bullet \Big]
     \\
     \label{eq:cov}
-    \pmb{\Sigma}(\pmb{\theta})
-     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}(\pmb{\theta}) \pmb{\mu}^\top(\pmb{\theta})
+    \pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
+     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \pmb{\theta}_\star, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet) \pmb{\mu}^\top(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
     \quad.
@@
 \begin{align}
-    \pmb{\mu}(\pmb{\theta})
-     & = \pmb{\mathcal{A}}(\pmb{\theta}) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta})
-    \\
-    \pmb{\Sigma}(\pmb{\theta})
-     & = \pmb{\mathcal{A}}(\pmb{\theta}) \, \pmb{\Sigma}_{\mathbf{y}} \, \pmb{\mathcal{A}}^\top(\pmb{\theta})
+    \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
+     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
+    \\
+    \pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
+     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet) \, \pmb{\mathcal{A}}^\top(\pmb{\theta_\star})
     \quad,
@@
     \label{eq:mean_y}
-    \pmb{\mu}_{\mathbf{y}}(\pmb{\theta})
-     & = \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta} \Big]
+    \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
+     & = \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
     \\
     \label{eq:cov_y}
-    \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta})
-     & = \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big] - \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}) \pmb{\mu}_{\mathbf{y}}^\top(\pmb{\theta})
+    \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet)
+     & = \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] - \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet) \pmb{\mu}_{\mathbf{y}}^\top(\pmb{\theta}_\bullet)
 \end{align}
@@
     \label{eq:exp_y}
-    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta} \Big]
-     & =
-    \int \mathbf{y}(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta})\mathrm{d}\mathbf{x}
+    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
+     & =
+    \int \mathbf{y}(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
     \\
     \label{eq:exp_yy}
-    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big]
-     & =
-    \int \mathbf{y}(\mathbf{x} ) \mathbf{y}^\top(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta})\mathrm{d}\mathbf{x}
+    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
+     & =
+    \int \mathbf{y}(\mathbf{x} ) \mathbf{y}^\top(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
     \quad,
@@
 distribution of features on the surface  (i.e., the size and location of star spots)
-and $p(\mathbf{x} \, \big| \, \pmb{\theta})$ is its probability density
+and $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$ is its probability density
 function (PDF).
@@
 In the sections that follow we will show that for suitable choices of $\mathbf{y}(\mathbf{x})$
-and $p(\mathbf{x} \, \big| \, \pmb{\theta})$, the integrals in the expressions
+and $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$, the integrals in the expressions
 above have closed form solutions that may be evaluated quickly.
@@
 the probability distribution governing the different spot properties $\mathbf{x}$,
-i.e., $p(\mathbf{x} \, \big| \, \pmb{\theta})$.
-%
-For simplicity, we assume that $p(\mathbf{x} \, \big| \, \pmb{\theta})$
+i.e., $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$.
+%
+For simplicity, we assume that $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$
 is separable in each of the four spot properties, and that all of the spots
@@
 \begin{align}
-    p(\mathbf{x} \, \big| \, \pmb{\theta})
+    p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)
     =
@@
 \begin{align}
-    \pmb{\theta} = \left(
+    \pmb{\theta}_\bullet = \left(
     \pmb{\theta}_{c} \, \,
@@
     \label{eq:exp_y_sep}
-    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta} \Big]
+    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
      & =
@@
     \label{eq:exp_yy_sep}
-    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta} \Big]
+    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
      & =
@@
 
-We devote the next several sections to the computation of these eight
+We devote the next section to the computation of these eight
 integrals.
 
-\section{The Size Integral}
+\section{The Expectation Integrals}
+\label{sec:integrals}
+\subsection{The Size Integral}
 \label{sec:size}
@@
 
-\subsection{Spot profile}
+\subsubsection{Spot profile}
 %
@@
 
-\subsection{Probability density function}
+\subsubsection{Probability density function}
 %
@@
 
-\subsection{First moment}
+\subsubsection{First moment}
 %
@@
 
-\section{The Latitude and Longitude Integrals}
+\subsection{The Latitude and Longitude Integrals}
 \label{sec:lat-lon}
@@
 
-\subsection{Rotation matrices}
+\subsubsection{Rotation matrices}
 %
@@
 
-\subsection{Latitude}
+\subsubsection{Probability density function for the latitude}
 %
@@
 
-\subsubsection{First moment}
+\subsubsection{First moment of the latitude}
 %
@@
 
-\subsubsection{Second moment}
+\subsubsection{Second moment of the latitude}
 %
@@
 
-\subsection{Longitude}
+\subsubsection{Probability density function for the longitude}
 %
@@
 
-\subsubsection{First moment}
+\subsubsection{First moment of the longitude}
 %
@@
 
-\subsubsection{Second moment}
+\subsubsection{Second moment of the longitude}
 %
@@
 
-\section{The Contrast Integral}
+\subsection{The Contrast Integral}
 \label{sec:contrast}
@@
 %
+
+\section{Marginalizing over inclination}
+\xxx{Coming soon.}
+\begin{align}
+    \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
+     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
+    \\
+    \pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
+     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet) \, \pmb{\mathcal{A}}^\top(\pmb{\theta_\star})
+    \quad,
+\end{align}
+%
+\begin{align}
+    \pmb{\mu}(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet)
+     & =
+    \int
+    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
+    \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
+    p(i) \, \mathrm{d}i
+    \\
+    \pmb{\Sigma}(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet)
+     & =
+    \int
+    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
+    \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet) \, \pmb{\mathcal{A}}^\top(i, \pmb{\theta_\star})
+    p(i) \, \mathrm{d}i
+    \quad,
+\end{align}
+
 \bibliography{bib}

### 2020-11-16 tex/gitlinks.tex
more text; structural changes; better batch

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/bea234c69e977f48e7acb13bb803f991e89aa4e9/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/bea234c69e977f48e7acb13bb803f991e89aa4e9/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/bea234c69e977f48e7acb13bb803f991e89aa4e9/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/b677d4b247da6117fd73f66b451765dac961c376/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/b677d4b247da6117fd73f66b451765dac961c376/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/b677d4b247da6117fd73f66b451765dac961c376/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-16 tex/ms.tex
more text; structural changes; better batch

@@
 %
-The spherical harmonics $Y_{l,m}$ are a complete set of orthogonal functions on the
+The spherical harmonics are a complete set of orthogonal functions on the
 surface of the sphere and, as such, are a convenient basis in which to construct
@@
 fully specified by a set of spherical harmonic coefficients
-$\{ y_{l,m}\}$, where $l \in [0, l_{\mathrm{max}}]$ is known as
+$\{ y^l_m \}$, where $l \in [0, l_{\mathrm{max}}]$ is known as
 the degree and
@@
 It is often convenient to collect the spherical harmonic coefficients of
-a given expansion into a vector $\mathbf{y}$ indexed by a single
-integer $n$, where
-%
-\begin{align}
-    \label{eq:n}
-    n = l^2 + l + m
-\end{align}
-%
-and, conversely,
-%
-\begin{align}
-    \label{eq:lm}
-    \begin{split}
-        l & = \floor{\sqrt{n}}
-        \\
-        m & = n - l^2 - l
-        \quad.
-    \end{split}
+a given expansion into a vector
+%
+\begin{align}
+    \label{eq:y}
+    \mathbf{y}
+     & =
+    \left(
+    \mathbf{y}^0 \,\,\,
+    \mathbf{y}^1 \,\,\,
+    \cdots \,\,\,
+    \mathbf{y}^{l_\mathrm{max}}
+    \right)^\top
+    \nonumber \\
+     & =
+    \left(
+    y^0_0 \,\,\,
+    \,\,\,
+    y^1_{-1} \,\,\,
+    y^1_{0} \,\,\,
+    y^1_{1} \,\,\,
+    \,\,\,
+    \cdots \,\,\,
+    \,\,\,
+    y^{l_\mathrm{max}}_{-l_\mathrm{max}} \,\,\,
+    \cdots \,\,\,
+    y^{l_\mathrm{max}}_{l_\mathrm{max}} \,\,\,
+    \right)^\top
+    \quad.
 \end{align}
@@
 An in-depth treatment of the spherical harmonics is beyond the scope of this
-paper. For more information, including closed-form expressions for $Y_{l,m}$
+paper. For more information, including closed-form expressions for
+the spherical harmonics
 and details on their usage in modeling stellar surfaces, the reader is
@@
 \begin{align}
+    \label{eq:mean_f}
     \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
@@
     \\
+    \label{eq:cov_f}
     \pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
-     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet) \, \pmb{\mathcal{A}}^\top(\pmb{\theta_\star})
+     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet) \, \pmb{\mathcal{A}}^\top(\pmb{\theta}_\star)
     \quad,
@@
     %
-    \quad,
+    \quad.
 \end{align}
@@
 \subsection{The Size Integral}
-\label{sec:size}
+\label{sec:size}+
 %
@@
 %
-The integral $\mathbf{p}^l_\phi$ defined above has a closed-form solution.
+The integral $\mathbf{p}_\phi^l$ defined above has a closed-form solution.
 To show this, we write the terms of $\mathbf{p}^l_\phi$ as
@@
 \begin{align}
-    {p^l_\phi}_m
+    {({p^l_\phi})_{}}_m
      & =
@@
     \sum\limits_{\mu=-l}^l
-    {D^l_{\hat{\mathbf{x}}}}_{m,\mu}(\phi) \,
-    {\bar{e}^l}{{}_r}_{\mu} \,
+    {({D^l_{\hat{\mathbf{x}}}})_{}}_{m,\mu}(\phi) \,
+    {({\bar{e}}^l_r)_{}}_{\mu} \,
     p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
@@
     \sum\limits_{\mu=-l}^l
-    {\bar{e}^l}{{}_r}_{\mu}
+    {({\bar{e}}^l_r)_{}}_{\mu}
     \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu)}
@@
     \,
-    {q^l_\phi}_i(\pmb{\theta}_{\phi})
+    {({q^l_\phi})_{}}_i(\pmb{\theta}_{\phi})
     \quad,
@@
 \begin{proof}{test_qlphii}
-    {q^l_\phi}_i(\pmb{\theta}_{\phi})
+    {({q^l_\phi})_{}}_i(\pmb{\theta}_{\phi})
     & \equiv
@@
     \label{eq:qlphii}
-    {q^l_\phi}_i(\pmb{\theta}_{\phi})
+    {({q^l_\phi})_{}}_i(\pmb{\theta}_{\phi})
     & =
@@
 \begin{align}
-    {P^l_\phi}_{m,m'}
+    {({P^l_\phi})_{}}_{m,m'}
      & =
@@
     \sum\limits_{\nu=-l}^l
-    {D^l_{\hat{\mathbf{x}}}}_{m,\mu}(\phi) \,
-    {\bar{E}^l}{{}_r}_{\mu,\nu} \,
-    {D^l_{\hat{\mathbf{x}}}}_{\nu,m'}(\phi) \,
+    {({D^l_{\hat{\mathbf{x}}}})_{}}_{m,\mu}(\phi) \,
+    {(\bar{E}^l_r)_{}}_{\mu,\nu} \,
+    {({D^l_{\hat{\mathbf{x}}}})_{}}_{\nu,m'}(\phi) \,
     p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
@@
         \sum\limits_{\nu=-l}^l
-        {\bar{E}^l}{{}_r}_{\mu,\nu}
+        {(\bar{E}^l_r)_{}}_{\mu,\nu} \,
         \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu + \nu - m')}
@@
         \,
-        {Q^l_\phi}_{i,j}(\pmb{\theta}_{\phi})
+        {({Q^l_\phi})_{}}_{i,j}(\pmb{\theta}_{\phi})
     $}
@@
 \begin{proof}{}
-    {Q^l_\phi}_{i,j}(\pmb{\theta}_{\phi})
+    {({Q^l_\phi})_{}}_{i,j}(\pmb{\theta}_{\phi})
     & \equiv
@@
     \label{eq:Qlphiij}
-    {Q^l_\phi}_{i,j}(\pmb{\theta}_{\phi})
+    {({Q^l_\phi})_{}}_{i,j}(\pmb{\theta}_{\phi})
     & =
@@
 \begin{align}
-    {p^l_\lambda}_m
+    {({p^l_\lambda})_{}}_m
      & =
@@
     \sum\limits_{\mu=-l}^l
-    {D^l_{\hat{\mathbf{y}}}}_{m,\mu}(\lambda) \,
-    {\bar{e}^l}{{}_\phi}_{\mu} \,
+    {({D^l_{\hat{\mathbf{y}}}})_{}}_{m,\mu}(\lambda) \,
+    {({\bar{e}^l_\phi})_{}}_{\mu} \,
     p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
@@
     \sum\limits_{\mu=-l}^l
-    {\bar{e}^l}{{}_\phi}_{\mu}
+    {({\bar{e}^l_\phi})_{}}_{\mu} \,
     \sum\limits_{i=0}^{2l} c_{m,\mu,i}^{l}
     \,
-    {q^l_\lambda}_i
+    {({q^l_\lambda})_{}}_i
     \quad,
@@
 \begin{proof}{test_qllam}
-    {q^l_\lambda}_i
+    {({q^l_\lambda})_{}}_i
     & \equiv
@@
 \begin{align}
-    {P^l_\lambda}_{m,m'}
+    {({P^l_\lambda})_{}}_{m,m'}
      & =
@@
     \sum\limits_{\nu=-l}^l
-    {D^l_{\hat{\mathbf{x}}}}_{m,\mu}(\lambda) \,
-    {\bar{E}^l}{{}_\phi}_{\mu,\nu} \,
-    {D^l_{\hat{\mathbf{x}}}}_{\nu,m'}(\lambda) \,
+    ({D^l_{\hat{\mathbf{x}}}})_{m,\mu}(\lambda) \,
+    (\bar{E}^l_\phi)_{\mu,\nu} \,
+    ({D^l_{\hat{\mathbf{x}}}})_{\nu,m'}(\lambda) \,
     p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
@@
     \sum\limits_{\nu=-l}^l
-    {\bar{E}^l}{{}_\phi}_{\mu,\nu}
+    (\bar{E}^l_\phi)_{\mu,\nu}
     \sum\limits_{i=0}^{2l}
@@
     \,
-    {Q^l_\lambda}_{i,j}
+    (Q^l_\lambda)_{i,j}
     \quad,
@@
 \begin{proof}{}
-    {Q^l_\lambda}_{i,j}
+    (Q^l_\lambda)_{i,j}
     & \equiv
@@
 \begin{align}
-    \mathbf{e}_c & = c \, \mathbf{e}_\lambda
-    \\
-    \mathbf{E}_c & = c^2 \mathbf{E}_\lambda
+    \mathbf{e}_c & = c_0 \, \mathbf{e}_\lambda
+    \\
+    \mathbf{E}_c & = c_0^2 \mathbf{E}_\lambda
     \quad.
@@
 \section{Marginalizing over inclination}
+%
+The equations for the mean and covariance of the flux GP
+(Equations~\ref{eq:mean_f} and \ref{eq:cov_f}, respectively) are conditioned
+on specific values of the stellar properties $\pmb{\theta}_\star$ and the
+spot properties $\pmb{\theta}_\bullet$. To obtain the posterior distribution
+for these parameters, we must typically resort to numerical sampling techniques,
+which often scale steeply with the number of parameters. It is therefore generally
+desirable to keep the total number of parameters small, especially when
+employing the GP in an ensemble setting.
+In such a setting, we might have light curves from $N_\star$ stars, all of which
+we believe to have similar spot properties (perhaps because they have
+similar spectral types and rotation periods, for example).
+The total number of parameters in our problem is therefore
+%
+\begin{align}
+    N_\theta = N_\star \, \mathrm{dim}(\pmb{\theta}_\star) + 5
+    \quad,
+\end{align}
+%
+since each of the stars will have their own set of stellar properties
+$\pmb{\theta}_\star$ but will all share the same 5 spot properties
+$\pmb{\theta}_\bullet$ (by assumption).
+In a typical problem, the stellar parameters might be
+%
+\begin{align}
+    \pmb{\theta}_\star =
+    \left(
+    i \,\,\,
+    P \,\,\,
+    u_1 \,\,\,
+    u_2
+    \right)^\top
+    \quad,
+\end{align}
+%
+i.e., the inclination, the rotational period, and some (usually two) limb
+darkening coefficients. Thus, for a reasonably sized ensemble of $N_\star=100$
+stars, we would have to sample over $N_\theta = 405$ parameters.
+While large, this number is certainly not absurd, especially by modern standards.
+However, it does pose
+a problem when considering how complex the posterior distribution for the
+spot mapping problem can be. In addition to strong nonlinear degeneracies
+between some of the parameters (such as the contrast $c$ and the
+number of spots $N$), the posterior is often multimodal, especially in the
+stellar inclinations. While modern sampling schemes such as Hamiltonian
+Monte Carlo and Nested Sampling may in principle be able to deal with these
+issues, in practice it can be very difficult to obtain convergence in a
+reasonable amount of time.
+
+One workaround is to fix the values of $\pmb{\theta}_\star$. This could be done,
+for instance, to the rotational period $P$, which can often be estimated with
+fairly good accuracy from a periodogram. The limb darkening coefficients could
+be fixed at theoretical values, or perhaps their values could be shared among
+all stars (and sampled over), given the similarity assumption above.
+
+The inclination, however, is a different matter. Absent prior information
+for a particular star (such as a measurement of its projected rotational
+velocity $v\sin i$ or the knowledge that it hosts a transiting planet), it
+is simply not possible to reliably estimate the inclination in a
+pre-processing step. Any light curve statistic one might argue should scale
+with inclination---such as the amplitude of the variability---is invariably
+degenerate with the spot parameters $\pmb{\theta}_\bullet$. If one knew
+$\pmb{\theta}_\bullet$, then perhaps a decent point estimate of $i$ could
+be obtained, but in that case the analysis wouldn't be needed in the first place!
+%
+
+Fortunately, there is a better way to reduce the number of parameters in
+the problem: we can explicitly marginalize over the stellar inclination.
+That is, we can write the first two moments of our flux GP as
+%
+\begin{align}
+    \label{eq:mu_marg}
+    \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+     & =
+    \int
+    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
+    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
+    p(i) \, \mathrm{d}i \\
+    \label{eq:cov_marg}
+    \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+     & =
+    \int
+    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
+    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
+    \pmb{\mathcal{A}}^\top(i, \bar{\pmb{\theta}}_\star)
+    p(i) \, \mathrm{d}i    \quad,
+\end{align}
+%
+where the expectations inside the integrals are given by
+Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), respectively, and
+$\bar{\pmb{\theta}}_\star$ are all the stellar parameters \emph{except}
+the inclination, i.e.,
+%
+\begin{align}
+    \bar{\pmb{\theta}}_\star = \left(
+    P \,\,\,
+    u_1 \,\,\,
+    u_2
+    \right)^\top
+    \quad,
+\end{align}
+%
+where perhaps $P$ is fixed for each star and the limb darkening coefficients
+are shared among all stars, as described above.
+%
+Thus, if we are able to perform the integrals in Equations~(\ref{eq:mu_marg})
+and (\ref{eq:cov_marg}), we can \emph{dramatically} reduce the number of
+parameters in our ensemble problem.
+
+As we will see below, these integrals do in fact have closed-form solutions,
+since they are (again) integrals over Wigner matrices. In our derivation,
+we will assume that stellar rotation axes are distributed isotropically. In
+that case, the probability density for the inclination is simply
+%
+\begin{align}
+    p(i) = \sin i
+    \quad.
+\end{align}
+
+\subsection{First moment}
 \xxx{Coming soon.}
-\begin{align}
-    \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
-     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
-    \\
-    \pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
-     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet) \, \pmb{\mathcal{A}}^\top(\pmb{\theta_\star})
-    \quad,
-\end{align}
-%
-\begin{align}
-    \pmb{\mu}(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet)
-     & =
-    \int
-    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
-    \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
-    p(i) \, \mathrm{d}i
-    \\
-    \pmb{\Sigma}(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet)
-     & =
-    \int
-    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
-    \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet) \, \pmb{\mathcal{A}}^\top(i, \pmb{\theta_\star})
-    p(i) \, \mathrm{d}i
-    \quad,
-\end{align}
+
+\subsection{Second moment}
+\xxx{Coming soon.}
 

### 2020-11-16 starry_process/calibrate/__init__.py

@@ -5,8 +5,10 @@ from .plot import (
     plot_corner,
     plot_latitude_pdf,
     plot_inclination_pdf,
+    plot_batch,
 )
 from .sample import sample
 from .log_prob import get_log_prob
-from .optimize import optimize
 from .run import run
+from .inclination import compute_inclination_pdf
+from .batch import run_batch

### 2020-11-16 starry_process/calibrate/optimize.py

@@ -1,140 +1,0 @@
-from .defaults import update_with_defaults
-from .. import StarryProcess
-import pymc3 as pm
-from pymc3_ext import optim as op
-from ..math import cho_factor, cho_solve
-import numpy as np
-import theano
-import theano.tensor as tt
-from theano.ifelse import ifelse
-from tqdm.auto import tqdm
-
-
-def optimize(data, **kwargs):
-
-    # Get kwargs
-    kwargs = update_with_defaults(**kwargs)
-    seed = kwargs["seed"]
-    optim_kwargs = kwargs["optimize"]
-    gen_kwargs = kwargs["generate"]
-    normalized = gen_kwargs["normalized"]
-    niter = optim_kwargs["niter"]
-    ntries = optim_kwargs["ntries"]
-    adam_kwargs = dict(lr=optim_kwargs["lr"])
-    rmin = optim_kwargs["rmin"]
-    rmax = optim_kwargs["rmax"]
-    amin = optim_kwargs["amin"]
-    amax = optim_kwargs["amax"]
-    bmin = optim_kwargs["bmin"]
-    bmax = optim_kwargs["bmax"]
-    cmin = optim_kwargs["cmin"]
-    cmax = optim_kwargs["cmax"]
-    nmin = optim_kwargs["nmin"]
-    nmax = optim_kwargs["nmax"]
-    apply_jac = optim_kwargs["apply_jac"]
-    ydeg = optim_kwargs["ydeg"]
-    baseline_var = optim_kwargs["baseline_var"]
-
-    def get_guesses():
-        r0 = rmin + np.random.random() * (rmax - rmin)
-        a0 = amin + np.random.random() * (amax - amin)
-        b0 = bmin + np.random.random() * (bmax - bmin)
-        c0 = cmin + np.random.random() * (cmax - cmin)
-        n0 = nmin + np.random.random() * (nmax - nmin)
-        return r0, a0, b0, c0, n0
-
-    # Get the data
-    t = data["t"]
-    flux = data["flux"]
-    ferr = data["ferr"]
-    p = data["period"]
-    nlc = len(flux)
-
-    # Init
-    np.random.seed(seed)
-    r0, a0, b0, c0, n0 = get_guesses()
-
-    with pm.Model() as model:
-
-        # Vars
-        r = pm.Uniform("r", lower=rmin, upper=rmax, testval=r0)
[…]

### 2020-11-16 starry_process/calibrate/plot.py

@@ -2,6 +2,7 @@ from .defaults import update_with_defaults
 from .log_prob import get_log_prob
 import matplotlib
 import matplotlib.pyplot as plt
+from matplotlib import colors
 from mpl_toolkits.axes_grid1.axes_divider import make_axes_locatable
 from starry_process import StarryProcess
 from starry_process.latitude import beta2gauss, gauss2beta
@@ -13,9 +14,17 @@ import dynesty.plotting as dyplot
 from dynesty import utils as dyfunc
 from corner import corner as _corner
 from tqdm.auto import tqdm
+import glob
+import os
+import json
+import pickle
 
 
 def corner(*args, **kwargs):
+    """
+    Override `corner.corner` by making some appearance tweaks.
+
+    """
     # Get the usual corner plot
     figure = _corner(*args, **kwargs)
 
@@ -29,13 +38,17 @@ def corner(*args, **kwargs):
             tick.label.set_fontsize(10)
         formatter = matplotlib.ticker.ScalarFormatter(useOffset=False)
         ax.yaxis.set_major_formatter(formatter)
-        ax.set_ylabel(ax.get_ylabel(), fontsize=kwargs["corner_label_size"])
+        ax.set_ylabel(
+            ax.get_ylabel(), fontsize=kwargs.get("corner_label_size", 12)
+        )
     for ax in axes[-1, :]:
         for tick in ax.xaxis.get_major_ticks():
             tick.label.set_fontsize(10)
         formatter = matplotlib.ticker.ScalarFormatter(useOffset=False)
         ax.xaxis.set_major_formatter(formatter)
-        ax.set_xlabel(ax.get_xlabel(), fontsize=kwargs["corner_label_size"])
+        ax.set_xlabel(
+            ax.get_xlabel(), fontsize=kwargs.get("corner_label_size", 12)
+        )
 
     # Pad the axes to always include the truths
     truths = kwargs.get("truths", None)
@@ -82,7 +95,10 @@ def lat2y(lat):
 
 
 def plot_data(data, **kwargs):
+    """
+    Plot a synthetic dataset.
 
+    """
     # Get kwargs
     kwargs = update_with_defaults(**kwargs)
     plot_kwargs = kwargs["plot"]
@@ -171,6 +187,10 @@ def plot_data(data, **kwargs):
 
 
 def plot_latitude_pdf(results, **kwargs):
+    """
+    Plot posterior draws from the latitude hyperdistribution.
+
[…]

### 2020-11-16 starry_process/calibrate/run.py

@@ -16,9 +16,11 @@ def run(
     plot_latitude_pdf=True,
     plot_trace=False,
     plot_corner=False,
+    plot_corner_no_baseline=False,
     plot_corner_transformed=True,
-    plot_inclination=False,
-    **kwargs
+    plot_corner_transformed_no_baseline=True,
+    plot_inclination_pdf=True,
+    **kwargs,
 ):
     if not os.path.exists(path):
         os.makedirs(path)
@@ -51,6 +53,21 @@ def run(
     else:
         results = pickle.load(open(os.path.join(path, "results.pkl"), "rb"))
 
+    # Compute inclination pdf
+    compute_inclination_pdf = kwargs["sample"]["compute_inclination_pdf"]
+    if compute_inclination_pdf:
+        if clobber or not os.path.exists(
+            os.path.join(path, "inclinations.npz")
+        ):
+            inc_results = calibrate.compute_inclination_pdf(
+                data, results, **kwargs
+            )
+            np.savez(os.path.join(path, "inclinations.npz"), **inc_results)
+        else:
+            inc_results = np.load(os.path.join(path, "inclinations.npz"))
+    else:
+        inc_results = None
+
     # Transform latitude params and store posterior mean and cov
     samples = np.array(results.samples)
     samples[:, 1], samples[:, 2] = beta2gauss(samples[:, 1], samples[:, 2])
@@ -74,12 +91,31 @@ def run(
         fig = calibrate.plot_corner(results, transform_beta=False, **kwargs)
         fig.savefig(os.path.join(path, "corner.pdf"), bbox_inches="tight")
 
+    if plot_all or plot_corner_no_baseline:
+        fig = calibrate.plot_corner(
+            results, transform_beta=False, include_baseline=False, **kwargs
+        )
+        fig.savefig(
+            os.path.join(path, "corner_no_baseline.pdf"), bbox_inches="tight"
+        )
+
     if plot_all or plot_corner_transformed:
         fig = calibrate.plot_corner(results, transform_beta=True, **kwargs)
         fig.savefig(
             os.path.join(path, "corner_transformed.pdf"), bbox_inches="tight"
         )
 
-    if plot_all or plot_inclination:
-        fig = calibrate.plot_inclination_pdf(data, results, **kwargs)
-        fig.savefig(os.path.join(path, "inclination.pdf"), bbox_inches="tight")
+    if plot_all or plot_corner_transformed_no_baseline:
[…]

### 2020-11-25 tex/gitlinks.tex
tex updates

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/b677d4b247da6117fd73f66b451765dac961c376/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/b677d4b247da6117fd73f66b451765dac961c376/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/b677d4b247da6117fd73f66b451765dac961c376/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/09b20fbf11c7967ce0510c149223cc7406ca9eca/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/09b20fbf11c7967ce0510c149223cc7406ca9eca/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/09b20fbf11c7967ce0510c149223cc7406ca9eca/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-25 tex/ms.tex
tex updates

@@
 
-\appendix
-
-\section{Spherical Harmonics}
+\section{A Gaussian Process for Star Spots}
+
+\subsection{Spherical Harmonics}
 \label{sec:ylm}
@@
 our gaussian process. As in \citet{Luger2019}, here we describe the intensity
-on the surface of a star as an expansion in the real spherical harmonics, which is
-fully specified by a set of spherical harmonic coefficients
-$\{ y^l_m \}$, where $l \in [0, l_{\mathrm{max}}]$ is known as
-the degree and
-$m \in [-l, l]$ is known as the order, with $l_{\mathrm{max}}$ equal to the highest
-degree in the expansion.
-It is often convenient to collect the spherical harmonic coefficients of
-a given expansion into a vector
-%
-\begin{align}
-    \label{eq:y}
-    \mathbf{y}
-     & =
-    \left(
-    \mathbf{y}^0 \,\,\,
-    \mathbf{y}^1 \,\,\,
-    \cdots \,\,\,
-    \mathbf{y}^{l_\mathrm{max}}
-    \right)^\top
-    \nonumber \\
-     & =
-    \left(
-    y^0_0 \,\,\,
-    \,\,\,
-    y^1_{-1} \,\,\,
-    y^1_{0} \,\,\,
-    y^1_{1} \,\,\,
-    \,\,\,
-    \cdots \,\,\,
-    \,\,\,
-    y^{l_\mathrm{max}}_{-l_\mathrm{max}} \,\,\,
-    \cdots \,\,\,
-    y^{l_\mathrm{max}}_{l_\mathrm{max}} \,\,\,
-    \right)^\top
-    \quad.
-\end{align}
+on the surface of a star as an expansion in the real spherical harmonics.
 %
@@
 rotation period $P$, the stellar limb darkening coefficients, etc.
-For reference, the rows of $\pmb{\mathcal{A}}$ are given by
-Equation~(18) in \citet{Luger2019}.
-
-An in-depth treatment of the spherical harmonics is beyond the scope of this
-paper. For more information, including closed-form expressions for
-the spherical harmonics
-and details on their usage in modeling stellar surfaces, the reader is
-encouraged to consult \citet{Luger2019}.
-
-\section{A Gaussian Process for Star Spots}
+(see Appendix~\ref{sec:starry} for details).
+
+\subsection{Computing the GP}
 \label{sec:gp}
@@
 Let
-$\mathbf{f} = \left( f_0 \, f_1 \, \cdots \,  f_K \right)^\top$
+$\mathbf{f} = \left( f_0 \, f_1 \, \cdots \,  f_{K-1} \right)^\top$
 denote a vector of $K$ flux measurements at times
-$\mathbf{t} = \left( t_0 \,  t_1 \,  \cdots \, t_K \right)^\top$.
+$\mathbf{t} = \left( t_0 \,  t_1 \,  \cdots \, t_{K-1} \right)^\top$.
 Conditioned on certain physical properties of the
@@
 As with any random variable, the mean and covariance may be computed from
-the expectation value of $\mathbf{f}$ and
+the expectation values of $\mathbf{f}$ and
 $\mathbf{f}\,\mathbf{f}^\top$, respectively:
@@
 are the mean and covariance of the GP in the spherical harmonics basis.
-The bulk of the math in this paper is devoted to computing
+The bulk of the math in this paper (Appendix~\ref{sec:integrals})
+is devoted to computing
 the expectations in the expressions above, which
@@
 where $\mathbf{x}$ is a random vector-valued variable corresponding to a particular
-distribution of features on the surface  (i.e., the size and location of star spots)
+distribution of features on the surface
 and $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$ is its probability density
 function (PDF).
-
-In the sections that follow we will show that for suitable choices of $\mathbf{y}(\mathbf{x})$
+%
+In the Appendix we show that for suitable choices of $\pmb{\theta}_\bullet$,
+$\mathbf{y}(\mathbf{x})$,
 and $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$, the integrals in the expressions
@@
 %
+In particular, we let the spot hyperparameters of our GP be
+%
+\begin{align}
+    \pmb{\theta}_\bullet
+     & =
+    \left(
+    N
+    \,\,\,
+    c
+    \,\,\,
+    \alpha
+    \,\,\,
+    \beta
+    \,\,\,
+    r
+    \,\,\,
+    \Delta
+    \right)^\top
+    \quad,
+\end{align}
+%
+where $N$ is the number of star spots, $c$ is their contrast,
+$\alpha$ and $\beta$ describe the latitudinal
+distribution of the spots, and $r$ and $\Delta$ are the mean and
+half-width, respectively, of the distribution of spot radii.
+%
+The PDF for the radius is a uniform distribution
+between $r - \Delta$ and $r + \Delta$ (Appendix \ref{sec:size}),
+while
+the PDF for the latitude $\phi$ of the spots is a Beta distribution in
+$\cos\phi$ with shape parameters $\alpha$ and $\beta$ (Appendix \ref{sec:lat-lon}).
+The spot longitude is assumed to be uniformly distributed. Finally,
+the PDFs for the spot contrast and the number of spots are delta functions
+centered at $c$ and $N$, respectively.
+
+\xxx{More soon.}
+
+\subsection{Marginalizing over inclination}
+%
+The equations for the mean and covariance of the flux GP
+(Equations~\ref{eq:mean_f} and \ref{eq:cov_f}, respectively) are conditioned
+on specific values of the stellar properties $\pmb{\theta}_\star$ and the
+spot properties $\pmb{\theta}_\bullet$. To obtain the posterior distribution
+for these parameters, we must typically resort to numerical sampling techniques,
+which often scale steeply with the number of parameters. It is therefore generally
+desirable to keep the total number of parameters small, especially when
+employing the GP in an ensemble setting.
+In such a setting, we might have light curves from $N_\star$ stars, all of which
+we believe to have similar spot properties (perhaps because they have
+similar spectral types and rotation periods, for example).
+The total number of parameters in our problem is therefore
+%
+\begin{align}
+    N_\theta = N_\star \, \mathrm{dim}(\pmb{\theta}_\star) + 5
+    \quad,
+\end{align}
+%
+since each of the stars will have their own set of stellar properties
+$\pmb{\theta}_\star$ but will all share the same 5 spot properties
+$\pmb{\theta}_\bullet$ (by assumption).
+In a typical problem, the stellar parameters might be
+%
+\begin{align}
+    \pmb{\theta}_\star =
+    \left(
+    i \,\,\,
+    P \,\,\,
+    u_1 \,\,\,
+    u_2
+    \right)^\top
+    \quad,
+\end{align}
+%
+i.e., the inclination, the rotational period, and some (usually two) limb
+darkening coefficients. Thus, for a reasonably sized ensemble of $N_\star=100$
+stars, we would have to sample over $N_\theta = 405$ parameters.
+While large, this number is certainly not absurd, especially by modern standards.
+However, it does pose
+a problem when considering how complex the posterior distribution for the
+spot mapping problem can be. In addition to strong nonlinear degeneracies
+between some of the parameters (such as the contrast $c$ and the
+number of spots $N$), the posterior is often multimodal, especially in the
+stellar inclinations. While modern sampling schemes such as Hamiltonian
+Monte Carlo and Nested Sampling may in principle be able to deal with these
+issues, in practice it can be very difficult to obtain convergence in a
+reasonable amount of time.
+
+One workaround is to fix the values of $\pmb{\theta}_\star$. This could be done,
+for instance, to the rotational period $P$, which can often be estimated with
+fairly good accuracy from a periodogram. The limb darkening coefficients could
+be fixed at theoretical values, or perhaps their values could be shared among
+all stars (and sampled over), given the similarity assumption above.
+
+The inclination, however, is a different matter. Absent prior information
+for a particular star (such as a measurement of its projected rotational
+velocity $v\sin i$ or the knowledge that it hosts a transiting planet), it
+is simply not possible to reliably estimate the inclination in a
+pre-processing step. Any light curve statistic one might argue should scale
+with inclination---such as the amplitude of the variability---is invariably
+degenerate with the spot parameters $\pmb{\theta}_\bullet$. If one knew
+$\pmb{\theta}_\bullet$, then perhaps a decent point estimate of $i$ could
+be obtained, but in that case the analysis wouldn't be needed in the first place!
+%
+
+Fortunately, there is a better way to reduce the number of parameters in
+the problem: we can explicitly marginalize over the stellar inclination.
+That is, we can write the first two moments of our flux GP as
+%
+\begin{align}
+    \label{eq:mu_marg}
+    \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+     & =
+    \int
+    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
+    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
+    p(i) \, \mathrm{d}i \\
+    \label{eq:cov_marg}
+    \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+     & =
+    \int
+    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
+    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
+    \pmb{\mathcal{A}}^\top(i, \bar{\pmb{\theta}}_\star)
+    p(i) \, \mathrm{d}i    \quad,
+\end{align}
+%
+where the expectations inside the integrals are given by
+Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), respectively, and
+$\bar{\pmb{\theta}}_\star$ are all 0524the stellar parameters \emph{except}
+the inclination, i.e.,
+%
+\begin{align}
+    \bar{\pmb{\theta}}_\star = \left(
+    P \,\,\,
+    u_1 \,\,\,
+    u_2
+    \right)^\top
+    \quad,
+\end{align}
+%
+where perhaps $P$ is fixed for each star and the limb darkening coefficients
+are shared among all stars, as described above.
+%
+Thus, if we are able to perform the integrals in Equations~(\ref{eq:mu_marg})
+and (\ref{eq:cov_marg}), we can \emph{dramatically} reduce the number of
+parameters in our ensemble problem.
+%
+As we show in Appendix~\ref{sec:inclination}, if we assume that stellar
+inclinations are distributed isotropically, these integrals
+do in fact have closed-form solutions,
+since they are (again) integrals over Wigner matrices.
+%
+
+\section{Stuff}
+
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/nullspace.pdf}
+        \oscaption{nullspace}{%
+            The null space.
+            \label{fig:nullspace}
+        }
+    \end{centering}
+\end{figure}
+
+%
+%
+%
+%
+
+\appendix
+
+%
+%
+%
+%
+
+\section{Notation}
+\label{sec:notation}
+%
+Unless otherwise noted, we adopt
+the following conventions throughout this paper:
+integers are represented by italic uppercase letters (i.e., $N$),
+scalars are represented by italic lowercase
+letters (i.e., $x$), column vectors are
+represented by boldface lowercase letters
+($\mathbf{x}$), and matrices and other linear operators are represented
+by boldface capital letters ($\mathbf{X}$). In general, the elements of a vector
+$\mathbf{x}$ are denoted $x_i$ and the elements of a matrix $\mathbf{X}$
+are denoted $X_{i,j}$. The exception to this is if the vector or matrix
+represents a quantity in the spherical harmonic basis, in which case we use
+\emph{two} indices to represent a vector element, $x^l_m$, and \emph{four}
+indices to represent a matrix element, $X^{l,l'}_{m,m'}$.
+%
+The upper indices corresponds to the spherical harmonic degree,
+$l \in [0, l_{\mathrm{max}}]$ and $l' \in [0, l_{\mathrm{max}}]$,
+while the lower indices correspond to the
+spherical harmonic order, $m \in [-l, l]$ and $m' \in [-l', l']$.
+%
+Vector elements are arranged in order of increasing $l$ and,
+within each $l$, in order of increasing $m$.
+For example, a vector $\mathbf{x}$
+representing a quantity in the spherical harmonic basis up to degree
+$l_\mathrm{max}$ has components given by
+%
+\begin{align}
+    \mathbf{x}
+     & =
+    \left(
+    x^0_0 \,\,\,
+    \,\,\,\,\,\,
+    x^1_{-1} \,\,\,
+    x^1_{0} \,\,\,
+    x^1_{1} \,\,\,
+    \,\,\,\,\,\,
+    \cdots \,\,\,
+    \,\,\,\,\,\,
+    x^{l_\mathrm{max}}_{-l_\mathrm{max}}
+    \cdots \,\,
+    x^{l_\mathrm{max}}_{l_\mathrm{max}} \,\,\,
+    \right)^\top
+    \quad,
+\end{align}
+%
+while a matrix $\mathbf{X}$ in the same basis has components given by
+%
+\begin{align}
+    \setstackgap{L}{1.25\baselineskip}
+    \fixTABwidth{T}
+    \mathbf{X} =
+    \parenMatrixstack{
+    X^{0,0}_{0,0}  &  & X^{0,1}_{0,-1}  & X^{0,1}_{0,0}  & X^{0,1}_{0,1}  &        \\
+                   &  &                 &                &                &        \\
+    X^{1,0}_{-1,0} &  & X^{1,1}_{-1,-1} & X^{1,1}_{-1,0} & X^{1,1}_{-1,1} &        \\
+    X^{1,0}_{0,0}  &  & X^{1,1}_{0,-1}  & X^{1,1}_{0,0}  & X^{1,1}_{0,1}  &        \\
+    X^{1,0}_{1,0}  &  & X^{1,1}_{1,-1}  & X^{1,1}_{1,0}  & X^{1,1}_{1,1}  &        \\
+                   &  &                 &                &                & \ddots
+    }\quad.
+\end{align}
+%
+For completeness, the element of a spherical harmonic vector $\mathbf{x}$ with
+degree $l$ and order $m$ is at (flattened) index
+%
+\begin{align}
+    \label{eq:n}
+    n = l^2 + l + m
+    \quad.
+\end{align}
+%
+Conversely, the element at (flattened) index $n$ has degree and order
+%
+\begin{align}
+    \label{eq:lm}
+    \begin{split}
+        l & = \floor{\sqrt{n}}
+        \\
+        m & = n - l^2 - l
+        \quad,
+    \end{split}
+\end{align}
+%
+respectively.
+%
+Note, finally, that our use of upper and lower indices is purely a
+notational convenience,
+and should not be confused with
+exponentiation or a distinction between covariant and contravariant
+tensors. It should also not be confused with the notation used for the complex
+spherical harmonics, which also uses upper and lower indexing.
+%
+
+\section{Computing the flux}
+\label{sec:starry}
+\xxx{Coming soon.}
+
+\section{The Expectation Integrals}
+\label{sec:integrals}
+
+Our goal in this section is to find closed-form solutions to the
+first and second moments of the spherical harmonic representation of the
+stellar surface $\mathbf{y}$,
+%
+\begin{align}
+    \label{eq:exp_y_app}
+    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
+     & =
+    \int \mathbf{y}(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
+    \\
+    \label{eq:exp_yy_app}
+    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
+     & =
+    \int \mathbf{y}(\mathbf{x} ) \mathbf{y}^\top(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
+    \quad,
+\end{align}
+%
+which are linearly related to the mean and the covariance of our GP in flux
+(\S\ref{sec:gp}).
+%
+Recall that $\mathbf{x}$ is a vector of parameters describing the particular
+configuration of features on the surface of a star, and
+$p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$ is its probability
+density function conditioned on hyperparameters $\pmb{\theta}_\bullet$,
+which describe the \emph{distribution} of the features on the surface
+of one or many stars.
+%
 As we are specifically interested in modeling the effect of star spots
@@
     \left(
-    c_0 \,\, \cdots \,\, c_{N-1} \,\,
-    ,
-    \lambda_0 \,\, \cdots \,\, \lambda_{N-1} \,\,
-    ,
-    \phi_0 \,\, \cdots \,\, \phi_{N-1} \,\,
-    ,
-    r_0 \,\, \cdots \,\, r_{N-1} \,\,
+    \eta \,\,\,\,
+    \xi_0 \,\, \cdots \,\, \xi_{\eta-1} \,\,\,\,
+    \lambda_0 \,\, \cdots \,\, \lambda_{\eta-1} \,\,\,\,
+    \phi_0 \,\, \cdots \,\, \phi_{\eta-1} \,\,\,\,
+    \rho_0 \,\, \cdots \,\, \rho_{\eta-1}
     \right)^\top
@@
     \mathbf{y}(\mathbf{x}) =
-    \sum_{n=0}^{N-1}
-    c_n
+    \sum_{n=0}^{\eta-1}
+    \xi_n
     \,
@@
     \,
-    \mathbf{s}(r_n)
-    \quad,
-\end{align}
-%
-where $c_n$ is the contrast of the $n^\mathrm{th}$ spot,
+    \mathbf{s}(\rho_n)
+    \quad,
+\end{align}
+%
+where $\eta$ is the total number of spots,
+$\xi_n$ is the contrast of the $n^\mathrm{th}$ spot,
 $\lambda_n$ is its longitude, $\phi_n$ is its latitude,
-$r_n$ is its radius, and $N$ is the total number of spots.
+and $\rho_n$ is its radius.
 The vector function $\mathbf{s}(r_n)$
@@
 %
-For simplicity, we assume that $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$
-is separable in each of the four spot properties, and that all of the spots
+
+For simplicity, we assume that
+the total number of spots is fixed to a value $N$, i.e.,
+%
+\begin{align}
+    p(\eta \, \big| \, \pmb{\theta}_\bullet)
+     & =
+    \delta(\eta - N)
+    \quad,
+\end{align}
+%
+where $\delta$ is the delta function.%
+\footnote{%
+    When modeling a single star using the GP, this assumption is justified by
+    definition. It is less justified when the GP is used to model an ensemble of
+    stars, where each star may have a different total number of spots $\eta$.
+    However, as we argue in the text, $\eta$ is extremely difficult to constrain
+    from light curves, in particular because of how degenerate it is with
+    the spot contrast. In practice, we find that assuming that all stars in the
+    ensemble have the same number of spots $N$ leads to higher variance in the
+    estimate of $N$, but it does not lead to noticeable bias in $N$ or in any of
+    the other hyperparameters.
+}
+We further asume that
+$p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$
+is separable in each of the four other spot properties, and that all of the spots
 are drawn from the same distribution:
@@
     \prod_{n=0}^{N-1}
-    p(c_n \, \big| \, \pmb{\theta}_{c}) \,
+    p(\xi_n \, \big| \, \pmb{\theta}_{\xi}) \,
     p(\lambda_n \, \big| \, \pmb{\theta}_{\lambda}) \,
     p(\phi_n \, \big| \, \pmb{\theta}_{\phi})\,
-    p(r_n \, \big| \, \pmb{\theta}_{r})
+    p(\rho_n \, \big| \, \pmb{\theta}_{\rho})
     \quad,
@@
     \pmb{\theta}_\bullet = \left(
-    \pmb{\theta}_{c} \, \,
+    N \, \,
+    \pmb{\theta}_{\xi} \, \,
     \pmb{\theta}_{\lambda} \, \,
     \pmb{\theta}_{\phi} \, \,
-    \pmb{\theta}_{r} \right)^\top
-\end{align}
-%
-is the vector of yet-to-be-specified hyperparameters describing the process.
-%
-This allows us to rewrite the expectation integrals (\ref{eq:exp_y})
-and (\ref{eq:exp_yy}) as
+    \pmb{\theta}_{\rho} \right)^\top
+\end{align}
+%
+is the vector of hyperparameters describing the process and
+$\pmb{\theta}_{\xi}$,
+$\pmb{\theta}_{\lambda}$,
+$\pmb{\theta}_{\phi}$, and
+$\pmb{\theta}_{\rho}$ are yet to be specified.
+%
+This allows us to rewrite the expectation integrals (\ref{eq:exp_y_app})
+and (\ref{eq:exp_yy_app}) as
 %
@@
      & =
-    N \mathbf{e}_c
+    N \mathbf{e}_\xi
     \\[1em]
@@
      & =
-    N \mathbf{E}_c
+    N \mathbf{E}_\xi
 \end{align}
@@
     \label{eq:e1}
-    \mathbf{e}_r
-     & \equiv
-    \int
-    \mathbf{s}(r) \,
-    p(r \, \big| \, \pmb{\theta}_{r}) \,
-    \mathrm{d}r
+    \mathbf{e}_\rho
+     & \equiv
+    \int
+    \mathbf{s}(\rho) \,
+    p(\rho \, \big| \, \pmb{\theta}_{\rho}) \,
+    \mathrm{d}\rho
     %
@@
     \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{e}_r \,
+    \mathbf{e}_\rho \,
     p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
@@
     \label{eq:e4}
-    \mathbf{e}_c
-     & \equiv
-    \int
-    c \,
+    \mathbf{e}_\xi
+     & \equiv
+    \int
+    \xi \,
     \mathbf{e}_\lambda \,
-    p(c \, \big| \, \pmb{\theta}_{c}) \,
-    \mathrm{d}c
+    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
+    \mathrm{d}\xi
     %
@@
     \label{eq:E1}
-    \mathbf{E}_r
-     & \equiv
-    \int
-    \mathbf{s}(r) \, \mathbf{s}^\top(r) \,
-    p(r \, \big| \, \pmb{\theta}_{r}) \,
-    \mathrm{d}r
+    \mathbf{E}_\rho
+     & \equiv
+    \int
+    \mathbf{s}(\rho) \, \mathbf{s}^\top(\rho) \,
+    p(\rho \, \big| \, \pmb{\theta}_{\rho}) \,
+    \mathrm{d}\rho
     %
@@
     \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{E}_r \,
+    \mathbf{E}_\rho \,
     \mathbf{R}_{\hat{\mathbf{x}}}^\top(\phi) \,
@@
     \label{eq:E4}
-    \mathbf{E}_c
-     & \equiv
-    \int
-    c^2 \,
+    \mathbf{E}_\xi
+     & \equiv
+    \int
+    \xi^2 \,
     \mathbf{E}_\lambda \,
-    p(c \, \big| \, \pmb{\theta}_c)
-    \mathrm{d}c
+    p(\xi \, \big| \, \pmb{\theta}_\xi)
+    \mathrm{d}\xi
     %
@@
 
-We devote the next section to the computation of these eight
+We devote the remainder of this section to the computation of these eight
 integrals.
 
-\section{The Expectation Integrals}
-\label{sec:integrals}
-\subsection{The Size Integral}
-\label{sec:size}+
-%
-Our goal in this section is to compute the first and second moments
-of the size distribution ($\mathbf{e}_\phi$, $\mathbf{E}_\phi$) under
-a suitable spherical harmonic expansion $\mathbf{s}(r)$ of the spot profile
+\subsection{The Radius Integral}
+\label{sec:size}
+%
+Below we compute the first and second moments
+of the radius distribution ($\mathbf{e}_\rho$, $\mathbf{E}_\rho$) under
+a suitable spherical harmonic expansion $\mathbf{s}(\rho)$ of the spot profile
 and a suitable probability distribution function for the spot radius,
-$p(r \, \big| \, \pmb{\theta}_{r})$.
+$p(\rho \, \big| \, \pmb{\theta}_{\rho})$.
 
@@
 We model the brightness $b$ an angle $\vartheta$ away from the
-center of a spot of negative unit intensity and radius $r$ be
+center of a spot of negative unit intensity and radius $\rho$ as
 %
@@
     \label{eq:brvartheta}
-    b(r; \vartheta) & = \frac{1}{1 + \exp\left(\dfrac{r-\vartheta}{s}\right)} - 1
+    b(\rho; \vartheta) & = \frac{1}{1 + \exp\left(\dfrac{\rho-\vartheta}{s}\right)} - 1
 \end{align}
@@
 for some (constant) shape parameter $s$. In the limit $s \rightarrow 0$, $b$ approaches an
-inverted top-hat function with half-width equal to $r$,
+inverted top-hat function with half-width equal to $\rho$,
 corresponding to a circular spot of uniform intensity. For $s > 0$, each half
-of $b$ is a sigmoid with half-width at half-minimum equal to $r$.
+of $b$ is a sigmoid with half-width at half-minimum equal to $\rho$.
 In our implementation of the algorithm we choose $s = 0.2^\circ$,
@@
 \begin{align}
-    \mathbf{B} \, \tilde{\mathbf{s}}(r) = \mathbf{b}(r)
-\end{align}
-%
-%
-where $\mathbf{b}(r)$ is computed by evaluating Equation~(\ref{eq:brvartheta})
+    \mathbf{B} \, \tilde{\mathbf{s}}(\rho) = \mathbf{b}(\rho)
+\end{align}
+%
+%
+where $\mathbf{b}(\rho)$ is computed by evaluating Equation~(\ref{eq:brvartheta})
 at each of the $\vartheta_k$,
@@
 %
-and $\tilde{\mathbf{s}}(r)$ are the coefficients of the expansion.
+and $\tilde{\mathbf{s}}(\rho)$ are the coefficients of the expansion.
 These are related to the full vector of spherical harmonic coefficients describing
-the spot, $\mathbf{s}(r)$, by
-%
-\begin{align}
-    s^l_m(r) = \tilde{s}_l(r) \delta_{m,0}
+the spot, $\mathbf{s}(\rho)$, by
+%
+\begin{align}
+    s^l_m(\rho) = \tilde{s}_l(\rho) \delta_{m,0}
     \quad,
@@
 where $\delta$ is the Kronecker delta function.
-To find the coefficients $\tilde{\mathbf{s}}(r)$, we solve the (linear) inverse problem,
-%
-\begin{align}
-    \tilde{\mathbf{s}}(r) & = \mathbf{B}^+ \, \mathbf{b}(r)
+To find the coefficients $\tilde{\mathbf{s}}(\rho)$, we solve the (linear) inverse problem,
+%
+\begin{align}
+    \tilde{\mathbf{s}}(\rho) & = \mathbf{B}^+ \, \mathbf{b}(\rho)
 \end{align}
@@
             Intensity profiles for spots with different
-            radii $r$ computed at spherical harmonic
+            radii $\rho$ computed at spherical harmonic
             degree $l_{\mathrm{max}} = \LMAX$.
-            For $r \gtrsim 10^\circ$, the spherical harmonic expansion
+            For $\rho \gtrsim 10^\circ$, the spherical harmonic expansion
             captures the spot shape and intensity reasonably well,
@@
 different radii expanded to spherical harmonic degree $l_{\mathrm{max}} = \LMAX$.
-% computed from
-% %
-% \begin{align}
-%     \mathbf{b'}(r) & = \mathbf{B}
-%     \mathbf{B}^+ \, \mathbf{b}(r) \quad.
-% \end{align}
-%
 The average intensity
@@
 without the smoothing term (i.e., $\sigma = \infty$). However, for
-$r \lesssim 10^\circ$, an expansion to $l_{\mathrm{max}} = \LMAX$ is
+$\rho \lesssim 10^\circ$, an expansion to $l_{\mathrm{max}} = \LMAX$ is
 insufficient to correctly model the spot, as can be seen from the
-$r = 5^\circ$ profile (dashed curve). Expansions to higher spherical harmonic
+$\rho = 5^\circ$ profile (dashed curve). Expansions to higher spherical harmonic
 degree allow one to model spots with radii smaller than $10^\circ$, although
@@
 For simplicity, we will adopt a uniform probability distribution for the
-spot radius, characterized by a mean radius $r_0$ and a half-width $\Delta$:
-%
-\begin{align}
-    p(r \, \big| \, \pmb{\theta}_{r})
+spot radius, characterized by a mean radius $r$ and a half-width $\Delta$:
+%
+\begin{align}
+    p(\rho \, \big| \, \pmb{\theta}_{\rho})
      & =
     \begin{cases}
-        \frac{1}{2\Delta} & r_0 - \Delta \leq r \leq r_0 + \Delta
+        \frac{1}{2\Delta} & r - \Delta \leq r \leq r + \Delta
         \\
@@
 \begin{align}
-    \pmb{\theta}_r = \left(
-    r_0 \, \, \, \,
+    \pmb{\theta}_\rho = \left(
+    r \, \, \, \,
     \Delta \right)^\top
@@
 \begin{align}
-    p(r \, \big| \, \pmb{\theta}_{r}, \Delta = 0)
-     & =
-    \delta(r - r_0)
+    p(\rho \, \big| \, \pmb{\theta}_{\rho}, \Delta = 0)
+     & =
+    \delta(\rho - r)
     \quad,
@@
 \begin{align}
-    \mathbf{e}_r
-     & \equiv
-    \int
-    \mathbf{s}(r) \,
-    p(r \, \big| \, \pmb{\theta}_{r}) \,
-    \mathrm{d}r
+    \mathbf{e}_\rho
+     & \equiv
+    \int
+    \mathbf{s}(\rho) \,
+    p(\rho \, \big| \, \pmb{\theta}_{\rho}) \,
+    \mathrm{d}\rho
     \nonumber \\
@@
     \int_{r_0 - \Delta}^{r_0 + \Delta}
-    \mathbf{s}(r) \,
-    \mathrm{d}r
+    \mathbf{s}(\rho) \,
+    \mathrm{d}\rho
     \quad.
@@
 \begin{align}
-    (e_r)^l_m
+    (e_\rho)^l_m
      & =
     \frac{\delta_{m,0}}{2\Delta}
-    \int_{r_0 - \Delta}^{r_0 + \Delta}
-    \tilde{s}_{l}(r) \,
-    \mathrm{d}r
+    \int_{r - \Delta}^{r+ \Delta}
+    \tilde{s}_{l}(\rho) \,
+    \mathrm{d}\rho
     \nonumber \\
@@
     \sum_{k=0}^{K-1} B^+_{l,k}
-    \int_{r_0 - \Delta}^{r_0 + \Delta}
-    b_{k}(r)
-    \mathrm{d}r
+    \int_{r - \Delta}^{r + \Delta}
+    b_{k}(\rho)
+    \mathrm{d}\rho
     \nonumber \\
@@
     \lim_{\Delta \rightarrow 0}
-    \mathbf{e}_r
+    \mathbf{e}_\rho
      & =
@@
 \begin{align}
-    \mathbf{E}_r
-     & \equiv
-    \int
-    \mathbf{s}(r) \,
-    \mathbf{s}^\top(r) \,
-    p(r \, \big| \, \pmb{\theta}_{r}) \,
-    \mathrm{d}r
+    \mathbf{E}_\rho
+     & \equiv
+    \int
+    \mathbf{s}(\rho) \,
+    \mathbf{s}^\top(\rho) \,
+    p(\rho \, \big| \, \pmb{\theta}_{\rho}) \,
+    \mathrm{d}\rho
     \nonumber \\
@@
     \frac{1}{2\Delta}
-    \int_{r_0 - \Delta}^{r_0 + \Delta}
-    \mathbf{s}(r) \,
-    \mathbf{s}^\top(r) \,
-    \mathrm{d}r
+    \int_{r - \Delta}^{r + \Delta}
+    \mathbf{s}(\rho) \,
+    \mathbf{s}^\top(\rho) \,
+    \mathrm{d}\rho
     \quad.
@@
 \begin{align}
-    (E_r)^{l,l'}_{m,m'}
+    (E_\rho)^{l,l'}_{m,m'}
      & =
     \frac{\delta_{m,0}\delta_{m',0}}{2\Delta}
-    \int_{r_0 - \Delta}^{r_0 + \Delta}
-    \tilde{s}^l_{m}(r) \,
-    \tilde{s}^{l'}_{m'}(r) \,
-    \mathrm{d}r
+    \int_{r - \Delta}^{r + \Delta}
+    \tilde{s}^l_{m}(\rho) \,
+    \tilde{s}^{l'}_{m'}(\rho) \,
+    \mathrm{d}\rho
     \nonumber \\
@@
     \sum_{k'=0}^{K-1} B^+_{l',k'}
-    \int_{r_0 - \Delta}^{r_0 + \Delta}
-    b_{k}(r)
-    b_{k'}(r)
-    \mathrm{d}r
+    \int_{r - \Delta}^{r + \Delta}
+    b_{k}(\rho)
+    b_{k'}(\rho)
+    \mathrm{d}\rho
     \nonumber \\
@@
     \sum_{k'=0}^{K-1} B^+_{l',k'}
-    C_{k,k'}(r)
+    C_{k,k'}(r,\Delta)
     \quad,
@@
 \begin{align}
-    C_{k,k'}(r)
+    C_{k,k'}(r,\Delta)
      & \equiv
@@
     \lim_{\Delta \rightarrow 0}
-    \mathbf{E}_r
+    \mathbf{E}_\rho
      & =
@@
 The final integrals we must take in our computation of
-$\pmb{\mu}_{\mathbf{y}}(\pmb{\theta})$
-amd $\pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta})$ are the integrals over
-the spot contrast distribution, $\mathbf{e}_c$ and $\mathbf{E}_c$.
+$\mathrm{E} \big[ \mathbf{y} \, \big| \, \pmb{\theta}_\bullet \big]$
+and $\mathrm{E} \big[ \mathbf{y} \mathbf{y}^\top \, \big| \, \pmb{\theta}_\bullet \big]$
+are the integrals over
+the spot contrast distribution, $\mathbf{e}_\xi$ and $\mathbf{E}_\xi$.
 These are by far the easiest, since the spot contrast is a scalar
@@
 \begin{align}
-    \mathbf{e}_c
+    \mathbf{e}_\xi
      & \equiv
@@
     \int
-    c \,
-    p(c \, \big| \, \pmb{\theta}_{c}) \,
-    \mathrm{d}c
-    \\
-    \mathbf{E}_c
+    \xi \,
+    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
+    \mathrm{d}\xi
+    \\
+    \mathbf{E}_\xi
      & \equiv
@@
     \int
-    c^2 \,
-    p(c \, \big| \, \pmb{\theta}_c)
-    \mathrm{d}c
+    \xi^2 \,
+    p(\xi \, \big| \, \pmb{\theta}_\xi)
+    \mathrm{d}\xi
     \quad.
@@
 These integrals may be computed analytically for any choice of
-probability density function $p(c \, \big| \, \pmb{\theta}_c)$
+probability density function $p(\xi \, \big| \, \pmb{\theta}_\xi)$
 with closed-form moments.
@@
 spot contrast from light curves, let alone higher moments of its
-distribution; this is due largely to the fact that the contrast $c$
-is extremely degenerate with the total number of spots $N$
+distribution; this is due largely to the fact that the contrast
+is extremely degenerate with the total number of spots
 (see \S\ref{sec:cN-degeneracy}).
 %
-In our imeplementation of the algorithm,
+In our implementation of the algorithm,
 we therefore choose the simplest possible probability distribution,
@@
 \begin{align}
-    p(c \, \big| \, \pmb{\theta}_{c}) = \delta(c - c_0)
+    p(\xi \, \big| \, \pmb{\theta}_{\xi}) = \delta(\xi - c)
 \end{align}
@@
 \begin{align}
-    \pmb{\theta}_c = \left( c_0 \right)^\top
+    \pmb{\theta}_\xi = \left( \, c \, \right)^\top
     \quad.
@@
 \begin{align}
-    \mathbf{e}_c & = c_0 \, \mathbf{e}_\lambda
-    \\
-    \mathbf{E}_c & = c_0^2 \mathbf{E}_\lambda
-    \quad.
-\end{align}
-%
-
-\section{Marginalizing over inclination}
-%
-The equations for the mean and covariance of the flux GP
-(Equations~\ref{eq:mean_f} and \ref{eq:cov_f}, respectively) are conditioned
-on specific values of the stellar properties $\pmb{\theta}_\star$ and the
-spot properties $\pmb{\theta}_\bullet$. To obtain the posterior distribution
-for these parameters, we must typically resort to numerical sampling techniques,
-which often scale steeply with the number of parameters. It is therefore generally
-desirable to keep the total number of parameters small, especially when
-employing the GP in an ensemble setting.
-In such a setting, we might have light curves from $N_\star$ stars, all of which
-we believe to have similar spot properties (perhaps because they have
-similar spectral types and rotation periods, for example).
-The total number of parameters in our problem is therefore
-%
-\begin{align}
-    N_\theta = N_\star \, \mathrm{dim}(\pmb{\theta}_\star) + 5
-    \quad,
-\end{align}
-%
-since each of the stars will have their own set of stellar properties
-$\pmb{\theta}_\star$ but will all share the same 5 spot properties
-$\pmb{\theta}_\bullet$ (by assumption).
-In a typical problem, the stellar parameters might be
-%
-\begin{align}
-    \pmb{\theta}_\star =
-    \left(
-    i \,\,\,
-    P \,\,\,
-    u_1 \,\,\,
-    u_2
-    \right)^\top
-    \quad,
-\end{align}
-%
-i.e., the inclination, the rotational period, and some (usually two) limb
-darkening coefficients. Thus, for a reasonably sized ensemble of $N_\star=100$
-stars, we would have to sample over $N_\theta = 405$ parameters.
-While large, this number is certainly not absurd, especially by modern standards.
-However, it does pose
-a problem when considering how complex the posterior distribution for the
-spot mapping problem can be. In addition to strong nonlinear degeneracies
-between some of the parameters (such as the contrast $c$ and the
-number of spots $N$), the posterior is often multimodal, especially in the
-stellar inclinations. While modern sampling schemes such as Hamiltonian
-Monte Carlo and Nested Sampling may in principle be able to deal with these
-issues, in practice it can be very difficult to obtain convergence in a
-reasonable amount of time.
-
-One workaround is to fix the values of $\pmb{\theta}_\star$. This could be done,
-for instance, to the rotational period $P$, which can often be estimated with
-fairly good accuracy from a periodogram. The limb darkening coefficients could
-be fixed at theoretical values, or perhaps their values could be shared among
-all stars (and sampled over), given the similarity assumption above.
-
-The inclination, however, is a different matter. Absent prior information
-for a particular star (such as a measurement of its projected rotational
-velocity $v\sin i$ or the knowledge that it hosts a transiting planet), it
-is simply not possible to reliably estimate the inclination in a
-pre-processing step. Any light curve statistic one might argue should scale
-with inclination---such as the amplitude of the variability---is invariably
-degenerate with the spot parameters $\pmb{\theta}_\bullet$. If one knew
-$\pmb{\theta}_\bullet$, then perhaps a decent point estimate of $i$ could
-be obtained, but in that case the analysis wouldn't be needed in the first place!
-%
-
-Fortunately, there is a better way to reduce the number of parameters in
-the problem: we can explicitly marginalize over the stellar inclination.
-That is, we can write the first two moments of our flux GP as
-%
-\begin{align}
-    \label{eq:mu_marg}
-    \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
-     & =
-    \int
-    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
-    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
-    p(i) \, \mathrm{d}i \\
-    \label{eq:cov_marg}
-    \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
-     & =
-    \int
-    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
-    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
-    \pmb{\mathcal{A}}^\top(i, \bar{\pmb{\theta}}_\star)
-    p(i) \, \mathrm{d}i    \quad,
-\end{align}
-%
-where the expectations inside the integrals are given by
-Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), respectively, and
-$\bar{\pmb{\theta}}_\star$ are all the stellar parameters \emph{except}
-the inclination, i.e.,
-%
-\begin{align}
-    \bar{\pmb{\theta}}_\star = \left(
-    P \,\,\,
-    u_1 \,\,\,
-    u_2
-    \right)^\top
-    \quad,
-\end{align}
-%
-where perhaps $P$ is fixed for each star and the limb darkening coefficients
-are shared among all stars, as described above.
-%
-Thus, if we are able to perform the integrals in Equations~(\ref{eq:mu_marg})
-and (\ref{eq:cov_marg}), we can \emph{dramatically} reduce the number of
-parameters in our ensemble problem.
-
-As we will see below, these integrals do in fact have closed-form solutions,
-since they are (again) integrals over Wigner matrices. In our derivation,
-we will assume that stellar rotation axes are distributed isotropically. In
-that case, the probability density for the inclination is simply
-%
+    \mathbf{e}_\xi & = c \, \mathbf{e}_\lambda
+    \\
+    \mathbf{E}_\xi & = c^2 \mathbf{E}_\lambda
+    \quad.
+\end{align}
+%
+
+\section{Inclination}
+\label{sec:inclination}
+
 \begin{align}

### 2020-11-25 tex/figures/spot_profile.py

@@ -60,7 +60,7 @@ for k in range(len(r)):
         ls=ls[k],
         alpha=alpha[k],
         color=color[k],
-        label=r"$r = {}^\circ$".format(r[k]),
+        label=r"$\rho = {}^\circ$".format(r[k]),
     )
 
 ax.legend(loc="lower right", fontsize=12)

### 2020-11-30 tex/gitlinks.tex
baseline stuff

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/09b20fbf11c7967ce0510c149223cc7406ca9eca/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/09b20fbf11c7967ce0510c149223cc7406ca9eca/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/09b20fbf11c7967ce0510c149223cc7406ca9eca/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/b0a01ed1228fab145ceab240603ae662ab96e2d5/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/b0a01ed1228fab145ceab240603ae662ab96e2d5/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/b0a01ed1228fab145ceab240603ae662ab96e2d5/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-30 tex/listofitems.tex
baseline stuff

% !TeX encoding = ISO-8859-1
% Ce fichier contient le code de l'extension "listofitems"
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                    %
\def\loiname                 {listofitems}                           %
\def\loiver                      {1.61}                              %
%                                                                    %
\def\loidate                  {2019/03/03}                           %
%                                                                    %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                    %
% Author     : Christian Tellechea, Steven B. Segletes               %
% Status     : Maintained                                            %
% Maintainer : Christian Tellechea                                   %
% Email      : unbonpetit@netc.fr                                    %
%              steven.b.segletes.civ@mail.mil                        %
% Package URL: https://www.ctan.org/pkg/listofitems                  %
% Bug tracker: https://framagit.org/unbonpetit/listofitems/issues    %
% Repository : https://framagit.org/unbonpetit/listofitems/tree/master
% Copyright  : Christian Tellechea 2016-2019                         %
% Licence    : Released under the LaTeX Project Public License v1.3c %
%              or later, see http://www.latex-project.org/lppl.txt   %
% Files      : 1) listofitems.tex                                    %
%              2) listofitems.sty                                    %
%              3) listofitems-fr.tex                                 %
%              4) listofitems-fr.pdf                                 %
%              5) listofitems-en.tex                                 %
%              6) listofitems-en.pdf                                 %
%              7) README                                             %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ifdefined\ProvidesPackage\else
	\immediate\write -1 {%
	Package: \loidate\space v\loiver\space Grab items in lists using user-specified sep char (CT)}%
\fi

\expandafter\edef\csname loi_restorecatcode\endcsname{\catcode\number`\_=\number\catcode`\_\relax}
\catcode`\_11

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%% gestion des erreurs %%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ifdefined\PackageError
	\def\loi_error#1{\PackageError\loiname{#1}{Read the manual}}% pour LaTeX
\else
	\def\loi_error#1{\errmessage{Package \loiname\space Error: #1^^J}}% pour TeX
\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%% vérification de la présence de etex %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begingroup
\edef\__tempa{\meaning\eTeXversion}\edef\__tempb{\string\eTeXversion}%
\ifx\__tempa\__tempb
	\endgroup
\else
	\endgroup
	\loi_error{You are not using an eTeX engine, listofitems cannot work.}%
	\loi_restorecatcode\expandafter\endinput
\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%% macros auxiliaires %%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chardef\loi_stop=0
\def\loi_quark{\loi_quark}
\long\def\loi_identity#1{#1}
\long\def\loi_gobarg#1{}
\long\def\loi_first#1#2{#1}
\long\def\loi_second#1#2{#2}
\long\def\loi_firsttonil#1#2\_nil{#1}
\long\def\loi_antefi#1#2\fi{#2\fi#1}
\long\def\loi_exparg#1#2{\expandafter\loi_exparg_a\expandafter{#2}{#1}}% \loi_exparg{<a>}{<b>} devient <a>{<*b>}
\long\def\loi_exparg_a#1#2{#2{#1}}
\long\def\loi_expafter#1#2{\expandafter\loi_expafter_a\expandafter{#2}{#1}}% \loi_expafter{<a>}{<b>} devient <a><*b>
\long\def\loi_expafter_a#1#2{#2#1}
\def\loi_macroname{\loi_ifinrange\escapechar[[0:255]]{\expandafter\loi_gobarg}{}\string}
\def\loi_argcsname#1#{\loi_argcsname_a{#1}}
\def\loi_argcsname_a#1#2{\loi_expafter{#1}{\csname#2\endcsname}}
\long\def\loi_addtomacro#1#2{\loi_exparg{\def#1}{#1#2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%% macros de test %%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\long\def\loi_ifnum#1{\ifnum#1\expandafter\loi_first\else\expandafter\loi_second\fi}
\long\def\loi_ifx#1{\ifx#1\expandafter\loi_first\else\expandafter\loi_second\fi}
\long\def\loi_ifempty#1{\loi_exparg\loi_ifx{\expandafter\relax\detokenize{#1}\relax}}
\def\loi_ifstar#1#2{\def\loi_ifstar_a{\loi_ifx{*\loi_nxttok}{\loi_first{#1}}{#2}}\futurelet\loi_nxttok\loi_ifstar_a}
\long\def\loi_ifstuffexpandable#1{\def\loi_tempa{#1}\loi_exparg{\def\loi_tempb}{#1}\expandafter\unless\loi_ifx{\loi_tempa\loi_tempb}}
\long\def\loi_ifcsexpandable#1{% #1 est-il constitué d'une sc _développable_ ?
	\loi_ifempty{#1}
		{\loi_second
		}
		{\loi_ifspacefirst{#1}
			{\loi_second% si espace en 1er, faux
			}
			{%
			\loi_exparg\loi_ifempty{\loi_gobarg#1}% 1 seul token ?
				{\begingroup\escapechar`\_\def\_{#1}\expandafter\endgroup
					\csname loi_\if\expandafter\expandafter\expandafter\loi_firsttonil\expandafter\string\_\_nil\string _first\else second\fi\endcsname
						{\loi_ifstuffexpandable{#1}}
						{\loi_second}%
				}
				{\loi_second% si plusieurs tokens, faux
				}%
			}%
		}%
}
\def\loi_ifinrange#1[[#2:#3]]{\loi_ifnum{\numexpr(#1-#2)*(#1-#3)>0 }\loi_second\loi_first}
\def\loi_ifstring#1\in#2{% si la chaine #1 est contenue dans #2
	\def\loi_ifstring_a##1#1##2\_nil{\loi_ifempty{##2}\loi_second\loi_first}%
	\loi_ifstring_a#2#1\@nil% appel de la macro auxiliaire
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%% macros \loi_foreach %%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcount\loi_cnt_foreach_nest \loi_cnt_foreach_nest=0
\def\end_foreach{\end_foreach}
\def\loi_def_foreachsep#1{%
	\long\def\loi_foreach##1\in##2##3{%
		\global\advance\loi_cnt_foreach_nest1
		\loi_argcsname\def{loop_code_\number\loi_cnt_foreach_nest}{##3}%
		\loi_foreach_a##1##2#1\end_foreach#1%
		\loi_argcsname\let{loop_code_\number\loi_cnt_foreach_nest}\empty
		\global\advance\loi_cnt_foreach_nest-1
	}%
	\long\def\loi_foreach_a##1##2#1{%
		\def##1{##2}%
		\loi_ifx{\end_foreach##1}
			{}
			{\csname loop_code_\number\loi_cnt_foreach_nest\endcsname% exécute le code
			\loi_foreach_a##1%
			}%
	}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%% macros gérant l'appariement %%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\long\def\defpair#1{%
	\let\loi_listofpair\empty
	\loi_ifempty{#1}
		{}
		{\defpair_a{}#1\loi_quark\loi_quark}%
}
\long\def\defpair_a#1#2#3{%
	\loi_ifx{\loi_quark#2}
		{\def\loi_sanitizelist##1,\_nil{\def\loi_listofpair{##1}}%
		\loi_sanitizelist#1\_nil
		}
		{\loi_if_validpair#2#3%
			{\long\def\loi_paired_a{#2}\long\def\loi_paired_b{#3}%
			\loi_ifx{\loi_paired_a\loi_paired_b}
				{\loi_error{Paired tokens must not be equal, the pair \detokenize{#2#3} is ignored}%
				\defpair_a{#1}%
				}
				{\defpair_a{#1#2#3,}%
				}%
			}
			{\loi_error{Invalid paired tokens, the pair "\detokenize{#2}" and "\detokenize{#3}" is ignored}%
			\defpair_a{#1}%
			}%
		}%
}
\long\def\loi_if_validpair#1#2{%
	\def\loi_validpair{1}%
	\loi_if_invalid_pairtoken{#1}{\def\loi_validpair{0}}%
	\loi_if_invalid_pairtoken{#2}{\def\loi_validpair{0}}%
	\loi_ifnum{\loi_validpair=1 }
}
\long\def\loi_if_invalid_pairtoken#1{%
	\loi_ifempty{#1}
		{\loi_identity
		}
		{\loi_ifspacefirst{#1}
			{\loi_identity
			}
			{\loi_exparg\loi_ifempty{\loi_gobarg#1}% 1 seul token ?
				{\ifcat\relax\noexpand#1\expandafter\loi_identity\else\expandafter\loi_gobarg\fi}
				{\loi_identity}% si plusieurs tokens, faux
			}%
		}%
}
\long\def\loi_count_occur#1\in#2:#3{% compte le nombre d'occurrences de #1 dans #2 et met le résultat dans la macro #3
	\long\def\loi_count_occur_a##1##2#1##3\_nil{%
		\loi_ifempty{##3}
			{\def#3{##1}}
			{\expandafter\loi_count_occur_a\number\numexpr##1+1\relax##3\_nil}%
	}%
	\loi_count_occur_a0#2#1\_nil
}
\long\def\loi_check_pair#1#2\in#3{% teste l'appariement de #1 et #2 dans #3
	\loi_ifempty{#3}
		{\loi_second
		}
		{\loi_count_occur#1\in#3:\loi_tempa
		\loi_count_occur#2\in#3:\loi_tempb
		\loi_ifnum{\loi_tempa=\loi_tempb\relax}%
		}%
}
\long\def\loi_grabpaired_expr#1#2#3#4#5{% #1=liste de paires  #2=expression  #3=séparateur   #4=résultat    #5=ce qui reste
	\let#4\empty
	\def\loi_remain{#2#3}%
	\loi_foreach\loi_pair\in{#1}{\expandafter\loi_grabpaired_expr_a\loi_pair{#3}#4}%
	\def\loi_remove_lastsep##1#3\_nil{\def#4{##1}}%
	\expandafter\loi_remove_lastsep#4\_nil
	\expandafter\long\expandafter\def\expandafter\loi_grab_remain#4##1\_nil{\loi_ifempty{##1}{\let#5\empty}{\loi_exparg{\def#5}{\loi_gobarg##1}}}%
	\loi_grab_remain#2\_nil
}
\long\def\loi_grabpaired_expr_a#1#2#3#4{% #1#2=paire en cours  #3=séparateur   #4=résultat
	\loi_exparg{\loi_check_pair#1#2\in}#4% si les paires sont appariées dans le résultat
		{}% passer à la paire suivante
		{\long\def\loi_grabpaired_expr_b##1#3##2\_nil{%
			\loi_addtomacro#4{##1#3}% ajouter au résultat ce qui est jusqu'au prochain séparateur
			\def\loi_remain{##2}%
			\loi_exparg{\loi_check_pair#1#2\in}{#4}
				{}
				{\loi_ifempty{##2}
					{\loi_error{"\detokenize{#1}" and "\detokenize{#2}" are not paired}}
					{\loi_grabpaired_expr_b##2\_nil}%
				}%
			}%
		\expandafter\loi_grabpaired_expr_b\loi_remain\_nil
		}%
}
\def\insidepair#1#2#3#4{% #1#2=paire  #3=expr  #4=macro reçevant le resultat
	\loi_if_validpair#1#2%
		{\loi_ifcsexpandable{#3}
			{\loi_exparg{\insidepair#1#2}{#3}#4%
			}
			{\loi_check_pair#1#2\in{#3}% si les paires sont appariées dans le résultat
				{\def\insidepair_a##1#1##2\_nil{\insidepair_b##2\_nil{#1}}%
				\def\insidepair_b##1#2##2\_nil##3{%
					\loi_check_pair#1#2\in{##3##1#2}
						{\loi_exparg{\def#4}{\loi_gobarg##3##1}}%
						{\insidepair_b##2\_nil{##3##1#2}}%
					}%
				\insidepair_a#3\_nil
				}
				{\loi_error{"\detokenize{#1}" and "\detokenize{#2}" are not paired in "#3"}%
				}%
			}%
		}
		{\loi_error{Invalid paired tokens "\detokenize{#1}" and "\detokenize{#2}", empty \string#4 returned}% et bim
		\let#4\empty% voilà, bien fait pour vos gueules
		}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%% macro \loi_fornum %%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\loi_fornum#1=#2to#3\do{%
	\edef#1{\number\numexpr#2}\edef\loi_sgncmp{\ifnum#1<\numexpr#3\relax>+\else<-\fi}%
	\expandafter\loi_fornum_a\csname loi_fornum_\string#1\expandafter\endcsname\expandafter{\number\numexpr#3\expandafter}\loi_sgncmp#1%
}
\long\def\loi_fornum_a#1#2#3#4#5#6{\def#1{\unless\ifnum#5#3#2\relax\loi_antefi{#6\edef#5{\number\numexpr#5#41\relax}#1}\fi}#1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%% macro retirant les espaces extrêmes %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\long\def\loi_ifspacefirst#1{\expandafter\loi_ifspacefirst_a\detokenize{#10} \_nil}
\long\def\loi_ifspacefirst_a#1 #2\_nil{\loi_ifempty{#1}}
\expandafter\def\expandafter\loi_gobspace\space{}
\def\loi_removefirstspaces{\romannumeral\loi_removefirstspaces_a}
\long\def\loi_removefirstspaces_a#1{\loi_ifspacefirst{#1}{\expandafter\loi_removefirstspaces_a\expandafter{\loi_gobspace#1}}{\loi_stop#1}}
\edef\loi_restorezerocatcode{\catcode0=\number\catcode0 \relax}
\catcode0 12
\long\def\loi_removelastspaces#1{\romannumeral\loi_removelastspaces_a#1^^00 ^^00\_nil}
\long\def\loi_removelastspaces_a#1 ^^00{\loi_removelastspaces_b#1^^00}
\long\def\loi_removelastspaces_b#1^^00#2\_nil{\loi_ifspacefirst{#2}{\loi_removelastspaces_a#1^^00 ^^00\_nil}{\loi_stop#1}}
\loi_restorezerocatcode
\long\def\loi_removeextremespaces#1{% #1=texte où les espaces extrêmes sont retirés
	\romannumeral\expandafter\expandafter\expandafter\loi_removelastspaces\expandafter\expandafter\expandafter
	{\expandafter\expandafter\expandafter\loi_stop\loi_removefirstspaces{#1}}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%% macro publique \setsepchar %%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\setsepchar{\futurelet\loi_nxttok\setsepchar_a}
\def\setsepchar_a{\loi_ifx{[\loi_nxttok}\setsepchar_b{\setsepchar_b[/]}}
\long\def\setsepchar_b[#1]#2{% #1=sepcar de <liste des sepcar>  #2=<liste des sepcar>
	\loi_ifempty{#1}
		{\loi_error{Empty separator not allowed, separator "/" used}%
		\setsepchar_b[/]{#2}%
		}
		{\def\loi_currentsep{#1}%
		\_removeextremespacesfalse
		\loi_nestcnt1 % réinitaliser niveau initial à 1
		\def\nestdepth{1}%
		\loi_argcsname\let{loi_previndex[\number\loi_nestcnt]}\empty
		\def\loi_listname{loi_listofsep}%
		\let\loi_def\def \let\loi_edef\edef \let\loi_let\let
		\let\loi_listofpair_saved\loi_list_ofpair
		\let\loi_list_ofpair\empty
		\loi_ifempty{#2}
			{\loi_error{Empty list of separators not allowed, "," used}%
			\readlist_e1{,}%
			}
			{\readlist_e1{#2}%
			}%
		\loi_argcsname\let\nestdepth{loi_listofseplen[0]}%
		\loi_argcsname\let\loi_currentsep{loi_listofsep[1]}% 1er car de séparation
		\let\loi_listofpair\loi_listofpair_saved
		}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%% macro normalisant l'index %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\loi_normalizeindex#1#2#3{% #1=macroname  #2=liste d'index  #3=profondeur max --> renvoie {err}{indx norm}
	\loi_ifempty{#2}
		{\loi_stop{}{}}
		{\loi_normalizeindex_a1{}{#3}{#1}#2,\loi_quark,}%
}%
\def\loi_normalizeindex_a#1#2#3#4#5,{% #1=compteur de profondeur #2=index précédents  #3=profondeur max #4=macroname  #5=index courant
	\loi_ifx{\loi_quark#5}
		{\loi_normalizeindex_c#2\loi_quark% supprimer la dernière virgule
		}
		{\loi_ifnum{#1>#3 }
			{\loi_invalidindex{Too deeply nested index, index [.] retained}{#2}% si profondeur trop grande
			}
			{\loi_ifinrange\ifnum\numexpr#5<0 -1*\fi(#5)[[1:\csname #4len[#20]\endcsname]]% si abs(#5) hors de [1,len]
				{\loi_exparg\loi_normalizeindex_b{\number\numexpr#5\ifnum\numexpr#5<0 +\csname #4len[#20]\endcsname+1\fi}{#1}{#2}{#3}{#4}}
				{\loi_invalidindex{#5 is an invalid index, index [.] retained}{#2}}%
			}%
		}%
}
\def\loi_normalizeindex_b#1#2#3{\loi_exparg\loi_normalizeindex_a{\number\numexpr#2+1}{#3#1,}}% #1=index à rajouter  #2=compteur de profondeur #3=index précédents
\def\loi_normalizeindex_c#1,\loi_quark{\loi_stop{}{#1}}
\def\loi_invalidindex#1#2{\loi_ifempty{#2}{\loi_invalidindex_a{#1},}\loi_invalidindex_a{#1}{#2}}
\def\loi_invalidindex_a#1#2{\loi_invalidindex_b#1\loi_quark#2\loi_quark}
\def\loi_invalidindex_b#1[.]#2\loi_quark#3,\loi_quark#4\loi_quark,{\loi_stop{#1[#3]#2}{#3}}% #4= index ignorés

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% macro publique \readlist %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcount\loi_nestcnt
\def\greadlist{\let\loi_def\gdef\let\loi_edef\xdef\def\loi_let{\global\let}\readlist_a}%
\def\readlist{\let\loi_def\def\let\loi_edef\edef\let\loi_let\let\readlist_a}
\def\readlist_a{%
	\loi_nestcnt1 % niveau initial = 1
	\loi_argcsname\let{loi_previndex[\number\loi_nestcnt]}\empty
	\loi_ifstar{\_removeextremespacestrue\readlist_b}{\_removeextremespacesfalse\readlist_b}%
}
\long\def\readlist_b#1#2{% #1=macro stockant les éléments  #2=liste des éléments
	\loi_ifcsexpandable{#2}
		{\loi_exparg{\readlist_b#1}{#2}%
		}
		{\loi_edef\loi_listname{\loi_macroname#1}%
		\loi_argcsname\loi_let{\loi_listname nest}\nestdepth
		\loi_argcsname\loi_def{\loi_listname[]}{#2}% la liste entière
		\loi_argcsname\loi_def{\loi_listname sep[]}{}% séparateur vide
		\loi_ifempty{#2}
			{\loi_def#1[##1]{}%
			\loi_argcsname\loi_def{\loi_listname len}{0}\loi_argcsname\loi_def{\loi_listname len[0]}{0}%
			\loi_error{Empty list ignored, nothing to do}%
			}
			{\loi_edef#1[##1]{\unexpanded{\romannumeral\expandafter\loi_checkindex\romannumeral\loi_normalizeindex}{\loi_listname}{##1}{\csname\loi_listname nest\endcsname}{\loi_listname}}%
			\loi_argcsname\loi_edef{\loi_listname sep}[##1]{\unexpanded{\romannumeral\expandafter\loi_checkindex\romannumeral\loi_normalizeindex}{\loi_listname}{##1}{\csname\loi_listname nest\endcsname}{\loi_listname sep}}%
			\readlist_c{#2}%
			\loi_argcsname\loi_argcsname\loi_let{\loi_listname len}{\loi_listname len[0]}% longueur du niveau 0
			}%
		}%
}
\def\loi_checkindex#1#2#3{%
	\expandafter\expandafter\expandafter\loi_stop\csname#3[#2]\expandafter\endcsname
	\romannumeral\loi_ifempty{#1}{\loi_stop}{\loi_stop\loi_error{#1}}%
}
\def\readlist_c{%
	\loi_argcsname\loi_let\loi_currentsep{loi_listofsep[\number\loi_nestcnt]}%
	\expandafter\readlist_d\loi_currentsep||\_nil
}
\long\def\readlist_d#1||#2\_nil#3{\readlist_e1{#3#1}}% #1=<sep courant simple>  #3=liste -> rajoute un élément vide pour le test \ifempty ci dessous
\long\def\readlist_e#1#2{% #1=compteur d'index  #2=liste d'éléments à examiner terminée par <sep courant simple> >>RIEN laissé après
	\loi_ifempty{#2}
		{\loi_argcsname\loi_edef{\loi_listname len[\csname loi_previndex[\number\loi_nestcnt]\endcsname0]}{\number\numexpr#1-1\relax}%
		\loi_argcsname\loi_let{\loi_listname sep[\csname loi_previndex[\number\loi_nestcnt]\endcsname\number\numexpr#1-1\relax]}\empty% le dernier <sep> est <vide> ##NEW v1.52
		\advance\loi_nestcnt-1
		\loi_argcsname\loi_let\loi_currentsep{loi_listofsep[\number\loi_nestcnt]}%
		}
		{\loi_expafter{\readlist_f{#2}{}}\loi_currentsep||\loi_quark||#2\_nil{#1}% aller isoler le 1er item
		}%
}
\long\def\readlist_f#1#2#3||{% #1=liste restante   #2=<dernier sep utilisé>  #3=<sep courant>
	\loi_ifx{\loi_quark#3}% on a épuisé tous les <séparateurs> ? RESTE à lire <expr+sep1>\_nil{<compteur>}
		{\loi_ifempty{#2}% si #2 vide, aucun <sep utilisé> n'a été trouvé, il reste à lire "<liste complète>\_nil"
			{\long\def\readlist_g##1\_nil##2{\loi_exparg{\readlist_h{##2}{}}{\loi_gobarg##1}{#2}}% ##2=compteur d'index
			}
			{\loi_ifx{\loi_listofpair\empty}% paires définies ?
				{\long\def\readlist_g##1#2##2\_nil##3{\loi_exparg{\readlist_h{##3}{##2}}{\loi_gobarg##1}{#2}}%
				}
				{\long\def\readlist_g##1\_nil##2{%
					\loi_exparg{\loi_exparg\loi_grabpaired_expr\loi_listofpair}{\loi_gobarg##1}{#2}\loi_grabpaired_result\loi_grabpaired_remain
					\loi_exparg{\loi_exparg{\readlist_h{##2}}{\loi_grabpaired_remain}}{\loi_grabpaired_result}{#2}}%
				}%
			}%
		\readlist_g\relax% le \relax meuble l'argument délimité
		}
		{\long\def\readlist_g##1#3##2\_nil{%
			\loi_ifempty{##2}% si <liste restante> ne contient pas le <sep courant>
				{\readlist_f{#1}{#2}% recommencer avec le même <sep utile>
				}%
				{\loi_ifx{\loi_listofpair\empty}% si pas de paires définies
					{\loi_exparg\readlist_f{\loi_gobarg##1#3}{#3}% raccourcir <liste restante> et <sep courant>:=<sep utile>% ##BUGFIX v1.53
					}%
					{\loi_exparg\loi_grabpaired_expr\loi_listofpair{#1}{#3}\loi_grabpaired_result\loi_grabpaired_remain
					\loi_exparg\readlist_f{\loi_grabpaired_result#3}{#3}%
					}%
				}%
			}%
		\readlist_g\relax#1#3\_nil% ##BUGFIX v1.53
		}%
}
\long\def\readlist_h#1#2#3{% #1=compteur d'index  #2=liste restante  #3=élément courant
	\loi_ifnum{0\loi_exparg\loi_ifspacefirst{\loi_currentsep}{}1\if_removeextremespaces1\fi=11 }% s'il faur retirer les espaces extrêmes
		{\loi_exparg{\loi_exparg{\readlist_i{#1}{#2}}}{\loi_removeextremespaces{#3}}}% redéfinir l'élément courant
		{\readlist_i{#1}{#2}{#3}}%
}
\long\def\readlist_i#1#2#3#4{% #1=compteur d'index  #2=liste restante  #3=élément courant   #4=sep utilisé
	\loi_ifnum{0\if_ignoreemptyitems1\fi\loi_ifempty{#3}1{}=11 }
		{\readlist_e{#1}{#2}% si l'on n'ignore pas les éléments vides
		}%
		{\loi_argcsname\loi_def{\loi_listname[\csname loi_previndex[\number\loi_nestcnt]\endcsname#1]}{#3}% assignation de l'item ctuel à la macro
		\loi_argcsname\loi_def{\loi_listname sep[\csname loi_previndex[\number\loi_nestcnt]\endcsname#1]}{#4}% assignation du <sep> actuel à la macro \<macrolist>sep
		\loi_ifnum{\loi_nestcnt<\nestdepth\relax}% si imbrication max non atteinte
			{\advance\loi_nestcnt1
			\loi_argcsname\edef{loi_previndex[\number\loi_nestcnt]}{\csname loi_previndex[\number\numexpr\loi_nestcnt-1]\endcsname#1,}%
			\readlist_c{#3}% recommencer avec l'élément courant
			}
			{}%
		\loi_exparg\readlist_e{\number\numexpr#1+1}{#2}% puis chercher l'élément suivant dans la liste restante
		}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%% macro \listlen %%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\listlen#1[#2]{%
	\romannumeral\loi_ifempty{#2}
		{\expandafter\expandafter\expandafter\loi_stop\csname\loi_macroname#1len[0]\endcsname}
		{\loi_exparg\listlen_a{\romannumeral-`\.\loi_macroname#1}{#2}}%
}
\def\listlen_a#1#2{% #1=macro name   #2=index non normalisé    prendre <profondeur max-1>
	\loi_exparg{\expandafter\listlen_b\romannumeral\loi_normalizeindex{#1}{#2}}{\number\numexpr\csname#1nest\endcsname-1}{#1}%
}
\def\listlen_b#1#2#3{% #1=err   #2=index normalisé  #3=macroname
	\expandafter\expandafter\expandafter\loi_stop\csname#3len[#2,0]\expandafter\endcsname
	\romannumeral\loi_ifempty{#1}{\loi_stop}{\loi_stop\loi_error{#1}}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%% macro \foreachitem %%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\foreachitem#1\in#2{%
	\edef\foreachitem_a{\noexpand\foreachitem_c\noexpand#1{\expandafter\noexpand\csname\loi_macroname#1cnt\endcsname}{\loi_macroname#2}}%
	\futurelet\loi_nxttok\foreachitem_b
}
\def\foreachitem_b{\loi_ifx{\loi_nxttok[}\foreachitem_a{\foreachitem_a[]}}
\def\foreachitem_c#1#2#3[#4]{% prendre <profondeur max-1>
	\loi_exparg{\expandafter\foreachitem_d\romannumeral\loi_normalizeindex{#3}{#4}}{\number\numexpr\csname#3nest\endcsname-1}#1{#2}{#3}%
}
\def\foreachitem_d#1#2{\loi_ifempty{#2}{\foreachitem_e{#1}{}}{\foreachitem_e{#1}{#2,}}}% #1=err  #2=index norm
\long\def\foreachitem_e#1#2#3#4#5#6{% #1=err  #2=index norm  #3=macroiter  #4=compteur associé  #5=nom de macrolist  #6=code
	\loi_ifnum{\csname#5len[#20]\endcsname>0 }
		{\loi_ifempty{#1}{}{\loi_error{#1}}%
		\loi_fornum#4=1to\csname#5len[#20]\endcsname\do{\loi_argcsname\let#3{#5[#2#4]}#6}%
		}
		{}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%% macro \showitem %%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\showitems{\loi_ifstar{\let\showitems_cmd\detokenize\showitems_a}{\let\showitems_cmd\loi_identity\showitems_a}}
\def\showitems_a#1{\def\showitems_b{\showitems_d#1}\futurelet\loi_nxttok\showitems_c}
\def\showitems_c{\loi_ifx{\loi_nxttok[}\showitems_b{\showitems_b[]}}
\def\showitems_d#1[#2]{\foreachitem\showitems_ater\in#1[#2]{\showitemsmacro{\expandafter\showitems_cmd\expandafter{\showitems_ater}}}}
\unless\ifdefined\fbox
	\newdimen\fboxrule \newdimen\fboxsep \fboxrule=.4pt \fboxsep=3pt % réglages identiques à LaTeX
	\def\fbox#1{% imitation de la macro \fbox de LaTeX, voir pages 271 à 274 de "Apprendre à programmer en TeX"
		\hbox{%
			\vrule width\fboxrule
			\vtop{%
				\vbox{\hrule height\fboxrule \kern\fboxsep \hbox{\kern\fboxsep#1\kern\fboxsep}}%
				\kern\fboxsep \hrule height\fboxrule
			}\vrule width\fboxrule
		}%
	}
\fi
\def\showitemsmacro#1{% encadrement par défaut
	\begingroup\fboxsep=0.25pt \fboxrule=0.5pt \fbox{\strut#1}\endgroup
	\hskip0.25em\relax
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%% macro \itemtomacro %%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\itemtomacro#1[#2]{% #1[#2]=item  non encore lu: #3=macro
	\edef\loi_listname{\loi_macroname#1}%
	\loi_exparg{\expandafter\itemtomacro_a\romannumeral\expandafter\loi_normalizeindex\expandafter{\loi_listname}{#2}}{\csname\loi_listname nest\endcsname}\let
}
\def\gitemtomacro#1[#2]{% #1[#2]=item
	\xdef\loi_listname{\loi_macroname#1}%
	\loi_exparg{\expandafter\itemtomacro_a\romannumeral\expandafter\loi_normalizeindex\expandafter{\loi_listname}{#2}}{\csname\loi_listname nest\endcsname}{\global\let}%
}
\def\itemtomacro_a#1#2#3#4{%
	\loi_ifempty{#1}{}{\loi_error{#1}}%
	\loi_argcsname#3#4{\loi_listname[#2]}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%% réglages par défaut %%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\if_removeextremespaces
\newif\if_ignoreemptyitems
\let\ignoreemptyitems\_ignoreemptyitemstrue
\let\reademptyitems\_ignoreemptyitemsfalse
\setsepchar{,}
\defpair{}
\loi_def_foreachsep{,}
\reademptyitems

\loi_restorecatcode
\endinput

######################################################################
############################  Historique #############################
######################################################################

v1.0  19/8/2016
  - Première version publique

v1.1  01/09/2016
  - Stockage des séparateurs dans <macrolist>sep
  - bug corrigé dans \loi_restorecatcode

v1.2  22/10/2016
  - macros \greadlist et \gitemtomacro pour la globalité

v1.3  18/11/2016
  - bugs corrigés dans la gestion de la globalité

v1.4  05/10/2017
  - test \loi_ifprimitive ajouté au test \loi_ifcs
  - suppression de \loi_expafternil, création de \loi_expafter,
    modification de \loi_argcsname
  - correction d'un bug : \setsepchar{\par} ne provoque plus
    d'erreur. \loi_ifnum devient \long

v1.5  06/10/2017
  - correction d'un bug dans \loi_ifcs

v1.51 24/10/2017
  - correction d'un bug dans \loi_ifcs

v1.52 13/01/2018
  - le dernier séparateur est <vide>

v1.53 13/03/2018
  - correction d'un bug dans \readlist_g

v1.6 01/11/2018
  - possibilité d'appariement de tokens dans les items

v1.61 03/03/2019
    - la macro \loi_ifcs contient une erreur de conception.
      Il faut tester si le token est un sc && s'il est
      développable pour renvoyer vrai car il existe des sc
      non développables && qui ne sont _pas_ des primitives.
      Macro rebaptisée \loi_ifcsexpandable

### 2020-11-30 tex/ms.tex
baseline stuff

@@
 
+\section{Motivation}
+\label{sec:motivation}
+See Figure~\ref{fig:nullspace_ensemble}.
+
+\begin{figure}[ht!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/nullspace_ensemble.pdf}
+        \oscaption{nullspace_ensemble}{%
+            Posterior shrinkage as a function of spherical harmonic degree
+            for a hypothetical scenario in which an observer can measure a
+            stellar light curve at a single inclination (thin blue curves),
+            at 3 different inclinations (thin orange curves), and at 10 different
+            inclinations (thin green curves). The thicker curves correspond to
+            the mean shrinkage for each experiment. The inclinations are
+            drawn randomly from an isotropic distribution.
+            The information content in the light curve of a star observed
+            from a single vantage point approaches zero as $\mathrm{l}$
+            increases. However, observing the star from many vantage points
+            allows one to recover nearly all of the information in the
+            even spherical harmonic modes.
+            While this thought experiment is not possible in practice for
+            a single star, it is extremely relevant to ensemble analyses of statistically
+            similar stars.
+            \label{fig:nullspace_ensemble}
+        }
+    \end{centering}
+\end{figure}
+
 \section{A Gaussian Process for Star Spots}
+\label{sec:main}
 
@@
     \label{eq:fAy}
-    \mathbf{f} = \pmb{\mathcal{A}} \, \mathbf{y}
+    \mathbf{f} = \mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y}
     \quad,
@@
 denote a vector of $K$ flux measurements at times
-$\mathbf{t} = \left( t_0 \,  t_1 \,  \cdots \, t_{K-1} \right)^\top$.
+$\mathbf{t} = \left( t_0 \,  t_1 \,  \cdots \, t_{K-1} \right)^\top$,
+defined in units such that a star with no spots on it will have
+$\mathbf{f} = \mathbf{1}$.
 Conditioned on certain physical properties of the
@@
     \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
-     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
+     & = \mathbf{1} + \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
     \\
@@
 
-\xxx{More soon.}
+\xxx{More soon. Include likelihood function.}
 
 \subsection{Marginalizing over inclination}
+\label{sec:inclination}
 %
@@
     \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+    - \mathbf{1} \mathbf{1}^\top
      & =
@@
     \pmb{\mathcal{A}}^\top(i, \bar{\pmb{\theta}}_\star)
-    p(i) \, \mathrm{d}i    \quad,
+    p(i) \, \mathrm{d}i
+    \quad,
 \end{align}
@@
 Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), respectively, and
-$\bar{\pmb{\theta}}_\star$ are all 0524the stellar parameters \emph{except}
+$\bar{\pmb{\theta}}_\star$ are all the stellar parameters \emph{except}
 the inclination, i.e.,
@@
 
-\section{Stuff}
-
-\begin{figure}[t!]
+\subsection{The baseline problem}
+\label{sec:baseline}
+
+There is one final subtle---but extremely important---point that remains to
+be discussed in our derivation of a Gaussian process for stellar
+rotation. Thus far we have cast the light curve problem in a typical GP inference
+setting: we model our observations $\mathbf{f}$ as
+a multidimensional random variable distributed as a multivariate Gaussian with mean vector
+$\pmb{\mu}$ and covariance matrix $\pmb{\Sigma}$. However, in the context of
+relative photometry, the mean $\pmb{\mu}$ is typically \emph{unobservable}.
+%
+To understand why, consider what
+$\pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$ actually represents:
+it is the expected value
+of the flux
+for a star with certain properties $\pmb{\theta}_\star$ and $\pmb{\theta}_\bullet$,
+\emph{in units of the flux one would measure if the star had no spots}.
+We do not observe in these units: rather, we observe in units of counts on
+the detector, which depend on the luminosity of the star, the distance to
+the star, and various properties of the telescope. Even if we knew all these
+things, we would still need to know the intrinsic photospheric brightness
+in order to properly normalize our dataset for direct application of
+the Gaussian process.
+
+Instead, what astronomers typically do is to self-normalize their data:
+i.e., divide the flux by the mean, median, maximum, or some similar
+statistic of the light curve. This operation folds the unknowability of
+the flux baseline under the rug and transforms the light curve into a \emph{relative}
+measurement of the star's temporal variability. While relative measurements
+are typically what we are interested in anyways, this normalization procedure
+has two significant drawbacks: (1) it introduces a new set of perfect degeneracies
+in the modeling and (2) it can substantially change the covariance structure
+of the dataset.
+
+The first point is illustrated in the toy example in
+Figure~\ref{fig:mean_normalization}. As it has been pointed out before \xxx{(?)},
+we will not spend too much time discussing it, other than to say that the
+flux normalization introduces a degeneracy between the total spot coverage
+of a star and the contrast of any individual feature on the surface.
+While this is a formally perfect degeneracy for a single light curve, we can
+again resort to ensemble statistics across many light curves to break it;
+we discuss this further in \S\ref{sec:breaking_baseline_degeneracy}.
+
+The second point is of more immediate concern, since we must correct our
+expression for the covariance matrix
+$\pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$ if we are to
+use our GP to model normalized light curves. For definiteness, let us
+consider only mean-normalized light curves.
+Since we are dividing the
+flux by the mean, one might imagine that we could simply divide the
+covariance matrix by the square of the mean and call it a day. However,
+this is incorrect, since we must be careful to distinguish between the
+\emph{sample} mean and the \emph{process} mean. When normalizing a light
+curve by the mean, the operation we perform is
+%
+\begin{align}
+    \tilde{\mathbf{f}} & = \frac{\mathbf{f}}{\left<f\right>}
+    \quad,
+\end{align}
+%
+where $\tilde{\mathbf{f}}$ is the normalized, unit-mean light curve,
+$\mathbf{f}$ is the measured light curve (in detector counts), and
+$\left<f\right>$ is the \emph{sample} mean: i.e., the average value of
+a given star's light curve (which we model as a sample from our GP).
+This may be close to but is in general different from the \emph{process} mean,
+$\pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$, since the mean of
+a draw from the GP is itself normally distributed with a variance that scales
+with the GP variance.
+
+Computing the covariance of the normalized process is tricky, especially
+because the normalized process is \emph{not} strictly Gaussian: the distribution
+has heavy tails due to the fact that $\tilde{\mathbf{f}}$ diverges as
+the sample mean approaches zero. In fact, because of these tails, the covariance
+of the normalized process is formally \emph{infinite}, since the probability of
+drawing a sample whose mean is arbitrarily close to zero is finite.
+
+If this is all starting to sound like a bad idea, that's because it is!
+A much safer approach is to resist the temptation to normalize the light curve
+and instead model the unknown baseline as a latent variable. However,
+this would require an extra parameter \emph{for every light curve}, so the
+computational savings we achieved by marginalizing out the inclination
+would be gone. Fortunately, in practice, the variance of a stellar light curve
+is usually small compared to its mean: stellar variability amplitudes are
+typically at the level of a few percent or lower. When this is the case,
+the probability of drawing a GP sample whose mean is close to zero is
+extremely small, and we can make use of the approximate expression derived
+in \citet{Luger2020} for the covariance of a normalized Gaussian process:
+%
+\begin{align}
+    \label{eq:SigmaTilde}
+    \tilde{\pmb{\Sigma}}
+     & \approx
+    \frac{A}{\mu^2} \pmb{\Sigma} +
+    z \Big(
+    (A + B) \, (\mathbf{1} - \mathbf{q}) \, (\mathbf{1} - \mathbf{q})^\top
+    - A \, \mathbf{q} \, \mathbf{q}^\top
+    \Big)
+    \quad,
+\end{align}
+%
+where
+%
+\begin{align}
+    z & \equiv \frac{\left< \Sigma \right>}{\mu^2}
+\end{align}
+%
+is the ratio of the average element in $\pmb{\Sigma}$
+to the square of the mean of the Gaussian process,
+$\mathbf{q}$ is the ratio of the average of each row in $\pmb{\Sigma}$
+to the average element in $\pmb{\Sigma}$, and $A$, $B$ are
+order unity and zero scalars, respectively,
+given by the optimally-truncated diverging series
+%
+\begin{align}
+    \label{eq:baseline_alpha}
+    A
+     & \equiv
+    \sum\limits_{n=0}^N
+    \frac{(2n + 1)!}{2^n \, n!}
+    z^n
+    \\[1em]
+    \label{eq:baseline_beta}
+    B
+     & \equiv
+    \sum\limits_{n=0}^N
+    \frac{2n(2n + 1)!}{2^n \, n!}
+    z^n
+    \quad,
+\end{align}
+%
+where $N$ is the largest value for which the series coefficient at $N$ is
+smaller than the coefficient at $N - 1$.
+
+\xxx{More soon.}
+
+\begin{figure}[ht!]
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/nullspace.pdf}
-        \oscaption{nullspace}{%
-            The null space.
-            \label{fig:nullspace}
+        \includegraphics[width=\linewidth]{figures/mean_normalization.pdf}
+        \oscaption{mean_normalization}{%
+            An example of the baseline problem.
+            \emph{Top:} Consider a star with a single
+            equatorial spot of contrast $c$ viewed at a certain inclination.
+            The total flux (in some units) as a function of time is shown as
+            the blue curve. Now, consider a second star,
+            identical in all respects to the first, except that (1) the equatorial spot
+            has half the contrast (i.e., $\nicefrac{c}{2}$); and (2) there is
+            a second, large spot centered on the pole. The corresponding light
+            curve is shown as the dashed orange curve.
+            The orange light curve is different from the blue one in two ways:
+            (1) since the equatorial spot has half the contrast, the amplitude of the associated
+            dips in the light curve is half that of the first star; and
+            (2) since the polar spot is azimuthally symmetric, its only
+            contribution is a net darkening at all phases.
+            \emph{Bottom:} The true baseline level of a stellar
+            light curve, which corresponds to the flux one would measure in the
+            absence of any spots, is almost always unknown. Photometric measurements
+            are therefore meaningful only in a relative sense, i.e., as deviations from
+            the mean, median, or maximum level of the light curve.
+            The bottom panel shows the same two light curves, this time plotted as
+            deviations in parts per thousand (ppt) from their respective maxima.
+            To the observer, the two light curves are \emph{indistinguishable}.
+            In the absence of baseline information, there
+            exists a perfect degeneracy between the total spot coverage
+            and the contrast of any individual feature on the surface.
+            \label{fig:mean_normalization}
         }
@@
 %
-Note, importantly, that we are not interested in any specific value of
+Regarding this equation,
+two things should be noted. First, we define $\mathbf{y}$ relative to
+a baseline of zero: i.e., a star with no spots on it will have
+$\mathbf{y} = \mathbf{0}$. Second, and more importantly,
+we are not interested in any specific value of
 $\mathbf{y}$; rather, we would like to know its expectation value under
@@
 
+\section{Latitude Distribution Transforms}
+
+Given the shape parameters $\alpha$ and $\beta$,
+the mode $\mu$ and standard deviation $\sigma$ of the
+distribution of spot latitudes may be written
+%
+\begin{proof}{}
+    \mu &= 2 \tan^{-1}
+    \left(
+    \sqrt{
+        2 \alpha + \beta - 2 -
+        \sqrt{
+            4 \alpha^2
+            - 8 \alpha
+            - 6 \beta
+            + 4 \alpha \beta
+            + \beta^2
+            + 5
+        }
+    }
+    \right)
+    \\
+    \sigma &= \frac{\sin\mu}{
+        \sqrt{1
+            - \alpha
+            + \beta
+            + \left(\beta - 1\right) \cos\mu
+            + \dfrac{\alpha - 1}{\cos\mu^2}}}
+    \quad.
+\end{proof}
+%
+The reverse transform is given by
+%
+\begin{proof}{}
+    \alpha &= \frac{2 + 4 \sigma^2 + (3 + 8 \sigma^2) \cos\mu + 2 \cos\left( 2\mu \right) + \cos\left( 3\mu \right)}{16 \sigma^2 \cos\left( \frac{\mu}{2} \right)^4} \\
+    \beta &= \frac{\cos\mu + 2 \sigma^2 (3 + \cos\left( 2\mu \right)) - \cos\left( 3\mu \right)}{16 \sigma^2 \cos\left( \frac{\mu}{2} \right)^4}
+    \quad.
+\end{proof}
+%
+These expressions may be used to compute the Jacobian of the transform:
+%
+\begin{proof}{}
+    J &=
+    \frac{\partial{\mu}}{\partial{\alpha}}
+    \frac{\partial{\sigma}}{\partial{\beta}} -
+    \frac{\partial{\mu}}{\partial{\beta}}
+    \frac{\partial{\sigma}}{\partial{\alpha}}
+    %
+    \\
+    %
+    &=\resizebox{.925\hsize}{!}{$
+            \frac{
+                \alpha \beta \left(1 + \cos\mu\right)^3
+                \sin\left(2\mu\right)^3
+            }{
+                \sigma
+                \left(
+                2 \alpha + \beta - 3 +
+                \left(2 \alpha + \beta - 1\right) \cos\mu
+                \right)
+                \left(
+                2 \left( \alpha + \beta - 1\right) +
+                3 (\beta - 1) \cos\mu
+                - 2 \left( \alpha - \beta - 1 \right) \cos\left( 2\mu \right)
+                + \left( \beta - 1\right) \cos\left( 3\mu \right)
+                \right)^2
+            }
+        $}
+    \nonumber
+\end{proof}
+
 \bibliography{bib}

### 2020-11-30 tex/figures/mean_normalization.py

@@ -1,0 +1,105 @@
+import starry
+import matplotlib.pyplot as plt
+from matplotlib.colors import Normalize
+import numpy as np
+
+
+class Star(object):
+    def __init__(self, nlon=300, ydeg=30, linear=True):
+        # Generate a uniform intensity grid
+        self.nlon = nlon
+        self.nlat = nlon // 2
+        self.lon = np.linspace(-180, 180, self.nlon)
+        self.lat = np.linspace(-90, 90, self.nlat)
+        self.lon, self.lat = np.meshgrid(self.lon, self.lat)
+        self.intensity = np.zeros_like(self.lat)
+        self.linear = linear
+
+        # Instantiate a starry map
+        self.map = starry.Map(ydeg, lazy=False)
+
+    def _angular_distance(self, lam1, lam2, phi1, phi2):
+        # https://en.wikipedia.org/wiki/Great-circle_distance
+        return (
+            np.arccos(
+                np.sin(phi1 * np.pi / 180) * np.sin(phi2 * np.pi / 180)
+                + np.cos(phi1 * np.pi / 180)
+                * np.cos(phi2 * np.pi / 180)
+                * np.cos((lam2 - lam1) * np.pi / 180)
+            )
+            * 180
+            / np.pi
+        )
+
+    def reset(self):
+        self.intensity = np.zeros_like(self.lat)
+
+    def add_spot(self, lon, lat, radius, contrast):
+        idx = self._angular_distance(lon, self.lon, lat, self.lat) <= radius
+        if self.linear:
+            self.intensity[idx] -= contrast
+        else:
+            self.intensity[idx] = -contrast
+
+    def flux(self, t, period=1.0, inc=60.0, smoothing=0.1):
+        # Expand in Ylms
+        self.map.load(self.intensity)
+        self.map.inc = inc
+
+        # Smooth to get rid of ringing
+        if smoothing > 0:
+            l = np.concatenate(
+                [np.repeat(l, 2 * l + 1) for l in range(self.map.ydeg + 1)]
[…]

### 2020-12-01 tex/gitlinks.tex
more text; fixed tests

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/b0a01ed1228fab145ceab240603ae662ab96e2d5/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/b0a01ed1228fab145ceab240603ae662ab96e2d5/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/b0a01ed1228fab145ceab240603ae662ab96e2d5/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/d4cf57ce3f2a3645c0d52306ca56e4386eb7de9b/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/d4cf57ce3f2a3645c0d52306ca56e4386eb7de9b/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/d4cf57ce3f2a3645c0d52306ca56e4386eb7de9b/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-01 tex/ms.tex
more text; fixed tests

@@
             The information content in the light curve of a star observed
-            from a single vantage point approaches zero as $\mathrm{l}$
+            from a single vantage point approaches zero as $l$
             increases. However, observing the star from many vantage points
@@
 we discuss this further in \S\ref{sec:breaking_baseline_degeneracy}.
-
-The second point is of more immediate concern, since we must correct our
-expression for the covariance matrix
-$\pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$ if we are to
-use our GP to model normalized light curves. For definiteness, let us
-consider only mean-normalized light curves.
-Since we are dividing the
-flux by the mean, one might imagine that we could simply divide the
-covariance matrix by the square of the mean and call it a day. However,
-this is incorrect, since we must be careful to distinguish between the
-\emph{sample} mean and the \emph{process} mean. When normalizing a light
-curve by the mean, the operation we perform is
-%
-\begin{align}
-    \tilde{\mathbf{f}} & = \frac{\mathbf{f}}{\left<f\right>}
-    \quad,
-\end{align}
-%
-where $\tilde{\mathbf{f}}$ is the normalized, unit-mean light curve,
-$\mathbf{f}$ is the measured light curve (in detector counts), and
-$\left<f\right>$ is the \emph{sample} mean: i.e., the average value of
-a given star's light curve (which we model as a sample from our GP).
-This may be close to but is in general different from the \emph{process} mean,
-$\pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$, since the mean of
-a draw from the GP is itself normally distributed with a variance that scales
-with the GP variance.
-
-Computing the covariance of the normalized process is tricky, especially
-because the normalized process is \emph{not} strictly Gaussian: the distribution
-has heavy tails due to the fact that $\tilde{\mathbf{f}}$ diverges as
-the sample mean approaches zero. In fact, because of these tails, the covariance
-of the normalized process is formally \emph{infinite}, since the probability of
-drawing a sample whose mean is arbitrarily close to zero is finite.
-
-If this is all starting to sound like a bad idea, that's because it is!
-A much safer approach is to resist the temptation to normalize the light curve
-and instead model the unknown baseline as a latent variable. However,
-this would require an extra parameter \emph{for every light curve}, so the
-computational savings we achieved by marginalizing out the inclination
-would be gone. Fortunately, in practice, the variance of a stellar light curve
-is usually small compared to its mean: stellar variability amplitudes are
-typically at the level of a few percent or lower. When this is the case,
-the probability of drawing a GP sample whose mean is close to zero is
-extremely small, and we can make use of the approximate expression derived
-in \citet{Luger2020} for the covariance of a normalized Gaussian process:
-%
-\begin{align}
-    \label{eq:SigmaTilde}
-    \tilde{\pmb{\Sigma}}
-     & \approx
-    \frac{A}{\mu^2} \pmb{\Sigma} +
-    z \Big(
-    (A + B) \, (\mathbf{1} - \mathbf{q}) \, (\mathbf{1} - \mathbf{q})^\top
-    - A \, \mathbf{q} \, \mathbf{q}^\top
-    \Big)
-    \quad,
-\end{align}
-%
-where
-%
-\begin{align}
-    z & \equiv \frac{\left< \Sigma \right>}{\mu^2}
-\end{align}
-%
-is the ratio of the average element in $\pmb{\Sigma}$
-to the square of the mean of the Gaussian process,
-$\mathbf{q}$ is the ratio of the average of each row in $\pmb{\Sigma}$
-to the average element in $\pmb{\Sigma}$, and $A$, $B$ are
-order unity and zero scalars, respectively,
-given by the optimally-truncated diverging series
-%
-\begin{align}
-    \label{eq:baseline_alpha}
-    A
-     & \equiv
-    \sum\limits_{n=0}^N
-    \frac{(2n + 1)!}{2^n \, n!}
-    z^n
-    \\[1em]
-    \label{eq:baseline_beta}
-    B
-     & \equiv
-    \sum\limits_{n=0}^N
-    \frac{2n(2n + 1)!}{2^n \, n!}
-    z^n
-    \quad,
-\end{align}
-%
-where $N$ is the largest value for which the series coefficient at $N$ is
-smaller than the coefficient at $N - 1$.
-
-\xxx{More soon.}
 
@@
 
+The second point is of more immediate concern, since we must correct our
+expression for the covariance matrix
+$\pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$ if we are to
+use our GP to model normalized light curves. For definiteness, let us
+consider only mean-normalized light curves.
+Since we are dividing the
+flux by the mean, one might imagine that we could simply divide the
+covariance matrix by the square of the mean and call it a day. However,
+this is incorrect, since we must be careful to distinguish between the
+\emph{sample} mean and the \emph{process} mean. When normalizing a light
+curve by the mean, the operation we perform is
+%
+\begin{align}
+    \tilde{\mathbf{f}} & = \frac{\mathbf{f}}{\left<f\right>}
+    \quad,
+\end{align}
+%
+where $\tilde{\mathbf{f}}$ is the normalized, unit-mean light curve,
+$\mathbf{f}$ is the measured light curve (in detector counts), and
+$\left<f\right>$ is the \emph{sample} mean: i.e., the average value of
+a given star's light curve (which we model as a sample from our GP).
+This may be close to but is in general different from the \emph{process} mean,
+$\pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$, since the mean of
+a draw from the GP is itself normally distributed with a variance that scales
+with the GP variance.
+
+Computing the covariance of the normalized process is tricky, especially
+because the normalized process is \emph{not} strictly Gaussian: the distribution
+has heavy tails due to the fact that $\tilde{\mathbf{f}}$ diverges as
+the sample mean approaches zero. In fact, because of these tails, the covariance
+of the normalized process is formally \emph{infinite}, since the probability of
+drawing a sample whose mean is arbitrarily close to zero is finite.
+
+If this is all starting to sound like a bad idea, that's because it is!
+A much safer approach is to resist the temptation to normalize the light curve
+and instead model the unknown baseline as a latent variable. However,
+this would require an extra parameter \emph{for every light curve}, so the
+computational savings we achieved by marginalizing out the inclination
+would be gone. Fortunately, in practice, the variance of a stellar light curve
+is usually small compared to its mean: stellar variability amplitudes are
+typically at the level of a few percent or lower. When this is the case,
+the probability of drawing a GP sample whose mean is close to zero is
+extremely small, and we can make use of the approximate expression derived
+in \citet{Luger2020} for the covariance of a normalized Gaussian process:
+%
+\begin{align}
+    \label{eq:SigmaTilde}
+    \tilde{\pmb{\Sigma}}
+     & \approx
+    \frac{A}{\mu^2} \pmb{\Sigma} +
+    z \Big(
+    (A + B) \, (\mathbf{1} - \mathbf{q}) \, (\mathbf{1} - \mathbf{q})^\top
+    - A \, \mathbf{q} \, \mathbf{q}^\top
+    \Big)
+    \quad,
+\end{align}
+%
+where
+%
+\begin{align}
+    z & \equiv \frac{\left< \Sigma \right>}{\mu^2}
+\end{align}
+%
+is the ratio of the average element in $\pmb{\Sigma}$
+to the square of the mean of the Gaussian process,
+$\mathbf{q}$ is the ratio of the average of each row in $\pmb{\Sigma}$
+to the average element in $\pmb{\Sigma}$, and $A$, $B$ are
+order unity and zero scalars, respectively,
+given by the optimally-truncated diverging series
+%
+\begin{align}
+    \label{eq:baseline_alpha}
+    A
+     & \equiv
+    \sum\limits_{n=0}^N
+    \frac{(2n + 1)!}{2^n \, n!}
+    z^n
+    \\[1em]
+    \label{eq:baseline_beta}
+    B
+     & \equiv
+    \sum\limits_{n=0}^N
+    \frac{2n(2n + 1)!}{2^n \, n!}
+    z^n
+    \quad,
+\end{align}
+%
+where $N$ is the largest value for which the series coefficient at $N$ is
+smaller than the coefficient at $N - 1$. In the expressions above, it is
+assumed that the mean $\pmb{\mu}$ is constant, i.e., $\pmb{\mu} = \mu\, \mathbf{1}$.
+Since our Gaussian process is azimuthally isotropic (i.e., no preferred
+longitude), that is the case throughout this paper.
+
+What Equation~(\ref{eq:SigmaTilde}) allows us to do is effectively marginalize over
+the unknown baseline by modeling the \emph{normalized} flux as a draw
+from a Gaussian process:
+%
+\begin{align}
+    \tilde{\mathbf{f}} \sim \mathcal{N}\left( \mathbf{1}, \tilde{\pmb{\Sigma}} \right)
+    \quad.
+\end{align}
+%
+This is appropriate as long as the quantity $z \ll 1$, for which the true
+distribution of $\tilde{\mathbf{f}}$ is approximately Gaussian. In practice,
+we recommend employing this trick only for $z \lesssim 0.02$, for which the
+error in the approximation to the covariance is less than $10^{-6}$.
+In cases where the light curve variability exceeds a few percent, we recommend
+modeling the baseline in each light curve as a latent variable, as discussed
+above.
+
 %

### 2020-12-01 tests/test_lnlike.py

@@ -118,7 +118,8 @@ def test_lnlike_grad(param, marginalize_over_inclination):
                 return
             verify_grad(
                 lambda x: StarryProcess(
-                    marginalize_over_inclination=marginalize_over_inclination
+                    marginalize_over_inclination=marginalize_over_inclination,
+                    normalized=False,
                 ).log_likelihood(t, flux, data_cov, **{param: x}),
                 (defaults[param],),
                 n_tests=1,
@@ -127,6 +128,7 @@ def test_lnlike_grad(param, marginalize_over_inclination):
             verify_grad(
                 lambda x: StarryProcess(
                     marginalize_over_inclination=marginalize_over_inclination,
+                    normalized=False,
                     **{param: x}
                 ).log_likelihood(t, flux, data_cov),
                 (defaults[param],),

### 2020-12-01 tex/gitlinks.tex
more text

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/d4cf57ce3f2a3645c0d52306ca56e4386eb7de9b/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/d4cf57ce3f2a3645c0d52306ca56e4386eb7de9b/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/d4cf57ce3f2a3645c0d52306ca56e4386eb7de9b/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/709731d2744831cbf6cc4e9bf5a8473f0c7d29e3/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/709731d2744831cbf6cc4e9bf5a8473f0c7d29e3/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/709731d2744831cbf6cc4e9bf5a8473f0c7d29e3/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-01 tex/ms.tex
more text

@@
 
-\subsection{Spherical Harmonics}
-\label{sec:ylm}
-%
-The spherical harmonics are a complete set of orthogonal functions on the
-surface of the sphere and, as such, are a convenient basis in which to construct
-our gaussian process. As in \citet{Luger2019}, here we describe the intensity
-on the surface of a star as an expansion in the real spherical harmonics.
-%
-\citet{Luger2019} derived the expression for the total disk-integrated
+In this section, we present an overview of the computation of our Gaussian
+process, which boils down to computing the mean and covariance of the stellar
+flux conditioned on certain physical properties of the star and its star spot
+distribution. Most of the math is folded into the Appendix for readability;
+readers may want to refer to Appendix~\ref{sec:notation} in particular for
+a discussion of the notation and conventions we adopt.
+
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/ylms.pdf}
+        \oscaption{ylms}{%
+            The real spherical harmonics in the polar frame up to
+            $l = 5$. Rows correspond to the degree $l$ and columns to
+            the order $m$. The set of all spherical harmonics forms a
+            complete, orthogonal basis on the sphere.
+            \label{fig:ylms}
+        }
+    \end{centering}
+\end{figure}
+
+Before we dive in, it is useful to introduce the spherical harmonics,
+a set of orthogonal functions on the surface of the sphere which we will use
+to describe the intensity field on the surface of a star. As we will see below,
+the spherical harmonics are a particularly convenient basis in which to
+describe star spot distributions%
+\footnote{There are, of course, drawbacks to using this basis:
+    in particular, the spherical harmonics are smooth, continuous functions that
+    struggle (at finite degree $l$) to capture high resolution features such as
+    small star spots. We discuss this point at length in \S\ref{sec:tinyspots},
+    where we show that our model is useful even when applied to stars with spots
+    smaller than the effective resolution of the GP.}, as they will allow us to compute
+moments of the intensity distribution analytically. And of more immediate concern,
+\citet{Luger2019} showed that there is a linear relationship between the
+spherical harmonic expansion of a stellar surface and the total disk-integrated
 flux $\mathbf{f}$ (i.e., the light curve)
-one would observe from a star rotating about a fixed axis
-if the stellar surface intensity is well described by the spherical harmonic
-coefficient vector $\mathbf{y}$:
+one would observe as the star rotates about a fixed axis.
+If the stellar surface intensity is described by the spherical harmonic
+coefficient vector $\mathbf{y}$, the flux is given by
 %
@@
 %
-where $\pmb{\mathcal{A}}$ is the \starry
+where $\mathbf{1}$ is the ones vector and
+$\pmb{\mathcal{A}}$ is the \starry
 design matrix, a purely linear operator that transforms from the spherical
@@
 %
-
-% The vector $\pmb{\theta}$ includes physical properties of the star such
-% as its inclination $i$ and rotational period $P$ as well as parameters
-% describing the shape of the probability density function (PDF) governing
-% the distribution of features on the surface.
 
@@
 covariance $\pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$
-of $\mathbf{f}$, which together fully specify our GP in flux.
+of $\mathbf{f}$. Together, these specify a multidimensional Gaussian
+distribution, which we assume fully describes%
+\footnote{The \emph{true} distribution of stellar light curves conditioned
+    on $\pmb{\theta}_\star$ and $\pmb{\theta}_\bullet$ is not Gaussian, so our
+    assumption is formally wrong. But, as the saying goes, \emph{all models are
+        wrong; some are useful}. As we will show later, this turns out to be
+    an extremely useful assumption.}
+how our flux measurements
+are distributed:
+%
+\begin{align}
+    \mathbf{f}\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right) \sim
+    \mathcal{N}\Big(
+    \pmb{\mu}\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right),
+    \,
+    \pmb{\Sigma}\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+    \Big)
+\end{align}
 %
@@
     \,\,\,
-    \alpha
+    a
     \,\,\,
-    \beta
+    b
     \,\,\,
@@
 where $N$ is the number of star spots, $c$ is their contrast,
-$\alpha$ and $\beta$ describe the latitudinal
+$a$ and $b$ describe the latitudinal
 distribution of the spots, and $r$ and $\Delta$ are the mean and
@@
 the PDF for the latitude $\phi$ of the spots is a Beta distribution in
-$\cos\phi$ with shape parameters $\alpha$ and $\beta$ (Appendix \ref{sec:lat-lon}).
+$\cos\phi$ with (normalized) shape parameters $a$ and $b$,
+which have a one-to-one correspondence to the mean $\mu_\phi$ and
+standard deviation $\sigma_\phi$ of the distribution in $\phi$
+(Appendix \ref{sec:lat-lon}).
 The spot longitude is assumed to be uniformly distributed. Finally,
@@
 
-\xxx{More soon. Include likelihood function.}
+In this paper, we assume that the parameters $\pmb{\theta}_\bullet$
+described above are the \emph{physically interesting} ones. That is,
+given a set of $M$ light curves
+$\left( \mathbf{f}_0 \, \mathbf{f}_1 \, \cdots \,  \mathbf{f}_{M-1} \right)^\top$,
+we wish to infer the statistical properties of the star spots, encoded in
+the entries of the vector $\pmb{\theta}_\bullet$.
+%
+This is typically a tall order, since it requires marginalizing over all
+the nuisance parameters, which include the nitty-gritty details of the
+size, contrast, and location of \emph{every spot} on \emph{every star}
+in the sample. Fortunately, however, the Gaussian process we constructed
+does just that. Specifically, given the mean and covariance of the process,
+we are able to directly evaluate the log marginal likelihood of the $m^\mathrm{th}$
+dataset
+conditioned on a specific value of $\pmb{\theta}_\bullet$ (and $\pmb{\theta}_\star$):
+%
+\begin{align}
+    \label{eq:log-like}
+    \resizebox{.9\hsize}{!}{$
+        \ln \mathcal{L}_m\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+        =
+        -\frac{1}{2}
+        \mathbf{r}_m^\top\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+        \mathbf{K}_m^{-1}\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+        \mathbf{r}_m\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+        -
+        \frac{1}{2}
+        \ln \Big| \mathbf{K}_m\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right) \Big|
+        -
+        \frac{K_m}{2}
+        \ln \left( 2 \pi \right)
+        \quad,
+    $}
+\end{align}
+%
+where
+%
+\begin{align}
+    \mathbf{r}_m\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+     & \equiv
+    \mathbf{f}_m - \pmb{\mu}\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+\end{align}
+%
+is the residual vector,
+%
+\begin{align}
+    \mathbf{K}_m\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+     & \equiv
+    \pmb{\Sigma}\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+    +
+    \mathbf{C}_m
+\end{align}
+%
+is the sum of the GP covariance and the data covariance $\mathbf{C}_m$
+(which in most cases is a diagonal matrix whose entries
+are the squared uncertainty corresponding to each data point in the light curve),
+%
+$| \cdots |$ denotes the determinant, and $K_m$ is the number of data points.
+In an ensemble analysis, the joint marginal likelihood of all datasets is
+simply the product of the individual likelihoods, so in log space we have
+%
+\begin{align}
+    \ln \mathcal{L}\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+     & =
+    \sum_{m} \ln \mathcal{L}_m\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+    \quad.
+\end{align}
+%
+The marginal likelihood may be interpreted as the probability of the data
+given the model. Typically, we are interested in the reverse: the probability
+of the \emph{model} given the \emph{data}, i.e., the posterior probability
+distribution. In later sections we present a comprehensive suite of
+posterior inference exercises demonstrating that our GP model is correctly
+calibrated, allowing one to efficiently infer statistical properties of star spots
+from light curves with minimal bias.
 
@@
 %
-The equations for the mean and covariance of the flux GP
+As we mentioned in the previous section,
+the equations for the mean and covariance of the flux GP
 (Equations~\ref{eq:mean_f} and \ref{eq:cov_f}, respectively) are conditioned
@@
 %
+We may then replace our expressions for the GP mean and covariance with
+%
+\begin{align}
+    \pmb{\mu}(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet)
+     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+    \\
+    \pmb{\Sigma}(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet)
+     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet) \pmb{\mu}^\top(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet)
+    \quad.
+\end{align}
+%
+and use these to compute the marginal likelihood instead.
 
@@
 \begin{align}
-    \tilde{\mathbf{f}} & = \frac{\mathbf{f}}{\left<f\right>}
-    \quad,
-\end{align}
-%
-where $\tilde{\mathbf{f}}$ is the normalized, unit-mean light curve,
+    \tilde{\mathbf{f}\hspace{0.2em}} & = \frac{\mathbf{f}}{\left<f\right>}
+    \quad,
+\end{align}
+%
+where $\tilde{\mathbf{f}\hspace{0.2em}}$ is the normalized, unit-mean light curve,
 $\mathbf{f}$ is the measured light curve (in detector counts), and
@@
 because the normalized process is \emph{not} strictly Gaussian: the distribution
-has heavy tails due to the fact that $\tilde{\mathbf{f}}$ diverges as
+has heavy tails due to the fact that $\tilde{\mathbf{f}\hspace{0.2em}}$ diverges as
 the sample mean approaches zero. In fact, because of these tails, the covariance
@@
 \begin{align}
-    \tilde{\mathbf{f}} \sim \mathcal{N}\left( \mathbf{1}, \tilde{\pmb{\Sigma}} \right)
+    \tilde{\mathbf{f}\hspace{0.2em}}
+    \left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+    \sim
+    \mathcal{N}\left(
+    \mathbf{1},
+    \tilde{\pmb{\Sigma}} \left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+    \right)
     \quad.
@@
 This is appropriate as long as the quantity $z \ll 1$, for which the true
-distribution of $\tilde{\mathbf{f}}$ is approximately Gaussian. In practice,
+distribution of $\tilde{\mathbf{f}\hspace{0.2em}}$ is approximately Gaussian. In practice,
 we recommend employing this trick only for $z \lesssim 0.02$, for which the
@@
 above.
+
+\begin{figure}[ht!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/normgp.pdf}
+        \oscaption{normgp}{%
+            An example of a flux covariance matrix $\pmb{\Sigma}$ for a
+            \starry process (left) and the corresponding covariance of the
+            normalized process (right), computed from Equation~(\ref{eq:SigmaTilde}).
+            In addition to an offset and an overall scaling relative to the
+            original covariance matrix, the covariance of the normalized process
+            is discernibly non-stationary.
+            \label{fig:normgp}
+        }
+    \end{centering}
+\end{figure}
+
+Figure~\ref{fig:normgp} shows an example of a covariance matrix normalized
+according to the procedure outlined above. The principal difference between
+the normalized covariance and the original covariance is an overall
+scaling and a small offset. However, the normalization also results in
+the process becoming non-stationary: the covariance between two points in
+a light curve is now slightly dependent on their phases.
+
+\subsection{Other stuff}
+\label{sec:baseline}
+
+Figure~\ref{fig:nongaussianity}
+\begin{figure}[ht!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/nongaussianity.pdf}
+        \oscaption{nongaussianity}{%
+            Non-Gaussianity of the process.
+            \label{fig:nongaussianity}
+        }
+    \end{centering}
+\end{figure}
 

### 2020-12-01 starry_process/sp.py

@@ -395,9 +395,9 @@ class StarryProcess(object):
         K = Sig.shape[0]
         j = tt.ones((K, 1))
         m = tt.mean(Sig)
-        mvec = tt.dot(Sig, j) / K
+        q = tt.dot(Sig, j) / (K * m)
         z = m / mu ** 2
-        s = m * j - mvec
+        p = j - q
 
         # Coefficients
         fac = 1.0
@@ -409,12 +409,12 @@ class StarryProcess(object):
             fac *= z * (2 * n + 3)
 
         # We're done
-        ssT = tt.dot(s, tt.transpose(s))
-        migT = tt.dot(mvec, tt.transpose(mvec))
-        normSig = (
-            alpha * Sig + alpha / m * (ssT - migT) + beta / m * ssT
-        ) / mu ** 2
-        return ifelse(tt.gt(z, self._normzmax), Sig * np.inf, normSig,)
+        ppT = tt.dot(p, tt.transpose(p))
+        qqT = tt.dot(q, tt.transpose(q))
+        normSig = (alpha / mu ** 2) * Sig + z * (
+            (alpha + beta) * ppT - alpha * qqT
+        )
+        return ifelse(tt.gt(z, self._normzmax), Sig * np.inf, normSig)
 
     def sample(self, t, i=defaults["i"], p=defaults["p"], nsamples=1):
         """

### 2020-12-01 tex/figures/ylms.py

@@ -1,0 +1,65 @@
+import starry
+import matplotlib.pyplot as plt
+import numpy as np
+
+# Settings
+lmax = 5
+res = 300
+
+# Set up the plot
+fig, ax = plt.subplots(lmax + 1, 2 * lmax + 1, figsize=(9, 6))
+fig.subplots_adjust(hspace=0)
+for axis in ax.flatten():
+    axis.set_xticks([])
+    axis.set_yticks([])
+    axis.spines["top"].set_visible(False)
+    axis.spines["right"].set_visible(False)
+    axis.spines["bottom"].set_visible(False)
+    axis.spines["left"].set_visible(False)
+for l in range(lmax + 1):
+    ax[l, lmax - l].set_ylabel(
+        r"$l = %d$" % l,
+        rotation="horizontal",
+        labelpad=15,
+        y=0.35,
+        fontsize=11,
+        alpha=0.5,
+    )
+for j, m in enumerate(range(-lmax, lmax + 1)):
+    if m < 0:
+        ax[-1, j].set_xlabel(
+            r"$m {=} $-$%d$" % -m,
+            labelpad=5,
+            fontsize=11,
+            rotation="45",
+            x=0.3,
+            alpha=0.5,
+        )
+    else:
+        ax[-1, j].set_xlabel(
+            r"$m = %d$" % m,
+            labelpad=5,
+            fontsize=11,
+            rotation=45,
+            x=0.35,
+            alpha=0.5,
+        )
+
+# Loop over the orders and degrees
+map = starry.Map(lmax, lazy=False)
+for i, l in enumerate(range(lmax + 1)):
+    for j, m in enumerate(range(-l, l + 1)):
+
+        # Offset the index for centered plotting
+        j += lmax - l
+
+        # Compute the spherical harmonic
+        map.reset()
+        if l > 0:
+            map[l, m] = 1.0
+        map.show(ax=ax[i, j], grid=False)
+        ax[i, j].axis("on")
+        ax[i, j].set_rasterization_zorder(1)
+
+# We're done
+fig.savefig(__file__.replace(".py", ".pdf"), bbox_inches="tight", dpi=300)

### 2020-12-02 tex/gitlinks.tex
more text, new fig

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/709731d2744831cbf6cc4e9bf5a8473f0c7d29e3/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/709731d2744831cbf6cc4e9bf5a8473f0c7d29e3/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/709731d2744831cbf6cc4e9bf5a8473f0c7d29e3/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/ae47501186fb2c08dfe0ce5e5d8355fab2d9bfad/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/ae47501186fb2c08dfe0ce5e5d8355fab2d9bfad/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/ae47501186fb2c08dfe0ce5e5d8355fab2d9bfad/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-02 tex/ms.tex
more text, new fig

@@
 a set of orthogonal functions on the surface of the sphere which we will use
-to describe the intensity field on the surface of a star. As we will see below,
+to describe the intensity field on the surface of a star
+(Figure~\ref{fig:ylms}). As we will see below,
 the spherical harmonics are a particularly convenient basis in which to
@@
 standard deviation $\sigma_\phi$ of the distribution in $\phi$
-(Appendix \ref{sec:lat-lon}).
-The spot longitude is assumed to be uniformly distributed. Finally,
+(Appendix \ref{sec:lat}).
+The spot longitude is assumed to be uniformly distributed
+(Appendix \ref{sec:lon}). Finally,
 the PDFs for the spot contrast and the number of spots are delta functions
-centered at $c$ and $N$, respectively.
+centered at $c$ and $N$, respectively (Appendix \ref{sec:contrast}).
 
@@
         \oscaption{nongaussianity}{%
-            Non-Gaussianity of the process.
+            Corner plot showing the joint distributions of select spherical
+            harmonic coefficients corresponding to the $l=15$ expansion of
+            $5\times 10^{4}$ stellar surface maps drawn from a certain
+            star spot distribution (black points and contours).
+            At $l=15$ there are 256 coefficients in total;
+            we chose five of the coefficients with the most odd-looking
+            distributions to illustrate the non-Gaussianity of the process.
+            In addition to non-linear correlations, skewness, and the
+            existence of points of very high curvature, some of the distributions
+            are also multi-modal.
+            The orange contours show slices of the Gaussian approximation to
+            the joint distribution; this is the approximation adopted in
+            this paper.
             \label{fig:nongaussianity}
@@
 
-\subsection{The Radius Integral}
+\subsection{The Radius Integrals}
 \label{sec:size}
@@
 
-\subsection{The Latitude and Longitude Integrals}
-\label{sec:lat-lon}
+\subsection{The Latitude Integrals}
+\label{sec:lat}
 %
 Our goal in this section is to compute the first and second moments
-of the latitude ($\mathbf{e}_\phi$, $\mathbf{E}_\phi$)
-and longitude ($\mathbf{e}_\lambda$, $\mathbf{E}_\lambda$) distributions
-(Equations \ref{eq:e2}, \ref{eq:e3}, \ref{eq:E2}, and \ref{eq:E3},
-respectively). These involve integrals over the terms in the Wigner
+of the latitude distribution ($\mathbf{e}_\phi$ and $\mathbf{E}_\phi$,
+given by Equations \ref{eq:e2} and \ref{eq:E2}, respectively).
+These involve integrals over the terms in the Wigner
 rotation matrix for spherical harmonics, which we discuss below.
@@
 \subsubsection{Rotation matrices}
+\label{sec:wigner}
 %
@@
 
-\subsubsection{Probability density function for the latitude}
+\subsubsection{Probability density function}
 %
@@
 %
-The Beta distribution in $\cos\phi$ has hyperparameters
-%
-\begin{align}
-    \pmb{\theta}_\phi = \left(
-    \alpha \, \, \, \,
-    \beta \right)^\top
-\end{align}
-%
+The Beta distribution in $\cos\phi$ has hyperparameters $\alpha$ and $\beta$
 and PDF given by
@@
     \label{eq:cosphi-pdf}
-    p \big(\cos\phi \, \big| \, \pmb{\theta}_\phi \big)
+    p \big(\cos\phi \, \big| \, \alpha, \beta \big)
      & =
@@
     \label{eq:phi-pdf}
-    p \big(\phi \, \big| \, \pmb{\theta}_\phi \big)
+    p \big(\phi \, \big| \, \alpha, \beta \big)
     & =
@@
 for $\phi \in \left[ -\frac{\pi}{2}, \frac{\pi}{2} \right]$.
-
-\subsubsection{First moment of the latitude}
+%
+Both $\alpha$ and $\beta$ are restricted to $(0, \infty)$.
+However, in practice it is necessary to limit the values of these parameters
+to a finite range to ensure the numerical stability of the algorithm.
+It is also convenient
+to work with the log of these quantities because of their large dynamic range.
+We therefore introduce the modified shape parameters
+%
+\begin{align}
+    \label{eq:gauss2beta}
+    a & \equiv \frac{\ln\alpha - K_{00}}{K_{10} - K_{00}}
+    \nonumber                                             \\[0.5em]
+    b & \equiv \frac{\ln\beta - K_{10}}{K_{11} - K_{10}}
+\end{align}
+%
+with inverse transform
+%
+\begin{align}
+    \label{eq:beta2gauss}
+    \alpha & = e^{K_{00} + (K_{10} - K_{00})a}
+    \nonumber                                  \\
+    \beta  & = e^{K_{10} + (K_{11} - K_{10})b}
+    \quad,
+\end{align}
+%
+where the matrix
+%
+\begin{align}
+    \setstackgap{L}{1.25\baselineskip}
+    \fixTABwidth{T}
+    \mathbf{K}
+    =
+    \parenMatrixstack{
+    0              & 5 \\
+    \ln\frac{1}{2} & 5
+    }
+\end{align}
+%
+defines the minimum and maximum values of $\ln\alpha$ (top row) and $\ln\beta$
+(bottom row) we adopt in our implementation of the algorithm. The lower limits
+correspond to $\alpha > 1$ and $\beta > \frac{1}{2}$, which excludes
+distributions with unphysically sharp peaks at $\phi = 90^\circ$. Both $a$
+and $b$ are restricted to the domain $(0, 1)$, and together comprise the
+hyperparameter vector
+%
+\begin{align}
+    \pmb{\theta}_\phi = \left(
+    a \, \, \, \,
+    b \right)^\top
+    \quad.
+\end{align}
+%
+
+Although we use $a$ and $b$ as the inputs to our algorithm, below
+we derive expressions for the moments of the distribution in terms
+of $\alpha$ and $\beta$, as this is somewhat more convenient. Note that
+this requires us to explicitly add a Jacobian term to the likelihood
+when doing posterior inference to account for the change in volume due
+to the transformation from $\alpha, \beta$ to $a, b$. In practice, however,
+$\alpha$ and $\beta$ are no more natural a set of parameters to sample in,
+since they do not \emph{obviously} relate to physical quantities of interest.
+It is therefore desirable to introduce one final transform, from $\alpha, \beta$
+to parameters controlling the mean (or mode) $\mu_\phi$ and
+standard deviation  $\sigma_\phi$ of the
+distribution of spot latitudes, which \emph{are} of physical interest.
+We introduce this transform, and the associated Jacobian for the complete
+transform $\mu_\phi, \sigma_\phi \rightarrow a, b$,
+in \S\ref{sec:lat-transform} below.
+
+\subsubsection{First moment}
 %
@@
 
-\subsubsection{Second moment of the latitude}
+\subsubsection{Second moment}
 %
@@
 
-\subsubsection{Probability density function for the longitude}
-%
-The longitude integrals (Equations~\ref{eq:e3} and \ref{eq:E3}) involve
-rotations by an angle $\lambda$ about $\hat{\mathbf{y}}$, which
-may be accomplished by choosing
-Euler angles $\alpha = 0$, $\beta = \lambda$, and
-$\gamma = 0$, such that
-%
-\begin{align}
-    \mathbf{R}^l_{\hat{\mathbf{y}}}(\lambda)
-     & =
-    {\mathbf{U}^l}^{-1} \mathbf{D}^l_{\hat{\mathbf{y}}}(\lambda) \mathbf{U}^l
-\end{align}
-%
-with
-\begin{align}
-    \mathbf{D}^l_{\hat{\mathbf{y}}}(\lambda)
-     & =
-    \mathbf{D}^l\left(0, \lambda, 0\right)
-    \quad.
-\end{align}
-%
-Since we expect the longitudinal distribution of features on the surfaces
-of stars to be (on average) isotropic, we will place a uniform prior on
-$\lambda \in [-\pi, \pi)$:
-%
-\begin{align}
-    p(\lambda \big| \pmb{\theta}_\lambda)
-     & =
-    \begin{cases}
-        \frac{1}{2\pi} & -\pi \leq \lambda < \pi
-        \\
-        0              & \mathrm{otherwise} \quad.
-    \end{cases}
-\end{align}
-%
-We therefore have no hyperparameters
-controlling the longitudinal distribution, i.e.,
-%
-\begin{align}
-    \pmb{\theta}_\lambda = \left( \,\,\, \right)
-    \quad.
-\end{align}
-%
-
-\subsubsection{First moment of the longitude}
-%
-As before, we will solve for the terms of the moment integrals one
-spherical harmonic degree at a time:
-%
-\begin{proof}{test_elam}
-    \mathbf{e}_\lambda
-    & \equiv
-    \int
-    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
-    \mathbf{e}_\phi \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
-    \nonumber
-    \\
-    %
-    & =
-    %
-    \left(
-    \mathbf{e}_\lambda^0
-    \,\,\,
-    \mathbf{e}_\lambda^1
-    \,\,\,
-    \mathbf{e}_\lambda^2
-    \,\,\,
-    \cdots
-    \,\,\,
-    \mathbf{e}_\lambda^{l_{\mathrm{max}}}
-    \right)^\top
-    \quad,
-\end{proof}
-%
-where
-%
-\begin{proof}{test_elam}
-    \mathbf{e}_\lambda^l
-    & =
-    \int
-    \mathbf{R}^l_{\hat{\mathbf{y}}}(\lambda) \,
-    \mathbf{e}^l_\phi \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
-    \nonumber \\
-    %
-    & =
-    %
-    {\mathbf{U}^l}^{-1}
-    \mathbf{p}^l_\lambda
-    \quad,
-\end{proof}
-%
-and we define
-%
-\begin{align}
-    \label{eq:pllam}
-    \mathbf{p}^l_\lambda
-     & \equiv
-    \int
-    \mathbf{D}^l_{\hat{\mathbf{x}}}(\lambda) \,
-    \bar{\mathbf{e}}^l_\phi \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
-    \\
-    \bar{\mathbf{e}}^l_\phi
-     & \equiv
-    \mathbf{U}^l
-    \mathbf{e}^l_\phi
-    \quad.
-\end{align}
-%
-The integral $\mathbf{p}^l_\lambda$ defined above has a closed-form solution.
-To show this, we write the terms of $\mathbf{p}^l_\lambda$ as
-%
-\begin{align}
-    {({p^l_\lambda})_{}}_m
-     & =
-    \int
-    \sum\limits_{\mu=-l}^l
-    {({D^l_{\hat{\mathbf{y}}}})_{}}_{m,\mu}(\lambda) \,
-    {({\bar{e}^l_\phi})_{}}_{\mu} \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
-    %
-    \nonumber \\[0.5em]
-    %
-     & =
-    \sum\limits_{\mu=-l}^l
-    {({\bar{e}^l_\phi})_{}}_{\mu} \,
-    \sum\limits_{i=0}^{2l} c_{m,\mu,i}^{l}
-    \,
-    {({q^l_\lambda})_{}}_i
-    \quad,
-\end{align}
-%
-where
-%
-\begin{proof}{test_qllam}
-    {({q^l_\lambda})_{}}_i
-    & \equiv
-    \frac{1}{2\pi}
-    \int_{-\pi}^{\pi}
-    %
-    \mathrm{sgn}(\sin\lambda)^{i}
-    (1 - \cos\lambda)^{l - \frac{i}{2}}
-    (1 + \cos\lambda)^\frac{i}{2}
-    %
-    \,
-    \mathrm{d}\lambda
-    \nonumber \\[0.5em]
-    %
-    & =
-    %
-    \begin{cases}
-        \dfrac{2^l \Gamma \left(\frac{1+i}{2}\right)
-            \Gamma \left(l+\frac{1-i}{2}\right)}{\pi  \Gamma (l+1)}
-        %
-         & i \,\, \mathrm{even}
-        %
-        \\
-        %
-        0
-        %
-         & i \,\, \mathrm{odd} \quad,
-    \end{cases}
-\end{proof}
-%
-whose terms may easily be computed by upward recursion.
-
-\subsubsection{Second moment of the longitude}
-%
-We write the second moment integral as
+\subsubsection{Variable transforms}
+\label{sec:lat-transform}
+
+\xxx{...}
+
+Although we adopt $a$ and $b$ as the actual inputs to our Gaussian process
+(i.e., as elements of the input vector $\pmb{\theta}_\bullet$), these
+parameters are not themselves ``interpretable'', as they do not directly
+correspond to physical
+quantities of interest such as the mean and standard deviation of the spot
+latitude distribution. However, over most of the parameter space in $a$ and $b$,
+the spot latitude distribution is well approximated
+by a bimodal Gaussian. In particular, there exists
+a one-to-one relationship between $a$ and $b$ and the mean $\mu_\phi$
+and standard deviation $\sigma_\phi$ of a normal approximation to the
+distribution. Moreover, we find that
+if we let $\mu_\phi$ be the \emph{mode} of the PDF and
+$\sigma_\phi^2$ be a \emph{local} approximation to the variance of the PDF,
+the relationship has a convenient closed form.
+
+To compute $\mu_\phi$, we differentiate Equation~(\ref{eq:phi-pdf}) with respect
+to $\phi$, set the expression equal to zero, and solve
+for $\phi$ to obtain
 %
 \begin{proof}{}
-    \mathbf{E}_\lambda
-    & \equiv
-    \int
-    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
-    \mathbf{E}_\phi \,
-    \mathbf{R}_{\hat{\mathbf{y}}}^\top(\lambda) \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
-    \nonumber
-    \\
-    %
-    & =
-    %
-    \setstackgap{L}{1.25\baselineskip}
-    \fixTABwidth{T}
-    \parenMatrixstack{
-    \mathbf{E}_\lambda^{0,0} & \mathbf{E}_\lambda^{0,1} & \mathbf{E}_\lambda^{0,2} & \cdots & \mathbf{E}_\lambda^{0,l_{\mathrm{max}}} \\
-    \mathbf{E}_\lambda^{1,0} & \mathbf{E}_\lambda^{1,1} & \mathbf{E}_\lambda^{1,2} & \cdots & \mathbf{E}_\lambda^{1,l_{\mathrm{max}}} \\
-    \mathbf{E}_\lambda^{2,0} & \mathbf{E}_\lambda^{2,1} & \mathbf{E}_\lambda^{2,2} & \cdots & \mathbf{E}_\lambda^{2,l_{\mathrm{max}}} \\
-    \vdots & \vdots & \vdots & \ddots & \vdots \\
-    \mathbf{E}_\lambda^{l_{\mathrm{max}},0} & \mathbf{E}_\lambda^{l_{\mathrm{max}},1} & \mathbf{E}_\lambda^{l_{\mathrm{max}},2} & \cdots & \mathbf{E}_\lambda^{l_{\mathrm{max}},l_{\mathrm{max}}}
-    }
-    \quad,
-\end{proof}
-%
-where
-%
-\begin{proof}{}
-    \mathbf{E}_\lambda^{l,l'}
-    & =
-    \int
-    \mathbf{R}^l_{\hat{\mathbf{y}}}(\lambda) \,
-    \mathbf{E}^l_\phi \,
-    {\mathbf{R}^{l'}_{\hat{\mathbf{y}}}}^\top(\lambda) \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
-    \nonumber \\
-    %
-    & =
-    %
-    {\mathbf{U}^l}^{-1}
-    \mathbf{P}^l_\lambda
-    {{\mathbf{U}^l}^{-1}}^\top
-\end{proof}
-%
-and we define
-%
-\begin{align}
-    \label{eq:Pllam}
-    \mathbf{P}^l_\lambda
-     & \equiv
-    \int
-    \mathbf{D}^l_{\hat{\mathbf{x}}}(\lambda) \,
-    \bar{\mathbf{E}}^l_\phi \,
-    {\mathbf{D}^l_{\hat{\mathbf{x}}}}^\top(\lambda) \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
-    \\
-    \bar{\mathbf{E}}^l_\phi
-     & \equiv
-    \mathbf{U}^l
-    \mathbf{E}^l_\phi
-    {\mathbf{U}^l}^\top
-    \quad.
-\end{align}
-%
-We then express the terms of $\mathbf{P}^l_\lambda$ as
-%
-\begin{align}
-    {({P^l_\lambda})_{}}_{m,m'}
-     & =
-    \int
-    \sum\limits_{\mu=-l}^l
-    \sum\limits_{\nu=-l}^l
-    ({D^l_{\hat{\mathbf{x}}}})_{m,\mu}(\lambda) \,
-    (\bar{E}^l_\phi)_{\mu,\nu} \,
-    ({D^l_{\hat{\mathbf{x}}}})_{\nu,m'}(\lambda) \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
-    %
-    \nonumber \\[0.5em]
-    %
-     & =
-    \sum\limits_{\mu=-l}^l
-    \sum\limits_{\nu=-l}^l
-    (\bar{E}^l_\phi)_{\mu,\nu}
-    \sum\limits_{i=0}^{2l}
-    \sum\limits_{j=0}^{2l}
-    c_{m,\mu,i}^{l}
-    c_{\nu,m',j}^{l}
-    \,
-    (Q^l_\lambda)_{i,j}
-    \quad,
-\end{align}
-%
-where
-%
-\begin{proof}{}
-    (Q^l_\lambda)_{i,j}
-    & \equiv
-    \frac{1}{2\pi}
-    \int_{-\pi}^{\pi}
-    %
-    \mathrm{sgn}(\sin\lambda)^{i+j}
-    (1 - \cos\lambda)^{2l - \frac{i+j}{2}}
-    (1 + \cos\lambda)^\frac{i+j}{2}
-    %
-    \,
-    \mathrm{d}\lambda
-    \nonumber \\[0.5em]
-    %
-    & =
-    %
-    \begin{cases}
-        \dfrac{2^{2l} \Gamma \left(\frac{1+i+j}{2}\right)
-            \Gamma \left(2l+\frac{1-i-j}{2}\right)}{\pi  \Gamma (2l+1)}
-        %
-         & i+j \,\, \mathrm{even}
-        %
-        \\
-        %
-        0
-        %
-         & i+j \,\, \mathrm{odd} \quad,
-    \end{cases}
-\end{proof}
-%
-whose terms may again be computed by upward recursion.
-
-\subsection{The Contrast Integral}
-\label{sec:contrast}
-%
-The final integrals we must take in our computation of
-$\mathrm{E} \big[ \mathbf{y} \, \big| \, \pmb{\theta}_\bullet \big]$
-and $\mathrm{E} \big[ \mathbf{y} \mathbf{y}^\top \, \big| \, \pmb{\theta}_\bullet \big]$
-are the integrals over
-the spot contrast distribution, $\mathbf{e}_\xi$ and $\mathbf{E}_\xi$.
-These are by far the easiest, since the spot contrast is a scalar
-multiplier of the spherical harmonic coefficient vector, so we can
-pull the terms $\mathbf{e}_\lambda$ and $\mathbf{E}_\lambda$ out
-of the
-integrals in Equations (\ref{eq:e4}) and (\ref{eq:E4}) to write
-%
-\begin{align}
-    \mathbf{e}_\xi
-     & \equiv
-    \mathbf{e}_\lambda \,
-    \int
-    \xi \,
-    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
-    \mathrm{d}\xi
-    \\
-    \mathbf{E}_\xi
-     & \equiv
-    \mathbf{E}_\lambda \,
-    \int
-    \xi^2 \,
-    p(\xi \, \big| \, \pmb{\theta}_\xi)
-    \mathrm{d}\xi
-    \quad.
-\end{align}
-%
-These integrals may be computed analytically for any choice of
-probability density function $p(\xi \, \big| \, \pmb{\theta}_\xi)$
-with closed-form moments.
-%
-However, in practice, it is quite difficult to constrain the
-spot contrast from light curves, let alone higher moments of its
-distribution; this is due largely to the fact that the contrast
-is extremely degenerate with the total number of spots
-(see \S\ref{sec:cN-degeneracy}).
-%
-In our implementation of the algorithm,
-we therefore choose the simplest possible probability distribution,
-a delta function:
-%
-\begin{align}
-    p(\xi \, \big| \, \pmb{\theta}_{\xi}) = \delta(\xi - c)
-\end{align}
-%
-characterized by a single parameter, the contrast of the spots:
-%
-\begin{align}
-    \pmb{\theta}_\xi = \left( \, c \, \right)^\top
-    \quad.
-\end{align}
-%
-The moment integrals are then trivial to evaluate:
-%
-\begin{align}
-    \mathbf{e}_\xi & = c \, \mathbf{e}_\lambda
-    \\
-    \mathbf{E}_\xi & = c^2 \mathbf{E}_\lambda
-    \quad.
-\end{align}
-%
-
-\section{Inclination}
-\label{sec:inclination}
-
-\begin{align}
-    p(i) = \sin i
-    \quad.
-\end{align}
-
-\subsection{First moment}
-\xxx{Coming soon.}
-
-\subsection{Second moment}
-\xxx{Coming soon.}
-
-\section{Latitude Distribution Transforms}
-
-Given the shape parameters $\alpha$ and $\beta$,
-the mode $\mu$ and standard deviation $\sigma$ of the
-distribution of spot latitudes may be written
-%
-\begin{proof}{}
-    \mu &= 2 \tan^{-1}
+    \mu_\phi &= 2 \tan^{-1}
     \left(
@@
     \right)
-    \\
-    \sigma &= \frac{\sin\mu}{
+\end{proof}
+%
+where $\alpha$ and $\beta$ may be computed from $a$ and $b$ via
+Equation~(\ref{eq:beta2gauss}).
+%
+To compute $\sigma_\phi$, we note that the variance of a Gaussian distribution
+$\varphi(\phi)$
+is the negative reciprocal of its curvature in log space:
+%
+\begin{align}
+    \sigma^2 = -\left(\frac{\mathrm{d}^2 \ln \varphi(\phi)}{\mathrm{d}\phi^2}\right)^{-1}
+\end{align}
+%
+We therefore twice differentiate the log of Equation~(\ref{eq:phi-pdf}), negate it,
+take the reciprocal, and evaluate it at $\phi = \mu_\phi$ to obtain a local
+approximation to the standard deviation of the distribution at the mode:
+%
+\begin{proof}{}
+    \sigma_\phi &= \frac{\sin\mu_\phi}{
         \sqrt{1
@@
             + \beta
-            + \left(\beta - 1\right) \cos\mu
-            + \dfrac{\alpha - 1}{\cos\mu^2}}}
+            + \left(\beta - 1\right) \cos\mu_\phi
+            + \dfrac{\alpha - 1}{\cos\mu_\phi^2}}}
     \quad.
@@
 %
-The reverse transform is given by
+For completeness, the inverse transform also has a closed form:
 %
 \begin{proof}{}
-    \alpha &= \frac{2 + 4 \sigma^2 + (3 + 8 \sigma^2) \cos\mu + 2 \cos\left( 2\mu \right) + \cos\left( 3\mu \right)}{16 \sigma^2 \cos\left( \frac{\mu}{2} \right)^4} \\
-    \beta &= \frac{\cos\mu + 2 \sigma^2 (3 + \cos\left( 2\mu \right)) - \cos\left( 3\mu \right)}{16 \sigma^2 \cos\left( \frac{\mu}{2} \right)^4}
+    \alpha &= \frac{2 + 4 \sigma_\phi^2 + (3 + 8 \sigma_\phi^2) \cos\mu_\phi + 2 \cos\left( 2\mu_\phi \right) + \cos\left( 3\mu_\phi \right)}{16 \sigma_\phi^2 \cos\left( \frac{\mu_\phi}{2} \right)^4} \\
+    \beta &= \frac{\cos\mu_\phi + 2 \sigma_\phi^2 (3 + \cos\left( 2\mu_\phi \right)) - \cos\left( 3\mu_\phi \right)}{16 \sigma_\phi^2 \cos\left( \frac{\mu_\phi}{2} \right)^4}
     \quad.
 \end{proof}
-%
-These expressions may be used to compute the Jacobian of the transform:
+
+These transformations allow one to easily transform constraints on $a$ and
+$b$ to constraints on the mode and approximate variance of the distribution
+of spot latitudes. Importantly, they also allow us to compute the Jacobian
+of the transform, given by
 %
@@
     J &=
-    \frac{\partial{\mu}}{\partial{\alpha}}
-    \frac{\partial{\sigma}}{\partial{\beta}} -
-    \frac{\partial{\mu}}{\partial{\beta}}
-    \frac{\partial{\sigma}}{\partial{\alpha}}
+    \frac{\partial{\mu_\phi}}{\partial{a}}
+    \frac{\partial{\sigma_\phi}}{\partial{b}} -
+    \frac{\partial{\mu_\phi}}{\partial{b}}
+    \frac{\partial{\sigma_\phi}}{\partial{a}}
     %
@@
             \frac{
-                \alpha \beta \left(1 + \cos\mu\right)^3
-                \sin\left(2\mu\right)^3
+                \alpha \beta \left(1 + \cos\mu_\phi\right)^3
+                \sin\left(2\mu_\phi\right)^3
             }{
-                \sigma
+                \sigma_\phi
                 \left(
                 2 \alpha + \beta - 3 +
-                \left(2 \alpha + \beta - 1\right) \cos\mu
+                \left(2 \alpha + \beta - 1\right) \cos\mu_\phi
                 \right)
@@
                 2 \left( \alpha + \beta - 1\right) +
-                3 (\beta - 1) \cos\mu
-                - 2 \left( \alpha - \beta - 1 \right) \cos\left( 2\mu \right)
-                + \left( \beta - 1\right) \cos\left( 3\mu \right)
+                3 (\beta - 1) \cos\mu_\phi
+                - 2 \left( \alpha - \beta - 1 \right) \cos\left( 2\mu_\phi \right)
+                + \left( \beta - 1\right) \cos\left( 3\mu_\phi \right)
                 \right)^2
@@
 \end{proof}
+%
+While one could sample in $\mu_\phi$ and $\sigma_\phi$ directly, we instead
+recommend sampling in the quantities $a$ and $b$ under a uniform prior on
+$(0, 1)$ for both of them, then correcting for the implied prior on
+$\mu_\phi$ and $\sigma_\phi$ by adding the log of the absolute value of $J$ to the log
+likelihood. This corrects for the \emph{ad hoc}
+prior on the latitude parameters introduced by our particular choice of
+parametrization, enforcing instead a uniform prior on the quantities $\mu_\phi$
+and $\sigma_\phi$.
+
+\begin{figure}[ht!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/latitude_pdf.pdf}
+        \oscaption{latitude_pdf}{%
+            Probability density function for the spot latitude (blue curves)
+            for different values of the mode $\mu_\phi$ (columns)
+            and local standard deviation $\sigma_\phi$ (rows).
+            The corresponding values of $a$ and $b$ are
+            indicated within each panel. The bimodal
+            normal distribution with mean $\mu_\phi$ and standard deviation
+            $\sigma_\phi$ is shown as the dashed orange curves; for mid-latitude
+            modes and low standard deviations, the Gaussian approximation is
+            quite good.
+            The shaded panel in the lower right
+            ($\mu_\phi = 0^\circ$, $\sigma_\phi = 0^\circ$)
+            corresponds to an approximately isotropic distribution of
+            spots over the surface of the star; in this panel only, the
+            dashed orange curve corresponds to a cosine distribution in
+            $\phi$ (i.e., the exact isotropic distribution).
+            \label{fig:latitude_pdf}
+        }
+    \end{centering}
+\end{figure}
+
+For reference, Figure~\ref{fig:latitude_pdf} shows the latitude PDF
+and the corresponding Gaussian approximation for different values of
+$\mu_\phi$ and $\sigma_\phi$. The corresponding values of $a$ and $b$ are
+indicated in the top left of each panel. For $\mu_\phi$ at intermediate
+latitudes and moderate values of $\sigma_\phi$, the approximation is quite good.
+However, for $\mu_\phi$ very close to the equator or to the poles, the curvature of
+the distribution changes significantly as a function of $\phi$, so the variance is somewhat
+underestimated by the approximation; and for $\sigma_\phi$ large, the
+distribution becomes noticeably non-Gaussian.
+
+The shaded panel at the lower left is a special case of the distribution
+($\mu_\phi = 0^\circ$, $\sigma_\phi \approx 40^\circ$; or equivalently,
+$a \approx 0.06$, $b = 0$), which is approximately isotropic in latitude. In
+this panel, the orange curve instead corresponds to an isotropic (cosine)
+distribution in $\phi$; note the excellent agreement.
+
+Thus, in addition to having closed-form moments, the Beta distribution is
+quite flexible and, via the transforms outlined above, intuitive in how
+it affects the distribution of spots on the surface of a star.
+
+\subsection{The Longitude Integrals}
+\label{sec:lon}
+
+In this section we will compute the first and second moments
+of the longitude distribution
+($\mathbf{e}_\lambda$ and $\mathbf{E}_\lambda$, given by Equations
+\ref{eq:e3} and \ref{eq:E3}, respectively). The math here is
+very similar to that in the previous section, as we are again
+dealing with integrals of Wigner matrices (\S\ref{sec:wigner}).
+
+\subsubsection{Probability density function}
+%
+The longitude integrals (Equations~\ref{eq:e3} and \ref{eq:E3}) involve
+rotations by an angle $\lambda$ about $\hat{\mathbf{y}}$, which
+may be accomplished by choosing
+Euler angles $\alpha = 0$, $\beta = \lambda$, and
+$\gamma = 0$, such that
+%
+\begin{align}
+    \mathbf{R}^l_{\hat{\mathbf{y}}}(\lambda)
+     & =
+    {\mathbf{U}^l}^{-1} \mathbf{D}^l_{\hat{\mathbf{y}}}(\lambda) \mathbf{U}^l
+\end{align}
+%
+with
+\begin{align}
+    \mathbf{D}^l_{\hat{\mathbf{y}}}(\lambda)
+     & =
+    \mathbf{D}^l\left(0, \lambda, 0\right)
+    \quad.
+\end{align}
+%
+Since we expect the longitudinal distribution of features on the surfaces
+of stars to be (on average) isotropic, we will place a uniform prior on
+$\lambda \in [-\pi, \pi)$:
+%
+\begin{align}
+    p(\lambda \big| \pmb{\theta}_\lambda)
+     & =
+    \begin{cases}
+        \frac{1}{2\pi} & -\pi \leq \lambda < \pi
+        \\
+        0              & \mathrm{otherwise} \quad.
+    \end{cases}
+\end{align}
+%
+We therefore have no hyperparameters
+controlling the longitudinal distribution, i.e.,
+%
+\begin{align}
+    \pmb{\theta}_\lambda = \left( \,\,\, \right)
+    \quad.
+\end{align}
+%
+
+\subsubsection{First moment}
+%
+As before, we will solve for the terms of the moment integrals one
+spherical harmonic degree at a time:
+%
+\begin{proof}{test_elam}
+    \mathbf{e}_\lambda
+    & \equiv
+    \int
+    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{e}_\phi \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    \nonumber
+    \\
+    %
+    & =
+    %
+    \left(
+    \mathbf{e}_\lambda^0
+    \,\,\,
+    \mathbf{e}_\lambda^1
+    \,\,\,
+    \mathbf{e}_\lambda^2
+    \,\,\,
+    \cdots
+    \,\,\,
+    \mathbf{e}_\lambda^{l_{\mathrm{max}}}
+    \right)^\top
+    \quad,
+\end{proof}
+%
+where
+%
+\begin{proof}{test_elam}
+    \mathbf{e}_\lambda^l
+    & =
+    \int
+    \mathbf{R}^l_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{e}^l_\phi \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    \nonumber \\
+    %
+    & =
+    %
+    {\mathbf{U}^l}^{-1}
+    \mathbf{p}^l_\lambda
+    \quad,
+\end{proof}
+%
+and we define
+%
+\begin{align}
+    \label{eq:pllam}
+    \mathbf{p}^l_\lambda
+     & \equiv
+    \int
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(\lambda) \,
+    \bar{\mathbf{e}}^l_\phi \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    \\
+    \bar{\mathbf{e}}^l_\phi
+     & \equiv
+    \mathbf{U}^l
+    \mathbf{e}^l_\phi
+    \quad.
+\end{align}
+%
+The integral $\mathbf{p}^l_\lambda$ defined above has a closed-form solution.
+To show this, we write the terms of $\mathbf{p}^l_\lambda$ as
+%
+\begin{align}
+    {({p^l_\lambda})_{}}_m
+     & =
+    \int
+    \sum\limits_{\mu=-l}^l
+    {({D^l_{\hat{\mathbf{y}}}})_{}}_{m,\mu}(\lambda) \,
+    {({\bar{e}^l_\phi})_{}}_{\mu} \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    %
+    \nonumber \\[0.5em]
+    %
+     & =
+    \sum\limits_{\mu=-l}^l
+    {({\bar{e}^l_\phi})_{}}_{\mu} \,
+    \sum\limits_{i=0}^{2l} c_{m,\mu,i}^{l}
+    \,
+    {({q^l_\lambda})_{}}_i
+    \quad,
+\end{align}
+%
+where
+%
+\begin{proof}{test_qllam}
+    {({q^l_\lambda})_{}}_i
+    & \equiv
+    \frac{1}{2\pi}
+    \int_{-\pi}^{\pi}
+    %
+    \mathrm{sgn}(\sin\lambda)^{i}
+    (1 - \cos\lambda)^{l - \frac{i}{2}}
+    (1 + \cos\lambda)^\frac{i}{2}
+    %
+    \,
+    \mathrm{d}\lambda
+    \nonumber \\[0.5em]
+    %
+    & =
+    %
+    \begin{cases}
+        \dfrac{2^l \Gamma \left(\frac{1+i}{2}\right)
+            \Gamma \left(l+\frac{1-i}{2}\right)}{\pi  \Gamma (l+1)}
+        %
+         & i \,\, \mathrm{even}
+        %
+        \\
+        %
+        0
+        %
+         & i \,\, \mathrm{odd} \quad,
+    \end{cases}
+\end{proof}
+%
+whose terms may easily be computed by upward recursion.
+
+\subsubsection{Second moment}
+%
+We write the second moment integral as
+%
+\begin{proof}{}
+    \mathbf{E}_\lambda
+    & \equiv
+    \int
+    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{E}_\phi \,
+    \mathbf{R}_{\hat{\mathbf{y}}}^\top(\lambda) \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    \nonumber
+    \\
+    %
+    & =
+    %
+    \setstackgap{L}{1.25\baselineskip}
+    \fixTABwidth{T}
+    \parenMatrixstack{
+    \mathbf{E}_\lambda^{0,0} & \mathbf{E}_\lambda^{0,1} & \mathbf{E}_\lambda^{0,2} & \cdots & \mathbf{E}_\lambda^{0,l_{\mathrm{max}}} \\
+    \mathbf{E}_\lambda^{1,0} & \mathbf{E}_\lambda^{1,1} & \mathbf{E}_\lambda^{1,2} & \cdots & \mathbf{E}_\lambda^{1,l_{\mathrm{max}}} \\
+    \mathbf{E}_\lambda^{2,0} & \mathbf{E}_\lambda^{2,1} & \mathbf{E}_\lambda^{2,2} & \cdots & \mathbf{E}_\lambda^{2,l_{\mathrm{max}}} \\
+    \vdots & \vdots & \vdots & \ddots & \vdots \\
+    \mathbf{E}_\lambda^{l_{\mathrm{max}},0} & \mathbf{E}_\lambda^{l_{\mathrm{max}},1} & \mathbf{E}_\lambda^{l_{\mathrm{max}},2} & \cdots & \mathbf{E}_\lambda^{l_{\mathrm{max}},l_{\mathrm{max}}}
+    }
+    \quad,
+\end{proof}
+%
+where
+%
+\begin{proof}{}
+    \mathbf{E}_\lambda^{l,l'}
+    & =
+    \int
+    \mathbf{R}^l_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{E}^l_\phi \,
+    {\mathbf{R}^{l'}_{\hat{\mathbf{y}}}}^\top(\lambda) \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    \nonumber \\
+    %
+    & =
+    %
+    {\mathbf{U}^l}^{-1}
+    \mathbf{P}^l_\lambda
+    {{\mathbf{U}^l}^{-1}}^\top
+\end{proof}
+%
+and we define
+%
+\begin{align}
+    \label{eq:Pllam}
+    \mathbf{P}^l_\lambda
+     & \equiv
+    \int
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(\lambda) \,
+    \bar{\mathbf{E}}^l_\phi \,
+    {\mathbf{D}^l_{\hat{\mathbf{x}}}}^\top(\lambda) \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    \\
+    \bar{\mathbf{E}}^l_\phi
+     & \equiv
+    \mathbf{U}^l
+    \mathbf{E}^l_\phi
+    {\mathbf{U}^l}^\top
+    \quad.
+\end{align}
+%
+We then express the terms of $\mathbf{P}^l_\lambda$ as
+%
+\begin{align}
+    {({P^l_\lambda})_{}}_{m,m'}
+     & =
+    \int
+    \sum\limits_{\mu=-l}^l
+    \sum\limits_{\nu=-l}^l
+    ({D^l_{\hat{\mathbf{x}}}})_{m,\mu}(\lambda) \,
+    (\bar{E}^l_\phi)_{\mu,\nu} \,
+    ({D^l_{\hat{\mathbf{x}}}})_{\nu,m'}(\lambda) \,
+    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\lambda
+    %
+    \nonumber \\[0.5em]
+    %
+     & =
+    \sum\limits_{\mu=-l}^l
+    \sum\limits_{\nu=-l}^l
+    (\bar{E}^l_\phi)_{\mu,\nu}
+    \sum\limits_{i=0}^{2l}
+    \sum\limits_{j=0}^{2l}
+    c_{m,\mu,i}^{l}
+    c_{\nu,m',j}^{l}
+    \,
+    (Q^l_\lambda)_{i,j}
+    \quad,
+\end{align}
+%
+where
+%
+\begin{proof}{}
+    (Q^l_\lambda)_{i,j}
+    & \equiv
+    \frac{1}{2\pi}
+    \int_{-\pi}^{\pi}
+    %
+    \mathrm{sgn}(\sin\lambda)^{i+j}
+    (1 - \cos\lambda)^{2l - \frac{i+j}{2}}
+    (1 + \cos\lambda)^\frac{i+j}{2}
+    %
+    \,
+    \mathrm{d}\lambda
+    \nonumber \\[0.5em]
+    %
+    & =
+    %
+    \begin{cases}
+        \dfrac{2^{2l} \Gamma \left(\frac{1+i+j}{2}\right)
+            \Gamma \left(2l+\frac{1-i-j}{2}\right)}{\pi  \Gamma (2l+1)}
+        %
+         & i+j \,\, \mathrm{even}
+        %
+        \\
+        %
+        0
+        %
+         & i+j \,\, \mathrm{odd} \quad,
+    \end{cases}
+\end{proof}
+%
+whose terms may again be computed by upward recursion.
+
+\subsection{The Contrast Integrals}
+\label{sec:contrast}
+%
+The final integrals we must take in our computation of
+$\mathrm{E} \big[ \mathbf{y} \, \big| \, \pmb{\theta}_\bullet \big]$
+and $\mathrm{E} \big[ \mathbf{y} \mathbf{y}^\top \, \big| \, \pmb{\theta}_\bullet \big]$
+are the integrals over
+the spot contrast distribution, $\mathbf{e}_\xi$ and $\mathbf{E}_\xi$.
+These are by far the easiest, since the spot contrast is a scalar
+multiplier of the spherical harmonic coefficient vector, so we can
+pull the terms $\mathbf{e}_\lambda$ and $\mathbf{E}_\lambda$ out
+of the
+integrals in Equations (\ref{eq:e4}) and (\ref{eq:E4}) to write
+%
+\begin{align}
+    \mathbf{e}_\xi
+     & \equiv
+    \mathbf{e}_\lambda \,
+    \int
+    \xi \,
+    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
+    \mathrm{d}\xi
+    \\
+    \mathbf{E}_\xi
+     & \equiv
+    \mathbf{E}_\lambda \,
+    \int
+    \xi^2 \,
+    p(\xi \, \big| \, \pmb{\theta}_\xi)
+    \mathrm{d}\xi
+    \quad.
+\end{align}
+%
+These integrals may be computed analytically for any choice of
+probability density function $p(\xi \, \big| \, \pmb{\theta}_\xi)$
+with closed-form moments.
+%
+However, in practice, it is quite difficult to constrain the
+spot contrast from light curves, let alone higher moments of its
+distribution; this is due largely to the fact that the contrast
+is extremely degenerate with the total number of spots
+(see \S\ref{sec:cN-degeneracy}).
+%
+In our implementation of the algorithm,
+we therefore choose the simplest possible probability distribution,
+a delta function:
+%
+\begin{align}
+    p(\xi \, \big| \, \pmb{\theta}_{\xi}) = \delta(\xi - c)
+\end{align}
+%
+characterized by a single parameter, the contrast of the spots:
+%
+\begin{align}
+    \pmb{\theta}_\xi = \left( \, c \, \right)^\top
+    \quad.
+\end{align}
+%
+The moment integrals are then trivial to evaluate:
+%
+\begin{align}
+    \mathbf{e}_\xi & = c \, \mathbf{e}_\lambda
+    \\
+    \mathbf{E}_\xi & = c^2 \mathbf{E}_\lambda
+    \quad.
+\end{align}
+%
+
+\section{Inclination}
+\label{sec:inclination}
+
+\begin{align}
+    p(i) = \sin i
+    \quad.
+\end{align}
+
+\subsection{First moment}
+\xxx{Coming soon.}
+
+\subsection{Second moment}
+\xxx{Coming soon.}
 

### 2020-12-02 starry_process/defaults.py

@@ -16,7 +16,7 @@ defaults = dict(
     driver="numpy",
     eps1=1e-12,
     eps2=1e-9,
-    eps3=1e-8,
+    eps3=1e-12,
     covpts=300,
     log_alpha_max=10,
     log_beta_max=10,

### 2020-12-02 tex/figures/nullspace_ensemble.py

@@ -7,7 +7,7 @@ from tqdm import tqdm
 ntheta = 50
 ydeg = 15
 ydeg_pad = 3
-L = 1e12
+L = 1e9
 C = 1
 ninc = [1, 3, 10]
 kpn = 300

### 2020-12-03 tex/gitlinks.tex
revert eps3, rename to eps

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/ae47501186fb2c08dfe0ce5e5d8355fab2d9bfad/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/ae47501186fb2c08dfe0ce5e5d8355fab2d9bfad/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/ae47501186fb2c08dfe0ce5e5d8355fab2d9bfad/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/56f89997cccb24bacd2c3b8d1b6d990961902783/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/56f89997cccb24bacd2c3b8d1b6d990961902783/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/56f89997cccb24bacd2c3b8d1b6d990961902783/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-03 tex/ms.tex
revert eps3, rename to eps

@@
             for a hypothetical scenario in which an observer can measure a
-            stellar light curve at a single inclination (thin blue curves),
-            at 3 different inclinations (thin orange curves), and at 10 different
+            stellar light curve at a single random inclination (thin blue curves),
+            at 3 random inclinations (thin orange curves), and at 10 random
             inclinations (thin green curves). The thicker curves correspond to
-            the mean shrinkage for each experiment. The inclinations are
-            drawn randomly from an isotropic distribution.
+            the mean shrinkage for each experiment.
             The information content in the light curve of a star observed
@@
 standard deviation $\sigma_\phi$ of the distribution in $\phi$
-(Appendix \ref{sec:lat}).
+(Appendix \ref{sec:lat-transform}).
 The spot longitude is assumed to be uniformly distributed
@@
 stellar inclinations. While modern sampling schemes such as Hamiltonian
-Monte Carlo and Nested Sampling may in principle be able to deal with these
+M0onte Carlo and Nested Sampling may in principle be able to deal with these
 issues, in practice it can be very difficult to obtain convergence in a
@@
 \begin{align}
-    \label{eq:mu_marg}
+    \label{eq:einc}
     \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
@@
     p(i) \, \mathrm{d}i \\
-    \label{eq:cov_marg}
+    \label{eq:Einc}
     \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
-    - \mathbf{1} \mathbf{1}^\top
      & =
@@
 \begin{align}
+    \label{eq:mu_marg}
     \pmb{\mu}(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet)
@@
     \\
+    \label{eq:cov_marg}
     \pmb{\Sigma}(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet)
@@
 \label{sec:starry}
-\xxx{Coming soon.}
+
+\subsection{Basic expression}
+%
+As we mentioned in \S\ref{sec:gp}, the flux $\mathbf{f}$ is a purely
+linear function of the spherical harmonic coefficient vector $\mathbf{y}$:
+%
+\begin{align}
+    \mathbf{f} = \mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y}
+    \quad.
+\end{align}
+%
+Even though this is derived in detail in \citet{Luger2019}, it is useful to
+expand on the computation of the design matrix $\pmb{\mathcal{A}}$ in more
+detail here. Let $\mathbf{a}_k^\top$ denote the $k^{th}$ row of $\pmb{\mathcal{A}}$,
+such that
+%
+\begin{align}
+    \label{eq:Arows}
+    \pmb{\mathcal{A}}
+     & =
+    \setstackgap{L}{1.25\baselineskip}
+    \fixTABwidth{T}
+    \parenMatrixstack{
+        \mathbf{a}_0^\top \\
+        \mathbf{a}_1^\top \\
+        \vdots            \\
+        \mathbf{a}_{K-1}^\top
+    }\quad.
+\end{align}
+%
+The row vector $\mathbf{a}_k^\top$ encodes how the spherical harmonic
+coefficient vector projects onto the $k^\mathrm{th}$ cadence in the flux timeseries, and
+may be computed from
+%
+\begin{align}
+    \label{eq:akT}
+    \mathbf{a}_k^\top = \mathbf{r}^\top \,
+    \mathbf{A_1} \,
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(-i\right) \,
+    \mathbf{R}_{\hat{\mathbf{z}}}\left(\frac{2\pi}{P}t_k\right) \,
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(\frac{\pi}{2}\right) \,
+    \mathbf{y}
+    \quad.
+\end{align}
+%
+To understand the expression above, let us proceed from right to left,
+starting with the spherical harmonic vector $\mathbf{y}$, which we assume
+describes the surface intensity of the star at time $t = 0$
+in a frame where $\hat{\mathbf{x}}$
+points to the right, $\hat{\mathbf{y}}$ points up, and $\hat{\mathbf{z}}$
+points out of the page. The quantity $\mathbf{R}_{\hat{\mathbf{x}}}$
+is a Wigner rotation matrix
+(described in detail in \S\ref{sec:wigner}),
+which in this case rotates the spherical harmonic representation
+of the star by an angle $\nicefrac{\pi}{2}$ counter-clockwise about $\hat{\mathbf{x}}$
+such that the north pole of the star points along $\hat{\mathbf{z}}$. In this
+frame, we apply a second Wigner rotation matrix, $\mathbf{R}_{\hat{\mathbf{z}}}$,
+to rotate the star about $\hat{\mathbf{z}}$ counter-clockwise (i.e., eastward) by an angle
+$\nicefrac{2\pi t_k}{P}$, where $P$ is the rotation period and $t_k$ is the
+time at cadence $t$.
+Next, we rotate the star by a \emph{clockwise} angle of $i$ about $\hat{\mathbf{x}}$,
+where $i$ is the stellar inclination ($i = 0$ corresponding to a pole-on view and
+$i = \nicefrac{\pi}{2}$ corresponding to an edge-on view). With this last rotation,
+we are now in the observer's frame%
+\footnote{In principle, one last rotation could be performed about $\hat{\mathbf{z}}$
+    to orient the projected disk of the star on the plane of the sky; however, the disk-integrated
+    flux is independent of the rotation angle along the plane of the sky
+    (which we refer to as the \emph{obliquity}), so this step is unnecessary.}.
+
+Following \citet{Luger2019}, the next step is to project the representation of
+the star into a more convenient basis for performing the integration over the stellar
+disk. The change-of-basis matrix $\mathbf{A}_1$ \citep[c.f. Appendix~B in][]{Luger2019}
+projects the stellar map into the \emph{polynomial basis}
+\citep[Equation~7 in][]{Luger2019}, comprised of the sequence of monomials in Cartesian
+coordinates
+$\left( 1 \,\, x \,\, z \,\, y \,\, x^2 \,\, xz \,\, xy \,\, yz \,\, y^2 \cdots \right)$
+where $z = \sqrt{1 - x^2 - y^2}$ on the surface of the unit sphere. We
+can now compute the disk-integrated flux by integrating each of the terms in the basis
+over the unit disk, which is straightforward in the polynomial basis; the
+individual terms integrate to simple ratios of Gamma functions. These are then
+assembled into the row vector $\mathbf{r}^\top$, given by Equation~(20) in
+\citet{Luger2019}, which we dot into our expression (and add one) to obtain the
+flux at the $k^\mathrm{th}$ cadence.
+
+\subsection{With limb darkening}
+%
+\xxx{TODO.}
 
@@
 
-\xxx{...}
-
-Although we adopt $a$ and $b$ as the actual inputs to our Gaussian process
+As we mentioned earlier,
+although we adopt $a$ and $b$ as the actual inputs to our Gaussian process
 (i.e., as elements of the input vector $\pmb{\theta}_\bullet$), these
@@
 
+In this section we will compute the first and second moments of the
+inclination distribution (Equations~\ref{eq:einc} and \ref{eq:Einc}),
+which allows us to compute the mean and covariance of the process
+that describes the flux marginalized over all values of the inclination
+(Equations~\ref{eq:mu_marg} and \ref{eq:cov_marg}).
+
+\subsection{Probability density function}
+%
+Similar to the latitude integrals, the process of inclining a star relative to the
+observer (see Equation~\ref{eq:akT}) involves
+rotations by an angle $-i$ about $\hat{\mathbf{x}}$, which
+may be accomplished by choosing
+Euler angles $\alpha = \nicefrac{\pi}{2}$, $\beta = -i$, and
+$\gamma = -\nicefrac{\pi}{2}$, such that
+%
+\begin{align}
+    \mathbf{R}^l_{\hat{\mathbf{x}}}\left(-i\right)
+     & =
+    {\mathbf{U}^l}^{-1} \mathbf{D}^l_{\hat{\mathbf{x}}}\left(-i\right) \mathbf{U}^l
+\end{align}
+%
+with
+\begin{align}
+    \mathbf{D}^l_{\hat{\mathbf{x}}}\left(-i\right)
+     & =
+    \mathbf{D}^l\left(\frac{\pi}{2}, -i, -\frac{\pi}{2}\right)
+    \quad.
+\end{align}
+%
+Since we expect an isotropic distribution of stellar rotation axes (absent
+prior constraints on individual stars), the prior probability density for
+the inclination $i$ is simply
+%
 \begin{align}
     p(i) = \sin i
-    \quad.
-\end{align}
+\end{align}
+%
+for $i \in [0, \frac{\pi}{2}]$.
 
 \subsection{First moment}
-\xxx{Coming soon.}
+The expression for the first moment is
+%
+\begin{proof}{}
+    \mathbf{e}_i
+    & \equiv
+    \int
+    %\mathbf{R}_{\hat{\mathbf{x}}}(-i) \,
+    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
+    \mathbf{e}_y \,
+    p(i) \,
+    \mathrm{d}i
+\end{proof}
+%
+where
+%
+\begin{proof}{}
+    \mathbf{e}_y
+    & \equiv
+    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
+\end{proof}
+%
+is the first moment of the GP in the spherical harmonic basis
+(Equation~\ref{eq:exp_y}). Let us expand the expression above using
+Equation~(\ref{eq:akT})
+
+\begin{align}
+    \label{eq:akT}
+    \mathbf{a}_k^\top = \mathbf{r}^\top \,
+    \mathbf{A_1} \,
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(-i\right) \,
+    \mathbf{R}_{\hat{\mathbf{z}}}\left(\frac{2\pi}{P}t_k\right) \,
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(\frac{\pi}{2}\right) \,
+    \mathbf{y}
+    \quad.
+\end{align}
+
+\xxx{More soon.}
 

### 2020-12-03 starry_process/defaults.py

@@ -14,9 +14,9 @@ defaults = dict(
     baseline_mean=0.0,
     baseline_var=0.0,
     driver="numpy",
-    eps1=1e-12,
-    eps2=1e-9,
-    eps3=1e-12,
+    eps=1e-6,
+    epsy=1e-12,
+    epsy15=1e-9,
     covpts=300,
     log_alpha_max=10,
     log_beta_max=10,

### 2020-12-03 starry_process/sp.py

@@ -127,17 +127,16 @@ class StarryProcess(object):
                 in the header file
                 `starry_process/ops/include/constants.h 
                 <https://github.com/rodluger/starry_process/blob/master/starry_process/ops/include/constants.h>`_.
-            eps1 (float, optional): A small number added to the diagonal of the
+            eps (float, optional): A small number added to the diagonal of the
+                flux covariance matrix when marginalizing over inclination
+                for extra stability. Default is %%defaults["eps"]%%.
+            epsy (float, optional): A small number added to the diagonal of the
                 spherical harmonic covariance matrix for stability.
-                Default is %%defaults["eps1"]%%.
-            eps2 (float, optional): A small number added to terms in the 
+                Default is %%defaults["epsy"]%%.
+            epsy15 (float, optional): A small number added to terms in the 
                 diagonal of the spherical harmonic covariance matrix 
                 above degree ``15``, which become particularly unstable.
-                Default is %%defaults["eps2"]%%.
-            eps3 (float, optional): A small number added to the diagonal of the
-                flux covariance matrix when marginalizing over inclination
-                for extra stability. Default is %%defaults["eps3"]%%.
-
+                Default is %%defaults["epsy15"]%%.
         """
         # Spherical harmonic degree of the process
         self._ydeg = kwargs.get("ydeg", defaults["ydeg"])
@@ -380,7 +379,7 @@ class StarryProcess(object):
         if self._normalized:
             cov = self._flux.cov(t, i, p)
             mean = self._flux.mean(t, i, p)[0]
-            return self._normalize(1.0 + mean, cov)  # cov / (1 + mean) ** 2
+            return self._normalize(1.0 + mean, cov)
         else:
             return self._flux.cov(t, i, p)
 
@@ -416,15 +415,21 @@ class StarryProcess(object):
         )
         return ifelse(tt.gt(z, self._normzmax), Sig * np.inf, normSig)
 
-    def sample(self, t, i=defaults["i"], p=defaults["p"], nsamples=1):
+    def sample(
+        self,
+        t,
+        i=defaults["i"],
+        p=defaults["p"],
+        nsamples=1,
+        eps=defaults["eps"],
+    ):
         """
-        TODO: Samples are noisy because of `eps3`.
 
         """
         if self._marginalize_over_inclination:
             t = cast(t)
             u = self.random.normal((t.shape[0], nsamples))
[…]

### 2020-12-03 tex/gitlinks.tex
ufff; wrote inclination section

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/56f89997cccb24bacd2c3b8d1b6d990961902783/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/56f89997cccb24bacd2c3b8d1b6d990961902783/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/56f89997cccb24bacd2c3b8d1b6d990961902783/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/3fb4d04da2b805c28ab3df4b1bdaf4ee4c772c8c/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/3fb4d04da2b805c28ab3df4b1bdaf4ee4c772c8c/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/3fb4d04da2b805c28ab3df4b1bdaf4ee4c772c8c/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-03 tex/ms.tex
ufff; wrote inclination section

@@
 harmonic basis to the flux basis; it is implicitly
-a function of the stellar inclination $i$, the stellar
+a function of the stellar inclination $I$, the stellar
 rotation period $P$, the stellar limb darkening coefficients, etc.
@@
     \left(
-    i \,\,\,
+    I \,\,\,
     P \,\,\,
@@
     \int
-    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
+    \pmb{\mathcal{A}}(I, \bar{\pmb{\theta}}_\star) \,
     \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
-    p(i) \, \mathrm{d}i \\
+    p(I) \, \mathrm{d}I \\
     \label{eq:Einc}
@@
     \int
-    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
+    \pmb{\mathcal{A}}(I, \bar{\pmb{\theta}}_\star) \,
     \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
-    \pmb{\mathcal{A}}^\top(i, \bar{\pmb{\theta}}_\star)
-    p(i) \, \mathrm{d}i
+    \pmb{\mathcal{A}}^\top(I, \bar{\pmb{\theta}}_\star)
+    p(I) \, \mathrm{d}I
     \quad,
@@
 represented by boldface lowercase letters
-($\mathbf{x}$), and matrices and other linear operators are represented
+($\mathbf{x}$), and matrices are represented
 by boldface capital letters ($\mathbf{X}$). In general, the elements of a vector
 $\mathbf{x}$ are denoted $x_i$ and the elements of a matrix $\mathbf{X}$
-are denoted $X_{i,j}$. The exception to this is if the vector or matrix
+are denoted $X_{i,j}$. Importantly, we make a distinction between
+quantities like $X_{i,j}$ and $\mathbf{X}_{i,j}$: the former is a scalar
+element of a matrix, while the latter is a \emph{matrix}, which is itself
+a component of a higher-dimensional (in this case, 4-dimensional) linear
+operator. Thus, lowercase bold symbols \emph{always} represent vectors, and
+uppercase bold symbols \emph{always} represent matrices.
+
+The exception to our indexing notation is if the vector or matrix
 represents a quantity in the spherical harmonic basis, in which case we use
-\emph{two} indices to represent a vector element, $x^l_m$, and \emph{four}
-indices to represent a matrix element, $X^{l,l'}_{m,m'}$.
+\emph{two} indices to represent a scalar vector element, $x^l_m$, and \emph{four}
+indices to represent a scalar matrix element, $X^{l,l'}_{m,m'}$.
 %
@@
     \mathbf{A_1} \,
-    \mathbf{R}_{\hat{\mathbf{x}}}\left(-i\right) \,
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(-I\right) \,
     \mathbf{R}_{\hat{\mathbf{z}}}\left(\frac{2\pi}{P}t_k\right) \,
-    \mathbf{R}_{\hat{\mathbf{x}}}\left(\frac{\pi}{2}\right) \,
-    \mathbf{y}
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(\frac{\pi}{2}\right)
     \quad.
@@
 time at cadence $t$.
-Next, we rotate the star by a \emph{clockwise} angle of $i$ about $\hat{\mathbf{x}}$,
-where $i$ is the stellar inclination ($i = 0$ corresponding to a pole-on view and
-$i = \nicefrac{\pi}{2}$ corresponding to an edge-on view). With this last rotation,
+Next, we rotate the star by a \emph{clockwise} angle of $I$ about $\hat{\mathbf{x}}$,
+where $I$ is the stellar inclination ($I = 0$ corresponding to a pole-on view and
+$I = \nicefrac{\pi}{2}$ corresponding to an edge-on view). With this last rotation,
 we are now in the observer's frame%
@@
     {\mathbf{U}^l}^{-1}
-    \mathbf{p}^l
+    \mathbf{p}^l_\phi
     \quad,
@@
 \begin{proof}{}
+    \label{eq:Ephi}
     \mathbf{E}_\phi
@@
     \mathbf{R}^l_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{E}^l_r \,
+    \mathbf{E}^{l,l'}_r \,
     {\mathbf{R}^{l'}_{\hat{\mathbf{x}}}}^\top(\phi) \,
@@
     {\mathbf{U}^l}^{-1}
-    \mathbf{P}^l_\phi
-    {{\mathbf{U}^l}^{-1}}^\top
+    \mathbf{P}^{l,l'}_\phi
+    {{\mathbf{U}^{l'}}^{-1}}^\top
 \end{proof}
@@
     \label{eq:Plphi}
-    \mathbf{P}^l_\phi
+    \mathbf{P}^{l,l'}_\phi
      & \equiv
@@
     \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi) \,
-    \bar{\mathbf{E}}^l_r \,
-    {\mathbf{D}^l_{\hat{\mathbf{x}}}}^\top(\phi) \,
+    \bar{\mathbf{E}}^{l,l'}_r \,
+    {\mathbf{D}^{l'}_{\hat{\mathbf{x}}}}^\top(\phi) \,
     p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
@@
     \\
-    \bar{\mathbf{E}}^l_r
+    \bar{\mathbf{E}}^{l,l'}_r
      & \equiv
     \mathbf{U}^l
-    \mathbf{E}^l_r
-    {\mathbf{U}^l}^\top
+    \mathbf{E}^{l,l'}_r
+    {\mathbf{U}^{l'}}^\top
     \quad.
@@
 %
-As before, we may express the solution to the integral $\mathbf{P}^l_\phi$ in
-closed form. Let us write the terms of $\mathbf{P}^l_\phi$ as
-%
-\begin{align}
-    {({P^l_\phi})_{}}_{m,m'}
+As before, we may express the solution to the integral $\mathbf{P}^{l,l'}_\phi$ in
+closed form. Let us write the terms of $\mathbf{P}^{l,l'}_\phi$ as
+%
+\begin{align}
+    {({P^{l,l'}_\phi})_{}}_{m,m'}
      & =
@@
     \sum\limits_{\mu=-l}^l
-    \sum\limits_{\nu=-l}^l
+    \sum\limits_{{\mu'}=-l'}^{l'}
     {({D^l_{\hat{\mathbf{x}}}})_{}}_{m,\mu}(\phi) \,
-    {(\bar{E}^l_r)_{}}_{\mu,\nu} \,
-    {({D^l_{\hat{\mathbf{x}}}})_{}}_{\nu,m'}(\phi) \,
+    {(\bar{E}^{l,l'}_r)_{}}_{\mu,{\mu'}} \,
+    {({D^{l'}_{\hat{\mathbf{x}}}})_{}}_{{\mu'},m'}(\phi) \,
     p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
@@
         \sum\limits_{\mu=-l}^l
-        \sum\limits_{\nu=-l}^l
-        {(\bar{E}^l_r)_{}}_{\mu,\nu} \,
-        \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu + \nu - m')}
+        \sum\limits_{{\mu'}=-l'}^{l'}
+        {(\bar{E}^{l,l'}_r)_{}}_{\mu,{\mu'}} \,
+        \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu + {\mu'} - m')}
         \sum\limits_{i=0}^{2l}
-        \sum\limits_{j=0}^{2l}
+        \sum\limits_{i'=0}^{2l'}
         c_{m,\mu,i}^{l}
-        c_{\nu,m',j}^{l}
+        c_{{\mu'},m',i'}^{l'}
         \,
-        {({Q^l_\phi})_{}}_{i,j}(\pmb{\theta}_{\phi})
+        {({Q^{l,l'}_\phi})_{}}_{i,i'}(\pmb{\theta}_{\phi})
     $}
@@
 \begin{proof}{}
-    {({Q^l_\phi})_{}}_{i,j}(\pmb{\theta}_{\phi})
+    {({Q^{l,l'}_\phi})_{}}_{i,i'}(\pmb{\theta}_{\phi})
     & \equiv
@@
     %
-    \mathrm{sgn}(\sin\phi)^{i+j}
+    \mathrm{sgn}(\sin\phi)^{i+i'}
     \big| \sin\phi \big|
     (\cos\phi)^{\alpha - 1}
-    (1 - \cos\phi)^{2l + \beta - \frac{i+j}{2} - 1}
-    (1 + \cos\phi)^\frac{i+j}{2}
+    (1 - \cos\phi)^{l + l' + \beta - \frac{i+i'}{2} - 1}
+    (1 + \cos\phi)^\frac{i+i'}{2}
     %
@@
         x^{\alpha - 1}
-        (1 - x)^{2l + \beta - \frac{i+j}{2} - 1 }
-        (1 + x)^\frac{i+j}{2}
+        (1 - x)^{l + l' + \beta - \frac{i+i'}{2} - 1 }
+        (1 + x)^\frac{i+i'}{2}
         %
@@
         %
-         & i+j \,\, \mathrm{even}
+         & i+i' \,\, \mathrm{even}
         %
@@
         %
-         & i+j \,\, \mathrm{odd} \quad.
+         & i+i' \,\, \mathrm{odd} \quad.
     \end{cases}
@@
     \label{eq:Qlphiij}
-    {({Q^l_\phi})_{}}_{i,j}(\pmb{\theta}_{\phi})
+    {({Q^{l,l'}_\phi})_{}}_{i,i'}(\pmb{\theta}_{\phi})
     & =
@@
             \begin{cases}
-                \dfrac{\Gamma(\alpha)\Gamma(2l + \beta - \frac{i+j}{2})}{\Gamma(2l + \alpha + \beta - \frac{i+j}{2})}
+                \dfrac{\Gamma(\alpha)\Gamma(l + l' + \beta - \frac{i+i'}{2})}{\Gamma(l + l' + \alpha + \beta - \frac{i+i'}{2})}
                 \,
                 _2F_1\Big(
-                -\frac{i+j}{2},
+                -\frac{i+i'}{2},
                 \alpha;
-                2l + \alpha + \beta - \frac{i+j}{2};
+                l + l' + \alpha + \beta - \frac{i+i'}{2};
                 -1
@@
                 %
-                 & i+j \,\, \mathrm{even}
+                 & i+i' \,\, \mathrm{even}
                 %
@@
                 %
-                 & i+j \,\, \mathrm{odd} \quad.
+                 & i+i' \,\, \mathrm{odd} \quad.
             \end{cases}
@@
 whose terms may easily be computed by upward recursion.
+Since $\mathbf{q}^l_I$ does not depend on any user inputs,
+it may be computed a single time as a pre-processing step
+for efficiency.
 
@@
     \mathbf{R}^l_{\hat{\mathbf{y}}}(\lambda) \,
-    \mathbf{E}^l_\phi \,
+    \mathbf{E}^{l,l'}_\phi \,
     {\mathbf{R}^{l'}_{\hat{\mathbf{y}}}}^\top(\lambda) \,
@@
     {\mathbf{U}^l}^{-1}
-    \mathbf{P}^l_\lambda
-    {{\mathbf{U}^l}^{-1}}^\top
+    \mathbf{P}^{l,l'}_\lambda
+    {{\mathbf{U}^{l'}}^{-1}}^\top
 \end{proof}
@@
     \label{eq:Pllam}
-    \mathbf{P}^l_\lambda
+    \mathbf{P}^{l,l'}_\lambda
      & \equiv
@@
     \mathbf{D}^l_{\hat{\mathbf{x}}}(\lambda) \,
-    \bar{\mathbf{E}}^l_\phi \,
-    {\mathbf{D}^l_{\hat{\mathbf{x}}}}^\top(\lambda) \,
+    \bar{\mathbf{E}}^{l,l'}_\phi \,
+    {\mathbf{D}^{l'}_{\hat{\mathbf{x}}}}^\top(\lambda) \,
     p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
@@
     \\
-    \bar{\mathbf{E}}^l_\phi
+    \bar{\mathbf{E}}^{l,l'}_\phi
      & \equiv
     \mathbf{U}^l
-    \mathbf{E}^l_\phi
-    {\mathbf{U}^l}^\top
+    \mathbf{E}^{l,l'}_\phi
+    {\mathbf{U}^{l'}}^\top
     \quad.
@@
 %
-We then express the terms of $\mathbf{P}^l_\lambda$ as
-%
-\begin{align}
-    {({P^l_\lambda})_{}}_{m,m'}
+We then express the terms of $\mathbf{P}^{l,l'}_\lambda$ as
+%
+\begin{align}
+    {({P^{l,l'}_\lambda})_{}}_{m,m'}
      & =
@@
     \sum\limits_{\mu=-l}^l
-    \sum\limits_{\nu=-l}^l
+    \sum\limits_{{\mu'}=-l'}^{l'}
     ({D^l_{\hat{\mathbf{x}}}})_{m,\mu}(\lambda) \,
-    (\bar{E}^l_\phi)_{\mu,\nu} \,
-    ({D^l_{\hat{\mathbf{x}}}})_{\nu,m'}(\lambda) \,
+    (\bar{E}^{l,l'}_\phi)_{\mu,{\mu'}} \,
+    ({D^{l'}_{\hat{\mathbf{x}}}})_{{\mu'},m'}(\lambda) \,
     p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
@@
     \sum\limits_{\mu=-l}^l
-    \sum\limits_{\nu=-l}^l
-    (\bar{E}^l_\phi)_{\mu,\nu}
+    \sum\limits_{{\mu'}=-l'}^{l'}
+    (\bar{E}^{l,l'}_\phi)_{\mu,{\mu'}}
     \sum\limits_{i=0}^{2l}
-    \sum\limits_{j=0}^{2l}
+    \sum\limits_{i'=0}^{2l'}
     c_{m,\mu,i}^{l}
-    c_{\nu,m',j}^{l}
+    c_{{\mu'},m',i'}^{l'}
     \,
-    (Q^l_\lambda)_{i,j}
+    (Q^{l,l'}_\lambda)_{i,i'}
     \quad,
@@
 \begin{proof}{}
-    (Q^l_\lambda)_{i,j}
+    (Q^{l,l'}_\lambda)_{i,i'}
     & \equiv
@@
     %
-    \mathrm{sgn}(\sin\lambda)^{i+j}
-    (1 - \cos\lambda)^{2l - \frac{i+j}{2}}
-    (1 + \cos\lambda)^\frac{i+j}{2}
+    \mathrm{sgn}(\sin\lambda)^{i+i'}
+    (1 - \cos\lambda)^{l + l' - \frac{i+i'}{2}}
+    (1 + \cos\lambda)^\frac{i+i'}{2}
     %
@@
     \begin{cases}
-        \dfrac{2^{2l} \Gamma \left(\frac{1+i+j}{2}\right)
-            \Gamma \left(2l+\frac{1-i-j}{2}\right)}{\pi  \Gamma (2l+1)}
-        %
-         & i+j \,\, \mathrm{even}
+        \dfrac{2^{l+l'} \Gamma \left(\frac{1+i+i'}{2}\right)
+            \Gamma \left(l+l'+\frac{1-i-i'}{2}\right)}{\pi  \Gamma (l+l'+1)}
+        %
+         & i+i' \,\, \mathrm{even}
         %
@@
         %
-         & i+j \,\, \mathrm{odd} \quad,
+         & i+i' \,\, \mathrm{odd} \quad,
     \end{cases}
@@
 %
-whose terms may again be computed by upward recursion.
+whose terms may again be computed by upward recursion in a single
+pre-processing step.
 
@@
 observer (see Equation~\ref{eq:akT}) involves
-rotations by an angle $-i$ about $\hat{\mathbf{x}}$, which
+rotations by an angle $-I$ about $\hat{\mathbf{x}}$, which
 may be accomplished by choosing
-Euler angles $\alpha = \nicefrac{\pi}{2}$, $\beta = -i$, and
+Euler angles $\alpha = \nicefrac{\pi}{2}$, $\beta = -I$, and
 $\gamma = -\nicefrac{\pi}{2}$, such that
@@
 \begin{align}
-    \mathbf{R}^l_{\hat{\mathbf{x}}}\left(-i\right)
-     & =
-    {\mathbf{U}^l}^{-1} \mathbf{D}^l_{\hat{\mathbf{x}}}\left(-i\right) \mathbf{U}^l
+    \mathbf{R}^l_{\hat{\mathbf{x}}}\left(-I\right)
+     & =
+    {\mathbf{U}^l}^{-1} \mathbf{D}^l_{\hat{\mathbf{x}}}\left(-I\right) \mathbf{U}^l
 \end{align}
@@
 \begin{align}
-    \mathbf{D}^l_{\hat{\mathbf{x}}}\left(-i\right)
-     & =
-    \mathbf{D}^l\left(\frac{\pi}{2}, -i, -\frac{\pi}{2}\right)
+    \mathbf{D}^l_{\hat{\mathbf{x}}}\left(-I\right)
+     & =
+    \mathbf{D}^l\left(\frac{\pi}{2}, -I, -\frac{\pi}{2}\right)
     \quad.
@@
 prior constraints on individual stars), the prior probability density for
-the inclination $i$ is simply
-%
-\begin{align}
-    p(i) = \sin i
-\end{align}
-%
-for $i \in [0, \frac{\pi}{2}]$.
+the inclination $I$ is simply
+%
+\begin{align}
+    p(I) = \sin I
+\end{align}
+%
+for $I \in [0, \frac{\pi}{2}]$.
 
 \subsection{First moment}
+\label{sec:inc-mom1}
 The expression for the first moment is
@@
 \begin{proof}{}
-    \mathbf{e}_i
-    & \equiv
-    \int
-    %\mathbf{R}_{\hat{\mathbf{x}}}(-i) \,
-    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
+    \mathbf{e}_I
+    & \equiv
+    \int
+    \pmb{\mathcal{A}}(I, \bar{\pmb{\theta}}_\star) \,
     \mathbf{e}_y \,
-    p(i) \,
-    \mathrm{d}i
+    p(I) \,
+    \mathrm{d}I
 \end{proof}
@@
 is the first moment of the GP in the spherical harmonic basis
-(Equation~\ref{eq:exp_y}). Let us expand the expression above using
-Equation~(\ref{eq:akT})
-
-\begin{align}
-    \label{eq:akT}
-    \mathbf{a}_k^\top = \mathbf{r}^\top \,
+(Equation~\ref{eq:exp_y}). We can use Equations~(\ref{eq:Arows})
+and (\ref{eq:akT})
+to express the element at index $k$ (corresponding to the mean of the
+GP at time $t = t_k$) as
+%
+\begin{align}
+    \left(e_I\right)_k
+     & =
+    \int
+    \mathbf{a}_k^\top
+    \mathbf{e}_y \,
+    p(I) \,
+    \mathrm{d}I
+    \nonumber \\
+     & =
+    \mathbf{r}^\top \,
     \mathbf{A_1} \,
-    \mathbf{R}_{\hat{\mathbf{x}}}\left(-i\right) \,
+    \int
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(-I\right) \,
+    \,
+    \left(\mathbf{e}_{y'}\right)_k \,
+    p(I) \,
+    \mathrm{d}I
+    \quad,
+\end{align}
+%
+where we define
+%
+\begin{align}
+    \left(\mathbf{e}_{y'}\right)_k
+     & \equiv
     \mathbf{R}_{\hat{\mathbf{z}}}\left(\frac{2\pi}{P}t_k\right) \,
     \mathbf{R}_{\hat{\mathbf{x}}}\left(\frac{\pi}{2}\right) \,
-    \mathbf{y}
+    \mathbf{e}_y
+\end{align}
+%
+as the expectation of $\mathbf{y}$ in the polar frame at time $t = t_k$.
+%
+At this point, it is convenient to invoke the fact that our GP
+is longitudinally isotropic: there is no preferred longitude on
+the surface of the star, or, equivalently, no preferred phase
+in the light curve. The rotation about $\hat{\mathbf{z}}$ (i.e., the
+rotational axis of the star) therefore cannot change the expectation
+of $\mathbf{y}$, so
+%
+\begin{proof}{}
+    \left(\mathbf{e}_{y'}\right)_k
+    &=
+    \left(\mathbf{e}_{y'}\right)_0
+    \nonumber \\
+    &=
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(\frac{\pi}{2}\right) \,
+    \mathbf{e}_y
+    \nonumber \\
+    &\equiv \mathbf{e}_{y'}
     \quad.
-\end{align}
-
-\xxx{More soon.}
+\end{proof}
+%
+We therefore have
+%
+\begin{proof}{}
+    \mathbf{e}_I
+    & \equiv
+    e_I \, \mathbf{1}
+\end{proof}
+%
+where
+%
+\begin{align}
+    e_I
+     & =
+    \mathbf{r}^\top \,
+    \mathbf{A_1} \,
+    \mathbf{e}_{y''}
+\end{align}
+%
+where
+%
+\begin{align}
+    \mathbf{e}_{y''}
+     & =
+    \int
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(-I\right) \,
+    \,
+    \mathbf{e}_{y'} \,
+    p(I) \,
+    \mathrm{d}I
+    \nonumber \\
+    %
+     & =
+    %
+    \left(
+    \mathbf{e}_{y''}^0
+    \,\,\,
+    \mathbf{e}_{y''}^1
+    \,\,\,
+    \mathbf{e}_{y''}^2
+    \,\,\,
+    \cdots
+    \,\,\,
+    \mathbf{e}_{y''}^{l_{\mathrm{max}}}
+    \right)^\top
+\end{align}
+%
+is the expectation of $\mathbf{y}$ in the observer's frame,
+and as before we explicitly separate it out by spherical harmonic degree.
+As in \S\ref{sec:lat}, we may write
+%
+\begin{proof}{}
+    \mathbf{e}_{y''}^l
+    & =
+    \int
+    \mathbf{R}^l_{\hat{\mathbf{x}}}(-I) \,
+    \mathbf{e}^l_{y'} \,
+    p(I) \,
+    \mathrm{d}I
+    \nonumber \\
+    %
+    & =
+    %
+    {\mathbf{U}^l}^{-1}
+    \mathbf{p}^l_I
+    \quad,
+\end{proof}
+%
+and we define
+%
+\begin{align}
+    \mathbf{p}^l_I
+     & \equiv
+    \int
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(-I) \,
+    \bar{\mathbf{e}}^l_{y'} \,
+    p(I) \,
+    \mathrm{d}I
+    \\
+    \bar{\mathbf{e}}^l_{y'}
+     & \equiv
+    \mathbf{U}^l
+    \mathbf{e}^l_{y'}
+    \quad.
+\end{align}
+%
+The integral $\mathbf{p}^l_I$ defined above has a closed-form solution.
+To show this, we write the terms of $\mathbf{p}^l_I$ as
+%
+\begin{align}
+    {({p^l_I})_{}}_m
+     & =
+    \int
+    \sum\limits_{\mu=-l}^l
+    {({D^l_{\hat{\mathbf{x}}}})_{}}_{m,\mu}(-I) \,
+    {({\bar{e}}^l_{y'})_{}}_{\mu} \,
+    p(I) \,
+    \mathrm{d}I
+    %
+    \nonumber \\[0.5em]
+    %
+     & =
+    \sum\limits_{\mu=-l}^l
+    {({\bar{e}}^l_{y'})_{}}_{\mu}
+    \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu)}
+    \sum\limits_{i=0}^{2l} c_{m,\mu,i}^{l}
+    \,
+    {({q^l_I})_{}}_i
+    \quad,
+\end{align}
+%
+where
+%
+\begin{proof}{}
+    {({q^l_I})_{}}_i
+    & \equiv
+    \int_{0}^{\frac{\pi}{2}}
+    %
+    (-1)^{i}
+    (1 - \cos I)^{\frac{2l - i}{2}}
+    (1 + \cos I)^\frac{i}{2}
+    \sin I
+    %
+    \,
+    \mathrm{d}I
+    \nonumber \\[0.5em]
+    %
+    & =
+    %
+    (-1)^{i + 1}
+    \displaystyle\int_{0}^{1}
+    %
+    (1 - x)^{\frac{2l - i}{2}}
+    (1 + x)^\frac{i}{2}
+    \,
+    \mathrm{d}x
+    \nonumber \\[0.5em]
+    &=
+    \frac{(-1)^{i+1}}{l-\frac{i}{2}+1}
+    \,
+    {_2F_1}\left(
+    1, -\frac{i}{2}; 2 + l - \frac{i}{2}; -1
+    \right)
+    \quad,
+\end{proof}
+%
+which may easily be computed recursively.
+As with the longitude integrals, the vector
+$\mathbf{q}^l_I$ need only be computed a single
+time as a pre-processing step, as it does not
+depend on any user inputs.
 
 \subsection{Second moment}
-\xxx{Coming soon.}
+%
+The expression for the second moment is
+%
+\begin{proof}{}
+    \mathbf{E}_I
+    & \equiv
+    \int
+    \pmb{\mathcal{A}}(I, \bar{\pmb{\theta}}_\star) \,
+    \mathbf{E}_y \,
+    \pmb{\mathcal{A}}^\top(I, \bar{\pmb{\theta}}_\star) \,
+    p(I) \,
+    \mathrm{d}I
+\end{proof}
+%
+where
+%
+\begin{proof}{}
+    \mathbf{E}_y
+    & \equiv
+    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
+\end{proof}
+%
+is the second moment of the GP in the spherical harmonic basis
+(Equation~\ref{eq:exp_yy}). We can use Equations~(\ref{eq:Arows})
+and (\ref{eq:akT})
+to express the element at index $k, k'$ (corresponding to the covariance of the
+GP between times $t = t_k$ and $t' = t_{k'}$) as
+%
+\begin{align}
+    \label{eq:Eikkp}
+    \left(E_I\right)_{k, k'}
+     & =
+    \int
+    \mathbf{a}_k^\top
+    \mathbf{E}_y \,
+    \mathbf{a}_{k'}
+    p(I) \,
+    \mathrm{d}I
+    \nonumber \\
+     & =
+    \mathbf{r}^\top \,
+    \mathbf{A_1} \,
+    \left(\mathbf{E}_{y''}\right)_{k,k'}
+    \mathbf{A_1}^\top \,
+    \mathbf{r}
+    \quad,
+\end{align}
+%
+where
+%
+\begin{align}
+    \label{eq:Eyppkkp}
+    \left(\mathbf{E}_{y''}\right)_{k,k'}
+     & =
+    \int
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(-I\right) \,
+    \left(\mathbf{E}_{y'}\right)_{k,k'} \,
+    \mathbf{R}_{\hat{\mathbf{x}}}^\top\left(-I\right) \,
+    p(I) \,
+    \mathrm{d}I
+\end{align}
+%
+is the expectation of $\mathbf{y}\,\mathbf{y}^\top$ in the
+observer's frame at times $t = t_k$ and $t' = t_{k'}$
+and
+%
+\begin{align}
+    \left(\mathbf{E}_{y'}\right)_{k,k'}
+     & \equiv
+    \mathbf{R}_{\hat{\mathbf{z}}}\left(\frac{2\pi}{P}t_k\right) \,
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(\frac{\pi}{2}\right) \,
+    \mathbf{E}_y \,
+    \mathbf{R}_{\hat{\mathbf{x}}}^\top\left(\frac{\pi}{2}\right) \,
+    \mathbf{R}_{\hat{\mathbf{z}}}^\top\left(\frac{2\pi}{P}t_{k'}\right)
+\end{align}
+%
+is the expectation of $\mathbf{y}\,\mathbf{y}^\top$ in the
+polar frame at times $t = t_k$ and $t' = t_{k'}$.
+%
+The rest of the computation follows what we did in \S\ref{sec:lat},
+except that the number of operations required to compute $\mathbf{E}_I$
+is a factor of $K^2$ larger than in the computation of expectations
+like $\mathbf{E}_\phi$ (Equation~\ref{eq:Ephi}). That is because we
+must compute the integral of all terms of a matrix
+for \emph{each} of the $K^2$ elements of $\mathbf{E}_I$.
+We discuss in \S\ref{sec:inc-speedup} below strategies that can drastically
+improve the computational scaling of marginalizing over the inclination.
+
+Let us write Equation~(\ref{eq:Eyppkkp}) in terms of its spherical harmonic
+components:
+%
+\begin{proof}{}
+    \resizebox{.85\hsize}{!}{$
+            \left(\mathbf{E}_{y''}\right)_{k,k'}
+            =
+            \setstackgap{L}{1.25\baselineskip}
+            \fixTABwidth{T}
+            \parenMatrixstack{
+                \left(\mathbf{E}_{y''}^{0,0}\right)_{k,k'} & \left(\mathbf{E}_{y''}^{0,1}\right)_{k,k'} & \left(\mathbf{E}_{y''}^{0,2}\right)_{k,k'} & \cdots & \left(\mathbf{E}_{y''}^{0,l_{\mathrm{max}}}\right)_{k,k'} \\
+                \left(\mathbf{E}_{y''}^{1,0}\right)_{k,k'} & \left(\mathbf{E}_{y''}^{1,1}\right)_{k,k'} & \left(\mathbf{E}_{y''}^{1,2}\right)_{k,k'} & \cdots & \left(\mathbf{E}_{y''}^{1,l_{\mathrm{max}}}\right)_{k,k'} \\
+                \left(\mathbf{E}_{y''}^{2,0}\right)_{k,k'} & \left(\mathbf{E}_{y''}^{2,1}\right)_{k,k'} & \left(\mathbf{E}_{y''}^{2,2}\right)_{k,k'} & \cdots & \left(\mathbf{E}_{y''}^{2,l_{\mathrm{max}}}\right)_{k,k'} \\
+                \vdots & \vdots & \vdots & \ddots & \vdots \\
+                \left(\mathbf{E}_{y''}^{l_{\mathrm{max}},0}\right)_{k,k'} & \left(\mathbf{E}_{y''}^{l_{\mathrm{max}},1}\right)_{k,k'} & \left(\mathbf{E}_{y''}^{l_{\mathrm{max}},2}\right)_{k,k'} & \cdots & \left(\mathbf{E}_{y''}^{l_{\mathrm{max}},l_{\mathrm{max}}}\right)_{k,k'}
+            }
+        $}
+    \quad,
+\end{proof}
+%
+where
+%
+\begin{proof}{}
+    \left(\mathbf{E}_{y''}^{l,l'}\right)_{k,k'}
+    & =
+    \int
+    \mathbf{R}^l_{\hat{\mathbf{x}}}(-I) \,
+    \left(\mathbf{E}_{y'}^{l,l'}\right)_{k,k'} \,
+    {\mathbf{R}^{l'}_{\hat{\mathbf{x}}}}^\top(-I) \,
+    p(I) \,
+    \mathrm{d}I
+    \nonumber \\
+    %
+    & =
+    %
+    {\mathbf{U}^l}^{-1}
+    \left(\mathbf{P}^{l,l'}_I\right)_{k,k'}
+    {{\mathbf{U}^{l'}}^{-1}}^\top
+\end{proof}
+%
+and we define
+%
+\begin{align}
+    \label{eq:PllpIkkp}
+    \left(\mathbf{P}^{l,l'}_I\right)_{k,k'}
+     & \equiv
+    \int
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(-I) \,
+    \left(\bar{\mathbf{E}}^{l,l'}_{y'}\right)_{k,k'} \,
+    {\mathbf{D}^{l'}_{\hat{\mathbf{x}}}}^\top(-I) \,
+    p(I) \,
+    \mathrm{d}I
+    \\
+    \left(\bar{\mathbf{E}}^{l,l'}_{y'}\right)_{k,k'}
+     & \equiv
+    \mathbf{U}^l
+    \left(\mathbf{E}^{l,l'}_{y'}\right)_{k,k'}
+    {\mathbf{U}^{l'}}^\top
+    \quad.
+\end{align}
+%
+As before, we may express the solution to the integral in Equation~(\ref{eq:PllpIkkp}) in
+closed form. Let us write its terms as
+%
+\begin{align}
+    \label{eq:nightmare}
+    \left[{\left({P^{l,l'}_I}\right)_{}}_{k,k'}\right]_{m,m'}
+     & =
+    \int
+    \sum\limits_{\mu=-l}^l
+    \sum\limits_{{\mu'}=-l}^l
+    {({D^l_{\hat{\mathbf{x}}}})_{}}_{m,\mu}(-I) \,
+    \left[{\left({\bar{E}^{l,l'}_{y'}}\right)_{}}_{k,k'}\right]_{\mu,\mu'}
+    {({D^{l'}_{\hat{\mathbf{x}}}})_{}}_{{\mu'},m'}(-I) \,
+    p(I) \,
+    \mathrm{d}I
+    %
+    \nonumber \\[0.5em]
+    %
+     & =
+    \resizebox{.75\hsize}{!}{$
+        \sum\limits_{\mu=-l}^l
+        \sum\limits_{{\mu'}=-l'}^{l'}
+        \left[{\left({\bar{E}^{l,l'}_{y'}}\right)_{}}_{k,k'}\right]_{\mu,\mu'}
+        \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu + {\mu'} - m')}
+        \sum\limits_{i=0}^{2l}
+        \sum\limits_{i'=0}^{2l'}
+        c_{m,\mu,i}^{l}
+        c_{{\mu'},m',i'}^{l'}
+        \,
+        {({Q^{l,l'}_I})_{}}_{i,i'}
+    $}
+    \quad,
+\end{align}
+%
+where, similarly to before,
+%
+\begin{proof}{}
+    \label{eq:QllpIiip}
+    {({Q^{l,l'}_I})_{}}_{i,i'}
+    & \equiv
+    \int_{0}^{\frac{\pi}{2}}
+    %
+    (-1)^{i+i'}
+    (1 - \cos\phi)^{l + l' - \frac{i+i'}{2}}
+    (1 + \cos\phi)^\frac{i+i'}{2}
+    \sin I
+    %
+    \,
+    \mathrm{d}I
+    \nonumber \\[0.5em]
+    %
+    & =
+    %
+    (-1)^{i+i'+1}
+    \int_{0}^{\frac{\pi}{2}}
+    %
+    (1 - x)^{l + l' - \frac{i+i'}{2}}
+    (1 + x)^\frac{i+i'}{2}
+    %
+    \,
+    \mathrm{d}I
+    \nonumber \\[0.5em]
+    %
+    &=
+    %
+    \frac{(-1)^{i+i'+1}}{l+l'-\frac{i+i'}{2}+1}
+    \,
+    {_2F_1}\left(
+    1, -\frac{i+i'}{2}; 2 + l + l' - \frac{i+i'}{2}; -1
+    \right)
+    \quad,
+\end{proof}
+%
+which may again be computed recursively in a pre-processing step.
+
+\subsection{Speeding up the computation}
+\label{sec:speedup}
+%
+The expressions in the previous section are a bit of a nightmare,
+particularly because of the dimensionality of some of the linear
+operators involved. The complexity of the expressions is due to
+the fact that the second moment of the spherical harmonic vector
+projected onto the sky (Equation~\ref{eq:Eyppkkp}) is time-dependent:
+it changes as the star rotates. Computing the second moment of the
+\emph{flux} requires computing the outer product of this tensor
+with itself, leading to expressions like those in
+Equation~(\ref{eq:nightmare}). In addition to being cumbersome to
+evaluate, the full second moment matrix $\mathbf{E}_I$
+(and hence the flux covariance matrix) is costly to compute.
+It is helpful that Equation~(\ref{eq:QllpIiip}) does not depend
+on any user inputs and thus may be pre-computed, but even still
+we require evaluating the four nested sums
+in Equation~(\ref{eq:Eyppkkp})
+$\mathcal{O}(l_\mathrm{max}^3)$ times
+\emph{for each entry} in the $K \times K$ matrix $\mathbf{E}_I$.
+
+Fortunately, the inner two sums in Equation~(\ref{eq:Eyppkkp})
+do not depend on user inputs, so those may be pre-computed,
+and Equation~(\ref{eq:Eyppkkp}) may be cast as a straightforward
+matrix dot product. However, in practice we find it helpful to
+take advantage of the phase independence (i.e., stationarity)
+of the covariance of our GP, as we did in \S\ref{sec:inc-mom1}:
+any two entries $(E_I)_{k,k'}$ and $(E_I)_{j,j'}$ are
+the same if $t_{k} - t_{k'} = t_{j} - t_{j'}$. If the data
+happen to be evenly sampled, such that the time difference between
+adjacent cadences is constant, then we need only compute the
+covariance at a total of $K$ points (as opposed to $K^2$), as the
+covariance is a circulant matrix which is fully specified by a
+single vector of length $K$.
+
+In the more general case where the data are not evenly sampled,
+we may still evaluate the covariance at a fixed number of points
+$K' < K^2$ and approximate the full covariance matrix via
+interpolation. As long as the data are roughly evenly sampled,
+as is the case with \emph{Kepler} or \emph{TESS} light curves,
+this approximation leads to negligible error when $K' \approx K$,
+affording the same $\mathcal{O}(K)$ computational savings.
+Note that even in the case where the flux is normalized
+(c.f. \S\ref{sec:baseline}), the non-stationary correction to the
+covariance is applied \emph{after} the step where we marginalize
+over the inclination, so this approach is still valid.
 

### 2020-12-03 tex/gitlinks.tex
misc tweaks

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/3fb4d04da2b805c28ab3df4b1bdaf4ee4c772c8c/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/3fb4d04da2b805c28ab3df4b1bdaf4ee4c772c8c/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/3fb4d04da2b805c28ab3df4b1bdaf4ee4c772c8c/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/3fb283601c2627f1ce16864ad46a6aad5a4438cc/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/3fb283601c2627f1ce16864ad46a6aad5a4438cc/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/3fb283601c2627f1ce16864ad46a6aad5a4438cc/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-03 tex/ms.tex
misc tweaks

@@
 stellar inclinations. While modern sampling schemes such as Hamiltonian
-M0onte Carlo and Nested Sampling may in principle be able to deal with these
+Monte Carlo and Nested Sampling may in principle be able to deal with these
 issues, in practice it can be very difficult to obtain convergence in a
@@
 %
+\xxx{FIX THE NOTATION IN THIS SECTION}
+%
 \begin{align}
@@
      & =
+    \mathrm{E} \Big[ \mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+    \nonumber \\[0.5em]
+     & =
+    \mathbf{1} + FOO
+\end{align}
+%
+where
+%
+\begin{align}
+    FOO
+     & \equiv
+    \mathrm{E} \Big[ \pmb{\mathcal{A}} \, \mathbf{y} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+    \nonumber \\[0.5em]
+     & =
+    \mathbf{1} +
     \int
@@
     \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
-    p(I) \, \mathrm{d}I \\
+    p(I) \, \mathrm{d}I
+\end{align}
+%
+and
+%
+\begin{align}
     \label{eq:Einc}
@@
      & =
+    \mathrm{E} \Big[ \left(\mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y}\right) \left(\mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y}\right)^\top  \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+    \nonumber \\[0.5em]
+     & =
+    \mathbf{1} \, \mathbf{1}^\top +
+    2 \sum FOO +
+    BAR
+\end{align}
+%
+where
+%
+\begin{align}
+    BAR
+     & \equiv
     \int
@@
     p(I) \, \mathrm{d}I
-    \quad,
-\end{align}
-%
-where the expectations inside the integrals are given by
+    \quad.
+\end{align}
+%
+The expectations inside the integrals are given by
 Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), respectively, and

### 2020-12-04 tex/gitlinks.tex
better inclination notation

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/3fb283601c2627f1ce16864ad46a6aad5a4438cc/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/3fb283601c2627f1ce16864ad46a6aad5a4438cc/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/3fb283601c2627f1ce16864ad46a6aad5a4438cc/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/966f526cbc69956e8beac5c550a653955342aa28/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/966f526cbc69956e8beac5c550a653955342aa28/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/966f526cbc69956e8beac5c550a653955342aa28/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-04 tex/ms.tex
better inclination notation

@@
 the problem: we can explicitly marginalize over the stellar inclination.
-That is, we can write the first two moments of our flux GP as
-%
-\xxx{FIX THE NOTATION IN THIS SECTION}
-%
-\begin{align}
-    \label{eq:einc}
-    \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
-     & =
-    \mathrm{E} \Big[ \mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
-    \nonumber \\[0.5em]
-     & =
-    \mathbf{1} + FOO
-\end{align}
-%
-where
-%
-\begin{align}
-    FOO
-     & \equiv
-    \mathrm{E} \Big[ \pmb{\mathcal{A}} \, \mathbf{y} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
-    \nonumber \\[0.5em]
-     & =
-    \mathbf{1} +
-    \int
-    \pmb{\mathcal{A}}(I, \bar{\pmb{\theta}}_\star) \,
-    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
-    p(I) \, \mathrm{d}I
-\end{align}
-%
-and
-%
-\begin{align}
-    \label{eq:Einc}
-    \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
-     & =
-    \mathrm{E} \Big[ \left(\mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y}\right) \left(\mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y}\right)^\top  \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
-    \nonumber \\[0.5em]
-     & =
-    \mathbf{1} \, \mathbf{1}^\top +
-    2 \sum FOO +
-    BAR
-\end{align}
-%
-where
-%
-\begin{align}
-    BAR
-     & \equiv
-    \int
-    \pmb{\mathcal{A}}(I, \bar{\pmb{\theta}}_\star) \,
-    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
-    \pmb{\mathcal{A}}^\top(I, \bar{\pmb{\theta}}_\star)
-    p(I) \, \mathrm{d}I
-    \quad.
-\end{align}
-%
-The expectations inside the integrals are given by
-Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), respectively, and
-$\bar{\pmb{\theta}}_\star$ are all the stellar parameters \emph{except}
-the inclination, i.e.,
+Let
 %
@@
 %
-where perhaps $P$ is fixed for each star and the limb darkening coefficients
+be the vector of all stellar parameters \emph{except}
+the inclination, where perhaps $P$ is fixed for each star and the limb darkening coefficients
 are shared among all stars, as described above.
-%
-Thus, if we are able to perform the integrals in Equations~(\ref{eq:mu_marg})
-and (\ref{eq:cov_marg}), we can \emph{dramatically} reduce the number of
+Then we can write the first two moments of our flux GP as
+%
+\begin{align}
+    \label{eq:einc}
+    \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+     & =
+    \mathrm{E} \Big[ \mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+    \nonumber \\[0.5em]
+     & =
+    \mathbf{1} + \mathbf{e}_I
+\end{align}
+%
+where we define the first moment integral
+%
+\begin{align}
+    \label{eq:eI}
+    \mathbf{e}_I
+     & \equiv
+    \mathrm{E} \Big[ \pmb{\mathcal{A}} \, \mathbf{y} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+    \nonumber \\[0.5em]
+     & =
+    \int
+    \pmb{\mathcal{A}}(I, \bar{\pmb{\theta}}_\star) \,
+    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
+    p(I) \, \mathrm{d}I
+\end{align}
+%
+and
+%
+\begin{align}
+    \label{eq:Einc}
+    \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+     & =
+    \mathrm{E} \Big[ \left(\mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y}\right) \left(\mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y}\right)^\top  \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+    \nonumber \\[0.5em]
+     & =
+    \mathbf{1} \, \mathbf{1}^\top +
+    \mathbf{1} \, \mathbf{e}_I^\top +
+    \mathbf{e}_I \, \mathbf{1}^\top +
+    \mathbf{E}_I
+\end{align}
+%
+where we define the second moment integral
+%
+\begin{align}
+    \label{eq:EI}
+    \mathbf{E}_I
+     & \equiv
+    \int
+    \pmb{\mathcal{A}}(I, \bar{\pmb{\theta}}_\star) \,
+    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
+    \pmb{\mathcal{A}}^\top(I, \bar{\pmb{\theta}}_\star)
+    p(I) \, \mathrm{d}I
+    \quad.
+\end{align}
+%
+The expectations inside the integrals in the expressions for
+$\mathbf{e}_I$ and $\mathbf{E}_I$
+are given by
+Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), respectively.
+%
+If we are able to perform the integrals in those expressions,
+we can \emph{dramatically} reduce the number of
 parameters in our ensemble problem.
@@
 inclinations are distributed isotropically, these integrals
-do in fact have closed-form solutions,
-since they are (again) integrals over Wigner matrices.
+do in fact have closed-form solutions.
 %
@@
      & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
-    \\
+    \nonumber                                                                                                                                                                                                                                       \\
+     & = \mathbf{1} + \mathbf{e}_I
+    \\[0.5em]
     \label{eq:cov_marg}
@@
      & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet) \pmb{\mu}^\top(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet)
+    \nonumber
+    \\
+     & =
+    \mathbf{E}_I
+    -
+    \mathbf{e}_I \,
+    \mathbf{e}_I^\top
     \quad.
@@
 
-In this section we will compute the first and second moments of the
-inclination distribution (Equations~\ref{eq:einc} and \ref{eq:Einc}),
-which allows us to compute the mean and covariance of the process
+In this section we will compute the first and second moment integrals of the
+inclination distribution (Equations~\ref{eq:eI} and \ref{eq:EI}),
+which allow us to compute the mean and covariance of the process
 that describes the flux marginalized over all values of the inclination
@@
 \emph{flux} requires computing the outer product of this tensor
-with itself, leading to expressions like those in
+with itself, leading to multi-indexed quantities like those in
 Equation~(\ref{eq:nightmare}). In addition to being cumbersome to
@@
 and Equation~(\ref{eq:Eyppkkp}) may be cast as a straightforward
-matrix dot product. However, in practice we find it helpful to
+matrix dot product. In practice we also find it helpful to
 take advantage of the phase independence (i.e., stationarity)

### 2020-12-04 tex/gitlinks.tex
more text

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/966f526cbc69956e8beac5c550a653955342aa28/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/966f526cbc69956e8beac5c550a653955342aa28/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/966f526cbc69956e8beac5c550a653955342aa28/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/64ae7e87c3615ea94e611dbe98f372be309bb3d6/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/64ae7e87c3615ea94e611dbe98f372be309bb3d6/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/64ae7e87c3615ea94e611dbe98f372be309bb3d6/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-04 tex/ms.tex
more text

@@
 
-% \begin{abstract}
-%     Abstract here.
-%     %
-%     \href{https://github.com/rodluger/starry_process}{\color{linkcolor}\faGithub}
-% \end{abstract}
+\begin{abstract}
+    The use of Gaussian processes (GPs) in stellar light curve modeling has recently
+    become almost ubiquitous, given their ease of use, flexibility, and
+    computational efficiency. GPs excel in particular at marginalization
+    of the stellar signal in cases where the stellar variability is treated
+    as a nuisance, such as in exoplanet transit modeling.
+    However, in cases where the stellar signal is of primary interest,
+    GPs have thus far been less useful, as it is often difficult to
+    relate their hyperparameters to the physical properties of interest, such
+    as the size, contrast, and latitudinal distribution of star spots.
+    Instead, it is common practice to explicitly model the effect
+    of individual star spots on the light curve and attempt to infer their
+    properties via optimization or posterior inference, a process that is
+    ill-posed, ridden by degeneracies, and computationally intractable when
+    applied to stars with more than a few spots and/or to ensembles of many
+    light curves.
+    %
+    In this paper, we derive a closed-form expression for the
+    mean and covariance of the Gaussian process that best describes
+    the light curve of a rotating star with a given distribution of
+    star spot sizes, contrasts, and latitudes.
+    %
+    We show that, when employed in an inference setting, our GP is unbiased,
+    allowing one to correctly infer physical parameters of interest from one
+    or more stellar light curves, including
+    the typical radii and the mean and variance of the latitude
+    distribution of star spots.
+    %
+    Our GP has far-ranging implications for understanding the variability and
+    magnetic activity of stars from both light curves and radial velocity (RV)
+    measurements, as well as for modeling correlated noise in both transiting
+    and RV exoplanet searches.
+    %
+    Our implementation is efficient, user-friendly, and open source, available
+    as the \Python package \starryprocess.
+    \href{https://github.com/rodluger/starry_process}{\color{linkcolor}\faGithub}
+\end{abstract}
 
@@
 \begin{align}
+    \label{eq:thetaspot}
     \pmb{\theta}_\bullet
@@
 \begin{align}
+    \label{eq:thetastarbar}
     \bar{\pmb{\theta}}_\star = \left(
@@
      & = \mathbf{1} + \mathbf{e}_I
+    \nonumber                                                                                                                                                                                                                                       \\
+     & = \mu \, \mathbf{1}
     \\[0.5em]
@@
     \mathbf{e}_I^\top
-    \quad.
-\end{align}
-%
-and use these to compute the marginal likelihood instead.
+\end{align}
+%
+and use these to compute the marginal likelihood instead. Note that in the
+expression for the mean we made explicit use of the fact that the longitudinal
+isotropy of the process results in $\mathbf{e}_I = e_I \, \mathbf{1}$ being
+constant (see Appendix~\ref{sec:inc-mom1}), so the process mean is likewise
+constant, with $\mu = 1 + e_I$.
 
@@
 use our GP to model normalized light curves. For definiteness, let us
-consider only mean-normalized light curves.
+consider only mean-normalized light curves.%
+\footnote{In practice, the expressions derived below also work well
+    for median-normalized light curves, since the distribution of the GP sample median
+    is usually close to the distribution of sample mean.}%
 Since we are dividing the
@@
 \begin{align}
+    \label{eq:ftilde}
     \tilde{\mathbf{f}\hspace{0.2em}} & = \frac{\mathbf{f}}{\left<f\right>}
@@
 
-\subsection{Other stuff}
-\label{sec:baseline}
-
-Figure~\ref{fig:nongaussianity}
+\subsection{Summary}
+\label{sec:summary}
+%
+As the computation of the GP relies on many interdependent equations
+scattered throughout the previous sections and the Appendix, it
+is useful to summarize the procedure for the case where we marginalize
+over the inclination (\S\ref{sec:inclination}) and the light curves are
+normalized to their means (\S\ref{sec:baseline}), which is likely to be the primary
+use case for our algorithm.
+
+We model the mean-normalized flux $\tilde{\mathbf{f}\hspace{0.2em}}$
+(Equation~\ref{eq:ftilde}) as a Gaussian process:
+%
+\begin{align}
+    \tilde{\mathbf{f}\hspace{0.2em}}
+    \left(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet\right)
+    \sim
+    \mathcal{N}\left(
+    \mathbf{1},
+    \tilde{\pmb{\Sigma}} \left(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet\right)
+    \right)
+    \quad,
+\end{align}
+%
+where $\bar{\pmb{\theta}}_\star$ are hyperparameters describing the star, such
+as the rotation period and limb darkening coefficients
+(Equation~\ref{eq:thetastarbar}),
+$\pmb{\theta}_\bullet$ are the hyperparameters describing the star spot
+distribution, including the number of spots, their contrast, parameters
+describing their latitudinal distribution, and parameters describing their
+radius distribution (Equation~\ref{eq:thetaspot}),
+and $\tilde{\pmb{\Sigma}}$ is the covariance of the normalized process
+(Equation~\ref{eq:SigmaTilde}). This covariance is a straightforward correction to the
+true covariance of the process, which accounts for
+changes of scale and phase dependence introduced by the common process of
+normalizing light curves to a mean of unity. It depends on the
+true (constant) mean $\mu$
+and true covariance $\pmb{\Sigma}$,
+given by Equations~(\ref{eq:mu_marg}) and (\ref{eq:cov_marg}),
+respectively.
+Those expressions in turn depend on the inclination expectation integrals
+$e_I$ (Appendix~\ref{sec:inc-mom1}) and $\mathbf{E}_I$
+(Appendix~\ref{sec:inc-mom2}). Those, in turn, depend on the first and
+second moments of the distribution of spherical harmonic coefficient vectors,
+$\mathrm{E}\left[ \mathbf{y} \big| \pmb{\theta}_\bullet \right]$
+and
+$\mathrm{E}\left[ \mathbf{y} \, \mathbf{y}^\top \big| \pmb{\theta}_\bullet \right]$,
+given by Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), respectively.
+To compute those, we must evaluate four nested integrals
+(Equations~\ref{eq:e1}--\ref{eq:e4} for the first moment
+and \ref{eq:E1}--\ref{eq:E4} for the second moment), corresponding to integrals
+over the radius, latitude, longitude, and contrast distributions, respectively.
+The computation of these integrals is discussed at length in Appendix~\ref{sec:integrals}.
+
+While lengthy (and quite tedious), all of the computations described above rely
+on equations whose solutions have a closed form.%
+%
+\footnote{The exception to this is the
+    normalization correction (\S\ref{sec:baseline}), which depends on a rapidly
+    converging series and thus adds negligible overhead to the computation.}
+%
+Moreover, most of the terms in the expectation vectors and matrices may
+be computed recursively, and many may be pre-computed, as they do not
+depend on user inputs.
+%
+It is therefore possible to evaluate $\tilde{\pmb{\Sigma}}$ in an extremely
+efficient manner. In \S\ref{sec:implementation} we discuss our
+implementation of the algorithm in a user-friendly \Python package.
+
+\section{Discussion}
+\label{sec:discussion}
+
+\subsection{Caveats}
+\label{sec:caveats}
+The true process is not Gaussian. See Figure~\ref{fig:nongaussianity}.
 \begin{figure}[ht!]
@@
 $I = \nicefrac{\pi}{2}$ corresponding to an edge-on view). With this last rotation,
-we are now in the observer's frame%
+we are now in the observer's frame.%
 \footnote{In principle, one last rotation could be performed about $\hat{\mathbf{z}}$
@@
     flux is independent of the rotation angle along the plane of the sky
-    (which we refer to as the \emph{obliquity}), so this step is unnecessary.}.
+    (which we refer to as the \emph{obliquity}), so this step is unnecessary.}
 
@@
 \section{Inclination}
-\label{sec:inclination}
+\label{sec:inc}
 
@@
 \subsection{Second moment}
+\label{sec:inc-mom2}
 %

### 2020-12-04 tex/gitlinks.tex
more text; features

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/64ae7e87c3615ea94e611dbe98f372be309bb3d6/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/64ae7e87c3615ea94e611dbe98f372be309bb3d6/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/64ae7e87c3615ea94e611dbe98f372be309bb3d6/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/b705a027358ae5fc187985032ec1bcb39175deee/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/b705a027358ae5fc187985032ec1bcb39175deee/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/b705a027358ae5fc187985032ec1bcb39175deee/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-04 tex/ms.tex
more text; features

@@
 \keywords{methods: analytic}
+
+\features{open-source figures \codeicon; equation unit tests \testpassicon}
 
@@
 \label{sec:motivation}
-See Figure~\ref{fig:nullspace_ensemble}.
+
+Figure~\ref{fig:nullspace_ensemble} shows the posterior shrinkage
+as a function of spherical harmonic degree for three different
+(purely hypothetical) scenarios: a single high signal-to-noise ratio (SNR)
+observation of a stellar light curve at a random
+observer inclination (thin blue curves), a set of three observations
+of the same star at three \emph{different} random inclinations (thin
+orange curves), and a set of ten observations at ten different
+random inclinations (thin green curves).
+The denser, opaque curves show the average for each scenario over 300 trials.
+The posterior shrinkage $R$ for a given
+spherical harmonic coefficient is computed from
+%
+\begin{align}
+    R \equiv 1 - \lim\limits_{\sigma_0^2 \rightarrow \infty}
+    \frac{\sigma^2}{\sigma_0^2}
+\end{align}
+%
+where $\sigma_0^2$ is the prior variance
+and $\sigma^2$ is the posterior variance.
+It is a measure of how informative a measurement is about a
+given mode on the surface in the limit of infinite SNR
+and is independent of what the stellar surface actually looks like.
+If, at infinite SNR and with a completely uninformative prior,
+a particular mode can be learned exactly from a dataset, the posterior
+shrinkage is defined to be unity. Conversely, if the data is completely
+unconstraining, $R$ will tend to zero.
 
@@
 \end{figure}
+
+Let us first consider the blue curves, in which a light curve is
+collected from a star viewed at a single random inclination (which, of course, is
+all we can do from our fixed vantage point on Earth).
+For some of the low-degree modes, the shrinkage is relatively high: it is
+fairly easy to constrain the dipole moment from a light curve, as this is
+usually the dominant sinusoidal signal. However, as the degree $l$
+increases, the shrinkage decreases dramatically: at $l = 14$, corresponding
+to features on scales of roughly $15^\circ$, the light curve
+can only tell us about $\sim 10\%$ of the total information about what the
+surface looks like. As $l$ increases further, $R$ tends to zero.
+Another important feature of the shrinkage is that it is exactly zero for
+all odd-degree modes above $l = 1$. This is a well-known fact: all odd spherical
+harmonics other than the dipole are in the null space \emph{regardless of
+    inclination} \citep[e.g.,][]{Luger2019}. In other words, these spherical
+harmonics are perfectly antisymmetric in projection over the unit disk
+when viewed from any orientation. Absent structure to break these symmetries
+(see \S\ref{sec:limbdark}), we simply cannot learn anything about these modes from
+stellar light curves. If we average over the shrinkage for all modes up to $l=15$,
+we find that a single light curve measurement can only tell us $\sim 9\%$
+of the information about the surface.
+
+The orange curves correspond to the scenario in which we can collect light curves
+from a star viewed at three different random inclinations. This is obviously
+impossible in practice, but it is still instructive as a thought experiment.
+Interestingly, the shrinkage increased at all even spherical harmonic degrees
+(the odd degrees, as we mentioned above, are always invisible).
+Note that since we are in the limit of infinite SNR, the fact that we have
+three times the data as we did in the previous scenario is irrelevant: the
+increase in the shrinkage is instead due to the fact that our observations
+from different vantage points broke some degeneracies in the problem.
+This is a consequence of the fact that the null space (for the even modes)
+is a strong function of the inclination: modes that are invisbile
+at one inclination can actually project into the flux at a different inclination.
+If we average over all modes, we obtain a total shrinkage of $\sim 24\%$
+for $l\leq15$ in this case.
+
+Finally, the green curves correspond to observations taken at ten random
+inclinations. In this scenario, the posterior shrinkage is nearly $100\%$
+for all even modes (it asymptotically approaches $100\%$ as the number
+of light curves increases further). The total shrinkage is $\sim 47\%$
+and approaches $50\%$ as both $l_\mathrm{max}$  and the number of light curves
+tend to infinity. Thus, if we were able to measure a light curve of a star
+from many different inclinations, the null space would consist \emph{only}
+of the odd modes, and our data would tell us exactly half of all the
+information about the surface.
+
+As we mentioned above, it is obviously not possible to observe a \emph{single}
+star from many different inclinations (at least not at present!).
+However, it is certainly the case that we can observe \emph{many} stars
+at a wide range of inclinations. Those stars certainly have different
+surface intensity distributions, but what if we restricted ourselves to
+light curves of ``similar'' stars?
 

### 2020-12-04 tex/gitlinks.tex
more text

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/b705a027358ae5fc187985032ec1bcb39175deee/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/b705a027358ae5fc187985032ec1bcb39175deee/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/b705a027358ae5fc187985032ec1bcb39175deee/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/c50415978a292c49f67eb6752e3f57c606f8bf6f/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/c50415978a292c49f67eb6752e3f57c606f8bf6f/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/c50415978a292c49f67eb6752e3f57c606f8bf6f/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-04 tex/ms.tex
more text

@@
     \textbf{
-        Interpretable Gaussian Processes for Stellar Light Curves
+        Interpretable Gaussian Processes for Stellar Variability
     }
@@
 \label{sec:intro}
-\xxx{Talk about starry, gps, etc.}
-
-\section{Motivation}
-\label{sec:motivation}
-
-Figure~\ref{fig:nullspace_ensemble} shows the posterior shrinkage
-as a function of spherical harmonic degree for three different
-(purely hypothetical) scenarios: a single high signal-to-noise ratio (SNR)
-observation of a stellar light curve at a random
-observer inclination (thin blue curves), a set of three observations
-of the same star at three \emph{different} random inclinations (thin
-orange curves), and a set of ten observations at ten different
-random inclinations (thin green curves).
-The denser, opaque curves show the average for each scenario over 300 trials.
-The posterior shrinkage $R$ for a given
-spherical harmonic coefficient is computed from
-%
-\begin{align}
-    R \equiv 1 - \lim\limits_{\sigma_0^2 \rightarrow \infty}
-    \frac{\sigma^2}{\sigma_0^2}
-\end{align}
-%
-where $\sigma_0^2$ is the prior variance
-and $\sigma^2$ is the posterior variance.
-It is a measure of how informative a measurement is about a
-given mode on the surface in the limit of infinite SNR
-and is independent of what the stellar surface actually looks like.
-If, at infinite SNR and with a completely uninformative prior,
-a particular mode can be learned exactly from a dataset, the posterior
-shrinkage is defined to be unity. Conversely, if the data is completely
-unconstraining, $R$ will tend to zero.
-
-\begin{figure}[ht!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/nullspace_ensemble.pdf}
-        \oscaption{nullspace_ensemble}{%
-            Posterior shrinkage as a function of spherical harmonic degree
-            for a hypothetical scenario in which an observer can measure a
-            stellar light curve at a single random inclination (thin blue curves),
-            at 3 random inclinations (thin orange curves), and at 10 random
-            inclinations (thin green curves). The thicker curves correspond to
-            the mean shrinkage for each experiment.
-            The information content in the light curve of a star observed
-            from a single vantage point approaches zero as $l$
-            increases. However, observing the star from many vantage points
-            allows one to recover nearly all of the information in the
-            even spherical harmonic modes.
-            While this thought experiment is not possible in practice for
-            a single star, it is extremely relevant to ensemble analyses of statistically
-            similar stars.
-            \label{fig:nullspace_ensemble}
-        }
-    \end{centering}
-\end{figure}
-
-Let us first consider the blue curves, in which a light curve is
-collected from a star viewed at a single random inclination (which, of course, is
-all we can do from our fixed vantage point on Earth).
-For some of the low-degree modes, the shrinkage is relatively high: it is
-fairly easy to constrain the dipole moment from a light curve, as this is
-usually the dominant sinusoidal signal. However, as the degree $l$
-increases, the shrinkage decreases dramatically: at $l = 14$, corresponding
-to features on scales of roughly $15^\circ$, the light curve
-can only tell us about $\sim 10\%$ of the total information about what the
-surface looks like. As $l$ increases further, $R$ tends to zero.
-Another important feature of the shrinkage is that it is exactly zero for
-all odd-degree modes above $l = 1$. This is a well-known fact: all odd spherical
-harmonics other than the dipole are in the null space \emph{regardless of
-    inclination} \citep[e.g.,][]{Luger2019}. In other words, these spherical
-harmonics are perfectly antisymmetric in projection over the unit disk
-when viewed from any orientation. Absent structure to break these symmetries
-(see \S\ref{sec:limbdark}), we simply cannot learn anything about these modes from
-stellar light curves. If we average over the shrinkage for all modes up to $l=15$,
-we find that a single light curve measurement can only tell us $\sim 9\%$
-of the information about the surface.
-
-The orange curves correspond to the scenario in which we can collect light curves
-from a star viewed at three different random inclinations. This is obviously
-impossible in practice, but it is still instructive as a thought experiment.
-Interestingly, the shrinkage increased at all even spherical harmonic degrees
-(the odd degrees, as we mentioned above, are always invisible).
-Note that since we are in the limit of infinite SNR, the fact that we have
-three times the data as we did in the previous scenario is irrelevant: the
-increase in the shrinkage is instead due to the fact that our observations
-from different vantage points broke some degeneracies in the problem.
-This is a consequence of the fact that the null space (for the even modes)
-is a strong function of the inclination: modes that are invisbile
-at one inclination can actually project into the flux at a different inclination.
-If we average over all modes, we obtain a total shrinkage of $\sim 24\%$
-for $l\leq15$ in this case.
-
-Finally, the green curves correspond to observations taken at ten random
-inclinations. In this scenario, the posterior shrinkage is nearly $100\%$
-for all even modes (it asymptotically approaches $100\%$ as the number
-of light curves increases further). The total shrinkage is $\sim 47\%$
-and approaches $50\%$ as both $l_\mathrm{max}$  and the number of light curves
-tend to infinity. Thus, if we were able to measure a light curve of a star
-from many different inclinations, the null space would consist \emph{only}
-of the odd modes, and our data would tell us exactly half of all the
-information about the surface.
-
-As we mentioned above, it is obviously not possible to observe a \emph{single}
-star from many different inclinations (at least not at present!).
-However, it is certainly the case that we can observe \emph{many} stars
-at a wide range of inclinations. Those stars certainly have different
-surface intensity distributions, but what if we restricted ourselves to
-light curves of ``similar'' stars?
+
+Over the past decade, Gaussian processes (GPs) have gained traction as
+a leading tool for modeling correlated signals in astrophysical timeseries.
+Despite whatever mystique the words ``Gaussian process'' may evoke, a GP
+is nothing but a Gaussian distribution in many (formally infinite)
+dimensions. Specifically, it is a Gaussian distribution over
+\emph{functions} spanning a continuous domain (in our case, the time domain).
+Like a multivariate
+Gaussian, a GP is fully specified by a $(K \times 1)$ vector $\pmb{\mu}$ characterizing
+the mean of the process and a $(K \times K)$
+covariance matrix $\pmb{\Sigma}$ characterizing how the data points
+are correlated with each other. To say that a measurement $\mathbf{f}$ is
+``distributed as a GP'' means that we may write
+%
+\begin{align}
+    \mathbf{f} \sim \mathcal{N}\left( \pmb{\mu}, \pmb{\Sigma} \right)
+    \quad,
+\end{align}
+%
+i.e., the measurement is modeled as a random draw from a multivariate Gaussian.
+GPs are therefore extremely easy to sample from.%
+\footnote{Given a 1-d array \texttt{mean} and a 2-d array \texttt{cov} in \Python,
+    sampling from the corresponding GP (if it exists)
+    can be done in a single line of code by calling
+    \texttt{numpy.random.multivariate\_normal(mean, cov)}.}
+But the real showstopper is the application of GPs to inference problems.
+Multivariate Gaussian distributions have a closed-form (marginal) likelihood
+function, so it is easy (and extremely efficient) to compute the probability
+of one's data conditioned on a given value of $\pmb{\mu}$ and $\pmb{\Sigma}$
+(i.e., the ``likelihood''). This can in turn be maximized
+to infer the optimal values of the model parameters
+or used in a
+numerical sampling scheme to compute the probability of those parameters
+given the data (i.e., the ``posterior'').
+Thanks to modern computer architectures, linear algebra packages, and
+GP algorithms,
+evaluating the GP likelihood may typically be done in a fraction of a second
+for a reasonably-sized dataset (i.e., $K \lesssim 10^5$ datapoints).
+
+Another big advantage of GPs is their flexibility. GPs are often dubbed
+a class of ``non-parametric'' models, given that nowhere in the specification
+of the GP is there an explicit functional form for $\mathbf{f}$. Rather, GPs are a stochastic
+process whose draws can in principle take on \emph{any} functional form,
+subject, however, to certain smoothness and correlation criteria
+of tunable strictness
+that are fully
+encoded in the covariance $\pmb{\Sigma}$. A common approach to constructing
+$\pmb{\Sigma}$ is to define a continuous \emph{kernel function} $k(t_i, t_j)$
+such that the covariance between points at times $t = t_i$ and $t = t_j$ is
+%
+\begin{align}
+    \Sigma_{i,j} = k(t_i, t_j)
+    \quad.
+\end{align}
+%
+In many applications, it is customary to further restrict the problem by
+assuming that the process is \emph{stationary}, such that we may write
+%
+\begin{align}
+    \Sigma_{i,j} & = k(t_i, t_j)
+    \nonumber                                    \\
+                 & = k(\left| t_i - t_j \right|)
+    \nonumber                                    \\
+                 & \equiv k(\Delta t)
+    \quad.
+\end{align}
+%
+A stationary process is one that is independent of phase (or, in this case,
+the actual value of the time $t$); rather, it depends only on the \emph{difference}
+between the phases of two data points. The kernel of a stationary process is
+therefore a one-dimensional function, which is typically chosen from a set of
+standard functions with desirable smoothness and spectral properties. These
+include, for example, the squared exponential kernel
+%
+\begin{align}
+    k_\mathrm{E^2}(\Delta t) = \sigma^2 e^{-\frac{\Delta t}{2\tau}}
+    \quad,
+\end{align}
+%
+the Mat\'ern-3/2 kernel
+%
+\begin{align}
+    k_\mathrm{M}(\Delta t) = \sigma^2 \left( 1 + \frac{\sqrt{3}\Delta t}{\tau} \right) e^{-\frac{\sqrt{3}\Delta t}{\tau}}
+    \quad,
+\end{align}
+%
+the exponential sine squared kernel,
+%
+\begin{align}
+    k_\mathrm{S^2}(\Delta t) = \sigma^2 e^{-\Gamma\sin^2\left( \frac{\pi \Delta t}{P} \right)}
+    \quad,
+\end{align}
+%
+the cosine kernel,
+%
+\begin{align}
+    k_\mathrm{C}(\Delta t) = \cos\left( \frac{2\pi \Delta t}{P} \right)
+    \quad,
+\end{align}
+%
+or any linear combination or product of those (as well as many other kernels not
+listed here). In the expressions above, $\sigma$, $\tau$, $\Gamma$, and $P$ are
+the \emph{hyperparameters} of the kernels, scalar quantities that (together with
+the assumed functional form for the kernel) fully specify the covariance of the
+Gaussian process.
+
+\xxx{More on this soon!}
 
@@
 
+\subsection{Ensemble Analyses}
+\label{sec:ensemble}
+
+Figure~\ref{fig:nullspace_ensemble} shows the posterior shrinkage
+as a function of spherical harmonic degree for three different
+(purely hypothetical) scenarios: a single high signal-to-noise ratio (SNR)
+observation of a stellar light curve at a random
+observer inclination (thin blue curves), a set of three observations
+of the same star at three \emph{different} random inclinations (thin
+orange curves), and a set of ten observations at ten different
+random inclinations (thin green curves).
+The denser, opaque curves show the average for each scenario over 300 trials.
+The posterior shrinkage $R$ for a given
+spherical harmonic coefficient is computed from
+%
+\begin{align}
+    R \equiv 1 - \lim\limits_{\sigma_0^2 \rightarrow \infty}
+    \frac{\sigma^2}{\sigma_0^2}
+\end{align}
+%
+where $\sigma_0^2$ is the prior variance
+and $\sigma^2$ is the posterior variance.
+It is a measure of how informative a measurement is about a
+given mode on the surface in the limit of infinite SNR
+and is independent of what the stellar surface actually looks like.
+If, at infinite SNR and with a completely uninformative prior,
+a particular mode can be learned exactly from a dataset, the posterior
+shrinkage is defined to be unity. Conversely, if the data is completely
+unconstraining, $R$ will tend to zero.
+
+\begin{figure}[ht!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/nullspace_ensemble.pdf}
+        \oscaption{nullspace_ensemble}{%
+            Posterior shrinkage as a function of spherical harmonic degree
+            for a hypothetical scenario in which an observer can measure a
+            stellar light curve at a single random inclination (thin blue curves),
+            at 3 random inclinations (thin orange curves), and at 10 random
+            inclinations (thin green curves). The thicker curves correspond to
+            the mean shrinkage for each experiment.
+            The information content in the light curve of a star observed
+            from a single vantage point approaches zero as $l$
+            increases. However, observing the star from many vantage points
+            allows one to recover nearly all of the information in the
+            even spherical harmonic modes.
+            While this thought experiment is not possible in practice for
+            a single star, it is extremely relevant to ensemble analyses of statistically
+            similar stars.
+            \label{fig:nullspace_ensemble}
+        }
+    \end{centering}
+\end{figure}
+
+Let us first consider the blue curves, in which a light curve is
+collected from a star viewed at a single random inclination (which, of course, is
+all we can do from our fixed vantage point on Earth).
+For some of the low-degree modes, the shrinkage is relatively high: it is
+fairly easy to constrain the dipole moment from a light curve, as this is
+usually the dominant sinusoidal signal. However, as the degree $l$
+increases, the shrinkage decreases dramatically: at $l = 14$, corresponding
+to features on scales of roughly $15^\circ$, the light curve
+can only tell us about $\sim 10\%$ of the total information about what the
+surface looks like. As $l$ increases further, $R$ tends to zero.
+Another important feature of the shrinkage is that it is exactly zero for
+all odd-degree modes above $l = 1$. This is a well-known fact: all odd spherical
+harmonics other than the dipole are in the null space \emph{regardless of
+    inclination} \citep[e.g.,][]{Luger2019}. In other words, these spherical
+harmonics are perfectly antisymmetric in projection over the unit disk
+when viewed from any orientation. Absent structure to break these symmetries
+(see \S\ref{sec:limbdark}), we simply cannot learn anything about these modes from
+stellar light curves. If we average over the shrinkage for all modes up to $l=15$,
+we find that a single light curve measurement can only tell us $\sim 9\%$
+of the information about the surface.
+
+The orange curves correspond to the scenario in which we can collect light curves
+from a star viewed at three different random inclinations. This is obviously
+impossible in practice, but it is still instructive as a thought experiment.
+Interestingly, the shrinkage increased at all even spherical harmonic degrees
+(the odd degrees, as we mentioned above, are always invisible).
+Note that since we are in the limit of infinite SNR, the fact that we have
+three times the data as we did in the previous scenario is irrelevant: the
+increase in the shrinkage is instead due to the fact that our observations
+from different vantage points broke some degeneracies in the problem.
+This is a consequence of the fact that the null space (for the even modes)
+is a strong function of the inclination: modes that are invisbile
+at one inclination can actually project into the flux at a different inclination.
+If we average over all modes, we obtain a total shrinkage of $\sim 24\%$
+for $l\leq15$ in this case.
+
+Finally, the green curves correspond to observations taken at ten random
+inclinations. In this scenario, the posterior shrinkage is nearly $100\%$
+for all even modes (it asymptotically approaches $100\%$ as the number
+of light curves increases further). The total shrinkage is $\sim 47\%$
+and approaches $50\%$ as both $l_\mathrm{max}$  and the number of light curves
+tend to infinity. Thus, if we were able to measure a light curve of a star
+from many different inclinations, the null space would consist \emph{only}
+of the odd modes, and our data would tell us exactly half of all the
+information about the surface.
+
+As we mentioned above, it is obviously not possible to observe a \emph{single}
+star from many different inclinations (at least not at present!).
+However, it is certainly the case that we can observe \emph{many} stars
+at a wide range of inclinations. Those stars certainly have different
+surface intensity distributions, but what if we restricted ourselves to
+light curves of ``similar'' stars?
+
 \subsection{Caveats}

### 2020-12-05 tex/gitlinks.tex
more text; fix test stability issue

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/c50415978a292c49f67eb6752e3f57c606f8bf6f/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/c50415978a292c49f67eb6752e3f57c606f8bf6f/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/c50415978a292c49f67eb6752e3f57c606f8bf6f/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/41f024ad5662f691f48103c4041eb3e915523e38/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/41f024ad5662f691f48103c4041eb3e915523e38/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/41f024ad5662f691f48103c4041eb3e915523e38/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-05 tex/ms.tex
more text; fix test stability issue

@@
 of one's data conditioned on a given value of $\pmb{\mu}$ and $\pmb{\Sigma}$
-(i.e., the ``likelihood''). This can in turn be maximized
+(i.e., the ``likelihood''; see Equation~\ref{eq:log-like} below).
+This can in turn be maximized
 to infer the optimal values of the model parameters
@@
 a class of ``non-parametric'' models, given that nowhere in the specification
-of the GP is there an explicit functional form for $\mathbf{f}$. Rather, GPs are a stochastic
+of the GP is there an explicit functional form for $\mathbf{f}$. Rather, a GP
+is a stochastic
 process whose draws can in principle take on \emph{any} functional form,
@@
 that are fully
-encoded in the covariance $\pmb{\Sigma}$. A common approach to constructing
-$\pmb{\Sigma}$ is to define a continuous \emph{kernel function} $k(t_i, t_j)$
-such that the covariance between points at times $t = t_i$ and $t = t_j$ is
+encoded in the covariance $\pmb{\Sigma}$. Importantly, the covariance
+$\pmb{\Sigma}$ of a GP is assembled via a continuous \emph{kernel function} $k(t_i, t_j)$
+that specifies the covariance between points at times $t = t_i$ and $t = t_j$:
 %
@@
 %
-In many applications, it is customary to further restrict the problem by
+In many applications, particularly when modeling stellar light curves,
+it is customary to restrict the problem by
 assuming that the process is \emph{stationary}, such that we may write
@@
 between the phases of two data points. The kernel of a stationary process is
-therefore a one-dimensional function, which is typically chosen from a set of
-standard functions with desirable smoothness and spectral properties. These
-include, for example, the squared exponential kernel
+therefore a one-dimensional function, typically chosen from a set of
+standard functions with desirable smoothness and spectral properties.
+Among a vast number of kernels are a few that are commonly used
+in modeling stellar variability: the squared exponential kernel,
 %
@@
 %
-the Mat\'ern-3/2 kernel
-%
-\begin{align}
-    k_\mathrm{M}(\Delta t) = \sigma^2 \left( 1 + \frac{\sqrt{3}\Delta t}{\tau} \right) e^{-\frac{\sqrt{3}\Delta t}{\tau}}
+the Mat\'ern-3/2 kernel,
+%
+\begin{align}
+    k_\mathrm{M^\frac{3}{2}}(\Delta t) = \sigma^2 \left( 1 + \frac{\sqrt{3}\Delta t}{\tau} \right) e^{-\frac{\sqrt{3}\Delta t}{\tau}}
     \quad,
@@
 \begin{align}
-    k_\mathrm{C}(\Delta t) = \cos\left( \frac{2\pi \Delta t}{P} \right)
-    \quad,
-\end{align}
-%
-or any linear combination or product of those (as well as many other kernels not
-listed here). In the expressions above, $\sigma$, $\tau$, $\Gamma$, and $P$ are
-the \emph{hyperparameters} of the kernels, scalar quantities that (together with
-the assumed functional form for the kernel) fully specify the covariance of the
-Gaussian process.
+    k_\mathrm{C}(\Delta t) = \sigma^2 \cos\left( \frac{2\pi \Delta t}{P} \right)
+    \quad,
+\end{align}
+%
+or any linear combination and/or product of these. In the expressions above,
+$\sigma$, $\tau$, $\Gamma$, and $P$ are
+the \emph{hyperparameters} of the kernels.
+Specifically, $\sigma^2$ is the variance (the value
+of the kernel when $t_i = t_j$), $\tau$ is a typical timescale over which
+data are correlated, $\Gamma$ is a dimensionless scale parameter, and $P$ is
+a period. Once a kernel has been chosen, these scalar quantities fully
+specify the covariance of the Gaussian process. GP-based inference
+entails optimizing or sampling over these few quantities, which in many
+cases can be done easily and efficiently.
+
+Which brings us, finally, to an important drawback of GPs when applied
+to modeling stellar variability. While it is straightforward to derive
+posterior constraints on the hyperparameters of a GP, it is usually
+not obvious what those constraints actually tell us about the star
+or the actual source of the variability. While the GP period $P$
+may usually be interpreted as the rotational period of the star,%
+\footnote{An exception to this is in the presence of strong differential
+    rotation, in which case many periods may be present in the data, or
+    when spots evolve coherently, which can also introduce weak periodicities
+    in the light curve.}%
+the relationship between quantities like the amplitude $\sigma^2$
+and the timescale $\tau$ and physical quantites of interest---such
+as the properties of the star spots---are much fuzzier.
+It may be tempting to interpret $\sigma^2$ as some measure of the
+contrast or number of spots, or $\tau$ as the timescale on which the
+spots evolve, but there are no guarantees these interpretations will
+hold in general.
 

### 2020-12-06 tex/gitlinks.tex
finished intro

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/41f024ad5662f691f48103c4041eb3e915523e38/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/41f024ad5662f691f48103c4041eb3e915523e38/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/41f024ad5662f691f48103c4041eb3e915523e38/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/1ceb09cbd12c72831cc9ea7579c67ae74c9220eb/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/1ceb09cbd12c72831cc9ea7579c67ae74c9220eb/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/1ceb09cbd12c72831cc9ea7579c67ae74c9220eb/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-06 tex/ms.tex
finished intro

@@
 or the actual source of the variability. While the GP period $P$
-may usually be interpreted as the rotational period of the star,%
+may sometimes be interpreted as the rotational period of the star,%
 \footnote{An exception to this is in the presence of strong differential
@@
     when spots evolve coherently, which can also introduce weak periodicities
-    in the light curve.}%
+    in the light curve.}
 the relationship between quantities like the amplitude $\sigma^2$
 and the timescale $\tau$ and physical quantites of interest---such
-as the properties of the star spots---are much fuzzier.
-It may be tempting to interpret $\sigma^2$ as some measure of the
-contrast or number of spots, or $\tau$ as the timescale on which the
+as the properties of the star spots---is much fuzzier.
+It may be tempting to interpret $\sigma^2$ as some measure of the spot
+contrast or the total number of spots, or $\tau$ as the timescale on which the
 spots evolve, but there are no guarantees these interpretations will
-hold in general.
-
-\xxx{More on this soon!}
+hold in general. After all, the choice of kernel is quite often \emph{ad hoc},
+providing an \emph{effective}---as opposed to \emph{interpretable}---description
+of the physics.
+
+For this reason, GPs are usually employed to model stellar variability
+when the variability itself is a nuisance parameter. For example, if the
+goal is to constrain properties of a transiting exoplanet, a GP might be
+used to remove (or, better yet, to marginalize over) the stellar signal.
+In this case, the physics behind the variability is irrelevant, so an
+effective model of this sort is generally sufficient. When the goal is to understand
+the variability itself, however, GPs have been far less useful to date. Instead,
+the usual approach has been to explicitly model the stellar signal: in the
+case of star spot-induced variability, one must construct a physical model
+for the stellar surface and for how it projects into the light curve as the
+star rotates. This might entail directly modeling the star spots,
+with free parameters controling the spot size, contrast, position,
+and temporal evolution of each of the $N$ spots, where $N$ is also
+a free parameter. In addition to this being a numerically challenging
+problem for all but the smallest values of $N$, the problem is
+intrinsically ill-posed, meaning there are strong degeneracies at play
+that formally preclude a unique solution to the stellar map.
+After all, a stellar light curve is a one-dimensional representation of a
+two-dimensional (or, if the star evolves in time, three-dimensional)
+surface, and thus it simply cannot encode all the information about the stellar surface.
+In fact, the \emph{vast majority} of the modes on the stellar surface
+are in the \emph{null space}, the set of surface features that have identically
+zero effect on the light curve.
+The underdetermined nature of the stellar mapping problem is therefore fundamental: it
+is simply not possible to uniquely determine the properties of all the spots on
+a star from its rotational light curve.
+
+However, it is hardly ever the case that this is actually our end goal.
+After all, physics predicts the properties of stellar surfaces at a fairly
+high level: i.e., typical spot sizes, active spot latitudes, or approximate
+timescales on which spots evolve. We are hardly ever interested in the
+\emph{particular} properties of a \emph{particular} spot, as we wouldn't really
+know what to do with that information! Instead, we often treat
+(whether explicitly or not)
+the properties of a star spot as a draw from some parent distribution
+controlling (say) the average and spread in the radii of the spots.
+The parameters controling this distribution are the ones that we can
+predict with physics; they are therefore also the ones we are usually
+interested in.
+
+Thus, if we were able to derive robust posterior constraints
+on the properties of each of the spots on a star, we could then
+\emph{marginalize} (integrate) over them to infer the properties
+describing the distribution of all the spots as a whole.
+%In principle,
+%this would work even if our posterior distributions contain
+%strong degeneracies: as we will see, the statistical properties of
+%star spots have a much smaller null space than the properties of
+%individual spots. 
+We could do this in the usual way described above, by
+modeling the properties of each of the spots in a posterior sampling
+scheme, but also including a few
+\emph{hyperparameters} controlling the distribution of those properties
+across all spots (acting, therefore, as a prior term): i.e., a one-level
+hierarchical model. The marginal posteriors for the hyperparameters, then,
+would encode what we actually wish to know.
+%
+In practice, however, the degeneracies and often extreme multi-modality
+of the distributions of individual spot properties would make this
+quite hard (and expensive) to perform.
+%
+If only we could use the elegant machinery of Gaussian processes to
+perform this marginalization for us!
+
+\textbf{In this paper, we derive an exact, closed-form expression for
+    the Gaussian approximation to the marginal likelihood of a light curve
+    conditioned on the statistical properties of star spots.} Our Gaussian
+process analytically marginalizes over the (degenerate, often unknowable)
+distributions of properties of individual star spots, revealing the
+constraints imposed on the bulk spot properties without the need to
+explicitly model or sample over properties of individual spots. It inherits
+the speed, ease-of-use, and all other properties of traditionally-used
+GPs, with the added benefit of direct physical interpretability of its
+hyperparameters.
+%
 

### 2020-12-07 tex/gitlinks.tex
added limb darkening; still needs testing

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/1ceb09cbd12c72831cc9ea7579c67ae74c9220eb/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/1ceb09cbd12c72831cc9ea7579c67ae74c9220eb/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/1ceb09cbd12c72831cc9ea7579c67ae74c9220eb/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/a27b8de7b8ca0c6ba3f072e82097cd00d7a6e643/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/a27b8de7b8ca0c6ba3f072e82097cd00d7a6e643/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/a27b8de7b8ca0c6ba3f072e82097cd00d7a6e643/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-07 tex/ms.tex
added limb darkening; still needs testing

@@
 %
-\xxx{TODO.}
+\label{sec:ld}
+We must adjust our expression for the flux in the presence of limb darkening.
+For any polynomial limb darkening law of the form
+%
+\begin{align}
+    \label{eq:ld:I}
+    \frac{I(\mu)}{I(\mu = 1)} = 1 - \sum_{n=1}^{n_\mathrm{max}} u_n(1 - \mu)^n
+    \quad,
+\end{align}
+%
+where $I$ is the intensity on the stellar surface,
+$\mu = z = \sqrt{1 - x^2 - y^2}$ is the radial coordinate on the
+projected disk, and $u_n$ is a limb darkening coefficient, the effect of
+limb darkening on the stellar map can be expressed exactly as a linear
+operation on the spherical harmonic coefficient vector
+\citep{Luger2019}.
+This includes the popular linear and quadractic limb darkening laws
+and generalizes to \emph{any} limb darkening law in the limit
+$n_\mathrm{max} \rightarrow \infty$. The linearity of the problem
+can be understood by noting that all terms
+in Equation~(\ref{eq:ld:I}) are strictly polynomials in $x$, $y$, and $z$,
+all of which can be expressed exactly as sums of spherical harmonics
+\citep{Luger2019}. When weighting the surface intensity by the limb darkening
+profile, the resulting intensity is simply a product of spherical harmonics,
+which is itself a linear combination of spherical harmonics.
+Thus, given a limb darkening law
+of degree $n_\mathrm{max}$ with coefficients $\mathbf{u}$,
+we can construct a matrix $\mathbf{L}(\mathbf{u})$
+that transforms a spherical harmonic vector $\mathbf{y}$ of degree
+$l_\mathrm{max}$ to a limb-darkened spherical harmonic vector $\mathbf{y'}$
+of degree $l_\mathrm{max} + n_\mathrm{max}$. As an example, consider
+a map of degree $l_\mathrm{max} = 1$ and the linear limb darkening law
+($n_\mathrm{max} = 1$) with coefficient vector $\mathbf{u} = ( u_1 )$.
+The transformation matrix from $\mathbf{y}$ to $\mathbf{y'}$ is
+%
+\begin{proof}{}
+    \label{eq:ld:L}
+    \setstackgap{L}{1.5\baselineskip}
+    \mathbf{L}(\mathbf{u}) =
+    \frac{1}{1 - \frac{u_1}{3}}
+    \begin{pmatrix}
+        \quadquad1-u_1\quadquad & 0                       & \frac{u_1}{\sqrt{3}}    & 0                       \\
+        0                       & \quadquad1-u_1\quadquad & 0                       & 0                       \\
+        \frac{u_1}{\sqrt{3}}    & 0                       & \quadquad1-u_1\quadquad & 0                       \\
+        0                       & 0                       & 0                       & \quadquad1-u_1\quadquad \\
+        0                       & 0                       & 0                       & 0                       \\
+        0                       & \frac{u_1}{\sqrt{5}}    & 0                       & 0                       \\
+        0                       & 0                       & \frac{2u_1}{\sqrt{15}}  & 0                       \\
+        0                       & 0                       & 0                       & \frac{u_1}{\sqrt{5}}    \\
+        0                       & 0                       & 0                       & 0
+    \end{pmatrix}
+    \quad.
+\end{proof}
+%
+The columns of $\mathbf{L}$ are constructed from the coefficient vectors of
+each transformed spherical harmonic, which are in turn computed by
+multiplying each spherical harmonic by the spherical harmonic decomposition
+of the particular limb darkening law. \xxx{A simple algorithm for computing
+    $\mathbf{L}$ for limb darkening of any order is given in the link next to
+    Equation~(\ref{eq:ld:L}).}
+
+In the presence of limb darkening, we may therefore replace our expression for the
+$k^\mathrm{th}$ row of the flux design matrix $\mathcal{A}$ (Equation~\ref{eq:akT})
+with
+%
+\begin{align}
+    \label{eq:akTld}
+    \mathbf{a}_k^\top = \mathbf{r}^\top \,
+    \mathbf{A_1} \,
+    \mathbf{L}(\mathbf{u})
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(-I\right) \,
+    \mathbf{R}_{\hat{\mathbf{z}}}\left(\frac{2\pi}{P}t_k\right) \,
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(\frac{\pi}{2}\right)
+\end{align}
+%
+for a given limb darkening coefficient vector $\mathbf{u}$.
 

### 2020-12-07 starry_process/ops/base_op.py

@@ -1,5 +1,6 @@
 # -*- coding: utf-8 -*-
 from ..starry_process_version import __version__
+from ..defaults import defaults
 from theano import gof
 import theano
 import theano.tensor as tt
@@ -18,16 +19,21 @@ __all__ = ["BaseOp", "IntegralOp"]
 
 class BaseOp(gof.COp):
 
-    __props__ = ("ydeg", "compile_args")
+    __props__ = ("ydeg", "udeg", "compile_args")
     func_file = None
     func_name = None
 
-    def __init__(self, ydeg=None, compile_args=[], **kwargs):
-        if ydeg is not None:
-            self.ydeg = ydeg
-            self.N = (self.ydeg + 1) ** 2
-        else:
-            self.ydeg = None
+    def __init__(
+        self,
+        ydeg=defaults["ydeg"],
+        udeg=defaults["udeg"],
+        compile_args=[],
+        **kwargs
+    ):
+        self.ydeg = ydeg
+        self.N = (self.ydeg + 1) ** 2
+        self.udeg = udeg
+        self.NLU = (self.ydeg + self.udeg + 1) ** 2
         assert type(compile_args) is list, "arg `compile_args` must be a list"
         for item in compile_args:
             assert (
@@ -76,8 +82,8 @@ class BaseOp(gof.COp):
         args = ["-std=c++14", "-O2", "-DNDEBUG"]
         if sys.platform == "darwin":
             args += ["-stdlib=libc++", "-mmacosx-version-min=10.7"]
-        if self.ydeg is not None:
-            args += ["-DSP__LMAX={0}".format(self.ydeg)]
+        args += ["-DSP__LMAX={0}".format(self.ydeg)]
+        args += ["-DSP__UMAX={0}".format(self.udeg)]
         for (key, value) in self.compile_args:
             if key.startswith("SP_"):
                 args += ["-D{0}={1}".format(key, value)]

### 2020-12-07 starry_process/ops/flux/rTA1L_rev.cc

@@ -1,0 +1,70 @@
+#section support_code_struct
+
+// Shorthand
+#define DI0 DTYPE_INPUT_0
+#define DI1 DTYPE_INPUT_1
+#define DO0 DTYPE_OUTPUT_0
+#define TO0 TYPENUM_OUTPUT_0
+
+sp::flux::LimbDark<DO0> *APPLY_SPECIFIC(LD_rev);
+
+#section init_code_struct
+
+{ APPLY_SPECIFIC(LD_rev) = NULL; }
+
+#section cleanup_code_struct
+
+if (APPLY_SPECIFIC(LD_rev) != NULL) {
+  delete APPLY_SPECIFIC(LD_rev);
+}
+
+#section support_code_struct
+
+int APPLY_SPECIFIC(rTA1L_rev)(PyArrayObject *input0, PyArrayObject *input1,
+                              PyArrayObject **output0) {
+
+  using namespace sp::theano;
+  using namespace sp::flux;
+  using namespace sp::utils;
+
+  // Get the inputs
+  int success = 0;
+  int ndim = -1;
+  npy_intp *shape;
+  auto u_in = get_input<DI0>(&ndim, &shape, input0, &success);
+  if (ndim != 1) {
+    PyErr_Format(PyExc_ValueError, "u must be a vector");
+    return 1;
+  }
+  Map<Vector<DI0, SP__UMAX>> u(u_in);
+
+  ndim = -1;
+  auto bf_in = get_input<DI1>(&ndim, &shape, input1, &success);
+  if (ndim != 1) {
+    PyErr_Format(PyExc_ValueError, "bf must be a vector");
+    return 1;
+  }
+  Map<RowVector<DI1, SP__N>> bf(bf_in);
+
+  // Allocate the outputs
+  ndim = 1;
+  std::vector<npy_intp> shape_vec(ndim);
+  shape_vec[0] = SP__UMAX;
+  shape = &(shape_vec[0]);
+  auto bu_out = allocate_output<DO0>(ndim, shape, TO0, output0, &success);
+  if (success) {
+    return 1;
+  }
+  Map<Vector<DO0, SP__UMAX>> bu(bu_out);
+
+  // Initialize the class if needed
+  if (APPLY_SPECIFIC(LD_rev) == NULL) {
+    APPLY_SPECIFIC(LD_rev) = new LimbDark<DO0>();
+  }
+
+  // Compute
[…]

### 2020-12-07 starry_process/ops/include/constants.h

@@ -1,7 +1,7 @@
 /**
  * \file constants.h
  * \brief Hard-coded constants.
-*/
+ */
 
 #ifndef _SP_CONSTANTS_H_
 #define _SP_CONSTANTS_H_
@@ -15,8 +15,19 @@
 #define SP__LMAX 15
 #endif
 
+//! Degree of limb darkening
+#ifndef SP__UMAX
+#define SP__UMAX 2
+#endif
+
+//! Total degree
+#define SP__LUMAX (SP__LMAX + SP__UMAX)
+
 //! Number of Ylm terms
 #define SP__N ((SP__LMAX + 1) * (SP__LMAX + 1))
+
+//! Number of Ylm + LD terms
+#define SP__NLU ((SP__LMAX + SP__UMAX + 1) * (SP__LMAX + SP__UMAX + 1))
 
 //! Number of terms in the Wigner rotation matrix
 #define SP__NWIG                                                               \

### 2020-12-07 starry_process/ops/include/flux.h

@@ -19,21 +19,23 @@ using namespace utils;
 Compute the `r^T` phase curve solution vector.
 
 */
-template <typename Scalar> void computerT(RowVector<Scalar, SP__N> &rT) {
+template <typename Scalar>
+void computerT(const int deg, RowVector<Scalar, Dynamic> &rT) {
   Scalar amp0, amp, lfac1, lfac2;
+
   int mu, nu;
-  rT.setZero();
+  rT.setZero((deg + 1) * (deg + 1));
   amp0 = M_PI;
   lfac1 = 1.0;
   lfac2 = 2.0 / 3.0;
-  for (int l = 0; l < SP__LMAX + 1; l += 4) {
+  for (int l = 0; l < deg + 1; l += 4) {
     amp = amp0;
     for (int m = 0; m < l + 1; m += 4) {
       mu = l - m;
       nu = l + m;
       rT(l * l + l + m) = amp * lfac1;
       rT(l * l + l - m) = amp * lfac1;
-      if (l < SP__LMAX) {
+      if (l < deg) {
         rT((l + 1) * (l + 1) + l + m + 1) = amp * lfac2;
         rT((l + 1) * (l + 1) + l - m + 1) = amp * lfac2;
       }
@@ -46,14 +48,14 @@ template <typename Scalar> void computerT(RowVector<Scalar, SP__N> &rT) {
   amp0 = 0.5 * M_PI;
   lfac1 = 0.5;
   lfac2 = 4.0 / 15.0;
-  for (int l = 2; l < SP__LMAX + 1; l += 4) {
+  for (int l = 2; l < deg + 1; l += 4) {
     amp = amp0;
     for (int m = 2; m < l + 1; m += 4) {
       mu = l - m;
       nu = l + m;
       rT(l * l + l + m) = amp * lfac1;
       rT(l * l + l - m) = amp * lfac1;
-      if (l < SP__LMAX) {
+      if (l < deg) {
         rT((l + 1) * (l + 1) + l + m + 1) = amp * lfac2;
         rT((l + 1) * (l + 1) + l - m + 1) = amp * lfac2;
       }
@@ -70,13 +72,13 @@ Multiply a polynomial vector/matrix by `z`.
 
 */
 template <typename T1, typename T2>
-inline void polymulz(int lmax, const Eigen::MatrixBase<T1> &p,
+inline void polymulz(int deg, const Eigen::MatrixBase<T1> &p,
[…]

### 2020-12-07 starry_process/ops/include/utils.h

@@ -1,7 +1,7 @@
 /**
  * \file utils.h
  * \brief Miscellaneous utilities and definitions used throughout the code.
-*/
+ */
 
 #ifndef _SP_UTILS_H_
 #define _SP_UTILS_H_
@@ -9,6 +9,7 @@
 #include <Eigen/Core>
 #include <Eigen/Dense>
 #include <Eigen/SparseCore>
+#include <Eigen/SparseLU>
 #include <cmath>
 #include <exception>
 #include <iomanip>
@@ -27,8 +28,8 @@ using std::isnan;
 using std::max;
 
 //! Eigen shorthand
-using Eigen::Map;
 using Eigen::Dynamic;
+using Eigen::Map;
 template <typename Scalar, int M, int N>
 using RowMatrix = Eigen::Matrix<Scalar, M, N, Eigen::RowMajor>;
 template <typename Scalar, int M, int N>
@@ -62,7 +63,7 @@ void for_constexpr_impl(F &&f, std::index_sequence<Is...> /*meta*/) {
 /**
  * Compile-time for loop. From https://nilsdeppe.com/posts/for-constexpr
  * Requires C++14
-*/
+ */
 template <class Bounds0, class F> void for_constexpr(F &&f) {
   for_constexpr_detail::for_constexpr_impl<Bounds0::lower>(
       std::forward<F>(f),
@@ -71,7 +72,7 @@ template <class Bounds0, class F> void for_constexpr(F &&f) {
 
 /**
  * Generic starry exception class.
-*/
+ */
 class StarryProcessException : public std::exception {
 
   std::string m_msg;

### 2020-12-07 tests/test_variance.py

@@ -1,0 +1,11 @@
+from starry_process import StarryProcess
+import numpy as np
+
+# BROKEN!
+def test_variance():
+
+    sp = StarryProcess()
+    cov = sp.cov([0.0, 0.1])
+    var = sp.cov([0.0])
+
+    assert np.allclose(cov[0, 0], var)

### 2020-12-08 tex/gitlinks.tex
fix theano install issues

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/a27b8de7b8ca0c6ba3f072e82097cd00d7a6e643/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/a27b8de7b8ca0c6ba3f072e82097cd00d7a6e643/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/a27b8de7b8ca0c6ba3f072e82097cd00d7a6e643/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/bcfdb4752a408aca0ca122743ff193f5a910bd5a/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/bcfdb4752a408aca0ca122743ff193f5a910bd5a/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/bcfdb4752a408aca0ca122743ff193f5a910bd5a/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-08 tex/ms.tex
fix theano install issues

@@
 each transformed spherical harmonic, which are in turn computed by
-multiplying each spherical harmonic by the spherical harmonic decomposition
-of the particular limb darkening law. \xxx{A simple algorithm for computing
-    $\mathbf{L}$ for limb darkening of any order is given in the link next to
-    Equation~(\ref{eq:ld:L}).}
+multiplying each spherical harmonic by the spherical harmonic representation
+of the particular limb darkening law.
 

### 2020-12-09 starry_process/app/css.py
much better loading screen for app

@@ -3,7 +3,7 @@ from bokeh.palettes import Plasma256, Plasma6
 import numpy as np
 import os
 
-__all__ = ["plasma", "svg_mu", "svg_nu", "loader", "description", "style"]
+__all__ = ["plasma", "svg_mu", "svg_nu", "description", "style"]
 
 # jinja template
 TEMPLATE = """
@@ -31,6 +31,20 @@ TEMPLATE = """
   {% block body %}
   <body style="min-width: 1400px; !important">
     <div style="width:90%; height:90%; margin-left: auto; margin-right: auto;">
+
+    <div class="preloader">
+        <div class="spinner">
+            <div class="dot1"></div>
+            <div class="dot2"></div>
+            <div class="loader-message">
+                &nbsp;starry process
+                <div style="font-size: 10pt; font-weight: 100; margin-top: 10px;">
+                    sit tight while the page renders...
+                </div>
+            </div>
+        </div>
+    </div>
+
     {% block inner_body %}
       {% block contents %}
         {% for doc in docs %}
@@ -60,25 +74,6 @@ for v, hex_color in enumerate(Plasma100):
 plasma = plasma[::3]
 plasma = ",\n    ".join(plasma)
 PLASMA = "background: linear-gradient(to left, \n{:s}\n);".format(plasma)
-
-
-# Loading screen
-loader = lambda: Div(
-    text="""
-<div class="preloader">
-    <div class="spinner">
-        <div class="dot1"></div>
-        <div class="dot2"></div>
-        <div class="loader-message">
-            &nbsp;starry process
-            <div style="font-size: 10pt; font-weight: 100; margin-top: 10px;">
-                sit tight while the page renders...
-            </div>
-        </div>
-    </div>
-</div>
-"""
-)
 
 
 # Description textbox
@@ -234,9 +229,10 @@ style = lambda: Div(
         flex-flow: row nowrap;
         justify-content: center;
         align-items: center;
-        background: none repeat scroll 0 0 #ffffff;
+        background: none repeat scroll 0 0 #ffffff; 
     }
     .loader-message {
+        font-family: Helvetica, Arial, sans-serif;
         font-size: 20pt;
         font-weight: 600;
         color: #999;

### 2020-12-09 starry_process/calibrate/generate.py

@@ -5,7 +5,9 @@ from tqdm import tqdm
 
 
 class Star(object):
-    def __init__(self, nlon=300, ydeg=30, linear=True, smoothing=0.1):
+    def __init__(
+        self, nlon=300, ydeg=30, linear=True, smoothing=0.1, eps=1e-12
+    ):
         # Generate a uniform intensity grid
         self.nlon = nlon
         self.nlat = nlon // 2
@@ -20,7 +22,10 @@ class Star(object):
 
         # cos(lat)-weighted SHT
         w = np.cos(self.lat.flatten() * np.pi / 180)
-        PTSinv = P.T * (w ** 2)[None, :]  # @ np.diag(w ** 2)
+        P = self.map.intensity_design_matrix(
+            lat=self.lat.flatten(), lon=self.lon.flatten()
+        )
+        PTSinv = P.T * (w ** 2)[None, :]
         self.Q = np.linalg.solve(PTSinv @ P + eps * np.eye(P.shape[1]), PTSinv)
         if smoothing > 0:
             l = np.concatenate(

### 2020-12-09 starry_process/defaults.py

@@ -2,6 +2,7 @@ defaults = dict(
     ydeg=15,
     udeg=2,
     r=20.0,
+    d=0.0,
     a=0.40,
     b=0.27,
     c=0.1,

### 2020-12-10 tex/gitlinks.tex
misc updates

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/bcfdb4752a408aca0ca122743ff193f5a910bd5a/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/bcfdb4752a408aca0ca122743ff193f5a910bd5a/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/bcfdb4752a408aca0ca122743ff193f5a910bd5a/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/03e335802a72714ed06f39e85f3d9c14193a5e32/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/03e335802a72714ed06f39e85f3d9c14193a5e32/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/03e335802a72714ed06f39e85f3d9c14193a5e32/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-10 starry_process/app/css.py

@@ -32,6 +32,68 @@ TEMPLATE = """
   <body style="min-width: 1400px; !important">
     <div style="width:90%; height:90%; margin-left: auto; margin-right: auto;">
 
+    <style>
+        .preloader {
+            position: fixed;
+            top: 0;
+            left: 0;
+            width: 100%%;
+            height: 100%%;
+            z-index: 99999;
+            display: flex;
+            flex-flow: row nowrap;
+            justify-content: center;
+            align-items: center;
+            background: none repeat scroll 0 0 #ffffff; 
+        }
+        .loader-message {
+            font-family: Helvetica, Arial, sans-serif;
+            font-size: 20pt;
+            font-weight: 600;
+            color: #999;
+            margin: 100px auto;
+            position: relative;
+        }
+        .spinner {
+            margin: 100px auto;
+            width: 320px;
+            height: 320px;
+            position: relative;
+            text-align: center;
+        }
+        .dot1 {
+            width: 25%%;
+            height: 25%%;
+            display: inline-block;
+            position: absolute;
+            left: 120px;
+            top: 0;
+            background-color: #999;
+            border-radius: 100%%;
+            z-index: 1;
+        }
+        .dot2 {
+            display: inline-block;
+            position: absolute;
+            background-color: #666;
+            border-radius: 100%%;
+            width: 10%%;
+            height: 10%%;
+            top: 60px;
+            left: 80px;
+            z-index: 2;
+            -webkit-animation: sk-orbit 0.75s infinite linear;
+            animation: sk-orbit 0.75s infinite linear;
+        }
+        @keyframes sk-orbit {
+            0%% { z-index: 2; transform: translate(0%%, 0%%); }
+            49%% { z-index: 2; transform: translate(400%%, -200%%); }
+            50%% { z-index: 0; transform: translate(400%%, -200%%); }
+            99%% { z-index: 0; transform: translate(0%%, 0%%); }
+            100%% { z-index: 2; transform: translate(0%%, 0%%); }
+        }
+    </style>
+
     <div class="preloader">
         <div class="spinner">
             <div class="dot1"></div>
@@ -216,67 +278,6 @@ style = lambda: Div(
         top: 2px !important;
     }
 
-    /* Loading screen */
-
-    .preloader {
-        position: fixed;
-        top: 0;
-        left: 0;
-        width: 100%%;
[…]

### 2020-12-10 starry_process/calibrate/sample.py

@@ -26,16 +26,17 @@ def sample(data, clobber=False, **kwargs):
     nmax = sample_kwargs["nmax"]
     bmmin = sample_kwargs["bmmin"]
     bmmax = sample_kwargs["bmmax"]
-    bvmin = sample_kwargs["bvmin"]
-    bvmax = sample_kwargs["bvmax"]
+    blvmin = sample_kwargs["blvmin"]
+    blvmax = sample_kwargs["blvmax"]
     fit_bm = sample_kwargs["fit_bm"]
-    fit_bv = sample_kwargs["fit_bv"]
+    fit_blv = sample_kwargs["fit_blv"]
     bm = sample_kwargs["bm"]
     blv = sample_kwargs["blv"]
     sampler = sample_kwargs["sampler"]
     sampler_kwargs = sample_kwargs["sampler_kwargs"]
     run_nested_kwargs = sample_kwargs["run_nested_kwargs"]
     apply_jac = sample_kwargs["apply_jac"]
+    u = sample_kwargs["u"]
 
     # Get the log prob function for the dataset
     t = data["t"]
@@ -47,20 +48,21 @@ def sample(data, clobber=False, **kwargs):
         ferr,
         period,
         ydeg=ydeg,
-        baseline_log_var=None if fit_bv else blv,
+        baseline_log_var=None if fit_blv else blv,
         baseline_mean=None if fit_bm else bm,
         apply_jac=apply_jac,
         normalized=normalized,
         marginalize_over_inclination=True,
+        u=u,
     )
 
     # Extra stuff if we're solving for the baseline
-    if fit_bm and fit_bv:
+    if fit_bm and fit_blv:
         ndim = 7
 
         def baseline_ptform(u, x):
             x[5] = bmmin + u[5] * (bmmax - bmmin)
-            x[6] = bvmin + u[6] * (bvmax - bvmin)
+            x[6] = blvmin + u[6] * (blvmax - blvmin)
 
     elif fit_bm:
         ndim = 6
@@ -68,11 +70,11 @@ def sample(data, clobber=False, **kwargs):
         def baseline_ptform(u, x):
             x[5] = bmmin + u[5] * (bmmax - bmmin)
 
-    elif fit_bv:
+    elif fit_blv:
         ndim = 6
 
         def baseline_ptform(u, x):
-            x[5] = bvmin + u[5] * (bvmax - bvmin)
+            x[5] = blvmin + u[5] * (blvmax - blvmin)
 
     else:
         ndim = 5

### 2020-12-13 tex/gitlinks.tex
new tests

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/03e335802a72714ed06f39e85f3d9c14193a5e32/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/03e335802a72714ed06f39e85f3d9c14193a5e32/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/03e335802a72714ed06f39e85f3d9c14193a5e32/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/5781bddaacfcc41fd945fb4973484b36d4d5e68b/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/5781bddaacfcc41fd945fb4973484b36d4d5e68b/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/5781bddaacfcc41fd945fb4973484b36d4d5e68b/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-13 tex/tests/test_bigErho.py

@@ -1,0 +1,88 @@
+import numpy as np
+from scipy.integrate import quad
+from scipy.special import legendre as P
+
+
+def b(rho, K=1000, s=0.0033 * 180 / np.pi, **kwargs):
+    """
+    The sigmoid spot profile.
+
+    """
+    theta = np.linspace(0, np.pi, K)
+    return 1 / (1 + np.exp((rho - theta) / s)) - 1
+
+
+def get_Bp(K=1000, lmax=5, eps=1e-9, sigma=15, **kwargs):
+    """
+    Return the matrix B+. This expands the 
+    spot profile `b` in Legendre polynomials.
+
+    """
+    theta = np.linspace(0, np.pi, K)
+    cost = np.cos(theta)
+    B = np.hstack(
+        [
+            np.sqrt(2 * l + 1) * P(l)(cost).reshape(-1, 1)
+            for l in range(lmax + 1)
+        ]
+    )
+    BInv = np.linalg.solve(B.T @ B + eps * np.eye(lmax + 1), B.T)
+    l = np.arange(lmax + 1)
+    i = l * (l + 1)
+    S = np.exp(-0.5 * i / sigma ** 2)
+    BInv = S[:, None] * BInv
+    return BInv
+
+
+def bigErho_dzero(r, s=0.0033 * 180 / np.pi, **kwargs):
+    L = (get_Bp(**kwargs) @ b(r)).reshape(-1, 1)
+    return L @ L.T
+
+
+def bigErho(r, d, s=0.0033 * 180 / np.pi, **kwargs):
+    theta = np.linspace(0, np.pi, kwargs.get("K", 1000)).reshape(-1, 1)
+    num = 1 + np.exp((r - d - theta) / s)
+    den = 1 + np.exp((r + d - theta) / s)
+    exp = np.exp((theta - theta.T) / s)
+    C = (exp * np.log(num / den) - np.log(num / den).T) / (1 - exp)
+
+    # When k = kp, we must take the limit, given below
+    C[np.diag_indices_from(C)] = (
+        (1 + np.exp((r + d - theta) / s)) ** -1
+        + (1 + np.exp((d - r + theta) / s)) ** -1
+        - np.log(
+            (np.exp((r - d) / s) + np.exp(theta / s))
+            / (np.exp((r + d) / s) + np.exp(theta / s))
+        )
+        - 1
+    ).flatten()
+    Bp = get_Bp(**kwargs)
[…]

### 2020-12-13 tex/gitlinks.tex
set randomizer seed for test, duh

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/5781bddaacfcc41fd945fb4973484b36d4d5e68b/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/5781bddaacfcc41fd945fb4973484b36d4d5e68b/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/5781bddaacfcc41fd945fb4973484b36d4d5e68b/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/1b565f496cf77b50c5eb4b5279cee8f687f36b27/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/1b565f496cf77b50c5eb4b5279cee8f687f36b27/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/1b565f496cf77b50c5eb4b5279cee8f687f36b27/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-13 tex/ms.tex
much better loading screen for app
new tests
set randomizer seed for test, duh

@@
     \textbf{
-        Interpretable Gaussian Processes for Stellar Variability
+        Interpretable Gaussian Processes for Stellar Variability\\
+        I: Static Surfaces
     }
@@
 \begin{align}
-    k_\mathrm{E^2}(\Delta t) = \sigma^2 e^{-\frac{\Delta t}{2\tau}}
+    k_\mathrm{E^2}(\Delta t) = \sigma^2 \exp\left(-\frac{\Delta t}{2\tau}\right)
     \quad,
@@
 \begin{align}
-    k_\mathrm{M^\frac{3}{2}}(\Delta t) = \sigma^2 \left( 1 + \frac{\sqrt{3}\Delta t}{\tau} \right) e^{-\frac{\sqrt{3}\Delta t}{\tau}}
+    k_\mathrm{M^\frac{3}{2}}(\Delta t) = \sigma^2 \left( 1 + \frac{\sqrt{3}\Delta t}{\tau} \right) \exp\left(-\frac{\sqrt{3}\Delta t}{\tau}\right)
     \quad,
@@
 \begin{align}
-    k_\mathrm{S^2}(\Delta t) = \sigma^2 e^{-\Gamma\sin^2\left( \frac{\pi \Delta t}{P} \right)}
+    k_\mathrm{S^2}(\Delta t) = \sigma^2 \exp\left[-\Gamma\sin^2\left( \frac{\pi \Delta t}{P} \right)\right]
     \quad,
@@
 %
-where $\delta$ is the Kronecker delta function.
-To find the coefficients $\tilde{\mathbf{s}}(\rho)$, we solve the (linear) inverse problem,
+or, in vector form,
+%
+\begin{align}
+    \mathbf{s}(\rho) = \pmb{\mathcal{I}} \, \tilde{\mathbf{s}}(\rho)
+\end{align}
+%
+where $\pmb{\mathcal{I}}$ is a rectangular
+$\left( \left(l_\mathrm{max} + 1\right)^2 \times \left(l_\mathrm{max} + 1\right) \right)$
+identity-like matrix with components
+%
+\begin{align}
+    \mathcal{I}_{n, l} = \delta_{n, l^2 + l}
+    \quad,
+\end{align}
+%
+and $\delta$ is the Kronecker delta function.
+To find the coefficients $\tilde{\mathbf{s}}(\rho)$
+(and hence $\mathbf{s}(\rho)$), we solve the (linear) inverse problem,
 %
@@
 %
-\begin{align}
+\begin{proof}{test_erho}
     (e_\rho)^l_m
@@
     & =
-    \frac{\delta_{m,0} s}{2\Delta}
+    \delta_{m,0}
     \sum_{k=0}^{K-1} B^+_{l,k}
+    c_k(r, \Delta)
+\end{proof}
+%
+where
+%
+\begin{proof}{test_erho}
+    c_k(r, \Delta) & =
+    \frac{s}{2\Delta}
     \ln
@@
     \frac{
-        1+\exp\left(\frac{r -\Delta -\vartheta_k}{s}\right)
+            1+\chi_k^-(r, \Delta)
         }
         {
-        1+\exp\left(\frac{r + \Delta -\vartheta_k}{s}\right)
+            1+\chi_k^+(r, \Delta)
         }
     \right)
+\end{proof}
+%
+and
+\begin{proof}{test_erho}
+    \chi_k^-(r, \Delta) & \equiv \exp\left(\frac{r -\Delta -\vartheta_k}{s}\right)
+    \nonumber                                                              \\
+    \chi_k^+(r, \Delta) & \equiv \exp\left(\frac{r +\Delta -\vartheta_k}{s}\right)
     \quad.
-\end{align}
-%
-In the limit $\Delta \rightarrow 0$,
-%
-\begin{align}
+\end{proof}
+%
+In vector form, we may simply write
+%
+\begin{proof}{test_erho}
+    \mathbf{e}_\rho
+    & =
+    \pmb{\mathcal{I}} \,
+    \mathbf{B}^+ \, \mathbf{c}(r, \Delta)
+    \quad.
+\end{proof}
+%
+Note, finally, that in the limit $\Delta \rightarrow 0$,
+%
+\begin{proof}{test_erho}
     \lim_{\Delta \rightarrow 0}
@@
     & =
+    \pmb{\mathcal{I}} \,
     \mathbf{B}^+ \mathbf{b}(r)
     \quad.
-\end{align}
+\end{proof}
 %
@@
 %
-\begin{align}
+\begin{proof}{test_bigErho}
     (E_\rho)^{l,l'}_{m,m'}
@@
     & =
-    \frac{\delta_{m,0}\delta_{m',0}s}{2\Delta}
+    \delta_{m,0}\delta_{m',0}
     \sum_{k=0}^{K-1} B^+_{l,k}
@@
     \quad,
-\end{align}
+\end{proof}
 %
@@
 %
-\begin{align}
+\begin{proof}{test_bigErho}
     C_{k,k'}(r,\Delta)
     & \equiv
+    \frac{s}{2\Delta}
+    \begin{cases}
+        \displaystyle
         \frac{
@@
             \frac{
-            1+\exp\left(\frac{r -\Delta -\vartheta_{k}}{s}\right)
+                1+\chi_k^-
             }
             {
-            1+\exp\left(\frac{r + \Delta -\vartheta_{k}}{s}\right)
+                1+\chi_k^+
             }
@@
             \frac{
-            1+\exp\left(\frac{r -\Delta -\vartheta_{k'}}{s}\right)
+                1+\chi_{k'}^-
             }
             {
-            1+\exp\left(\frac{r + \Delta -\vartheta_{k'}}{s}\right)
+                1+\chi_{k'}^+
             }
@@
         }
+         & k \ne k'
+        \\[0.5em]
+        \frac{1}{1 + \chi_k^+}
+        +
+        \frac{\chi_k^-}{1 + \chi_k^-}
+        -
+        \ln\left( \frac{
+            1 + \chi_k^-
+        }{
+            1 + \chi_k^+
+        } \right)
+        - 1
+         & k = k'
         \quad.
-\end{align}
-%
-In the limit $\Delta \rightarrow 0$,
-%
-\begin{align}
+    \end{cases}
+\end{proof}
+%
+In vector form, this may be written
+%
+\begin{proof}{test_bigErho}
+    \mathbf{E}_\rho
+    &=
+    \pmb{\mathcal{I}} \,
+    \mathbf{B}^+ \mathbf{C} (r, \Delta) {\mathbf{B}^+}^\top
+    \pmb{\mathcal{I}}^\top
+    \quad.
+\end{proof}
+%
+Finally, in the limit $\Delta \rightarrow 0$,
+%
+\begin{proof}{test_bigErho}
     \lim_{\Delta \rightarrow 0}
@@
     & =
+    \pmb{\mathcal{I}} \,
     \mathbf{B}^+ \mathbf{b}(r) \mathbf{b}^\top(r) {\mathbf{B}^+}^\top
+    \pmb{\mathcal{I}}^\top
     \quad.
-\end{align}
+\end{proof}
 %
@@
 \begin{align}
+    \label{eq:eIk}
     \left(e_I\right)_k
@@
 %
-as the expectation of $\mathbf{y}$ in the polar frame at time $t = t_k$.
+as the expectation of $\mathbf{y}$ in the polar frame at time $t = t_k$.%
+\footnote{In the presence of limb darkening, we must include the limb
+    darkening operator $\mathbf{L}(\mathbf{u})$ in Equation~(\ref{eq:eIk});
+    see \S\ref{sec:ld}.}
 %

### 2020-12-14 tex/gitlinks.tex
more unit tests

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/1b565f496cf77b50c5eb4b5279cee8f687f36b27/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/1b565f496cf77b50c5eb4b5279cee8f687f36b27/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/1b565f496cf77b50c5eb4b5279cee8f687f36b27/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/67220db9d930909e1b8a98ef092fc7b13860bd54/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/67220db9d930909e1b8a98ef092fc7b13860bd54/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/67220db9d930909e1b8a98ef092fc7b13860bd54/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-14 starry_process/size.py

@@ -3,6 +3,8 @@ from .ops import CheckBoundsOp
 import theano.tensor as tt
 import numpy as np
 from scipy.special import legendre as P
+from theano.ifelse import ifelse
+from .math import matrix_sqrt
 
 
 class Spot:
@@ -13,6 +15,7 @@ class Spot:
         eps4=1e-9,
         smoothing=0.075,
         sfac=300,
+        cutoff=1.5,
         **kwargs
     ):
         """
@@ -31,22 +34,64 @@ class Spot:
         l = np.arange(ydeg + 1)
         i = l * (l + 1)
         S = np.exp(-0.5 * i * smoothing ** 2)
-        A = S[:, None] * A
+        Bp = S[:, None] * A
         self.i = i
+        self.ij = np.ix_(i, i)
         self.N = (ydeg + 1) ** 2
         self.theta = tt.as_tensor_variable(theta)
-        self.A = tt.as_tensor_variable(A)
+        self.Bp = tt.as_tensor_variable(Bp)
         self.sfac = sfac
+        self.cutoff = cutoff
 
-    def S(self, r):
+    def b(self, r):
         z = self.sfac * (self.theta - r)
         return 1 / (1 + tt.exp(-z)) - 1
 
     def get_y(self, r):
-        I = self.S(r)
+        b = self.b(r)
         y = tt.zeros(self.N)
-        y = tt.set_subtensor(y[self.i], tt.dot(self.A, I))
+        y = tt.set_subtensor(y[self.i], tt.dot(self.Bp, b))
         return y
+
+    def get_e(self, r, d):
+        chim = tt.exp(self.sfac * (r - d - self.theta))
+        chip = tt.exp(self.sfac * (r + d - self.theta))
+        c = 1.0 / (2 * d * self.sfac) * tt.log((1 + chim) / (1 + chip))
+        e = tt.zeros(self.N)
+        e = tt.set_subtensor(e[self.i], tt.dot(self.Bp, c))
+        return e
+
+    def get_eigE(self, r, d):
+        # NOTE: For theta > r + d, the `C` matrix drops
+        # to zero VERY quickly. In practice we get better
+        # numerical stability if we just set those elements
+        # to zero without evaluating them, especially since
+        # we can get NaNs from operations involving the extremely
+        # large dynamic range between the `exp` and `ln` terms.
+        # TODO: Find a more numerically stable way of evaluating this.
[…]

### 2020-12-14 tex/gitlinks.tex
more unit tests, fixed eqn typos

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/67220db9d930909e1b8a98ef092fc7b13860bd54/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/67220db9d930909e1b8a98ef092fc7b13860bd54/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/67220db9d930909e1b8a98ef092fc7b13860bd54/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/ccd46f868c6bcdeb26f56055d8dc023476fcd21b/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/ccd46f868c6bcdeb26f56055d8dc023476fcd21b/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/ccd46f868c6bcdeb26f56055d8dc023476fcd21b/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-14 tex/tests/test_bigElam.py

@@ -1,9 +1,7 @@
 import numpy as np
 from scipy.special import gamma as Gamma
-from scipy.special import beta as B
 from scipy.special import hyp2f1
 from scipy.integrate import quad
-from scipy.stats import beta as Beta
 from wigner import R
 from tqdm.auto import tqdm
 
@@ -60,16 +58,20 @@ def clmmpi(l, m, mp, i):
         return 0.0
 
 
-def Qllpiip(l, lp, i, ip, alpha, beta):
+def Qllpiip(l, lp, i, ip):
     if (i + ip) % 2 == 0:
-        return B(alpha, l + lp + beta - (i + ip) / 2) * hyp2f1(
-            -(i + ip) / 2, alpha, l + lp + alpha + beta - (i + ip) / 2, -1
+        return (
+            2 ** (l + lp)
+            / np.pi
+            * Gamma(0.5 * (1 + i + ip))
+            * Gamma(l + lp + 0.5 * (1 - i - ip))
+            / Gamma(l + lp + 1)
         )
     else:
         return 0.0
 
 
-def Pllpmmp(l, lp, m, mp, alpha, beta, Ellpr):
+def Pllpmmp(l, lp, m, mp, Ellpphi):
     term1 = 0
     for mu in range(-l, l + 1):
         for mup in range(-lp, lp + 1):
@@ -79,73 +81,70 @@ def Pllpmmp(l, lp, m, mp, alpha, beta, Ellpr):
                     term2 += (
                         clmmpi(l, m, mu, i)
                         * clmmpi(lp, mp, mup, ip)
-                        * Qllpiip(l, lp, i, ip, alpha, beta)
+                        * Qllpiip(l, lp, i, ip)
                     )
-            term1 += (
-                Ellpr[l + mu, lp + mup]
-                * np.exp(1j * np.pi / 2 * (m - mu + mp - mup))
-                * term2
-            )
-    return term1 / B(alpha, beta)
+            term1 += Ellpphi[l + mu, lp + mup] * term2
+    return term1
 
 
-def Pllp(l, lp, alpha, beta, Ellpr):
+def Pllp(l, lp, Ellpphi):
     P = np.zeros((2 * l + 1, 2 * lp + 1), dtype="complex128")
     for m in range(-l, l + 1):
[…]

### 2020-12-14 tex/gitlinks.tex
more proofs

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/ccd46f868c6bcdeb26f56055d8dc023476fcd21b/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/ccd46f868c6bcdeb26f56055d8dc023476fcd21b/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/ccd46f868c6bcdeb26f56055d8dc023476fcd21b/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/080b9bdd98a2b9bbda73d188b36db3e9247ed3db/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/080b9bdd98a2b9bbda73d188b36db3e9247ed3db/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/080b9bdd98a2b9bbda73d188b36db3e9247ed3db/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-15 tex/gitlinks.tex
fix size integral

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/080b9bdd98a2b9bbda73d188b36db3e9247ed3db/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/080b9bdd98a2b9bbda73d188b36db3e9247ed3db/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/080b9bdd98a2b9bbda73d188b36db3e9247ed3db/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/edcf6a79f67f20eaa9e2e3caee030909a2a3d5f2/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/edcf6a79f67f20eaa9e2e3caee030909a2a3d5f2/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/edcf6a79f67f20eaa9e2e3caee030909a2a3d5f2/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-15 starry_process/size.py

@@ -101,29 +101,32 @@ class SizeIntegral(MomentIntegral):
         set up the transform and rotation operators.
 
         """
-        # Ingest it
-        self._r = CheckBoundsOp(name="r", lower=0, upper=0.5 * np.pi)(
-            r * self._angle_fac
-        )
-        self._d = CheckBoundsOp(name="d", lower=0, upper=0.5 * np.pi)(
-            d * self._angle_fac
-        )
-        self._params = [self._r, self._d]
-
         # Set up the spot operator
         self._spot = Spot(ydeg=self._ydeg, **kwargs)
 
-        # Compute the moments
-        self._q = ifelse(
-            tt.eq(self._d, 0),
-            self._spot.get_y(self._r),
-            self._spot.get_e(self._r, self._d),
+        # Ingest params
+        self._r = CheckBoundsOp(name="r", lower=0, upper=0.5 * np.pi)(
+            r * self._angle_fac
         )
-        self._eigQ = ifelse(
-            tt.eq(self._d, 0),
-            tt.set_subtensor(tt.zeros((self._nylm, 2))[:, 0], self._q),
-            self._spot.get_eigE(self._r, self._d),
-        )
+
+        # Different behavior depending on whether we're
+        # modeling a delta function or a uniform prior
+        if d is None:
+
+            # delta function
+            self._params = [self._r]
+            self._q = self._spot.get_y(self._r)
+            self._eigQ = tt.reshape(self._q, (-1, 1))
+
+        else:
+
+            # uniform
+            self._d = CheckBoundsOp(name="d", lower=0, upper=0.5 * np.pi)(
+                d * self._angle_fac
+            )
+            self._params = [self._r, self._d]
+            self._q = self._spot.get_e(self._r, self._d)
+            self._eigQ = self._spot.get_eigE(self._r, self._d)
 
     def _compute(self):
         pass

### 2020-12-15 tex/gitlinks.tex
more figures

@@
-\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/edcf6a79f67f20eaa9e2e3caee030909a2a3d5f2/tex/figures/#1.py}{\codeicon}\,\,}
-\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/edcf6a79f67f20eaa9e2e3caee030909a2a3d5f2/tex/figures/#1.gif}{\animicon}\,\,}
-\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/edcf6a79f67f20eaa9e2e3caee030909a2a3d5f2/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
+\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/8cc07b6268e28ce42b1ce63cb09147746d49c236/tex/figures/#1.py}{\codeicon}\,\,}
+\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/8cc07b6268e28ce42b1ce63cb09147746d49c236/tex/figures/#1.gif}{\animicon}\,\,}
+\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/8cc07b6268e28ce42b1ce63cb09147746d49c236/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
 \newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-12-15 tex/ms.tex
more unit tests
more unit tests, fixed eqn typos
more proofs
fix size integral
more figures

@@
 
-\features{open-source figures \codeicon; equation unit tests \testpassicon}
+\features{open-source figures \codeicon; equation unit tests: \input{tests/tally}}
 
@@
 %
+
+\xxx{Brief paper outline here.}
 
@@
 
-\subsection{Ensemble Analyses}
+\subsection{Ensemble analyses}
 \label{sec:ensemble}
@@
             for a hypothetical scenario in which an observer can measure a
-            stellar light curve at a single random inclination (thin blue curves),
-            at 3 random inclinations (thin orange curves), and at 10 random
-            inclinations (thin green curves). The thicker curves correspond to
+            stellar light curve at 1 (thin blue curves),
+            3 (thin orange curves), 10 (thin green curves), and 30 (thin red
+            curves) random inclinations. The thicker curves correspond to
             the mean shrinkage for each experiment.
@@
 light curves of ``similar'' stars?
+
+\subsection{Limb darkening}
+
+See Figures~\ref{fig:nullspace_ensemble_ld} and \ref{fig:nullspace_ensemble_ld_var}.
+
+\begin{figure}[ht!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/nullspace_ensemble_ld.pdf}
+        \oscaption{nullspace_ensemble_ld}{%
+            Same as Figure~\ref{fig:nullspace_ensemble}, but for limb-darkened
+            stars with coefficients $u_1 = 0.5$ and $u_2 = 0.25$.
+            \label{fig:nullspace_ensemble_ld}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[ht!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/nullspace_ensemble_ld_var.pdf}
+        \oscaption{nullspace_ensemble_ld_var}{%
+            Same as Figure~\ref{fig:nullspace_ensemble_ld}, but allowing for
+            a 10\% variation in each of the two limb darkening coefficients
+            across different observations.
+            \label{fig:nullspace_ensemble_ld_var}
+        }
+    \end{centering}
+\end{figure}
 
@@
 \subsection{With limb darkening}
+% http://dx.doi.org/10.1051/0004-6361/200913675
 %
@@
                 \dfrac{
-                \left(-1\right)^{\frac{2l + 3m + m' - i}{2}}
+                \left(-1\right)^{\frac{2l - m + m' - i}{2}}
                 \sqrt{\left(l - m\right)!
@@
 %
-\begin{proof}{}
+\begin{proof}{test_bigEphi}
     \label{eq:Ephi}
@@
 %
-\begin{proof}{}
+\begin{proof}{test_bigEphi}
     \mathbf{E}_\phi^{l,l'}
@@
     {(\bar{E}^{l,l'}_r)_{}}_{\mu,{\mu'}} \,
-    {({D^{l'}_{\hat{\mathbf{x}}}})_{}}_{{\mu'},m'}(\phi) \,
+    {({D^{l'}_{\hat{\mathbf{x}}}})_{}}_{m',{\mu'}}(\phi) \,
     p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
@@
         {(\bar{E}^{l,l'}_r)_{}}_{\mu,{\mu'}} \,
-        \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu + {\mu'} - m')}
+        \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu + m' - {\mu'})}
         \sum\limits_{i=0}^{2l}
@@
         c_{m,\mu,i}^{l}
-        c_{{\mu'},m',i'}^{l'}
+        c_{m',{\mu'},i'}^{l'}
         \,
@@
 %
-\begin{proof}{}
+\begin{proof}{test_bigEphi}
     {({Q^{l,l'}_\phi})_{}}_{i,i'}(\pmb{\theta}_{\phi})
@@
 %
-\begin{proof}{}
+\begin{proof}{test_bigEphi}
     %
@@
 %
-\begin{proof}{}
+\begin{proof}{test_lat_transforms}
     \mu_\phi &= 2 \tan^{-1}
@@
 %
-\begin{proof}{}
+\begin{proof}{test_lat_transforms}
     \sigma_\phi &= \frac{\sin\mu_\phi}{
@@
 %
-\begin{proof}{}
+\begin{proof}{test_lat_transforms}
     \alpha &= \frac{2 + 4 \sigma_\phi^2 + (3 + 8 \sigma_\phi^2) \cos\mu_\phi + 2 \cos\left( 2\mu_\phi \right) + \cos\left( 3\mu_\phi \right)}{16 \sigma_\phi^2 \cos\left( \frac{\mu_\phi}{2} \right)^4} \\
@@
 %
-\begin{proof}{}
+\begin{proof}{test_lat_jac}
     J &=
@@
             \frac{
+                \left(K_{10} - K_{00}\right) \left(K_{11} - K_{10}\right)
                 \alpha \beta \left(1 + \cos\mu_\phi\right)^3
@@
 %
-\begin{proof}{}
+\begin{proof}{test_bigElam}
     \mathbf{E}_\lambda
@@
 %
-\begin{proof}{}
+\begin{proof}{test_bigElam}
     \mathbf{E}_\lambda^{l,l'}
@@
     (\bar{E}^{l,l'}_\phi)_{\mu,{\mu'}} \,
-    ({D^{l'}_{\hat{\mathbf{x}}}})_{{\mu'},m'}(\lambda) \,
+    ({D^{l'}_{\hat{\mathbf{x}}}})_{m',{\mu'}}(\lambda) \,
     p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
@@
     c_{m,\mu,i}^{l}
-    c_{{\mu'},m',i'}^{l'}
+    c_{m',{\mu'},i'}^{l'}
     \,
@@
 %
-\begin{proof}{}
+\begin{proof}{test_bigElam}
     (Q^{l,l'}_\lambda)_{i,i'}
@@
 %
-\begin{proof}{}
+\begin{align}
     \mathbf{e}_I
@@
     \mathrm{d}I
-\end{proof}
+\end{align}
 %
@@
 %
-\begin{proof}{}
+\begin{align}
     \mathbf{e}_y
@@
     \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
-\end{proof}
+\end{align}
 %
@@
 %
-\begin{proof}{}
+\begin{align}
     \left(\mathbf{e}_{y'}\right)_k
-    &=
+     & =
     \left(\mathbf{e}_{y'}\right)_0
     \nonumber                 \\
-    &=
+     & =
     \mathbf{R}_{\hat{\mathbf{x}}}\left(\frac{\pi}{2}\right) \,
@@
     \nonumber                 \\
-    &\equiv \mathbf{e}_{y'}
+     & \equiv \mathbf{e}_{y'}
     \quad.
-\end{proof}
+\end{align}
 %
@@
     e_I \, \mathbf{1}
-\end{proof}
+\end{align}
 %
@@
 %
-where
+and
 %
@@
 %
-\begin{proof}{}
+\begin{proof}{test_eI}
     \mathbf{e}_{y''}^l
@@
 %
-\begin{align}
+\begin{proof}{test_eI}
     {({p^l_I})_{}}_m
@@
 %
-\begin{proof}{}
+\begin{proof}{test_eI}
     {({q^l_I})_{}}_i
@@
     %
-    (-1)^{i + 1}
+    (-1)^{i}
     \displaystyle\int_{0}^{1}
@@
     &=
-    \frac{(-1)^{i+1}}{l-\frac{i}{2}+1}
+    \frac{(-1)^{i}}{l-\frac{i}{2}+1}
     \,
@@
 %
-\begin{proof}{}
+\begin{align}
     \mathbf{E}_I
@@
 %
-\begin{proof}{}
+\begin{align}
     \mathbf{E}_y
@@
     \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
-\end{proof}
+\end{align}
 %
@@
 %
-\begin{proof}{}
+\begin{proof}{test_bigEI}
     \resizebox{.85\hsize}{!}{$
@@
 %
-\begin{proof}{}
+\begin{proof}{test_bigEI}
     \left(\mathbf{E}_{y''}^{l,l'}\right)_{k,k'}
@@
         \left[{\left({\bar{E}^{l,l'}_{y'}}\right)_{}}_{k,k'}\right]_{\mu,\mu'}
-        \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu + {\mu'} - m')}
+        \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu + m' - {\mu'})}
         \sum\limits_{i=0}^{2l}
@@
 %
-\begin{proof}{}
+\begin{proof}{test_bigEI}
     \label{eq:QllpIiip}
@@
     (-1)^{i+i'}
-    (1 - \cos\phi)^{l + l' - \frac{i+i'}{2}}
-    (1 + \cos\phi)^\frac{i+i'}{2}
+    (1 - \cos I)^{l + l' - \frac{i+i'}{2}}
+    (1 + \cos I)^\frac{i+i'}{2}
     \sin I
@@
     %
-    (-1)^{i+i'+1}
-    \int_{0}^{\frac{\pi}{2}}
+    (-1)^{i+i'}
+    \int_{0}^{1}
     %
@@
     \,
-    \mathrm{d}I
+    \mathrm{d}x
     \nonumber \\[0.5em]
@@
     %
-    \frac{(-1)^{i+i'+1}}{l+l'-\frac{i+i'}{2}+1}
+    \frac{(-1)^{i+i'}}{l+l'-\frac{i+i'}{2}+1}
     \,

### 2020-12-15 starry_process/calibrate/run.py

@@ -30,7 +30,12 @@ def run(
             open(os.path.join(path, "kwargs.json"), "w"),
         )
     else:
-        kwargs = json.load(open(os.path.join(path, "kwargs.json"), "r"))
+        input_kwargs = defaults.update_with_defaults(**kwargs)
+        saved_kwargs = json.load(open(os.path.join(path, "kwargs.json"), "r"))
+        assert (
+            input_kwargs == saved_kwargs
+        ), "Input kwargs don't match saved kwargs for this run."
+        kwargs = saved_kwargs
 
     # Generate
     if clobber or not os.path.exists(os.path.join(path, "data.npz")):
@@ -41,8 +46,9 @@ def run(
 
     # Plot the data
     if plot_all or plot_data:
-        fig = plot.plot_data(data, **kwargs)
-        fig.savefig(os.path.join(path, "data.pdf"), bbox_inches="tight")
+        if clobber or not os.path.exists(os.path.join(path, "data.pdf")):
+            fig = plot.plot_data(data, **kwargs)
+            fig.savefig(os.path.join(path, "data.pdf"), bbox_inches="tight")
 
     # Sample
     if clobber or not os.path.exists(os.path.join(path, "results.pkl")):
@@ -69,36 +75,53 @@ def run(
         inc_results = None
 
     # Transform latitude params and store posterior mean and cov
-    samples = np.array(results.samples)
-    samples[:, 1], samples[:, 2] = beta2gauss(samples[:, 1], samples[:, 2])
-    try:
-        weights = np.exp(results["logwt"] - results["logz"][-1])
-    except:
-        weights = results["weights"]
-    mean, cov = dyfunc.mean_and_cov(samples, weights)
-    np.savez(os.path.join(path, "mean_and_cov.npz"), mean=mean, cov=cov)
+    if clobber or not os.path.exists(os.path.join(path, "mean_and_cov.npz")):
+        samples = np.array(results.samples)
+        samples[:, 1], samples[:, 2] = beta2gauss(samples[:, 1], samples[:, 2])
+        try:
+            weights = np.exp(results["logwt"] - results["logz"][-1])
+        except:
+            weights = results["weights"]
+        mean, cov = dyfunc.mean_and_cov(samples, weights)
+        np.savez(os.path.join(path, "mean_and_cov.npz"), mean=mean, cov=cov)
+    else:
+        mean_and_cov = np.load(os.path.join(path, "mean_and_cov.npz"))
+        mean = mean_and_cov["mean"]
[…]

### 2020-12-16 starry_process/calibrate/inclination.py
new makefile structure

@@ -3,6 +3,7 @@ from .log_prob import get_log_prob
 import numpy as np
 from tqdm.auto import tqdm
 from dynesty import utils as dyfunc
+import os
 
 
 def compute_inclination_pdf(data, results, **kwargs):
@@ -62,7 +63,7 @@ def compute_inclination_pdf(data, results, **kwargs):
     samples = dyfunc.resample_equal(samples, weights)
 
     # Compute the posteriors
-    for n in tqdm(range(nlc)):
+    for n in tqdm(range(nlc), disable=bool(int(os.getenv("NOTQDM", "0")))):
         for j in range(ninc_samples):
             idx = np.random.randint(len(samples))
             lp[n, j] = np.array(

### 2020-12-16 tex/figures/calibration.py

@@ -2,8 +2,9 @@ from starry_process import calibrate
 import numpy as np
 import os
 
-# TODO: Not yet ready for Azure runs
-if not bool(os.getenv("MAKEFILE", False)):
+
+# TODO: Not yet ready for CI runs
+if not bool(os.getenv("CI", False)):
 
     # Default run
     calibrate.run(path="data/default")

### 2020-12-16 tex/figures/nullspace_ensemble_ld.py

@@ -1,8 +1,8 @@
 import starry
 import matplotlib.pyplot as plt
 import numpy as np
-from tqdm import tqdm
 import os
+from tqdm import tqdm
 
 # Settings
 ntheta = 50
@@ -16,14 +16,20 @@ kpn = 300
 clobber = False
 
 # Compute
-DATA_FILE = os.path.abspath(os.path.join("data", "nullspace_ensemble_ld.npz"))
+DATA_FILE = (
+    os.path.abspath(__file__)
+    .replace("figures", "figures/data")
+    .replace(".py", ".npz")
+)
 if clobber or not os.path.exists(DATA_FILE):
     map = starry.Map(ydeg=ydeg + ydeg_pad, udeg=len(u), lazy=False)
     map[1:] = u
     theta = np.linspace(0, 360, ntheta, endpoint=False)
     S = np.empty((len(ninc), kpn, map.Ny))
     np.random.seed(0)
-    for n in tqdm(range(len(ninc))):
+    for n in tqdm(
+        range(len(ninc)), disable=bool(int(os.getenv("NOTQDM", "0")))
+    ):
         for k in range(kpn):
             A = np.empty((0, map.Ny))
             for _ in range(ninc[n]):
@@ -44,7 +50,7 @@ else:
 
 # Plot
 fig, ax = plt.subplots(1)
-for n in tqdm(range(len(ninc))):
+for n in tqdm(range(len(ninc)), disable=bool(int(os.getenv("NOTQDM", "0")))):
     for k in range(kpn):
         ax.plot(S[n, k], color="C{}".format(n), lw=0.75, alpha=0.01, zorder=-1)
     ax.plot(
@@ -76,4 +82,8 @@ print(
     np.mean(S[:, :, : (ydeg + 1) ** 2], axis=(1, 2)),
 )
 
-fig.savefig(__file__.replace("py", "pdf"), bbox_inches="tight", dpi=300)
+fig.savefig(
+    os.path.abspath(__file__).replace(".py", ".pdf"),
+    bbox_inches="tight",
+    dpi=300,
+)

### 2020-12-16 tex/figures/ylms.py

@@ -1,6 +1,7 @@
 import starry
 import matplotlib.pyplot as plt
 import numpy as np
+import os
 
 # Settings
 lmax = 5
@@ -62,4 +63,8 @@ for i, l in enumerate(range(lmax + 1)):
         ax[i, j].set_rasterization_zorder(1)
 
 # We're done
-fig.savefig(__file__.replace(".py", ".pdf"), bbox_inches="tight", dpi=300)
+fig.savefig(
+    os.path.abspath(__file__).replace(".py", ".pdf"),
+    bbox_inches="tight",
+    dpi=300,
+)

### 2020-12-16 tex/tests/tally.py

@@ -1,21 +1,0 @@
-import glob
-
-test_files = glob.glob("test_*.tex")
-
-passed = 0
-failed = 0
-for file in test_files:
-    with open(file, "r") as f:
-        l = f.readline()
-    if r"\testpassicon" in l:
-        passed += 1
-    elif r"\testfailicon" in l:
-        failed += 1
-
-with open("tally.tex", "w") as f:
-    print(
-        r"{} passed \testpassicon, {} failed \testfailicon".format(
-            passed, failed
-        ),
-        file=f,
-    )

### 2020-12-16 tex/tests/test_bigEI.py

@@ -6,6 +6,7 @@ from wigner import R
 from starry_process.ops import rTA1Op
 import starry
 from tqdm.auto import tqdm
+import os
 
 
 def factorial(n):
@@ -60,18 +61,6 @@ def clmmpi(l, m, mp, i):
 
 def Qllpiip(l, lp, i, ip):
     """Return the scalar (Q_I^{l,l'})_{i,i'}."""
-    """
-    # DEBUG
-    l += lp
-    i += ip
-    f = (
-        lambda x: (-1) ** i
-        * (1 - np.cos(x)) ** (l - i / 2)
-        * (1 + np.cos(x)) ** (i / 2)
-        * np.sin(x)
-    )
-    return quad(f, 0, 0.5 * np.pi)[0]
-    """
     return (
         (-1) ** (i + ip)
         / (l + lp - 0.5 * (i + ip) + 1)
@@ -127,7 +116,7 @@ def bigEI(Ey, t, P=1):
     Rx = R(lmax, 0.5 * np.pi, 0.5 * np.pi, -0.5 * np.pi)
     bigEI = np.zeros(K, dtype="complex128")
     kp = 0
-    for k in tqdm(range(K)):
+    for k in tqdm(range(K), disable=bool(int(os.getenv("NOTQDM", "0")))):
         Rzk = R(lmax, 2 * np.pi / P * t[k], 0, 0)
         Rzkp = R(lmax, 2 * np.pi / P * t[kp], 0, 0)
         for l in range(lmax + 1):
@@ -153,7 +142,7 @@ def bigEI_numerical(Ey, t, P=1):
 
     bigEI = np.zeros(K)
     kp = 0
-    for k in tqdm(range(K)):
+    for k in tqdm(range(K), disable=bool(int(os.getenv("NOTQDM", "0")))):
 
         def integrand(I):
             map.inc = I * 180 / np.pi

### 2020-12-16 tex/tests/test_bigEphi.py

@@ -6,6 +6,7 @@ from scipy.integrate import quad
 from scipy.stats import beta as Beta
 from wigner import R
 from tqdm.auto import tqdm
+import os
 
 
 def factorial(n):
@@ -109,7 +110,9 @@ def bigEphi(alpha, beta, Er):
     lmax = int(np.sqrt(Er.shape[0]) - 1)
     N = (lmax + 1) ** 2
     E = np.zeros((N, N), dtype="complex128")
-    for l in tqdm(range(lmax + 1)):
+    for l in tqdm(
+        range(lmax + 1), disable=bool(int(os.getenv("NOTQDM", "0")))
+    ):
         for lp in range(lmax + 1):
             i = slice(l ** 2, (l + 1) ** 2)
             j = slice(lp ** 2, (lp + 1) ** 2)
@@ -122,7 +125,7 @@ def bigEphi_numerical(alpha, beta, Er):
     lmax = int(np.sqrt(Er.shape[0]) - 1)
     N = (lmax + 1) ** 2
     E = np.zeros((N, N))
-    for n1 in tqdm(range(N)):
+    for n1 in tqdm(range(N), disable=bool(int(os.getenv("NOTQDM", "0")))):
         for n2 in range(N):
 
             def func(phi):

### 2020-12-16 tex/ms.tex
new makefile structure
more text

@@
     \label{eq:fAy}
-    \mathbf{f} = \mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y}
+    \mathbf{f} = \mathbf{1} + \pmb{\mathcal{A}}(I, P, \mathbf{u}) \, \mathbf{y}
     \quad,
@@
 design matrix, a purely linear operator that transforms from the spherical
-harmonic basis to the flux basis; it is implicitly
+harmonic basis to the flux basis; it is
 a function of the stellar inclination $I$, the stellar
-rotation period $P$, the stellar limb darkening coefficients, etc.
+rotation period $P$, and the stellar limb darkening coefficients $\mathbf{u}$
 (see Appendix~\ref{sec:starry} for details).
@@
 $\mathbf{f} = \mathbf{1}$.
-Conditioned on certain physical properties of the
-star
-(such as its inclination, rotation period, etc.), $\pmb{\theta}_\star$,
-and on certain properties of the  star spots
-(such as their sizes, positions, etc.), $\pmb{\theta}_\bullet$,
-we wish to compute the mean $\pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$ and
-covariance $\pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$
+Conditioned on the stellar inclination $I$, the rotational period $P$,
+a set of limb darkening coefficients $\mathbf{u}$, and
+on certain properties of the star spots $\pmb{\theta}_\bullet$
+(including their number, sizes, positions, and contrasts),
+we wish to compute the mean $\pmb{\mu}(I, P, \mathbf{u}, \pmb{\theta}_\bullet)$ and
+covariance $\pmb{\Sigma}(I, P, \mathbf{u}, \pmb{\theta}_\bullet)$
 of $\mathbf{f}$. Together, these specify a multidimensional Gaussian
@@
 \footnote{The \emph{true} distribution of stellar light curves conditioned
-    on $\pmb{\theta}_\star$ and $\pmb{\theta}_\bullet$ is not Gaussian, so our
+    on $I$, $P$, $\mathbf{u}$, and $\pmb{\theta}_\bullet$ is not Gaussian, so our
     assumption is formally wrong. But, as the saying goes, \emph{all models are
@@
 \begin{align}
-    \mathbf{f}\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right) \sim
+    \mathbf{f}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right) \sim
     \mathcal{N}\Big(
-    \pmb{\mu}\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right),
+    \pmb{\mu}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right),
     \,
-    \pmb{\Sigma}\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+    \pmb{\Sigma}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
     \Big)
@@
     \label{eq:mean}
-    \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
-     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \pmb{\theta}_\star, \pmb{\theta}_\bullet \Big]
+    \pmb{\mu}(I, P, \mathbf{u}, \pmb{\theta}_\bullet)
+     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, I, P, \mathbf{u}, \pmb{\theta}_\bullet \Big]
     \\
     \label{eq:cov}
-    \pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
-     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \pmb{\theta}_\star, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet) \pmb{\mu}^\top(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
+    \pmb{\Sigma}(I, P, \mathbf{u}, \pmb{\theta}_\bullet)
+     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, I, P, \mathbf{u}, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(I, P, \mathbf{u}, \pmb{\theta}_\bullet) \pmb{\mu}^\top(I, P, \mathbf{u}, \pmb{\theta}_\bullet)
     \quad.
@@
     \label{eq:mean_f}
-    \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
-     & = \mathbf{1} + \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
+    \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)
+     & = \mathbf{1} + \pmb{\mathcal{A}}(I, P, \mathbf{u}) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
     \\
     \label{eq:cov_f}
-    \pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
-     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet) \, \pmb{\mathcal{A}}^\top(\pmb{\theta}_\star)
+    \pmb{\Sigma}(P, \mathbf{u}, \pmb{\theta}_\bullet)
+     & = \pmb{\mathcal{A}}(I, P, \mathbf{u}) \, \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet) \, \pmb{\mathcal{A}}^\top(I, P, \mathbf{u})
     \quad,
@@
 dataset
-conditioned on a specific value of $\pmb{\theta}_\bullet$ (and $\pmb{\theta}_\star$):
+conditioned on a specific value of $\pmb{\theta}_\bullet$ (as well as $I$,
+$P$, and $\mathbf{u}$):
 %
@@
     \label{eq:log-like}
-    \resizebox{.9\hsize}{!}{$
-        \ln \mathcal{L}_m\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+    \ln \mathcal{L}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
     =
-        -\frac{1}{2}
-        \mathbf{r}_m^\top\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
-        \mathbf{K}_m^{-1}\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
-        \mathbf{r}_m\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
-        -
+     & -\frac{1}{2}
+    \mathbf{r}_m^\top\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
+    \mathbf{K}_m^{-1}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
+    \mathbf{r}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
+    \nonumber       \\[0.75em]
+     & -
     \frac{1}{2}
-        \ln \Big| \mathbf{K}_m\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right) \Big|
+    \ln \Big| \mathbf{K}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right) \Big|
     -
@@
     \quad,
-    $}
 \end{align}
@@
 \begin{align}
-    \mathbf{r}_m\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
-     & \equiv
-    \mathbf{f}_m - \pmb{\mu}\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+    \mathbf{r}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
+     & \equiv
+    \mathbf{f}_m - \pmb{\mu}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
 \end{align}
@@
 \begin{align}
-    \mathbf{K}_m\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
-     & \equiv
-    \pmb{\Sigma}\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+    \mathbf{K}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
+     & \equiv
+    \pmb{\Sigma}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
     +
@@
 \begin{align}
-    \ln \mathcal{L}\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
-     & =
-    \sum_{m} \ln \mathcal{L}_m\left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+    \ln \mathcal{L}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
+     & =
+    \sum_{m} \ln \mathcal{L}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
     \quad.
@@
 (Equations~\ref{eq:mean_f} and \ref{eq:cov_f}, respectively) are conditioned
-on specific values of the stellar properties $\pmb{\theta}_\star$ and the
-spot properties $\pmb{\theta}_\bullet$. To obtain the posterior distribution
+on specific values of the stellar and spot properties. To obtain the posterior distribution
 for these parameters, we must typically resort to numerical sampling techniques,
@@
 employing the GP in an ensemble setting.
-In such a setting, we might have light curves from $N_\star$ stars, all of which
+In such a setting, we might have light curves from $M$ stars, all of which
 we believe to have similar spot properties (perhaps because they have
@@
 \begin{align}
-    N_\theta = N_\star \, \mathrm{dim}(\pmb{\theta}_\star) + 5
-    \quad,
-\end{align}
-%
-since each of the stars will have their own set of stellar properties
-$\pmb{\theta}_\star$ but will all share the same 5 spot properties
+    N_\theta = 4 M + 5
+    \quad,
+\end{align}
+%
+since each of the stars will have their own set of 4 stellar properties
+(an inclination, a period, and usually two limb darkening coefficients)
+but will all share the same 5 spot properties
 $\pmb{\theta}_\bullet$ (by assumption).
-In a typical problem, the stellar parameters might be
-%
-\begin{align}
-    \pmb{\theta}_\star =
-    \left(
-    I \,\,\,
-    P \,\,\,
-    u_1 \,\,\,
-    u_2
-    \right)^\top
-    \quad,
-\end{align}
-%
-i.e., the inclination, the rotational period, and some (usually two) limb
-darkening coefficients. Thus, for a reasonably sized ensemble of $N_\star=100$
+For a reasonably sized ensemble of $M=100$
 stars, we would have to sample over $N_\theta = 405$ parameters.
@@
 
-One workaround is to fix the values of $\pmb{\theta}_\star$. This could be done,
+One workaround is to fix the values of the stellar parameters. This could be done,
 for instance, to the rotational period $P$, which can often be estimated with
@@
 for a particular star (such as a measurement of its projected rotational
-velocity $v\sin i$ or the knowledge that it hosts a transiting planet), it
+velocity $v\sin I$ or the knowledge that it hosts a transiting planet), it
 is simply not possible to reliably estimate the inclination in a
@@
 degenerate with the spot parameters $\pmb{\theta}_\bullet$. If one knew
-$\pmb{\theta}_\bullet$, then perhaps a decent point estimate of $i$ could
+$\pmb{\theta}_\bullet$, then perhaps a decent point estimate of $I$ could
 be obtained, but in that case the analysis wouldn't be needed in the first place!
@@
 the problem: we can explicitly marginalize over the stellar inclination.
-Let
-%
-\begin{align}
-    \label{eq:thetastarbar}
-    \bar{\pmb{\theta}}_\star = \left(
-    P \,\,\,
-    u_1 \,\,\,
-    u_2
-    \right)^\top
-    \quad,
-\end{align}
-%
-be the vector of all stellar parameters \emph{except}
-the inclination, where perhaps $P$ is fixed for each star and the limb darkening coefficients
-are shared among all stars, as described above.
-Then we can write the first two moments of our flux GP as
-%
-\begin{align}
-    \label{eq:einc}
-    \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
-     & =
-    \mathrm{E} \Big[ \mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
-    \nonumber \\[0.5em]
-     & =
-    \mathbf{1} + \mathbf{e}_I
-\end{align}
-%
-where we define the first moment integral
-%
-\begin{align}
-    \label{eq:eI}
-    \mathbf{e}_I
-     & \equiv
-    \mathrm{E} \Big[ \pmb{\mathcal{A}} \, \mathbf{y} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
-    \nonumber \\[0.5em]
-     & =
-    \int
-    \pmb{\mathcal{A}}(I, \bar{\pmb{\theta}}_\star) \,
-    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
-    p(I) \, \mathrm{d}I
-\end{align}
-%
-and
-%
-\begin{align}
-    \label{eq:Einc}
-    \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
-     & =
-    \mathrm{E} \Big[ \left(\mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y}\right) \left(\mathbf{1} + \pmb{\mathcal{A}} \, \mathbf{y}\right)^\top  \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
-    \nonumber \\[0.5em]
-     & =
-    \mathbf{1} \, \mathbf{1}^\top +
-    \mathbf{1} \, \mathbf{e}_I^\top +
-    \mathbf{e}_I \, \mathbf{1}^\top +
-    \mathbf{E}_I
-\end{align}
-%
-where we define the second moment integral
-%
-\begin{align}
-    \label{eq:EI}
-    \mathbf{E}_I
-     & \equiv
-    \int
-    \pmb{\mathcal{A}}(I, \bar{\pmb{\theta}}_\star) \,
-    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
-    \pmb{\mathcal{A}}^\top(I, \bar{\pmb{\theta}}_\star)
-    p(I) \, \mathrm{d}I
-    \quad.
-\end{align}
-%
-The expectations inside the integrals in the expressions for
-$\mathbf{e}_I$ and $\mathbf{E}_I$
-are given by
-Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), respectively.
-%
-If we are able to perform the integrals in those expressions,
-we can \emph{dramatically} reduce the number of
-parameters in our ensemble problem.
-%
-As we show in Appendix~\ref{sec:inclination}, if we assume that stellar
-inclinations are distributed isotropically, these integrals
-do in fact have closed-form solutions.
-%
-We may then replace our expressions for the GP mean and covariance with
+That is, we may write the mean and covariance of our GP as
+%
 %
@@
     \label{eq:mu_marg}
-    \pmb{\mu}(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet)
-     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
+    \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)
+     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big]
     \nonumber                                                                                                                                                                                                      \\
      & = \mathbf{1} + \mathbf{e}_I
-    \nonumber                                                                                                                                                                                                                                       \\
-     & = \mu \, \mathbf{1}
     \\[0.5em]
     \label{eq:cov_marg}
-    \pmb{\Sigma}(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet)
-     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet) \pmb{\mu}^\top(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet)
-    \nonumber
-    \\
+    \pmb{\Sigma}(P, \mathbf{u}, \pmb{\theta}_\bullet)
+     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet) \pmb{\mu}^\top(P, \mathbf{u}, \pmb{\theta}_\bullet)
+    \nonumber                                                                                                                                                                                                      \\
      & =
@@
 %
-and use these to compute the marginal likelihood instead. Note that in the
-expression for the mean we made explicit use of the fact that the longitudinal
-isotropy of the process results in $\mathbf{e}_I = e_I \, \mathbf{1}$ being
-constant (see Appendix~\ref{sec:inc-mom1}), so the process mean is likewise
-constant, with $\mu = 1 + e_I$.
+where we define the inclination first moment integral
+%
+\begin{align}
+    \label{eq:eI}
+    \mathbf{e}_I
+     & \equiv
+    \mathrm{E} \Big[ \pmb{\mathcal{A}} \, \mathbf{y} \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big]
+    \nonumber \\[0.5em]
+     & =
+    \int
+    \pmb{\mathcal{A}}(I, P, \mathbf{u}) \,
+    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
+    p(I) \, \mathrm{d}I
+\end{align}
+%
+and the inclination second moment integral
+%
+\begin{align}
+    \label{eq:EI}
+    \mathbf{E}_I
+     & \equiv
+    \int
+    \pmb{\mathcal{A}}(I, P, \mathbf{u}) \,
+    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
+    \pmb{\mathcal{A}}^\top(I, P, \mathbf{u})
+    p(I) \, \mathrm{d}I
+    \quad.
+\end{align}
+%
+The expectations inside the integrals in the expressions for
+$\mathbf{e}_I$ and $\mathbf{E}_I$
+are given by
+Equations~(\ref{eq:exp_y}) and (\ref{eq:exp_yy}), respectively, and
+are computed in Appendix~\ref{sec:integrals}.
+%
+If we are able to perform the integrals in those expressions,
+we can \emph{dramatically} reduce the number of
+parameters in our ensemble problem.
+%
+As we show in Appendix~\ref{sec:inclination}, if we assume that stellar
+inclinations are distributed isotropically, these integrals
+do in fact have closed-form solutions.
+
+%
+Finally, for future reference, it is useful to note that the mean of the GP
+is constant:
+%
+\begin{align}
+    \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)
+     & =  (1 + e_I)\mathbf{1}
+    \nonumber                 \\
+     & \equiv
+    \mu \, \mathbf{1}
+    \quad,
+\end{align}
+%
+since by construction our GP is longitudinally isotropic
+(see Appendix~\ref{sec:inc-mom1}).
 
@@
 To understand why, consider what
-$\pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$ actually represents:
+$\pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)$ actually represents:
 it is the expected value
-of the flux
-for a star with certain properties $\pmb{\theta}_\star$ and $\pmb{\theta}_\bullet$,
+of the flux,
 \emph{in units of the flux one would measure if the star had no spots}.
@@
 
-\begin{figure}[ht!]
+\begin{figure}[t!]
     \begin{centering}
@@
 expression for the covariance matrix
-$\pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$ if we are to
+$\pmb{\Sigma}(P, \mathbf{u}, \pmb{\theta}_\bullet)$ if we are to
 use our GP to model normalized light curves. For definiteness, let us
@@
     for median-normalized light curves, since the distribution of the GP sample median
-    is usually close to the distribution of sample mean.}%
+    is usually close to the distribution of sample mean.}
 Since we are dividing the
@@
 This may be close to but is in general different from the \emph{process} mean,
-$\pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$, since the mean of
+$\pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)$, since the mean of
 a draw from the GP is itself normally distributed with a variance that scales
@@
     \tilde{\mathbf{f}\hspace{0.2em}}
-    \left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+    \left(P, \mathbf{u}, \pmb{\theta}_\bullet\right)
     \sim
@@
     \mathbf{1},
-    \tilde{\pmb{\Sigma}} \left(\pmb{\theta}_\star, \pmb{\theta}_\bullet\right)
+    \tilde{\pmb{\Sigma}} \left(P, \mathbf{u}, \pmb{\theta}_\bullet\right)
     \right)
@@
 %
-This is appropriate as long as the quantity $z \ll 1$, for which the true
+This is appropriate as long as $z \ll 1$, for which the true
 distribution of $\tilde{\mathbf{f}\hspace{0.2em}}$ is approximately Gaussian. In practice,
@@
     \tilde{\mathbf{f}\hspace{0.2em}}
-    \left(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet\right)
+    \left(P, \mathbf{u}, \pmb{\theta}_\bullet\right)
     \sim
@@
     \mathbf{1},
-    \tilde{\pmb{\Sigma}} \left(\bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet\right)
+    \tilde{\pmb{\Sigma}} \left(P, \mathbf{u}, \pmb{\theta}_\bullet\right)
     \right)
@@
 %
-where $\bar{\pmb{\theta}}_\star$ are hyperparameters describing the star, such
-as the rotation period and limb darkening coefficients
-(Equation~\ref{eq:thetastarbar}),
+where $P$ is the stellar rotation period, $\mathbf{u}$ are its limb darkening
+coefficients,
 $\pmb{\theta}_\bullet$ are the hyperparameters describing the star spot
@@
 true covariance of the process, which accounts for
-changes of scale and phase dependence introduced by the common process of
+changes in scale and phase introduced by the common process of
 normalizing light curves to a mean of unity. It depends on the
@@
 implementation of the algorithm in a user-friendly \Python package.
+
+\subsection{Examples}
+\label{sec:examples}
+
+Examples. Figure~\ref{fig:samples}.
+
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/samples.pdf}
+        \oscaption{samples}{%
+            Samples.
+            \label{fig:samples}
+        }
+    \end{centering}
+\end{figure}
+
+\section{Calibration}
+\label{sec:calibration}
+
+In the previous section we derived a closed form solution to the (best possible)
+Gaussian approximation to the marginal likelihood function for stellar
+light curves conditioned on a vector $\pmb{\theta}_\bullet$
+of spot hyperparameters. The true likelihood, however, is certainly \emph{not}
+Gaussian.
 
@@
 \subsection{Limb darkening}
+\label{sec:limbdark}
 
@@
 \end{figure}
+
+\subsection{Small spots}
+\label{sec:tinyspots}
+
+\subsection{The baseline degeneracy}
+\label{sec:breaking_baseline_degeneracy}
+
+\subsection{The number-contrast degeneracy}
+\label{sec:cN-degeneracy}
 
@@
 
+\section{Implementation}
+\label{sec:implementation}
+
 %
@@
 at increased computational cost and potential numerical stability issues;
-we discuss this point at length in \S\ref{sec:tiny-spots}.
+we discuss this point at length in \S\ref{sec:tinyspots}.
 
@@
     \int
-    \pmb{\mathcal{A}}(I, \bar{\pmb{\theta}}_\star) \,
+    \pmb{\mathcal{A}}(I, P, \mathbf{u}) \,
     \mathbf{e}_y \,
@@
     \int
-    \pmb{\mathcal{A}}(I, \bar{\pmb{\theta}}_\star) \,
+    \pmb{\mathcal{A}}(I, P, \mathbf{u}) \,
     \mathbf{E}_y \,
-    \pmb{\mathcal{A}}^\top(I, \bar{\pmb{\theta}}_\star) \,
+    \pmb{\mathcal{A}}^\top(I, P, \mathbf{u}) \,
     p(I) \,
@@
 \subsection{Speeding up the computation}
-\label{sec:speedup}
+\label{sec:inc-speedup}
 %

### 2020-12-16 tex/figures/samples.py

@@ -54,7 +54,7 @@ for n in range(len(kwargs)):
         for i, inc in enumerate(incs):
             map.inc = inc
             flux = map.flux(theta=360.0 * t)
-            flux -= np.median(flux)
+            flux -= np.mean(flux)
             flux *= 1e3
             ax[n, 1, k].plot(t, flux, color=color(i), lw=0.75)
 
@@ -94,6 +94,7 @@ for n in range(len(kwargs)):
     ax[n, 1, -1].axis("off")
 
     ax[n, 1, 0].get_shared_y_axes().join(*ax[n, 1, :])
+    ax[n, 1, 0].get_shared_x_axes().join(*ax[n, 1, :])
 
 # We're done
 fig.savefig(

### 2020-12-17 tex/ms.tex
more text

@@
     \right)
-    \quad,
-\end{align}
-%
-where $P$ is the stellar rotation period, $\mathbf{u}$ are its limb darkening
-coefficients,
-$\pmb{\theta}_\bullet$ are the hyperparameters describing the star spot
-distribution, including the number of spots, their contrast, parameters
-describing their latitudinal distribution, and parameters describing their
-radius distribution (Equation~\ref{eq:thetaspot}),
-and $\tilde{\pmb{\Sigma}}$ is the covariance of the normalized process
-(Equation~\ref{eq:SigmaTilde}). This covariance is a straightforward correction to the
+    \quad.
+\end{align}
+%
+The hyperparameters of the GP are
+the stellar rotation period $P$,
+the vector of limb darkening
+coefficients $\mathbf{u}$, and the vector of parameters describing the
+spot distribution
+%
+\begin{align}
+    \pmb{\theta}_\bullet
+     & =
+    \left(
+    N
+    \,\,\,
+    c
+    \,\,\,
+    a
+    \,\,\,
+    b
+    \,\,\,
+    r
+    \,\,\,
+    \Delta
+    \right)^\top
+    \quad,
+\end{align}
+%
+consisting of the number of spots $N$, their contrast $c$,
+the shape parameters of the latitude distribution $a$ and $b$, and
+the mean $r$ and half-width $\Delta$ of the spot radius
+distribution.
+The quantity $\tilde{\pmb{\Sigma}}$ is the covariance of the (normalized) process
+(Equation~\ref{eq:SigmaTilde}), which is a straightforward correction to the
 true covariance of the process, which accounts for
@@
 
-\subsection{Examples}
+\subsection{An example}
 \label{sec:examples}
-
-Examples. Figure~\ref{fig:samples}.
 
@@
         \oscaption{samples}{%
-            Samples.
+            Five random samples from the GP prior (columns) conditioned
+            on four different values of $\pmb{\theta}_\bullet$ (rows).
+            The samples are shown on the surface of
+            the star in a Mollweide projection (upper panels) alongside
+            their corresponding light curves viewed over four rotation
+            cycles at several different inclinations (lower panels).
+            From top to bottom, the GP corresponds to a star with
+            (a) small mid-latitude spots; (b) small circumpolar spots;
+            (c) large equatorial spots; and (d) small isotropic spots.
+            See text for details.
             \label{fig:samples}
@@
 
+A concrete example of the GP derived above is presented in Figure~\ref{fig:samples},
+where we show random samples from the process
+evaluated up to spherical harmonic degree $l_\mathrm{max} = 15$ and
+conditioned on different values
+of the hyperparameter vector $\pmb{\theta}_\bullet$.
+Each column corresponds to a different random draw from the GP, while each
+row corresponds to a different value of $\pmb{\theta}_\bullet$.
+The images are intensity maps of the stellar surface seen in an equal-area
+Mollweide projection, in units such that a spotless star would have intensity
+equal to 1 everywhere. Below them are the corresponding light curves (in units of
+parts per thousand deviation from the mean) over four
+rotation cycles, seen at inclinations varying from $15^\circ$ (yellow) to
+$90^\circ$ (dark blue), and assuming no limb darkening (i.e., $\mathbf{u} = \mathbf{0}$).
+From top to bottom, the hyperparameter vectors $\pmb{\theta}_\bullet$
+for each row are
+%
+\begin{subequations}
+    \begin{align}
+        \left( N \,\,\, c \,\,\, a \,\,\, b \,\,\, r \,\,\, \Delta \right)^\top
+         & =\left( 10 \,\,\,\,\,\,\,\,\, 0.10 \,\,\,\,\,\,\,\,\, 0.40 \,\,\,\,\,\,\,\,\, 0.27 \,\,\,\,\,\,\,\,\, 10.0 \,\,\,\,\,\,\,\,\, 0.00 \right)^\top \\
+         & =\left( 10 \,\,\,\,\,\,\,\,\, 0.10 \,\,\,\,\,\,\,\,\, 0.31 \,\,\,\,\,\,\,\,\, 0.36 \,\,\,\,\,\,\,\,\, 10.0 \,\,\,\,\,\,\,\,\, 0.00 \right)^\top \\
+         & =\left( 10 \,\,\,\,\,\,\,\,\, 0.05 \,\,\,\,\,\,\,\,\, 0.28 \,\,\,\,\,\,\,\,\, 0.00 \,\,\,\,\,\,\,\,\, 30.0 \,\,\,\,\,\,\,\,\, 0.00 \right)^\top \\
+         & =\left( 20 \,\,\,\,\,\,\,\,\, 0.10 \,\,\,\,\,\,\,\,\, 0.06 \,\,\,\,\,\,\,\,\, 0.00 \,\,\,\,\,\,\,\,\, 10.0 \,\,\,\,\,\,\,\,\, 0.00 \right)^\top
+        \quad.
+    \end{align}
+\end{subequations}
+%
+These correspond to
+(a) 10 spots of radius $10^\circ$ centered at
+$30^\circ \pm 5^\circ$ latitude with a contrast of $10\%$;
+(b) 10 spots of radius $10^\circ$ centered at
+$60^\circ \pm 5^\circ$ latitude with a contrast of $10\%$;
+(c) 10 spots of radius $30^\circ$ centered at
+$0^\circ \pm 5^\circ$ latitude with a contrast of $5\%$;
+and
+(d) 20 spots of radius $10^\circ$ centered at
+$30^\circ \pm 5^\circ$ latitude with a contrast of $10\%$.
+Note that the conversion between the shape parameters $a$ and $b$ and
+the more intuitive mean $\mu_\phi$ and standard deviation $\sigma_\phi$
+of the spot latitude distribution is given in \S\ref{sec:lat-transform}.
+
+The surface maps in the figure show dark, compact features of roughly the expected
+size and contrast and at the expected latitudes. However, there are important
+differences between these maps and what we would obtain by procedurally adding discrete
+circular spots to a gridded stellar surface:
+%
+\begin{enumerate}
+    \item \textbf{The spots are not circular.} This is most evident in row (c),
+          where some spots are distinctly asymmetric.
+    \item \textbf{There is significant variance in contrast from one spot to another}
+          even though our model
+          implicitly treats $c$ as constant. Within spots, the contrast is also
+          not constant, even though (again) our model implictly treats it as such.
+    \item \textbf{There is ringing in the background.} This is apparent to some extent in
+          row (b), where there are small fluctuations in the brightness at low
+          latitudes where no spots are present.
+    \item \textbf{There aren't exactly 10 (or 20) spots in those maps.} This is
+          most obvious in row (c), where only a few large distinct spots,
+          plus maybe (?) a few smaller ones, are visible.
+    \item \textbf{There are bright spots in addition to dark spots.} This may be
+          the most glaring issue. We explicitly model spots as being dark, and yet
+          there are (almost) just as many bright spots, particularly in row (d).
+          While bright spots (such as plages) certainly exist in reality, we did not
+          explicitly ask for them here!
+\end{enumerate}
+%
+While these may appear to be critical shortcomings of our model, it is
+important to keep in mind that a model consisting of discrete, circular, constant
+contrast spots is likely just as far (or perhaps even farther) from the truth. In fact,
+points (1) and (2) above suggest our model is more flexible than the discrete
+spot model and thus (potentially) better suited to modeling real stellar surfaces.
+Points (3), (4), and (5), on the other hand, are more concerning, since they are due,
+respectively, to truncation error in the spherical harmonic expansion,
+to an intrinsic limitation of our Gaussian approximation,
+and to the fact that Gaussian distributions are symmetric about the mean: a positive
+deviation is just as likely as a negative deviation of the same magnitude.%
+\footnote{
+    Even still, the model favors dark spots over bright spots because the GP mean
+    itself is lower than unity, in practice making
+    positive deviations \emph{from unity} less likely than negative
+    deviations. This is why there are usually more dark spots than bright spots in
+    the samples shown in the figure.
+}
+However, as we have argued before, the true power of the GP is in its
+applicability to inference problems. In other words, while our GP has some
+undesirable features when used as a prior sampling distribution, the real test
+of the GP is when it is faced with data in an inference setting.
+As long as the data is sufficiently informative, it does not matter that the
+prior has finite support for unphysical configurations, as those
+will be confidently rejected.
+
+In \S\ref{sec:calibration} below, we exhaustively test the performance of our
+GP as an inference tool when used to model synthetic light curves. We will
+show that, despite the issues raised above, the GP is in general unbiased and
+correctly estimates the posterior variance when used to infer the spot
+properties $\pmb{\theta}_\bullet$.
+
+But before we dive in, it is worth pausing for a moment to take another
+look at Figure~\ref{fig:samples}. While we focused on the shortcomings of
+the GP as a prior in the discussion above, it is important to appreciate
+that it even works in the first place! Despite some issues, the fact that
+the surface map samples \emph{actually look like maps of spotted stellar surfaces}
+is fairly amazing. A Gaussian process, after all, is a non-parameteric
+\emph{process} describing a smooth and continuous function only via its
+covariance structure. The GP knows nothing about the existence of discrete
+spots---only how any two points on the surface are correlated.
+Because spherical harmonics are smooth functions with support over the
+entire sphere, the GP also does not know about features restricted
+to certain latitudes; in fact, in most applications of GPs to mapping
+problems in astronomy \citep[such as in models of the cosmic microwave
+    background][]{?}, the process is assumed to be isotropic,
+with no preferred angular direction.
+The fact that we are able to construct a GP that captures discrete
+features restricted to certain regions on the surface (albeit with
+some issues) is remarkable.
+
 \section{Calibration}
@@
 
+\subsection{Why we need calibration tests}
+\label{sec:why-calibrate}
+
 In the previous section we derived a closed form solution to the (best possible)
-Gaussian approximation to the marginal likelihood function for stellar
-light curves conditioned on a vector $\pmb{\theta}_\bullet$
-of spot hyperparameters. The true likelihood, however, is certainly \emph{not}
-Gaussian.
+Gaussian approximation to the distribution of stellar surfaces (and their corresponding
+light curves) conditioned on a vector $\pmb{\theta}_\bullet$
+of spot hyperparameters. As we mentioned, the real test of this GP is
+in how well it performs as a likelihood function for stellar light curves.
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/nongaussianity.pdf}
+        \oscaption{nongaussianity}{%
+            Corner plot showing the joint distributions of select spherical
+            harmonic coefficients corresponding to the $l=15$ expansion of
+            $5\times 10^{4}$ stellar surface maps drawn from a certain
+            star spot distribution (black points and contours).
+            At $l=15$ there are 256 coefficients in total;
+            we chose five of the coefficients with the most odd-looking
+            distributions to illustrate the non-Gaussianity of the process.
+            In addition to non-linear correlations, skewness, and the
+            existence of points of very high curvature, some of the distributions
+            are also multi-modal.
+            The orange contours show slices of the Gaussian approximation to
+            the joint distribution; this is the approximation adopted in
+            this paper.
+            \label{fig:nongaussianity}
+        }
+    \end{centering}
+\end{figure}
+
+It is not immediately obvious that the GP approach should work, because the
+true marginal likelihood function
+%
+$p ( \tilde{\mathbf{f}\hspace{0.2em}} \, \big| \, P, \mathbf{u}, \pmb{\theta}_\bullet )$
+%
+is certainly \emph{not} Gaussian. To see why, let us generate $10^{4}$ stellar
+surfaces sampled from the true distribution we are trying to model: that is,
+a surface with $N$ discrete circular spots of fixed, uniform contrast $c$
+and radius $r \pm \Delta$ at latitudes $\mu_\phi \pm \sigma_\phi$.
+Let us then expand each surface in spherical harmonics and visualize the
+distribution of coefficients $y^l_m$. Figure~\ref{fig:nongaussianity}
+shows the joint distribution for five of the coefficients with the most
+non-Gaussian marginal distributions (selected by eye). Different slices
+through this distribution
+(in black) are skewed, strongly peaked, non-linearly correlated, and even bimodal.
+Our approach in this paper is to model this distribution as a Gaussian
+(orange contours). While this may be a good approximation in certain regions
+of parameter space, it is certainly a poor approximation in others.
+In all fairness, the vast majority of the spherical harmonic coefficients
+have marginal distributions that \emph{look} to be very Gaussian; however,
+those that do not could in principle lead to errors and biases when using
+the Gaussian approximation to the likelihood function.
 
@@
 \label{sec:cN-degeneracy}
-
-\subsection{Caveats}
-\label{sec:caveats}
-The true process is not Gaussian. See Figure~\ref{fig:nongaussianity}.
-\begin{figure}[ht!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/nongaussianity.pdf}
-        \oscaption{nongaussianity}{%
-            Corner plot showing the joint distributions of select spherical
-            harmonic coefficients corresponding to the $l=15$ expansion of
-            $5\times 10^{4}$ stellar surface maps drawn from a certain
-            star spot distribution (black points and contours).
-            At $l=15$ there are 256 coefficients in total;
-            we chose five of the coefficients with the most odd-looking
-            distributions to illustrate the non-Gaussianity of the process.
-            In addition to non-linear correlations, skewness, and the
-            existence of points of very high curvature, some of the distributions
-            are also multi-modal.
-            The orange contours show slices of the Gaussian approximation to
-            the joint distribution; this is the approximation adopted in
-            this paper.
-            \label{fig:nongaussianity}
-        }
-    \end{centering}
-\end{figure}
 

### 2020-12-18 tex/ms.tex
more text

@@
           where some spots are distinctly asymmetric.
-    \item \textbf{There is significant variance in contrast from one spot to another}
+    \item \textbf{There is significant variance in contrast from one spot to another},
           even though our model
@@
 the Gaussian approximation to the likelihood function.
+
+\xxx{An example using the 1d Beta distribution.}
 

### 2020-12-18 tex/ms.tex
more text, last test

@@
 
-\xxx{An example using the 1d Beta distribution.}
+\subsection{Setup}
+\label{sec:calibration-setup}
+
+\begin{table}[t!]
+    \begin{center}
+        \begin{longtable}{W{c}{1cm} W{l}{6cm} W{l}{2cm}}
+            \label{tab:synthetic}
+            \\
+            %
+            \toprule
+            \multicolumn{1}{c}{\textbf{Symbol}}
+             &
+            \multicolumn{1}{c}{\textbf{Description}}
+             &
+            \multicolumn{1}{c}{\textbf{Value}}
+            \\
+            \midrule
+            \endhead
+            \bottomrule                                 \\
+            %
+            \caption{%
+                Default parameters used to generate synthetic light curves in
+                the calibration tests.
+            }
+            \endfoot
+            %
+            %
+            $N$
+             & number of spots
+             & $\sim\mathcal{N}(20, 0^2)$
+            %
+            \\
+            %
+            $c$
+             & spot contrast
+             & $\sim\mathcal{N}(0.05, 0^2)$
+            %
+            \\
+            %
+            $\phi$
+             & spot latitude
+             & $\sim\mathcal{N}(30^\circ, {5^\circ}^2)$
+            %
+            \\
+            $\lambda$
+             & spot longitude
+             & $\sim\mathcal{U}(0^\circ, 360^\circ)$
+            %
+            \\
+            %
+            $r$
+             & spot radius
+             & $\sim\mathcal{N}(15^\circ, {0^\circ}^2)$
+            %
+            \\
+            %
+            $I$
+             & stellar inclination
+             & $\sim\sin$
+            %
+            \\
+            %
+            $P$
+             & rotational period
+             & $1$ day
+            %
+            \\
+            %
+            $\mathbf{u}$
+             & limb darkening coefficients
+             & $\left( 0 \,\,\,\, 0 \right)^\top$
+            %
+            \\
+            %
+            $\sigma_f$
+             & photometric uncertainty
+             & $10^{-3}$
+            %
+            \\
+            %
+            $K$
+             & number of cadences per light curve
+             & $10^3$
+            %
+            \\
+            %
+            $\Delta t$
+             & time baseline
+             & $4$ days
+            %
+            \\
+            %
+            $M$
+             & number of light curves in ensemble
+             & $50$
+            %
+            \\
+            %
+        \end{longtable}
+    \end{center}
+\end{table}
+
+We seek to demonstrate that our model is correctly calibrated by testing it
+on synthetic data, which we generate as follows. For each of $M$ synthetic
+light curves in a given ensemble,
+we create a rectangular
+$(150 \times 300)$ latitude-longitude grid of surface intensity values, all
+intialized
+to zero. We then add $N$ spots to this grid, each of fractional contrast
+$c$ and radius $r$ centered at latitude $\phi$ and longitude $\lambda$,
+by decreasing the intensity at all points within an angular distance $r$
+(measured along the surface of the sphere) by an amount $c$.
+In order to compute the corresponding light curve, we expand the surface
+in spherical harmonics, although at much higher degree
+($l_\mathrm{max}^{(0)} = 30$)
+than the degree we will use in the inference step ($l_\mathrm{max} = 15$)
+to minimize potential ringing effects or other artefacts in the synthetic
+data. For reference, the chosen degree $l_\mathrm{max} = 30$ is large
+enough to resolve features on the order of
+$\nicefrac{180^\circ}{30} = 6^\circ$ across, but small enough that
+the algorithm for computing the light curve is numerically stable. We
+compute the light curve at $K$ points equally spaced over a baseline $\Delta t$
+using the \starry algorithm
+(Appendix \ref{sec:starry}), assuming an inclination $I$, a rotational period $P$,
+and limb darkening coefficients $\mathbf{u}$. Finally, we divide the
+light curve by the mean and add Gaussian noise with standard deviation $\sigma_f$
+to emulate photon noise.%
+\footnote{
+    In theory, we should do this in the reverse order: i.e., we should add photon
+    noise and \emph{then} normalize the light curve to the mean, since that is the order in which
+    those steps occur in reality. However, if we did that, we would have to normalize $\sigma_f$
+    in our inference step, such that the variance of each of the normalized light curves
+    in the ensemble would be different, requiring us to invert a different matrix for
+    each light curve when computing the log likelihood (see Equation~\ref{eq:log-like}).
+    This would significantly increase the computational cost of our tests. Fortunately, in
+    practice, the difference between these two approaches has negligible effect on
+    our inference results, so we opt for the faster of the two methods. Note, of course,
+    that when applying our GP to real data, we won't have this choice!
+}
+The default values of each of the parameters mentioned above are given in
+Table~\ref{tab:synthetic}. Some, like the number of spots, their contrast,
+etc., are drawn from fiducial distributions, while others, like the
+photometric uncertainty, the rotational period, and the limb darkening
+coefficients,
+are fixed across all $M$ light curves in an ensemble. These fixed values
+are not realistic, but they greatly speed up the inference step, since they
+allow us to invert a single covariance matrix for all light curves when computing
+the log likelihood.
 
@@
 %
-\begin{proof}{}
+\begin{proof}{test_ld}
     \label{eq:ld:L}

### 2020-12-18 starry_process/ops/__init__.py

@@ -1,6 +1,6 @@
 from .latitude import LatitudeIntegralOp
 from .wigner import RxOp, RyOp, tensordotRzOp, special_tensordotRzOp
-from .flux import rTA1Op, rTA1LOp
+from .flux import rTA1Op, rTA1LOp, LOp
 from .exceptions import CheckBoundsOp, CheckVectorSizeOp
 from .eigh import EighOp
 from .norm import AlphaBetaOp

### 2020-12-18 starry_process/ops/flux/L.py

@@ -1,0 +1,26 @@
+# -*- coding: utf-8 -*-
+from ..base_op import BaseOp
+from theano import gof
+import theano
+import theano.tensor as tt
+
+__all__ = ["LOp"]
+
+
+class LOp(BaseOp):
+    func_file = "./L.cc"
+    func_name = "APPLY_SPECIFIC(L)"
+
+    def make_node(self, u):
+        in_args = [
+            tt.as_tensor_variable(arg).astype(tt.config.floatX) for arg in [u]
+        ]
+        out_args = [
+            tt.TensorType(
+                dtype=tt.config.floatX, broadcastable=[False, False]
+            )()
+        ]
+        return gof.Apply(self, in_args, out_args)
+
+    def infer_shape(self, node, shapes):
+        return ([self.NLU, self.N],)

### 2020-12-18 starry_process/ops/include/flux.h

@@ -278,6 +278,21 @@ inline void computeA1(int deg, Eigen::SparseMatrix<Scalar> &A1) {
   A1.setFromTriplets(coeffs.begin(), coeffs.end());
 }
 
+template <typename Scalar>
+inline void computeA1Inv(int deg, const Eigen::SparseMatrix<Scalar> &A1,
+                         Eigen::SparseMatrix<Scalar> &A1Inv) {
+  int N = (deg + 1) * (deg + 1);
+  Eigen::SparseLU<Eigen::SparseMatrix<Scalar>> solver;
+  solver.compute(A1);
+
+  if (solver.info() != Eigen::Success)
+    throw std::runtime_error(
+        "Error computing the change of basis matrix `A1Inv`.");
+  Eigen::SparseMatrix<Scalar> I =
+      RowMatrix<Scalar, Dynamic, Dynamic>::Identity(N, N).sparseView();
+  A1Inv = solver.solve(I);
+}
+
 /**
 Compute the starry flux operator `rTA1`.
 
@@ -300,6 +315,7 @@ template <typename Scalar> class LimbDark {
 public:
   RowVector<Scalar, Dynamic> rT;
   Eigen::SparseMatrix<Scalar> A1;
+  Eigen::SparseMatrix<Scalar> A1Inv;
   Eigen::SparseMatrix<Scalar> A1NbyN;
   Eigen::SparseMatrix<Scalar> U1;
   Vector<Scalar, Dynamic> p;
@@ -392,11 +408,6 @@ public:
 
   /**
   Compute the limb darkening operator in the polynomial basis.
-  This is equal to
-
-    Lp = A1^-1 . L . A1
-
-  where `L` is the limb darkening operator in the spherical harmonic basis.
 
   */
   inline void computeLp() {
@@ -473,6 +484,7 @@ public:
       // Pre-compute a bunch of stuff
       computerT(SP__LUMAX, rT);
       computeA1(SP__LUMAX, A1);
+      computeA1Inv(SP__LUMAX, A1, A1Inv);
       computeU1();
       A1NbyN = A1.topLeftCorner(SP__N, SP__N);
       computeDDp();
@@ -502,7 +514,7 @@ public:
     // Compute the limb darkening operator `L` in the polynomial basis
     computeLp();
 
-    // Recall that `Lp = A1^-1 . L . A1`
+    // Change of basis
     rTA1L = (rT * Lp) * A1NbyN;
   }
 
[…]

### 2020-12-19 tex/ms.tex
more text

@@
 
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_default_data.pdf}
+        \oscaption{calibration}{%
+            A synthetic dataset generated by adding discrete spots to a stellar
+            surface with parameters given by the default values listed in
+            Table~\ref{tab:synthetic}. The surface maps for 50 synthetic stars are
+            shown in a Mollweide equal-area projection above their corresponding
+            light curves. All maps and light curves are plotted on the same scale.
+            The $\times$ on each map indicates the sub-observer latitude assumed
+            when generating the light curve. The blue curves correspond to the
+            exact light curve, while the black dots are the observed light curve.
+            \label{fig:calibration_default_data}
+        }
+    \end{centering}
+\end{figure}
+
+Figure~\ref{fig:calibration_default_data} shows a synthetic dataset generated
+from the default values listed in Table~\ref{tab:synthetic}. While the 50
+light curves correspond to surfaces with the same statistical spot properties,
+they all look qualitatively different: the mapping from star spot properties
+to flux is (very) nontrivial. In the inference step below, we assume we observe
+only these 50 light curves, with no knowledge of the inclination of any individual
+star, and attempt to infer the spot properties.
+
+\subsection{Inference}
+\label{sec:calibration-inference}
+\xxx{TODO.}
+
 \section{Discussion}

### 2020-12-19 tex/figures/calibration.py

@@ -2,12 +2,22 @@ from starry_process import calibrate
 import numpy as np
 import os
 
+# Utility funcs to move figures to this directory
+abspath = lambda *args: os.path.join(
+    os.path.dirname(os.path.abspath(__file__)), *args
+)
+move = lambda name, src, dest: os.rename(
+    abspath("data", name, src), abspath(dest)
+)
+
+
+# Default run
+calibrate.run(path=abspath("data/default"))
+move("default", "data.pdf", "calibration_default_data.pdf")
+
 
 # TODO: Not yet ready for CI runs
-if not bool(os.getenv("CI", False)):
-
-    # Default run
-    calibrate.run(path="data/default")
+if not int(os.getenv("CI", 0)):
 
     # Equatorial spots
     calibrate.run(

### 2020-12-21 tex/ms.tex
better notation for random vars

@@
 \emph{functions} spanning a continuous domain (in our case, the time domain).
-Like a multivariate
-Gaussian, a GP is fully specified by a $(K \times 1)$ vector $\pmb{\mu}$ characterizing
+Similar to a multivariate Gaussian, which is described by
+a $(K \times 1)$ vector $\pmb{\mu}$ characterizing
 the mean of the process and a $(K \times K)$
-covariance matrix $\pmb{\Sigma}$ characterizing how the data points
-are correlated with each other. To say that a measurement $\mathbf{f}$ is
-``distributed as a GP'' means that we may write
-%
-\begin{align}
-    \mathbf{f} \sim \mathcal{N}\left( \pmb{\mu}, \pmb{\Sigma} \right)
-    \quad,
-\end{align}
-%
-i.e., the measurement is modeled as a random draw from a multivariate Gaussian.
-GPs are therefore extremely easy to sample from.%
+matrix $\pmb{\Sigma}$ describing its covariance,
+a GP is fully specified by a mean
+function $m(t)$ and a kernel function $k(t, t')$.
+To say that a random vector-valued variable $\mathbbb{f}$
+defined on a $(K \times 1)$ time array $\mathbf{t}$
+is ``distributed as a GP'' means that we may write
+%
+\begin{align}
+    \mathbbb{f} \sim \mathcal{N}\left( \pmb{\mu}, \pmb{\Sigma} \right)
+    \quad,
+\end{align}
+%
+where the elements of the mean and covariance are given by
+$\mu_i = m(t_i)$ and $\Sigma_{i,j} = k(t_i, t_j)$, respectively.%
+\footnote{%
+    In this paper, we will use blackboard font (i.e., $\mathbbb{f}$) to
+    denote random variables and serif font (i.e., $\mathbf{f}$) to denote
+    particular realizations of those variables.
+    See Appendix~\ref{sec:notation} for a detailed explanation of our notation.
+}
+Because of this relationship to multivariate Gaussians,
+GPs are extremely easy to sample from.%
 \footnote{Given a 1-d array \texttt{mean} and a 2-d array \texttt{cov} in \Python,
@@
 a class of ``non-parametric'' models, given that nowhere in the specification
-of the GP is there an explicit functional form for $\mathbf{f}$. Rather, a GP
+of the GP is there an explicit functional form for $\mathbbb{f}$. Rather, a GP
 is a stochastic
@@
 that are fully
-encoded in the covariance $\pmb{\Sigma}$. Importantly, the covariance
-$\pmb{\Sigma}$ of a GP is assembled via a continuous \emph{kernel function} $k(t_i, t_j)$
-that specifies the covariance between points at times $t = t_i$ and $t = t_j$:
-%
-\begin{align}
-    \Sigma_{i,j} = k(t_i, t_j)
-    \quad.
-\end{align}
+encoded in the covariance $\pmb{\Sigma}$.
 %
@@
 with free parameters controling the spot size, contrast, position,
-and temporal evolution of each of the $N$ spots, where $N$ is also
+and temporal evolution of each of the $n$ spots, where $n$ is also
 a free parameter. In addition to this being a numerically challenging
-problem for all but the smallest values of $N$, the problem is
+problem for all but the smallest values of $n$, the problem is
 intrinsically ill-posed, meaning there are strong degeneracies at play
@@
     smaller than the effective resolution of the GP.}, as they will allow us to compute
-moments of the intensity distribution analytically. And of more immediate concern,
+moments of the intensity distribution analytically. Of more immediate concern,
 \citet{Luger2019} showed that there is a linear relationship between the
@@
 Let
-$\mathbf{f} = \left( f_0 \, f_1 \, \cdots \,  f_{K-1} \right)^\top$
-denote a vector of $K$ flux measurements at times
-$\mathbf{t} = \left( t_0 \,  t_1 \,  \cdots \, t_{K-1} \right)^\top$,
+$\mathbbb{f} = \left( \mathbb{f}_0 \,\, \mathbb{f}_1 \,\, \cdots \,\,  \mathbb{f}_{K-1} \right)^\top$
+denote a random vector of $K$ flux measurements at times
+$\mathbf{t} = \left( t_0 \,\,  t_1 \,\,  \cdots \,\, t_{K-1} \right)^\top$,
 defined in units such that a star with no spots on it will have
-$\mathbf{f} = \mathbf{1}$.
+unit flux.
 Conditioned on the stellar inclination $I$, the rotational period $P$,
@@
 covariance $\pmb{\Sigma}(I, P, \mathbf{u}, \pmb{\theta}_\bullet)$
-of $\mathbf{f}$. Together, these specify a multidimensional Gaussian
+of $\mathbbb{f}$. Together, these specify a multidimensional Gaussian
 distribution, which we assume fully describes%
@@
 \begin{align}
-    \mathbf{f}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right) \sim
+    \mathbbb{f}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right) \sim
     \mathcal{N}\Big(
@@
 As with any random variable, the mean and covariance may be computed from
-the expectation values of $\mathbf{f}$ and
-$\mathbf{f}\,\mathbf{f}^\top$, respectively:
+the expectation values of $\mathbbb{f}$ and
+$\mathbbb{f}\,\mathbbb{f}^\top$, respectively:
 %
@@
     \pmb{\mu}(I, P, \mathbf{u}, \pmb{\theta}_\bullet)
-     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, I, P, \mathbf{u}, \pmb{\theta}_\bullet \Big]
+     & = \mathrm{E} \Big[ \mathbbb{f} \, \Big| \, I, P, \mathbf{u}, \pmb{\theta}_\bullet \Big]
     \\
@@
     \pmb{\Sigma}(I, P, \mathbf{u}, \pmb{\theta}_\bullet)
-     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, I, P, \mathbf{u}, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(I, P, \mathbf{u}, \pmb{\theta}_\bullet) \pmb{\mu}^\top(I, P, \mathbf{u}, \pmb{\theta}_\bullet)
+     & = \mathrm{E} \Big[ \mathbbb{f} \, \mathbbb{f}^\top \, \Big| \, I, P, \mathbf{u}, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(I, P, \mathbf{u}, \pmb{\theta}_\bullet) \pmb{\mu}^\top(I, P, \mathbf{u}, \pmb{\theta}_\bullet)
     \quad.
@@
     \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
-     & = \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
+     & = \mathrm{E} \Big[ \mathbbb{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
     \\
@@
     \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet)
-     & = \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] - \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet) \pmb{\mu}_{\mathbf{y}}^\top(\pmb{\theta}_\bullet)
+     & = \mathrm{E} \Big[ \mathbbb{y} \, \mathbbb{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] - \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet) \pmb{\mu}_{\mathbf{y}}^\top(\pmb{\theta}_\bullet)
 \end{align}
@@
     \label{eq:exp_y}
-    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
-     & =
-    \int \mathbf{y}(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
+    \mathrm{E} \Big[ \mathbbb{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
+     & =
+    \int \mathbbb{y}(\mathbbb{x} ) \, p(\mathbbb{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbbb{x}
     \\
     \label{eq:exp_yy}
-    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
-     & =
-    \int \mathbf{y}(\mathbf{x} ) \mathbf{y}^\top(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
-    \quad,
-\end{align}
-%
-where $\mathbf{x}$ is a random vector-valued variable corresponding to a particular
+    \mathrm{E} \Big[ \mathbbb{y} \, \mathbbb{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
+     & =
+    \int \mathbbb{y}(\mathbbb{x} ) \mathbbb{y}^\top(\mathbbb{x} ) \, p(\mathbbb{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbbb{x}
+    \quad,
+\end{align}
+%
+where $\mathbbb{x}$ is a random vector-valued variable corresponding to a particular
 distribution of features on the surface
-and $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$ is its probability density
+and $p(\mathbbb{x} \, \big| \, \pmb{\theta}_\bullet)$ is its probability density
 function (PDF).
@@
 In the Appendix we show that for suitable choices of $\pmb{\theta}_\bullet$,
-$\mathbf{y}(\mathbf{x})$,
-and $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$, the integrals in the expressions
+$\mathbbb{y}(\mathbbb{x})$,
+and $p(\mathbbb{x} \, \big| \, \pmb{\theta}_\bullet)$, the integrals in the expressions
 above have closed form solutions that may be evaluated quickly.
 %
-In particular, we let the spot hyperparameters of our GP be
+While we present a few different ways of specifying $\pmb{\theta}_\bullet$,
+our default representation of the GP hyperparameters is
 %
@@
     \left(
-    N
+    n
     \,\,\,
@@
     \,\,\,
-    a
+    \mu_\phi
     \,\,\,
-    b
+    \sigma_\phi
     \,\,\,
     r
-    \,\,\,
-    \Delta
     \right)^\top
@@
 %
-where $N$ is the number of star spots, $c$ is their contrast,
-$a$ and $b$ describe the latitudinal
-distribution of the spots, and $r$ and $\Delta$ are the mean and
-half-width, respectively, of the distribution of spot radii.
-%
-The PDF for the radius is a uniform distribution
-between $r - \Delta$ and $r + \Delta$ (Appendix \ref{sec:size}),
+where $n$ is the number of star spots, $c$ is their contrast,
+$\mu$ and $\sigma$ are parameters describing the
+distribution of the spots in latitude, and $r$ is the radius
+of the spots.
+%
+The PDFs for the spot radius, the spot contrast, and the number
+of spots are delta functions centered at $r$, $c$, and $n$, respectively
+(Appendices \ref{sec:size} and \ref{sec:contrast}),
 while
@@
 $\cos\phi$ with (normalized) shape parameters $a$ and $b$,
-which have a one-to-one correspondence to the mean $\mu_\phi$ and
+which have a one-to-one correspondence to the mode $\mu_\phi$ and
 standard deviation $\sigma_\phi$ of the distribution in $\phi$
-(Appendix \ref{sec:lat-transform}).
-The spot longitude is assumed to be uniformly distributed
-(Appendix \ref{sec:lon}). Finally,
-the PDFs for the spot contrast and the number of spots are delta functions
-centered at $c$ and $N$, respectively (Appendix \ref{sec:contrast}).
+(Appendix \ref{sec:lat}). Finally,
+the spot longitude is assumed to be uniformly distributed
+(Appendix \ref{sec:lon}).
 
@@
 given a set of $M$ light curves
-$\left( \mathbf{f}_0 \, \mathbf{f}_1 \, \cdots \,  \mathbf{f}_{M-1} \right)^\top$,
+$\left( \mathbf{f}_0 \,\, \mathbf{f}_1 \,\, \cdots \,\,  \mathbf{f}_{M-1} \right)^\top$,
 we wish to infer the statistical properties of the star spots, encoded in
@@
 between some of the parameters (such as the contrast $c$ and the
-number of spots $N$), the posterior is often multimodal, especially in the
+number of spots $n$), the posterior is often multimodal, especially in the
 stellar inclinations. While modern sampling schemes such as Hamiltonian
@@
     \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)
-     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big]
+     & = \mathrm{E} \Big[ \mathbbb{f} \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big]
     \nonumber                                                                                                                                                                                                        \\
@@
     \pmb{\Sigma}(P, \mathbf{u}, \pmb{\theta}_\bullet)
-     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet) \pmb{\mu}^\top(P, \mathbf{u}, \pmb{\theta}_\bullet)
+     & = \mathrm{E} \Big[ \mathbbb{f} \, \mathbbb{f}^\top \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet) \pmb{\mu}^\top(P, \mathbf{u}, \pmb{\theta}_\bullet)
     \nonumber                                                                                                                                                                                                        \\
@@
      & \equiv
-    \mathrm{E} \Big[ \pmb{\mathcal{A}} \, \mathbf{y} \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big]
-    \nonumber \\[0.5em]
-     & =
     \int
-    \pmb{\mathcal{A}}(I, P, \mathbf{u}) \,
-    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
-    p(I) \, \mathrm{d}I
+    \pmb{\mathcal{A}}(\mathbb{I}, P, \mathbf{u}) \,
+    \mathbf{E} \Big[ \mathbbb{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
+    p(\mathbb{I}) \, \mathrm{d}\mathbb{I}
 \end{align}
@@
     \int
-    \pmb{\mathcal{A}}(I, P, \mathbf{u}) \,
-    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
-    \pmb{\mathcal{A}}^\top(I, P, \mathbf{u})
-    p(I) \, \mathrm{d}I
+    \pmb{\mathcal{A}}(\mathbb{I}, P, \mathbf{u}) \,
+    \mathbf{E} \Big[ \mathbbb{y} \, \mathbbb{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
+    \pmb{\mathcal{A}}^\top(\mathbb{I}, P, \mathbf{u})
+    p(\mathbb{I}) \, \mathrm{d}\mathbb{I}
     \quad.
@@
 rotation. Thus far we have cast the light curve problem in a typical GP inference
-setting: we model our observations $\mathbf{f}$ as
+setting: we model our observations $\mathbbb{f}$ as
 a multidimensional random variable distributed as a multivariate Gaussian with mean vector
@@
     \label{eq:ftilde}
-    \tilde{\mathbf{f}\hspace{0.2em}} & = \frac{\mathbf{f}}{\left<f\right>}
-    \quad,
-\end{align}
-%
-where $\tilde{\mathbf{f}\hspace{0.2em}}$ is the normalized, unit-mean light curve,
-$\mathbf{f}$ is the measured light curve (in detector counts), and
-$\left<f\right>$ is the \emph{sample} mean: i.e., the average value of
+    \tilde{\mathbbb{f}\hspace{0.2em}} & = \frac{\mathbbb{f}}{\left<\mathbb{f}\right>}
+    \quad,
+\end{align}
+%
+where $\tilde{\mathbbb{f}\hspace{0.2em}}$ is the normalized, unit-mean light curve,
+$\mathbbb{f}$ is the measured light curve (in detector counts), and
+$\left<\mathbb{f}\right>$ is the \emph{sample} mean: i.e., the average value of
 a given star's light curve (which we model as a sample from our GP).
@@
 because the normalized process is \emph{not} strictly Gaussian: the distribution
-has heavy tails due to the fact that $\tilde{\mathbf{f}\hspace{0.2em}}$ diverges as
+has heavy tails due to the fact that $\tilde{\mathbbb{f}\hspace{0.2em}}$ diverges as
 the sample mean approaches zero. In fact, because of these tails, the covariance
@@
      & \equiv
-    \sum\limits_{n=0}^N
-    \frac{(2n + 1)!}{2^n \, n!}
-    z^n
+    \sum\limits_{i=0}^{i_\mathrm{max}}
+    \frac{(2i + 1)!}{2^i \, i!}
+    z^i
     \\[1em]
@@
      & \equiv
-    \sum\limits_{n=0}^N
-    \frac{2n(2n + 1)!}{2^n \, n!}
-    z^n
-    \quad,
-\end{align}
-%
-where $N$ is the largest value for which the series coefficient at $N$ is
-smaller than the coefficient at $N - 1$. In the expressions above, it is
+    \sum\limits_{i=0}^{i_\mathrm{max}}
+    \frac{2i(2i + 1)!}{2^i \, i!}
+    z^i
+    \quad,
+\end{align}
+%
+where $i_\mathrm{max}$ is the largest value for which the series coefficient at $i_\mathrm{max}$ is
+smaller than the coefficient at $i_\mathrm{max} - 1$. In the expressions above, it is
 assumed that the mean $\pmb{\mu}$ is constant, i.e., $\pmb{\mu} = \mu\, \mathbf{1}$.
@@
 \begin{align}
-    \tilde{\mathbf{f}\hspace{0.2em}}
+    \tilde{\mathbbb{f}\hspace{0.2em}}
     \left(P, \mathbf{u}, \pmb{\theta}_\bullet\right)
@@
 This is appropriate as long as $z \ll 1$, for which the true
-distribution of $\tilde{\mathbf{f}\hspace{0.2em}}$ is approximately Gaussian. In practice,
+distribution of $\tilde{\mathbbb{f}\hspace{0.2em}}$ is approximately Gaussian. In practice,
 we recommend employing this trick only for $z \lesssim 0.02$, for which the
@@
 
-We model the mean-normalized flux $\tilde{\mathbf{f}\hspace{0.2em}}$
+We model the mean-normalized flux $\tilde{\mathbbb{f}\hspace{0.2em}}$
 (Equation~\ref{eq:ftilde}) as a Gaussian process:
@@
 %
-consisting of the number of spots $N$, their contrast $c$,
-the shape parameters of the latitude distribution $a$ and $b$, and
-the mean $r$ and half-width $\Delta$ of the spot radius
-distribution.
+consisting of the number of spots $n$, their contrast $c$,
+the mode $\mu_\phi$ and standard deviation $\sigma_\phi$ of the latitude
+distribution, and the radius of the spots $r$.
 The quantity $\tilde{\pmb{\Sigma}}$ is the covariance of the (normalized) process
@@
 second moments of the distribution of spherical harmonic coefficient vectors,
-$\mathrm{E}\left[ \mathbf{y} \big| \pmb{\theta}_\bullet \right]$
+$\mathbf{E}\left[ \mathbbb{y} \, \big| \, \pmb{\theta}_\bullet \right]$
 and
-$\mathrm{E}\left[ \mathbf{y} \, \mathbf{y}^\top \big| \pmb{\theta}_\bullet \right]$,
+$\mathbf{E}\left[ \mathbbb{y} \, \mathbbb{y}^\top \, \big| \, \pmb{\theta}_\bullet \right]$,
 given by Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), respectively.
@@
     \begin{align}
-        \left( N \,\,\, c \,\,\, a \,\,\, b \,\,\, r \,\,\, \Delta \right)^\top
-         & =\left( 10 \,\,\,\,\,\,\,\,\, 0.10 \,\,\,\,\,\,\,\,\, 0.40 \,\,\,\,\,\,\,\,\, 0.27 \,\,\,\,\,\,\,\,\, 10.0 \,\,\,\,\,\,\,\,\, 0.00 \right)^\top \\
-         & =\left( 10 \,\,\,\,\,\,\,\,\, 0.10 \,\,\,\,\,\,\,\,\, 0.31 \,\,\,\,\,\,\,\,\, 0.36 \,\,\,\,\,\,\,\,\, 10.0 \,\,\,\,\,\,\,\,\, 0.00 \right)^\top \\
-         & =\left( 10 \,\,\,\,\,\,\,\,\, 0.05 \,\,\,\,\,\,\,\,\, 0.28 \,\,\,\,\,\,\,\,\, 0.00 \,\,\,\,\,\,\,\,\, 30.0 \,\,\,\,\,\,\,\,\, 0.00 \right)^\top \\
-         & =\left( 20 \,\,\,\,\,\,\,\,\, 0.10 \,\,\,\,\,\,\,\,\, 0.06 \,\,\,\,\,\,\,\,\, 0.00 \,\,\,\,\,\,\,\,\, 10.0 \,\,\,\,\,\,\,\,\, 0.00 \right)^\top
+        \left( n \,\,\, c \,\,\, \mu_\phi \,\,\, \sigma_\phi \,\,\, r\right)^\top
+         & =\left( 10.0 \,\,\,\,\,\,\,\,\, 0.10 \,\,\,\,\,\,\,\,\, 30.0 \,\,\,\,\,\,\,\,\, 5.00 \,\,\,\,\,\,\,\,\, 10.0 \right)^\top \\
+         & =\left( 10.0 \,\,\,\,\,\,\,\,\, 0.10 \,\,\,\,\,\,\,\,\, 60.0 \,\,\,\,\,\,\,\,\, 5.00 \,\,\,\,\,\,\,\,\, 10.0 \right)^\top \\
+         & =\left( 10.0 \,\,\,\,\,\,\,\,\, 0.05 \,\,\,\,\,\,\,\,\, 0.00 \,\,\,\,\,\,\,\,\, 5.00 \,\,\,\,\,\,\,\,\, 30.0 \right)^\top \\
+         & =\left( 20.0 \,\,\,\,\,\,\,\,\, 0.10 \,\,\,\,\,\,\,\,\, 0.00 \,\,\,\,\,\,\,\,\, 40.0 \,\,\,\,\,\,\,\,\, 10.0 \right)^\top
         \quad.
@@
 (d) 20 spots of radius $10^\circ$ centered at
-$30^\circ \pm 5^\circ$ latitude with a contrast of $10\%$.
-Note that the conversion between the shape parameters $a$ and $b$ and
-the more intuitive mean $\mu_\phi$ and standard deviation $\sigma_\phi$
-of the spot latitude distribution is given in \S\ref{sec:lat-transform}.
+$0^\circ \pm 40^\circ$ latitude (a good approximation to a perfectly
+isotropic distribution; see Appendix~\ref{sec:lat}) with a contrast of $10\%$.
 
@@
 %
-$p ( \tilde{\mathbf{f}\hspace{0.2em}} \, \big| \, P, \mathbf{u}, \pmb{\theta}_\bullet )$
+$p ( \tilde{\mathbbb{f}\hspace{0.2em}} \, \big| \, P, \mathbf{u}, \pmb{\theta}_\bullet )$
 %
@@
 surfaces sampled from the true distribution we are trying to model: that is,
-a surface with $N$ discrete circular spots of fixed, uniform contrast $c$
-and radius $r \pm \Delta$ at latitudes $\mu_\phi \pm \sigma_\phi$.
+a surface with $n=5$ discrete circular spots of fixed, uniform contrast $c=0.1$
+and radius $r=20^\circ$ at latitudes $\mu_\phi \pm \sigma_\phi = 30^\circ \pm 5^\circ$.
 Let us then expand each surface in spherical harmonics and visualize the
-distribution of coefficients $y^l_m$. Figure~\ref{fig:nongaussianity}
+distribution of coefficients $\mathbbb{y}$. Figure~\ref{fig:nongaussianity}
 shows the joint distribution for five of the coefficients with the most
@@
             %
-            $N$
+            $\mathbb{n}$
              & number of spots
@@
             %
-            $c$
+            $\mathbb{c}$
              & spot contrast
@@
             %
-            $\phi$
+            $\bbphi$
              & spot latitude
@@
             \\
-            $\lambda$
+            $\bblambda$
              & spot longitude
@@
             %
-            $r$
+            $\mathbb{r}$
              & spot radius
@@
             %
-            $I$
+            $\mathbb{I}$
              & stellar inclination
@@
 intialized
-to zero. We then add $N$ spots to this grid, each of fractional contrast
-$c$ and radius $r$ centered at latitude $\phi$ and longitude $\lambda$,
-by decreasing the intensity at all points within an angular distance $r$
-(measured along the surface of the sphere) by an amount $c$.
+to zero. We then add $\mathbb{n}$ spots to this grid, each of fractional contrast
+$\mathbb{c}$ and radius $\mathbb{r}$ centered at latitude $\bbphi$ and longitude $\bblambda$,
+by decreasing the intensity at all points within an angular distance $\mathbb{r}$
+(measured along the surface of the sphere) by an amount $\mathbb{c}$.
 In order to compute the corresponding light curve, we expand the surface
@@
 using the \starry algorithm
-(Appendix \ref{sec:starry}), assuming an inclination $I$, a rotational period $P$,
+(Appendix \ref{sec:starry}), assuming an inclination $\mathbb{I}$, a rotational period $P$,
 and limb darkening coefficients $\mathbf{u}$. Finally, we divide the
@@
 \footnote{
-    In theory, we should do this in the reverse order: i.e., we should add photon
-    noise and \emph{then} normalize the light curve to the mean, since that is the order in which
+    In theory, we should do this in the reverse order: we should add photon
+    noise and \emph{then} normalize the light curve to the mean, as that is the order in which
     those steps occur in reality. However, if we did that, we would have to normalize $\sigma_f$
@@
     practice, the difference between these two approaches has negligible effect on
-    our inference results, so we opt for the faster of the two methods. Note, of course,
+    our results, so we opt for the faster of the two methods. Note, of course,
     that when applying our GP to real data, we won't have this choice!
@@
 we find that a single light curve measurement can only tell us $\sim 9\%$
-of the information about the surface.
+of the information about the surface on those scales.
 
@@
 
+\section{Caveats}
+\label{sec:caveats}
+The latitude distribution is not Gaussian. The number of spots does not have to be an integer.
+
+\subsection{Other distributions}
+\label{sec:other-dist}
+Uniform spot radius distribution.
+
 \section{Implementation}
@@
 the following conventions throughout this paper:
-integers are represented by italic uppercase letters (i.e., $N$),
+integers are represented by italic uppercase letters (i.e., $K$),
 scalars are represented by italic lowercase
@@
 
-The exception to our indexing notation is if the vector or matrix
-represents a quantity in the spherical harmonic basis, in which case we use
+We also make an explicit distinction
+between numerical quantities and random variables. The former are typeset
+in serif font (as above), while the latter are typeset in blackboard font.
+For example, the quantity $\mathbb{x}$ denotes a scalar random variable,
+while $x$ denotes a particular realization of that variable. The same
+applies to vector-valued ($\mathbf{x}$ is a realization of $\mathbbb{x}$)
+and matrix-valued random variables
+($\mathbf{X}$ is a realization of $\mathbbb{X}$).
+
+Much of the math in this paper involves vectors representing
+coefficients in the spherical harmonic basis, which are
+customarily indexed by two integers $l$ and $m$.
+We therefore make an exception to our indexing notation for
+quantities in the spherical harmonic basis: we use
 \emph{two} indices to represent a scalar vector element, $x^l_m$, and \emph{four}
@@
 %
-For completeness, the element of a spherical harmonic vector $\mathbf{x}$ with
+For completeness, the element of a spherical harmonic vector $\mathbbb{x}$ with
 degree $l$ and order $m$ is at (flattened) index
@@
 To understand the expression above, let us proceed from right to left,
-starting with the spherical harmonic vector $\mathbf{y}$, which we assume
+starting with the spherical harmonic vector $\mathbbb{y}$, which we assume
 describes the surface intensity of the star at time $t = 0$
@@
 we can construct a matrix $\mathbf{L}(\mathbf{u})$
-that transforms a spherical harmonic vector $\mathbf{y}$ of degree
+that transforms a spherical harmonic vector $\mathbbb{y}$ of degree
 $l_\mathrm{max}$ to a limb-darkened spherical harmonic vector $\mathbf{y'}$
@@
 ($n_\mathrm{max} = 1$) with coefficient vector $\mathbf{u} = ( u_1 )$.
-The transformation matrix from $\mathbf{y}$ to $\mathbf{y'}$ is
+The transformation matrix from $\mathbbb{y}$ to $\mathbf{y'}$ is
 %
@@
 first and second moments of the spherical harmonic representation of the
-stellar surface $\mathbf{y}$,
+stellar surface $\mathbbb{y}$,
 %
@@
     \label{eq:exp_y_app}
-    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
-     & =
-    \int \mathbf{y}(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
+    \mathrm{E} \Big[ \mathbbb{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
+     & =
+    \int \mathbbb{y}(\mathbbb{x} ) \, p(\mathbbb{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbbb{x}
     \\
     \label{eq:exp_yy_app}
-    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
-     & =
-    \int \mathbf{y}(\mathbf{x} ) \mathbf{y}^\top(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
+    \mathrm{E} \Big[ \mathbbb{y} \, \mathbbb{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
+     & =
+    \int \mathbbb{y}(\mathbbb{x} ) \mathbbb{y}^\top(\mathbbb{x} ) \, p(\mathbbb{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbbb{x}
     \quad,
@@
 %
-Recall that $\mathbf{x}$ is a vector of parameters describing the particular
+Recall that $\mathbbb{x}$ is a vector of parameters describing the exact
 configuration of features on the surface of a star, and
-$p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$ is its probability
+$p(\mathbbb{x} \, \big| \, \pmb{\theta}_\bullet)$ is its probability
 density function conditioned on hyperparameters $\pmb{\theta}_\bullet$,
@@
 \begin{align}
-    \mathbf{x} =
+    \mathbbb{x} =
     \left(
-    \eta \,\,\,\,
-    \xi_0 \,\, \cdots \,\, \xi_{\eta-1} \,\,\,\,
-    \lambda_0 \,\, \cdots \,\, \lambda_{\eta-1} \,\,\,\,
-    \phi_0 \,\, \cdots \,\, \phi_{\eta-1} \,\,\,\,
-    \rho_0 \,\, \cdots \,\, \rho_{\eta-1}
+    \mathbb{n} \,\,\,\,
+    \mathbb{c}_0 \,\, \cdots \,\, \mathbb{c}_{\mathbb{n}-1} \,\,\,\,
+    \bblambda_0 \,\, \cdots \,\, \bblambda_{\mathbb{n}-1} \,\,\,\,
+    \bbphi_0 \,\, \cdots \,\, \bbphi_{\mathbb{n}-1} \,\,\,\,
+    \mathbb{r}_0 \,\, \cdots \,\, \mathbb{r}_{\mathbb{n}-1}
     \right)^\top
@@
     \label{eq:RRs}
-    \mathbf{y}(\mathbf{x}) =
-    \sum_{n=0}^{\eta-1}
-    \xi_n
+    \mathbbb{y}(\mathbbb{x}) =
+    \sum_{i=0}^{\mathbb{n}-1}
+    \mathbb{c}_i
     \,
-    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda_n)
+    \mathbf{R}_{\hat{\mathbf{y}}}(\bblambda_i)
     \,
-    \mathbf{R}_{\hat{\mathbf{x}}}(\phi_n)
+    \mathbf{R}_{\hat{\mathbf{x}}}(\bbphi_i)
     \,
-    \mathbf{s}(\rho_n)
-    \quad,
-\end{align}
-%
-where $\eta$ is the total number of spots,
-$\xi_n$ is the contrast of the $n^\mathrm{th}$ spot,
-$\lambda_n$ is its longitude, $\phi_n$ is its latitude,
-and $\rho_n$ is its radius.
-The vector function $\mathbf{s}(r_n)$
+    \mathbf{s}(\mathbb{r}_i)
+    \quad,
+\end{align}
+%
+where $\mathbb{n}$ is the total number of spots,
+$\mathbb{c}_i$ is the contrast of the $i^\mathrm{th}$ spot,
+$\bblambda_i$ is its longitude, $\bbphi_i$ is its latitude,
+and $\mathbb{r}_i$ is its radius.
+The vector function $\mathbf{s}(\mathbb{r}_i)$
 returns the spherical harmonic expansion of a negative unit brightness
-circular spot of radius $r_n$ at $\lambda = \phi = 0$,
-$\mathbf{R}_{\hat{\mathbf{x}}}(\phi_n)$ is the Wigner matrix that rotates the
+circular spot of radius $\mathbb{r}_i$ at $\bblambda = \bbphi = 0$,
+$\mathbf{R}_{\hat{\mathbf{x}}}(\bbphi_i)$ is the Wigner matrix that rotates the
 expansion about $\hat{\mathbf{x}}$ such that the spot is centered at a
-latitude $\phi_n$, and $\mathbf{R}_{\hat{\mathbf{y}}}(\lambda_n)$ is the Wigner
+latitude $\bbphi_i$, and $\mathbf{R}_{\hat{\mathbf{y}}}(\bblambda_i)$ is the Wigner
 matrix that then rotates the
 expansion about $\hat{\mathbf{y}}$ such that the spot is centered at a
-longitude $\lambda_n$; these three functions are detailed in the sections below.
+longitude $\bblambda_i$; these three functions are detailed in the sections below.
 %
 Equation~(\ref{eq:RRs}) thus provides a way of converting a random variable
-$\mathbf{x}$ describing the size, brightness, and position of spots to the
+$\mathbbb{x}$ describing the size, brightness, and position of spots to the
 corresponding representation in terms of spherical harmonics.
@@
 Regarding this equation,
-two things should be noted. First, we define $\mathbf{y}$ relative to
+two things should be noted. First, we define $\mathbbb{y}$ relative to
 a baseline of zero: i.e., a star with no spots on it will have
@@
 we are not interested in any specific value of
-$\mathbf{y}$; rather, we would like to know its expectation value under
-the probability distribution governing the different spot properties $\mathbf{x}$,
-i.e., $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$.
+$\mathbbb{y}$; rather, we would like to know its expectation value under
+the probability distribution governing the different spot properties $\mathbbb{x}$,
+i.e., $p(\mathbbb{x} \, \big| \, \pmb{\theta}_\bullet)$.
 %
@@
 For simplicity, we assume that
-the total number of spots is fixed to a value $N$, i.e.,
-%
-\begin{align}
-    p(\eta \, \big| \, \pmb{\theta}_\bullet)
-     & =
-    \delta(\eta - N)
+the total number of spots is fixed to a value $n$, i.e.,
+%
+\begin{align}
+    p(\mathbb{n} \, \big| \, \pmb{\theta}_\bullet)
+     & =
+    \delta(\mathbb{n} - n)
     \quad,
@@
     definition. It is less justified when the GP is used to model an ensemble of
-    stars, where each star may have a different total number of spots $\eta$.
-    However, as we argue in the text, $\eta$ is extremely difficult to constrain
+    stars, where each star may have a different total number of spots $\mathbb{n}$.
+    However, as we argue in the text, $\mathbb{n}$ is extremely difficult to constrain
     from light curves, in particular because of how degenerate it is with
     the spot contrast. In practice, we find that assuming that all stars in the
-    ensemble have the same number of spots $N$ leads to higher variance in the
-    estimate of $N$, but it does not lead to noticeable bias in $N$ or in any of
+    ensemble have the same number of spots $n$ leads to higher variance in the
+    estimate of $n$, but it does not lead to noticeable bias in $n$ or in any of
     the other hyperparameters.
@@
 We further asume that
-$p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$
+$p(\mathbbb{x} \, \big| \, \pmb{\theta}_\bullet)$
 is separable in each of the four other spot properties, and that all of the spots
@@
 \begin{align}
-    p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)
+    p(\mathbbb{x} \, \big| \, \pmb{\theta}_\bullet)
     =
-    \prod_{n=0}^{N-1}
-    p(\xi_n \, \big| \, \pmb{\theta}_{\xi}) \,
-    p(\lambda_n \, \big| \, \pmb{\theta}_{\lambda}) \,
-    p(\phi_n \, \big| \, \pmb{\theta}_{\phi})\,
-    p(\rho_n \, \big| \, \pmb{\theta}_{\rho})
+    \prod_{i=0}^{n-1}
+    p(\mathbb{c}_i \, \big| \, \pmb{\theta}_{c}) \,
+    p(\bblambda_i \, \big| \, \pmb{\theta}_{\lambda}) \,
+    p(\bbphi_i \, \big| \, \pmb{\theta}_{\phi})\,
+    p(\mathbb{r}_i \, \big| \, \pmb{\theta}_{r})
     \quad,
@@
     \pmb{\theta}_\bullet = \left(
-    N \, \,
-    \pmb{\theta}_{\xi} \, \,
+    n \, \,
+    \pmb{\theta}_{c} \, \,
     \pmb{\theta}_{\lambda} \, \,
     \pmb{\theta}_{\phi} \, \,
-    \pmb{\theta}_{\rho} \right)^\top
+    \pmb{\theta}_{r} \right)^\top
 \end{align}
@@
 is the vector of hyperparameters describing the process and
-$\pmb{\theta}_{\xi}$,
+$\pmb{\theta}_{c}$,
 $\pmb{\theta}_{\lambda}$,
 $\pmb{\theta}_{\phi}$, and
-$\pmb{\theta}_{\rho}$ are yet to be specified.
+$\pmb{\theta}_{r}$ are yet to be specified.
 %
@@
     \label{eq:exp_y_sep}
-    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
-     & =
-    N \mathbf{e}_\xi
+    \mathrm{E} \Big[ \mathbbb{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
+     & =
+    n \, \mathbf{e}_c
     \\[1em]
     \label{eq:exp_yy_sep}
-    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
-     & =
-    N \mathbf{E}_\xi
+    \mathrm{E} \Big[ \mathbbb{y} \, \mathbbb{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
+     & =
+    n \, \mathbf{E}_c
 \end{align}
@@
     \label{eq:e1}
-    \mathbf{e}_\rho
+    \mathbf{e}_r
      & \equiv
     \int
-    \mathbf{s}(\rho) \,
-    p(\rho \, \big| \, \pmb{\theta}_{\rho}) \,
-    \mathrm{d}\rho
+    \mathbf{s}(\mathbb{r}) \,
+    p(\mathbb{r} \, \big| \, \pmb{\theta}_{r}) \,
+    \mathrm{d}\mathbb{r}
     %
@@
     \int
-    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{e}_\rho \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
-    \mathrm{d}\phi
+    \mathbf{R}_{\hat{\mathbf{x}}}(\bbphi) \,
+    \mathbf{e}_r \,
+    p(\bbphi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\bbphi
     %
@@
     \int
-    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{R}_{\hat{\mathbf{y}}}(\bblambda) \,
     \mathbf{e}_\phi \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
+    p(\bblambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\bblambda
     \\[1em]
     \label{eq:e4}
-    \mathbf{e}_\xi
+    \mathbf{e}_c
      & \equiv
     \int
-    \xi \,
+    \mathbb{c} \,
     \mathbf{e}_\lambda \,
-    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
-    \mathrm{d}\xi
+    p(\mathbb{c} \, \big| \, \pmb{\theta}_{c}) \,
+    \mathrm{d}\mathbb{c}
     %
@@
     \label{eq:E1}
-    \mathbf{E}_\rho
+    \mathbf{E}_r
      & \equiv
     \int
-    \mathbf{s}(\rho) \, \mathbf{s}^\top(\rho) \,
-    p(\rho \, \big| \, \pmb{\theta}_{\rho}) \,
-    \mathrm{d}\rho
+    \mathbf{s}(\mathbb{r}) \, \mathbf{s}^\top(\mathbb{r}) \,
+    p(\mathbb{r} \, \big| \, \pmb{\theta}_{r}) \,
+    \mathrm{d}\mathbb{r}
     %
@@
     \int
-    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{E}_\rho \,
-    \mathbf{R}_{\hat{\mathbf{x}}}^\top(\phi) \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi})
-    \mathrm{d}\phi
+    \mathbf{R}_{\hat{\mathbf{x}}}(\bbphi) \,
+    \mathbf{E}_r \,
+    \mathbf{R}_{\hat{\mathbf{x}}}^\top(\bbphi) \,
+    p(\bbphi \, \big| \, \pmb{\theta}_{\phi})
+    \mathrm{d}\bbphi
     %
@@
     \int
-    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{R}_{\hat{\mathbf{y}}}(\bblambda) \,
     \mathbf{E}_\phi \,
-    \mathbf{R}_{\hat{\mathbf{y}}}^\top(\lambda) \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda})
-    \mathrm{d}\phi
+    \mathbf{R}_{\hat{\mathbf{y}}}^\top(\bblambda) \,
+    p(\bblambda \, \big| \, \pmb{\theta}_{\lambda})
+    \mathrm{d}\bbphi
     \\[1em]
@@
     \label{eq:E4}
-    \mathbf{E}_\xi
+    \mathbf{E}_c
      & \equiv
     \int
-    \xi^2 \,
+    \mathbb{c}^2 \,
     \mathbf{E}_\lambda \,
-    p(\xi \, \big| \, \pmb{\theta}_\xi)
-    \mathrm{d}\xi
+    p(\mathbb{c} \, \big| \, \pmb{\theta}_c)
+    \mathrm{d}\mathbb{c}
     %
@@
 In Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), we used the fact
-that both the mean and the variance of the sum of $N$
-independent, identically-distributed random variables are equal to $N$
+that both the mean and the variance of the sum of $n$
+independent, identically-distributed random variables are equal to $n$
 times the individual mean and the variance, respectively.
@@
 Below we compute the first and second moments
-of the radius distribution ($\mathbf{e}_\rho$, $\mathbf{E}_\rho$) under
-a suitable spherical harmonic expansion $\mathbf{s}(\rho)$ of the spot profile
+of the radius distribution ($\mathbf{e}_r$, $\mathbf{E}_r$) under
+a suitable spherical harmonic expansion $\mathbf{s}(\mathbb{r})$ of the spot profile
 and a suitable probability distribution function for the spot radius,
-$p(\rho \, \big| \, \pmb{\theta}_{\rho})$.
+$p(\mathbb{r} \, \big| \, \pmb{\theta}_{r})$.
 
@@
 We model the brightness $b$ an angle $\vartheta$ away from the
-center of a spot of negative unit intensity and radius $\rho$ as
+center of a spot of negative unit intensity and radius $\mathbb{r}$ as
 %
@@
     \label{eq:brvartheta}
-    b(\rho; \vartheta) & = \frac{1}{1 + \exp\left(\dfrac{\rho-\vartheta}{s}\right)} - 1
+    b(r; \vartheta) & = \frac{1}{1 + \exp\left(\dfrac{r-\vartheta}{s}\right)} - 1
 \end{align}
@@
 for some (constant) shape parameter $s$. In the limit $s \rightarrow 0$, $b$ approaches an
-inverted top-hat function with half-width equal to $\rho$,
+inverted top-hat function with half-width equal to $r$,
 corresponding to a circular spot of uniform intensity. For $s > 0$, each half
-of $b$ is a sigmoid with half-width at half-minimum equal to $\rho$.
+of $b$ is a sigmoid with half-width at half-minimum equal to $r$.
 In our implementation of the algorithm we choose $s = 0.2^\circ$,
@@
 \begin{align}
-    \mathbf{B} \, \tilde{\mathbf{s}}(\rho) = \mathbf{b}(\rho)
-\end{align}
-%
-%
-where $\mathbf{b}(\rho)$ is computed by evaluating Equation~(\ref{eq:brvartheta})
+    \mathbf{B} \, \tilde{\mathbf{s}}(r) = \mathbf{b}(r)
+\end{align}
+%
+%
+where $\mathbf{b}(r)$ is computed by evaluating Equation~(\ref{eq:brvartheta})
 at each of the $\vartheta_k$,
@@
 %
-and $\tilde{\mathbf{s}}(\rho)$ are the coefficients of the expansion.
+and $\tilde{\mathbf{s}}(r)$ are the coefficients of the expansion.
 These are related to the full vector of spherical harmonic coefficients describing
-the spot, $\mathbf{s}(\rho)$, by
-%
-\begin{align}
-    s^l_m(\rho) = \tilde{s}_l(\rho) \delta_{m,0}
+the spot, $\mathbf{s}(r)$, by
+%
+\begin{align}
+    s^l_m(r) = \tilde{s}_l(r) \delta_{m,0}
     \quad,
@@
 \begin{align}
-    \mathbf{s}(\rho) = \pmb{\mathcal{I}} \, \tilde{\mathbf{s}}(\rho)
+    \mathbf{s}(r) = \pmb{\mathcal{I}} \, \tilde{\mathbf{s}}(r)
 \end{align}
@@
 and $\delta$ is the Kronecker delta function.
-To find the coefficients $\tilde{\mathbf{s}}(\rho)$
-(and hence $\mathbf{s}(\rho)$), we solve the (linear) inverse problem,
-%
-\begin{align}
-    \tilde{\mathbf{s}}(\rho) & = \mathbf{B}^+ \, \mathbf{b}(\rho)
+To find the coefficients $\tilde{\mathbf{s}}(r)$
+(and hence $\mathbf{s}(r)$), we solve the (linear) inverse problem,
+%
+\begin{align}
+    \tilde{\mathbf{s}}(r) & = \mathbf{B}^+ \, \mathbf{b}(r)
 \end{align}
@@
             Intensity profiles for spots with different
-            radii $\rho$ computed at spherical harmonic
+            radii $r$ computed at spherical harmonic
             degree $l_{\mathrm{max}} = \LMAX$.
-            For $\rho \gtrsim 10^\circ$, the spherical harmonic expansion
+            For $r \gtrsim 10^\circ$, the spherical harmonic expansion
             captures the spot shape and intensity reasonably well,
@@
 without the smoothing term (i.e., $\sigma = \infty$). However, for
-$\rho \lesssim 10^\circ$, an expansion to $l_{\mathrm{max}} = \LMAX$ is
+$r \lesssim 10^\circ$, an expansion to $l_{\mathrm{max}} = \LMAX$ is
 insufficient to correctly model the spot, as can be seen from the
-$\rho = 5^\circ$ profile (dashed curve). Expansions to higher spherical harmonic
+$r = 5^\circ$ profile (dashed curve). Expansions to higher spherical harmonic
 degree allow one to model spots with radii smaller than $10^\circ$, although
@@
 For simplicity, we will adopt a uniform probability distribution for the
-spot radius, characterized by a mean radius $r$ and a half-width $\Delta$:
-%
-\begin{align}
-    p(\rho \, \big| \, \pmb{\theta}_{\rho})
+spot radius, characterized by a mean radius $r$ and a half-width $\Delta r$:
+%
+\begin{align}
+    p(\mathbb{r} \, \big| \, \pmb{\theta}_{r})
      & =
     \begin{cases}
-        \frac{1}{2\Delta} & r - \Delta \leq r \leq r + \Delta
+        \frac{1}{2\Delta r} & r - \Delta r \leq r \leq r + \Delta r
         \\
@@
 \begin{align}
-    \pmb{\theta}_\rho = \left(
+    \pmb{\theta}_\mathbb{r} = \left(
     r \, \, \, \,
-    \Delta \right)^\top
+    \Delta r \right)^\top
     \quad.
@@
 therefore useful to also consider the limiting case of the radius distribution
-as $\Delta \rightarrow 0$, in which case the PDF becomes
-%
-\begin{align}
-    p(\rho \, \big| \, \pmb{\theta}_{\rho}, \Delta = 0)
-     & =
-    \delta(\rho - r)
+as $\Delta r \rightarrow 0$, in which case the PDF becomes
+%
+\begin{align}
+    p(\mathbb{r} \, \big| \, \pmb{\theta}_{r}, \Delta r = 0)
+     & =
+    \delta(\mathbb{r} - r)
     \quad,
@@
 \begin{align}
-    \mathbf{e}_\rho
+    \mathbf{e}_r
      & \equiv
     \int
-    \mathbf{s}(\rho) \,
-    p(\rho \, \big| \, \pmb{\theta}_{\rho}) \,
-    \mathrm{d}\rho
+    \mathbf{s}(\mathbb{r}) \,
+    p(\mathbb{r} \, \big| \, \pmb{\theta}_{r}) \,
+    \mathrm{d}\mathbb{r}
     \nonumber \\
      & =
-    \frac{1}{2\Delta}
-    \int_{r_0 - \Delta}^{r_0 + \Delta}
-    \mathbf{s}(\rho) \,
-    \mathrm{d}\rho
+    \frac{1}{2\Delta r}
+    \int_{r_0 - \Delta r}^{r_0 + \Delta r}
+    \mathbf{s}(\mathbb{r}) \,
+    \mathrm{d}\mathbb{r}
     \quad.
@@
 \begin{proof}{test_erho}
-    (e_\rho)^l_m
-    & =
-    \frac{\delta_{m,0}}{2\Delta}
-    \int_{r - \Delta}^{r+ \Delta}
-    \tilde{s}_{l}(\rho) \,
-    \mathrm{d}\rho
+    (e_\mathbb{r})^l_m
+    & =
+    \frac{\delta_{m,0}}{2\Delta r}
+    \int_{r - \Delta r}^{r+ \Delta r}
+    \tilde{s}_{l}(\mathbb{r}) \,
+    \mathrm{d}\mathbb{r}
     \nonumber \\
     & =
-    \frac{\delta_{m,0}}{2\Delta}
+    \frac{\delta_{m,0}}{2\Delta r}
     \sum_{k=0}^{K-1} B^+_{l,k}
-    \int_{r - \Delta}^{r + \Delta}
-    b_{k}(\rho)
-    \mathrm{d}\rho
+    \int_{r - \Delta r}^{r + \Delta r}
+    b_{k}(\mathbb{r})
+    \mathrm{d}\mathbb{r}
     \nonumber \\
@@
     \sum_{k=0}^{K-1} B^+_{l,k}
-    c_k(r, \Delta)
+    c_k(r, \Delta r)
 \end{proof}
@@
 \begin{proof}{test_erho}
-    c_k(r, \Delta) & =
-    \frac{s}{2\Delta}
+    c_k(r, \Delta r) & =
+    \frac{s}{2\Delta r}
     \ln
@@
     \frac{
-            1+\chi_k^-(r, \Delta)
+            1+\chi_k^-(r, \Delta r)
         }
         {
-            1+\chi_k^+(r, \Delta)
+            1+\chi_k^+(r, \Delta r)
         }
@@
 \begin{proof}{test_erho}
-    \chi_k^-(r, \Delta) & \equiv \exp\left(\frac{r -\Delta -\vartheta_k}{s}\right)
+    \chi_k^-(r, \Delta r) & \equiv \exp\left(\frac{r -\Delta r -\vartheta_k}{s}\right)
     \nonumber                                                              \\
-    \chi_k^+(r, \Delta) & \equiv \exp\left(\frac{r +\Delta -\vartheta_k}{s}\right)
+    \chi_k^+(r, \Delta r) & \equiv \exp\left(\frac{r +\Delta r -\vartheta_k}{s}\right)
     \quad.
@@
 \begin{proof}{test_erho}
-    \mathbf{e}_\rho
+    \mathbf{e}_r
     & =
     \pmb{\mathcal{I}} \,
-    \mathbf{B}^+ \, \mathbf{c}(r, \Delta)
+    \mathbf{B}^+ \, \mathbf{c}(r, \Delta r)
     \quad.
@@
 %
-Note, finally, that in the limit $\Delta \rightarrow 0$,
+Note, finally, that in the limit $\Delta r \rightarrow 0$,
 %
 \begin{proof}{test_erho}
-    \lim_{\Delta \rightarrow 0}
-    \mathbf{e}_\rho
+    \lim_{\Delta r \rightarrow 0}
+    \mathbf{e}_r
     & =
@@
 \begin{align}
-    \mathbf{E}_\rho
+    \mathbf{E}_r
      & \equiv
     \int
-    \mathbf{s}(\rho) \,
-    \mathbf{s}^\top(\rho) \,
-    p(\rho \, \big| \, \pmb{\theta}_{\rho}) \,
-    \mathrm{d}\rho
+    \mathbf{s}(\mathbb{r}) \,
+    \mathbf{s}^\top(\mathbb{r}) \,
+    p(\mathbb{r} \, \big| \, \pmb{\theta}_{r}) \,
+    \mathrm{d}\mathbb{r}
     \nonumber \\
      & =
-    \frac{1}{2\Delta}
-    \int_{r - \Delta}^{r + \Delta}
-    \mathbf{s}(\rho) \,
-    \mathbf{s}^\top(\rho) \,
-    \mathrm{d}\rho
+    \frac{1}{2\Delta r}
+    \int_{r - \Delta r}^{r + \Delta r}
+    \mathbf{s}(\mathbb{r}) \,
+    \mathbf{s}^\top(\mathbb{r}) \,
+    \mathrm{d}\mathbb{r}
     \quad.
@@
 \begin{proof}{test_bigErho}
-    (E_\rho)^{l,l'}_{m,m'}
-    & =
-    \frac{\delta_{m,0}\delta_{m',0}}{2\Delta}
-    \int_{r - \Delta}^{r + \Delta}
-    \tilde{s}^l_{m}(\rho) \,
-    \tilde{s}^{l'}_{m'}(\rho) \,
-    \mathrm{d}\rho
+    (E_\mathbb{r})^{l,l'}_{m,m'}
+    & =
+    \frac{\delta_{m,0}\delta_{m',0}}{2\Delta r}
+    \int_{r - \Delta r}^{r + \Delta r}
+    \tilde{s}^l_{m}(\mathbb{r}) \,
+    \tilde{s}^{l'}_{m'}(\mathbb{r}) \,
+    \mathrm{d}\mathbb{r}
     \nonumber \\
     & =
-    \frac{\delta_{m,0}\delta_{m',0}}{2\Delta}
+    \frac{\delta_{m,0}\delta_{m',0}}{2\Delta r}
     \sum_{k=0}^{K-1} B^+_{l,k}
     \sum_{k'=0}^{K-1} B^+_{l',k'}
-    \int_{r - \Delta}^{r + \Delta}
-    b_{k}(\rho)
-    b_{k'}(\rho)
-    \mathrm{d}\rho
+    \int_{r - \Delta r}^{r + \Delta r}
+    b_{k}(\mathbb{r})
+    b_{k'}(\mathbb{r})
+    \mathrm{d}\mathbb{r}
     \nonumber \\
@@
     \sum_{k'=0}^{K-1} B^+_{l',k'}
-    C_{k,k'}(r,\Delta)
+    C_{k,k'}(r,\Delta r)
     \quad,
@@
 \begin{proof}{test_bigErho}
-    C_{k,k'}(r,\Delta)
-    & \equiv
-    \frac{s}{2\Delta}
+    C_{k,k'}(r,\Delta r)
+    & \equiv
+    \frac{s}{2\Delta r}
     \begin{cases}
@@
         }
-         & k \ne k'
+         & \hspace{-2em} k \ne k'
         \\[0.5em]
@@
         - 1
-         & k = k'
+         & \hspace{-2em} k = k'
         \quad.
@@
 \begin{proof}{test_bigErho}
-    \mathbf{E}_\rho
+    \mathbf{E}_r
     &=
     \pmb{\mathcal{I}} \,
-    \mathbf{B}^+ \mathbf{C} (r, \Delta) {\mathbf{B}^+}^\top
+    \mathbf{B}^+ \mathbf{C} (r, \Delta r) {\mathbf{B}^+}^\top
     \pmb{\mathcal{I}}^\top
@@
 %
-Finally, in the limit $\Delta \rightarrow 0$,
+Finally, in the limit $\Delta r \rightarrow 0$,
 %
 \begin{proof}{test_bigErho}
-    \lim_{\Delta \rightarrow 0}
-    \mathbf{E}_\rho
+    \lim_{\Delta r \rightarrow 0}
+    \mathbf{E}_r
     & =
@@
 \begin{align}
-    D_{m,m'}^l(\alpha, \beta, \gamma) = \mathrm{e}^{-\imag m' \alpha}
-    d_{m,m'}^l(\beta) \mathrm{e}^{-\imag m \gamma}
-\end{align}
-%
-where $\alpha$, $\beta$, and $\gamma$ are the Euler angles describing
+    D_{m,m'}^l(\upalpha, \upbeta, \upgamma) = \exp\left({-\imag m' \upalpha}\right)
+    d_{m,m'}^l(\upbeta) \exp\left({-\imag m \upgamma}\right)
+\end{align}
+%
+where $\upalpha$, $\upbeta$, and $\upgamma$ are the Euler angles describing
 the rotation in the
@@
 and $\imag$ is the imaginary unit. The terms of the $d$-matrix
-depend on powers of $\sin\left(\nicefrac{\beta}{2}\right)$
-and $\cos\left(\nicefrac{\beta}{2}\right)$
+depend on powers of $\sin\left(\nicefrac{\upbeta}{2}\right)$
+and $\cos\left(\nicefrac{\upbeta}{2}\right)$
 \citep[c.f. Equation C15 in][]{Luger2019}, but it is convenient to use
@@
 \begin{proof}{test_dlmmp}
-    d_{m,m'}^l(\beta) =
+    d_{m,m'}^l(\upbeta) =
     \sum\limits_{i=0}^{2l} c_{m,m',i}^{l}
-    \mathrm{sgn}(\sin\beta)^{i}
-    (1 - \cos\beta)^{\frac{2l - i}{2}}
-    (1 + \cos\beta)^\frac{i}{2}
+    \mathrm{sgn}(\sin\upbeta)^{i}
+    (1 - \cos\upbeta)^{\frac{2l - i}{2}}
+    (1 + \cos\upbeta)^\frac{i}{2}
 \end{proof}
@@
                 %
-                 & m - m' - i \,\, \mathrm{even}
+                 & \hspace{-2em} m - m' - i \,\, \mathrm{even}
                 %
@@
                 %
-                 & m - m' - i \,\, \mathrm{odd}
+                 & \hspace{-2em} m - m' - i \,\, \mathrm{odd}
             \end{cases}
@@
 The latitude integrals (Equations~\ref{eq:e2} and \ref{eq:E2}) involve
-rotations by an angle $\phi$ about $\hat{\mathbf{x}}$, which
+rotations by an angle $\bbphi$ about $\hat{\mathbf{x}}$, which
 may be accomplished by choosing
-Euler angles $\alpha = \nicefrac{\pi}{2}$, $\beta = \phi$, and
-$\gamma = -\nicefrac{\pi}{2}$, such that
-%
-\begin{align}
-    \mathbf{R}^l_{\hat{\mathbf{x}}}(\phi)
-     & =
-    {\mathbf{U}^l}^{-1} \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi) \mathbf{U}^l
+Euler angles $\upalpha = \nicefrac{\pi}{2}$, $\upbeta = \bbphi$, and
+$\upgamma = -\nicefrac{\pi}{2}$, such that
+%
+\begin{align}
+    \mathbf{R}^l_{\hat{\mathbf{x}}}(\bbphi)
+     & =
+    {\mathbf{U}^l}^{-1} \mathbf{D}^l_{\hat{\mathbf{x}}}(\bbphi) \mathbf{U}^l
 \end{align}
@@
 \begin{align}
-    \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi)
-     & =
-    \mathbf{D}^l\left(\frac{\pi}{2}, \phi, -\frac{\pi}{2}\right)
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(\bbphi)
+     & =
+    \mathbf{D}^l\left(\frac{\pi}{2}, \bbphi, -\frac{\pi}{2}\right)
     \quad.
@@
 From the expressions above, it is clear that all terms in
-$\mathbf{R}_{\hat{\mathbf{x}}}(\phi)$ are equal to (weighted) sums of powers
-of $(1 \pm \cos\phi)$.
+$\mathbf{R}_{\hat{\mathbf{x}}}(\bbphi)$ are equal to (weighted) sums of powers
+of $(1 \pm \cos\bbphi)$.
 %
@@
 a probability density function, it is convenient to model
-$\cos\phi$ as a Beta-distributed variable. As we will see, this
+$\cos\bbphi$ as a Beta-distributed variable. As we will see, this
 choice will allow us to
 analytically compute the first two moments of the distribution of
-$\phi$ conditioned on $\pmb{\theta}_\phi$.
-
-%
-The Beta distribution in $\cos\phi$ has hyperparameters $\alpha$ and $\beta$
-and PDF given by
+$\bbphi$ conditioned on $\pmb{\theta}_\phi$.
+
+%
+The Beta distribution in $\cos\bbphi$ has hyperparameters $\alpha$ and $\beta$
+(not to be confused with the Euler angles $\upalpha$ and $\upbeta$) and PDF given by
 %
@@
     \label{eq:cosphi-pdf}
-    p \big(\cos\phi \, \big| \, \alpha, \beta \big)
+    p \big(\cos\bbphi \, \big| \, \alpha, \beta \big)
      & =
     \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
-    (\cos\phi)^{\alpha - 1}
-    (1 - \cos\phi)^{\beta - 1}
-    \quad,
-\end{align}
-%
-where $\Gamma$ is the Gamma function. The implied distribution for $\phi$
+    (\cos\bbphi)^{\alpha - 1}
+    (1 - \cos\bbphi)^{\beta - 1}
+    \quad,
+\end{align}
+%
+where $\Gamma$ is the Gamma function. The implied distribution for $\bbphi$
 may be computed by a straightforward change of variable:
@@
     \label{eq:phi-pdf}
-    p \big(\phi \, \big| \, \alpha, \beta \big)
+    p \big(\bbphi \, \big| \, \alpha, \beta \big)
     & =
@@
     \big|
-    \sin\phi
+    \sin\bbphi
     \big|
-    (\cos\phi)^{\alpha - 1}
-    (1 - \cos\phi)^{\beta - 1}
+    (\cos\bbphi)^{\alpha - 1}
+    (1 - \cos\bbphi)^{\beta - 1}
     \quad,
@@
 %
-for $\phi \in \left[ -\frac{\pi}{2}, \frac{\pi}{2} \right]$.
+for $\bbphi \in \left[ -\frac{\pi}{2}, \frac{\pi}{2} \right]$.
 %
@@
     \label{eq:beta2gauss}
-    \alpha & = e^{K_{00} + (K_{10} - K_{00})a}
+    \alpha & = \exp\left({K_{00} + (K_{10} - K_{00})a}\right)
     \nonumber                                                 \\
-    \beta  & = e^{K_{10} + (K_{11} - K_{10})b}
+    \beta  & = \exp\left({K_{10} + (K_{11} - K_{10})b}\right)
     \quad,
@@
 \begin{align}
+    \label{eq:ab}
     \pmb{\theta}_\phi = \left(
@@
 %
-
-Although we use $a$ and $b$ as the inputs to our algorithm, below
-we derive expressions for the moments of the distribution in terms
-of $\alpha$ and $\beta$, as this is somewhat more convenient. Note that
-this requires us to explicitly add a Jacobian term to the likelihood
-when doing posterior inference to account for the change in volume due
-to the transformation from $\alpha, \beta$ to $a, b$. In practice, however,
-$\alpha$ and $\beta$ are no more natural a set of parameters to sample in,
-since they do not \emph{obviously} relate to physical quantities of interest.
-It is therefore desirable to introduce one final transform, from $\alpha, \beta$
-to parameters controlling the mean (or mode) $\mu_\phi$ and
-standard deviation  $\sigma_\phi$ of the
-distribution of spot latitudes, which \emph{are} of physical interest.
-We introduce this transform, and the associated Jacobian for the complete
-transform $\mu_\phi, \sigma_\phi \rightarrow a, b$,
-in \S\ref{sec:lat-transform} below.
-
-\subsubsection{First moment}
-%
-Since the Wigner matrices are block diagonal, we may evaluate the moments of the
-distribution one spherical harmonic degree at a time. To that end, let us
-write the first moment integral as
-%
-\begin{proof}{test_ephi}
-    \mathbf{e}_\phi
-    & \equiv
-    \int
-    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{e}_r \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
-    \mathrm{d}\phi
-    \nonumber
-    \\
-    %
-    & =
-    %
-    \left(
-    \mathbf{e}_\phi^0
-    \,\,\,
-    \mathbf{e}_\phi^1
-    \,\,\,
-    \mathbf{e}_\phi^2
-    \,\,\,
-    \cdots
-    \,\,\,
-    \mathbf{e}_\phi^{l_{\mathrm{max}}}
-    \right)^\top
-    \quad,
-\end{proof}
-%
-where
-%
-\begin{proof}{test_ephi}
-    \mathbf{e}_\phi^l
-    & =
-    \int
-    \mathbf{R}^l_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{e}^l_r \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
-    \mathrm{d}\phi
-    \nonumber \\
-    %
-    & =
-    %
-    {\mathbf{U}^l}^{-1}
-    \mathbf{p}^l_\phi
-    \quad,
-\end{proof}
-%
-and we define
-%
-\begin{align}
-    \label{eq:plphi}
-    \mathbf{p}^l_\phi
-     & \equiv
-    \int
-    \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi) \,
-    \bar{\mathbf{e}}^l_r \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
-    \mathrm{d}\phi
-    \\
-    \bar{\mathbf{e}}^l_r
-     & \equiv
-    \mathbf{U}^l
-    \mathbf{e}^l_r
+Because of their trivial domain, these parameters are convenient to sample
+in when doing inference (provided we account for their implied prior on
+the spot latitudes; see below). However, $a$ and $b$ do not intuitively
+relate to physical quantities of interest. In many cases it is more
+desirable to parametrize the latitude distribution in terms of a parameter
+$\mu_\phi$ controlling the central latitude and a parameter $\sigma_\phi$
+controlling the dispersion in latitude among the spots.
+In this case, we may choose instead
+%
+\begin{align*}
+    \tag{\ref*{eq:ab}*}
+    \pmb{\theta}_\phi = \left(
+    \mu_\phi \, \, \, \,
+    \sigma_\phi \right)^\top
     \quad.
-\end{align}
-%
-The integral $\mathbf{p}_\phi^l$ defined above has a closed-form solution.
-To show this, we write the terms of $\mathbf{p}^l_\phi$ as
-%
-\begin{align}
-    {({p^l_\phi})_{}}_m
-     & =
-    \int
-    \sum\limits_{\mu=-l}^l
-    {({D^l_{\hat{\mathbf{x}}}})_{}}_{m,\mu}(\phi) \,
-    {({\bar{e}}^l_r)_{}}_{\mu} \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
-    \mathrm{d}\phi
-    %
-    \nonumber \\[0.5em]
-    %
-     & =
-    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
-    \sum\limits_{\mu=-l}^l
-    {({\bar{e}}^l_r)_{}}_{\mu}
-    \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu)}
-    \sum\limits_{i=0}^{2l} c_{m,\mu,i}^{l}
-    \,
-    {({q^l_\phi})_{}}_i(\pmb{\theta}_{\phi})
-    \quad,
-\end{align}
-%
-where
-%
-\begin{proof}{test_qlphii}
-    {({q^l_\phi})_{}}_i(\pmb{\theta}_{\phi})
-    & \equiv
-    \frac{1}{2}
-    \int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}
-    %
-    \mathrm{sgn}(\sin\phi)^{i}
-    \big| \sin\phi \big|
-    (\cos\phi)^{\alpha - 1}
-    (1 - \cos\phi)^{l + \beta - \frac{i}{2} - 1}
-    (1 + \cos\phi)^\frac{i}{2}
-    %
-    \,
-    \mathrm{d}\phi
-    \nonumber \\[0.5em]
-    %
-    & =
-    %
-    \begin{cases}
-        \displaystyle\int_{0}^{1}
-        %
-        x^{\alpha - 1}
-        (1 - x)^{l + \beta - \frac{i}{2} - 1 }
-        (1 + x)^\frac{i}{2}
-        %
-        \,
-        \mathrm{d}x
-        %
-         & i \,\, \mathrm{even}
-        %
-        \\
-        %
-        0
-        %
-         & i \,\, \mathrm{odd} \quad,
-    \end{cases}
-\end{proof}
-%
-and in the last line we made use of the transformation $x = \cos\phi$.
-%
-The integral in the expression above has a closed-form solution in terms
-of the hypergeometric function $_2F_1$:
-%
-\begin{proof}{test_qlphii}
-    %
-    \label{eq:qlphii}
-    {({q^l_\phi})_{}}_i(\pmb{\theta}_{\phi})
-    & =
-    \resizebox{.75\hsize}{!}{$
-            \begin{cases}
-                \dfrac{\Gamma(\alpha)\Gamma(l + \beta - \frac{i}{2})}{\Gamma(l + \alpha + \beta - \frac{i}{2})}
-                \,
-                _2F_1\Big(
-                -\frac{i}{2},
-                \alpha;
-                l + \alpha + \beta - \frac{i}{2};
-                -1
-                \Big)
-                %
-                 & i \,\, \mathrm{even}
-                %
-                \\
-                %
-                0
-                %
-                 & i \,\, \mathrm{odd} \quad.
-            \end{cases}
-        $}
-\end{proof}
-%
-In order to compute the integral $\mathbf{e}_\phi$
-(Equation~\ref{eq:e2}) we must evaluate Equation~(\ref{eq:qlphii})
-for all $0 \leq l \leq l_{\mathrm{max}}$, $0 \leq i \leq 2l$,
-which can be done efficiently via upward recursion relations for both
-the gamma function and the hypergeometric function.
-
-\subsubsection{Second moment}
-%
-Similarly as before, let us write the second moment integral as
-%
-\begin{proof}{test_bigEphi}
-    \label{eq:Ephi}
-    \mathbf{E}_\phi
-    & \equiv
-    \int
-    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{E}_r \,
-    \mathbf{R}_{\hat{\mathbf{x}}}^\top(\phi) \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
-    \mathrm{d}\phi
-    \nonumber
-    \\
-    %
-    & =
-    %
-    \setstackgap{L}{1.25\baselineskip}
-    \fixTABwidth{T}
-    \parenMatrixstack{
-    \mathbf{E}_\phi^{0,0} & \mathbf{E}_\phi^{0,1} & \mathbf{E}_\phi^{0,2} & \cdots & \mathbf{E}_\phi^{0,l_{\mathrm{max}}} \\
-    \mathbf{E}_\phi^{1,0} & \mathbf{E}_\phi^{1,1} & \mathbf{E}_\phi^{1,2} & \cdots & \mathbf{E}_\phi^{1,l_{\mathrm{max}}} \\
-    \mathbf{E}_\phi^{2,0} & \mathbf{E}_\phi^{2,1} & \mathbf{E}_\phi^{2,2} & \cdots & \mathbf{E}_\phi^{2,l_{\mathrm{max}}} \\
-    \vdots & \vdots & \vdots & \ddots & \vdots \\
-    \mathbf{E}_\phi^{l_{\mathrm{max}},0} & \mathbf{E}_\phi^{l_{\mathrm{max}},1} & \mathbf{E}_\phi^{l_{\mathrm{max}},2} & \cdots & \mathbf{E}_\phi^{l_{\mathrm{max}},l_{\mathrm{max}}}
-    }
-    \quad,
-\end{proof}
-%
-where
-%
-\begin{proof}{test_bigEphi}
-    \mathbf{E}_\phi^{l,l'}
-    & =
-    \int
-    \mathbf{R}^l_{\hat{\mathbf{x}}}(\phi) \,
-    \mathbf{E}^{l,l'}_r \,
-    {\mathbf{R}^{l'}_{\hat{\mathbf{x}}}}^\top(\phi) \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
-    \mathrm{d}\phi
-    \nonumber \\
-    %
-    & =
-    %
-    {\mathbf{U}^l}^{-1}
-    \mathbf{P}^{l,l'}_\phi
-    {{\mathbf{U}^{l'}}^{-1}}^\top
-\end{proof}
-%
-and we define
-%
-\begin{align}
-    \label{eq:Plphi}
-    \mathbf{P}^{l,l'}_\phi
-     & \equiv
-    \int
-    \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi) \,
-    \bar{\mathbf{E}}^{l,l'}_r \,
-    {\mathbf{D}^{l'}_{\hat{\mathbf{x}}}}^\top(\phi) \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
-    \mathrm{d}\phi
-    \\
-    \bar{\mathbf{E}}^{l,l'}_r
-     & \equiv
-    \mathbf{U}^l
-    \mathbf{E}^{l,l'}_r
-    {\mathbf{U}^{l'}}^\top
-    \quad.
-\end{align}
-%
-As before, we may express the solution to the integral $\mathbf{P}^{l,l'}_\phi$ in
-closed form. Let us write the terms of $\mathbf{P}^{l,l'}_\phi$ as
-%
-\begin{align}
-    {({P^{l,l'}_\phi})_{}}_{m,m'}
-     & =
-    \int
-    \sum\limits_{\mu=-l}^l
-    \sum\limits_{{\mu'}=-l'}^{l'}
-    {({D^l_{\hat{\mathbf{x}}}})_{}}_{m,\mu}(\phi) \,
-    {(\bar{E}^{l,l'}_r)_{}}_{\mu,{\mu'}} \,
-    {({D^{l'}_{\hat{\mathbf{x}}}})_{}}_{m',{\mu'}}(\phi) \,
-    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
-    \mathrm{d}\phi
-    %
-    \nonumber \\[0.5em]
-    %
-     & =
-    \resizebox{.75\hsize}{!}{$
-        \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
-        \sum\limits_{\mu=-l}^l
-        \sum\limits_{{\mu'}=-l'}^{l'}
-        {(\bar{E}^{l,l'}_r)_{}}_{\mu,{\mu'}} \,
-        \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu + m' - {\mu'})}
-        \sum\limits_{i=0}^{2l}
-        \sum\limits_{i'=0}^{2l'}
-        c_{m,\mu,i}^{l}
-        c_{m',{\mu'},i'}^{l'}
-        \,
-        {({Q^{l,l'}_\phi})_{}}_{i,i'}(\pmb{\theta}_{\phi})
-    $}
-    \quad,
-\end{align}
-%
-where, similarly to before,
-%
-\begin{proof}{test_bigEphi}
-    {({Q^{l,l'}_\phi})_{}}_{i,i'}(\pmb{\theta}_{\phi})
-    & \equiv
-    \frac{1}{2}
-    \int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}
-    %
-    \mathrm{sgn}(\sin\phi)^{i+i'}
-    \big| \sin\phi \big|
-    (\cos\phi)^{\alpha - 1}
-    (1 - \cos\phi)^{l + l' + \beta - \frac{i+i'}{2} - 1}
-    (1 + \cos\phi)^\frac{i+i'}{2}
-    %
-    \,
-    \mathrm{d}\phi
-    \nonumber \\[0.5em]
-    %
-    & =
-    %
-    \begin{cases}
-        \displaystyle\int_{0}^{1}
-        %
-        x^{\alpha - 1}
-        (1 - x)^{l + l' + \beta - \frac{i+i'}{2} - 1 }
-        (1 + x)^\frac{i+i'}{2}
-        %
-        \,
-        \mathrm{d}x
-        %
-         & i+i' \,\, \mathrm{even}
-        %
-        \\
-        %
-        0
-        %
-         & i+i' \,\, \mathrm{odd} \quad.
-    \end{cases}
-\end{proof}
-%
-We may again express this integral in closed form:
-%
-\begin{proof}{test_bigEphi}
-    %
-    \label{eq:Qlphiij}
-    {({Q^{l,l'}_\phi})_{}}_{i,i'}(\pmb{\theta}_{\phi})
-    & =
-    \resizebox{.75\hsize}{!}{$
-            \begin{cases}
-                \dfrac{\Gamma(\alpha)\Gamma(l + l' + \beta - \frac{i+i'}{2})}{\Gamma(l + l' + \alpha + \beta - \frac{i+i'}{2})}
-                \,
-                _2F_1\Big(
-                -\frac{i+i'}{2},
-                \alpha;
-                l + l' + \alpha + \beta - \frac{i+i'}{2};
-                -1
-                \Big)
-                %
-                 & i+i' \,\, \mathrm{even}
-                %
-                \\
-                %
-                0
-                %
-                 & i+i' \,\, \mathrm{odd} \quad.
-            \end{cases}
-        $}
-\end{proof}
-%
-As before, this integral may be evaluated recursively to efficiently compute all
-of the terms in $\mathbf{E}_\phi$.
-
-\subsubsection{Variable transforms}
-\label{sec:lat-transform}
-
-As we mentioned earlier,
-although we adopt $a$ and $b$ as the actual inputs to our Gaussian process
-(i.e., as elements of the input vector $\pmb{\theta}_\bullet$), these
-parameters are not themselves ``interpretable'', as they do not directly
-correspond to physical
-quantities of interest such as the mean and standard deviation of the spot
-latitude distribution. However, over most of the parameter space in $a$ and $b$,
-the spot latitude distribution is well approximated
+\end{align*}
+%
+Over most of the parameter space in $a$ and $b$,
+the spot latitude distribution defined above is well approximated
 by a bimodal Gaussian. In particular, there exists
@@
     \right)
+    \quad.
 \end{proof}
-%
-where $\alpha$ and $\beta$ may be computed from $a$ and $b$ via
-Equation~(\ref{eq:beta2gauss}).
 %
@@
 \end{proof}
-
-These transformations allow one to easily transform constraints on $a$ and
-$b$ to constraints on the mode and approximate variance of the distribution
-of spot latitudes. Importantly, they also allow us to compute the Jacobian
-of the transform, given by
-%
-\begin{proof}{test_lat_jac}
-    J &=
-    \frac{\partial{\mu_\phi}}{\partial{a}}
-    \frac{\partial{\sigma_\phi}}{\partial{b}} -
-    \frac{\partial{\mu_\phi}}{\partial{b}}
-    \frac{\partial{\sigma_\phi}}{\partial{a}}
-    %
-    \\
-    %
-    &=\resizebox{.925\hsize}{!}{$
-            \frac{
-                \left(K_{10} - K_{00}\right) \left(K_{11} - K_{10}\right)
-                \alpha \beta \left(1 + \cos\mu_\phi\right)^3
-                \sin\left(2\mu_\phi\right)^3
-            }{
-                \sigma_\phi
-                \left(
-                2 \alpha + \beta - 3 +
-                \left(2 \alpha + \beta - 1\right) \cos\mu_\phi
-                \right)
-                \left(
-                2 \left( \alpha + \beta - 1\right) +
-                3 (\beta - 1) \cos\mu_\phi
-                - 2 \left( \alpha - \beta - 1 \right) \cos\left( 2\mu_\phi \right)
-                + \left( \beta - 1\right) \cos\left( 3\mu_\phi \right)
-                \right)^2
-            }
-        $}
-    \nonumber
-\end{proof}
-%
-While one could sample in $\mu_\phi$ and $\sigma_\phi$ directly, we instead
-recommend sampling in the quantities $a$ and $b$ under a uniform prior on
-$(0, 1)$ for both of them, then correcting for the implied prior on
-$\mu_\phi$ and $\sigma_\phi$ by adding the log of the absolute value of $J$ to the log
-likelihood. This corrects for the \emph{ad hoc}
-prior on the latitude parameters introduced by our particular choice of
-parametrization, enforcing instead a uniform prior on the quantities $\mu_\phi$
-and $\sigma_\phi$.
 
@@
 distribution in $\phi$; note the excellent agreement.
-
+%
 Thus, in addition to having closed-form moments, the Beta distribution is
@@
 it affects the distribution of spots on the surface of a star.
+
+In the following sections we derive expressions
+for the moments of the distribution in terms
+of $\alpha$ and $\beta$, as this is somewhat more convenient; these
+can easily be transformed into expressions involving either $\mu_\phi$ and $\sigma_\phi$
+or $a$ and $b$ via the equations above. The former parametrization is convenient
+when these properties are known and can be fixed; i.e., when using the GP as
+a restrictive prior for the light curve of a star whose spot distribution is
+already understood.
+However, for the purposes of posterior inference---that is, when trying to
+constrain the hyperparameters of the GP---we recommend sampling in the
+shape parameters $a$ and $b$, since their domains are trivial,
+with uncorrelated boundaries. Posterior constraints on these quantities may
+easily be transformed into constraints on $\mu_\phi$ and $\sigma_\phi$
+via the equations above. Note, importantly, that this
+requires us to explicitly add a Jacobian term to the likelihood
+to account for the prior implied by sampling uniformly in $a$ and $b$
+in the range $(0, 1)$.
+The Jacobian is given by
+%
+\begin{proof}{test_lat_jac}
+    J &=
+    \frac{\partial{\mu_\phi}}{\partial{a}}
+    \frac{\partial{\sigma_\phi}}{\partial{b}} -
+    \frac{\partial{\mu_\phi}}{\partial{b}}
+    \frac{\partial{\sigma_\phi}}{\partial{a}}
+    %
+    \\
+    %
+    &=\resizebox{.925\hsize}{!}{$
+            \frac{
+                \left(K_{10} - K_{00}\right) \left(K_{11} - K_{10}\right)
+                \alpha \beta \left(1 + \cos\mu_\phi\right)^3
+                \sin\left(2\mu_\phi\right)^3
+            }{
+                \sigma_\phi
+                \left(
+                2 \alpha + \beta - 3 +
+                \left(2 \alpha + \beta - 1\right) \cos\mu_\phi
+                \right)
+                \left(
+                2 \left( \alpha + \beta - 1\right) +
+                3 (\beta - 1) \cos\mu_\phi
+                - 2 \left( \alpha - \beta - 1 \right) \cos\left( 2\mu_\phi \right)
+                + \left( \beta - 1\right) \cos\left( 3\mu_\phi \right)
+                \right)^2
+            }
+        $}
+    \nonumber
+\end{proof}
+%
+Adding the log of the absolute value of $J$ to the log likelihood
+corrects for the \emph{ad hoc}
+prior on the latitude parameters introduced by our particular choice of
+parametrization, enforcing instead a uniform prior on the quantities $\mu_\phi$
+and $\sigma_\phi$.
+
+\subsubsection{First moment}
+%
+Since the Wigner matrices are block diagonal, we may evaluate the moments of the
+distribution one spherical harmonic degree at a time. To that end, let us
+write the first moment integral as
+%
+\begin{proof}{test_ephi}
+    \mathbf{e}_\phi
+    & \equiv
+    \int
+    \mathbf{R}_{\hat{\mathbf{x}}}(\bbphi) \,
+    \mathbf{e}_r \,
+    p(\bbphi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\bbphi
+    \nonumber
+    \\
+    %
+    & =
+    %
+    \left(
+    \mathbf{e}_\phi^0
+    \,\,\,
+    \mathbf{e}_\phi^1
+    \,\,\,
+    \mathbf{e}_\phi^2
+    \,\,\,
+    \cdots
+    \,\,\,
+    \mathbf{e}_\phi^{l_{\mathrm{max}}}
+    \right)^\top
+    \quad,
+\end{proof}
+%
+where
+%
+\begin{proof}{test_ephi}
+    \mathbf{e}_\phi^l
+    & =
+    \int
+    \mathbf{R}^l_{\hat{\mathbf{x}}}(\bbphi) \,
+    \mathbf{e}^l_r \,
+    p(\bbphi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\bbphi
+    \nonumber \\
+    %
+    & =
+    %
+    {\mathbf{U}^l}^{-1}
+    \mathbf{p}^l_\phi
+    \quad,
+\end{proof}
+%
+and we define
+%
+\begin{align}
+    \label{eq:plphi}
+    \mathbf{p}^l_\phi
+     & \equiv
+    \int
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(\bbphi) \,
+    \bar{\mathbf{e}}^l_r \,
+    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\bbphi
+    \\
+    \bar{\mathbf{e}}^l_r
+     & \equiv
+    \mathbf{U}^l
+    \mathbf{e}^l_r
+    \quad.
+\end{align}
+%
+The integral $\mathbf{p}_\phi^l$ defined above has a closed-form solution.
+To show this, we write the terms of $\mathbf{p}^l_\phi$ as
+%
+\begin{align}
+    {({p^l_\phi})_{}}_m
+     & =
+    \int
+    \sum\limits_{\mu=-l}^l
+    {({D^l_{\hat{\mathbf{x}}}})_{}}_{m,\mu}(\bbphi) \,
+    {({\bar{e}}^l_r)_{}}_{\mu} \,
+    p(\bbphi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\bbphi
+    %
+    \nonumber \\[0.5em]
+    %
+     & =
+    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+    \sum\limits_{\mu=-l}^l
+    {({\bar{e}}^l_r)_{}}_{\mu}
+    \exp\left[{\frac{\imag \pi}{2}(m - \mu)}\right]
+    \sum\limits_{i=0}^{2l} c_{m,\mu,i}^{l}
+    \,
+    {({q^l_\phi})_{}}_i(\pmb{\theta}_{\phi})
+    \quad,
+\end{align}
+%
+where
+%
+\begin{proof}{test_qlphii}
+    {({q^l_\phi})_{}}_i(\pmb{\theta}_{\phi})
+    & \equiv
+    \frac{1}{2}
+    \int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}
+    %
+    \mathrm{sgn}(\sin\bbphi)^{i}
+    \big| \sin\bbphi \big|
+    (\cos\bbphi)^{\alpha - 1}
+    (1 - \cos\bbphi)^{l + \beta - \frac{i}{2} - 1}
+    (1 + \cos\bbphi)^\frac{i}{2}
+    %
+    \,
+    \mathrm{d}\bbphi
+    \nonumber \\[0.5em]
+    %
+    & =
+    %
+    \begin{cases}
+        \displaystyle\int_{0}^{1}
+        %
+        \mathbb{x}^{\alpha - 1}
+        (1 - \mathbb{x})^{l + \beta - \frac{i}{2} - 1 }
+        (1 + \mathbb{x})^\frac{i}{2}
+        %
+        \,
+        \mathrm{d}\mathbb{x}
+        %
+         & i \,\, \mathrm{even}
+        %
+        \\
+        %
+        0
+        %
+         & i \,\, \mathrm{odd} \quad,
+    \end{cases}
+\end{proof}
+%
+and in the last line we made use of the transformation $\mathbb{x} = \cos\bbphi$.
+%
+The integral in the expression above has a closed-form solution in terms
+of the hypergeometric function $_2F_1$:
+%
+\begin{proof}{test_qlphii}
+    %
+    \label{eq:qlphii}
+    {({q^l_\phi})_{}}_i(\pmb{\theta}_{\phi})
+    & =
+    \resizebox{.72\hsize}{!}{$
+            \begin{cases}
+                \dfrac{\Gamma(\alpha)\Gamma(l + \beta - \frac{i}{2})}{\Gamma(l + \alpha + \beta - \frac{i}{2})}
+                \,
+                _2F_1\Big(
+                -\frac{i}{2},
+                \alpha;
+                l + \alpha + \beta - \frac{i}{2};
+                -1
+                \Big)
+                %
+                 & \hspace{-3em} i \,\, \mathrm{even}
+                %
+                \\
+                %
+                0
+                %
+                 & \hspace{-3em} i \,\, \mathrm{odd} \quad.
+            \end{cases}
+        $}
+\end{proof}
+%
+In order to compute the integral $\mathbf{e}_\phi$
+(Equation~\ref{eq:e2}) we must evaluate Equation~(\ref{eq:qlphii})
+for all $0 \leq l \leq l_{\mathrm{max}}$, $0 \leq i \leq 2l$,
+which can be done efficiently via upward recursion relations for both
+the gamma function and the hypergeometric function.
+
+\subsubsection{Second moment}
+%
+Similarly as before, let us write the second moment integral as
+%
+\begin{proof}{test_bigEphi}
+    \label{eq:Ephi}
+    \mathbf{E}_\phi
+    & \equiv
+    \int
+    \mathbf{R}_{\hat{\mathbf{x}}}(\bbphi) \,
+    \mathbf{E}_r \,
+    \mathbf{R}_{\hat{\mathbf{x}}}^\top(\bbphi) \,
+    p(\bbphi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\bbphi
+    \nonumber
+    \\
+    %
+    & =
+    %
+    \setstackgap{L}{1.25\baselineskip}
+    \fixTABwidth{T}
+    \parenMatrixstack{
+    \mathbf{E}_\phi^{0,0} & \mathbf{E}_\phi^{0,1} & \mathbf{E}_\phi^{0,2} & \cdots & \mathbf{E}_\phi^{0,l_{\mathrm{max}}} \\
+    \mathbf{E}_\phi^{1,0} & \mathbf{E}_\phi^{1,1} & \mathbf{E}_\phi^{1,2} & \cdots & \mathbf{E}_\phi^{1,l_{\mathrm{max}}} \\
+    \mathbf{E}_\phi^{2,0} & \mathbf{E}_\phi^{2,1} & \mathbf{E}_\phi^{2,2} & \cdots & \mathbf{E}_\phi^{2,l_{\mathrm{max}}} \\
+    \vdots & \vdots & \vdots & \ddots & \vdots \\
+    \mathbf{E}_\phi^{l_{\mathrm{max}},0} & \mathbf{E}_\phi^{l_{\mathrm{max}},1} & \mathbf{E}_\phi^{l_{\mathrm{max}},2} & \cdots & \mathbf{E}_\phi^{l_{\mathrm{max}},l_{\mathrm{max}}}
+    }
+    \quad,
+\end{proof}
+%
+where
+%
+\begin{proof}{test_bigEphi}
+    \mathbf{E}_\phi^{l,l'}
+    & =
+    \int
+    \mathbf{R}^l_{\hat{\mathbf{x}}}(\bbphi) \,
+    \mathbf{E}^{l,l'}_r \,
+    {\mathbf{R}^{l'}_{\hat{\mathbf{x}}}}^\top(\bbphi) \,
+    p(\bbphi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\bbphi
+    \nonumber \\
+    %
+    & =
+    %
+    {\mathbf{U}^l}^{-1}
+    \mathbf{P}^{l,l'}_\phi
+    {{\mathbf{U}^{l'}}^{-1}}^\top
+\end{proof}
+%
+and we define
+%
+\begin{align}
+    \label{eq:Plphi}
+    \mathbf{P}^{l,l'}_\phi
+     & \equiv
+    \int
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(\bbphi) \,
+    \bar{\mathbf{E}}^{l,l'}_r \,
+    {\mathbf{D}^{l'}_{\hat{\mathbf{x}}}}^\top(\bbphi) \,
+    p(\bbphi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\bbphi
+    \\
+    \bar{\mathbf{E}}^{l,l'}_r
+     & \equiv
+    \mathbf{U}^l
+    \mathbf{E}^{l,l'}_r
+    {\mathbf{U}^{l'}}^\top
+    \quad.
+\end{align}
+%
+As before, we may express the solution to the integral $\mathbf{P}^{l,l'}_\phi$ in
+closed form. Let us write the terms of $\mathbf{P}^{l,l'}_\phi$ as
+%
+\begin{align}
+    {({P^{l,l'}_\phi})_{}}_{m,m'}
+     & =
+    \int
+    \sum\limits_{\mu=-l}^l
+    \sum\limits_{{\mu'}=-l'}^{l'}
+    {({D^l_{\hat{\mathbf{x}}}})_{}}_{m,\mu}(\bbphi) \,
+    {(\bar{E}^{l,l'}_r)_{}}_{\mu,{\mu'}} \,
+    {({D^{l'}_{\hat{\mathbf{x}}}})_{}}_{m',{\mu'}}(\bbphi) \,
+    p(\bbphi \, \big| \, \pmb{\theta}_{\phi}) \,
+    \mathrm{d}\bbphi
+    %
+    \\[0.5em]
+    %
+     & =
+    \resizebox{.75\hsize}{!}{$
+            \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
+            \sum\limits_{\mu=-l}^l
+            \sum\limits_{{\mu'}=-l'}^{l'}
+            {(\bar{E}^{l,l'}_r)_{}}_{\mu,{\mu'}} \,
+            \exp\left[{\frac{\imag \pi}{2}(m - \mu + m' - {\mu'})}\right]
+            \sum\limits_{i=0}^{2l}
+            \sum\limits_{i'=0}^{2l'}
+            c_{m,\mu,i}^{l}
+            c_{m',{\mu'},i'}^{l'}
+            \,
+            {({Q^{l,l'}_\phi})_{}}_{i,i'}(\pmb{\theta}_{\phi})
+        $}
+    \quad,
+    \nonumber
+\end{align}
+%
+where, similarly to before,
+%
+\begin{proof}{test_bigEphi}
+    {({Q^{l,l'}_\phi})_{}}_{i,i'}(\pmb{\theta}_{\phi})
+    & \equiv
+    \resizebox{.8\hsize}{!}{$
+            \frac{1}{2}
+            \int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}
+            %
+            \mathrm{sgn}(\sin\bbphi)^{i+i'}
+            \big| \sin\bbphi \big|
+            (\cos\bbphi)^{\alpha - 1}
+            (1 - \cos\bbphi)^{l + l' + \beta - \frac{i+i'}{2} - 1}
+            (1 + \cos\bbphi)^\frac{i+i'}{2}
+            %
+            \,
+            \mathrm{d}\bbphi
+        $}
+    \nonumber \\[0.5em]
+    %
+    & =
+    %
+    \begin{cases}
+        \displaystyle\int_{0}^{1}
+        %
+        \mathbb{x}^{\alpha - 1}
+        (1 - \mathbb{x})^{l + l' + \beta - \frac{i+i'}{2} - 1 }
+        (1 + \mathbb{x})^\frac{i+i'}{2}
+        %
+        \,
+        \mathrm{d}\mathbb{x}
+        %
+         & \hspace{-2em} i+i' \,\, \mathrm{even}
+        %
+        \\
+        %
+        0
+        %
+         & \hspace{-2em} i+i' \,\, \mathrm{odd} \quad.
+    \end{cases}
+\end{proof}
+%
+We may again express this integral in closed form:
+%
+\begin{proof}{test_bigEphi}
+    %
+    \label{eq:Qlphiij}
+    {({Q^{l,l'}_\phi})_{}}_{i,i'}(\pmb{\theta}_{\phi})
+    & =
+    \resizebox{.68\hsize}{!}{$
+            \begin{cases}
+                \dfrac{\Gamma(\alpha)\Gamma(l {+} l' {+} \beta {-} \frac{i+i'}{2})}{\Gamma(l {+} l' {+} \alpha {+} \beta {-} \frac{i+i'}{2})}
+                \,
+                _2F_1\Big(
+                -\frac{i+i'}{2},
+                \alpha;
+                l {+} l' {+} \alpha {+} \beta {-} \frac{i+i'}{2};
+                -1
+                \Big)
+                %
+                 & \hspace{-3em} i+i' \,\, \mathrm{even}
+                %
+                \\
+                %
+                0
+                %
+                 & \hspace{-3em} i+i' \,\, \mathrm{odd} \quad.
+            \end{cases}
+        $}
+\end{proof}
+%
+As before, this integral may be evaluated recursively to efficiently compute all
+of the terms in $\mathbf{E}_\phi$.
 
@@
 The longitude integrals (Equations~\ref{eq:e3} and \ref{eq:E3}) involve
-rotations by an angle $\lambda$ about $\hat{\mathbf{y}}$, which
+rotations by an angle $\bblambda$ about $\hat{\mathbf{y}}$, which
 may be accomplished by choosing
-Euler angles $\alpha = 0$, $\beta = \lambda$, and
-$\gamma = 0$, such that
-%
-\begin{align}
-    \mathbf{R}^l_{\hat{\mathbf{y}}}(\lambda)
-     & =
-    {\mathbf{U}^l}^{-1} \mathbf{D}^l_{\hat{\mathbf{y}}}(\lambda) \mathbf{U}^l
+Euler angles $\upalpha = 0$, $\upbeta = \bblambda$, and
+$\upgamma = 0$, such that
+%
+\begin{align}
+    \mathbf{R}^l_{\hat{\mathbf{y}}}(\bblambda)
+     & =
+    {\mathbf{U}^l}^{-1} \mathbf{D}^l_{\hat{\mathbf{y}}}(\bblambda) \mathbf{U}^l
 \end{align}
@@
 \begin{align}
-    \mathbf{D}^l_{\hat{\mathbf{y}}}(\lambda)
-     & =
-    \mathbf{D}^l\left(0, \lambda, 0\right)
+    \mathbf{D}^l_{\hat{\mathbf{y}}}(\bblambda)
+     & =
+    \mathbf{D}^l\left(0, \bblambda, 0\right)
     \quad.
@@
 of stars to be (on average) isotropic, we will place a uniform prior on
-$\lambda \in [-\pi, \pi)$:
-%
-\begin{align}
-    p(\lambda \big| \pmb{\theta}_\lambda)
+$\bblambda \in [-\pi, \pi)$:
+%
+\begin{align}
+    p(\bblambda \big| \pmb{\theta}_\lambda)
      & =
     \begin{cases}
-        \frac{1}{2\pi} & -\pi \leq \lambda < \pi
+        \frac{1}{2\pi} & -\pi \leq \bblambda < \pi
         \\
@@
     \int
-    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{R}_{\hat{\mathbf{y}}}(\bblambda) \,
     \mathbf{e}_\phi \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
+    p(\bblambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\bblambda
     \nonumber
@@
     \int
-    \mathbf{R}^l_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{R}^l_{\hat{\mathbf{y}}}(\bblambda) \,
     \mathbf{e}^l_\phi \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
+    p(\bblambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\bblambda
     \nonumber \\
@@
     \int
-    \mathbf{D}^l_{\hat{\mathbf{x}}}(\lambda) \,
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(\bblambda) \,
     \bar{\mathbf{e}}^l_\phi \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
+    p(\bblambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\bblambda
     \\
@@
     \sum\limits_{\mu=-l}^l
-    {({D^l_{\hat{\mathbf{y}}}})_{}}_{m,\mu}(\lambda) \,
+    {({D^l_{\hat{\mathbf{y}}}})_{}}_{m,\mu}(\bblambda) \,
     {({\bar{e}^l_\phi})_{}}_{\mu} \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
+    p(\bblambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\bblambda
     %
@@
     %
-    \mathrm{sgn}(\sin\lambda)^{i}
-    (1 - \cos\lambda)^{l - \frac{i}{2}}
-    (1 + \cos\lambda)^\frac{i}{2}
+    \mathrm{sgn}(\sin\bblambda)^{i}
+    (1 - \cos\bblambda)^{l - \frac{i}{2}}
+    (1 + \cos\bblambda)^\frac{i}{2}
     %
     \,
-    \mathrm{d}\lambda
+    \mathrm{d}\bblambda
     \nonumber \\[0.5em]
@@
     \int
-    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{R}_{\hat{\mathbf{y}}}(\bblambda) \,
     \mathbf{E}_\phi \,
-    \mathbf{R}_{\hat{\mathbf{y}}}^\top(\lambda) \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
+    \mathbf{R}_{\hat{\mathbf{y}}}^\top(\bblambda) \,
+    p(\bblambda \, \big| \, \pmb{\theta}_{\bblambda}) \,
+    \mathrm{d}\bblambda
     \nonumber
@@
     \int
-    \mathbf{R}^l_{\hat{\mathbf{y}}}(\lambda) \,
+    \mathbf{R}^l_{\hat{\mathbf{y}}}(\bblambda) \,
     \mathbf{E}^{l,l'}_\phi \,
-    {\mathbf{R}^{l'}_{\hat{\mathbf{y}}}}^\top(\lambda) \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
+    {\mathbf{R}^{l'}_{\hat{\mathbf{y}}}}^\top(\bblambda) \,
+    p(\bblambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\bblambda
     \nonumber \\
@@
     \int
-    \mathbf{D}^l_{\hat{\mathbf{x}}}(\lambda) \,
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(\bblambda) \,
     \bar{\mathbf{E}}^{l,l'}_\phi \,
-    {\mathbf{D}^{l'}_{\hat{\mathbf{x}}}}^\top(\lambda) \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
+    {\mathbf{D}^{l'}_{\hat{\mathbf{x}}}}^\top(\bblambda) \,
+    p(\bblambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\bblambda
     \\
@@
     \sum\limits_{{\mu'}=-l'}^{l'}
-    ({D^l_{\hat{\mathbf{x}}}})_{m,\mu}(\lambda) \,
+    ({D^l_{\hat{\mathbf{x}}}})_{m,\mu}(\bblambda) \,
     (\bar{E}^{l,l'}_\phi)_{\mu,{\mu'}} \,
-    ({D^{l'}_{\hat{\mathbf{x}}}})_{m',{\mu'}}(\lambda) \,
-    p(\lambda \, \big| \, \pmb{\theta}_{\lambda}) \,
-    \mathrm{d}\lambda
+    ({D^{l'}_{\hat{\mathbf{x}}}})_{m',{\mu'}}(\bblambda) \,
+    p(\bblambda \, \big| \, \pmb{\theta}_{\lambda}) \,
+    \mathrm{d}\bblambda
     %
@@
     %
-    \mathrm{sgn}(\sin\lambda)^{i+i'}
-    (1 - \cos\lambda)^{l + l' - \frac{i+i'}{2}}
-    (1 + \cos\lambda)^\frac{i+i'}{2}
+    \mathrm{sgn}(\sin\bblambda)^{i+i'}
+    (1 - \cos\bblambda)^{l + l' - \frac{i+i'}{2}}
+    (1 + \cos\bblambda)^\frac{i+i'}{2}
     %
     \,
-    \mathrm{d}\lambda
+    \mathrm{d}\bblambda
     \nonumber \\[0.5em]
@@
 The final integrals we must take in our computation of
-$\mathrm{E} \big[ \mathbf{y} \, \big| \, \pmb{\theta}_\bullet \big]$
-and $\mathrm{E} \big[ \mathbf{y} \mathbf{y}^\top \, \big| \, \pmb{\theta}_\bullet \big]$
+$\mathrm{E} \big[ \mathbbb{y} \, \big| \, \pmb{\theta}_\bullet \big]$
+and $\mathrm{E} \big[ \mathbbb{y} \mathbbb{y}^\top \, \big| \, \pmb{\theta}_\bullet \big]$
 are the integrals over
-the spot contrast distribution, $\mathbf{e}_\xi$ and $\mathbf{E}_\xi$.
+the spot contrast distribution, $\mathbf{e}_c$ and $\mathbf{E}_c$.
 These are by far the easiest, since the spot contrast is a scalar
@@
 \begin{align}
-    \mathbf{e}_\xi
+    \mathbf{e}_c
      & \equiv
@@
     \xi \,
-    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
+    p(\xi \, \big| \, \pmb{\theta}_{c}) \,
     \mathrm{d}\xi
     \\
-    \mathbf{E}_\xi
+    \mathbf{E}_c
      & \equiv
@@
 \begin{align}
-    p(\xi \, \big| \, \pmb{\theta}_{\xi}) = \delta(\xi - c)
+    p(\xi \, \big| \, \pmb{\theta}_{c}) = \delta(\xi - c)
 \end{align}
@@
 \begin{align}
-    \mathbf{e}_\xi & = c \, \mathbf{e}_\lambda
+    \mathbf{e}_c & = c \, \mathbf{e}_\lambda
     \\
-    \mathbf{E}_\xi & = c^2 \mathbf{E}_\lambda
+    \mathbf{E}_c & = c^2 \mathbf{E}_\lambda
     \quad.
@@
 observer (see Equation~\ref{eq:akT}) involves
-rotations by an angle $-I$ about $\hat{\mathbf{x}}$, which
+rotations by an angle $-\mathbb{I}$ about $\hat{\mathbf{x}}$, which
 may be accomplished by choosing
-Euler angles $\alpha = \nicefrac{\pi}{2}$, $\beta = -I$, and
-$\gamma = -\nicefrac{\pi}{2}$, such that
-%
-\begin{align}
-    \mathbf{R}^l_{\hat{\mathbf{x}}}\left(-I\right)
-     & =
-    {\mathbf{U}^l}^{-1} \mathbf{D}^l_{\hat{\mathbf{x}}}\left(-I\right) \mathbf{U}^l
+Euler angles $\upalpha = \nicefrac{\pi}{2}$, $\upbeta = -\mathbb{I}$, and
+$\upgamma = -\nicefrac{\pi}{2}$, such that
+%
+\begin{align}
+    \mathbf{R}^l_{\hat{\mathbf{x}}}\left(-\mathbb{I}\right)
+     & =
+    {\mathbf{U}^l}^{-1} \mathbf{D}^l_{\hat{\mathbf{x}}}\left(-\mathbb{I}\right) \mathbf{U}^l
 \end{align}
@@
 \begin{align}
-    \mathbf{D}^l_{\hat{\mathbf{x}}}\left(-I\right)
-     & =
-    \mathbf{D}^l\left(\frac{\pi}{2}, -I, -\frac{\pi}{2}\right)
+    \mathbf{D}^l_{\hat{\mathbf{x}}}\left(-\mathbb{I}\right)
+     & =
+    \mathbf{D}^l\left(\frac{\pi}{2}, -\mathbb{I}, -\frac{\pi}{2}\right)
     \quad.
@@
 \begin{align}
-    p(I) = \sin I
-\end{align}
-%
-for $I \in [0, \frac{\pi}{2}]$.
+    p(\mathbb{I}) = \sin \mathbb{I}
+\end{align}
+%
+for $\mathbb{I} \in [0, \frac{\pi}{2}]$.
 
@@
     \int
-    \pmb{\mathcal{A}}(I, P, \mathbf{u}) \,
+    \pmb{\mathcal{A}}(\mathbb{I}, P, \mathbf{u}) \,
     \mathbf{e}_y \,
-    p(I) \,
-    \mathrm{d}I
+    p(\mathbb{I}) \,
+    \mathrm{d}\mathbb{I}
 \end{align}
@@
      & \equiv
-    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
+    \mathrm{E} \Big[ \mathbbb{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
 \end{align}
@@
     \int
-    \mathbf{a}_k^\top
+    \mathbf{a}_k^\top(\mathbb{I})
     \mathbf{e}_y \,
-    p(I) \,
-    \mathrm{d}I
+    p(\mathbb{I}) \,
+    \mathrm{d}\mathbb{I}
     \nonumber \\
@@
     \int
-    \mathbf{R}_{\hat{\mathbf{x}}}\left(-I\right) \,
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(-\mathbb{I}\right) \,
     \,
     \left(\mathbf{e}_{y'}\right)_k \,
-    p(I) \,
-    \mathrm{d}I
+    p(\mathbb{I}) \,
+    \mathrm{d}\mathbb{I}
     \quad,
@@
 %
-as the expectation of $\mathbf{y}$ in the polar frame at time $t = t_k$.%
+as the expectation of $\mathbbb{y}$ in the polar frame at time $t = t_k$.%
 \footnote{In the presence of limb darkening, we must include the limb
@@
 rotational axis of the star) therefore cannot change the expectation
-of $\mathbf{y}$, so
+of $\mathbbb{y}$, so
 %
@@
     \int
-    \mathbf{R}_{\hat{\mathbf{x}}}\left(-I\right) \,
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(-\mathbb{I}\right) \,
     \,
     \mathbf{e}_{y'} \,
-    p(I) \,
-    \mathrm{d}I
+    p(\mathbb{I}) \,
+    \mathrm{d}\mathbb{I}
     \nonumber \\
@@
 %
-is the expectation of $\mathbf{y}$ in the observer's frame,
+is the expectation of $\mathbbb{y}$ in the observer's frame,
 and as before we explicitly separate it out by spherical harmonic degree.
@@
     \int
-    \mathbf{R}^l_{\hat{\mathbf{x}}}(-I) \,
+    \mathbf{R}^l_{\hat{\mathbf{x}}}(-\mathbb{I}) \,
     \mathbf{e}^l_{y'} \,
-    p(I) \,
-    \mathrm{d}I
+    p(\mathbb{I}) \,
+    \mathrm{d}\mathbb{I}
     \nonumber \\
@@
     \int
-    \mathbf{D}^l_{\hat{\mathbf{x}}}(-I) \,
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(-\mathbb{I}) \,
     \bar{\mathbf{e}}^l_{y'} \,
-    p(I) \,
-    \mathrm{d}I
+    p(\mathbb{I}) \,
+    \mathrm{d}\mathbb{I}
     \\
@@
     \sum\limits_{\mu=-l}^l
-    {({D^l_{\hat{\mathbf{x}}}})_{}}_{m,\mu}(-I) \,
+    {({D^l_{\hat{\mathbf{x}}}})_{}}_{m,\mu}(-\mathbb{I}) \,
     {({\bar{e}}^l_{y'})_{}}_{\mu} \,
-    p(I) \,
-    \mathrm{d}I
+    p(\mathbb{I}) \,
+    \mathrm{d}\mathbb{I}
     %
@@
     {({\bar{e}}^l_{y'})_{}}_{\mu}
-    \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu)}
+    \exp\left[{\frac{\imag \pi}{2}(m - \mu)}\right]
     \sum\limits_{i=0}^{2l} c_{m,\mu,i}^{l}
@@
     (-1)^{i}
-    (1 - \cos I)^{\frac{2l - i}{2}}
-    (1 + \cos I)^\frac{i}{2}
-    \sin I
+    (1 - \cos \mathbb{I})^{\frac{2l - i}{2}}
+    (1 + \cos \mathbb{I})^\frac{i}{2}
+    \sin \mathbb{I}
     %
     \,
-    \mathrm{d}I
+    \mathrm{d}\mathbb{I}
     \nonumber \\[0.5em]
@@
     %
-    (1 - x)^{\frac{2l - i}{2}}
-    (1 + x)^\frac{i}{2}
+    (1 - \mathbb{x})^{\frac{2l - i}{2}}
+    (1 + \mathbb{x})^\frac{i}{2}
     \,
-    \mathrm{d}x
+    \mathrm{d}\mathbb{x}
     \nonumber \\[0.5em]
@@
     \int
-    \pmb{\mathcal{A}}(I, P, \mathbf{u}) \,
+    \pmb{\mathcal{A}}(\mathbb{I}, P, \mathbf{u}) \,
     \mathbf{E}_y \,
     \pmb{\mathcal{A}}^\top(I, P, \mathbf{u}) \,
-    p(I) \,
-    \mathrm{d}I
+    p(\mathbb{I}) \,
+    \mathrm{d}\mathbb{I}
 \end{align}
@@
      & \equiv
-    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
+    \mathrm{E} \Big[ \mathbbb{y} \, \mathbbb{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
 \end{align}
@@
     \int
-    \mathbf{a}_k^\top
+    \mathbf{a}_k^\top(\mathbb{I})
     \mathbf{E}_y \,
-    \mathbf{a}_{k'}
-    p(I) \,
-    \mathrm{d}I
+    \mathbf{a}_{k'}(\mathbb{I})
+    p(\mathbb{I}) \,
+    \mathrm{d}\mathbb{I}
     \nonumber \\
@@
     \int
-    \mathbf{R}_{\hat{\mathbf{x}}}\left(-I\right) \,
+    \mathbf{R}_{\hat{\mathbf{x}}}\left(-\mathbb{I}\right) \,
     \left(\mathbf{E}_{y'}\right)_{k,k'} \,
-    \mathbf{R}_{\hat{\mathbf{x}}}^\top\left(-I\right) \,
-    p(I) \,
-    \mathrm{d}I
-\end{align}
-%
-is the expectation of $\mathbf{y}\,\mathbf{y}^\top$ in the
+    \mathbf{R}_{\hat{\mathbf{x}}}^\top\left(-\mathbb{I}\right) \,
+    p(\mathbb{I}) \,
+    \mathrm{d}\mathbb{I}
+\end{align}
+%
+is the expectation of $\mathbbb{y}\,\mathbbb{y}^\top$ in the
 observer's frame at times $t = t_k$ and $t' = t_{k'}$
@@
 %
-is the expectation of $\mathbf{y}\,\mathbf{y}^\top$ in the
+is the expectation of $\mathbbb{y}\,\mathbbb{y}^\top$ in the
 polar frame at times $t = t_k$ and $t' = t_{k'}$.
@@
 \begin{proof}{test_bigEI}
-    \resizebox{.85\hsize}{!}{$
+    \resizebox{.83\hsize}{!}{$
             \left(\mathbf{E}_{y''}\right)_{k,k'}
@@
     \int
-    \mathbf{R}^l_{\hat{\mathbf{x}}}(-I) \,
+    \mathbf{R}^l_{\hat{\mathbf{x}}}(-\mathbb{I}) \,
     \left(\mathbf{E}_{y'}^{l,l'}\right)_{k,k'} \,
-    {\mathbf{R}^{l'}_{\hat{\mathbf{x}}}}^\top(-I) \,
-    p(I) \,
-    \mathrm{d}I
+    {\mathbf{R}^{l'}_{\hat{\mathbf{x}}}}^\top(-\mathbb{I}) \,
+    p(\mathbb{I}) \,
+    \mathrm{d}\mathbb{I}
     \nonumber \\
@@
     \int
-    \mathbf{D}^l_{\hat{\mathbf{x}}}(-I) \,
+    \mathbf{D}^l_{\hat{\mathbf{x}}}(-\mathbb{I}) \,
     \left(\bar{\mathbf{E}}^{l,l'}_{y'}\right)_{k,k'} \,
-    {\mathbf{D}^{l'}_{\hat{\mathbf{x}}}}^\top(-I) \,
-    p(I) \,
-    \mathrm{d}I
+    {\mathbf{D}^{l'}_{\hat{\mathbf{x}}}}^\top(-\mathbb{I}) \,
+    p(\mathbb{I}) \,
+    \mathrm{d}\mathbb{I}
     \\
@@
      & =
+    \resizebox{.65\hsize}{!}{$
         \int
@@
         \sum\limits_{{\mu'}=-l}^l
-    {({D^l_{\hat{\mathbf{x}}}})_{}}_{m,\mu}(-I) \,
+        {({D^l_{\hat{\mathbf{x}}}})_{}}_{m,\mu}(-\mathbb{I}) \,
         \left[{\left({\bar{E}^{l,l'}_{y'}}\right)_{}}_{k,k'}\right]_{\mu,\mu'}
-    {({D^{l'}_{\hat{\mathbf{x}}}})_{}}_{{\mu'},m'}(-I) \,
-    p(I) \,
-    \mathrm{d}I
-    %
-    \nonumber \\[0.5em]
+        {({D^{l'}_{\hat{\mathbf{x}}}})_{}}_{{\mu'},m'}(-\mathbb{I}) \,
+        p(\mathbb{I}) \,
+        \mathrm{d}\mathbb{I}
+    $}
+    %
+    \\[0.5em]
     %
@@
             \left[{\left({\bar{E}^{l,l'}_{y'}}\right)_{}}_{k,k'}\right]_{\mu,\mu'}
-        \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu + m' - {\mu'})}
+            \exp\left[{\frac{\imag \pi}{2}(m - \mu + m' - {\mu'})}\right]
             \sum\limits_{i=0}^{2l}
@@
     \quad,
+    \nonumber
 \end{align}
@@
     (-1)^{i+i'}
-    (1 - \cos I)^{l + l' - \frac{i+i'}{2}}
-    (1 + \cos I)^\frac{i+i'}{2}
-    \sin I
+    (1 - \cos \mathbb{I})^{l + l' - \frac{i+i'}{2}}
+    (1 + \cos \mathbb{I})^\frac{i+i'}{2}
+    \sin \mathbb{I}
     %
     \,
-    \mathrm{d}I
+    \mathrm{d}\mathbb{I}
     \nonumber \\[0.5em]
@@
     %
-    (1 - x)^{l + l' - \frac{i+i'}{2}}
-    (1 + x)^\frac{i+i'}{2}
+    (1 - \mathbb{x})^{l + l' - \frac{i+i'}{2}}
+    (1 + \mathbb{x})^\frac{i+i'}{2}
     %
     \,
-    \mathrm{d}x
+    \mathrm{d}\mathbb{x}
     \nonumber \\[0.5em]

### 2020-12-21 tex/ms.tex
more figs

@@
 \begin{align}
-    k_\mathrm{E^2}(\Delta t) = \sigma^2 \exp\left(-\frac{\Delta t}{2\tau}\right)
-    \quad,
-\end{align}
-%
-the Mat\'ern-3/2 kernel,
-%
-\begin{align}
-    k_\mathrm{M^\frac{3}{2}}(\Delta t) = \sigma^2 \left( 1 + \frac{\sqrt{3}\Delta t}{\tau} \right) \exp\left(-\frac{\sqrt{3}\Delta t}{\tau}\right)
-    \quad,
-\end{align}
-%
-the exponential sine squared kernel,
-%
-\begin{align}
-    k_\mathrm{S^2}(\Delta t) = \sigma^2 \exp\left[-\Gamma\sin^2\left( \frac{\pi \Delta t}{P} \right)\right]
-    \quad,
-\end{align}
-%
-the cosine kernel,
-%
-\begin{align}
-    k_\mathrm{C}(\Delta t) = \sigma^2 \cos\left( \frac{2\pi \Delta t}{P} \right)
+    k_\mathrm{E^2}(\Delta t)           & = \sigma^2 \exp\left(-\frac{\Delta t}{2\tau}\right)
+    \quad,                                                                                                                                           \\
+    \intertext{the Mat\'ern-3/2 kernel,}
+    k_\mathrm{M^\frac{3}{2}}(\Delta t) & = \sigma^2 \left( 1 + \frac{\sqrt{3}\Delta t}{\tau} \right) \exp\left(-\frac{\sqrt{3}\Delta t}{\tau}\right)
+    \quad,                                                                                                                                           \\
+    \intertext{the exponential sine squared kernel,}
+    k_\mathrm{S^2}(\Delta t)           & = \sigma^2 \exp\left[-\Gamma\sin^2\left( \frac{\pi \Delta t}{P} \right)\right]
+    \quad,                                                                                                                                           \\
+    \intertext{the cosine kernel,}
+    k_\mathrm{C}(\Delta t)             & = \sigma^2 \cos\left( \frac{2\pi \Delta t}{P} \right)
     \quad,
@@
 where $n$ is the number of star spots, $c$ is their contrast,
-$\mu$ and $\sigma$ are parameters describing the
+$\mu_\phi$ and $\sigma_\phi$ are parameters describing the
 distribution of the spots in latitude, and $r$ is the radius
@@
 \begin{align}
-    N_\theta = 4 M + 5
+    N = 4 M + 5
     \quad,
@@
 For a reasonably sized ensemble of $M=100$
-stars, we would have to sample over $N_\theta = 405$ parameters.
+stars, we would have to sample over $N = 405$ parameters.
 While large, this number is certainly not absurd, especially by modern standards.
@@
 The first point is illustrated in the toy example in
-Figure~\ref{fig:mean_normalization}. As it has been pointed out before \xxx{(?)},
+Figure~\ref{fig:mean_normalization}. As it has been pointed out before \citep{?},
 we will not spend too much time discussing it, other than to say that the
@@
 \footnote{The exception to this is the
-    normalization correction (\S\ref{sec:baseline}), which depends on a rapidly
-    converging series and thus adds negligible overhead to the computation.}
+    normalization correction (\S\ref{sec:baseline}), which depends on a
+    fast-to-evaluate series and thus adds negligible overhead to the computation.}
 %
@@
 }
-The default values of each of the parameters mentioned above are given in
+The default values/distributions of each of the parameters mentioned above are given in
 Table~\ref{tab:synthetic}. Some, like the number of spots, their contrast,
@@
 \label{sec:calibration-inference}
-\xxx{TODO.}
+We use our \Python-based implementation of the GP (\S\ref{sec:implementation})
+to perform inference on the synthetic datasets. For simplicity, we assume we
+know the value of the period $P$, which is fixed at unity for all stars, as well
+as the value of the limb darkening coefficients (fixed at zero for the default
+run). In practice these will not be known exactly; we discuss this further in
+\S\ref{sec:other-marg}. Since we explicitly marginalize over the inclinations
+of individual stars, the only quantities we must constrain are the five
+parameters in the spot parameter vector $\pmb{\theta}_\bullet$
+(Equation~\ref{eq:thetaspot}). We experimented with three different methods
+for doing posterior inference on our synthetic datasets: No-U Turn Sampling,
+a variant of Hamiltonian Monte Carlo
+\citep[NUTS;][]{?}, automatic differentiation
+variational inference \citep[ADVI;][]{?}, and nested sampling \citep{?}.
+We obtained the best performance using the nested sampling algorithm
+implemented in the \textsf{dynesty} package \citep{?}, so that is what
+we will use below.
+
+Our sampling parameters are the number of spots $n$, their contrast $c$,
+their radius $r$, and the shape parameters $a$ and $b$ describing their
+distribution in latitude. As we discuss in Appendix~\ref{sec:lat},
+the parameters $a$ and $b$ are easier to sample in than
+the mode $\mu_\phi$ and standard deviation $\sigma_\phi$, provided we
+account for the Jacobian of the transformation (Equation~\ref{eq:J})
+in our log probability function, which maps a uniform prior on $a$ and $b$
+to a uniform prior on $\mu_\phi$ and $\sigma_\phi$.
+
+We place uniform priors on all five quantities, with support in
+$1 \leq n \leq 50$, $0 < c \leq 1$,
+$10^\circ \leq r \leq 30^\circ$,
+$0 \leq a \leq 1$, and
+$0 \leq b \leq 1$. Note that while $n$ formally represents an integer,
+its effect on the GP is a scaling of the covariance (see Equation~\ref{eq:exp_yy_sep});
+as such, it has support over all real numbers within the bounds listed
+above. We \emph{could} restrict it to integer values, but this would make
+sampling quite tricky. Moreover, in practice it is useful to allow for
+noninteger values to add some flexibility to the model; we discuss this
+in more detail in \S\ref{sec:integerspots}
+
+We use Equation~(\ref{eq:log-like}) as our log likelihood
+term, adding the log of the absolute value of Equation~(\ref{eq:J}) to
+enforce a uniform prior on $\mu_\phi$ and $\sigma_\phi$. As we mentioned
+above, the fact that $P$, $\mathbf{u}$, and $\sigma_f$ are shared among
+all $M$ light curves means that $\mathbf{K}_m$ is the same for all
+of them, greatly speeding up the likelihood evaluation since we need
+only invert (or factorize) it a single time per sample. Our
+covariance is the covariance of the \emph{normalized} process,
+given by Equation~(\ref{eq:SigmaTilde}). Since we only consider light curves
+with variability limited to a few percent or less, the approximation for
+$\tilde{\pmb{\Sigma}}$ is always valid.
+
+We use the standard implementation of the nested sampler,
+\textsf{dynesty.NestedSampler}, with all arguments set to their
+default values
+(multi-ellipsoidal decomposition for bounds determination,
+uniform sampling within the bounds,
+500 live points,
+and no gradients), to perform our inference step.
+Convergence---defined as when the estimate of the remaining evidence $\Delta\ln \mathcal{Z}$ drops
+below $0.5$---is usually attained after $5{,}000$ to $10{,}000$ samples and
+within a couple hours on a typical machine for most of the trials we perform.
+
+\subsection{Default run}
+\label{sec:inference-default}
+
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_default_corner.pdf}
+        \oscaption{calibration_default}{%
+            Posterior distributions for the spot parameters $\pmb{\theta}_\bullet$
+            (radius $r$ in degrees, latitude mode $\mu_\phi$ and standard deviation $\sigma_\phi$ in degrees,
+            fractional contrast $c$, and number of spots $n$)
+            for the default run (Table~\ref{tab:synthetic} and Figure~\ref{fig:calibration_default_data}).
+            The axes span the entire prior volume, and the blue lines and markers
+            indicate the true (input) values.
+            \label{fig:calibration_default_corner}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_default_latitude.pdf}
+        \oscaption{calibration_default}{%
+            Posterior distributions for the spot latitudes for the default run. Each blue curve corresponds
+            to the Beta distribution PDF (Equation~\ref{eq:phi-pdf}) for the spot latitude
+            with shape parameters drawn from the posterior in $\mu_\phi$ and $\sigma_\phi$
+            (Figure~\ref{fig:calibration_default_corner}); the (hyper)distribution of blue curves
+            quantifies our beliefs about how spots are distributed on any given star. The
+            orange curve is the true distribution used to generate the spots (see Table~\ref{tab:synthetic}).
+            \label{fig:calibration_default_latitude}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_default_inclination.pdf}
+        \oscaption{calibration_default}{%
+            Posterior distributions for the inclinations of individual stars; individual
+            panels correspond directly to those in Figure~\ref{fig:calibration_default_data}.
+            As in Figure~\ref{fig:calibration_default_latitude}, each blue curve is a sample
+            from the hyperdistribution of inclination PDFs, conditioned on a specific value
+            of $\pmb{\theta}_\bullet$ drawn from the posterior. The orange lines indicate the
+            true inclination of each star.
+            \label{fig:calibration_default_inclination}
+        }
+    \end{centering}
+\end{figure}
+
+The input parameters for the default run are shown in Table~\ref{tab:synthetic},
+and the corresponding light curves in Figure~\ref{fig:calibration_default_data}.
+We run the nested sampler as described in the previous section and transform
+the posteriors in $a$ and $b$ to posteriors in $\mu_\phi$ and $\sigma_\phi$ via
+Equations~(\ref{eq:beta2gauss}), (\ref{eq:muphi}), and (\ref{eq:sigmaphi}).
+The results are shown in Figure~\ref{fig:calibration_default_corner}, where
+we correctly infer all five parameters within $2-3$ standard deviations.
+Posterior distributions for the spot radius $r$, the central spot latitude $\mu_\phi$,
+and the spot contrast $c$ are fairly tight, while the distribution for
+the latitudinal scatter $\sigma_\phi$ has wider tails and the distribution for
+the number of spots $n$ is very poorly constrained. The latter, in particular, is
+degenerate with the spot contrast $c$; we discuss this at length in
+\S\ref{sec:cN-degeneracy}.
+
+Figure~\ref{fig:calibration_default_latitude} shows samples from the spot
+latitude posterior (hyper)distribution. Since the parameters $\mu_\phi$
+and $\sigma_\phi$ characterize a distribution over spot latitudes, uncertainty
+in their values translates to uncertainty in the actual shape of the spot
+latitude distribution. Thus, the ensemble of blue curves in
+Figure~\ref{fig:calibration_default_latitude} quantifies our knowledge of
+how spots are distributed on the surfaces of the stars in the dataset.
+These distributions are again consistent with the true distribution used
+to generate the spots (orange curve) within less than 2 standard deviations.
+
+Even though we explicitly marginalized over inclination, we can still
+derive posterior constraints on the inclinations of the individual stars in
+our sample by computing the log-likelihood as a function of $I$
+conditioned on the value of $\pmb{\theta}_\bullet$ from a particular draw from
+the posterior. We do this in Figure~\ref{fig:calibration_default_inclination},
+where blue curves again correspond to samples from the posterior hyperdistribution,
+and orange lines indicate the true inclination; the panels are arranged in the
+same order as those in Figure~\ref{fig:calibration_default_data}.
+In almost all cases, we are able to constrain the inclinations of individual
+stars to within about $10^\circ$, consistent with the truth at less than
+$2{-}3\sigma$.
+
+\xxx{TODO: discuss the batch calibration default run.}
+
+\subsection{Other runs}
+In this section we test the robustness of our model by changing one or
+more of the fiducial values listed in Table~\ref{tab:synthetic}.
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_midlat_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_midlat_latitude.pdf}
+        \oscaption{calibration_midlat}{%
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for
+            mid-latitude spots with $\mu_\phi = 45^\circ$.
+            \label{fig:calibration_midlat}
+        }
+    \end{centering}
+\end{figure}
+
+\xxx{TODO}
 
@@
 Uniform spot radius distribution.
+
+\subsection{Gaussians all the way down}
+\label{sec:other-marg}
+We can also marginalize over period and limb darkening coefficients.
+
+\subsection{Miscellaneous}
+\label{sec:integerspots}
+Shrinkage prior on sigma. Inclination distribution is bimodal.
 
@@
 \begin{proof}{test_lat_transforms}
+    \label{eq:muphi}
     \mu_\phi &= 2 \tan^{-1}
@@
 To compute $\sigma_\phi$, we note that the variance of a Gaussian distribution
-$\varphi(\phi)$
+$\varphi(\phi; \mu, \sigma^2)$
 is the negative reciprocal of its curvature in log space:
@@
 \begin{align}
-    \sigma^2 = -\left(\frac{\mathrm{d}^2 \ln \varphi(\phi)}{\mathrm{d}\phi^2}\right)^{-1}
+    \sigma^2 = -\left(\frac{\mathrm{d}^2 \ln \varphi(\phi; \mu, \sigma^2)}{\mathrm{d}\phi^2}\right)^{-1}
 \end{align}
@@
 \begin{proof}{test_lat_transforms}
+    \label{eq:sigmaphi}
     \sigma_\phi &= \frac{\sin\mu_\phi}{
@@
 \begin{proof}{test_lat_jac}
+    \label{eq:J}
     J &=

### 2020-12-21 tex/figures/calibration_midlat.py

@@ -1,0 +1,19 @@
+from starry_process import calibrate
+import numpy as np
+import os
+import shutil
+
+# Utility funcs to move figures to this directory
+abspath = lambda *args: os.path.join(
+    os.path.dirname(os.path.abspath(__file__)), *args
+)
+copy = lambda name, src, dest: shutil.copyfile(
+    abspath("data", name, src), abspath(dest)
+)
+
+# Run
+calibrate.run(path=abspath("data/midlat"), generate=dict(latitude=dict(mu=45)))
+
+# Copy output to this directory
+copy("midlat", "corner_transformed.pdf", "calibration_midlat_corner.pdf")
+copy("midlat", "latitude.pdf", "calibration_midlat_latitude.pdf")

### 2020-12-22 tex/ms.tex
outsourcing data tarball

@@
 \citep[NUTS;][]{?}, automatic differentiation
-variational inference \citep[ADVI;][]{?}, and nested sampling \citep{?}.
+variational inference \citep[ADVI;][]{?}, and nested sampling \citep{Skilling2004,Skilling2006}.
 We obtained the best performance using the nested sampling algorithm
-implemented in the \textsf{dynesty} package \citep{?}, so that is what
+implemented in the \textsf{dynesty} package \citep{Speagle2020}, so that is what
 we will use below.
@@
 default values
-(multi-ellipsoidal decomposition for bounds determination,
+(multi-ellipsoidal decomposition for bounds determination \citep{Feroz2009},
 uniform sampling within the bounds,
@@
 
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_hilat_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_hilat_latitude.pdf}
+        \oscaption{calibration_hilat}{%
+            Same as Figure~\ref{fig:calibration_midlat}, but for
+            high-latitude spots with $\mu_\phi = 60^\circ$.
+            \label{fig:calibration_hilat}
+        }
+    \end{centering}
+\end{figure}
+
 \xxx{TODO}
@@
 
-\begin{figure}[t!]
+\begin{figure}[p!]
     \begin{centering}
         \includegraphics[width=\linewidth]{figures/nullspace_ensemble.pdf}
+        \\
+        \includegraphics[width=\linewidth]{figures/nullspace_ensemble_ld.pdf}
+        \\
+        \includegraphics[width=\linewidth]{figures/nullspace_ensemble_ld_var.pdf}
         \oscaption{nullspace_ensemble}{%
-            Posterior shrinkage as a function of spherical harmonic degree
+            \emph{Top:} Posterior shrinkage as a function of spherical harmonic degree
             for a hypothetical scenario in which an observer can measure a
-            stellar light curve at 1 (thin blue curves),
-            3 (thin orange curves), 10 (thin green curves), and 30 (thin red
-            curves) random inclinations. The thicker curves correspond to
+            stellar light curve at 1 (blue),
+            3 (orange), 10 (green), and 30 (red)
+            random inclinations. The thicker curves correspond to
             the mean shrinkage for each experiment.
@@
             allows one to recover nearly all of the information in the
-            even spherical harmonic modes.
-            While this thought experiment is not possible in practice for
-            a single star, it is extremely relevant to ensemble analyses of statistically
-            similar stars.
+            even spherical harmonic modes. This is why ensemble analyses of
+            many similar stars at different inclinations allows us to infer
+            their surface properties.
+            \emph{Middle:} Same as the top panel, but for limb-darkened
+            stars with coefficients $u_1 = 0.5$ and $u_2 = 0.25$. Odd
+            modes can now be probed, at the expense of the even modes.
+            \emph{Bottom:} Same as the middle panel, but allowing for
+            a 10\% variation in each of the two limb darkening coefficients
+            across different observations. For ${\gtrsim}30$ light curves,
+            there is virtually no null space up to $l_\mathrm{max} = 15$.
             \label{fig:nullspace_ensemble}
@@
 light curves of ``similar'' stars?
-
-\subsection{Limb darkening}
-\label{sec:limbdark}
-
-See Figures~\ref{fig:nullspace_ensemble_ld} and \ref{fig:nullspace_ensemble_ld_var}.
-
-\begin{figure}[t!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/nullspace_ensemble_ld.pdf}
-        \oscaption{nullspace_ensemble_ld}{%
-            Same as Figure~\ref{fig:nullspace_ensemble}, but for limb-darkened
-            stars with coefficients $u_1 = 0.5$ and $u_2 = 0.25$.
-            \label{fig:nullspace_ensemble_ld}
-        }
-    \end{centering}
-\end{figure}
-
-\begin{figure}[t!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/nullspace_ensemble_ld_var.pdf}
-        \oscaption{nullspace_ensemble_ld_var}{%
-            Same as Figure~\ref{fig:nullspace_ensemble_ld}, but allowing for
-            a 10\% variation in each of the two limb darkening coefficients
-            across different observations.
-            \label{fig:nullspace_ensemble_ld_var}
-        }
-    \end{centering}
-\end{figure}
 
@@
 $\hat{\mathbf{z}}{-}\hat{\mathbf{y}}{-}\hat{\mathbf{z}}$ convention
-and $\imag$ is the imaginary unit. The terms of the $d$-matrix
+and $\imag$ is the imaginary unit
+\citep{AlvarezCollado1989}. The terms of the $d$-matrix
 depend on powers of $\sin\left(\nicefrac{\upbeta}{2}\right)$

### 2020-12-22 tex/ms.tex
chugging along

@@
 
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=0.85\linewidth]{figures/calibration_batch_default_corner.pdf}
+        \oscaption{calibration_batch_default}{%
+            Similar to Figure~\ref{fig:calibration_default_corner}, but
+            showing the posterior distribution for 100 different synthetic
+            datasets, all generated from the same default input parameters
+            (Table~\ref{tab:synthetic}). Each colored histogram corresponds to a single
+            run, with the corresponding $1\sigma$ contours shown as shaded ellipses for each pair
+            of parameters. The black histograms correspond to the distributions of all
+            samples from each of the 100 runs, and the black curves again indicate
+            $1\sigma$ contours in the joint posterior.
+            \label{fig:calibration_batch_default_corner}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=0.5\linewidth]{figures/calibration_batch_default_inclinations.pdf}
+        \oscaption{calibration_batch_default}{%
+            Distribution of stellar inclination residuals normalized to the posterior
+            standard deviation for all $5{,}000$ stars across the 100 trials in
+            Figure~\ref{fig:calibration_batch_default_corner} (blue). The
+            standard normal distribution is shown in orange for comparison.
+            The inclination posteriors inferred with our GP are largely unbiased and
+            have the expected variance.
+            \label{fig:calibration_batch_default_inclinations}
+        }
+    \end{centering}
+\end{figure}
+
 \xxx{TODO: discuss the batch calibration default run.}
+Stellar inclination residuals (posterior mean minus true value) normalized to the
+posterior standard deviation. Only \emph{marginally} overestimate the variance.
 
@@
             \label{fig:calibration_hilat}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_equatorial_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_equatorial_latitude.pdf}
+        \oscaption{calibration_equatorial}{%
+            Same as Figure~\ref{fig:calibration_midlat}, but for
+            equatorial spots with $\mu_\phi = 0^\circ$.
+            \label{fig:calibration_equatorial}
         }
@@
             \emph{Bottom:} Same as the middle panel, but allowing for
-            a 10\% variation in each of the two limb darkening coefficients
+            a 10\% variation in $u_1$ and $u_2$
             across different observations. For ${\gtrsim}30$ light curves,
             there is virtually no null space up to $l_\mathrm{max} = 15$.
+            \codelink{nullspace_ensemble_ld}
+            \codelink{nullspace_ensemble_ld_var}
             \label{fig:nullspace_ensemble}

### 2020-12-22 tex/figures/batch.py

@@ -1,13 +1,0 @@
-from starry_process import calibrate
-import numpy as np
-import os
-
-
-# TODO: Not yet ready for CI runs
-if not bool(os.getenv("CI", False)):
-
-    # Default run
-    calibrate.run_batch(
-        path="data/batch_default", compute_inclination_pdf=False
-    )
-

### 2020-12-22 tex/figures/calibration_bigspots.py
new figs

@@ -1,0 +1,29 @@
+from starry_process import calibrate
+import numpy as np
+import os
+import shutil
+
+# TODO: Not yet ready for CI runs
+if not int(os.getenv("CI", 0)):
+
+    # Utility funcs to move figures to this directory
+    abspath = lambda *args: os.path.join(
+        os.path.dirname(os.path.abspath(__file__)), *args
+    )
+    copy = lambda name, src, dest: shutil.copyfile(
+        abspath("data", name, src), abspath(dest)
+    )
+
+    # Run
+    calibrate.run(
+        path=abspath("data/bigspots"),
+        generate=dict(radius=dict(mu=30, sigma=5), nspots=dict(mu=5)),
+        plot_data=True,  # TODO
+        plot_inclination_pdf=False,
+    )
+
+    # Copy output to this directory
+    copy(
+        "bigspots", "corner_transformed.pdf", "calibration_bigspots_corner.pdf"
+    )
+    copy("bigspots", "latitude.pdf", "calibration_bigspots_latitude.pdf")

### 2020-12-22 tex/ms.tex
new figs
more figures

@@
 posterior standard deviation. Only \emph{marginally} overestimate the variance.
+Tiny sunspots are about $70{,}000$ km across, comparable to large sunspots.
 
@@
             mid-latitude spots with $\mu_\phi = 45^\circ$.
+            The radius and latitude parameters are again inferred correctly.
             \label{fig:calibration_midlat}
@@
         \oscaption{calibration_hilat}{%
-            Same as Figure~\ref{fig:calibration_midlat}, but for
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for
             high-latitude spots with $\mu_\phi = 60^\circ$.
+            The radius and latitude parameters are again inferred correctly,
+            although the model cannot rule out the presence of polar spots.
             \label{fig:calibration_hilat}
@@
         \oscaption{calibration_equatorial}{%
-            Same as Figure~\ref{fig:calibration_midlat}, but for
-            equatorial spots with $\mu_\phi = 0^\circ$.
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for
+            equatorial spots with $\mu_\phi = 0^\circ$. Even though the model
+            favors a bimodal distribution at low latitudes, the
+            posterior strongly supports the presence of equatorial spots.
             \label{fig:calibration_equatorial}
@@
 
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_isotropic_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_isotropic_latitude.pdf}
+        \oscaption{calibration_isotropic}{%
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for
+            isotropically-distributed spots. The posterior accurately
+            captures the cosine-like distribution of spot latitudes.
+            \label{fig:calibration_isotropic}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_tinyspots_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_tinyspots_latitude.pdf}
+        \oscaption{calibration_tinyspots}{%
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for high contrast ($c = 1$)
+            small ($r = 3^\circ$) spots (significantly lower than the
+            effective resolution of the model).
+            The hatched regions in the posterior
+            plots for the radius ($r < 10^\circ$) are excluded by the prior, since
+            the model cannot capture features that small.
+            Despite this, the spot latitude distribution is still inferred correctly,
+            although the spot contrast is off by more than $10\sigma$.
+            \label{fig:calibration_tinyspots}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_default_1_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_default_1_latitude.pdf}
+        \oscaption{calibration_default_1}{%
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for
+            inference based on a single light curve ($M = 1$).
+            The constraints on all of the parameters are dramatically weaker.
+            \label{fig:calibration_default_1}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_default_1000_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_default_1000_latitude.pdf}
+        \oscaption{calibration_default_1000}{%
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for
+            inference based on one thousand light curves ($M = 1{,}000$).
+            The constraints on the radius and the latitude parameters are
+            dramatically tighter.
+            \label{fig:calibration_default_1000}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_ld_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_ld_latitude.pdf}
+        \oscaption{calibration_ld}{%
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for \xxx{...}
+            \label{fig:calibration_ld}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_ld_1000_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_ld_1000_latitude.pdf}
+        \oscaption{calibration_ld_1000}{%
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for \xxx{...}
+            \label{fig:calibration_ld_1000}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_ld_500_no_ld_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_ld_500_no_ld_latitude.pdf}
+        \oscaption{calibration_ld_500_no_ld}{%
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for \xxx{...}
+            \label{fig:calibration_ld_500_no_ld}
+        }
+    \end{centering}
+\end{figure}
+
 \xxx{TODO}
+
+\xxx{More runs: 1) Unnormalized (to show what we can learn if we know the baseline, as in
+    multi-band photometry. 2) Big spots. 3) Single spot. 4) Variance in everything.)}
 

### 2020-12-22 tex/figures/calibration_bigspots.py

@@ -17,7 +17,8 @@ if not int(os.getenv("CI", 0)):
     # Run
     calibrate.run(
         path=abspath("data/bigspots"),
-        generate=dict(radius=dict(mu=30, sigma=5), nspots=dict(mu=5)),
+        generate=dict(radius=dict(mu=45, sigma=5), nspots=dict(mu=1)),
+        sample=dict(rmax=60),
         plot_data=True,  # TODO
         plot_inclination_pdf=False,
     )

### 2020-12-22 tex/figures/calibration_ld_500_no_ld.py

@@ -1,0 +1,29 @@
+from starry_process import calibrate
+import numpy as np
+import os
+import shutil
+
+# Utility funcs to move figures to this directory
+abspath = lambda *args: os.path.join(
+    os.path.dirname(os.path.abspath(__file__)), *args
+)
+copy = lambda name, src, dest: shutil.copyfile(
+    abspath("data", name, src), abspath(dest)
+)
+
+# Run
+calibrate.run(
+    path=abspath("data/ld_500_no_ld"),
+    generate=dict(nlc=500, u=[0.5, 0.25]),
+    sample=dict(u=[0.0, 0.0], compute_inclination_pdf=False),
+    plot_data=False,
+    plot_inclination_pdf=False,
+)
+
+# Copy output to this directory
+copy(
+    "ld_500_no_ld",
+    "corner_transformed.pdf",
+    "calibration_ld_500_no_ld_corner.pdf",
+)
+copy("ld_500_no_ld", "latitude.pdf", "calibration_ld_500_no_ld_latitude.pdf")

### 2020-12-22 tex/figures/calibration_tinyspots.py

@@ -1,0 +1,39 @@
+import numpy as np
+import os
+import shutil
+import corner
+
+old_corner = corner.corner
+
+
+def new_corner(*args, **kwargs):
+    fig = old_corner(*args, **kwargs)
+    ax = np.reshape(fig.axes, (5, 5))
+    for axis in ax[:, 0]:
+        axis.axvspan(0, 10, hatch="//", facecolor="none", lw=0.5, alpha=0.25)
+    return fig
+
+
+corner.corner = new_corner
+
+from starry_process import calibrate
+
+# Utility funcs to move figures to this directory
+abspath = lambda *args: os.path.join(
+    os.path.dirname(os.path.abspath(__file__)), *args
+)
+copy = lambda name, src, dest: shutil.copyfile(
+    abspath("data", name, src), abspath(dest)
+)
+
+# Run
+calibrate.run(
+    path=abspath("data/tinyspots"),
+    generate=dict(radius=dict(mu=3.0, sigma=0.25), contrast=dict(mu=1)),
+    plot_data=False,
+    plot_inclination_pdf=False,
+)
+
+# Copy output to this directory
+copy("tinyspots", "corner_transformed.pdf", "calibration_tinyspots_corner.pdf")
+copy("tinyspots", "latitude.pdf", "calibration_tinyspots_latitude.pdf")

### 2020-12-22 tex/ms.tex
more figs

@@
         \oscaption{calibration_hilat}{%
-            Same as Figures~\ref{fig:calibration_default_corner}
-            and \ref{fig:calibration_default_latitude}, but for
+            Same as Figure~\ref{fig:calibration_midlat}, but for
             high-latitude spots with $\mu_\phi = 60^\circ$.
@@
         \oscaption{calibration_equatorial}{%
-            Same as Figures~\ref{fig:calibration_default_corner}
-            and \ref{fig:calibration_default_latitude}, but for
+            Same as Figure~\ref{fig:calibration_midlat}, but for
             equatorial spots with $\mu_\phi = 0^\circ$. Even though the model
@@
         \oscaption{calibration_isotropic}{%
-            Same as Figures~\ref{fig:calibration_default_corner}
-            and \ref{fig:calibration_default_latitude}, but for
+            Same as Figure~\ref{fig:calibration_midlat}, but for
             isotropically-distributed spots. The posterior accurately
@@
             Same as Figures~\ref{fig:calibration_default_corner}
-            and \ref{fig:calibration_default_latitude}, but for \xxx{...}
+            and \ref{fig:calibration_default_latitude}, but for
+            stars with (assumed known) limb darkening coefficients
+            $u_1 = 0.50$ and $u_2 = 0.25$. Limb darkening makes it
+            much harder to infer the variance of the distribution of
+            star spot latitudes.
             \label{fig:calibration_ld}
@@
         \oscaption{calibration_ld_1000}{%
-            Same as Figures~\ref{fig:calibration_default_corner}
-            and \ref{fig:calibration_default_latitude}, but for \xxx{...}
+            Same as Figure~\ref{fig:calibration_ld}, but for an ensemble
+            consisting of $M = 1{,}000$ light curves. For a
+            sufficiently large ensemble, it is possible to correctly infer the spot
+            radii and latitudes in the presence of limb darkening.
             \label{fig:calibration_ld_1000}
@@
         \oscaption{calibration_ld_500_no_ld}{%
-            Same as Figures~\ref{fig:calibration_default_corner}
-            and \ref{fig:calibration_default_latitude}, but for \xxx{...}
+            Same as Figure~\ref{fig:calibration_ld}, but for $M = 500$
+            light curves and assuming no limb darkening when doing inference.
+            Neglecting limb darkening leads to significant bias in the inferred
+            spot radii and to a lesser extent in the mean spot latitude.
             \label{fig:calibration_ld_500_no_ld}

### 2020-12-22 starry_process/calibrate/plot.py

@@ -245,7 +245,7 @@ def plot_latitude_pdf(results, **kwargs):
     # Plot
     fig, ax = plt.subplots(1)
     for k in range(nlat_samples):
-        ax.plot(x, pdf[k], "C0-", lw=1, alpha=0.05)
+        ax.plot(x, pdf[k], "C0-", lw=1, alpha=0.05, zorder=-1)
     ax.plot(x, pdf_true, "C1-", label="truth")
     ax.plot(x, np.nan * x, "C0-", label="samples")
     ax.legend(loc="upper right")

### 2020-12-22 tex/figures/calibration_unnorm.py

@@ -16,15 +16,12 @@ if not int(os.getenv("CI", 0)):
 
     # Run
     calibrate.run(
-        path=abspath("data/bigspots"),
-        generate=dict(radius=dict(mu=45, sigma=5), nspots=dict(mu=1)),
-        sample=dict(rmax=60),
+        path=abspath("data/unnorm"),
+        generate=dict(normalized=False),
         plot_data=False,
         plot_inclination_pdf=False,
     )
 
     # Copy output to this directory
-    copy(
-        "bigspots", "corner_transformed.pdf", "calibration_bigspots_corner.pdf"
-    )
-    copy("bigspots", "latitude.pdf", "calibration_bigspots_latitude.pdf")
+    copy("unnorm", "corner_transformed.pdf", "calibration_unnorm_corner.pdf")
+    copy("unnorm", "latitude.pdf", "calibration_unnorm_latitude.pdf")

### 2020-12-22 tex/figures/calibration_variance.py

@@ -16,15 +16,16 @@ if not int(os.getenv("CI", 0)):
 
     # Run
     calibrate.run(
-        path=abspath("data/bigspots"),
-        generate=dict(radius=dict(mu=45, sigma=5), nspots=dict(mu=1)),
-        sample=dict(rmax=60),
+        path=abspath("data/variance"),
+        generate=dict(
+            nspots=dict(mu=5), radius=dict(mu=3), contrast=dict(mu=0.01)
+        ),
         plot_data=False,
         plot_inclination_pdf=False,
     )
 
     # Copy output to this directory
     copy(
-        "bigspots", "corner_transformed.pdf", "calibration_bigspots_corner.pdf"
+        "variance", "corner_transformed.pdf", "calibration_variance_corner.pdf"
     )
-    copy("bigspots", "latitude.pdf", "calibration_bigspots_latitude.pdf")
+    copy("variance", "latitude.pdf", "calibration_variance_latitude.pdf")

### 2020-12-23 tex/ms.tex
more figures, more text

@@
 
-\xxx{TODO: discuss the batch calibration default run.}
-Stellar inclination residuals (posterior mean minus true value) normalized to the
-posterior standard deviation. Only \emph{marginally} overestimate the variance.
-Tiny sunspots are about $70{,}000$ km across, comparable to large sunspots.
+The results in
+Figures~\ref{fig:calibration_default_corner}---\ref{fig:calibration_default_inclination}
+are based on a single realization of the light curve ensemble conditioned on
+the properties of Table~\ref{tab:synthetic}. To properly gauge potential biases
+in our model, it is useful to re-run the experiment under many different realizations
+of the dataset. We therefore generate 100 ensembles of light curves in
+exactly the same way as above and run inference on each of them.
+Figure~\ref{fig:calibration_batch_default_corner} shows the marginal and joint
+posterior distributions for $\pmb{\theta}_\bullet$ for all 100 trials.
+Posteriors for individual trials are shown as the transluscent colored curves
+(in the marginal plots) and as ellipses bounding the $1\sigma$ posterior level
+(in the joint posterior plots). The black curves show the marginal distributions of all
+samples across all trials, and the black contours show the corresponding $1\sigma$
+levels in the joint posterior plots.
+
+If our model is truly unbiased, in the limit of an infinite number of realizations
+of the data, the expectation value of the distribution
+of samples across all ensembles (the mean of the black curves in the marginal
+posterior plots) should coincide with the true values (blue lines).
+This is the case with the spot size $r$ posterior: on average, the posterior
+distributions are centered on the correct value.
+However, this is not the case for
+the spot latitude parameters $\mu_\phi$ and $\sigma_\phi$, for which our
+posterior means are biased high. While the \emph{modes} of their
+posteriors are very close to the true values, the distributions have long
+tails extending to high latitudes and high variance, respectively.
+
+The reason for this bias has to do with the baseline degeneracy discussed in
+\S\ref{sec:baseline} and illustrated in Figure~\ref{fig:mean_normalization}:
+the total spottiness of a stellar surface cannot be probed in relative
+photometry. In particular, this means spots near the poles lie almost
+entirely in the null space. Applied to the problem at hand, this degeneracy
+makes it difficult to distinguish between stars with spots concentrated
+exclusively at mid-latitudes (in this case, the truth) and
+stars with spots centered closer to the poles but with large latitudinal
+variance. The latter configuration leads to many spots close to the poles,
+whose effect on the (relative) light curve is negligible, and some spots
+at mid-latitudes, whose effect on the light curve is similar to that of
+the former configuration.
+Thus, the data alone cannot be used to discriminate between these two
+scenarios, introducing the degeneracy we see in the posterior. In fact,
+it is clear that in the tails of the distribution, the mean spot latitude
+and the standard deviation of spot latitudes are positively correlated.
+The bias we see is therefore not a shortcoming of the model, but of the
+\emph{data} itself. To get around this, we either need to impose a
+shrinkage prior on $\sigma_\phi$ (\S\ref{sec:sigma-prior}),
+observe in a different mode that is sensitive to baseline information
+(\S\ref{sec:breaking_baseline_degeneracy}), or simply collect more data.
+As we will see below, the degeneracy described above is not perfect:
+for very large ($M \sim 1{,}000$) ensembles, high-variance polar spots
+can confidently be ruled out.
+
+Lastly, the posteriors for the contrast $c$ and the number of spots $n$
+are mostly unbiased. The contrast distribution has a bit of
+a tail; inspection of Figure~\ref{fig:calibration_batch_default_corner}
+reveals that it too is positively correlated with the mean spot latitude
+and therefore suffers from the same degeneracy as above. And while the
+mean of the spot number distribution is roughly correct, the posterior
+is nearly unchanged across all runs, and equally uninformative in all
+of them. This is yet another manifestation of the baseline degeneracy:
+the total number of spots is not an observable in relative photometry.
+
+There is one final distribution that is instructive to consider: the
+distribution of errors on the inferred stellar inclination.
+Figure~\ref{fig:calibration_batch_default_inclinations} shows a histogram
+of stellar inclination residuals (posterior mean minus true value)
+normalized to the posterior standard deviation for all stars across the
+100 trials described above. For a correctly calibrated model, this distribution
+should equal the standard normal $\mathcal{N}(0, 1^2)$ in the limit of
+infinite trials. This, in fact, is roughly what we find (compare to the
+orange histogram in the figure). Our posterior has marginally
+heavier tails, meaning we tend to \emph{slightly} underestimate the
+posterior variance, but in general it is an excellent estimator of
+individual stellar inclinations.
 
@@
 
-\begin{figure}[p!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/calibration_midlat_corner.pdf}
-        \\[1em]
-        \includegraphics[width=\linewidth]{figures/calibration_midlat_latitude.pdf}
-        \oscaption{calibration_midlat}{%
-            Same as Figures~\ref{fig:calibration_default_corner}
-            and \ref{fig:calibration_default_latitude}, but for
-            mid-latitude spots with $\mu_\phi = 45^\circ$.
-            The radius and latitude parameters are again inferred correctly.
-            \label{fig:calibration_midlat}
-        }
-    \end{centering}
-\end{figure}
-
-\begin{figure}[p!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/calibration_hilat_corner.pdf}
-        \\[1em]
-        \includegraphics[width=\linewidth]{figures/calibration_hilat_latitude.pdf}
-        \oscaption{calibration_hilat}{%
-            Same as Figure~\ref{fig:calibration_midlat}, but for
-            high-latitude spots with $\mu_\phi = 60^\circ$.
-            The radius and latitude parameters are again inferred correctly,
-            although the model cannot rule out the presence of polar spots.
-            \label{fig:calibration_hilat}
-        }
-    \end{centering}
-\end{figure}
-
-\begin{figure}[p!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/calibration_equatorial_corner.pdf}
-        \\[1em]
-        \includegraphics[width=\linewidth]{figures/calibration_equatorial_latitude.pdf}
-        \oscaption{calibration_equatorial}{%
-            Same as Figure~\ref{fig:calibration_midlat}, but for
-            equatorial spots with $\mu_\phi = 0^\circ$. Even though the model
-            favors a bimodal distribution at low latitudes, the
-            posterior strongly supports the presence of equatorial spots.
-            \label{fig:calibration_equatorial}
-        }
-    \end{centering}
-\end{figure}
-
-\begin{figure}[p!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/calibration_isotropic_corner.pdf}
-        \\[1em]
-        \includegraphics[width=\linewidth]{figures/calibration_isotropic_latitude.pdf}
-        \oscaption{calibration_isotropic}{%
-            Same as Figure~\ref{fig:calibration_midlat}, but for
-            isotropically-distributed spots. The posterior accurately
-            captures the cosine-like distribution of spot latitudes.
-            \label{fig:calibration_isotropic}
-        }
-    \end{centering}
-\end{figure}
-
-\begin{figure}[p!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/calibration_tinyspots_corner.pdf}
-        \\[1em]
-        \includegraphics[width=\linewidth]{figures/calibration_tinyspots_latitude.pdf}
-        \oscaption{calibration_tinyspots}{%
-            Same as Figures~\ref{fig:calibration_default_corner}
-            and \ref{fig:calibration_default_latitude}, but for high contrast ($c = 1$)
-            small ($r = 3^\circ$) spots (significantly lower than the
-            effective resolution of the model).
-            The hatched regions in the posterior
-            plots for the radius ($r < 10^\circ$) are excluded by the prior, since
-            the model cannot capture features that small.
-            Despite this, the spot latitude distribution is still inferred correctly,
-            although the spot contrast is off by more than $10\sigma$.
-            \label{fig:calibration_tinyspots}
-        }
-    \end{centering}
-\end{figure}
-
-\begin{figure}[p!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/calibration_default_1_corner.pdf}
-        \\[1em]
-        \includegraphics[width=\linewidth]{figures/calibration_default_1_latitude.pdf}
-        \oscaption{calibration_default_1}{%
-            Same as Figures~\ref{fig:calibration_default_corner}
-            and \ref{fig:calibration_default_latitude}, but for
-            inference based on a single light curve ($M = 1$).
-            The constraints on all of the parameters are dramatically weaker.
-            \label{fig:calibration_default_1}
-        }
-    \end{centering}
-\end{figure}
-
-\begin{figure}[p!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/calibration_default_1000_corner.pdf}
-        \\[1em]
-        \includegraphics[width=\linewidth]{figures/calibration_default_1000_latitude.pdf}
-        \oscaption{calibration_default_1000}{%
-            Same as Figures~\ref{fig:calibration_default_corner}
-            and \ref{fig:calibration_default_latitude}, but for
-            inference based on one thousand light curves ($M = 1{,}000$).
-            The constraints on the radius and the latitude parameters are
-            dramatically tighter.
-            \label{fig:calibration_default_1000}
-        }
-    \end{centering}
-\end{figure}
-
-\begin{figure}[p!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/calibration_ld_corner.pdf}
-        \\[1em]
-        \includegraphics[width=\linewidth]{figures/calibration_ld_latitude.pdf}
-        \oscaption{calibration_ld}{%
-            Same as Figures~\ref{fig:calibration_default_corner}
-            and \ref{fig:calibration_default_latitude}, but for
-            stars with (assumed known) limb darkening coefficients
-            $u_1 = 0.50$ and $u_2 = 0.25$. Limb darkening makes it
-            much harder to infer the variance of the distribution of
-            star spot latitudes.
-            \label{fig:calibration_ld}
-        }
-    \end{centering}
-\end{figure}
-
-\begin{figure}[p!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/calibration_ld_1000_corner.pdf}
-        \\[1em]
-        \includegraphics[width=\linewidth]{figures/calibration_ld_1000_latitude.pdf}
-        \oscaption{calibration_ld_1000}{%
-            Same as Figure~\ref{fig:calibration_ld}, but for an ensemble
-            consisting of $M = 1{,}000$ light curves. For a
-            sufficiently large ensemble, it is possible to correctly infer the spot
-            radii and latitudes in the presence of limb darkening.
-            \label{fig:calibration_ld_1000}
-        }
-    \end{centering}
-\end{figure}
-
-\begin{figure}[p!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/calibration_ld_500_no_ld_corner.pdf}
-        \\[1em]
-        \includegraphics[width=\linewidth]{figures/calibration_ld_500_no_ld_latitude.pdf}
-        \oscaption{calibration_ld_500_no_ld}{%
-            Same as Figure~\ref{fig:calibration_ld}, but for $M = 500$
-            light curves and assuming no limb darkening when doing inference.
-            Neglecting limb darkening leads to significant bias in the inferred
-            spot radii and to a lesser extent in the mean spot latitude.
-            \label{fig:calibration_ld_500_no_ld}
-        }
-    \end{centering}
-\end{figure}
+Figure~\ref{fig:calibration_hilat}
+Figure~\ref{fig:calibration_equatorial}
+Figure~\ref{fig:calibration_isotropic}
+Figure~\ref{fig:calibration_tinyspots}
+Figure~\ref{fig:calibration_default_1}
+Figure~\ref{fig:calibration_default_1000}
+Figure~\ref{fig:calibration_ld}
+Figure~\ref{fig:calibration_ld_1000}
+Figure~\ref{fig:calibration_ld_500_no_ld}
+Figure~\ref{fig:calibration_unnorm}
+Figure~\ref{fig:calibration_bigspots}
+Figure~\ref{fig:calibration_variance}
+Figure~\ref{fig:calibration_batch_default_1000}
+
+Tiny sunspots are about $70{,}000$ km across, comparable to large sunspots.
 
 \xxx{TODO}
-
-\xxx{More runs: 1) Unnormalized (to show what we can learn if we know the baseline, as in
-    multi-band photometry. 2) Big spots. 3) Single spot. 4) Variance in everything.)}
 
@@
 
+\subsection{Prior on sigma}
+\label{sec:sigma-prior}
+Shrinkage prior on sigma.
+
 \subsection{Miscellaneous}
-\label{sec:integerspots}
-Shrinkage prior on sigma. Inclination distribution is bimodal.
+Inclination distribution is bimodal.
 
@@
 
+\section{Supplementary Figures}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_midlat_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_midlat_latitude.pdf}
+        \oscaption{calibration_midlat}{%
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for
+            mid-latitude spots with $\mu_\phi = 45^\circ$.
+            The radius and latitude parameters are again inferred correctly.
+            \label{fig:calibration_midlat}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_hilat_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_hilat_latitude.pdf}
+        \oscaption{calibration_hilat}{%
+            Same as Figure~\ref{fig:calibration_midlat}, but for
+            high-latitude spots with $\mu_\phi = 60^\circ$.
+            The radius and latitude parameters are again inferred correctly,
+            although the model cannot rule out the presence of polar spots.
+            \label{fig:calibration_hilat}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_equatorial_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_equatorial_latitude.pdf}
+        \oscaption{calibration_equatorial}{%
+            Same as Figure~\ref{fig:calibration_midlat}, but for
+            equatorial spots with $\mu_\phi = 0^\circ$. Even though the model
+            favors a bimodal distribution at low latitudes, the
+            posterior strongly supports the presence of equatorial spots.
+            \label{fig:calibration_equatorial}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_isotropic_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_isotropic_latitude.pdf}
+        \oscaption{calibration_isotropic}{%
+            Same as Figure~\ref{fig:calibration_midlat}, but for
+            isotropically-distributed spots. The posterior accurately
+            captures the cosine-like distribution of spot latitudes.
+            \label{fig:calibration_isotropic}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_tinyspots_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_tinyspots_latitude.pdf}
+        \oscaption{calibration_tinyspots}{%
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for high contrast ($c = 1$)
+            small ($r = 3^\circ$) spots (significantly lower than the
+            effective resolution of the model).
+            The hatched regions in the posterior
+            plots for the radius ($r < 10^\circ$) are excluded by the prior, since
+            the model cannot capture features that small.
+            Despite this, the spot latitude distribution is still inferred correctly,
+            although the spot contrast is off by more than $10\sigma$.
+            \label{fig:calibration_tinyspots}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_default_1_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_default_1_latitude.pdf}
+        \oscaption{calibration_default_1}{%
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for
+            inference based on a single light curve ($M = 1$).
+            The constraints on all of the parameters are dramatically weaker.
+            \label{fig:calibration_default_1}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_default_1000_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_default_1000_latitude.pdf}
+        \oscaption{calibration_default_1000}{%
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for
+            inference based on one thousand light curves ($M = 1{,}000$).
+            The constraints on the radius and the latitude parameters are
+            dramatically tighter.
+            \label{fig:calibration_default_1000}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_ld_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_ld_latitude.pdf}
+        \oscaption{calibration_ld}{%
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for
+            stars with (assumed known) limb darkening coefficients
+            $u_1 = 0.50$ and $u_2 = 0.25$. Limb darkening makes it
+            much harder to infer the variance of the distribution of
+            star spot latitudes.
+            \label{fig:calibration_ld}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_ld_1000_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_ld_1000_latitude.pdf}
+        \oscaption{calibration_ld_1000}{%
+            Same as Figure~\ref{fig:calibration_ld}, but for an ensemble
+            consisting of $M = 1{,}000$ light curves. For a
+            sufficiently large ensemble, it is possible to correctly infer the spot
+            radii and latitudes in the presence of limb darkening.
+            \label{fig:calibration_ld_1000}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_ld_500_no_ld_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_ld_500_no_ld_latitude.pdf}
+        \oscaption{calibration_ld_500_no_ld}{%
+            Same as Figure~\ref{fig:calibration_ld}, but for $M = 500$
+            light curves and assuming no limb darkening when doing inference.
+            Neglecting limb darkening leads to significant bias in the inferred
+            spot radii and to a lesser extent in the mean spot latitude.
+            \label{fig:calibration_ld_500_no_ld}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \oscaption{calibration_unnorm}{%
+            \xxx{TODO: Unnormalized.}
+            \label{fig:calibration_unnorm}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \oscaption{calibration_bigspots}{%
+            \xxx{TODO: Large single spot.}
+            \label{fig:calibration_bigspots}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \oscaption{calibration_variance}{%
+            \xxx{TODO: Variance in all input parameters.}
+            \label{fig:calibration_variance}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \oscaption{calibration_batch_default_1000}{%
+            \xxx{TODO: Batch run with $M=1000$.}
+            \label{fig:calibration_batch_default_1000}
+        }
+    \end{centering}
+\end{figure}
+
 \bibliography{bib}

### 2020-12-23 tex/ms.tex
more text, more figs

@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/nongaussianity.pdf}
+        \includegraphics[width=\linewidth]{figures/nongaussianity.png}
         \oscaption{nongaussianity}{%
@@
 with variability limited to a few percent or less, the approximation for
-$\tilde{\pmb{\Sigma}}$ is always valid.
+$\tilde{\pmb{\Sigma}}$ is always valid. Finally, we restrict our spherical
+harmonic expansion to $l_\mathrm{max} = 15$ as a compromise between
+resolution, computational speed, and numerical stability (see \S\ref{sec:implementation}).
 
@@
             \label{fig:calibration_batch_default_inclinations}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \oscaption{calibration_batch_default_1000}{%
+        \xxx{TODO: Batch run with $M=1{,}000$.}
+        \label{fig:calibration_batch_default_1000}
         }
@@
 
-Lastly, the posteriors for the contrast $c$ and the number of spots $n$
+The posteriors for the contrast $c$ and the number of spots $n$
 are mostly unbiased. The contrast distribution has a bit of
@@
 
+Finally, Figure~\ref{fig:calibration_batch_default_1000} shows
+the same posterior distributions as in Figure~\ref{fig:calibration_batch_default_corner},
+but for 100 trials each with $M=1{,}000$ light curves. \xxx{TODO.}
+
 \subsection{Other runs}
@@
 more of the fiducial values listed in Table~\ref{tab:synthetic}.
-
-Figure~\ref{fig:calibration_hilat}
-Figure~\ref{fig:calibration_equatorial}
-Figure~\ref{fig:calibration_isotropic}
-Figure~\ref{fig:calibration_tinyspots}
-Figure~\ref{fig:calibration_default_1}
-Figure~\ref{fig:calibration_default_1000}
+Each of the runs below correspond to a single realization of the ensemble
+dataset, and the corresponding figures are presented at the end of the Appendix.
+
+Figures~\ref{fig:calibration_midlat}---\ref{fig:calibration_isotropic}
+show the results for different latitudinal
+distributions, keeping all other values in Table~\ref{tab:synthetic} the same.
+Specifically, Figure~\ref{fig:calibration_midlat} corresponds to a trial
+with mid-latitude ($\mu_\phi = 45^\circ$ and $\sigma_\phi = 5^\circ$) spots,
+Figure~\ref{fig:calibration_hilat} to a trial with high-latitude
+($\mu_\phi = 60^\circ$ and $\sigma_\phi = 5^\circ$) spots,
+Figure~\ref{fig:calibration_equatorial} to a trial with equatorial
+($\mu_\phi = 0^\circ$ and $\sigma_\phi = 5^\circ$) spots, and
+Figure~\ref{fig:calibration_isotropic} to a trial with isotropically-distributed
+spots ($\phi \sim \cos$).
+The results are largely consistent with those of the default run: in all
+cases we infer the correct spot radius, the mean and standard deviation of
+the spot latitude, and the spot contrast within $2-3\sigma$; the number of
+spots is equally unconstrained in all runs.
+In Figure~\ref{fig:calibration_hilat} and to a lesser extent in
+Figure~\ref{fig:calibration_midlat}, the polar spot degeneracy discussed
+above is evident, particularly in the lower panels showing the latitudinal
+distribution of spots. Nevertheless, the distribution peaks near the correct
+latitude in both cases.
+Figures~\ref{fig:calibration_equatorial} and \ref{fig:calibration_isotropic}
+are interesting because, while the true latitude distribution is unimodal,
+most of the posterior samples are not. In the equatorial case, the posterior
+peaks at very low (but nonzero) latitudes and $\sigma_\phi$ appears to be
+inconsistent with the true value at many standard deviations; however, recall
+that $\sigma_\phi$ is a \emph{local} approximation the standard deviation of the
+PDF at the mode (Appendix~\ref{sec:lat}), which deviates from the true
+standard deviation (i.e., the
+square root of the variance, computed from the expectation of the second
+moment of the distribution) when the two modes are very closely spaced.
+In fact, the latitude PDF samples (lower panel in the figure) nearly span
+the true distribution, to the extent that our parametrization of the
+latitude distribution can approximate a zero-mean Gaussian. While the Beta
+distribution in $\cos\phi$ \emph{can} be unimodal in $\phi$ (see
+Equation~\ref{eq:phi-pdf} and the first
+column of Figure~\ref{fig:latitude_pdf}), this happens \emph{only} when
+$\beta = 0$, which occupies an infinitesimally thin hyperplane in
+parameter space. In practice, the majority of the posterior mass will be close
+to but not exactly at $\beta = 0$, leading to the bimodality in the figure.
+The same argument applies to Figure~\ref{fig:calibration_isotropic}.
+In both cases, the posterior approximates the true distribution as best it
+can given the constraints of the adopted PDF.
+
+Figures~\ref{fig:calibration_tinyspots} tests the performance of the model
+on light curves of stars with spots much smaller than the effective resolution
+of the GP. Our expansion to $l_\mathrm{max} = 15$ only allows us to model spots
+with radii $r \gtrsim 10^\circ$ (see Figure~\ref{fig:spot_profile}), so we
+place zero prior mass below this value. The figure shows the results of inference
+on a dataset generated from spots with $r = 3^\circ$ (and an increased
+contrast $c = 1$ to enforce a comparable signal-to-noise to the other trials).
+On the Sun, these would correspond to spots with diameters of about
+$70{,}000$ km---typical of the larger spots during solar maximum.
+While the radius posterior is biased (as it \emph{must} be, given our prior),
+the fact that it peaks at the lower bound of the prior suggests the presence
+of spots smaller than the model can capture. More importantly, however, the
+latitudinal parameters are inferred correctly and at fairly high precision:
+even though or model is biased against small spots, this does not affect
+inference about their latitudes. On the other hand, the spot contrast is wrong
+by many standard deviations, since the model must compensate for the fact
+that the radii are biased high with a lower contrast to match the variability
+amplitude of the light curves.
+
+Figures~\ref{fig:calibration_default_1} and \ref{fig:calibration_default_1000}
+show results for the default run but with extreme values of the number of
+light curves in the ensemble: $M = 1$ and $M = 1{,}000$, respectively.
+These two figures underscore the power of ensemble analyses: a single light
+curve (Figure~\ref{fig:calibration_default_1}) is simply not informative
+enough about the properties of its spots. On the other hand, a very large
+ensemble can be \emph{extremely} informative: the radius, latitude, and even
+the contrast are inferred correctly at high precision.
+
+Figure~\ref{fig:calibration_ld}---\ref{fig:calibration_ld_500_no_ld}
+show results for the default run but with limb darkening. In all cases
+we assume quadratic limb darkening with fiducial values $u_1 = 0.5$
+and $u_2 = 0.25$ for all stars. From Figure~\ref{fig:calibration_ld}, in which we
+assume we know the limb darkening coefficients exactly, it is clear that
+the presence of limb darkening significantly degrades our ability to
+infer both the radii and latitudes of the spots. Limb darkening has a complicated
+effect on the mapping between surface features and disk-integrated flux, as it 
+reveals information about odd harmonics at
+the expense of introducing strong degeneracies with the even harmonics
+(\S\ref{sec:limbdark}).
+
 Figure~\ref{fig:calibration_ld}
@@
 Figure~\ref{fig:calibration_variance}
-Figure~\ref{fig:calibration_batch_default_1000}
-
-Tiny sunspots are about $70{,}000$ km across, comparable to large sunspots.
 
@@
 
+\subsection{Limb darkening}
+\label{sec:limbdark}
+
 \subsection{Small spots}
@@
 
+\renewcommand{\thefigure}{S\arabic{figure}}
+
 \begin{figure}[p!]
@@
 
-\begin{figure}[p!]
-    \begin{centering}
-        \oscaption{calibration_batch_default_1000}{%
-            \xxx{TODO: Batch run with $M=1000$.}
-            \label{fig:calibration_batch_default_1000}
-        }
-    \end{centering}
-\end{figure}
-
 \bibliography{bib}

### 2020-12-23 tex/figures/calibration_batch_ld_1000.py

@@ -1,0 +1,33 @@
+from starry_process import calibrate
+import numpy as np
+import os
+import shutil
+
+# TODO: Not yet ready for CI runs
+if not int(os.getenv("CI", 0)):
+
+    # Utility funcs to move figures to this directory
+    abspath = lambda *args: os.path.join(
+        os.path.dirname(os.path.abspath(__file__)), *args
+    )
+    copy = lambda name, src, dest: shutil.copyfile(
+        abspath("data", name, src), abspath(dest)
+    )
+
+    # Run
+    calibrate.run_batch(
+        path=abspath("data/batch_ld_1000"),
+        generate=dict(nlc=1000, u=[0.5, 0.25]),
+        sample=dict(compute_inclination_pdf=False, u=[0.5, 0.25]),
+        plot_data=False,
+        plot_corner_transformed=False,
+        plot_latitude_pdf=False,
+        plot_inclination_pdf=False,
+    )
+
+    # Copy output to this directory
+    copy(
+        "batch_ld_1000",
+        "calibration_corner.pdf",
+        "calibration_batch_ld_1000_corner.pdf",
+    )

### 2020-12-23 tex/figures/calibration_hicontrast.py

@@ -1,0 +1,29 @@
+from starry_process import calibrate
+import numpy as np
+import os
+import shutil
+
+# TODO: Not yet ready for CI runs
+if not int(os.getenv("CI", 0)):
+
+    # Utility funcs to move figures to this directory
+    abspath = lambda *args: os.path.join(
+        os.path.dirname(os.path.abspath(__file__)), *args
+    )
+    copy = lambda name, src, dest: shutil.copyfile(
+        abspath("data", name, src), abspath(dest)
+    )
+
+    # Run
+    calibrate.run(
+        path=abspath("data/hicontrast"),
+        generate=dict(nspots=dict(mu=2), contrast=dict(mu=0.5)),
+        sample=dict(compute_inclination_pdf=False),
+    )
+
+    # Copy output to this directory
+    copy(
+        "hicontrast",
+        "corner_transformed.pdf",
+        "calibration_hicontrast_corner.pdf",
+    )

### 2020-12-23 tex/ms.tex
cleaning stuff a bit

@@
 
-\xxx{Brief paper outline here.}
+\xxx{TODO: Brief paper outline here.}
+
+\xxx{TODO: Discuss our open-source figures and equation ``unit tests.''}
 
@@
 
-In the previous section we derived a closed form solution to the (best possible)
+In the previous section we derived a closed form solution to the
 Gaussian approximation to the distribution of stellar surfaces (and their corresponding
@@
 degenerate with the spot contrast $c$; we discuss this at length in
-\S\ref{sec:cN-degeneracy}.
+\S\ref{sec:breaking_baseline_degeneracy}.
 
@@
 the expense of introducing strong degeneracies with the even harmonics
-(\S\ref{sec:limbdark}).
-
-Figure~\ref{fig:calibration_ld}
-Figure~\ref{fig:calibration_ld_1000}
-Figure~\ref{fig:calibration_ld_500_no_ld}
-Figure~\ref{fig:calibration_unnorm}
-Figure~\ref{fig:calibration_bigspots}
-Figure~\ref{fig:calibration_variance}
-
-\xxx{TODO}
+(\S\ref{sec:limbdark}). In practice, this leads to higher uncertainty in
+the spot radii and latitudes relative to the same dataset without limb
+darkening (Figure~\ref{fig:calibration_default_corner}). Fortunately,
+this uncertainty can be dramatically reduced with more data, as evident
+in Figure~\ref{fig:calibration_ld_1000}, which shows the results of the
+same run but with $M=1{,}000$ light curves in the ensemble. The constraints
+on $r$, $\mu_\phi$, $\sigma_\phi$, and $c$ are now much tighter and in
+good agreement with the truth. \xxx{TODO: Is the radius biased high? Discuss the
+    result of the batch run on 1000 limb-darkened light curves.}
+
+\xxx{TODO: Figure~\ref{fig:calibration_unnorm}, Figure~\ref{fig:calibration_hicontrast},
+    Figure~\ref{fig:calibration_bigspots}, Figure~\ref{fig:calibration_variance}.}
 
@@
 
-\subsection{Ensemble analyses}
+\subsection{Why ensemble analyses work}
 \label{sec:ensemble}
 
-Figure~\ref{fig:nullspace_ensemble} shows the posterior shrinkage
-as a function of spherical harmonic degree for three different
-(purely hypothetical) scenarios: a single high signal-to-noise ratio (SNR)
-observation of a stellar light curve at a random
-observer inclination (thin blue curves), a set of three observations
-of the same star at three \emph{different} random inclinations (thin
-orange curves), and a set of ten observations at ten different
-random inclinations (thin green curves).
-The denser, opaque curves show the average for each scenario over 300 trials.
-The posterior shrinkage $R$ for a given
-spherical harmonic coefficient is computed from
+\xxx{TODO: A gentler introduction before diving in.}
+
+Let us define the \emph{posterior shrinkage} $R$ for a given
+spherical harmonic coefficient as
 %
@@
     \frac{\sigma^2}{\sigma_0^2}
+    \quad,
 \end{align}
@@
 and $\sigma^2$ is the posterior variance.
-It is a measure of how informative a measurement is about a
-given mode on the surface in the limit of infinite SNR
+The posterior shrinkage is a measure of how informative a measurement is about a
+given mode on the surface in the limit of infinite signal-to-noise ratio (SNR)
 and is independent of what the stellar surface actually looks like.
@@
 
-\begin{figure}[p!]
+\begin{figure}[t!]
     \begin{centering}
         \includegraphics[width=\linewidth]{figures/nullspace_ensemble.pdf}
-        \\
-        \includegraphics[width=\linewidth]{figures/nullspace_ensemble_ld.pdf}
-        \\
-        \includegraphics[width=\linewidth]{figures/nullspace_ensemble_ld_var.pdf}
         \oscaption{nullspace_ensemble}{%
-            \emph{Top:} Posterior shrinkage as a function of spherical harmonic degree
-            for a hypothetical scenario in which an observer can measure a
-            stellar light curve at 1 (blue),
-            3 (orange), 10 (green), and 30 (red)
-            random inclinations. The thicker curves correspond to
+            Posterior shrinkage as a function of spherical harmonic degree
+            for a hypothetical scenario in which an observer can measure the light curves
+            of 1 (blue), 3 (orange), 10 (green), and 30 (red) \emph{identical} stars
+            viewed at random inclinations.
+            The thicker curves correspond to
             the mean shrinkage for each experiment.
@@
             from a single vantage point approaches zero as $l$
-            increases. However, observing the star from many vantage points
+            increases. However, observing many identical stars from different vantage points
             allows one to recover nearly all of the information in the
@@
             their surface properties.
-            \emph{Middle:} Same as the top panel, but for limb-darkened
-            stars with coefficients $u_1 = 0.5$ and $u_2 = 0.25$. Odd
-            modes can now be probed, at the expense of the even modes.
-            \emph{Bottom:} Same as the middle panel, but allowing for
-            a 10\% variation in $u_1$ and $u_2$
-            across different observations. For ${\gtrsim}30$ light curves,
-            there is virtually no null space up to $l_\mathrm{max} = 15$.
-            \codelink{nullspace_ensemble_ld}
-            \codelink{nullspace_ensemble_ld_var}
             \label{fig:nullspace_ensemble}
@@
 
+Figure~\ref{fig:nullspace_ensemble} shows the posterior shrinkage
+as a function of spherical harmonic degree for a few different hypothetical
+scenarios in which we observe one or more stellar light curves.
 Let us first consider the blue curves, in which a light curve is
-collected from a star viewed at a single random inclination (which, of course, is
+collected from a single star viewed at a single random inclination (which, of course, is
 all we can do from our fixed vantage point on Earth).
+Each thin blue curve corresponds to a particular random draw from the (isotropic)
+inclination distribution; the thick blue curve is the average over 300 trials.
 For some of the low-degree modes, the shrinkage is relatively high: it is
@@
 when viewed from any orientation. Absent structure to break these symmetries
-(see \S\ref{sec:limbdark}), we simply cannot learn anything about these modes from
+(see below), we simply cannot learn anything about these modes from
 stellar light curves. If we average over the shrinkage for all modes up to $l=15$,
@@
 
-The orange curves correspond to the scenario in which we can collect light curves
-from a star viewed at three different random inclinations. This is obviously
-impossible in practice, but it is still instructive as a thought experiment.
+In an ensemble analysis, we assume we observe the lightcurves of many stars
+that are ``similar'' in some statistical sense. As a thought experiment,
+let us consider an extreme version of ensemble analysis in which all the
+stars in our dataset happen to have \emph{identical} surfaces. We will
+still assume they are oriented at random inclinations, as we would expect for
+field stars. The orange curves in the top panel of the figure correspond to
+the case in which we analyze ensembles consisting of three identical stars
+viewed at three different random inclinations.
 Interestingly, the shrinkage increased at all even spherical harmonic degrees
@@
 
-Finally, the green curves correspond to observations taken at ten random
-inclinations. In this scenario, the posterior shrinkage is nearly $100\%$
+Finally, the green and red curves correspond respectively to ensembles of ten and thirty
+identical stars observed at random inclinations.
+In these scenarios, the posterior shrinkage
+is nearly $100\%$
 for all even modes (it asymptotically approaches $100\%$ as the number
@@
 and approaches $50\%$ as both $l_\mathrm{max}$  and the number of light curves
-tend to infinity. Thus, if we were able to measure a light curve of a star
+tend to infinity. Thus, if we were able to measure light curves of identical stars
 from many different inclinations, the null space would consist \emph{only}
 of the odd modes, and our data would tell us exactly half of all the
-information about the surface.
-
-As we mentioned above, it is obviously not possible to observe a \emph{single}
-star from many different inclinations (at least not at present!).
-However, it is certainly the case that we can observe \emph{many} stars
-at a wide range of inclinations. Those stars certainly have different
-surface intensity distributions, but what if we restricted ourselves to
-light curves of ``similar'' stars?
+information about the surfaces of these objects.
 
@@
 
+\xxx{TODO: Discuss Figure~\ref{fig:nullspace_ensemble_ld} and how limb darkening
+    changes things.}
+
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/nullspace_ensemble_ld.pdf}
+        \\
+        \includegraphics[width=\linewidth]{figures/nullspace_ensemble_ld_var.pdf}
+        \oscaption{nullspace_ensemble_ld}{%
+            \emph{Top:} Same as Figure~\ref{fig:nullspace_ensemble},
+            but for limb-darkened
+            stars with coefficients $u_1 = 0.5$ and $u_2 = 0.25$. Odd
+            modes can now be probed, at the expense of the even modes.
+            \emph{Bottom:} Same as the top panel, but allowing for
+            a 10\% variation in $u_1$ and $u_2$
+            across different observations. For ${\gtrsim}30$ light curves,
+            there is virtually no null space up to $l_\mathrm{max} = 15$.
+            \codelink{nullspace_ensemble_ld_var}
+            \label{fig:nullspace_ensemble_ld}
+        }
+    \end{centering}
+\end{figure}
+
 \subsection{Small spots}
@@
 
+\xxx{TODO: Discuss what we should do about tiny spots.}
+
 \subsection{The baseline degeneracy}
@@
 
-\subsection{The number-contrast degeneracy}
-\label{sec:cN-degeneracy}
+\xxx{TODO: Discuss the baseline degeneracy and how to break it. Discuss the
+    degeneracy between $c$ and $N$ and how to break it.}
 
@@
 \label{sec:caveats}
-The latitude distribution is not Gaussian. The number of spots does not have to be an integer.
+\xxx{
+    \begin{enumerate}
+        \item The latitude distribution is not Gaussian.
+        \item The number of spots does not have to be an integer.
+    \end{enumerate}
+}
 
@@
 Inclination distribution is bimodal.
+Careful with light curves with large amplitudes, although
+Figure~\ref{fig:calibration_hicontrast}
+has max amplitudes of more than 10\% and works great.
 
@@
 is extremely degenerate with the total number of spots
-(see \S\ref{sec:cN-degeneracy}).
+(see \S\ref{sec:breaking_baseline_degeneracy}).
 %
@@
     \begin{centering}
+        \oscaption{calibration_hicontrast}{%
+            \xxx{TODO: Two high contrast spots.}
+            \label{fig:calibration_hicontrast}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
         \oscaption{calibration_bigspots}{%

### 2020-12-23 tex/ms.tex
better test comments

@@
 
-\section{Caveats}
+\subsection{Gaussians all the way down}
+\label{sec:other-marg}
+\xxx{TODO: We can also marginalize over period and limb darkening coefficients.}
+
+\subsection{Caveats}
 \label{sec:caveats}
@@
         \item The number of spots does not have to be an integer.
+        \item Careful with light curves with large amplitudes, although
+        Figure~\ref{fig:calibration_hicontrast}
+        has max amplitudes of more than 10\% and works great.
+        \item No differential rotation / spot evolution yet!
     \end{enumerate}
@@
 \label{sec:other-dist}
-Uniform spot radius distribution.
-
-\subsection{Gaussians all the way down}
-\label{sec:other-marg}
-We can also marginalize over period and limb darkening coefficients.
+\xxx{TODO: Discuss the uniform spot radius distribution.}
 
@@
 \label{sec:sigma-prior}
-Shrinkage prior on sigma.
-
-\subsection{Miscellaneous}
-Inclination distribution is bimodal.
-Careful with light curves with large amplitudes, although
-Figure~\ref{fig:calibration_hicontrast}
-has max amplitudes of more than 10\% and works great.
+\xxx{TODO: We might want a shrinkage prior on sigma.}
+
+\subsection{Inferring inclinations}
+\xxx{TODO: Works great. But inclination distribution is bimodal!}
 
@@
 
+\xxx{TODO: Discuss the Python code here.}
+
+\section{Conclusions}
+\xxx{TODO: Summarize. Discuss future work, including differential rotation
+and spot evolution.}
+
 %
@@
 %
+
+\input{table}
 
@@
 %
-\begin{proof}{test_erho}
+\begin{proof}{test_er}
     (e_\mathbb{r})^l_m
@@
 %
-\begin{proof}{test_erho}
+\begin{proof}{test_er}
     c_k(r, \Delta r) & =
@@
 and
-\begin{proof}{test_erho}
+\begin{proof}{test_er}
     \chi_k^-(r, \Delta r) & \equiv \exp\left(\frac{r -\Delta r -\vartheta_k}{s}\right)
@@
 %
-\begin{proof}{test_erho}
+\begin{proof}{test_er}
     \mathbf{e}_r
@@
 %
-\begin{proof}{test_erho}
+\begin{proof}{test_er}
     \lim_{\Delta r \rightarrow 0}
@@
 %
-\begin{proof}{test_bigErho}
+\begin{proof}{test_bigEr}
     (E_\mathbb{r})^{l,l'}_{m,m'}
@@
 %
-\begin{proof}{test_bigErho}
+\begin{proof}{test_bigEr}
     C_{k,k'}(r,\Delta r)
@@
 %
-\begin{proof}{test_bigErho}
+\begin{proof}{test_bigEr}
     \mathbf{E}_r
@@
 %
-\begin{proof}{test_bigErho}
+\begin{proof}{test_bigEr}
     \lim_{\Delta r \rightarrow 0}

### 2020-12-23 tex/tests/test_bigEphi.py

@@ -40,6 +40,7 @@ def Ul(l):
 
 
 def clmmpi(l, m, mp, i):
+    """Return the scalar c^l_{m,mu,i}."""
     if (m - mp - i) % 2 == 0:
         return (
             (-1) ** ((2 * l - m + mp - i) / 2)
@@ -62,6 +63,7 @@ def clmmpi(l, m, mp, i):
 
 
 def Qllpiip(l, lp, i, ip, alpha, beta):
+    """Return the scalar (Q_\phi^{l,l'})_{i,i'}."""
     if (i + ip) % 2 == 0:
         return B(alpha, l + lp + beta - (i + ip) / 2) * hyp2f1(
             -(i + ip) / 2, alpha, l + lp + alpha + beta - (i + ip) / 2, -1
@@ -71,6 +73,7 @@ def Qllpiip(l, lp, i, ip, alpha, beta):
 
 
 def Pllpmmp(l, lp, m, mp, alpha, beta, Ellpr):
+    """Return the scalar (P_\phi^{l,l'})_{m,m'}."""
     term1 = 0
     for mu in range(-l, l + 1):
         for mup in range(-lp, lp + 1):
@@ -91,6 +94,7 @@ def Pllpmmp(l, lp, m, mp, alpha, beta, Ellpr):
 
 
 def Pllp(l, lp, alpha, beta, Ellpr):
+    """Return the matrix P_\phi^{l,l'}."""
     P = np.zeros((2 * l + 1, 2 * lp + 1), dtype="complex128")
     for m in range(-l, l + 1):
         for mp in range(-lp, lp + 1):
@@ -99,6 +103,7 @@ def Pllp(l, lp, alpha, beta, Ellpr):
 
 
 def Ellpphi(l, lp, alpha, beta, Ellpr):
+    """Return the matrix E_\phi^{l,l'}."""
     return (
         np.linalg.inv(Ul(l))
         @ Pllp(l, lp, alpha, beta, Ul(l) @ Ellpr @ Ul(lp).T)
@@ -107,6 +112,7 @@ def Ellpphi(l, lp, alpha, beta, Ellpr):
 
 
 def bigEphi(alpha, beta, Er):
+    """Return the latitude expectation integral E_\phi."""
     lmax = int(np.sqrt(Er.shape[0]) - 1)
     N = (lmax + 1) ** 2
     E = np.zeros((N, N), dtype="complex128")
[…]

### 2020-12-23 tex/tests/test_eI.py

@@ -37,6 +37,7 @@ def Ul(l):
 
 
 def clmmpi(l, m, mp, i):
+    """Return the scalar c^l_{m,mu,i}."""
     if (m - mp - i) % 2 == 0:
         return (
             (-1) ** ((2 * l - m + mp - i) / 2)
@@ -59,6 +60,7 @@ def clmmpi(l, m, mp, i):
 
 
 def qli(l, i):
+    """Return the scalar (q_I^{l})_{i}."""
     return (
         (-1) ** i
         / (l - 0.5 * i + 1)
@@ -67,6 +69,7 @@ def qli(l, i):
 
 
 def plm(l, m, elyp):
+    """Return the scalar (P_I^{l})_{m}."""
     term1 = 0
     for mu in range(-l, l + 1):
         term2 = 0
@@ -77,6 +80,7 @@ def plm(l, m, elyp):
 
 
 def pl(l, elyp):
+    """Return the vector p_I^{l}."""
     p = np.zeros(2 * l + 1, dtype="complex128")
     for m in range(-l, l + 1):
         p[l + m] = plm(l, m, elyp)
@@ -84,11 +88,13 @@ def pl(l, elyp):
 
 
 def elypp(l, elyp):
+    """Return the vector e_{y''}^{l}."""
     U = Ul(l)
     return np.linalg.inv(U) @ pl(l, U @ elyp)
 
 
 def eI(ey):
+    """Return the first moment integral of the inclination distribution."""
     lmax = int(np.sqrt(len(ey)) - 1)
     rTA1 = rTA1Op(ydeg=lmax)().eval()
     e = []
@@ -101,7 +107,7 @@ def eI(ey):
 
 
 def eI_numerical(ey):
-
+    """Return the first moment integral of the inclination distribution, computed numerically."""
     lmax = int(np.sqrt(len(ey)) - 1)
     N = (lmax + 1) ** 2
     rTA1 = rTA1Op(ydeg=lmax)().eval()
@@ -124,6 +130,13 @@ def eI_numerical(ey):
 
 
 def test_eI(lmax=5):
+    """
+    Show that our expression for the first moment
+    expectation integral for the inclination agrees
+    with a numerical estimate. For simplicity, we're
+    making use of the isotropy assumption, so this
+    moment integral is actually just a scalar.
+    """
     np.random.seed(0)
     ey = np.random.randn((lmax + 1) ** 2)
     assert np.allclose(eI(ey), eI_numerical(ey))

### 2020-12-23 tex/tests/test_lat_jac.py

@@ -2,6 +2,7 @@ import numpy as np
 
 
 def get_mu(a, b):
+    """Compute the mode of the latitude distribution."""
     alpha = np.exp(a * 5)
     beta = np.exp(np.log(0.5) + b * (5 - np.log(0.5)))
     term = (
@@ -16,6 +17,7 @@ def get_mu(a, b):
 
 
 def get_sigma(a, b):
+    """Compute the local standard deviation of the latitude distribution."""
     alpha = np.exp(a * 5)
     beta = np.exp(np.log(0.5) + b * (5 - np.log(0.5)))
     mu = get_mu(a, b)
@@ -31,7 +33,7 @@ def get_sigma(a, b):
 
 def get_J(a, b):
     """
-    Jacobian for the transformation (a, b) <--> (mu, sigma)
+    Compute the Jacobian for the transformation (a, b) <--> (mu, sigma).
 
     """
     alpha = np.exp(a * 5)
@@ -57,6 +59,11 @@ def get_J(a, b):
 
 
 def test_jacobian(ntests=100, eps=1e-6):
+    """
+    Numerically show that our expression for the Jacobian agrees
+    with its definition in terms of derivatives.
+    
+    """
     np.random.seed(0)
 
     for n in range(ntests):

### 2020-12-23 tex/tests/test_lat_transforms.py

@@ -3,22 +3,26 @@ from scipy.stats import beta as Beta
 
 
 def pdf(phi, alpha, beta):
+    """Latitude PDF as a function of latitude."""
     return 0.5 * np.abs(np.sin(phi)) * Beta.pdf(np.cos(phi), alpha, beta)
 
 
 def deriv_pdf(phi, alpha, beta, eps=1e-8):
+    """Derivative of `pdf` with respect to latitude."""
     return (pdf(phi + eps, alpha, beta) - pdf(phi - eps, alpha, beta)) / (
         2 * eps
     )
 
 
 def log_pdf(phi, alpha, beta):
+    """Natural log of `pdf`."""
     return np.log(0.5 * np.abs(np.sin(phi))) + Beta.logpdf(
         np.cos(phi), alpha, beta
     )
 
 
 def second_deriv_log_pdf(phi, alpha, beta, eps=1e-4):
+    """Second derivative of `log_pdf` with respect to latitude."""
     return (
         log_pdf(phi + eps, alpha, beta)
         - 2 * log_pdf(phi, alpha, beta)
@@ -27,6 +31,7 @@ def second_deriv_log_pdf(phi, alpha, beta, eps=1e-4):
 
 
 def get_mu(alpha, beta):
+    """Compute the mode of the latitude distribution."""
     term = (
         4 * alpha ** 2
         - 8 * alpha
@@ -39,6 +44,7 @@ def get_mu(alpha, beta):
 
 
 def get_sigma(alpha, beta):
+    """Compute the local standard deviation of the latitude distribution."""
     mu = get_mu(alpha, beta)
     term = (
         1
@@ -51,6 +57,7 @@ def get_sigma(alpha, beta):
 
 
 def get_alpha(mu, sigma):
+    """Compute the Beta shape parameter #1."""
     v = sigma ** 2
     c1 = np.cos(mu)
     c2 = np.cos(2 * mu)
@@ -60,6 +67,7 @@ def get_alpha(mu, sigma):
 
 
 def get_beta(mu, sigma):
+    """Compute the Beta shape parameter #2."""
     v = sigma ** 2
     c1 = np.cos(mu)
     c2 = np.cos(2 * mu)
@@ -70,7 +78,8 @@ def get_beta(mu, sigma):
 
 def test_mu(ntests=100):
     """
-    Show that the derivative of the PDF is zero at phi = mu.
+    Show that the derivative of the PDF is zero at phi = mu;
[…]

### 2020-12-23 tex/ms.tex
misc

@@
     \begin{centering}
-        \includegraphics[width=0.85\linewidth]{figures/calibration_batch_default_corner.pdf}
+        \includegraphics[width=0.85\linewidth]{figures/calibration_batch_default_corner.png}
         \oscaption{calibration_batch_default}{%
@@
     \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_unnorm_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_unnorm_latitude.pdf}
         \oscaption{calibration_unnorm}{%
-            \xxx{TODO: Unnormalized.}
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but assuming
+            the light curves are not normalized and the true baseline is
+            known. Knowledge of the baseline breaks the $c-N$ degeneracy
+            and allows us to infer the total number of spots.
             \label{fig:calibration_unnorm}
@@
     \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_hicontrast_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_hicontrast_latitude.pdf}
         \oscaption{calibration_hicontrast}{%
-            \xxx{TODO: Two high contrast spots.}
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for stars with
+            $n = 2$ spots with high contrast $c = 0.5$. Our model correctly
+            captures the increased contrast, but it is still strongly degenerate
+            with the number of spots.
             \label{fig:calibration_hicontrast}
@@
     \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_bigspots_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_bigspots_latitude.pdf}
         \oscaption{calibration_bigspots}{%
-            \xxx{TODO: Large single spot.}
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but for a single ($n = 1$)
+            large ($r \sim \mathcal{N}(45^\circ, {5^\circ}^2)$) spot on each star.
+            Our model correctly infers the larger radius, and infers the
+            latitude within $3\sigma$, albeit with large uncertainty.
             \label{fig:calibration_bigspots}
@@
         \oscaption{calibration_variance}{%
-            \xxx{TODO: Variance in all input parameters.}
+            \xxx{
+                TODO: Same as Figures~\ref{fig:calibration_default_corner}
+                and \ref{fig:calibration_default_latitude}, but for stars
+                with variance in their number of spots,
+                $n \sim \mathcal{N}(20, 3^2)$,
+                the spot radii, $r \sim \mathcal{N}(15^\circ, {3^\circ}^2)$,
+                and the spot contrasts,
+                $c \sim \mathcal{N}(0.05, 0.01^2)$.
+            }
             \label{fig:calibration_variance}

### 2020-12-25 tex/ms.tex
momre figs

@@
 
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/stability.pdf}
+        \oscaption{stability}{%
+            Log of the condition number of the covariance in the
+            spherical harmonic basis, $\pmb{\Sigma}_{\mathbf{y}}$,
+            as a function of the spherical harmonic degree of the
+            expansion, $l_\mathrm{max}$. Different lines correspond
+            to different values of $\pmb{\theta}_\bullet$ drawn
+            from a uniform prior (see text for details).
+            In the majority of the cases, the matrix becomes ill-conditioned
+            above $l_\mathrm{max} = 15$.
+            \label{fig:stability}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/speed.pdf}
+        \oscaption{speed}{%
+            Speed tests \xxx{(TODO)}.
+            \label{fig:speed}
+        }
+    \end{centering}
+\end{figure}
+
 \section{Conclusions}

### 2020-12-25 tex/figures/speed.py

@@ -1,0 +1,55 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from starry_process import StarryProcess
+import theano
+import theano.tensor as tt
+from tqdm import tqdm
+import os
+import time
+
+
+# Compile the function to compute the likelihood
+def get_loglike_function(ydeg, marginalize_over_inclination=False):
+    t = tt.dvector()
+    return theano.function(
+        [t],
+        StarryProcess(
+            ydeg=ydeg,
+            marginalize_over_inclination=marginalize_over_inclination,
+        ).log_likelihood(t, tt.ones_like(t), 1.0),
+    )
+
+
+ydegs = [5, 10, 15]
+npts = np.array(np.logspace(0, 4, 20), dtype=int)
+ncalls = 10
+tloglike = np.zeros((len(ydegs), len(npts)))
+
+for i in tqdm(range(len(ydegs)), disable=bool(int(os.getenv("NOTQDM", "0")))):
+    get_loglike = get_loglike_function(ydegs[i], False)
+    for j in range(len(npts)):
+        tstart = time.time()
+        t = np.linspace(0, 1, npts[j])
+        for k in range(ncalls):
+            get_loglike(t)
+        tloglike[i, j] = (time.time() - tstart) / ncalls
+
+
+fig, ax = plt.subplots(1)
+for i in range(len(ydegs)):
+    ax.plot(
+        npts, tloglike[i], label=r"$l_\mathrm{{max}} = {}$".format(ydegs[i])
+    )
+
+ax.legend(fontsize=10)
+ax.set_xlabel("number of points")
+ax.set_ylabel("likelihood evaluation time [s]")
+ax.set_yscale("log")
+ax.set_xscale("log")
+
+# We're done
+fig.savefig(
+    os.path.abspath(__file__).replace(".py", ".pdf"),
+    bbox_inches="tight",
+    dpi=300,
+)

### 2020-12-26 tex/ms.tex
more figs and more text

@@
     \begin{centering}
+        \includegraphics[width=0.85\linewidth]{figures/calibration_batch_default_1000_corner.png}
         \oscaption{calibration_batch_default_1000}{%
@@
         \oscaption{speed}{%
-            Speed tests \xxx{(TODO)}.
+            Evaluation time in seconds for a single log-likelihood evaluation
+            as a function of the number of points $K$ in each light curve
+            when conditioning on a value of the inclination (blue) and when
+            marginalizing over the inclination (orange). At $l_\mathrm{max} = 15$,
+            computation of the covariance matrix of the GP takes about 10ms
+            on a 2018 MacBook Pro. The dashed line shows the asymptotic scaling
+            of the algorithm, which is due to the Cholesky factorization and
+            solve operations.
             \label{fig:speed}
@@
 
+\bibliography{bib}
+
+%
+%
+%
+%
+
+\pagebreak
 \appendix
@@
 
+\begin{table}[p!]
     \input{table}
+\end{table}
+
+%
+%
+%
+%
 
@@
 
-\bibliography{bib}
 \end{document}

### 2020-12-26 tex/figures/speed.py

@@ -20,13 +20,15 @@ def get_loglike_function(ydeg, marginalize_over_inclination=False):
     )
 
 
-ydegs = [5, 10, 15]
-npts = np.array(np.logspace(0, 4, 20), dtype=int)
+# Settings
+ydeg = 15
+npts = np.sort(list(set(list(np.array(np.logspace(0, 4, 20), dtype=int)))))
 ncalls = 10
-tloglike = np.zeros((len(ydegs), len(npts)))
 
-for i in tqdm(range(len(ydegs)), disable=bool(int(os.getenv("NOTQDM", "0")))):
-    get_loglike = get_loglike_function(ydegs[i], False)
+# Compute
+tloglike = np.zeros((2, len(npts)))
+for i, marginalize_over_inclination in enumerate([False, True]):
+    get_loglike = get_loglike_function(ydeg, marginalize_over_inclination)
     for j in range(len(npts)):
         tstart = time.time()
         t = np.linspace(0, 1, npts[j])
@@ -34,16 +36,32 @@ for i in tqdm(range(len(ydegs)), disable=bool(int(os.getenv("NOTQDM", "0")))):
             get_loglike(t)
         tloglike[i, j] = (time.time() - tstart) / ncalls
 
-
+# Plot
 fig, ax = plt.subplots(1)
-for i in range(len(ydegs)):
-    ax.plot(
-        npts, tloglike[i], label=r"$l_\mathrm{{max}} = {}$".format(ydegs[i])
-    )
+ax.plot(npts, tloglike[0], label="conditional")
+ax.plot(npts, tloglike[1], label="marginal")
 
-ax.legend(fontsize=10)
+# Asymptotic
+n = np.array([2e2, 1.0e4])
+ax.plot(n, 3e-10 * n ** 2.5, "k--", alpha=0.5)
+ax.annotate(
+    r"$\propto K^{2.5}$",
+    fontsize=16,
+    xy=(2e3, 1e-2),
+    xycoords="data",
+    xytext=(0, 0),
+    textcoords="offset points",
+    ha="center",
+    va="center",
+    rotation=35,
+    alpha=0.5,
+)
+
+# Appearance
+ax.margins(0, 0)
+ax.legend(fontsize=12)
 ax.set_xlabel("number of points")
-ax.set_ylabel("likelihood evaluation time [s]")
+ax.set_ylabel("likelihood eval time [s]")
 ax.set_yscale("log")
 ax.set_xscale("log")
 

### 2020-12-26 tex/ms.tex
new test

@@
 %
-\begin{align}
+\begin{proof}{test_normgp}
     \label{eq:SigmaTilde}
@@
     \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_variance_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_variance_latitude.pdf}
         \oscaption{calibration_variance}{%
-            \xxx{
-                TODO: Same as Figures~\ref{fig:calibration_default_corner}
+            Same as Figures~\ref{fig:calibration_default_corner}
             and \ref{fig:calibration_default_latitude}, but for stars
@@
             $c \sim \mathcal{N}(0.05, 0.01^2)$.
-            }
             \label{fig:calibration_variance}

### 2020-12-27 tex/ms.tex
better download
more text

@@
 \begin{align}
+    \label{eq:fnormal}
     \mathbbb{f} \sim \mathcal{N}\left( \pmb{\mu}, \pmb{\Sigma} \right)
@@
 \begin{align}
+    \label{eq:kernel}
     \Sigma_{i,j} & = k(t_i, t_j)
@@
     \mathbf{r}_m^\top\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
-    \mathbf{K}_m^{-1}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
+    \big[
+        \pmb{\Sigma}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
+        +
+        \mathbf{C}_m
+        \big]^{-1}
     \mathbf{r}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
@@
     \frac{1}{2}
-    \ln \Big| \mathbf{K}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right) \Big|
+    \ln \Big|
+    \pmb{\Sigma}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
+    +
+    \mathbf{C}_m
+    \Big|
     -
@@
 %
-\begin{align}
-    \mathbf{K}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
-     & \equiv
-    \pmb{\Sigma}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
-    +
-    \mathbf{C}_m
-\end{align}
-%
-is the sum of the GP covariance and the data covariance $\mathbf{C}_m$
+$\mathbf{C}_m$ is the data covariance
 (which in most cases is a diagonal matrix whose entries
@@
     \pmb{\mathcal{A}}(\mathbb{I}, P, \mathbf{u}) \,
-    \mathbf{E} \Big[ \mathbbb{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
+    \mathrm{E} \Big[ \mathbbb{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
     p(\mathbb{I}) \, \mathrm{d}\mathbb{I}
@@
     \pmb{\mathcal{A}}(\mathbb{I}, P, \mathbf{u}) \,
-    \mathbf{E} \Big[ \mathbbb{y} \, \mathbbb{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
+    \mathrm{E} \Big[ \mathbbb{y} \, \mathbbb{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
     \pmb{\mathcal{A}}^\top(\mathbb{I}, P, \mathbf{u})
@@
 \begin{align}
+    \label{eq:scalar-mean}
     \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)
@@
 \begin{align}
+    \label{eq:z}
     z & \equiv \frac{\left< \Sigma \right>}{\mu^2}
@@
 second moments of the distribution of spherical harmonic coefficient vectors,
-$\mathbf{E}\left[ \mathbbb{y} \, \big| \, \pmb{\theta}_\bullet \right]$
+$\mathrm{E}\left[ \mathbbb{y} \, \big| \, \pmb{\theta}_\bullet \right]$
 and
-$\mathbf{E}\left[ \mathbbb{y} \, \mathbbb{y}^\top \, \big| \, \pmb{\theta}_\bullet \right]$,
+$\mathrm{E}\left[ \mathbbb{y} \, \mathbbb{y}^\top \, \big| \, \pmb{\theta}_\bullet \right]$,
 given by Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), respectively.
@@
 above, the fact that $P$, $\mathbf{u}$, and $\sigma_f$ are shared among
-all $M$ light curves means that $\mathbf{K}_m$ is the same for all
+all $M$ light curves means that $\pmb{\Sigma} + \mathbf{C}_m$ is the same for all
 of them, greatly speeding up the likelihood evaluation since we need
@@
 
-\begin{figure}[p!]
+\begin{figure}[t!]
     \begin{centering}
@@
         \oscaption{calibration_batch_default_1000}{%
-        \xxx{TODO: Batch run with $M=1{,}000$.}
+        Same as Figure~\ref{fig:calibration_batch_default_corner}, but
+        showing the posterior distributions of 100 realizations of ensembles
+        with $M=1{,}000$ (instead of 50) light curves each. The constraints on
+        mosts of the parameters are much tighter, and the polar spot degeneracy
+        is gone.
         \label{fig:calibration_batch_default_1000}
@@
 posterior plots) should coincide with the true values (blue lines).
-This is the case with the spot size $r$ posterior: on average, the posterior
+This is approximately
+the case with the spot size $r$ posterior: on average, the posterior
 distributions are centered on the correct value.
@@
 the same posterior distributions as in Figure~\ref{fig:calibration_batch_default_corner},
-but for 100 trials each with $M=1{,}000$ light curves. \xxx{TODO.}
+but for 100 trials each with $M=1{,}000$ light curves. In addition to the
+constraints on all parameters (except the number of spots) being much
+tighter, the larger amount of data breaks the polar spot degeneracy discussed
+above. Given enough light curves, the model is capable of differentiating between
+concentrated mid-latitude spots and high-latitude spots with large variance.
+Interestingly, the inferred radius appears to be biased high by a small amount
+in this run. This is likely due to the fact that our prescription for
+generating the spots (\S\ref{sec:calibration-setup}) is different from how
+we actually model these spots. While we generate the spots as compact
+circular disks expanded at high spherical harmonic degree, we model them
+as sigmoids (\S\ref{sec:size}) expanded at significantly lower spherical
+harmonic degree. Some minor disagreement is therefore to be expected in the inferred
+radii.
 
@@
 Figure~\ref{fig:calibration_isotropic} to a trial with isotropically-distributed
-spots ($\phi \sim \cos$).
+spots ($\bbphi \sim \cos$).
 The results are largely consistent with those of the default run: in all
@@
 on $r$, $\mu_\phi$, $\sigma_\phi$, and $c$ are now much tighter and in
-good agreement with the truth. \xxx{TODO: Is the radius biased high? Discuss the
+good agreement with the truth.
+
+\xxx{TODO: Discuss the
     result of the batch run on 1000 limb-darkened light curves.}
@@
 \begin{align}
+    \label{eq:shrinkage}
     R \equiv 1 - \lim\limits_{\sigma_0^2 \rightarrow \infty}
@@
 
-\begin{table}[p!]
+\afterpage{
     \input{table}
-\end{table}
+    \vfill
+    \clearpage
+}
 
@@
 %
+
+For reference, Table~\ref{tab:variables} lists the principal symbols,
+operators, and variables used throughout the paper, with links to
+the equations and/or section in which they are presented.
 
@@
     \label{eq:ld:I}
-    \frac{I(\mu)}{I(\mu = 1)} = 1 - \sum_{n=1}^{n_\mathrm{max}} u_n(1 - \mu)^n
+    \frac{I(\upmu)}{I(\upmu = 1)} = 1 - \sum_{n=1}^{n_\mathrm{max}} u_n(1 - \upmu)^n
     \quad,
@@
 where $I$ is the intensity on the stellar surface,
-$\mu = z = \sqrt{1 - x^2 - y^2}$ is the radial coordinate on the
+$\upmu = z = \sqrt{1 - x^2 - y^2}$ is the radial coordinate on the
 projected disk, and $u_n$ is a limb darkening coefficient, the effect of
@@
 \begin{align}
+    \label{eq:x}
     \mathbbb{x} =
@@
 \begin{align}
-    \mathbf{B} \, \tilde{\mathbf{s}}(r) = \mathbf{b}(r)
+    \mathbf{B} \, \mathbf{s_0}(r) = \mathbf{b}(r)
 \end{align}
@@
 %
-and $\tilde{\mathbf{s}}(r)$ are the coefficients of the expansion.
+and $\mathbf{s_0}(r)$ are the coefficients of the expansion.
 These are related to the full vector of spherical harmonic coefficients describing
@@
 \begin{align}
-    s^l_m(r) = \tilde{s}_l(r) \delta_{m,0}
+    s^l_m(r) = s^l_0 \delta_{m,0}
     \quad,
@@
 \begin{align}
-    \mathbf{s}(r) = \pmb{\mathcal{I}} \, \tilde{\mathbf{s}}(r)
+    \label{eq:sofr}
+    \mathbf{s}(r) = \pmb{\mathcal{I}} \, \mathbf{s_0}(r)
 \end{align}
@@
 and $\delta$ is the Kronecker delta function.
-To find the coefficients $\tilde{\mathbf{s}}(r)$
+To find the coefficients $\mathbf{s_0}(r)$
 (and hence $\mathbf{s}(r)$), we solve the (linear) inverse problem,
@@
 \begin{align}
-    \tilde{\mathbf{s}}(r) & = \mathbf{B}^+ \, \mathbf{b}(r)
+    \mathbf{s_0}(r) & = \mathbf{B}^+ \, \mathbf{b}(r)
 \end{align}
@@
 \begin{align}
-    S_{k,l} = \exp\left[-\frac{l(l + 1)}{2\sigma^2}\right] \delta_{k,l}
-\end{align}
-%
-for smoothing strength $\sigma$. For $\epsilon \rightarrow 0$ and
-$\sigma \rightarrow \infty$,
+    S_{k,l} = \exp\left[-\frac{l(l + 1)}{2\xi^2}\right] \delta_{k,l}
+\end{align}
+%
+for smoothing strength $\xi$. For $\epsilon \rightarrow 0$ and
+$\xi \rightarrow \infty$,
 $\mathbf{B}^+$ is the exact pseudo-inverse of
@@
 $\epsilon > 0$ is chosen for improved numerical stability and
-$\sigma > 0$ is chosen to mitigate the effect of ringing in the solution.
+$\xi > 0$ is chosen to mitigate the effect of ringing in the solution.
 In practice, we obtain good results with $\epsilon \approx 10^{-9}$
-and $\sigma \approx 15$.
+and $\xi \approx 15$.
 
@@
 evident, although it is strongly suppressed compared to an expansion
-without the smoothing term (i.e., $\sigma = \infty$). However, for
+without the smoothing term (i.e., $\xi = \infty$). However, for
 $r \lesssim 10^\circ$, an expansion to $l_{\mathrm{max}} = \LMAX$ is
@@
 \begin{align}
-    \pmb{\theta}_\mathbb{r} = \left(
+    \pmb{\theta}_r = \left(
     r \, \, \, \,
@@
 \begin{proof}{test_er}
-    (e_\mathbb{r})^l_m
+    (e_r)^l_m
     & =
@@
     \int
-    \xi \,
-    p(\xi \, \big| \, \pmb{\theta}_{c}) \,
-    \mathrm{d}\xi
+    \mathbb{c} \,
+    p(\mathbb{c} \, \big| \, \pmb{\theta}_{c}) \,
+    \mathrm{d}\mathbb{c}
     \\
@@
     \int
-    \xi^2 \,
-    p(\xi \, \big| \, \pmb{\theta}_\xi)
-    \mathrm{d}\xi
+    \mathbb{c}^2 \,
+    p(\mathbb{c} \, \big| \, \pmb{\theta}_c)
+    \mathrm{d}\mathbb{c}
     \quad.
@@
 These integrals may be computed analytically for any choice of
-probability density function $p(\xi \, \big| \, \pmb{\theta}_\xi)$
+probability density function $p(\mathbb{c} \, \big| \, \pmb{\theta}_c)$
 with closed-form moments.
@@
 \begin{align}
-    p(\xi \, \big| \, \pmb{\theta}_{c}) = \delta(\xi - c)
+    p(\mathbb{c} \, \big| \, \pmb{\theta}_{c}) = \delta(\mathbb{c} - c)
 \end{align}
@@
 \begin{align}
-    \pmb{\theta}_\xi = \left( \, c \, \right)^\top
+    \pmb{\theta}_c = \left( \, c \, \right)^\top
     \quad.
@@
 
-\section{Supplementary Figures}
+%
+%
+%
+% Supplementary Figures
+%
+%
+%
 

### 2020-12-28 tex/ms.tex
more text

@@
 
-\xxx{TODO: Brief paper outline here.}
-
-\xxx{TODO: Discuss our open-source figures and equation ``unit tests.''}
+The paper is organized as follows:
+we present an overview of the derivation of the GP in \S\ref{sec:gp}
+and a suite of tests on synthetic data to show the model is calibrated
+in \S\ref{sec:calibration}. We discuss our GP from an information
+theoretical perspective in \S\ref{sec:discussion}, focusing on
+the various degeneracies our model can and cannot break when applied
+to ensembles of light curves. In \S\ref{sec:implementation} we present
+our implementation of the algorithm in the open-source \starryprocess package
+alongside speed and stability tests, and in \S\ref{sec:conclusions} we
+summarize our results. Most of the math behind the algorithm
+is presented in the Appendix, followed by a series of supplementary figures
+from \S\ref{sec:calibration}. Appendix~\ref{sec:notation} discusses the
+notation we adopt throughout the paper and Table~\ref{tab:variables}
+lists the main symbols and variables, with links to their definitions.
+
+Finally, we note that all of the figures in this paper were auto-generated
+using the Azure Pipelines continuous integration (CI) service, which
+ensures they are up to date with the latest version of the
+\starryprocess code. In particular, icons next to each of the figures \codeicon \,
+link to the exact script used to generate them to ensure the reproducibility
+of our results. In this paper we also introduce the concept of equation
+``unit tests'': \textsf{pytest}-compatible test scripts associated
+with the principal equations that pass (fail) if the equation is correct (wrong),
+in which case a clickable \testpassicon \, (\testfailicon) is shown next to the equation
+label.
+In most cases, the validity of an equation is gauged by comparison to
+a numerical solution. Like the figure scripts, the equation unit tests are
+run on Azure Pipelines upon every commit of the code.%
+\footnote{
+    These unit tests are certainly not foolproof: in particular, there is
+    no guarantee against a mismatch in the \LaTeX \, version of an equation
+    and its \Python implementation (e.g., due to an uncaught typo). However, they
+    \emph{do} ensure that the linked \Python implementation is correct to
+    within the accuracy of the numerical solution, providing readers with a
+    valid implementation of the equation for purposes of reproducibility.
+}
 
@@
 a variant of Hamiltonian Monte Carlo
-\citep[NUTS;][]{?}, automatic differentiation
-variational inference \citep[ADVI;][]{?}, and nested sampling \citep{Skilling2004,Skilling2006}.
+\citep[NUTS;][]{Hoffman2011}, automatic differentiation
+variational inference \citep[ADVI;][]{Kucukelbir2016,Blei2016}, and nested sampling \citep{Skilling2004,Skilling2006}.
 We obtained the best performance using the nested sampling algorithm
@@
             for the default run (Table~\ref{tab:synthetic} and Figure~\ref{fig:calibration_default_data}).
-            The axes span the entire prior volume, and the blue lines and markers
+            The axes span the entire prior volume, and the orange lines and markers
             indicate the true (input) values.
@@
 of samples across all ensembles (the mean of the black curves in the marginal
-posterior plots) should coincide with the true values (blue lines).
+posterior plots) should coincide with the true values (orange lines).
 This is approximately
@@
 \subsection{Inferring inclinations}
-\xxx{TODO: Works great. But inclination distribution is bimodal!}
+\xxx{TODO: Works great. Note that inclination distribution is sometimes bimodal!}
 
@@
 \section{Conclusions}
+\label{sec:conclusions}
 \xxx{TODO: Summarize. Discuss future work, including differential rotation

### 2020-12-28 starry_process/calibrate/plot.py

@@ -387,6 +387,7 @@ def plot_corner(results, transform_beta=False, **kwargs):
         smooth1d=2.0,
         bins=100,
         hist_kwargs=dict(lw=1),
+        truth_color="#ff7f0e",
         **plot_kwargs
     )
 
@@ -624,7 +625,7 @@ def plot_batch(path):
             hist_kwargs=dict(alpha=0.5 * batch_alpha),
             range=ranges,
             truths=truths_,
-            truth_color="#4682b4",
+            truth_color="#ff7f0e",
         )
 
         # Running list

### 2020-12-28 starry_process/latitude.py

@@ -2,6 +2,7 @@ from .wigner import R
 from .integrals import WignerIntegral
 from .ops import LatitudeIntegralOp, CheckBoundsOp
 from .defaults import defaults
+from .math import is_tensor
 import theano.tensor as tt
 from theano.ifelse import ifelse
 from scipy.stats import beta as Beta
@@ -22,8 +23,8 @@ def gauss2beta(
     closest to the Gaussian with mean ``mu`` and standard deviation ``sigma``.
 
     Args:
-        mu (float or ndarray): The mean latitude in degrees.
-        sigma (float or ndarray): The latitude standard deviation in degrees.
+        mu (scalar or vector): The mean latitude in degrees.
+        sigma (scalar or vector): The latitude standard deviation in degrees.
         log_alpha_max (float, optional): The maximum value of ``log(alpha)``. 
             Default is %%defaults["log_alpha_max"]%%.
         log_beta_max (float, optional): The maximum value of ``log(beta)``. 
@@ -40,28 +41,36 @@ def gauss2beta(
 
     .. note:: 
         
-        This is a utility function that accepts and returns numeric values
-        (not tensors).
+        This is a utility function that can accept and return
+        either numeric values or tensors. If both `mu` and `sigma`
+        are numeric quantities, the result will be a numeric
+        quantity; otherwise it will be a tensor.
 
     """
-    is_vector = hasattr(mu, "__len__")
-    if is_vector:
-        assert hasattr(sigma, "__len__")
-        assert len(mu) == len(sigma)
+    if is_tensor(mu, sigma):
+        math = tt
+        is_vector = True
+        m = mu * np.pi / 180
+        v = (sigma * np.pi / 180) ** 2
     else:
-        assert not hasattr(sigma, "__len__")
-    angle_fac = np.pi / 180
-    m = np.atleast_1d(mu) * angle_fac
-    v = (np.atleast_1d(sigma) * angle_fac) ** 2
-    c1 = np.cos(m)
-    c2 = np.cos(2 * m)
-    c3 = np.cos(3 * m)
-    term = 1.0 / (16 * v * np.cos(0.5 * m) ** 4)
+        math = np
+        is_vector = hasattr(mu, "__len__")
+        if is_vector:
+            assert hasattr(sigma, "__len__")
[…]

### 2020-12-28 tex/ms.tex
new figure, more text

@@
 evaluating the GP likelihood may typically be done in a fraction of a second
-for a reasonably-sized dataset (i.e., $K \lesssim 10^5$ datapoints).
+for a reasonably-sized dataset (i.e., $K \lesssim 10^4$ datapoints).
 
@@
 is presented in the Appendix, followed by a series of supplementary figures
-from \S\ref{sec:calibration}. Appendix~\ref{sec:notation} discusses the
+(discussed in \S\ref{sec:calibration}). Appendix~\ref{sec:notation} discusses the
 notation we adopt throughout the paper and Table~\ref{tab:variables}
@@
             The real spherical harmonics in the polar frame up to
-            $l = 5$. Rows correspond to the degree $l$ and columns to
+            $l = 5$, where dark colors correspond to negative intensity
+            and bright colors to positive intensity.
+            Rows correspond to the degree $l$ and columns to
             the order $m$. The set of all spherical harmonics forms a
@@
 where $n$ is the number of star spots, $c$ is their contrast,
-$\mu_\phi$ and $\sigma_\phi$ are parameters describing the
-distribution of the spots in latitude, and $r$ is the radius
+$\mu_\phi$ and $\sigma_\phi$ are the mode and standard deviation
+of the spot latitude distribution, respectively, and $r$ is the radius
 of the spots.
@@
 a variant of Hamiltonian Monte Carlo
-\citep[NUTS;][]{Hoffman2011}, automatic differentiation
+\citep[NUTS;][]{Duane1987,Hoffman2011}, automatic differentiation
 variational inference \citep[ADVI;][]{Kucukelbir2016,Blei2016}, and nested sampling \citep{Skilling2004,Skilling2006}.
@@
 more of the fiducial values listed in Table~\ref{tab:synthetic}.
-Each of the runs below correspond to a single realization of the ensemble
+Each of the runs below corresponds to a single realization of the ensemble
 dataset, and the corresponding figures are presented at the end of the Appendix.
@@
 on $r$, $\mu_\phi$, $\sigma_\phi$, and $c$ are now much tighter and in
-good agreement with the truth.
-
-\xxx{TODO: Discuss the
-    result of the batch run on 1000 limb-darkened light curves.}
-
-\xxx{TODO: Figure~\ref{fig:calibration_unnorm}, Figure~\ref{fig:calibration_hicontrast},
-    Figure~\ref{fig:calibration_bigspots}, Figure~\ref{fig:calibration_variance}.}
+good agreement with the truth. Finally, Figure~\ref{fig:calibration_ld_500_no_ld}
+shows the results of inference on limb-darkened light curves under the
+(wrong) assumption that limb darkening is not present ($\mathbf{u} = \mathbf{0}$).
+Neglecting the effect of limb darkening leads to significant biases in the spot radius
+and latitude parameters. While the model still favors mid-latitude spots
+(at $\sim 45^\circ$ instead of $30^\circ$), the constraints are deceptively tight
+and discrepant by many standard deviations. We discuss these points in
+more detail in \S\ref{sec:limbdark}.
+
+The runs so far correspond to stars with many ($n = 20$) spots, for which the
+resulting light curves are smooth due to the fact that many spots are in
+view at any given time. Figures~\ref{fig:calibration_hicontrast} and
+\ref{fig:calibration_bigspots} show what happens when the model is applied to
+stars with $n=2$ and $n=1$ spots, respectively. Despite large portions of the
+light curves being flat (and therefore extremely non-stationary)
+in these scenarios, the GP does surprisingly well, recovering the radii
+and latitude parameters within $2-3\sigma$ in both cases.
+Note that in order to preserve the same signal-to-noise ratio relative to
+the other runs, the spots in Figure~\ref{fig:calibration_hicontrast}
+have much higher contrast ($c = 0.5$). Even though the contrast is degenerate
+with the number of spots (which is unobservable), the $c$ posterior
+has a much heavier tail than in the other runs, showing that the model is
+in fact sensitive to the spot contrast to some degree. In
+Figure~\ref{fig:calibration_bigspots}, we compensate for the smaller number
+of spots by increasing the spot radius to $r = 45^\circ \pm 5^\circ$ instead,
+showing that the model can accurately model large spots, even in the
+presence of some (unmodelled) scatter in their sizes.
+
+In Figure~\ref{fig:calibration_variance} we add variance to all the spot
+properties when generating the light curves: we add $n = 20 \pm 3$ spots
+to each star with radii $r = 15^\circ \pm 3^\circ$, contrasts
+$c = 0.05 \pm 0.01$, and at latitudes $\phi = 30^\circ \pm 5^\circ$.
+As before, we only explicitly account for the latitude variance in our
+model. We correctly infer the latitude parameters and the contrast, but
+our radii appear to be biased high. This is likely due to the fact that
+larger spots have a bigger impact on the signal, so our inferred radius
+is a weighted average of all spot radii. In Appendix~\ref{sec:size} we
+derive an expression for the moment integrals of the spot size distribution
+assuming a uniform distribution between $r - \Delta r$ and $r + \Delta r$
+(instead of a delta function at $r$), which can be used to compute the
+GP if one wishes to explicitly account for scatter in the spot sizes.
+We find that repeating the trial shown in Figure~\ref{fig:calibration_variance}
+while explicitly sampling over the distribution in $\Delta r$ shifts the
+posterior mass to lower radii, mitigating the bias described above.
+
+Our final trial is shown in Figure~\ref{fig:calibration_unnorm}, in which
+we assume we know the true baseline of each light curve. That is, we
+assume that we can measure all light curves in units of the flux we would
+measure if the stars had no spots on them, and we \emph{do not}
+normalize them (see \S\ref{sec:baseline}). In practice, this would require
+knowledge of the brightness (or temperature) of the unspotted
+photosphere, which is not an observable in single-band relative photometry.
+This value can be probed, however, in multi-band photometry
+(see \S\ref{sec:breaking_baseline_degeneracy}), for which this trial is
+extremely relevant. We again recover the radii and latitude parameters
+to within $2-3\sigma$, but most importantly, we also infer the correct
+spot contrast \emph{and} the correct number of spots with fairly high
+precision. In particular, knowledge of the baseline breaks the $c-N$ degeneracy.
+Photometric measurements in multiple bands (even just two!)
+are therefore extremely useful when inferring spot properties.
+We discuss this point at length in \S\ref{sec:breaking_baseline_degeneracy}.
 
@@
 
+\subsection{The null space}
+\label{sec:nullspace}
+
+In general, inferring the properties of a stellar surface from its light curve is not
+only difficult, but \emph{formally impossible}. To understand why, consider
+an expansion of the stellar surface intensity in the spherical
+harmonic basis (see Figure~\ref{fig:ylms}). Assuming (for the moment) that the
+star rotates about an axis that points up along the page, the observed
+light curve may be expressed as a weighted sum of the disk-integrated intensity
+of each of the spherical harmonics as they rotate about that same axis.
+However, not all spherical harmonics will contribute to the full light curve,
+as many (in fact, most) of the spherical harmonics are perfectly antisymmetric
+about the equator. This is the case for
+the $l = 1$, $m = -1$ harmonic, which integrates to zero regardless of
+the phase at which it is viewed. The same is true, in fact, for all other harmonics
+of order $m = -1$ and (perhaps less obviously) for all harmonics with odd
+$l = 3, 5, 7, ...$ Furthermore, there exist many linear combinations of
+spherical harmonics that similarly integrate to zero at all rotational
+phases. Together, these modes constitute the \emph{null space} of the problem:
+the set of modes on the surface that do not contribute to the observed
+light curve and therefore cannot be probed from photometry.
+As we will see below, the vast majority of the surface modes lie at least in part in the null
+space.
+
+But what if the star rotates about a different axis? A useful property of
+the spherical harmonics is that rotation about any axis can only change the
+order $m$ of the harmonic, while keeping the degree $l$ fixed: in other
+words, rotation simply mixes the spherical harmonics along rows in
+Figure~\ref{fig:ylms}. Therefore, if we rotate the spherical harmonics
+in the figure about a different axis, we are still subject to just as
+many degeneracies (i.e., the size of the null space is unchanged),
+but the particular linear combination of modes in the null space will
+be different.
+
+This all seems quite hopeless, and for analyses of individual stars,
+it pretty much is. Unless extremely strong (and likely incorrect) priors are employed
+(such as assuming the stellar surface consists of exactly two perfectly
+circular, uniform contrast spots), we simply cannot learn much about
+the surface of an individual star from photometry.
+
+However, we can use the dependence of the null space on inclination to our
+advantage. If we could observe a star from many different vantage points,
+we would be able to break many of the degeneracies at play, since we
+would get different constraints on the amplitude of each mode when viewed
+at different inclinations. This, of course, is not possible (at least not
+yet!). But what we \emph{can} do is observe many similar stars, each viewed
+at a different (random) inclination, and attempt to learn something about
+the properties of the ensemble of stars as a whole. This is exactly
+what we did in \S\ref{sec:calibration}, where we showed that this kind
+of ensemble analyses is not only possible but can also be extremely
+constraining of the spot properties at the population level. In the
+following section, we explore the role of ensemble analyses in breaking
+the degeneracies of the mapping problem in more detail.
+
 \subsection{Why ensemble analyses work}
@@
 
-\xxx{TODO: A gentler introduction before diving in.}
-
-Let us define the \emph{posterior shrinkage} $R$ for a given
+To quantify the power of ensemble analyses,
+let us define the \emph{posterior shrinkage} $R$ for a given
 spherical harmonic coefficient as
@@
 shrinkage is defined to be unity. Conversely, if the data is completely
-unconstraining, $R$ will tend to zero.
+unconstraining of that mode (i.e., it is entirely in the null space),
+$R$ will tend to zero.
 
@@
 surface looks like. As $l$ increases further, $R$ tends to zero.
-Another important feature of the shrinkage is that it is exactly zero for
+Another important feature of the shrinkage, which we hinted at above,
+is that it is exactly zero for
 all odd-degree modes above $l = 1$. This is a well-known fact: all odd spherical
@@
 from different vantage points broke some degeneracies in the problem.
-This is a consequence of the fact that the null space (for the even modes)
-is a strong function of the inclination: modes that are invisbile
-at one inclination can actually project into the flux at a different inclination.
+This is a consequence of the fact we mentioned in the previous section:
+the null space (for the even modes)
+is a strong function of the inclination.
 If we average over all modes, we obtain a total shrinkage of $\sim 24\%$
@@
 information about the surfaces of these objects.
+
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/evenmodes.pdf}
+        \oscaption{evenmodes}{%
+            A spotted stellar surface seen in Mollweide projection
+            (left), the same surface with all odd $l > 2$ harmonics
+            set to zero (center), and the same surface with $l = 1$
+            and all even harmonics set to zero (right). The plot
+            on the right represents the portion of the surface that is
+            in the null space of the light curve problem at all
+            inclinations; the plot in the center is the portion of the
+            surface that can be inferred at some value of the inclination.
+            \label{fig:evenmodes}
+        }
+    \end{centering}
+\end{figure}
+
+Figure~\ref{fig:evenmodes} shows a spotted stellar surface decomposed
+into its \emph{preimage} and its \emph{null space} in the limiting case
+in which its light curve can be observed at a large number of inclinations.
+The former is the set of all modes that map onto the light curve (i.e.,
+the complement of the null space). Because spots are compact features, they
+are necessarily made up of a continuum of spherical harmonic modes spanning
+many different values of $l$: they can therefore be seen in both the even
+modes (the preimage) and the odd modes (the null space). Interestingly,
+the symmetries at play require spots to be paired with antipodal dark mirror
+images in the even modes, and with \emph{bright} ones in the odd modes
+(which sum to perfectly cancel out in the true map). Since these reflections
+are at the same latitude (in the opposite hemisphere) as the true spots,
+they do not contribute any bias to our inference, which already assumes the
+latitude distribution is symmetric about the equator. Thus, even though
+we are only sensitive to the even modes in ensemble analyses, they are
+sufficient to learn about the properties we are interested in.
+
+In practice, the fact that stars are not identical complicates this picture.
+We can still learn about the \emph{distribution} of spot properties, but
+in general our ensemble must consist of far more light curves in order to
+place strong constraints on these properties. As we showed in \S\ref{sec:calibration},
+under ideal conditions an ensemble of $M = 50$ similar stars may be sufficient,
+but in practice we may need on the order of hundreds or even a thousand light curves
+in order to mitigate the effects of variance in the sample.
 
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/calibration_unnorm_corner.pdf}
-        \\[1em]
-        \includegraphics[width=\linewidth]{figures/calibration_unnorm_latitude.pdf}
-        \oscaption{calibration_unnorm}{%
-            Same as Figures~\ref{fig:calibration_default_corner}
-            and \ref{fig:calibration_default_latitude}, but assuming
-            the light curves are not normalized and the true baseline is
-            known. Knowledge of the baseline breaks the $c-N$ degeneracy
-            and allows us to infer the total number of spots.
-            \label{fig:calibration_unnorm}
-        }
-    \end{centering}
-\end{figure}
-
-\begin{figure}[p!]
-    \begin{centering}
         \includegraphics[width=\linewidth]{figures/calibration_hicontrast_corner.pdf}
@@
 
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/calibration_unnorm_corner.pdf}
+        \\[1em]
+        \includegraphics[width=\linewidth]{figures/calibration_unnorm_latitude.pdf}
+        \oscaption{calibration_unnorm}{%
+            Same as Figures~\ref{fig:calibration_default_corner}
+            and \ref{fig:calibration_default_latitude}, but assuming
+            the light curves are not normalized and the true baseline is
+            known. Knowledge of the baseline breaks the $c-N$ degeneracy
+            and allows us to infer the total number of spots.
+            \label{fig:calibration_unnorm}
+        }
+    \end{centering}
+\end{figure}
+
 \end{document}

### 2020-12-28 tex/ms.tex
more text

@@
 
-\xxx{TODO: Discuss the Python code here.}
+We implement our interpretable GP in the open-source, user-friendly
+\Python package \starryprocess, which can be installed via \textsf{pip}\xspace
+or from source on GitHub\footnote{\url{https://github.com/rodluger/starry_process}}.
+The code is thoroughly unit-tested and well documented, with examples on how
+to use the GP in custom inference problems. Users can choose, among other
+options, whether or not
+to marginalize over inclination (\S\ref{sec:inclination}) and
+whether or not to model a normalized
+process (\S\ref{sec:baseline}). Users can also choose the spherical harmonic
+degree of the expansion, although it is recommended to use $l_\mathrm{max} = 15$
+(see below).
+Users may compute the mean vector and covariance matrix in either the spherical
+harmonic basis or the flux basis, or they may sample from it or use it to
+compute marginal likelihoods.
+Arbitrary order limb darkening is implemented following
+\citet{Agol2020}.
+
+The code was designed to maximize the speed and numerical stability of the
+computation. Although most of the computations of the expectation integrals
+(see Appendix~\ref{sec:integrals}) involve many layers of nested sums over
+all spherical harmonic coefficients, these may be expressed as high-dimensional
+tensor products, which can be evaluated extremely efficiently on modern
+hardware. Many of the expressions can also be either pre-computed or
+computed recursively. To maximize the speed of the algorithm, the code is
+implemented in hybrid \cpp/\Python using the just-in-time compilation
+capability of the \theano package. Since all equations derived here
+have closed form expressions, these can be differentiated in a straightforward
+and numerically stable manner, enabling the computation of backpropagated
+gradients within \theano. As such, \starryprocess is designed to work
+out-of-the box with \theano-based inference tools such as \pymcthree
+for NUTS/HMC or ADVI sampling.
+
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/speed.pdf}
+        \oscaption{speed}{%
+            Evaluation time in seconds for a single log-likelihood evaluation
+            as a function of the number of points $K$ in each light curve
+            when conditioning on a value of the inclination (blue) and when
+            marginalizing over the inclination (orange). At $l_\mathrm{max} = 15$,
+            computation of the covariance matrix of the GP takes about 10ms
+            on a 2018 MacBook Pro. The dashed line shows the asymptotic scaling
+            of the algorithm, which is due to the Cholesky factorization and
+            solve operations.
+            \label{fig:speed}
+        }
+    \end{centering}
+\end{figure}
+
+Figure~\ref{fig:speed} shows the computational scaling of the \Python
+implementation of the algorithm for the case where we condition the GP
+on a specific value of the inclination (blue) and the case where we marginalize
+over inclination (orange). Both curves show the time in seconds to compute the
+likelihood (averaged over many trials) as a function of the number of points $K$
+in a single light curve. For $K \lesssim 100$, the computation time is constant
+at $10-20$ ms for both algorithms. This is the approximate time (on a typical
+modern laptop) taken to compute the GP covariance matrix given a set of
+hyperparameters $\pmb{\theta}_\bullet$. For larger values of $K$, the
+cost approaches a scaling of $K^{2.5}$, which is dominated by the factorization
+of the covariance matrix and the solve operation to compute the likelihood.
+The likelihood marginalized over inclination is only slightly slower to compute,
+thanks to the tricks discussed in Appendix~\ref{sec:inc-speedup}.
+
+Many modern GP packages \citep[e.g.,][]{Ambikasaran2015,ForemanMackey2017}
+have significantly better asymptotic scalings, but these are usually due to
+specific structure imposed on the kernel functions, such as the assumption
+of stationarity. Our kernel structure is determined by the physics (or perhaps
+more accurately, the geometry) of stellar surfaces, and its nonstationarity
+is a consequence of the normalization step in relative photometry. Moreover,
+and unlike the kernels listed in \S\ref{sec:intro}, our
+kernel is a nontrivial function of the hyperparameters
+$\pmb{\theta}_\bullet$, so its computation is necessarily more expensive.
+Nevertheless, the fact that our GP may be used for likelihood evaluation
+in a small fraction of a second for typical datasets ($K \sim 1{,000}$)
+makes it extremely useful for inference. Recall that we are implicitly
+marginalizing over all of the properties of \emph{every spot} on the surface of
+the star.
+
+In ensemble analyses, we must compute the likelihood of each of the $M$ light curves
+conditioned on $\pmb{\theta}_\bullet$. In practice, each star will have a
+different rotation period, different limb darkening coefficients, and different
+photometric uncertainty, meaning we must factorize $M$ different covariance
+matrices. Fortunately, the spherical harmonic covariance, $\pmb{\Sigma}_\mathbf{y}$
+(\S\ref{sec:gp}) need only be computed once; this can then be linearly
+transformed into the flux basis for each light curve. As the evaluation of
+$\pmb{\Sigma}_\mathbf{y}$ is the computationally-intensive step, the
+likelihood evaluation typically scales sub-linearly with $M$. Furthermore,
+as we discussed in \ref{sec:other-marg}, it is possible to marginalize over
+the period and limb darkening coefficients, which would remove the scaling
+with $M$ entirely if the photometric precision were the same for all
+light curves. Even if it is not, the algorithm can still be greatly sped up,
+although an implementation of this is deferred to the next paper.
+\xxx{DFM: Is this true? I.e., is there an efficient way to solve Equation~(\ref{eq:log-like})
+    when $\pmb{\Sigma}$ is the same for all stars but $\mathbf{C}$ is different? Is there
+    some limit in which an approximation of some sort can be made?}
 
@@
 
-\begin{figure}[t!]
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/speed.pdf}
-        \oscaption{speed}{%
-            Evaluation time in seconds for a single log-likelihood evaluation
-            as a function of the number of points $K$ in each light curve
-            when conditioning on a value of the inclination (blue) and when
-            marginalizing over the inclination (orange). At $l_\mathrm{max} = 15$,
-            computation of the covariance matrix of the GP takes about 10ms
-            on a 2018 MacBook Pro. The dashed line shows the asymptotic scaling
-            of the algorithm, which is due to the Cholesky factorization and
-            solve operations.
-            \label{fig:speed}
-        }
-    \end{centering}
-\end{figure}
+Our algorithm is also numerically stable over nearly all of the prior
+volume up to $l_\mathrm{max} = 15$. Figure~\ref{fig:stability}
+shows the log of the condition number of the covariance matrix in the
+spherical harmonic basis, $\pmb{\Sigma}_\mathbf{y}$, as a function of the
+spherical harmonic degree of the expansion for 100 draws from a uniform
+prior over $r \in [10^\circ, 45^\circ]$, $\mu_\phi \in [0^\circ, 85^\circ]$,
+$\sigma_\phi \in [5^\circ, 40^\circ]$, and $n \in [1, 50]$. The condition
+number is nearly constant up to $l_\mathrm{max} = 15$ in almost all cases;
+above this value, the algorithm suddenly becomes unstable and the
+covariance is ill-conditioned. The instability occurs within the computation
+of the latitude and longitude moment integrals (Appendix~\ref{sec:integrals})
+and is likely due to the large number of operations involving linear
+combinations of hypergeometric and gamma functions. While it may be possible to
+achieve stability at higher values of $l_\mathrm{max}$ via careful
+reparametrization of some of those equations, we find that $l_\mathrm{max} = 15$
+is high enough for most practical purposes. We plan to revisit this point
+in future work.
+
+Finally, instabilities can also occur if $\sigma_\phi$ is too small and/or
+$n$ is too large. Values of $\sigma_\phi \lesssim 1^\circ$ lead to instabilities
+in the computation of the hypergeometric functions (Appendix~\ref{sec:lat}),
+while values of $n \gtrsim 50$ can sometimes cause the Cholesky factorization
+of the covariance to fail (although this can be mitigated by adding a small
+quantity to the diagonal to ensure positive-semidefinitess). In cases where
+the algorithm goes (very) unstable, the log-likelihood evaluation returns
+$-\infty$: in other words, they are silently rejected by an implicit prior.
+Fortunately, these cases are likely unphysical: in practice, there should always be some
+finite amount of variance in the latitudes of spots, and stars with more than
+50 spots are likely too spotted for individual spots to be discernible in the
+first place. Instead, we are likely sensitive to \emph{groups} of spots, which
+our GP is flexible enough to model.
 

### 2020-12-28 tex/ms.tex
more words

@@
 
-Figures~\ref{fig:calibration_tinyspots} tests the performance of the model
+Figure~\ref{fig:calibration_tinyspots} tests the performance of the model
 on light curves of stars with spots much smaller than the effective resolution
@@
 as a function of spherical harmonic degree for a few different hypothetical
-scenarios in which we observe one or more stellar light curves.
+scenarios in which we observe one or more stellar light curves with
+no limb darkening.
 Let us first consider the blue curves, in which a light curve is
@@
 
-\xxx{TODO: Discuss Figure~\ref{fig:nullspace_ensemble_ld} and how limb darkening
-    changes things.}
-
 \begin{figure}[t!]
@@
             a 10\% variation in $u_1$ and $u_2$
-            across different observations. For ${\gtrsim}30$ light curves,
+            across different observations. For $\gtrsim 30$ light curves,
             there is virtually no null space up to $l_\mathrm{max} = 15$.
@@
 \end{figure}
+
+As we saw in Figure~\ref{fig:calibration_ld}, the presence of limb darkening
+makes it significantly more difficult to infer spot properties at fixed
+SNR. To understand
+why, let us repeat the experiment from the previous section,
+this time with moderate quadratic limb darkening ($u_1 = 0.5$ and $u_2 = 0.25$).
+The top panel of Figure~\ref{fig:nullspace_ensemble_ld} shows the
+posterior shrinkage plot (same as Figure~\ref{fig:nullspace_ensemble}, but
+with limb darkening). Interestingly, there is no longer a clean division of
+the null space between even and odd modes in the limit of large $M$.
+This is because limb darkening effectively lifts odd modes out of the null space, \emph{at the
+    expense of the even modes}. While no coefficient lies entirely in the
+null space ($R = 0$) when limb darkening is present, no coefficient
+can be uniquely inferred ($R = 1$), either. This can be understood by noting that
+a polynomial limb darkening law can be written
+exactly as a linear combination of the $m=0$ spherical harmonics
+up to a degree equal to the order of the limb darkening
+\citep[in this case, $l = 2$;][and Appendix~\ref{sec:ld}]{Luger2019,Agol2020}.
+Since the limb darkening operation is a (multiplicative) downweighting of
+the surface intensity, the map seen by the observer is just the product
+of the spherical harmonic representation of the surface ($\mathbf{y}$)
+and the spherical harmonic representation of the limb darkening profile.
+And since spherical harmonics are just polynomials on the surface of the sphere,
+the product of spherical harmonics of degree $l_1$ and $l_2$ is a spherical
+harmonic of degree $l_1 + l_2$. This means that the linear limb darkening
+component ($l = 1$) effectively raises the degree of all spherical harmonic
+coefficients of the surface map by one. This has the effect of reversing
+the null space: under \emph{only} linear limb darkening, it is the \emph{even}
+modes that would like in the null space. However, the quadratic
+limb darkening term ($l = 2$) raises the degree of all spherical harmonics by two,
+so its presence ensures that the even modes can still be probed to some extent.
+In reality, the true limb darkening profile of a stellar surface is more
+complicated than a two-parameter quadratic model can capture; but one may still
+expand it as an arbitrary order polynomial, in which case the argument still
+applies---limb darkening mixes the null space and the preimage in a nontrivial
+way. The fact that no coefficient can be determined uniquely---i.e., there are
+perfect degeneracies involving \emph{all} modes on the surface---is what makes it more
+difficult in practice to perform ensemble analyses on limb-darkened stars.
+Nevertheless, it is still possible to uniquely and precisely infer the spot properties
+under limb darkening,
+but as a rule of thumb, the stronger the limb darkening, the larger the
+ensemble must be (see Figure~\ref{fig:calibration_ld_1000}).
+
+In reality, it is unlikely that all stars in a given ensemble have exactly
+the same limb darkening coefficients, however ``similar'' the stars may be.
+The bottom panel of Figure~\ref{fig:nullspace_ensemble_ld} shows the same
+posterior shrinkage plot, but for the case where each star has coefficients
+$u_1 = 0.5 \pm 0.05$ and $u_2 = 0.25 \pm 0.025$; i.e., we add a scatter of
+10\% in the value of these coefficients. The plot shows the posterior
+shrinkage in the hypothetical case where we know the coefficients for
+each star exactly. Now, as the size of the ensemble increases, the posterior
+shrinkage approaches unity \emph{for all spherical harmonic modes}.
+In the same way that the null space is a strong function of the inclination,
+allowing us to chip away at it with observations at different inclinations,
+the null space is also a strong function of the limb darkening law. Even a
+small amount of variance in the coefficients is sufficient to constrain all surface
+modes exactly (in the limit of infinite SNR and large $M$).
+
+In practice, of course, we will never know the limb darkening coefficients
+exactly, so we must either model them or marginalize over them
+(see \S\ref{sec:other-marg}). This will dilute our ability to infer surface
+properties exactly.
 

### 2020-12-29 starry_process/latitude.py
adding tutorial

@@ -25,22 +25,22 @@ def gauss2beta(
     Args:
         mu (scalar or vector): The mean latitude in degrees.
         sigma (scalar or vector): The latitude standard deviation in degrees.
-        log_alpha_max (float, optional): The maximum value of ``log(alpha)``. 
+        log_alpha_max (float, optional): The maximum value of ``log(alpha)``.
             Default is %%defaults["log_alpha_max"]%%.
-        log_beta_max (float, optional): The maximum value of ``log(beta)``. 
+        log_beta_max (float, optional): The maximum value of ``log(beta)``.
             Default is %%defaults["log_alpha_max"]%%.
 
     The shape parameters ``a`` and ``b`` are related to the shape parameters of
     the Beta distribution in cosine latitude via the transformations
 
-        .. code-block::python 
-        
+        .. code-block::python
+
             alpha = exp(a * log_alpha_max)
             beta = exp(log(0.5) + b * (log_beta_max - log(0.5)))
 
 
-    .. note:: 
-        
+    .. note::
+
         This is a utility function that can accept and return
         either numeric values or tensors. If both `mu` and `sigma`
         are numeric quantities, the result will be a numeric
@@ -86,27 +86,27 @@ def beta2gauss(
 ):
     """
     Return the mode ``mu`` and standard deviation ``sigma`` of Laplace's
-    (Gaussian) approximation to the PDF of the latitude Beta distribution 
+    (Gaussian) approximation to the PDF of the latitude Beta distribution
     with shape parameters ``a`` and ``b``.
 
     Args:
         a (scalar or vector): Shape parameter.
         b (scalar or vector): Shape parameter.
-        log_alpha_max (float, optional): The maximum value of ``log(alpha)``. 
+        log_alpha_max (float, optional): The maximum value of ``log(alpha)``.
             Default is %%defaults["log_alpha_max"]%%.
-        log_beta_max (float, optional): The maximum value of ``log(beta)``. 
+        log_beta_max (float, optional): The maximum value of ``log(beta)``.
             Default is %%defaults["log_alpha_max"]%%.
 
     The shape parameters ``a`` and ``b`` are related to the shape parameters of
     the Beta distribution in cosine latitude via the transformations
 
-        .. code-block::python 
-        
+        .. code-block::python
+
             alpha = exp(a * log_alpha_max)
[…]

### 2020-12-29 tests/test_size.py

@@ -7,7 +7,7 @@ from theano.configparser import change_flags
 import pytest
 
 
-def test_size(ydeg=15, r=15.0, d=5.0, rtol=1e-7, ftol=1e-7, **kwargs):
+def test_size(ydeg=15, r=15.0, dr=5.0, rtol=1e-7, ftol=1e-7, **kwargs):
     # Indices of non-zero elements
     l = np.arange(ydeg + 1)
     i = l * (l + 1)
@@ -15,14 +15,14 @@ def test_size(ydeg=15, r=15.0, d=5.0, rtol=1e-7, ftol=1e-7, **kwargs):
 
     # Get analytic integrals
     print("Computing moments analytically...")
-    S = SizeIntegral(r, d, ydeg=ydeg, **kwargs)
+    S = SizeIntegral(r, dr, ydeg=ydeg, **kwargs)
     e = S._first_moment().eval()[i]
     eigE = S._second_moment().eval()
     E = (eigE @ eigE.T)[ij]
 
     # Get the first moment by numerical integration
     r_rad = r * np.pi / 180
-    d_rad = d * np.pi / 180
+    d_rad = dr * np.pi / 180
     e_num = np.zeros(ydeg + 1)
     Bp = S._spot.Bp.eval()
     theta = S._spot.theta.eval()
@@ -59,16 +59,16 @@ def test_size(ydeg=15, r=15.0, d=5.0, rtol=1e-7, ftol=1e-7, **kwargs):
 
 
 def test_size_grad(
-    ydeg=15, r=15.0, d=5.0, abs_tol=1e-5, rel_tol=1e-5, eps=1e-5,
+    ydeg=15, r=15.0, dr=5.0, abs_tol=1e-5, rel_tol=1e-5, eps=1e-5
 ):
     with change_flags(compute_test_value="off"):
 
-        S = SizeIntegral(r, d, ydeg=ydeg)
+        S = SizeIntegral(r, dr, ydeg=ydeg)
 
         # d/de
         verify_grad(
-            lambda r, d: SizeIntegral(r, d, ydeg=ydeg)._first_moment(),
-            (r, d),
+            lambda r, dr: SizeIntegral(r, dr, ydeg=ydeg)._first_moment(),
+            (r, dr),
             n_tests=1,
             abs_tol=abs_tol,
             rel_tol=rel_tol,
@@ -77,8 +77,8 @@ def test_size_grad(
 
         # d/dE
         verify_grad(
[…]

### 2020-12-29 tex/ms.tex
adding tutorial
docs fix; more text

@@
 noninteger values to add some flexibility to the model; we discuss this
-in more detail in \S\ref{sec:integerspots}
+in more detail in \S\ref{sec:caveats}
 
@@
 \emph{data} itself. To get around this, we either need to impose a
-shrinkage prior on $\sigma_\phi$ (\S\ref{sec:sigma-prior}),
+shrinkage prior on $\sigma_\phi$ (\S\ref{sec:caveats}),
 observe in a different mode that is sensitive to baseline information
@@
 \label{sec:tinyspots}
-
-\xxx{TODO: Discuss what we should do about tiny spots.}
+\xxx{TODO: The model is limited in resolution to features with radii $r \gtrsim 10^\circ$.
+    Typical sun spots are MUCH smaller than that. However, our model can still
+    correctly infer latitudinal parameters even when spots are really tiny
+    (i.e., no bias due to the limited resolution). We might try to improve the
+    numerical stability of the model at higher spherical harmonic degree, but
+    that's a task for a future paper.}
+
+\subsection{Bright spots}
+\label{sec:brightspots}
+\xxx{TODO: We can easily model bright spots and plages by setting $c < 0$. Show
+    an example. This model can be used to differentiate between spot-dominated
+    and plage-dominated surfaces, which is physically interesting. Cite
+    Brett's paper on TRAPPIST-1.}
 
@@
 \label{sec:breaking_baseline_degeneracy}
-
 \xxx{TODO: Discuss the baseline degeneracy and how to break it. Discuss the
-    degeneracy between $c$ and $N$ and how to break it.}
+    degeneracy between $c$ and $N$ and how to break it. The basic idea here
+    is that if we can constrain the (spotless) photospheric temperature, we
+    can uniquely infer $n$. I think this can be done by taking light curves
+    in two bands, but I'm checking with Gully. @HEDGES: Anything to add on this?}
 
@@
 \label{sec:other-marg}
-\xxx{TODO: We can also marginalize over period and limb darkening coefficients.}
+\xxx{TODO: We can also marginalize over period and limb darkening coefficients.
+    Argue that it's just integrals of Wigner matrices all over again, which we
+    can compute efficiently in closed form. I think we need a uniform prior on
+    the rotation \emph{frequency}, which is probably fine. And I think we can
+    parametrize limb darkening in terms of Kipping's $q_1$ and $q_2$ and place
+    a uniform prior in that space---I'm fairly confident the integrals can still
+    be solved exactly.}
+
+\subsection{Time evolution}
+\label{sec:temporal}
+\xxx{TODO: The model presented in this paper assumes statics stellar surfaces,
+    corresponding to perfectly periodic light curves. This is hardly ever the
+    case in practice, as spots evolve and differential rotation changes the surface
+    over time. We can still use this model, but we need to either restrict our analysis
+    to segments of light curves of duration shorter than the timescale on which
+    the surface evolves, or explicitly model the effect of time evolution on the covariance.
+    The next paper will deal with temporal evolution more
+    rigorously, but the way to do this is just take the elementwise product of the \starryprocess
+    flux covariance with a temporal covariance. This corresponds to the Kronecker product
+    in spherical harmonic space. Show a figure, give an example, etc.}
+
+\subsection{Application to transits and RV}
+\label{sec:other-work}
+\xxx{TODO: The covariance in the spherical harmonic basis can be used as a prior
+    in many other settings.}
+
+\subsection{Comparison to other work}
+\label{sec:other-work}
+\xxx{TODO: Discuss \citet{Basri2020}, how we agree and disagree. Compare to Brett's
+    paper using ABC to measure spot sizes as a function of spectral type. Discuss
+    Simon Wood's paper in ecology on synthetic likelihoods. Talk about similarities
+    to Gaussian Random Fields in CMB studies.}
 
@@
 \label{sec:caveats}
+\xxx{TODO: Miscellaneous caveats.}
 \xxx{
     \begin{enumerate}
-        \item The latitude distribution is not Gaussian.
-        \item The number of spots does not have to be an integer.
+        \item The latitude distribution is not Gaussian, so careful
+              when interpreting $\mu_\phi$ and $\sigma_\phi$
+        \item The number of spots does not have to be an integer. This is
+              fine, since it makes the model more flexible.
         \item Careful with light curves with large amplitudes, although
@@
               has max amplitudes of more than 10\% and works great.
-        \item No differential rotation / spot evolution yet!
+        \item No differential rotation / spot evolution yet! We're working on it.
+        \item We might want a shrinkage prior on sigma, since it's hard to
+              break the polar spot degeneracy with single-band light curves.
     \end{enumerate}
 }
-
-\subsection{Other distributions}
-\label{sec:other-dist}
-\xxx{TODO: Discuss the uniform spot radius distribution.}
-
-\subsection{Prior on sigma}
-\label{sec:sigma-prior}
-\xxx{TODO: We might want a shrinkage prior on sigma.}
-
-\subsection{Inferring inclinations}
-\xxx{TODO: Works great. Note that inclination distribution is sometimes bimodal!}
 
@@
 \Python package \starryprocess, which can be installed via \textsf{pip}\xspace
-or from source on GitHub\footnote{\url{https://github.com/rodluger/starry_process}}.
+or from source on \href{https://github.com/rodluger/starry_process}{GitHub}.
 The code is thoroughly unit-tested and well documented, with examples on how
@@
 although an implementation of this is deferred to the next paper.
-\xxx{DFM: Is this true? I.e., is there an efficient way to solve Equation~(\ref{eq:log-like})
+\xxx{@DFM: Is this true? I.e., is there an efficient way to solve Equation~(\ref{eq:log-like})
     when $\pmb{\Sigma}$ is the same for all stars but $\mathbf{C}$ is different? Is there
@@
 \label{sec:conclusions}
-\xxx{TODO: Summarize. Discuss future work, including differential rotation
+We presented a new Gaussian process (GP) model for stellar variability
+whose hyperparameters explicitly correspond to physical properties of the stellar
+surface. Our GP allows one to efficiently compute the likelihood function
+for stellar light curves
+marginalized over nuisance parameters such as the specific sizes,
+positions, and contrasts of individual spots,
+which are generally unknowable due to the extreme degeneracies
+involved in the light curve mapping problem. Our GP therefore
+makes it easy to
+do posterior inference on the real quantities of interest: parameters
+controlling the \emph{distribution} of spot sizes, latitudes, and
+contrasts within a star and/or across many stars in an ensemble.
+
+Because our expression for the GP covariance has an exact, closed-form
+solution as a function of the spot parameters, it can be computed
+extremely efficiently: a typical likelihood evaluation on a dataset
+consisting of $K \sim 1{,}000$ points takes about 10ms on a modern
+laptop. Our algorithm is implemented in the open-source, user-friendly
+\Python package \starryprocess, which is \textsf{pip}-installable and available on
+\href{https://github.com/rodluger/starry_process}{GitHub}.
+In keeping with other papers in the \starry series, all figures in this
+paper are generated automatically from open-source scripts linked to in
+each of the captions \codeicon, and the principal equations link to associated
+unit tests that ensure the accuracy and reproducibility of the algorithm
+presented here \testpassicon/\testfailicon.
+
+We devoted a large portion of this paper to testing the algorithm on a variety
+of synthetic datasets, showing that it is a well-calibrated and in most
+cases unbiased estimator for star spot properties. Below we list the main
+results presented in this paper:
+
+\begin{enumerate}
+    \item \textbf{Our GP works best for ensemble analyses.} The light curve mapping
+          problem is extremely degenerate, as light curves contain a vanishingly small
+          fraction of the total information about a stellar surface. However, the
+          degenerate surface modes are a strong function of the observer's viewing
+          angle, so light curves of stars seen at different inclinations constrain
+          different components of the surface. We have shown that if we jointly
+          analyze the light curves of many stars,
+          we can break many of the degeneracies at play and uniquely infer the
+          statistical properties of the spots across the ensemble. This type
+          of analysis works best if the stars in the ensemble are statistically
+          similar: i.e., the properties of their spots are all drawn from the
+          same parent distribution, whose parameters we can constrain.
+    \item \textbf{In general, an ensemble of at least $M \sim 50$ light curves
+              is needed to place meaningful constraints on the spot properties}.
+          This estimate is based on ensemble analyses of light curves with
+          $K = 1{,}000$ cadence each and per-cadence precision of one part
+          per thousand. Lower quality, shorter baseline observations, or
+          datasets contaminated by outliers will in general require larger values
+          of $M$ for the same constraining power. The presence of strong limb
+          darkening also degrades the information content of light curves, in
+          which case an ensemble of hundreds or even a thousand light curves
+          is recommended.
+    \item \textbf{Our GP is an unbiased estimator for the spot radius and
+              latitude distributions.} We showed that our GP can accurately infer the
+          angular size of spots and the mode and standard deviation of their
+          distribution in latitude from stellar light curves. For the fiducial
+          ensemble of $M = 50$ light curves described above, we are able to
+          constrain the average spot radii to within a couple degrees and the
+          average spot latitudes to within $5^\circ$.
+    \item \textbf{Relative single-band photometry cannot uniquely constrain
+              the spot contrast or the total number of spots}. These two quantities
+          are degnerate with each other. While the contrast can be very weakly
+          constrained with lots of data, the total number of spots is not
+          an observable in relative single-band photometry. Knowledge of the
+          unspotted photospheric brightness (or temperate) can break this
+          degeneracy, so analyses of multi-band light curves can in principle
+          infer the total spottiness of a stellar surface.
+    \item \textbf{Our GP can accurately infer stellar inclinations.} We presented
+          two versions of our GP: one conditioned on a specific value of the stellar
+          inclination, and one marginalized over inclination under an isotropic
+          prior. In both cases, we find that we can accurately infer the inclinations
+          of individual stars in an ensemble analysis (in the latter case, a
+          simple post-processing step can yield the inclination posterior distribution).
+          While the inclination is not an observable for an individual stellar light
+          curve, the population-level constraints on the spot properties achieved
+          by the GP can break the degeneracies involving the inclination, allowing
+          us to usually infer it to within about $10^\circ$ and without bias.
+    \item \textbf{Our GP can be used to model small, Sun-like spots.} The algorithm
+          presented here is limited to a surface resolution of about $10^\circ$,
+          corresponding to spots about an order of magnitude larger than typical
+          Sun spots. However, we have showed that when we apply our model to
+          light curves of stars with small ($r \sim 3^\circ$) spots, we can
+          still infer their latitudinal distribution without bias.
+    \item \textbf{Our GP can be extended to model time-variable surfaces.} The
+          algorithm presented here was derived for static stellar surfaces,
+          corresponding to perfectly periodic light curves. \xxx{...}
+    \item \textbf{Possible applications.} Spot-dominated versus plage-dominated
+\end{enumerate}
+
+\xxx{TODO: Discuss future work, including differential rotation
     and spot evolution.}

### 2020-12-29 tex/ms.tex
finished conclusions

@@
 \href{https://github.com/rodluger/starry_process}{GitHub}.
-In keeping with other papers in the \starry series, all figures in this
-paper are generated automatically from open-source scripts linked to in
-each of the captions \codeicon, and the principal equations link to associated
-unit tests that ensure the accuracy and reproducibility of the algorithm
-presented here \testpassicon/\testfailicon.
+The algorithm is implemented in a combination of \cpp and \Python,
+linked using the \theano package. Because our GP covariance has an exact
+representation, so too do its derivatives. We therefore implement
+backpropagated derivatives with respect to all input parameters for
+out-of-the-box usage with gradient-based inference and optimization
+tools such as Hamiltionain Monte Carlo (HMC), autodifferentation
+variational inference (ADVI), and gradient-based nested sampling.
 
@@
           light curves of stars with small ($r \sim 3^\circ$) spots, we can
-          still infer their latitudinal distribution without bias.
+          still infer their latitudinal distribution without bias, as well as
+          the presence of spots below our resolution limit.
     \item \textbf{Our GP can be extended to model time-variable surfaces.} The
           algorithm presented here was derived for static stellar surfaces,
-          corresponding to perfectly periodic light curves. \xxx{...}
-    \item \textbf{Possible applications.} Spot-dominated versus plage-dominated
+          corresponding to perfectly periodic light curves. However, time
+          variability can easily be modeled as the product of the \starryprocess
+          kernel and a kernel describing the covariance of the process in
+          time, such as a simple squared exponential or Mat\'ern-3/2 kernel.
+          The hyperparameters of the temporal covariance are then strictly
+          tied to the timescale on which the surface evolves. More
+          complex temporal variability, such as that induced by differential
+          rotation, will be the subject of the next paper in this series.
+    \item \textbf{Our GP can be used in exoplanet transit modeling and
+              radial velocity datasets.} At its core, the \starryprocess GP
+          defines a distribution over spherical harmonic representations
+          of stellar surfaces. It can therefore be used as a physically
+          interpretable prior when modeling transits of exoplanets across
+          spotted stars, either to marginalize over the stellar inhomogeneity
+          or to explicitly infer spot properties. When combined with the
+          \starry package, it can also be used as a
+          prior on the stellar surface in Doppler imaging, Doppler tomography,
+          or even radial velocity searches for exoplanets.
 \end{enumerate}
 
-\xxx{TODO: Discuss future work, including differential rotation
-    and spot evolution.}
+The GP presented here has far-ranging applications for stellar light
+curve studies. It serves as a drop-in replacement for commonly used
+GP kernels for stellar variability, which currently do not have
+physically interpretable parameters other than (perhaps) the rotation
+period. As such, it can be used to marginalize over stellar rotational variability
+signals in (say) transiting exoplanet searches, asteroseismic characterization
+of stars, radial velocity searches, etc. It can also be used to learn
+about stellar surfaces directly: to infer spot properties of main sequence
+stars as a function of spectral type and age, to differentiate between
+spot-dominated and plage-dominated stellar surfaces, to better understand
+chemically peculiar massive stars, and to better understand the spot
+properties of transiting exoplanet hosts for unbiased spectroscopic
+characterization of their atmospheres (to name a few).
+
+The next papers in this series will focus on (II) a more rigorous
+treatment of time variability and (III) explicit marginalization over
+the remaining stellar parameters. In particular, paper II will focus
+on modeling differential rotation, whose effect on the covariance of the
+process can be derived in a similar fashion to what we did here. This
+will enable direct inference about the differential rotation rates and
+spot evolution timescales of stars, processes whose effects on light curves
+are too similar for current methodology to reliably discern between them.
+Paper III will focus on marginalization over the stellar rotation period
+and limb darkening coefficients, which can also be done analytically
+under certain choices of prior. This will eliminate all per-star
+hyperparameters in the expression for the GP covariance, greatly
+speeding up inference for large ensembles of stellar light curves.
+
+In keeping with other papers in the \starry series, all figures in this
+paper are generated automatically from open-source scripts linked to in
+each of the captions \codeicon, and the principal equations link to associated
+unit tests that ensure the accuracy and reproducibility of the algorithm
+presented here \testpassicon/\testfailicon.
 
@@
     \parenMatrixstack{
-    0              & 5 \\
-    \ln\frac{1}{2} & 5
+    0              & 10 \\
+    \ln\frac{1}{2} & 10
     }

### 2020-12-30 tex/ms.tex
experimental time variability

@@
 Among a vast number of kernels are a few that are commonly used
-in modeling stellar variability: the squared exponential kernel,
+in modeling stellar variability: the exponential squared kernel,
 %
@@
     in two bands, but I'm checking with Gully. @HEDGES: Anything to add on this?}
+
+\subsection{Mixture models}
+\label{sec:mixture}
+\xxx{TODO: kernels can be summed!}
 
@@
           kernel and a kernel describing the covariance of the process in
-          time, such as a simple squared exponential or Mat\'ern-3/2 kernel.
+          time, such as a simple exponential squared or Mat\'ern-3/2 kernel.
           The hyperparameters of the temporal covariance are then strictly

### 2020-12-30 starry_process/defaults.py

@@ -1,3 +1,5 @@
+from .temporal import *
+
 defaults = dict(
     ydeg=15,
     udeg=2,
@@ -10,6 +12,8 @@ defaults = dict(
     p=1.0,
     i=60.0,
     u=None,
+    tau=None,
+    temporal_kernel=Matern32Kernel,
     normalized=True,
     normalization_order=20,
     normalization_zmax=0.023,
@@ -17,7 +21,7 @@ defaults = dict(
     baseline_mean=0.0,
     baseline_var=0.0,
     driver="numpy",
-    eps=1e-6,
+    eps=1e-8,
     epsy=1e-12,
     epsy15=1e-9,
     covpts=300,

### 2020-12-30 starry_process/flux.py

@@ -11,7 +11,6 @@ from .ops import (
 from .wigner import R
 from .defaults import defaults
 from .math import cast
-from .cache import cache
 from scipy.special import gamma, hyp2f1
 import numpy as np
 import theano
@@ -126,9 +125,9 @@ class FluxIntegral:
     def _G(self, j, i):
         """
         This is the integral of
-         
+
             cos(x / 2)^i sin(x / 2)^j sin(x)
-        
+
         from 0 to pi/2.
         """
         return 2 * gamma(1 + 0.5 * i) * gamma(1 + 0.5 * j) / gamma(
@@ -250,7 +249,7 @@ class FluxIntegral:
     def _precompute(self):
         """
         Pre-compute some vectors and matrices.
-        
+
         """
         # Pre-compute some matrices
         self._precompute_inclination_integrals()

### 2020-12-30 starry_process/temporal.py

@@ -1,0 +1,16 @@
+import theano.tensor as tt
+import numpy as np
+
+
+__all__ = ["ExpSquaredKernel", "Matern32Kernel"]
+
+
+def ExpSquaredKernel(t1, t2, tau):
+    dt = tt.abs_(tt.reshape(t1, (-1, 1)) - tt.reshape(t2, (1, -1)))
+    return tt.exp(-dt / (2 * tau))
+
+
+def Matern32Kernel(t1, t2, tau):
+    dt = tt.abs_(tt.reshape(t1, (-1, 1)) - tt.reshape(t2, (1, -1)))
+    x = np.sqrt(3) * dt / tau
+    return (1 + x) * tt.exp(-x)

### 2020-12-30 tex/ms.tex
more figures, new test

@@
 \begin{align}
-    k_\mathrm{E^2}(\Delta t)           & = \sigma^2 \exp\left(-\frac{\Delta t}{2\tau}\right)
+    k_\mathrm{E^2}(\Delta t)           & = \sigma^2 \exp\left(-\frac{\Delta t^2}{2\tau}\right)
     \quad,                                                                                                                                           \\
@@
 
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/gpsum.pdf}
+        \oscaption{gpsum}{%
+            Prior samples from a sum of two \starryprocess models. The
+            first model consists of small ($r=10^\circ$)
+            circumpolar ($\phi = 60^\circ \pm 3^\circ$) spots and the
+            second model consists of larger ($r=20^\circ$)
+            equatorial ($\phi = 0^\circ \pm 3^\circ$) spots.
+            The sum of two \starryprocess models is also a
+            \starryprocess, making it easy to model more complex
+            distributions of spots.
+            \label{fig:gpsum}
+        }
+    \end{centering}
+\end{figure}
+
 \subsection{Gaussians all the way down}
@@
     in spherical harmonic space. Show a figure, give an example, etc.}
+
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/temporal.pdf}
+        \oscaption{temporal}{%
+            Prior sample from a time-variable \starryprocess, assuming a rotation
+            period $P = 1$ day and a temporal evolution timescale $\tau = 20$ days
+            modeled with an exponential squared kernel. Evolving map samples are
+            shown at the top, with the corresponding light curve viewed at an
+            inclination $I = 60^\circ$ at the bottom. The GP hyperparameters are
+            set to their default values (Table~\ref{tab:synthetic}).
+            \label{fig:temporal}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{proof}{test_temporal}
+    \pmb{\mathcal{A}}^\ddagger
+    \left(\pmb{\Sigma}_\mathbf{t} \otimes \pmb{\Sigma}_\mathbf{y}\right)
+    {\pmb{\mathcal{A}}^\ddagger}^\top
+    &=
+    \pmb{\Sigma} \odot \pmb{\Sigma}_\mathbf{t}
+\end{proof}
+%
+where
+%
+\begin{align}
+    \pmb{\mathcal{A}}^\ddagger
+                      & \equiv
+    \setstackgap{L}{1.25\baselineskip}
+    \fixTABwidth{T}
+    \parenMatrixstack{
+    \mathbf{a}_0^\top &                   &        &                              \\
+                      & \mathbf{a}_1^\top &        &                              \\
+                      &                   & \ddots &                              \\
+                      &                   &        & \,\,\, \mathbf{a}_{K-1}^\top
+    }\quad.
+\end{align}
 

### 2020-12-30 starry_process/sp.py

@@ -518,11 +518,6 @@ class StarryProcess(object):
                 %%defaults["baseline_var"]%%.
             nsamples (int, optional): The number of samples to draw. Default 1.
 
-        .. note::
-
-            This method returns static map samples only (i.e., any time variability
-            is ignored).
-
         """
         # TODO
         if self._marginalize_over_inclination:
@@ -534,6 +529,12 @@ class StarryProcess(object):
         if self._normalized:
             raise NotImplementedError(
                 "Not yet implemented when the flux is normalized."
+            )
+
+        # TODO
+        if self._time_variable:
+            raise NotImplementedError(
+                "Not yet implemented for time-variable maps."
             )
 
         # Get the full data covariance

### 2020-12-30 starry_process/temporal.py

@@ -7,7 +7,7 @@ __all__ = ["ExpSquaredKernel", "Matern32Kernel"]
 
 def ExpSquaredKernel(t1, t2, tau):
     dt = tt.abs_(tt.reshape(t1, (-1, 1)) - tt.reshape(t2, (1, -1)))
-    return tt.exp(-dt / (2 * tau))
+    return tt.exp(-(dt ** 2) / (2 * tau))
 
 
 def Matern32Kernel(t1, t2, tau):

### 2021-01-07 starry_process/ops/__init__.py
add viz option

@@ -5,3 +5,4 @@ from .exceptions import CheckBoundsOp, CheckVectorSizeOp
 from .eigh import EighOp
 from .norm import AlphaBetaOp
 from .sample import SampleYlmTemporalOp
+from .poly import pTA1Op

### 2021-01-07 starry_process/ops/poly/__init__.py

@@ -1,0 +1,2 @@
+# -*- coding: utf-8 -*-
+from .pTA1 import pTA1Op

### 2021-01-07 starry_process/ops/poly/pTA1.cc

@@ -1,0 +1,83 @@
+#section support_code_struct
+
+// Shorthand
+#define DI0 DTYPE_INPUT_0
+#define DI1 DTYPE_INPUT_1
+#define DI2 DTYPE_INPUT_2
+#define DO0 DTYPE_OUTPUT_0
+#define TO0 TYPENUM_OUTPUT_0
+
+Eigen::SparseMatrix<DO0> *APPLY_SPECIFIC(A1);
+
+#section init_code_struct
+
+{ APPLY_SPECIFIC(A1) = NULL; }
+
+#section cleanup_code_struct
+
+if (APPLY_SPECIFIC(A1) != NULL) {
+  delete APPLY_SPECIFIC(A1);
+}
+
+#section support_code_struct
+
+int APPLY_SPECIFIC(pTA1)(PyArrayObject *input0, // x
+                         PyArrayObject *input1, // y
+                         PyArrayObject *input2, // z
+                         PyArrayObject **output0) {
+
+  using namespace sp::theano;
+  using namespace sp::flux;
+  using namespace sp::utils;
+
+  // Get the inputs
+  int success = 0;
+  int ndim = -1;
+  npy_intp *shape;
+  auto x_in = get_input<DI0>(&ndim, &shape, input0, &success);
+  if (ndim != 1) {
+    PyErr_Format(PyExc_ValueError, "x must be a vector");
+    return 1;
+  }
+  auto y_in = get_input<DI1>(&ndim, &shape, input1, &success);
+  if (ndim != 1) {
+    PyErr_Format(PyExc_ValueError, "y must be a vector");
+    return 1;
+  }
+  auto z_in = get_input<DI2>(&ndim, &shape, input2, &success);
+  if (ndim != 1) {
+    PyErr_Format(PyExc_ValueError, "z must be a vector");
+    return 1;
+  }
+  int K = shape[0];
+  Map<RowVector<DI0, Dynamic>> x(x_in, K);
+  Map<RowVector<DI0, Dynamic>> y(y_in, K);
+  Map<RowVector<DI0, Dynamic>> z(z_in, K);
+
+  // Allocate the outputs
+  success = 0;
+  ndim = 2;
+  std::vector<npy_intp> shape_vec(ndim);
+  shape_vec[0] = K;
+  shape_vec[1] = SP__N;
[…]

### 2021-01-08 tex/ms.tex
bugfixes
add viz option
figure tweak

@@
 \label{sec:intro}
+
+% CHECK OUT
+% Louden 2020 https://ui.adsabs.harvard.edu/abs/2020arXiv201200776L/abstract
+% Mazeh 2015
 
@@
 modeling the baseline in each light curve as a latent variable, as discussed
-above.
+above. \xxx{NOTE: I'm actually no long sure this would work! DFM, let's chat about this.}
 
@@
 likelihood evaluation typically scales sub-linearly with $M$. Furthermore,
-as we discussed in \ref{sec:other-marg}, it is possible to marginalize over
+as we discussed in \S\ref{sec:other-marg}, it is possible to marginalize over
 the period and limb darkening coefficients, which would remove the scaling

### 2021-01-08 tex/figures/speed.py

@@ -43,9 +43,9 @@ ax.plot(npts, tloglike[1], label="marginal")
 
 # Asymptotic
 n = np.array([2e2, 1.0e4])
-ax.plot(n, 3e-10 * n ** 2.5, "k--", alpha=0.5)
+ax.plot(n, 1.2e-10 * n ** 2.6, "k--", alpha=0.5)
 ax.annotate(
-    r"$\propto K^{2.5}$",
+    r"$\propto K^{2.6}$",
     fontsize=16,
     xy=(2e3, 1e-2),
     xycoords="data",

### 2021-01-11 tex/ms.tex
new nb

@@
     \textbf{
-        Interpretable Gaussian Processes for Stellar Variability\\
-        I: Static Surfaces
+        Interpretable Gaussian Processes for Stellar Variability
     }
@@
     The use of Gaussian processes (GPs) in stellar light curve modeling has recently
-    become almost ubiquitous, given their ease of use, flexibility, and
-    computational efficiency. GPs excel in particular at marginalization
+    become almost ubiquitous, given their ease of use and flexibility.
+    GPs excel in particular at marginalization
     of the stellar signal in cases where the stellar variability is treated
@@
     properties via optimization or posterior inference, a process that is
-    ill-posed, ridden by degeneracies, and computationally intractable when
+    ill-posed, plagued by degeneracies, and computationally intractable when
     applied to stars with more than a few spots and/or to ensembles of many
@@
     In this paper, we derive a closed-form expression for the
-    mean and covariance of the Gaussian process that best describes
-    the light curve of a rotating star with a given distribution of
+    mean and covariance of a Gaussian process model that describes
+    the light curve of a rotating star conditioned on a given distribution of
     star spot sizes, contrasts, and latitudes.
@@
 hyperparameters $\pmb{\theta}_\bullet$. For larger values of $K$, the
-cost approaches a scaling of $K^{2.5}$, which is dominated by the factorization
+cost approaches a scaling of $K^{2.6}$, which is dominated by the factorization
 of the covariance matrix and the solve operation to compute the likelihood.

### 2021-01-11 starry_process/sp.py

@@ -3,9 +3,9 @@ from .latitude import LatitudeIntegral, gauss2beta, beta2gauss
 from .longitude import LongitudeIntegral
 from .contrast import ContrastIntegral
 from .flux import FluxIntegral
-from .math import cho_factor, cho_solve, cast
+from .math import cho_factor, cho_solve, cast, is_tensor
 from .defaults import defaults
-from .visualize import mollweide_transform, visualize
+from .visualize import mollweide_transform, latlon_transform, visualize
 from .ops import CheckBoundsOp, AlphaBetaOp, SampleYlmTemporalOp
 import theano.tensor as tt
 from theano.ifelse import ifelse
@@ -447,6 +447,52 @@ class StarryProcess(object):
 
         """
         return self._cho_cov_ylm
+
+    def mean_pix(self, latlon):
+        """
+        The mean of the process in latitude-longitude pixel space.
+
+        Args:
+            latlon (ndarray): The latitude(s) and longitude(s) at which to
+                evaluate the mean in degrees. This must be a numerical
+                quantity (tensor inputs are not supported) of shape
+                ``(..., 2)``, where the last dimension corresponds to
+                a tuple of latitude-longitude values.
+
+        Returns:
+            The tensor-valued mean evaluated at the flattened vector of
+            input coordinates.
+            This will have shape ``(N,)`` where ``N`` is the number of
+            input lat-lon coordinates.
+
+        """
+        assert not is_tensor(latlon), "Input must be a numerical ndarray."
+        lat, lon = latlon.reshape(-1, 2).T
+        A = latlon_transform(lat * np.pi / 180, lon * np.pi / 180)
+        return tt.dot(A, self._mean_ylm)
+
+    def cov_pix(self, latlon):
+        """
+        The covariance of the process in latitude-longitude pixel space.
+
+        Args:
+            latlon (ndarray): The latitude(s) and longitude(s) at which to
+                evaluate the covariance in degrees. This must be a numerical
+                quantity (tensor inputs are not supported) of shape
+                ``(..., 2)``, where the last dimension corresponds to
+                a tuple of latitude-longitude values.
+
+        Returns:
+            The tensor-valued covariance evaluated at the flattened vector of
+            input coordinates.
+            This will have shape ``(N, N)`` where ``N`` is the number of
+            input lat-lon coordinates.
+
+        """
[…]

### 2021-01-11 tex/ms.tex
more text in discussion

@@
 A much safer approach is to resist the temptation to normalize the light curve
-and instead model the unknown baseline as a latent variable. However,
+and instead model the unknown multiplicative amplitude as a latent variable. However,
 this would require an extra parameter \emph{for every light curve}, so the
@@
 In cases where the light curve variability exceeds a few percent, we recommend
-modeling the baseline in each light curve as a latent variable, as discussed
-above. \xxx{NOTE: I'm actually no long sure this would work! DFM, let's chat about this.}
+modeling the multiplicative amplitude in each light curve as a latent variable, as discussed
+above.
 
@@
 
-\subsection{Mixture models}
+\subsection{Composite GPs}
 \label{sec:mixture}
-\xxx{TODO: kernels can be summed!}
+Thus far we have assumed that spots are concentrated at a single latitude
+(above and below the equator). We baked this assumption directly into our choice of
+distribution function for the latitude (\S\ref{sec:lat}), which has exactly
+two modes at $\pm \mu_\sigma$. However, it is possible (at least in principle)
+that certain stars could have two or more active latitudes, in which case
+our GP is not an appropriate description of the stellar surface.
 
@@
 
-\subsection{Gaussians all the way down}
-\label{sec:other-marg}
-\xxx{TODO: We can also marginalize over period and limb darkening coefficients.
-    Argue that it's just integrals of Wigner matrices all over again, which we
-    can compute efficiently in closed form. I think we need a uniform prior on
-    the rotation \emph{frequency}, which is probably fine. And I think we can
-    parametrize limb darkening in terms of Kipping's $q_1$ and $q_2$ and place
-    a uniform prior in that space---I'm fairly confident the integrals can still
-    be solved exactly.}
+Fortunately, Gaussian distributions (and thus also Gaussian processes) are
+closed under addition, meaning that the sum of two GPs is also a GP. We can thus
+construct more complex models for stellar variability by summing GPs with different
+spot hyperparameter vectors $\pmb{\theta}_\bullet$. The composite GP will then
+have a mean equal to the sum of the means of each GP, and a covariance matrix equal
+to the sum of the covariance matrices of each GP. One possible application of this
+is to model stars with multiple active latitudes, as described above; an example
+of this is shown in Figure~\ref{fig:gpsum}, where samples are drawn from a GP with
+small circumpolar spots and large equatorial spots (see caption for details).
+Since the composite GP inherits all of the properties of the standard GP, it can
+be used to do inference under more complex priors than those presented here.
+
+It is also worth noting that this technique may be employed to model arbitrary
+distributions for parameters like the spot radius and the number of spots. In
+the Appendix we present a formulation of our GP that admits a radius distribution
+half-width parameter $\Delta r$; this generalizes our delta function distribution to
+a uniform distribution between $r - \Delta r$ and $r + \Delta r$. One may then
+compute the weighted sum of several GPs
+with half-widths $\Delta \ne 0$ and central radii
+$r = r_0$,
+$r = r_0 + 2\Delta r$,
+$r = r_0 + 4\Delta r$,
+etc., to approximate \emph{any} distribution of spot radii. Similarly, one
+may compute the weighted sum of several GPs with different values of the
+number of spots $n$ to enforce any discrete distribution for that quantity.
+The reader should keep in mind that the cost of computing the GP
+covariance matrix will scale linearly with the number of GP components.
+In many cases, however, the computational bottleneck is the covariance factorization
+step (\S\ref{sec:implementation}), in which case adding components to the
+model will result in negligible overhead.
 
@@
 \end{align}
+
+\subsection{Gaussians all the way down}
+\label{sec:other-marg}
+
+In \S\ref{sec:inclination} we discussed the value of marginalizing over the
+stellar inclination when computing the GP covariance (and thus the likelihood).
+When jointly analyzing the light curves of many stars (which we have repeatedly
+argued is the best way to infer their spot properties), it can be extremely useful to
+minimize the number of latent variables associated with individual stars by
+analytically marginalizing over them. This can dramatically reduce the number
+of free parameters, turning a difficult inference problem in (possibly) hundreds
+or thousands of dimensions into a much easier problem in a handful of dimensions.
+We showed how it is possible to marginalize away the dependence of our GP on
+the inclinations of individual stars, which is a huge step in this direction.
+However, our GP remains a function of two other quantities that will in general
+be different for different stars: the stellar rotation period $P$ and the
+limb darkening coefficient vector $\mathbf{u}$. We argued that in some cases
+one may be able to fix the period of each star at an estimate obtained in a pre-processing
+step (i.e., from a periodogram) and fix the limb darkening coefficients at
+theoretical values, in which case the number of free parameters of the GP
+is equal to the size of the spot hyperparameter vector $\pmb{\theta}_\bullet$
+(five by default), and \emph{independent} of the number of light curves in
+the ensemble.
+%
+However, this procedure ignores any uncertainty in the period and limb darkening
+coefficients, which could be significant; it is also subject to bias due to
+the fact that theoretical models for limb darkening coefficients have been
+shown to be wrong, especially for low mass stars \citep{?}. A much better
+approach would be to analytically marginalize over these two quantities.
+
+Consider Equation~(\ref{eq:akT}) in the Appendix, from which the rows of the
+design matrix (which transforms vectors in the spherical harmonic basis to vectors
+representing the flux at a point in time) are computed. Marginalization over the
+inclination, which we demonstrated how to perform analytically in
+Appendix~\ref{sec:inc}, entails integrating over the term
+$\mathbf{R}_{\hat{\mathbf{x}}}\left(-I\right)$, the Wigner matrix that rotates
+the star by the inclination angle $I$ into the observer's frame.
+Similarly, marginalization over the
+rotation period would entail integration over the term
+$\mathbf{R}_{\hat{\mathbf{z}}}\left(\frac{2\pi}{P} t_k\right)$, another
+Wigner matrix that rotates the star to the correct rotational phase at
+time $t = t_k$. Given the appropriate prior on the rotational angular
+frequency $\frac{2\pi}{P}$, it should be possible to follow the same
+procedure outlined in Appendix~\ref{sec:inc} to analytically
+compute the first two moments of the distribution of light curves marginalized over the
+rotation period.
+
+Equation~(\ref{eq:akT}) also makes clear the dependence of the GP on
+the limb darkening cofficients, which enter via the limb darkening
+operator $\mathbf{L}(\mathbf{u})$. The coefficients
+of this matrix are linear in the limb darkening coefficients $u_1$ and $u_2$
+(c.f. Equation~\ref{eq:ld:L}), so it should be possible to derive closed
+form solutions to the relevant integrals over $\mathbf{L}(\mathbf{u})$ to
+yield the GP covariance marginalized over $\mathbf{u}$. It may be
+possible to compute the marginalized covariance even when parametrizing
+the quadratic limb darkening
+coefficients in terms of the uncorrelated $q_1$ and $q_2$ parameters from
+\citet{?}, which would allow us to incorporate hard constraints on positivity
+and strict limb darkening (as opposed to brightening) directly into the prior.
+
+Given the complexity of the operations involved,
+performing these marginalizations is beyond the scope of the present paper.
+However, given the significant computational benefits of this marginalization,
+and given the fact that the mapping problem is particularly sensitive to
+the limb darkening coefficients (see \S\ref{sec:limbdark}), marginalizing over
+$P$ and $\mathbf{u}$ will be the subject of the next paper in this series.
 

### 2021-01-12 tex/ms.tex
new notebook; more text

@@
             of 1 (blue), 3 (orange), 10 (green), and 30 (red) \emph{identical} stars
-            viewed at random inclinations.
+            viewed at random orientations.
             The thicker curves correspond to
@@
             inclinations; the plot in the center is the portion of the
-            surface that can be inferred at some value of the inclination.
+            surface that can be inferred by combining observations
+            made from all possible inclinations.
             \label{fig:evenmodes}
@@
             but for limb-darkened
-            stars with coefficients $u_1 = 0.5$ and $u_2 = 0.25$. Odd
+            stars with quadratic coefficients $u_1 = 0.5$ and $u_2 = 0.25$. Odd
             modes can now be probed, at the expense of the even modes.
@@
             across different observations. For $\gtrsim 30$ light curves,
-            there is virtually no null space up to $l_\mathrm{max} = 15$.
+            there is virtually no null space up to at least $l_\mathrm{max} = 15$.
             \codelink{nullspace_ensemble_ld_var}
@@
 
+\subsection{Application to transits and RV}
+\label{sec:other-work}
+\xxx{TODO: The covariance in the spherical harmonic basis can be used as a prior
+    in many other settings.}
+
+\subsection{Comparison to other work}
+\label{sec:other-work}
+\xxx{TODO: Discuss \citet{Basri2020}, how we agree and disagree. Compare to Brett's
+    paper using ABC to measure spot sizes as a function of spectral type. Discuss
+    Simon Wood's paper in ecology on synthetic likelihoods. Talk about similarities
+    to Gaussian Random Fields in CMB studies.}
+
+\subsection{Caveats}
+\label{sec:caveats}
+\xxx{TODO: Miscellaneous caveats.}
+\xxx{
+    \begin{enumerate}
+        \item The latitude distribution is not Gaussian, so careful
+              when interpreting $\mu_\phi$ and $\sigma_\phi$
+        \item The number of spots does not have to be an integer. This is
+              fine, since it makes the model more flexible.
+        \item Careful with light curves with large amplitudes, although
+              Figure~\ref{fig:calibration_hicontrast}
+              has max amplitudes of more than 10\% and works great.
+        \item No differential rotation / spot evolution yet! We're working on it.
+        \item We might want a shrinkage prior on sigma, since it's hard to
+              break the polar spot degeneracy with single-band light curves.
+    \end{enumerate}
+}
+
+\section{Extensions}
+\label{sec:extensions}
+
 \subsection{Composite GPs}
@@
 \label{sec:temporal}
+
+Another limitation of the base algorithm...
+
 \xxx{TODO: The model presented in this paper assumes statics stellar surfaces,
@@
 
-\subsection{Gaussians all the way down}
+\subsection{Marginalizing over period and limb darkening}
 \label{sec:other-marg}
@@
 the limb darkening coefficients (see \S\ref{sec:limbdark}), marginalizing over
-$P$ and $\mathbf{u}$ will be the subject of the next paper in this series.
-
-\subsection{Application to transits and RV}
-\label{sec:other-work}
-\xxx{TODO: The covariance in the spherical harmonic basis can be used as a prior
-    in many other settings.}
-
-\subsection{Comparison to other work}
-\label{sec:other-work}
-\xxx{TODO: Discuss \citet{Basri2020}, how we agree and disagree. Compare to Brett's
-    paper using ABC to measure spot sizes as a function of spectral type. Discuss
-    Simon Wood's paper in ecology on synthetic likelihoods. Talk about similarities
-    to Gaussian Random Fields in CMB studies.}
-
-\subsection{Caveats}
-\label{sec:caveats}
-\xxx{TODO: Miscellaneous caveats.}
-\xxx{
-    \begin{enumerate}
-        \item The latitude distribution is not Gaussian, so careful
-              when interpreting $\mu_\phi$ and $\sigma_\phi$
-        \item The number of spots does not have to be an integer. This is
-              fine, since it makes the model more flexible.
-        \item Careful with light curves with large amplitudes, although
-              Figure~\ref{fig:calibration_hicontrast}
-              has max amplitudes of more than 10\% and works great.
-        \item No differential rotation / spot evolution yet! We're working on it.
-        \item We might want a shrinkage prior on sigma, since it's hard to
-              break the polar spot degeneracy with single-band light curves.
-    \end{enumerate}
-}
+$P$ and $\mathbf{u}$ will be the subject of an upcoming paper.
 
@@
 although an implementation of this is deferred to the next paper.
-\xxx{@DFM: Is this true? I.e., is there an efficient way to solve Equation~(\ref{eq:log-like})
-    when $\pmb{\Sigma}$ is the same for all stars but $\mathbf{C}$ is different? Is there
-    some limit in which an approximation of some sort can be made?}
 

### 2021-01-12 tex/figures/nullspace_ensemble_ld_var.py

@@ -73,7 +73,7 @@ ax.set_xticks(l ** 2 + l)
 ax.set_xticklabels(["{}".format(l) for l in np.arange(2, ydeg + 1)])
 ax.set_xlabel("spherical harmonic degree")
 ax.set_ylabel("posterior shrinkage")
-leg = ax.legend(loc="upper right", title="light curves", fontsize=10)
+leg = ax.legend(loc="upper right", title="orientations", fontsize=10)
 leg.get_title().set_fontsize(8)
 leg.get_title().set_fontweight("bold")
 

### 2021-01-12 tex/ms.tex
trying to fix mem errors on azure

@@
 \begin{enumerate}
-    \item \textbf{The spots are not circular.} This is most evident in row (c),
+    \item \emph{The spots are not circular.} This is most evident in row (c),
           where some spots are distinctly asymmetric.
-    \item \textbf{There is significant variance in contrast from one spot to another},
+    \item \emph{There is significant variance in contrast from one spot to another},
           even though our model
@@
           not constant, even though (again) our model implictly treats it as such.
-    \item \textbf{There is ringing in the background.} This is apparent to some extent in
+    \item \emph{There is ringing in the background.} This is apparent to some extent in
           row (b), where there are small fluctuations in the brightness at low
           latitudes where no spots are present.
-    \item \textbf{There aren't exactly 10 (or 20) spots in those maps.} This is
+    \item \emph{There aren't exactly 10 (or 20) spots in those maps.} This is
           most obvious in row (c), where only a few large distinct spots,
           plus maybe (?) a few smaller ones, are visible.
-    \item \textbf{There are bright spots in addition to dark spots.} This may be
+    \item \emph{There are bright spots in addition to dark spots.} This may be
           the most glaring issue. We explicitly model spots as being dark, and yet
@@
 
-Another limitation of the base algorithm...
+Another big limitation of the base algorithm is the implicit assumption that
+stellar surfaces are static. Our GP hyperparameters $\pmb{\theta}_\bullet$
+describe the \emph{spatial} configuration of star spots, but they say nothing
+about their evolution in time. We know from observations of the Sun and of
+\emph{Kepler} stars that temporal variability is extremely common: spots appear,
+disappear, and even migrate in latitude over time. While it may be possible to
+parametrize their evolution in a way that is general enough to capture all the
+ways in which they may change over time, such an approach is beyond the scope
+of the present paper. It is, however, straightforward to implement an
+\emph{uninformative} temporal prior within the framework of our GP. To do this,
+we will make two simplifying assumptions:
+
+\begin{enumerate}
+    \item \emph{The temporal process is stationary.} This implies that there is no
+          preferred time (or phase) and that the spatial covariance is the
+          same at all points in time.
+    \item \emph{The temporal and spatial covariances are independent.} This implies
+          that the evolution of each spherical harmonic mode in time is independent of the
+          evolution of any of the other modes in time.
+\end{enumerate}
+
+There is some tension between these assumptions and our knowledge of how stellar
+surfaces evolve. Assumption (1) excludes surfaces whose total spottiness changes
+significantly
+or whose spots migrate in latitude, as both processes change the spatial covariance
+over time. Assumption (2) ignores the correlation between spherical harmonic modes
+due to the migration of spots, which requires the coherent evolution of many
+modes at once. These assumptions likely limit the ability of our GP to model
+light curves on very long baselines (i.e., on timescales of years) over which
+stellar activity cycles take place. However, given that the use case of our
+algorithm is likely to be the analysis of individual quarters of \emph{Kepler}
+and individual sectors of \emph{TESS} data,
+our assumptions are likely valid in most cases. Analyses of (say) all quarters of
+\emph{Kepler} data could process each quarter at a time in a hierarchical
+framework in which hyperparameters like the mean spot latitude in each quarter
+are treated as functions of time.
+
+The two assumptions listed above suggest a fairly straightforward form for the
+GP covariance in spherical harmonics and time:
+%
+\begin{align}
+    \pmb{\Sigma}_\mathbf{t,y} = \pmb{\Sigma}_\mathbf{t} \otimes \pmb{\Sigma}_\mathbf{y}
+    \quad,
+\end{align}
+%
+where $\pmb{\Sigma}_\mathbf{t}$ is a $(K \times K)$ matrix describing the
+covariance among the $K$ points in time,
+$\pmb{\Sigma}_\mathbf{y}$ is the $(N \times N)$ matrix describing describing
+the covariance among the $N \equiv (l_\mathrm{max} + 1)^2$ spherical
+harmonic coefficients,
+and $\otimes$ denotes the Kronecker product.
+The quantity $\pmb{\Sigma}_\mathbf{t,y}$
+is the $(NK \times NK)$ temporal-spatial covariance, whose coefficient at index
+$(Nk + n, Nk' + n')$ is the covariance between the spherical harmonic coefficient
+$y_{n}(t_k)$ and the spherical harmonic coefficient $y_{n'}(t_{k'})$.
 
@@
 $\mathcal{O}(l_\mathrm{max}^3)$ times
-\emph{for each entry} in the $K \times K$ matrix $\mathbf{E}_I$.
+\emph{for each entry} in the $(K \times K)$ matrix $\mathbf{E}_I$.
 

### 2021-01-12 tex/ms.tex
reorder yml

@@
 \begin{align}
+    \label{eq:expsq}
     k_\mathrm{E^2}(\Delta t)           & = \sigma^2 \exp\left(-\frac{\Delta t^2}{2\tau}\right)
@@
     \intertext{the Mat\'ern-3/2 kernel,}
+    \label{eq:mat32}
     k_\mathrm{M^\frac{3}{2}}(\Delta t) & = \sigma^2 \left( 1 + \frac{\sqrt{3}\Delta t}{\tau} \right) \exp\left(-\frac{\sqrt{3}\Delta t}{\tau}\right)
@@
 \begin{align}
-    \pmb{\Sigma}_\mathbf{t,y} = \pmb{\Sigma}_\mathbf{t} \otimes \pmb{\Sigma}_\mathbf{y}
+    \label{eq:Sigmaty}
+    \pmb{\Sigma}_\mathbf{y}^\mathbf{(t)} = \mathbf{K} \otimes \pmb{\Sigma}_\mathbf{y}
     \quad,
@@
 %
-where $\pmb{\Sigma}_\mathbf{t}$ is a $(K \times K)$ matrix describing the
+where $\mathbf{K}$ is a $(K \times K)$ matrix describing the
 covariance among the $K$ points in time,
@@
 the covariance among the $N \equiv (l_\mathrm{max} + 1)^2$ spherical
-harmonic coefficients,
+harmonic coefficients (Equation~\ref{eq:cov_y}),
 and $\otimes$ denotes the Kronecker product.
-The quantity $\pmb{\Sigma}_\mathbf{t,y}$
+The quantity $\pmb{\Sigma}_\mathbf{y}^\mathbf{(t)}$
 is the $(NK \times NK)$ temporal-spatial covariance, whose coefficient at index
 $(Nk + n, Nk' + n')$ is the covariance between the spherical harmonic coefficient
-$y_{n}(t_k)$ and the spherical harmonic coefficient $y_{n'}(t_{k'})$.
-
-\xxx{TODO: The model presented in this paper assumes statics stellar surfaces,
-    corresponding to perfectly periodic light curves. This is hardly ever the
-    case in practice, as spots evolve and differential rotation changes the surface
-    over time. We can still use this model, but we need to either restrict our analysis
-    to segments of light curves of duration shorter than the timescale on which
-    the surface evolves, or explicitly model the effect of time evolution on the covariance.
-    The next paper will deal with temporal evolution more
-    rigorously, but the way to do this is just take the elementwise product of the \starryprocess
-    flux covariance with a temporal covariance. This corresponds to the Kronecker product
-    in spherical harmonic space. Show a figure, give an example, etc.}
+$y_{n}(t_k)$ and the spherical harmonic coefficient $y_{n'}(t_{k'})$, where
+the index $n$ is related to the spherical harmonic indices $l$ and $m$ via
+Equation~(\ref{eq:n}). Finally, because of our assumption of stationarity,
+the mean of the GP is still constant and equal to the mean of the static
+process.
 
@@
 
-\begin{proof}{test_temporal}
+The covariance matrix in Equation~(\ref{eq:Sigmaty}) can be sampled from to
+yield time-variable surface maps, or it can be transformed into flux space
+for sampling light curves or computing likelihoods. The covariance in
+flux space is given by
+%
+\begin{align}
+    \label{eq:SigmatSlow}
+    \pmb{\Sigma}^\mathbf{(t)} & =
     \pmb{\mathcal{A}}^\ddagger
-    \left(\pmb{\Sigma}_\mathbf{t} \otimes \pmb{\Sigma}_\mathbf{y}\right)
+    \,
+    \pmb{\Sigma}_\mathbf{y}^\mathbf{(t)}
+    \,
     {\pmb{\mathcal{A}}^\ddagger}^\top
-    &=
-    \pmb{\Sigma} \odot \pmb{\Sigma}_\mathbf{t}
-\end{proof}
+\end{align}
 %
@@
                       &                   &        & \,\,\, \mathbf{a}_{K-1}^\top
-    }\quad.
-\end{align}
+    }
+\end{align}
+%
+is a design matrix constructed by staggering the rows of the standard design
+matrix $\pmb{\mathcal{A}}$ (see Appendix \ref{sec:starry:basic}). The diagonal
+structure of $\pmb{\mathcal{A}}^\ddagger$ is due to the fact that each snapshot of the surface is only observed
+at a single phase.
+
+In the temporally-variable version of our GP, the covariance matrix in
+Equation~(\ref{eq:SigmatSlow}) replaces the standard covariance
+$\pmb{\Sigma}$; it can similarly be modified (\S\ref{sec:baseline})
+to obtain the covariance of the normalized process. While $\pmb{\Sigma}^\mathbf{(t)}$
+is $(K \times K)$, it is computed from the contraction of a much larger
+$(NK \times NK)$ matrix, which is extremely inefficient to instantiate and
+operate on. Fortunately, it can be shown that
+%
+\begin{proof}{test_temporal}
+    \pmb{\Sigma}^\mathbf{(t)} & =
+    \pmb{\mathcal{A}}^\ddagger
+    \,
+    \pmb{\Sigma}_\mathbf{y}^\mathbf{(t)}
+    \,
+    {\pmb{\mathcal{A}}^\ddagger}^\top
+    \nonumber \\
+    &=
+    \pmb{\mathcal{A}}^\ddagger
+    \left(\mathbf{K} \otimes \pmb{\Sigma}_\mathbf{y}\right)
+    {\pmb{\mathcal{A}}^\ddagger}^\top
+    \nonumber \\
+    &=
+    \pmb{\Sigma} \odot \mathbf{K}
+    \quad,
+\end{proof}
+%
+that is, the flux covariance is just the elementwise $\odot$ product of the
+GP covariance $\pmb{\Sigma}$ and the temporal covariance $\mathbf{K}$.
+This fact makes the temporal GP \emph{just as efficient} to
+evaluate as the standard GP!
+
+Armed with this algorithm for computing $\pmb{\Sigma}^\mathbf{(t)}$, it only
+remains for us to decide on a structure for $\mathbf{K}$. While this can
+in principle be any covariance matrix constructed from a stationary kernel,
+we recommend one of the common radial kernels such as the exponential
+squared kernel (Equation~\ref{eq:expsq}) or the Mat\'ern-3/2 kernel
+(Equation~\ref{eq:mat32}) with variance parameter $\sigma^2$ set to
+unity (since the variance is already specified within $\pmb{\Sigma}$). In this
+case, the temporal covariance $\mathbf{K}$ is a function of a single
+parameter: the timescale of the variability, $\tau$.
 
@@
 %
+\label{sec:starry:basic}
 As we mentioned in \S\ref{sec:gp}, the flux $\mathbf{f}$ is a purely

### 2021-01-12 tex/ms.tex
more text

@@
 case, the temporal covariance $\mathbf{K}$ is a function of a single
-parameter: the timescale of the variability, $\tau$.
+parameter: the timescale of the variability, $\tau$. Similar to the other
+hyperparameters of our GP, this parameter can be estimated in an inference
+setting. However, a detailed investigation of the ability of our temporal
+GP to accurately capture spot variability is beyond the scope of this paper,
+and will be revisited in the future, along with an algorithm to explicitly
+model the effects of differential rotation on the covariance structure.
+
+Figure~\ref{fig:temporal} shows a single prior sample from the temporal
+GP described above, assuming a squared exponential covariance with $\tau = 20$ days
+and a rotation period $P = 1$ day. We compute both the map samples at five
+different times (top) and the light curve sample over ten rotations (bottom).
+The flux variability is qualitatively similar to that seen in
+temporally-variable \emph{Kepler} light curves.
 

### 2021-01-13 tex/ms.tex
more text; new fig; azure test hack

@@
 \label{sec:brightspots}
-\xxx{TODO: We can easily model bright spots and plages by setting $c < 0$. Show
-    an example. This model can be used to differentiate between spot-dominated
-    and plage-dominated surfaces, which is physically interesting. Cite
-    Brett's paper on TRAPPIST-1.}
+
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/brightspots.pdf}
+        \oscaption{brightspots}{%
+            Comparison of GP samples with dark spots (top) and bright
+            spots (center) alongside their corresponding light curves
+            viewed at $I = 75^\circ$ (bottom). Without knowledge of
+            the baseline (\S\ref{sec:baseline}), it is very
+            difficult to differentiate between the two from stellar
+            light curves.
+            \label{fig:brightspots}
+        }
+    \end{centering}
+\end{figure}
+
+All the calibration tests performed in the previous section assumed the
+stellar surface was dominated by dark spots. We can easily model the
+effect of bright spots by choosing negative values for the contrast
+parameter, i.e. $c < 0$. The top two panels of Figure~\ref{fig:brightspots}
+show five random samples from the GP with dark spots ($c = 0.1$) and
+bright spots ($c = -0.1$), respectively; the random seed is the same for
+both panels, so the maps are identical in all other respects.
+While the surface maps can be easily distinguished by eye, the same is not
+true for the corresponding light curve samples (bottom panel), which
+are almost identical. There is a slight difference in amplitude between
+the two cases: surfaces with dark spots have slightly higher light curve
+amplitudes than surfaces with bright spots of the same contrast magnitude.
+However, the magnitude of the bright spots can be increased slightly to get
+a near-perfect match to the dark spot light curves, meaning it may be
+difficult (if not impossible) to tell the difference between dark and bright
+spots via the GP approach.
+
+The reason for this degeneracy is rooted (once again) in the
+fundamental issue with relative photometry: we lack any information about
+the true baseline level of the light curve.
+Consider the dependence of the (unnormalized) GP covariance on the
+contrast: it enters in a single place, via Equation~(\ref{eq:bigEc}),
+as $c^2$, meaning dark and bright spots models have exactly the same
+covariance. These models differ only in the mean of the unnormalized
+process, since that is proportional to $c$ (via Equation~\ref{eq:ec}).
+However, as we argue in \S\ref{sec:baseline}, the mean is not a direct
+observable. Instead, in relative photometry, we are only sensitive to
+the \emph{ratio} of the covariance to the square of the mean
+(see Equation~\ref{eq:SigmaTilde}). From that equation, we can deduce
+that stars with dark spots (for which
+the light curve mean $\mu < 1$) will therefore have larger variance
+than stars with bright spots ($\mu > 1$), leading to the slight difference
+seen in the figure. However, since this is strictly a multiplicative
+factor affecting the covariance, it is degenerate with the two other
+properties that scale the covariance: the magnitude of the contrast and
+the total number of spots.%
+\footnote{There is also a small additive term in
+    Equation~(\ref{eq:SigmaTilde}), but this, too, depends only on the ratio
+    of entries in the covariance matrix to the mean, so it is of little help
+    in breaking the degeneracy.}
+
+We therefore conclude that \textbf{relative single-band photometry is largely insensitive to
+    the difference between bright spots and dark spots}. However, it is
+important to bear in mind that the
+degeneracy described above exists only for the \emph{Gaussian approximation}
+to the likelihood function. As we argued earlier, the true likelihood function
+is not a Gaussian; in particular, the true probability distribution
+has higher-order moments that we do not model here. These moments should
+in principle encode information about the sign of the spot contrast, but they
+may be very difficult to infer in practice. It may be possible to distinguish
+between dark spots and bright spots with traditional forward models of stellar
+surfaces, but (as we argued earlier) a statistically rigorous ensemble analysis
+of stellar light curves using such forward models is probably computationally
+intractable.
+
+We can, however, skirt this degeneracy with observations in multiple bands,
+which can provide limited information about the baseline level. Recently,
+\citet{Morris2018} used approximately coeval \emph{Kepler} and \emph{Spitzer}
+light curves of TRAPPIST-1 to argue that a bright spot model for the star
+is more consistent with the data. We discuss the ability of multi-band
+photometry to break these degeneracies in the following section.
 
@@
 \begin{align}
+    \label{eq:ec}
     \mathbf{e}_c & = c \, \mathbf{e}_\lambda
     \\
+    \label{eq:bigEc}
     \mathbf{E}_c & = c^2 \mathbf{E}_\lambda

### 2021-01-13 tests/test_lnlike.py

@@ -11,12 +11,13 @@ from theano.configparser import change_flags
 import pytest
 from tqdm import tqdm
 from itertools import product
+import os
+
+ON_AZURE = bool(int(os.getenv("ON_AZURE", "0")))
 
 
 def get_functions(marginalize_over_inclination=False, seed=42):
-    def _lnlike(
-        r, a, b, c, n, i, p, t, flux, data_cov,
-    ):
+    def _lnlike(r, a, b, c, n, i, p, t, flux, data_cov):
         gp = StarryProcess(
             r=r,
             a=a,
@@ -28,9 +29,7 @@ def get_functions(marginalize_over_inclination=False, seed=42):
         )
         return gp.log_likelihood(t, flux, data_cov, p=p, i=i)
 
-    def _sample(
-        r, a, b, c, n, i, p, t,
-    ):
+    def _sample(r, a, b, c, n, i, p, t):
         gp = StarryProcess(
             r=r,
             a=a,
@@ -100,6 +99,7 @@ def test_lnlike_array(marginalize_over_inclination, plot=False, seed=42):
     assert np.abs(b_arr[np.nanargmax(ll)] - defaults["b"]) < 0.10
 
 
+@pytest.mark.skipif(ON_AZURE, reason="consumes too much memory")
 @pytest.mark.parametrize(
     "param,marginalize_over_inclination",
     product(["r", "a", "b", "c", "n", "i", "p"], [False, True]),
@@ -134,4 +134,3 @@ def test_lnlike_grad(param, marginalize_over_inclination):
                 (defaults[param],),
                 n_tests=1,
             )
-

### 2021-01-14 tex/ms.tex
more text

@@
 coefficients (Equation~\ref{eq:fAy}),
-we may write the mean and covariance of our flux GP as
+we may write the mean and covariance of our GP as
 %
@@
 %
-are the mean and covariance of the GP in the spherical harmonics basis.
+are the mean and covariance of the distribution over spherical harmonic coefficients.
 The bulk of the math in this paper (Appendix~\ref{sec:integrals})
@@
 As we mentioned in the previous section,
-the equations for the mean and covariance of the flux GP
+the equations for the mean and covariance of our GP
 (Equations~\ref{eq:mean_f} and \ref{eq:cov_f}, respectively) are conditioned
@@
 \label{sec:tinyspots}
-\xxx{TODO: The model is limited in resolution to features with radii $r \gtrsim 10^\circ$.
-    Typical sun spots are MUCH smaller than that. However, our model can still
-    correctly infer latitudinal parameters even when spots are really tiny
-    (i.e., no bias due to the limited resolution). We might try to improve the
-    numerical stability of the model at higher spherical harmonic degree, but
-    that's a task for a future paper.}
+
+One of the biggest downsides of adopting a spherical harmonic representation
+of the stellar surface as the foundation of our GP is the inherent limitation
+it imposes on the resolution of surface features. In order to maximize
+computational efficiency and numerical stability, our default approach is to
+model the surface using an expansion up to degree $l_\mathrm{max} = 15$,
+which can model features only as small as
+$\sim \nicefrac{180^\circ}{15} = 12^\circ$ on the surface. Even on scales
+slightly larger than this, the presence of ringing can be seen
+(c.f. panel (b) in Figure~\ref{fig:samples}, where ringing can just barely be
+seen in the equatorial region of the maps). The spherical harmonic basis consists
+of global modes, all of which contribute to the intensity everywhere on the surface.
+Localized features require constructive interference of modes inside
+and destructive interference of modes outside, often leading to a wave-like ringing
+pattern that gets worse as the size of the features gets smaller. Taken at
+face value, this might suggest that a different basis---such as the common
+choice of pixels on a grid, or perhaps a localized wavelet basis---would be better
+at modeling small spots. While this is probably true, a different choice of
+basis would not admit the same analytic solutions to the integrals needed to
+compute the GP covariance. And it is precisely the ability to compute these
+integrals analytically (Appendix~\ref{sec:integrals}) that makes the evaluation
+of the marginal likelihood tractable; it is difficult to envision an approach
+that uses a \emph{local} basis and simultaneously enjoys the same
+analyticity under integration.
+
+In principle, small spots can be modeled with negligible ringing by simply
+increasing the degree of the spherical harmonic expansion. As we discuss in
+\S\ref{sec:implementation}, however, the algorithm presented here becomes
+unstable for $l \gtrsim 15$, so doing so would require a reparametrization
+of the equations in the Appendix. However, as we showed in
+\S\ref{sec:calibration}, the current implementation of the GP is suitable
+to modeling light curves of stars with spots smaller than the limiting
+resolution of $\sim 10^\circ$. Consider Figure~\ref{fig:calibration_tinyspots},
+which shows the results of doing inference on an ensemble of light curves
+of stars with small ($r = 3^\circ$) spots; these are comparable to some of
+the larger spots seen on the Sun with diameters of about $70,000$ km.
+Even though our inferred radius and contrast are wrong, the fact that the radius
+posterior peaks
+at the lower prior bound of $10^\circ$ is strongly suggestive of the presence
+of spots smaller than the resolution of the model. Moreover, the
+latitude mode and standard deviation posteriors are \emph{unbiased}, and we
+correctly infer the presence of low-variance, mid-latitude spots.
+With the above caveats in mind, our GP can therefore still be used to model
+stars with small spots.
 
@@
 
-\subsection{Application to transits and RV}
-\label{sec:other-work}
-\xxx{TODO: The covariance in the spherical harmonic basis can be used as a prior
-    in many other settings.}
-
 \subsection{Comparison to other work}
@@
 \label{sec:caveats}
-\xxx{TODO: Miscellaneous caveats.}
-\xxx{
+
+We conclude our discussion with a list of several caveats that should be kept
+in mind when using our algorithm and/or its \Python implementation.
+
+\xxx{TODO}
+
 \begin{enumerate}
-        \item The latitude distribution is not Gaussian, so careful
-              when interpreting $\mu_\phi$ and $\sigma_\phi$
-        \item The number of spots does not have to be an integer. This is
-              fine, since it makes the model more flexible.
-        \item Careful with light curves with large amplitudes, although
+    \item \emph{The assumed latitude distribution is not Gaussian}.
+          Careful when interpreting $\mu_\phi$ and $\sigma_\phi$.
+    \item \emph{The number of spots $n$ does not have to be an integer}.
+          This is fine, since it makes the model more flexible.
+    \item \emph{Care should be taken when modeling large-amplitude light curves}.
+          Careful with light curves with large amplitudes, although
           Figure~\ref{fig:calibration_hicontrast}
           has max amplitudes of more than 10\% and works great.
-        \item No differential rotation / spot evolution yet! We're working on it.
-        \item We might want a shrinkage prior on sigma, since it's hard to
+    \item \emph{Keep in mind the polar spot degeneracy}.
+          We might want a shrinkage prior on sigma, since it's hard to
           break the polar spot degeneracy with single-band light curves.
+    \item \emph{Careful with the data}. Jitter term, baseline, etc.
+    \item \emph{Careful with sample selection}. Even non-variable light curves.
+          If you're worried, use a mixture model.
 \end{enumerate}
-}
 
@@
 $P$ and $\mathbf{u}$ will be the subject of an upcoming paper.
+
+\subsection{Modeling transits and radial velocity datasets}
+\label{sec:transits-rvs}
+
+Finally, we would like to note that the GP developed in this paper is not
+limited to modeling (rotational) light curves of stars. We derived the
+covariance structure of spotted stellar surfaces in the spherical harmonic
+basis, which we then linearly transformed into flux space to obtain the
+light curve GP presented above. However, our GP may be used to model
+\emph{any} kind of observation that
+is linearly related to the spherical harmonic representation. If $\mathbf{A}$
+is the linear operator that transforms the spherical harmonic representation
+$\mathbf{y}$ to the data vector $\mathbf{d}$ via
+$\mathbf{d} = \mathbf{A} \mathbf{y}$, then the mean and covariance of the
+GP model for $\mathbf{d}$ are given by
+%
+\begin{align}
+    \pmb{\mu}_\mathbf{d}    & = \mathbf{A} \, \pmb{\mu}_\mathbf{y}
+    \\
+    \pmb{\Sigma}_\mathbf{d} & = \mathbf{A} \, \pmb{\Sigma}_\mathbf{y} \, \mathbf{A}^\top
+    \quad,
+\end{align}
+%
+respectively, where $\pmb{\mu}_\mathbf{y}$ and $\pmb{\Sigma}_\mathbf{y}$
+are the mean and covariance in the spherical harmonic basis derived in this paper.
+In \citet{Luger2019} we showed that
+occultation light curves are also linearly related to the
+spherical harmonic representation of the stellar surface, so it is
+straightforward to use our GP to model transits of planets across spotted
+stars, either to marginalize over the spot variability or to constrain the
+surface map of the star. In this context, $\mathbf{A}$ may be computed via the
+\textsf{design\_matrix()} method of a \textsf{Map} instance of the \starry
+package.
+
+The formalism developed here may also be extended to model radial
+velocity (RV) datasets, although this requires a bit more work. The
+instantaneous radial
+velocity shift $v$ induced by a rotating spotted star may be approximated as
+%
+\begin{align}
+    v & = \frac{\iint_S I \, V \, \mathrm{d} S}{\iint_S I \mathrm{d} S}
+\end{align}
+%
+where $I$ is the stellar intensity at a point on the surface, $V$ is the
+radial component of the rotational velocity vector at that point, and the
+integral is taken over the projected disk of the star.
+%
+If we expand the surface intensity distribution in spherical harmonics,
+the integral in the denominator is just a classical \starry integral, as it
+is just the disk-integrated intensity (i.e., the flux).
+%
+The numerator may also be computed following the \starry formalism,
+provided we weight the surface intensity representation by the velocity
+field $V$. In the case of rigid body rotation, $V$ is exactly a dipole ($l = 1$)
+field, so the quantity $I V$ can be expressed exactly as a product of
+spherical harmonics.%
+\footnote{
+    Higher order effects such as differential rotation and convective
+    blueshift can be easily modeled with a higher degree expansion of $u$.
+}
+Since spherical harmonics are closed under multiplication,
+we may write $I V$ as a linear combination of spherical harmonics, meaning
+the integral in the numerator is \emph{also} a \starry integral.
+For some linear operators $\mathbf{A}$ and $\mathbf{B}$, we may therefore
+write
+%
+\begin{align}
+    \mathbbb{v} & = \frac{\mathbf{B} \, \mathbbb{y}}{\mathbf{A} \, \mathbbb{y}}
+\end{align}
+%
+where $\mathbbb{v}$ is the random variable representing the observed radial
+velocity timeseries and
+$\mathbbb{y}$ is the Gaussian random variable representing the spherical harmonic
+coefficients describing the stellar surface.
+%
+Since the $\mathbbb{v}$ is the ratio of two Gaussian random variables, its
+distribution is not Gaussian. However, we can still compute the first two moments
+of the distribution of $\mathbbb{v}$ to derive a Gaussian approximation to it
+(similar to what we did in \S\ref{sec:baseline}), which will yield the mean
+and covariance of the Gaussian process representation of the radial velocity
+timeseries.
+%
+Given the complexity of the operations described above,
+we defer this calculation (and the calibration of the resulting GP model)
+to future work.
 
@@
           prior on the stellar surface in Doppler imaging, Doppler tomography,
-          or even radial velocity searches for exoplanets.
+          or even radial velocity searches for exoplanets. The latter will
+          be the subject of a future paper in this series.
 \end{enumerate}
@@
 
-The next papers in this series will focus on (II) a more rigorous
-treatment of time variability and (III) explicit marginalization over
-the remaining stellar parameters. In particular, paper II will focus
+The next papers in this series will focus on (in no particular order)
+\begin{itemize}
+    \item \emph{A more rigorous treatment of time variability}.
+          This paper will focus
           on modeling differential rotation, whose effect on the covariance of the
@@
           are too similar for current methodology to reliably discern between them.
-Paper III will focus on marginalization over the stellar rotation period
-and limb darkening coefficients, which can also be done analytically
+    \item \emph{Explicit marginalization over the remaining stellar parameters}.
+          These include the stellar rotation period
+          and limb darkening coefficients, which can be marginalized over analytically
           under certain choices of prior. This will eliminate all per-star
@@
           speeding up inference for large ensembles of stellar light curves.
+    \item \emph{Extension of this formalism to radial velocity datasets}.
+          As we discussed in \S\ref{sec:transits-rvs}, it is possible to extend
+          the methodology presented here to model the contribution of stellar
+          surface variability to radial velocity measurements, which can be used
+          (for instance) to mitigate systematics in extreme precision radial velocity
+          (EPRV) searches for exoplanets.
+    \item \emph{Extension of this formalism to Doppler imaging}.
+          As we show in upcoming work, it is possible to derive an exact
+          linear relationship between the wavelength-dependent spherical harmonic
+          representation of a rotating star and its time-variable spectrum.
+          This linearity makes it possible to adapt our GP formalism to the Doppler
+          imaging problem, providing an efficient marginal likelihood function
+          for rigorous inference studies.
+\end{itemize}
 
@@
 %
-which are linearly related to the mean and the covariance of our GP in flux
+which are linearly related to the mean and the covariance of our GP
 (\S\ref{sec:gp}).
@@
 %
-is the first moment of the GP in the spherical harmonic basis
+is the first moment of the distribution over spherical harmonic coefficients
 (Equation~\ref{eq:exp_y}). We can use Equations~(\ref{eq:Arows})
@@
 %
-is the second moment of the GP in the spherical harmonic basis
+is the second moment of the distribution over spherical harmonic coefficients
 (Equation~\ref{eq:exp_yy}). We can use Equations~(\ref{eq:Arows})

### 2021-01-14 tex/ms.tex
getting there

@@
 has a much heavier tail than in the other runs, showing that the model is
-in fact sensitive to the spot contrast to some degree. In
-Figure~\ref{fig:calibration_bigspots}, we compensate for the smaller number
+in fact sensitive to the spot contrast to some degree.%
+\footnote{The sensitivity to $c$ is for the most part
+    due to our hard lower prior bound $n \ge 1$.}
+In Figure~\ref{fig:calibration_bigspots}, we compensate for the smaller number
 of spots by increasing the spot radius to $r = 45^\circ \pm 5^\circ$ instead,
@@
 \label{sec:breaking_baseline_degeneracy}
-\xxx{TODO: Discuss the baseline degeneracy and how to break it. Discuss the
-    degeneracy between $c$ and $N$ and how to break it. The basic idea here
-    is that if we can constrain the (spotless) photospheric temperature, we
-    can uniquely infer $n$. I think this can be done by taking light curves
-    in two bands, but I'm checking with Gully. @HEDGES: Anything to add on this?}
+
+We showed in \S\ref{sec:calibration-inference} that there exists a
+strong degeneracy between the contrast of individual spots $c$ and
+the total number of spots $n$ on a star. For relative photometry, in
+which the true baseline level (\S\ref{sec:baseline}) is not known,
+this degeneracy is perfect: relative photometry cannot tell us anything
+about the total spottiness of a stellar surface. Inspection of the
+dependence of the GP on $c$ and $n$ (c.f. Appendix~\ref{sec:integrals}) reveals
+that the mean and variance of the GP scale as
+%
+\begin{align}
+    \mu      & \propto c \, n
+    \nonumber                   \\
+    \sigma^2 & \propto c^2 \, n
+    \quad,
+\end{align}
+%
+respectively. If our observations were sensitive to both the true mean and
+variance of stellar light curves,
+we could uniquely infer $c$ and $n$. However, as we discussed in \S\ref{sec:baseline},
+to very good approximation, relative photometry can only tell us about the ratio
+%
+\begin{align}
+    \frac{\sigma^2}{\mu^2} \propto \frac{c}{n}
+    \quad,
+\end{align}
+%
+i.e., the ratio of the spot contrast to the total number of spots.
+The reason for this is that the relationship between the amplitude of the
+light curve (measured in, e.g., counts per second on the detector)
+and the fractional intensity deficit due to spots on the surface of
+the star is not known. It is easy enough to normalize a light curve to its
+mean, median, or maximum to obtain what we typically refer to as
+``fractional'' or ``relative'' units, but this is \emph{not the same
+    thing}! Consider the two stellar surfaces in Figure~\ref{fig:mean_normalization},
+whose true light curves---measured in units where an unspotted star would
+have unit flux---have different baselines and different amplitudes (top panel).
+This is not what we observe when we do photometry---instead, we observe these
+light curves multiplied by an unknown amplitude, which depends on (potentially
+knowable) factors such as the luminosity of the star, the distance to it,
+and the efficiency of the detector but also on the (typically unknowable)
+unspotted photospheric brightness, which sets the true baseline level.
+%
+Division of these light curves by their (observed) baseline levels
+(which is the operation typically done on stellar light curves) yields the
+``relative'' light curves in the bottom panel, which are \emph{indistinguishable}
+from each other. In this example, we are unable to distinguish between
+models of the surface with different spot contrasts and number of spots (as well
+as total spot ``coverage''). In general, these quantities cannot be learned
+from single-band photometric measurements.
+%
+
+Fortunately, it may be possible to break this degeneracy (at least to some extent)
+by collecting photometric data in multiple wavelength bands, an idea that has
+been explored recently \citep[e.g.,][]{Gully2017,Guo2018}. Assuming the
+locations and sizes of a star's spots are constant in wavelength, the amplitude
+of the light curve in different bands (and in particular its slope as a function
+of wavelength) can be used to directly constrain the
+temperature, and hence the contrast, of the spots. This effectively breaks the
+$c-N$ degeneracy. In practice, the effective size of spots may be different
+at different wavelengths, which could complicate this picture somewhat. An extension
+of the formalism in this paper to multi-band photometry, and a detailed investigation
+of the information content of such observations, is left to future work.
 
@@
 
-We conclude our discussion with a list of several caveats that should be kept
+We conclude our discussion with a list of several notes and caveats that should be kept
 in mind when using our algorithm and/or its \Python implementation.
-
-\xxx{TODO}
 
@@
     \item \emph{The assumed latitude distribution is not Gaussian}.
-          Careful when interpreting $\mu_\phi$ and $\sigma_\phi$.
+          Because we require the first and second moment integrals
+          (Equations~\ref{eq:ephi} and \ref{eq:Ephi}) to have closed
+          form solutions, there are restrictions on the probability
+          density function we can assume for the spot latitude.
+          We find that a Beta distribution in the cosine of the latitude
+          is integrable in closed form and can be evaluated efficiently
+          in terms of recursion relations. In many cases, particularly
+          when $\mu_\phi \lesssim 75^\circ$ and $\sigma_\phi \lesssim 10^\circ$,
+          the distribution in the spot latitude
+          is close to a bi-modal Gaussian with mean $\pm\mu_\phi$ and
+          standard deviation $\sigma_\phi$ (see Figure~\ref{fig:latitude_pdf}).
+          In general, however, $\mu_\phi$ is formally equal to the \emph{mode}
+          (as opposed to the mean) of the distribution, and $\sigma_\phi$ is
+          the Laplace approximation to the local standard deviation at the mode.
     \item \emph{The number of spots $n$ does not have to be an integer}.
-          This is fine, since it makes the model more flexible.
+          Samples from our GP prior will not in general have exactly $n$
+          spots (see, e.g., Figure~\ref{fig:samples}). This is due to the fact
+          that our model is only an approximation to the true distribution
+          of stellar surfaces conditioned on the spot properties. A corollary
+          of this point is that $n$ need not be an integer, which makes it
+          easier in practice to sample over using modern inference
+          techniques such as MCMC, HMC, ADVI, and nested sampling.
     \item \emph{Care should be taken when modeling large-amplitude light curves}.
-          Careful with light curves with large amplitudes, although
-          Figure~\ref{fig:calibration_hicontrast}
-          has max amplitudes of more than 10\% and works great.
-    \item \emph{Keep in mind the polar spot degeneracy}.
-          We might want a shrinkage prior on sigma, since it's hard to
-          break the polar spot degeneracy with single-band light curves.
-    \item \emph{Careful with the data}. Jitter term, baseline, etc.
-    \item \emph{Careful with sample selection}. Even non-variable light curves.
-          If you're worried, use a mixture model.
+          We discussed this point at length in \S\ref{sec:baseline}. Modeling
+          a light curve that has been normalized to its mean (or median) as a
+          Gaussian process is conceptually a bad idea when the amplitude
+          of variability is large compared to the mean. As a rule of thumb, if
+          the amplitude of variability exceeds $\sim 10\%$, we recommend
+          not normalizing the light curve in this way, and instead modeling
+          the normalization amplitude as a latent variable.
+          % Note that Figure~\ref{fig:calibration_hicontrast}
+          % has max amplitudes of more than 10\% and works great.
+    \item \emph{Keep in mind the polar spot degeneracy}. Even when modeling
+          ensembles of light curves, there are still strong degeneracies at
+          play. In particular, spots centered on the poles are always in
+          the null space, so it can be difficult in practice to rule out
+          their presence. This can be seen in Figure~\ref{fig:calibration_hilat},
+          in which the model cannot distinguish between localized spots
+          at $60^\circ$ and polar spots with high latitude variance.
+          It may thus be advisable to adopt a prior that favors small values
+          of $\sigma_\phi$, such as the common inverse gamma prior for the
+          variance.
+    \item \emph{Limb darkening matters!} The null space is extremely sensitive
+          to limb darkening (\S\ref{sec:limbdark}). It is therefore
+          extremely important to model it correctly; otherwise, there may
+          be substantial bias in the inferred spot parameters.
+    \item \emph{Careful with the data}. In general, covariances can be very
+          hard to estimate from noisy data. This makes it especially important
+          to ensure one is correctly modeling the noise. When applying our GP
+          to model real data, we recommend the usual
+          inference practices of clipping outliers,
+          modeling a latent white noise (jitter) term, and modeling
+          a small latent additive offset term to minimize the risk of
+          bias in the posteriors of interest.
+    \item \emph{Careful with the sample selection}. When performing an ensemble
+          analysis of stellar variability, it is tempting to only analyze light curves
+          that show variability in the first place. \emph{This is extremely
+              dangerous}, since the lack of variability could simply be due to low
+          inclinations. It is extremely important to ensure the sample selection step does
+          not introduce bias. If there is reason to believe there are two
+          distinct populations within an ensemble---say, a population of active
+          stars and a population of quiet (spotless) stars---we strongly recommend
+          the use of a Gaussian mixture model.
 \end{enumerate}
@@
               the spot contrast or the total number of spots}. These two quantities
-          are degnerate with each other. While the contrast can be very weakly
-          constrained with lots of data, the total number of spots is not
-          an observable in relative single-band photometry. Knowledge of the
-          unspotted photospheric brightness (or temperate) can break this
+          are perfectly degnerate with each other in relative single-band photometry.
+          Knowledge of the
+          unspotted photospheric brightness (or temperature) can break this
           degeneracy, so analyses of multi-band light curves can in principle
@@
 \begin{proof}{test_ephi}
+    \label{eq:ephi}
     \mathbf{e}_\phi

### 2021-01-14 tex/ms.tex
more words

@@
     %
-    We show that, when employed in an inference setting, our GP is unbiased,
+    We show that our GP is unbiased,
     allowing one to correctly infer physical parameters of interest from one
@@
 \label{sec:other-work}
+
+The core idea behind the methodology presented in this paper---to compute the
+Gaussian approximation to an intractable multidimensional distribution in order
+to obtain a likelihood function for inference---is not new. Although the
+method likely goes by different names in different fields, it is a popular
+technique particularly in the field of ecological population dynamics,
+where it is referred to as the synthetic likelihood \citep[SL;][]{Wood2010}
+or Bayesian synthetic likelihood \citep[BSL;][]{Price2018} method. In many
+ecological systems, population growth is a chaotic process; observations of the
+size of a population over time can be dominated by steep spikes and drops
+in the population that occur due to sudden, random environmental pressure.
+While population growth can be forward modeled with ease, it is very
+difficult to use forward models to constrain basic growth parameters in an
+inference setting, since that requires marginalization over the extremely
+nonlinear noice processes. As a way around this, \citet{Wood2010} introduced
+the SL method, in which, conditioned on a set of parameters of interest
+$\pmb{\theta}$, one computes the forward model $f(\pmb{\theta})$ many times
+under different realizations of the noise, and adopts the sample mean
+and sample covariance as the mean and covariance of the Gaussian likelihood
+function $p(f \, | \, \pmb{\theta})$. \citet{Wood2010} showed that,
+provided the number of forward model samples is large enough, this
+``synthetic'' likelihood allows one to infer the population growth parameters
+efficiently and without bias.
+
+The method presented in this paper may be thought of as a synthetic likelihood
+method in the limit of an infinite number of forward model samples. Unlike
+\citet{Wood2010}, whose method determines the mean and covariance of the
+distribution of $f$ conditioned on $\pmb{\theta}$ by sampling, we are able
+to actually compute the mean and covariance of $f$ \emph{in closed form}.
+Unlike traditional SL methods, which are inherently noisy, our method employs
+the \emph{exact} Gaussian approximation to the likelihood function.
+
+Our GP is also closely related to techniques commonly employed in models
+of the cosmic microwave background. In particular, it is a type of
+Gaussian random field (GRF) on the sphere \citep{Wandelt2012} \xxx{...}
+
+Finally, our method is also related to various families of approximate inference, such as
+variational inference (VI), in which a multivariate Gaussian is used to
+approximate the \emph{posterior} distribution, or to approximate Bayesian
+computation (ABC) \xxx{...}
+
 \xxx{TODO: Discuss \citet{Basri2020}, how we agree and disagree. Compare to Brett's
-    paper using ABC to measure spot sizes as a function of spectral type. Discuss
-    Simon Wood's paper in ecology on synthetic likelihoods. Talk about similarities
-    to Gaussian Random Fields in CMB studies.}
+    paper using ABC to measure spot sizes as a function of spectral type.}
 

### 2021-01-15 tex/ms.tex
more text

@@
 
+\subsubsection{GPs, GRFs, and approximate inference}
+
 The core idea behind the methodology presented in this paper---to compute the
@@
 to actually compute the mean and covariance of $f$ \emph{in closed form}.
-Unlike traditional SL methods, which are inherently noisy, our method employs
+While traditional SL methods are inherently noisy, our method employs
 the \emph{exact} Gaussian approximation to the likelihood function.
@@
 Our GP is also closely related to techniques commonly employed in models
-of the cosmic microwave background. In particular, it is a type of
-Gaussian random field (GRF) on the sphere \citep{Wandelt2012} \xxx{...}
-
-Finally, our method is also related to various families of approximate inference, such as
+of the cosmic microwave background (CMB). In particular, it is a type of
+Gaussian random field (GRF) on the sphere, which are frequently used to
+model perturbations in the CMB \citep{Wandelt2012}. In general, however,
+GRFs used in cosmology are isotropic: when expressed in the spherical
+harmonic basis, their covariance matrix is diagonal and admits a
+representation as a (one-dimensional) power spectrum. Our GP, in contrast,
+is anisotropic in
+the polar coordinate (i.e., the latitude) by construction.
+
+Our method is also related to various families of approximate inference, such as
 variational inference (VI), in which a multivariate Gaussian is used to
-approximate the \emph{posterior} distribution, or to approximate Bayesian
-computation (ABC) \xxx{...}
-
-\xxx{TODO: Discuss \citet{Basri2020}, how we agree and disagree. Compare to Brett's
-    paper using ABC to measure spot sizes as a function of spectral type.}
+approximate the \emph{posterior} distribution \citep[e.g,][]{Blei2016},
+or to approximate Bayesian
+computation (ABC), in which an (often intractable) likelihood function is
+replaced with an approximation computed from simulations from the prior
+\citep[e.g.,][]{Beaumont2019}.
+
+\xxx{Finally, talk about \citet{Perger2020} on interpretable GPs}.
+
+\subsubsection{Spots and stellar variability}
+
+\xxx{%
+    \citet{Morris2020} uses the ``smoothed amplitude'', the difference between the
+    maximum and minimum flux in a de-trended light curve, as the summary statistic
+    in ABC. Try to infer spot coverage fractions. But this is \emph{not an
+        observable}. Note that his contrast $c'$ is defined in the opposite sense
+    as ours (our $c = 1 - c'$.)
+}
+
+\xxx{%
+    \begin{itemize}
+        \item Discuss \citet{Basri2020}, how we agree and disagree.
+        \item Discuss \citep{Morris2020} on ABC to measure spot sizes as a function of spectral type.
+        \item Discuss \citep{Jackson2013}.
+        \item Discuss \citep{Mazeh2015}.
+    \end{itemize}
+}
 

### 2021-01-17 tex/ms.tex
more text

@@
 The first point is illustrated in the toy example in
-Figure~\ref{fig:mean_normalization}. As it has been pointed out before \citep{?},
-we will not spend too much time discussing it, other than to say that the
+Figure~\ref{fig:mean_normalization}. \xxx{As it has been pointed out before \citep{?},
+    we will not spend too much time discussing it,} other than to say that the
 flux normalization introduces a degeneracy between the total spot coverage
 of a star and the contrast of any individual feature on the surface.
-While this is a formally perfect degeneracy for a single light curve, we can
+\xxx{While this is a formally perfect degeneracy for a single light curve, we can
     again resort to ensemble statistics across many light curves to break it;
     we discuss this further in \S\ref{sec:breaking_baseline_degeneracy}.
+    NO WE CAN'T. Talk about this in more detail.}
 
@@
             and the contrast of any individual feature on the surface.
+            As a consequence, the total spot coverage of a star cannot be
+            inferred from relative photometry.
             \label{fig:mean_normalization}
@@
 choice of pixels on a grid, or perhaps a localized wavelet basis---would be better
-at modeling small spots. While this is probably true, a different choice of
-basis would not admit the same analytic solutions to the integrals needed to
-compute the GP covariance. And it is precisely the ability to compute these
-integrals analytically (Appendix~\ref{sec:integrals}) that makes the evaluation
-of the marginal likelihood tractable; it is difficult to envision an approach
-that uses a \emph{local} basis and simultaneously enjoys the same
-analyticity under integration.
-
-In principle, small spots can be modeled with negligible ringing by simply
+at modeling small spots. While this is probably true, it may be quite
+difficult to find closed-form expressions for the expectation integrals
+(\S\ref{sec:integrals}) that make the GP covariance evaluation tractable.
+
+One option is to bypass the computation of the covariance on the surface of
+the star and to write down an expression for the flux directly in terms of
+the properties of a star spot. Circular spots of uniform intensity can be
+modeled as spherical caps, which are simply segments of ellipses when
+projeted onto the sky. It is possible to express in closed form the projected area
+they cover---and hence their contribution to the flux $\mathbbb{f}$---even in the case
+where part of the spot is on the hemisphere facing away from the observer.
+This was done in \citet{Luger2017}, who solved this problem in closed form
+(see \S3.3.2 and Appendix A.3 in that paper). Computing the GP is then a matter of
+integrating $\mathbbb{f}$ and $\mathbbb{f}\,\mathbbb{f}^\top$, weighted by
+the hyperparameter PDFs, as we did in \S\ref{sec:gp}.
+However, the expression for
+$\mathbbb{f}$ involves square roots and arctangents of functions of the spot
+latitude and longitude
+\citep[c.f. Equation~40 in][]{Luger2017},
+so computing these integrals in closed form is
+likely to be very difficult. Even if a closed form solution can be found,
+incorporating limb darkening (which we argued can be extremely important)
+poses an even greater challenge. It is likely that several simplifications
+must be made in order to make this approach tractable. This was done to some
+extent in \citet{Morris2020}, who derived a closed form expression for the flux by
+ignoring certain projection effects, such
+as the self-occultation of large spots by the limb of the star, and neglected
+variations in limb darkening within spots. Such a model could admit a closed-form
+solution to the GP covariance and may be better at capturing the effects
+of small spots, at the expense of the ability to model larger spots.
+
+In principle, small spots can be modeled under our current
+approach with negligible ringing by simply
 increasing the degree of the spherical harmonic expansion. As we discuss in
@@
 unstable for $l \gtrsim 15$, so doing so would require a reparametrization
-of the equations in the Appendix. However, as we showed in
+of the equations in the Appendix. Importantly, however, as we showed in
 \S\ref{sec:calibration}, the current implementation of the GP is suitable
@@
 
-We showed in \S\ref{sec:calibration-inference} that there exists a
+We showed in \S\ref{sec:calibration-inference} that in relative photometry there exists a
 strong degeneracy between the contrast of individual spots $c$ and
-the total number of spots $n$ on a star. For relative photometry, in
-which the true baseline level (\S\ref{sec:baseline}) is not known,
-this degeneracy is perfect: relative photometry cannot tell us anything
-about the total spottiness of a stellar surface. Inspection of the
+the total number of spots $n$ on a star. Inspection of the
 dependence of the GP on $c$ and $n$ (c.f. Appendix~\ref{sec:integrals}) reveals
@@
 \begin{align}
-    \mu      & \propto c \, n
+    \mu      & = 1 - c \, n \, g (I, \mathbf{u}, r, \mu_\phi, \sigma_\phi)
     \nonumber                                                              \\
-    \sigma^2 & \propto c^2 \, n
+    \sigma^2 & = c^2 \, n \, h (I, \mathbf{u}, r, \mu_\phi, \sigma_\phi)
     \quad,
@@
 %
-respectively. If our observations were sensitive to both the true mean and
+for some function $g$ and some function $h$ of the inclination,
+the limb darkening coefficients, the spot radius, and the spot latitude
+parameters.
+If our observations were sensitive to both the true mean and
 variance of stellar light curves,
 we could uniquely infer $c$ and $n$. However, as we discussed in \S\ref{sec:baseline},
-to very good approximation, relative photometry can only tell us about the ratio
-%
-\begin{align}
-    \frac{\sigma^2}{\mu^2} \propto \frac{c}{n}
+to good approximation, relative photometry can only tell us about the \emph{ratio}
+%
+\begin{align}
+    \frac{\sigma^2}{\mu^2}
+     & \propto \frac{c^2 n}{(1 - c n g)^2}
     \quad,
@@
 %
-i.e., the ratio of the spot contrast to the total number of spots.
+where $g$ is a complicated function of the (unknown) stellar and spot parameters;
+relative photometry cannot tell us about the values of $c$ and $n$ independently.
+
 The reason for this is that the relationship between the amplitude of the
@@
 
-Fortunately, it may be possible to break this degeneracy (at least to some extent)
+A direct consequence of the baseline degeneracy is that it is not in general
+possible to constrain the total spot coverage of a star from relative,
+single-band photometry. Assuming no spot overlap, the total spot coverage $f_S$ is simply the (average) area of a spot
+times the total number of spots divided by the total area of the sphere, which
+may be expressed as
+%
+\begin{proof}{test_fS}
+    f_S = \frac{1}{2}(1 - \cos r)n
+    \quad,
+\end{proof}
+%
+recalling that $r$ is the \emph{angular} radius of the spot.
+While $r$ may be constrained from the covariance structure of the data
+(\S\ref{sec:calibration-inference}), $n$ cannot.
+
+The arguments above are based on the \emph{Gaussian approximation} to the
+distribution of stellar light curves. As we discussed in \S\ref{sec:why-calibrate},
+the true distribution has higher order moments that we do not consider here.
+It is possible in principle that these moments could encode information to
+break the $c-N$ degeneracy, but these are in general much more difficult
+to constrain from the data. And even if this were the case, the fact remains
+that degeneracies like the polar spot effect are fundamental: recall
+Figure~\ref{fig:mean_normalization}, in which the two stars
+have identical relative light curves, but very different spot coverage
+fractions.
+
+Fortunately, it may be possible to break these degeneracies (at least to some extent)
 by collecting photometric data in multiple wavelength bands, an idea that has
@@
 \xxx{%
-    \citet{Morris2020} uses the ``smoothed amplitude'', the difference between the
+    \begin{itemize}
+        \item Discuss \citet{Morris2020} on ABC to measure spot sizes as a
+              function of spectral type.
+              Brett uses the ``smoothed amplitude'', the difference between the
               maximum and minimum flux in a de-trended light curve, as the summary statistic
-    in ABC. Try to infer spot coverage fractions. But this is \emph{not an
-        observable}. Note that his contrast $c'$ is defined in the opposite sense
+              in ABC to try to infer spot coverage fractions. But this is \emph{not an
+                  observable} as we discussed above. Poking around in his code reveals he
+              assumes $n=3$ spots for all stars, so his results are extremely
+              prior-dominated. He also fixes limb darkening and assumed a uniform
+              (not isotropic) prior on the inclinations. But there's probably an
+              easy way to turn his ABC code into a GP, since his expression for the flux
+              is closed-form and (probably) easily integrable.
+              Also note that his contrast $c'$ is defined in the opposite sense
               as ours (our $c = 1 - c'$.)
-}
-
-\xxx{%
-    \begin{itemize}
-        \item Discuss \citet{Basri2020}, how we agree and disagree.
-        \item Discuss \citep{Morris2020} on ABC to measure spot sizes as a function of spectral type.
-        \item Discuss \citep{Jackson2013}.
-        \item Discuss \citep{Mazeh2015}.
+        \item Discuss \citet{Basri2020}, how we agree and disagree.  Gibor has some very
+              good points about the null space, but he is overly pessimistic about what
+              can and can't be learned from light curves. He does not consider the
+              power of ensemble analysis, and claims there's just no way to disentangle
+              spot evolution from differential rotation.
+        \item Discuss \citet{Jackson2013}.
+        \item Discuss \citet{Mazeh2015}.
     \end{itemize}
@@
           of $\sigma_\phi$, such as the common inverse gamma prior for the
-          variance.
+          variance. \xxx{Or an isotropic prior on $\mu_\phi$.}
     \item \emph{Limb darkening matters!} The null space is extremely sensitive
@@
     \item \textbf{Relative single-band photometry cannot uniquely constrain
-              the spot contrast or the total number of spots}. These two quantities
+              the spot contrast, the total number of spots, or the total spot
+              coverage}. These two quantities
           are perfectly degnerate with each other in relative single-band photometry.
@@
     \mathbbb{y}(\mathbbb{x}) =
+    -
     \sum_{i=0}^{\mathbb{n}-1}
@@
      & \equiv
+    -
     \int
@@
      & \equiv
+    -
     \mathbf{e}_\lambda \,
@@
     \label{eq:ec}
-    \mathbf{e}_c & = c \, \mathbf{e}_\lambda
+    \mathbf{e}_c & = - c \, \mathbf{e}_\lambda
     \\

### 2021-01-17 tex/figures/mean_normalization.py

@@ -92,17 +92,42 @@ ax[1].plot(t - 0.5, 1e3 * (flux2 / flux2.max() - 1), "C1--", lw=2.5)
 ax[1].set_ylabel("observed flux [ppt]", fontsize=20)
 ax[1].set_xlabel("rotations", fontsize=20)
 
+# Compute the spot coverage: the fraction of the
+# surface that is darker than one-tenth the peak
+# spot darknesss
+moll1 = star1.map.render(projection="moll").flatten()
+moll1 = moll1[np.isfinite(moll1)]
+moll1[moll1 > -0.1] = 0
+moll1[moll1 < -0.1] = 1
+fs1 = np.count_nonzero(moll1) / np.size(moll1)
+
+moll2 = star2.map.render(projection="moll").flatten()
+moll2 = moll2[np.isfinite(moll2)]
+moll2[moll2 > -0.1] = 0
+moll2[moll2 < -0.1] = 1
+fs2 = np.count_nonzero(moll2) / np.size(moll2)
+
 # Legend
 ax1a = plt.axes([0.9, 0.55, 0.2, 0.2])
+for spines in ["top", "right", "left", "bottom"]:
+    ax1a.spines[spines].set_visible(False)
+ax1a.set_xticks([])
+ax1a.set_yticks([])
+ax1a.set_title("{:.0f}% coverage".format(fs1 * 100))
 star1.map.show(ax=ax1a, norm=norm)
-ax2a = plt.axes([0.9, 0.25, 0.2, 0.2])
+ax2a = plt.axes([0.9, 0.23, 0.2, 0.2])
+for spines in ["top", "right", "left", "bottom"]:
+    ax2a.spines[spines].set_visible(False)
+ax2a.set_xticks([])
+ax2a.set_yticks([])
+ax2a.set_title("{:.0f}% coverage".format(fs2 * 100))
 star2.map.show(ax=ax2a, norm=norm)
 ax1b = plt.axes([1.1, 0.55, 0.07, 0.2])
 ax1b.set_ylim(0, 1)
 ax1b.set_xlim(0, 1)
 ax1b.plot([0, 1], [0.5, 0.5], "C0-", lw=3)
 ax1b.axis("off")
-ax2b = plt.axes([1.1, 0.25, 0.07, 0.2])
+ax2b = plt.axes([1.1, 0.23, 0.07, 0.2])
 ax2b.set_ylim(0, 1)
[…]

### 2021-01-18 tex/ms.tex
new fig

@@
 \label{sec:intro}
+
+\begin{figure}[t!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/degeneracies.pdf}
+        \oscaption{degeneracies}{%
+            The fundamental limitations of the mapping problem. Each row
+            corresponds to a stellar surface with a different number of
+            dark spots seen at various phases at an inclination $I=60^\circ$.
+            The bottom panel shows the light curves of each of these stars.
+            All six light curves are indistinguishable from each other, even
+            at infinite signal to noise. See text for details.
+            \label{fig:degeneracies}
+        }
+    \end{centering}
+\end{figure}
 

### 2021-01-18 tex/ms.tex
finished discussion

@@
 \begin{proof}{test_fS}
-    f_S = \frac{1}{2}(1 - \cos r)n
+    \label{eq:fS}
+    f_S = \frac{1}{2}\left(1 - \left<\cos r\right>\right)n
     \quad,
@@
 
-\xxx{Finally, talk about \citet{Perger2020} on interpretable GPs}.
+Finally, the methodology developed here is closely related to that in
+\citet{Perger2020}, who studied the effect of different star spot
+configurations on the autocorrelation and covariance of stellar
+radial velocity (RV) measurements. The authors of that study compared the
+performance of various commonly used quasi-periodic kernels when applied
+to synthetic RV datasets, arguing that a new four-parameter quasi-periodic
+cosine kernel (QPC) can better capture the variability due to star spots.
+However, their study was empirical and related spot configurations to their
+effect on the covariance structure of the data primarily in a qualitative
+fashion. Their QPC kernel is a function of two interpretable hyperparameters
+(the rotation period and a spot timescale) as well as two amplitudes, which
+are not explicitly related to physical spot properties. Our GP, in contrast,
+is built from the ground up such that all of its hyperparameters directly
+correspond to physical spot properties, allowing one to use it in star spot
+inference (not just marginalization) problems. While the methodology presented
+here applies to photometry, it is possible to extend it to model RV
+datasets as well; we discuss this in \S\ref{sec:extensions}.
 
@@
 
-\xxx{%
-    \begin{itemize}
-        \item Discuss \citet{Morris2020} on ABC to measure spot sizes as a
-              function of spectral type.
-              Brett uses the ``smoothed amplitude'', the difference between the
-              maximum and minimum flux in a de-trended light curve, as the summary statistic
-              in ABC to try to infer spot coverage fractions. But this is \emph{not an
-                  observable} as we discussed above. Poking around in his code reveals he
-              assumes $n=3$ spots for all stars, so his results are extremely
-              prior-dominated. He also fixes limb darkening and assumed a uniform
-              (not isotropic) prior on the inclinations. But there's probably an
-              easy way to turn his ABC code into a GP, since his expression for the flux
-              is closed-form and (probably) easily integrable.
-              Also note that his contrast $c'$ is defined in the opposite sense
-              as ours (our $c = 1 - c'$.)
-        \item Discuss \citet{Basri2020}, how we agree and disagree.  Gibor has some very
-              good points about the null space, but he is overly pessimistic about what
-              can and can't be learned from light curves. He does not consider the
-              power of ensemble analysis, and claims there's just no way to disentangle
-              spot evolution from differential rotation.
-        \item Discuss \citet{Jackson2013}.
-        \item Discuss \citet{Mazeh2015}.
-    \end{itemize}
-}
+Recently, \citet{Morris2020} used \emph{Kepler}, \emph{K2}, and \emph{TESS}
+light curves to derive a relationship between stellar age and spot
+coverage. Because of the intractability of the marginal likelihood function,
+that study used an approximate Bayesian computation (ABC) method to
+infer spot properties from a large ensemble of stars. \citet{Morris2020}
+developed a fast, approximate forward model for light curves of spotted stars
+\citep[\textsf{fleck};][]{Morris2020b}, which they used to generate a large
+number of prior samples for different values of the spot radii, contrasts,
+and latitude distributions. For each collection of samples generated from a
+given set of hyperparameters, \citet{Morris2020} computed the distribution of
+the ``smoothed amplitude'',
+the peak-to-trough difference of the (normalized, de-trended) light curve.
+This distribution was then compared to the distribution of observed
+smoothed amplitude values among stellar clusters of different ages
+within an ABC algorithm, yielding approximate
+posterior distributions for the hyperparameters as a function of
+stellar age. \citet{Morris2020} found that the spot coverage $f_S$ as a
+function of age is well modeled by a simple power law, decreasing from
+${\sim}10\%$ for the youngest (${\sim}10$ Myr) stars
+to less than $1\%$ for the oldest (${\sim}5$ Gyr) stars.
+%
+While these results broadly agree with the expectation that stellar
+magnetic activity decreases over time, our work strongly suggests
+that these results are predominantly prior-dominated. As we discussed
+at length
+in \S\ref{sec:breaking_baseline_degeneracy} (and found in our inference
+runs in \S\ref{sec:calibration-inference}, as well as in the toy
+model presented in Figure~\ref{fig:mean_normalization}), the fractional
+spot coverage $f_S$ is not an observable in single-band photometry.
+%
+This is because the expression for $f_S$ (Equation~\ref{eq:fS}) depends on
+two quantities: the average spot radius $r$, which we showed can be
+constrained, and the total number of spots $n$, which
+we showed \emph{cannot} be constrained from single-band light curves.
+%
+In fact, \citet{Morris2020} assumed $n=3$ for simplicity when doing
+posterior inference. We therefore urge care in interpreting those results,
+as it is very likely this assumption does not hold across the large
+range of spectral types and stellar ages considered in that study.
+
+Nevertheless, the core idea in \citet{Morris2020}
+\citep[and in related studies such as][]{Jackson2013}
+is very similar
+to that advocated here: the use of ensemble analyses to constrain population-level
+parameters when invidual datasets are not sufficiently constraining.
+Our work builds on that of \citet{Morris2020} by deriving a closed form
+solution to the likelihood function (as opposed to a sample-based
+likelihood-free inference algorithm) and by harnessing the covariance
+structure of the data when doing inference (as opposed to relying solely
+on the amplitude of the data).
+
+Another recent paper relevant to our work is that of \citet{Basri2020}, who
+investigated the information content of stellar light curves, exploring what
+can and cannot be learned about star spot configurations from individual
+light curves. That paper strongly urges against the common practice
+of interpreting
+light curves with one or two dips as originating from one or two spots,
+respectively, a point we strongly agree with (c.f. Figure~\ref{fig:degeneracies}).
+It also reinforces our point about the additional degeneracies introduced
+by the unknown baseline inherent to single-band photometry.
+%
+Finally, \citet{Basri2020} present a large suite of forward models of light curves of
+spotted stars, which they use to discuss the (complicated) dependence of
+various light curve metrics on
+the physical spot parameters used to generate the data. They conclude that it
+is not possible to uniquely relate these metrics to the underlying star spot
+configuration. While we agree this is the case for individual stars, our work
+stresses that it is possible to break many of these degeneracies with
+ensemble analyses.
+%
+\citet{Basri2020} also conclude it is not in general possible to
+uniquely disentangle differential rotation from spot evolution when their
+timescales are comparable; nor is it possible to confidently measure a rotation
+period when the evolution timescale is very short. However, their study
+relies on the effect these processes have on simple light curve metrics,
+which are almost certainly not sufficient statistics of the data. Inference
+that takes into account the full covariance structure of the data, while
+considering large ensembles of light curves, could in principle break
+this degeneracy. While we do not explicitly model differential rotation in
+this paper, it will be the subject of a future paper in this series.
 
@@
           their presence. This can be seen in Figure~\ref{fig:calibration_hilat},
-          in which the model cannot distinguish between localized spots
+          in which the model cannot distinguish between spots localized
           at $60^\circ$ and polar spots with high latitude variance.
@@
           of $\sigma_\phi$, such as the common inverse gamma prior for the
-          variance. \xxx{Or an isotropic prior on $\mu_\phi$.}
+          variance. Alternatively, one could place an isotropic prior on
+          the latitude (with density proportional to $\cos\mu_\phi$)
+          to downweight very high latitude spots.
     \item \emph{Limb darkening matters!} The null space is extremely sensitive

### 2021-01-19 tex/ms.tex
finished discussion

@@
 The first point is illustrated in the toy example in
-Figure~\ref{fig:mean_normalization}. \xxx{As it has been pointed out before \citep{?},
-    we will not spend too much time discussing it,} other than to say that the
+Figure~\ref{fig:mean_normalization}. As it has been pointed out before
+\citep[e.g.,][]{Basri2018},
+we will not spend too much time discussing it, other than to say that the
 flux normalization introduces a degeneracy between the total spot coverage
 of a star and the contrast of any individual feature on the surface.
-\xxx{While this is a formally perfect degeneracy for a single light curve, we can
-    again resort to ensemble statistics across many light curves to break it;
-    we discuss this further in \S\ref{sec:breaking_baseline_degeneracy}.
-    NO WE CAN'T. Talk about this in more detail.}
+The two stars in the figure have different numbers of spots, different
+spot contrasts, and different total spot coverage. In ``absolute'' units, in which
+the flux is measured as a fraction of the flux of an unspotted star, the two
+light curves are distinct (top panel). However, we can only measure these light
+curves up to an unknown multiplicative factor. Normalizing these light curves
+to the ``continuum'' level yields the curves in the bottom panel, which are
+\emph{perfectly} indistinguishable from each other. Normalization to the mean,
+median, or any other statistic of the light curve does not help break this
+degeneracy, which is inherent to relative photometry. While one could instead \emph{fit}
+for the normalization as a latent variable, there is no information in the data
+that can help break it, so the result will be entirely determined by the prior.%
+\footnote{%
+In principle, the maximum level of a light curve could set a lower limit on
+the location of the baseline. However, this applies only if the surface is
+known to be made up of \emph{exclusively} dark spots. The presence of bright
+spots (faculae), which are common on the Sun, make it effectively impossible
+for one to infer the baseline from single-band relative photometry alone.
+}
+We discuss these issues further in \S\ref{sec:breaking_baseline_degeneracy}.
 

### 2021-01-19 tex/ms.tex
new figure, as if we needed more

@@
     GPs excel in particular at marginalization
-    of the stellar signal in cases where the stellar variability is treated
+    of the stellar signal in cases where the
+    variability due to star spots rotating in and out of view is treated
     as a nuisance, such as in exoplanet transit modeling.
-    However, in cases where the stellar signal is of primary interest,
-    GPs have thus far been less useful, as it is often difficult to
+    However, in cases where the star spot signal is of primary interest,
+    GPs have thus far been less useful: it is often difficult to
     relate their hyperparameters to the physical properties of interest, such
-    as the size, contrast, and latitudinal distribution of star spots.
+    as the size, contrast, and latitudinal distribution of the spots.
     Instead, it is common practice to explicitly model the effect
     of individual star spots on the light curve and attempt to infer their
-    properties via optimization or posterior inference, a process that is
-    ill-posed, plagued by degeneracies, and computationally intractable when
+    properties via optimization or posterior inference. Unfortunately,
+    this process is
+    degenerate, ill-posed, and often computationally intractable when
     applied to stars with more than a few spots and/or to ensembles of many
@@
     mean and covariance of a Gaussian process model that describes
-    the light curve of a rotating star conditioned on a given distribution of
+    the light curve of a rotating, evolving stellar surface
+    conditioned on a given distribution of
     star spot sizes, contrasts, and latitudes.
@@
     magnetic activity of stars from both light curves and radial velocity (RV)
-    measurements, as well as for modeling correlated noise in both transiting
+    measurements, as well as for robustly modeling correlated noise in both transiting
     and RV exoplanet searches.
@@
 Over the past decade, Gaussian processes (GPs) have gained traction as
-a leading tool for modeling correlated signals in astrophysical timeseries.
+a leading tool for modeling correlated signals in astrophysical timeseries
+\citep[e.g.,][]{RasmussenWilliams2005,BrewerStello2009,Luger2016,ForemanMackey2017,Angus2018}.
 Despite whatever mystique the words ``Gaussian process'' may evoke, a GP
@@
 the mean of the process and a $(K \times K)$
-matrix $\pmb{\Sigma}$ describing its covariance,
+matrix $\pmb{\Sigma}$ characterizing its covariance,
 a GP is fully specified by a mean
@@
 data are correlated, $\Gamma$ is a dimensionless scale parameter, and $P$ is
-a period. Once a kernel has been chosen, these scalar quantities fully
+a period.
+Once a kernel has been chosen, these scalar quantities fully
 specify the covariance of the Gaussian process. GP-based inference
@@
 Which brings us, finally, to an important drawback of GPs when applied
-to modeling stellar variability. While it is straightforward to derive
+to modeling stellar spot-induced variability. While it is straightforward to derive
 posterior constraints on the hyperparameters of a GP, it is usually
 not obvious what those constraints actually tell us about the star
-or the actual source of the variability. While the GP period $P$
+or its surface. While the GP period $P$
 may sometimes be interpreted as the rotational period of the star,%
@@
 the relationship between quantities like the amplitude $\sigma^2$
-and the timescale $\tau$ and physical quantites of interest---such
-as the properties of the star spots---is much fuzzier.
+and the timescale $\tau$ and physical properties of the spots is much fuzzier.
 It may be tempting to interpret $\sigma^2$ as some measure of the spot
@@
 and temporal evolution of each of the $n$ spots, where $n$ is also
-a free parameter. In addition to this being a numerically challenging
+a free parameter \xxx{(CITE ME)}. In addition to this being a numerically challenging
 problem for all but the smallest values of $n$, the problem is
@@
 are in the \emph{null space}, the set of surface features that have identically
-zero effect on the light curve.
+zero effect on the light curve \citep[e.g.,][]{Cowan2013,Luger2019}.
 The underdetermined nature of the stellar mapping problem is therefore fundamental: it
@@
 However, it is hardly ever the case that this is actually our end goal.
-After all, physics predicts the properties of stellar surfaces at a fairly
+After all, physics can be used to predict properties of stellar surfaces at a fairly
 high level: i.e., typical spot sizes, active spot latitudes, or approximate
-timescales on which spots evolve. We are hardly ever interested in the
+timescales on which spots evolve
+\citep[e.g.,][]{Schuessler1996,Solanki2006,Cantiello2019}.
+We are hardly ever interested in the
 \emph{particular} properties of a \emph{particular} spot, as we wouldn't really
@@
 one would observe as the star rotates about a fixed axis.
-If the stellar surface intensity is described by the spherical harmonic
-coefficient vector $\mathbf{y}$, the flux is given by
+If the stellar surface intensity is described by a spherical harmonic
+coefficient vector $\mathbf{y}$ (up to a certain degree $l_\mathrm{max}$),
+the flux is given by
 %
@@
 %
-are the mean and covariance of the distribution over spherical harmonic coefficients.
+are the mean and covariance of the distribution over spherical harmonic coefficient
+vectors $\mathbbb{y}$.
 The bulk of the math in this paper (Appendix~\ref{sec:integrals})
@@
 %
-The PDFs for the spot radius, the spot contrast, and the number
-of spots are delta functions centered at $r$, $c$, and $n$, respectively
+For simplicity, the PDFs for the spot radius, the spot contrast, and the number
+of spots are chosen to be delta functions centered at $r$, $c$, and $n$, respectively
 (Appendices \ref{sec:size} and \ref{sec:contrast}),
 while
-the PDF for the latitude $\phi$ of the spots is a Beta distribution in
-$\cos\phi$ with (normalized) shape parameters $a$ and $b$,
+the PDF for the latitude $\phi$ of the spots is chosen to be a Beta distribution in
+$\cos\phi$ with (normalized) parameters $a$ and $b$,
 which have a one-to-one correspondence to the mode $\mu_\phi$ and
@@
     for median-normalized light curves, since the distribution of the GP sample median
-    is usually close to the distribution of sample mean.}
+    is usually close to the distribution of the sample mean.}
 Since we are dividing the
 flux by the mean, one might imagine that we could simply divide the
-covariance matrix by the square of the mean and call it a day. However,
+covariance matrix by the square of the mean. However,
 this is incorrect, since we must be careful to distinguish between the
@@
 a draw from the GP is itself normally distributed with a variance that scales
-with the GP variance.
+with the GP variance.%
+\footnote{
+    Importantly, the sample mean and process mean will be different even in the
+    absence of measurement error! In other words, the mean flux of a given
+    star (i.e., the sample mean)
+    will in general be different from the mean flux \emph{across all stars} with similar
+    surface properties (the process mean).
+}
 
@@
 error in the approximation to the covariance is less than $10^{-6}$.
-In cases where the light curve variability exceeds a few percent, we recommend
+In cases where the light curve variability exceeds about ten percent, we recommend
 modeling the multiplicative amplitude in each light curve as a latent variable, as discussed
@@
 
-\begin{figure}[t!]
+\begin{figure}[p!]
     \begin{centering}
@@
             original covariance matrix, the covariance of the normalized process
-            is discernibly non-stationary.
+            is discernibly non-stationary (see Figure~\ref{fig:nonstationarity}).
             \label{fig:normgp}
+        }
+    \end{centering}
+\end{figure}
+
+\begin{figure}[p!]
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/nonstationarity.pdf}
+        \oscaption{nonstationarity}{%
+            An example of why normalized light curves are non-stationary.
+            The top panel shows $1{,}000$ samples from a unit-mean sinusoid with
+            an amplitude of 10\% and a period of 10 days, much longer than the
+            1 day observation baseline. The bottom panel shows the same light curves,
+            each normalized to its own mean. Because the mean tends to be near
+            the center of the observation window, points near $t=0.5$ are driven
+            to values very close to unity, while points near the edges have much
+            larger scatter.
+            \label{fig:nonstationarity}
         }
@@
 a light curve is now slightly dependent on their phases.
+
+To understand why normalized light curves are non-stationary, it is useful to
+consider the
+toy example in Figure~\ref{fig:nonstationarity}, in which we draw
+$1{,}000$ samples from a sinusoid with period ten times longer than the
+observational baseline. In this limit, each light curve is approximately
+linear, which causes its mean value to roughly coincide with the midpoint of the
+observation window. Division by the mean value results in points near the
+midpoint being driven to unity and points near the edges (whose values differ
+the most from the mean) to be driven to both large and small values. The
+variance across all samples is now distinctly phase-dependent.
+%
+In general, the non-stationarity will be most extreme in cases where the
+observation window is shorter than or comparable to one period, and
+asymptotically vanishes in the limit that a large number of cycles are
+observed.
 
@@
 the GP as a prior in the discussion above, it is important to appreciate
-that it even works in the first place! Despite some issues, the fact that
-the surface map samples \emph{actually look like maps of spotted stellar surfaces}
-is fairly amazing. A Gaussian process, after all, is a non-parameteric
+that it even works in the first place!
+%Despite some issues, the fact that
+%the surface map samples \emph{actually look like maps of spotted stellar surfaces}
+%is fairly amazing.
+A Gaussian process, after all, is a non-parameteric
 \emph{process} describing a smooth and continuous function only via its
@@
 problems in astronomy \citep[such as in models of the cosmic microwave
-    background][]{?}, the process is assumed to be isotropic,
+    background; e.g.,][]{Wandelt2012}, the process is assumed to be isotropic,
 with no preferred angular direction.
@@
             $5\times 10^{4}$ stellar surface maps drawn from a certain
-            star spot distribution (black points and contours).
+            distribution of discrete circular spots (black points and contours).
             At $l=15$ there are 256 coefficients in total;
@@
 of parameter space, it is certainly a poor approximation in others.
-In all fairness, the vast majority of the spherical harmonic coefficients
-have marginal distributions that \emph{look} to be very Gaussian; however,
-those that do not could in principle lead to errors and biases when using
-the Gaussian approximation to the likelihood function.
+%
+In this section, we will show that, fortunately, the non-Gaussianity of the
+distribution is not in general an issue when doing inference with our GP,
+as the resulting posteriors are unbiased.
 
@@
 $\nicefrac{180^\circ}{30} = 6^\circ$ across, but small enough that
-the algorithm for computing the light curve is numerically stable. We
-compute the light curve at $K$ points equally spaced over a baseline $\Delta t$
+the algorithm for computing the light curve is numerically stable.%
+\footnote{%
+    A more principled approach would perhaps be to geneate light curves using
+    a completely different model, such as by discretizing the surface at very
+    high resolution and computing the flux via a weighted sum of the visible
+    pixels. However, this would take orders of magnitude longer than the adopted
+    approach and would still be subject to artefacts due to the discretization
+    scheme. We have gone to great lengths in \citet{Luger2019} to show that
+    our flux computation from spherical harmonics is both accurate and precise,
+    so we are confident that our synthetic light curves correctly represent the
+    assumed spot distributions.
+}
+We compute the light curve at $K$ points equally spaced over a baseline $\Delta t$
 using the \starry algorithm
@@
 
-\begin{figure}[t!]
+\begin{figure}[p!]
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/calibration_default_data.pdf}
+        \includegraphics[width=0.9\linewidth]{figures/calibration_default_data.pdf}
         \oscaption{calibration}{%
@@
             surface with parameters given by the default values listed in
-            Table~\ref{tab:synthetic}. The surface maps for 50 synthetic stars are
+            Table~\ref{tab:synthetic}. The surface maps for 49 synthetic stars are
             shown in a Mollweide equal-area projection above their corresponding
@@
 Figure~\ref{fig:calibration_default_data} shows a synthetic dataset generated
-from the default values listed in Table~\ref{tab:synthetic}. While the 50
+from the default values listed in Table~\ref{tab:synthetic}. While the
 light curves correspond to surfaces with the same statistical spot properties,
 they all look qualitatively different: the mapping from star spot properties
-to flux is (very) nontrivial. In the inference step below, we assume we observe
-only these 50 light curves, with no knowledge of the inclination of any individual
+to flux is nontrivial. In the inference step below, we assume we observe
+only these 50 light curves (the figure only shows 49 of them),
+with no knowledge of the inclination of any individual
 star, and attempt to infer the spot properties.
@@
 Our sampling parameters are the number of spots $n$, their contrast $c$,
-their radius $r$, and the shape parameters $a$ and $b$ describing their
+their radius $r$, and the Beta distribution parameters $a$ and $b$ describing their
 distribution in latitude. As we discuss in Appendix~\ref{sec:lat},
@@
             to the Beta distribution PDF (Equation~\ref{eq:phi-pdf}) for the spot latitude
-            with shape parameters drawn from the posterior in $\mu_\phi$ and $\sigma_\phi$
+            with parameters drawn from the posterior in $\mu_\phi$ and $\sigma_\phi$
             (Figure~\ref{fig:calibration_default_corner}); the (hyper)distribution of blue curves
@@
 of them. This is yet another manifestation of the baseline degeneracy:
-the total number of spots is not an observable in relative photometry.
+the total number of spots is not an observable in relative photometry
+(see \S\ref{sec:breaking_baseline_degeneracy}).
 
@@
 concentrated mid-latitude spots and high-latitude spots with large variance.
-Interestingly, the inferred radius appears to be biased high by a small amount
-in this run. This is likely due to the fact that our prescription for
+Interestingly, the inferred radius appears to be biased high by a small amount.
+This is likely due to the fact that our prescription for
 generating the spots (\S\ref{sec:calibration-setup}) is different from how
@@
 %
-While these results broadly agree with the expectation that stellar
+While this broadly agrees with the expectation that stellar
 magnetic activity decreases over time, our work strongly suggests
@@
 range of spectral types and stellar ages considered in that study.
-
-Nevertheless, the core idea in \citet{Morris2020}
+Nevertheless, the power law scaling of the smoothed amplitude with
+age suggests it is still a good metric for stellar ages.
+
+Moreover, the core idea in \citet{Morris2020}
 \citep[and in related studies such as][]{Jackson2013}
@@
 to work with the log of these quantities because of their large dynamic range.
-We therefore introduce the modified shape parameters
+We therefore introduce the modified parameters
 %
@@
 constrain the hyperparameters of the GP---we recommend sampling in the
-shape parameters $a$ and $b$, since their domains are trivial,
+parameters $a$ and $b$, since their domains are trivial,
 with uncorrelated boundaries. Posterior constraints on these quantities may

### 2021-01-19 starry_process/calibrate/run.py

@@ -18,6 +18,8 @@ def run(
     plot_corner=False,
     plot_corner_transformed=True,
     plot_inclination_pdf=True,
+    ncols=10,
+    clip=False,
     **kwargs,
 ):
     if not os.path.exists(path):
@@ -47,7 +49,7 @@ def run(
     # Plot the data
     if plot_all or plot_data:
         if clobber or not os.path.exists(os.path.join(path, "data.pdf")):
-            fig = plot.plot_data(data, **kwargs)
+            fig = plot.plot_data(data, ncols=ncols, clip=clip, **kwargs)
             fig.savefig(
                 os.path.join(path, "data.pdf"), bbox_inches="tight", dpi=300
             )
@@ -125,5 +127,5 @@ def run(
         ):
             fig = plot.plot_inclination_pdf(data, inc_results, **kwargs)
             fig.savefig(
-                os.path.join(path, "inclination.pdf"), bbox_inches="tight",
+                os.path.join(path, "inclination.pdf"), bbox_inches="tight"
             )

### 2021-01-19 tex/figures/nonstationarity.py

@@ -1,0 +1,38 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import os
+
+np.random.seed(0)
+
+# Draw 1000 samples from a very long
+# period sinusoid
+t = np.linspace(0, 1, 300)
+z = 1 + 0.1 * np.sin(
+    2 * np.pi / 10 * t.reshape(1, -1)
+    + 2 * np.pi * np.random.random(size=(1000, 1))
+)
+
+# Normalize each sample to the mean
+znorm = z / np.mean(z, axis=1).reshape(-1, 1)
+
+# Figure setup
+fig, ax = plt.subplots(2, figsize=(12, 8), sharex=True)
+ax[1].set_xlabel("time", fontsize=18)
+ax[0].set_ylabel("true flux", fontsize=18)
+ax[1].set_ylabel("normalized flux", fontsize=18)
+ax[0].margins(0, 0.1)
+ax[1].margins(0, 0.1)
+ax[0].set_rasterization_zorder(2)
+ax[1].set_rasterization_zorder(2)
+
+# Plot
+for k in range(len(z)):
+    ax[0].plot(t, z[k], color="C0", lw=1, alpha=0.1)
+    ax[1].plot(t, znorm[k], color="C0", lw=1, alpha=0.1)
+
+# We're done
+fig.savefig(
+    os.path.abspath(__file__).replace(".py", ".pdf"),
+    bbox_inches="tight",
+    dpi=300,
+)

### 2021-01-20 tex/ms.tex
COMPLETE DRAFT

@@
     \textbf{
-        Interpretable Gaussian Processes for Stellar Variability
+        Interpretable Gaussian Processes for Starspots
     }
@@
 \email{rluger@flatironinstitute.org}
-\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY}
+\affil{Center~for~Computational~Astrophysics,~Flatiron~Institute,~New~York,~NY}
 \affil{Virtual~Planetary~Laboratory, University~of~Washington, Seattle, WA}
 %
-
-\keywords{methods: analytic}
+\author[0000-0002-9328-5652]{Daniel Foreman-Mackey}
+\affil{Center~for~Computational~Astrophysics,~Flatiron~Institute,~New~York,~NY}
+%
+\author[0000-0002-3385-8391]{Christina Hedges}
+\affil{Bay~Area~Environmental~Research~Institute,~P.O.~Box~25,~Moffett~Field,~CA~94035,~USA}
+\affil{NASA~Ames~Research~Center,~Moffett~Field,~CA}
+%
+
+\keywords{analytical mathematics --- time series analysis --- Gaussian processes regression --- starspots}
 
@@
     of the stellar signal in cases where the
-    variability due to star spots rotating in and out of view is treated
+    variability due to starspots rotating in and out of view is treated
     as a nuisance, such as in exoplanet transit modeling.
@@
     Instead, it is common practice to explicitly model the effect
-    of individual star spots on the light curve and attempt to infer their
+    of individual starspots on the light curve and attempt to infer their
     properties via optimization or posterior inference. Unfortunately,
@@
     the typical radii and the mean and variance of the latitude
-    distribution of star spots.
+    distribution of starspots.
     %
@@
 \label{sec:intro}
+
+\subsection{Stellar surface modeling}
+\label{sec:intro:spots}
+
+The advent of space-based precision photometry with missions such as
+\emph{Kepler} \citep{Borucki2010} and \emph{TESS} \citep{Ricker2015}
+has led to a renewed interest in the modeling of stellar light curves,
+and, in particular, in understanding what these light curves can tell
+us about the surfaces of stars across the HR diagram. One of the dominant
+sources of stellar light curve variability is the modulation caused
+by starspots rotating in and out of view.
+Dark spots arise due to the suppression of convection in regions of
+intense magnetic field, resulting in a locally cooler (and hence darker)
+photosphere. Bright spots can similarly
+arise in the photosphere as faculae or in the chromosphere as plages, and
+are also magnetically driven \citep[e.g.,][]{Berdyugina2005}.
+%
+Constraining the sizes,
+contrasts, locations, and number of spots on stars can therefore reveal
+information about stellar magnetic activity, stellar interior structure,
+and how these quantities vary across spectral type and over time.
+A detailed understanding of the stellar surface is also crucial to
+mitigating systematics in the radial velocity search for exoplanets
+\citep[e.g.,][]{Lanza2011} and in the spectroscopic characterization of their
+atmospheres \citep[e.g.,][]{Rackham2018}.
+
+To date, most studies aimed at inferring stellar surface
+properties from light curves follow one of two broad approaches. The
+first is to model the stellar surface as a collection of one or more
+discrete, circular, uniform contrast dark or bright spots on a uniform
+intensity photosphere.
+The advantage of this approach is that
+the light curve can be computed efficiently and in some cases even
+analytically \citep[e.g.,][]{Davenport2015,Morris2017,Morris2020b}.
+%
+The second approach is to discretize the surface at some resolution
+and compute the emergent flux as a weighted sum of the visible pixel intensities.
+This approach is more flexible, since it is not limited to surfaces
+composed of distinct circular spots \citep[e.g.,][]{Harmon2000,Roettenbacher2017}.
+%
+Both approaches rely on an explicit \emph{forward model}, a prescription for
+how to generate data given a set of parameters. In most cases, however, we are interested
+in the inverse problem: constraining the parameters given data.
+%
+Unfortunately, the inverse problem is not only difficult---as it requires
+a large number of forward model evaluations to find the parameter values
+that are most consistent with the
+data---but also formally \emph{ill-posed}.
+%
+While the mapping from a stellar surface to a light curve (the forward
+problem) is unique, the mapping from a light curve to a surface
+(the inverse problem) is not:
+given any light curve, there exist an infinite number of surfaces that
+could have generated it. These degeneracies are illustrated in
+Figure~\ref{fig:degeneracies}, where six synthetic stellar surfaces
+are shown (rows) at twelve different phases (columns), all at an
+inclination $I=60^\circ$. Each surface
+consists of a different number of dark spots on a brighter, heterogeneous
+background.
 
@@
             corresponds to a stellar surface with a different number of
-            dark spots seen at various phases at an inclination $I=60^\circ$.
+            dark spots seen at various phases at an inclination $I=60^\circ$;
+            all images are shown on the same color scale.
             The bottom panel shows the light curves of each of these stars.
@@
 
-% CHECK OUT
+While the stellar surfaces are all distinct, containing between one (top)
+and six (bottom) large dark spots,
+\textbf{their rotational light curves are identical}
+(lower panel). This is true even at infinite signal to
+noise: the mapping from a stellar surface
+to its rotational light curve is so degenerate that there exist an infinite number of solutions
+to the inverse problem. This fact has been pointed out recently in different contexts
+\citep[e.g.,][]{Cowan2013,Luger2019,Basri2020}, but it dates back at least to
+\citet{Russell1906}, who demonstrated it by expanding the surface
+intensity of a celestial body in terms of spherical harmonics
+(see \S\ref{sec:main} and Figure~\ref{fig:ylms}). \citet{Russell1906} showed
+that many of the modes comprising the intensity profile of a spherical
+object are in the \emph{null space}, the set of surface features that have identically
+zero effect on the light curve. In fact,
+the \emph{vast majority} of the modes are in the null space for rotational
+light curves of stars (\S\ref{sec:nullspace}). This is what allows us to construct pathological
+scenarios like that shown in Figure~\ref{fig:degeneracies}, where the light curve
+could be explained by any number of spots atop a heterogeneous bright background.
+
+Stellar mapping studies tackle these degeneracies in different ways, but
+it usually comes down to a choice of prior: when the data is not
+sufficiently informative, prior assumptions are needed to discriminate
+between competing solutions. In discrete spot models like the ones
+discussed above, the degeneracy-breaking prior is (typically) the assumption that the
+spots must be circular, have uniform contrast, and sit atop an otherwise
+uniform photosphere. In gridded stellar surface models, it is common to
+assume a regularization prior such as the maximum entropy penalty
+\citep[e.g.,][]{Vogt1987}, which typically favors solutions with the fewest
+number of dark pixels (usually referred to as the ``simplest'' solution).
+
+While these assumptions may be approximately valid in some cases, it is
+important to bear in mind that because of the light curve degeneracies discussed above,
+\textbf{most of the information about the stellar surface usually comes from the prior},
+so it is very important to get the
+prior right. In general, sunspots are not circular and do not have uniform
+contrast throughout; nor do spots always arrange themselves in the highest
+entropy configuration. The amount of bias introduced by these assumptions
+will in general vary, but in principle it could be quite significant.
+
+% Add to this photometric noise, the generally unknown stellar inclination,
+% the poorly constrained limb darkening parameters, and you get ... a mess.
+
+Fortunately, although individual light curves are not very constraining,
+we have \emph{a lot of them} thanks to missions like \emph{Kepler} and \emph{TESS}.
+In this paper, we will thoroughly explore the power of ensemble analyses
+in breaking the degeneracies inherent to the light curve mapping problem.
+By jointly analyzing the light curves of many stars, we will show that it
+is often possible to uniquely constrain statistical properties about their
+surfaces. This idea was recently explored to some extent
+in \citet{Morris2020}, who used ensemble analyses to derive constraints on
+spot coverage areas as a function of stellar age (see \S\ref{sec:other-work}).
+We will go a step further and
+show that it is possible to constrain the distribution of spot sizes, active
+latitudes, and other statistical properties of stellar surfaces using a novel
+interpretable Gaussian process approach, which we motivate in the following
+section.
+
+% CHECK OUT (for a future paper)
 % Louden 2020 https://ui.adsabs.harvard.edu/abs/2020arXiv201200776L/abstract
 % Mazeh 2015
+
+\subsection{Gaussian processes}
+\label{sec:intro:gps}
 
@@
 Because of this relationship to multivariate Gaussians,
-GPs are extremely easy to sample from.%
+GPs are easy to sample from.%
 \footnote{Given a 1-d array \texttt{mean} and a 2-d array \texttt{cov} in \Python,
@@
 Multivariate Gaussian distributions have a closed-form (marginal) likelihood
-function, so it is easy (and extremely efficient) to compute the probability
+function, so it is easy to compute the probability
 of one's data conditioned on a given value of $\pmb{\mu}$ and $\pmb{\Sigma}$
@@
 to modeling stellar spot-induced variability. While it is straightforward to derive
-posterior constraints on the hyperparameters of a GP, it is usually
+posterior constraints on the hyperparameters of a GP, it is often
 not obvious what those constraints actually tell us about the star
-or its surface. While the GP period $P$
-may sometimes be interpreted as the rotational period of the star,%
-\footnote{An exception to this is in the presence of strong differential
-    rotation, in which case many periods may be present in the data, or
-    when spots evolve coherently, which can also introduce weak periodicities
-    in the light curve.}
-the relationship between quantities like the amplitude $\sigma^2$
-and the timescale $\tau$ and physical properties of the spots is much fuzzier.
-It may be tempting to interpret $\sigma^2$ as some measure of the spot
+or its surface.
+Specifically, it is difficult to constrain the relationship between
+quantities like the amplitude $\sigma^2$
+and the timescale $\tau$ and physical properties of the surface.
+For instance, it may be tempting to interpret $\sigma^2$ as some measure of the spot
 contrast or the total number of spots, or $\tau$ as the timescale on which the
@@
 of the physics.
-
-For this reason, GPs are usually employed to model stellar variability
+%
+There are two important exceptions to this: asteroseismic studies, in which the
+the GP hyperparameters can offer direct insight into
+the behavior of complex pulsation modes and thus physical properties of the
+stellar interior \citep[e.g.,][]{BrewerStello2009,ForemanMackey2017}; and
+stellar rotation period studies, in which the period hyperparameter $P$ can usually
+be associated with the rotation period of the star \citep[e.g.,][]{Angus2018}.%
+\footnote{An exception to this is in the presence of strong differential
+    rotation, in which case many periods may be present in the data, or
+    when spots evolve coherently, which can also introduce weak periodicities
+    in the light curve.}
+%
+For spot-induced variability, on the other hand,
+GPs are usually used
 when the variability itself is a nuisance parameter. For example, if the
-goal is to constrain properties of a transiting exoplanet, a GP might be
-used to remove (or, better yet, to marginalize over) the stellar signal.
+goal is to constrain the properties of a transiting exoplanet or to
+search for a planetary signal in a radial velocity dataset, a GP might be
+used to remove (or, better yet, to marginalize over) the stellar variability
+\citep[e.g.,][]{Haywood2014,Rajpaul2015,Luger2017b}.
 In this case, the physics behind the variability is irrelevant, so an
-effective model of this sort is generally sufficient. When the goal is to understand
-the variability itself, however, GPs have been far less useful to date. Instead,
-the usual approach has been to explicitly model the stellar signal: in the
-case of star spot-induced variability, one must construct a physical model
-for the stellar surface and for how it projects into the light curve as the
-star rotates. This might entail directly modeling the star spots,
-with free parameters controling the spot size, contrast, position,
-and temporal evolution of each of the $n$ spots, where $n$ is also
-a free parameter \xxx{(CITE ME)}. In addition to this being a numerically challenging
-problem for all but the smallest values of $n$, the problem is
-intrinsically ill-posed, meaning there are strong degeneracies at play
-that formally preclude a unique solution to the stellar map.
-After all, a stellar light curve is a one-dimensional representation of a
-two-dimensional (or, if the star evolves in time, three-dimensional)
-surface, and thus it simply cannot encode all the information about the stellar surface.
-In fact, the \emph{vast majority} of the modes on the stellar surface
-are in the \emph{null space}, the set of surface features that have identically
-zero effect on the light curve \citep[e.g.,][]{Cowan2013,Luger2019}.
-The underdetermined nature of the stellar mapping problem is therefore fundamental: it
-is simply not possible to uniquely determine the properties of all the spots on
-a star from its rotational light curve.
+effective model of this sort is generally sufficient.
+
+When the goal is to understand
+the the stellar surface itself, however, GPs have been less useful to date.
+Instead, the common approach is that discussed in the previous section:
+to forward model the stellar surface and its effect on the light curve, then
+seek to constrain the surface properties by solving the inverse problem.
+As we discussed above, however, a unique solution to the surface map of
+the star from a rotational light curve is not possible without the use of
+aggressive (and often \emph{ad hoc}) priors. The degeneracies at play
+make it effectively impossible for one to know the exact configuration
+of starspots and other features on the surface of a star from its rotational
+light curve alone.
 
@@
 %strong degeneracies: as we will see, the statistical properties of
-%star spots have a much smaller null space than the properties of
+%starspots have a much smaller null space than the properties of
 %individual spots.
@@
 
+\subsection{Overview}
+\label{sec:intro:overview}
+
 \textbf{In this paper, we derive an exact, closed-form expression for
     the Gaussian approximation to the marginal likelihood of a light curve
-    conditioned on the statistical properties of star spots.} Our Gaussian
+    conditioned on the statistical properties of starspots.} Our Gaussian
 process analytically marginalizes over the (degenerate, often unknowable)
-distributions of properties of individual star spots, revealing the
+distributions of properties of individual starspots, revealing the
 constraints imposed on the bulk spot properties without the need to
@@
 in \S\ref{sec:calibration}. We discuss our GP from an information
-theoretical perspective in \S\ref{sec:discussion}, focusing on
+theory perspective in \S\ref{sec:discussion}, focusing on
 the various degeneracies our model can and cannot break when applied
@@
 
-\section{A Gaussian Process for Star Spots}
+\section{A Gaussian Process for starspots}
 \label{sec:main}
@@
     struggle (at finite degree $l$) to capture high resolution features such as
-    small star spots. We discuss this point at length in \S\ref{sec:tinyspots},
+    small starspots. We discuss this point at length in \S\ref{sec:tinyspots},
     where we show that our model is useful even when applied to stars with spots
@@
 a set of limb darkening coefficients $\mathbf{u}$, and
-on certain properties of the star spots $\pmb{\theta}_\bullet$
+on certain properties of the starspots $\pmb{\theta}_\bullet$
 (including their number, sizes, positions, and contrasts),
@@
 %
-where $n$ is the number of star spots, $c$ is their contrast,
+where $n$ is the number of starspots, $c$ is their contrast,
 $\mu_\phi$ and $\sigma_\phi$ are the mode and standard deviation
@@
 $\left( \mathbf{f}_0 \,\, \mathbf{f}_1 \,\, \cdots \,\,  \mathbf{f}_{M-1} \right)^\top$,
-we wish to infer the statistical properties of the star spots, encoded in
+we wish to infer the statistical properties of the starspots, encoded in
 the entries of the vector $\pmb{\theta}_\bullet$.
@@
 posterior inference exercises demonstrating that our GP model is correctly
-calibrated, allowing one to efficiently infer statistical properties of star spots
+calibrated, allowing one to efficiently infer statistical properties of starspots
 from light curves with minimal bias.
@@
 %
-It is therefore possible to evaluate $\tilde{\pmb{\Sigma}}$ in an extremely
+It is therefore possible to evaluate $\tilde{\pmb{\Sigma}}$ in an
 efficient manner. In \S\ref{sec:implementation} we discuss our
@@
 we will use below.
+
+% HACK HACK HACK HACK HACK
+\pagebreak
+% HACK HACK HACK HACK HACK
 
@@
 
+% HACK HACK HACK HACK HACK
+\pagebreak
+% HACK HACK HACK HACK HACK
+
 \subsection{Why ensemble analyses work}
@@
 
+% HACK HACK HACK HACK HACK
+\pagebreak
+% HACK HACK HACK HACK HACK
+
 \subsection{Comparison to other work}
@@
 to synthetic RV datasets, arguing that a new four-parameter quasi-periodic
-cosine kernel (QPC) can better capture the variability due to star spots.
+cosine kernel (QPC) can better capture the variability due to starspots.
 However, their study was empirical and related spot configurations to their
@@
 stellar surfaces are static. Our GP hyperparameters $\pmb{\theta}_\bullet$
-describe the \emph{spatial} configuration of star spots, but they say nothing
+describe the \emph{spatial} configuration of starspots, but they say nothing
 about their evolution in time. We know from observations of the Sun and of
@@
 temporally-variable \emph{Kepler} light curves.
+
+% HACK HACK HACK HACK HACK
+\pagebreak
+% HACK HACK HACK HACK HACK
 
@@
 coefficients, which could be significant; it is also subject to bias due to
-the fact that theoretical models for limb darkening coefficients have been
-shown to be wrong, especially for low mass stars \citep{?}. A much better
+the fact that there may be systematic errors in theoretical models for the limb darkening
+coefficients, particularly for low mass stars \citep[e.g.,][]{Kervella2017}. A much better
 approach would be to analytically marginalize over these two quantities.
@@
 coefficients in terms of the uncorrelated $q_1$ and $q_2$ parameters from
-\citet{?}, which would allow us to incorporate hard constraints on positivity
+\citet{Kipping2013}, which would allow us to incorporate hard constraints on positivity
 and strict limb darkening (as opposed to brightening) directly into the prior.
@@
           corresponding to spots about an order of magnitude larger than typical
-          Sun spots. However, we have showed that when we apply our model to
+          sunspots. However, we have showed that when we apply our model to
           light curves of stars with small ($r \sim 3^\circ$) spots, we can
@@
 
+\vspace{2em}
+
+We would like to thank David W. Hogg, Michael Gully-Santiago, and the
+Astronomical Data Group at the Center for Computational Astrophysics for
+many thought-provoking discussions that helped shape this paper.
+
 \bibliography{bib}
@@
 %
-As we are specifically interested in modeling the effect of star spots
+As we are specifically interested in modeling the effect of starspots
 on stellar light curves, we let

### 2021-01-20 tex/ms.tex
tweaks

@@
 all spherical harmonic coefficients, these may be expressed as high-dimensional
-tensor products, which can be evaluated extremely efficiently on modern
+tensor products, which can be evaluated efficiently on modern
 hardware. Many of the expressions can also be either pre-computed or
@@
             marginalizing over the inclination (orange). At $l_\mathrm{max} = 15$,
-            computation of the covariance matrix of the GP takes about 10ms
+            computation of the covariance matrix of the GP takes about 20ms
             on a 2018 MacBook Pro. The dashed line shows the asymptotic scaling
@@
 solution as a function of the spot parameters, it can be computed
-extremely efficiently: a typical likelihood evaluation on a dataset
-consisting of $K \sim 1{,}000$ points takes about 10ms on a modern
+efficiently: a typical likelihood evaluation on a dataset
+consisting of $K \sim 1{,}000$ points takes about 20ms on a modern
 laptop. Our algorithm is implemented in the open-source, user-friendly
@@
 cases unbiased estimator for star spot properties. Below we list the main
-results presented in this paper:
+results concerning our GP:
 
@@
           same parent distribution, whose parameters we can constrain.
-    \item \textbf{In general, an ensemble of at least $M \sim 50$ light curves
-              is needed to place meaningful constraints on the spot properties}.
-          This estimate is based on ensemble analyses of light curves with
-          $K = 1{,}000$ cadence each and per-cadence precision of one part
-          per thousand. Lower quality, shorter baseline observations, or
-          datasets contaminated by outliers will in general require larger values
-          of $M$ for the same constraining power. The presence of strong limb
-          darkening also degrades the information content of light curves, in
-          which case an ensemble of hundreds or even a thousand light curves
-          is recommended.
     \item \textbf{Our GP is an unbiased estimator for the spot radius and
@@
           average spot latitudes to within $5^\circ$.
-    \item \textbf{Relative single-band photometry cannot uniquely constrain
-              the spot contrast, the total number of spots, or the total spot
-              coverage}. These two quantities
-          are perfectly degnerate with each other in relative single-band photometry.
-          Knowledge of the
-          unspotted photospheric brightness (or temperature) can break this
-          degeneracy, so analyses of multi-band light curves can in principle
-          infer the total spottiness of a stellar surface.
     \item \textbf{Our GP can accurately infer stellar inclinations.} We presented
@@
           complex temporal variability, such as that induced by differential
-          rotation, will be the subject of the next paper in this series.
+          rotation, will be the subject of a future paper in this series.
     \item \textbf{Our GP can be used in exoplanet transit modeling and
@@
           be the subject of a future paper in this series.
+\end{enumerate}
+
+In this paper we also demonstrated and discussed four points that apply more
+broadly to stellar light curve variability modeling in general:
+
+\begin{enumerate}
+    \item \textbf{Typically, an ensemble of at least $M \sim 50$ light curves
+              is needed to place meaningful constraints on starspot properties}.
+          This estimate is based on ensemble analyses of light curves with
+          $K = 1{,}000$ cadence each and per-cadence precision of one part
+          per thousand. Lower quality, shorter baseline observations, or
+          datasets contaminated by outliers will in general require larger values
+          of $M$ for the same constraining power. The presence of strong limb
+          darkening also degrades the information content of light curves, in
+          which case an ensemble of hundreds or even a thousand light curves
+          is recommended.
+    \item \textbf{Limb darkening breaks degeneracies}. The presence of limb darkening
+          changes the structure of the null space. Even small variations in the
+          limb darkening coefficients among stars in an ensemble can effectively
+          increase the information content of their light curves when analyzed
+          jointly. However, this is both a blessing and a curse: it is very
+          important to accurately model limb darkening to infer starspot
+          properties without bias.
+    \item \textbf{Relative single-band photometry cannot uniquely constrain
+              the spot contrast, the total number of spots, or the total spot
+              coverage}. We showed that these two quantities
+          are perfectly degenerate with each other in relative single-band photometry.
+          Knowledge of the
+          unspotted photospheric brightness (or temperature) can break this
+          degeneracy, so analyses of multi-band light curves can in principle
+          infer the total spottiness of a stellar surface.
+    \item \textbf{Care should be taken when analyzing normalized light curves}.
+        We showed that the widely-adopted practice of normalizing light curves
+        to their mean, median, or maximum changes the covariance structure of
+        the data in potentially complex ways. In addition to changing the total
+        variance of the data, this normalization step introduces non-stationarity
+        in the covariance, which can lead to biases particularly when
+        modeling light curves of stars with very long rotation periods.
 \end{enumerate}

### 2021-02-17 joss/paper.tex
building JOSS paper

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{colorlinks,breaklinks=true,
            urlcolor=[rgb]{0.0, 0.5, 1.0},
            linkcolor=[rgb]{0.0, 0.5, 1.0}}
\usepackage{caption}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{xstring}

\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

\usepackage{fixltx2e} % provides \textsubscript
\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{bib.bib}

% --- Splitting \texttt --------------------------------------------------

\let\textttOrig=\texttt
\def\texttt#1{\expandafter\textttOrig{\seqsplit{#1}}}
\renewcommand{\seqinsert}{\ifmmode
  \allowbreak
  \else\penalty6000\hspace{0pt plus 0.02em}\fi}

% --- Pandoc does not distinguish between links like [foo](bar) and
% --- [foo](foo) -- a simplistic Markdown model.  However, this is
% --- wrong:  in links like [foo](foo) the text is the url, and must
% --- be split correspondingly.
% --- Here we detect links \href{foo}{foo}, and also links starting
% --- with https://doi.org, and use path-like splitting (but not
% --- escaping!) with these links.
% --- Another vile thing pandoc does is the different escaping of
% --- foo and bar.  This may confound our detection.
% --- This problem we do not try to solve at present, with the exception
% --- of doi-like urls, which we detect correctly.

\makeatletter
\let\href@Orig=\href
\def\href@Urllike#1#2{\href@Orig{#1}{\begingroup
    \def\Url@String{#2}\Url@FormatString
    \endgroup}}
\def\href@Notdoi#1#2{\def\tempa{#1}\def\tempb{#2}%
  \ifx\tempa\tempb\relax\href@Urllike{#1}{#2}\else
  \href@Orig{#1}{#2}\fi}
\def\href#1#2{%
  \IfBeginWith{#1}{https://doi.org}%
  {\href@Urllike{#1}{#2}}{\href@Notdoi{#1}{#2}}}
\makeatother

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Default font ------------------------------------------------------------
\renewcommand\familydefault{\sfdefault}

% --- Style -------------------------------------------------------------------
\renewcommand{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{figures/joss-logo.png}}
\fancyhead[C]{}
\fancyhead[R]{}
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Luger, (2021). starry\_process:
Interpretable Gaussian processes for stellar light
curves. \textit{Journal of Open Source Software}, TBD(TBD), TBD. \url{https://doi.org/DOI unavailable}}}}

\fancyfoot[R]{\sffamily \thepage}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Macros ---------

\definecolor{linky}{rgb}{0.0, 0.5, 1.0}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

\newcommand{\ExternalLink}{%
   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
       \begin{scope}[x=1ex, y=1ex]
           \clip (-0.1,-0.1)
               --++ (-0, 1.2)
               --++ (0.6, 0)
               --++ (0, -0.6)
               --++ (0.6, 0)
               --++ (0, -1);
           \path[draw,
               line width = 0.5,
               rounded corners=0.5]
               (0,0) rectangle (1,1);
       \end{scope}
       \path[draw, line width = 0.5] (0.5, 0.5)
           -- (1, 1);
       \path[draw, line width = 0.5] (0.6, 1)
           -- (1, 1) -- (1, 0.6);
       }
   }

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\renewcommand{\affil}[2][]%
   {\newaffiltrue\let\AB@blk@and\AB@pand
      \if\relax#1\relax\def\AB@note{\AB@thenote}\else\def\AB@note{#1}%
        \setcounter{Maxaffil}{0}\fi
        \begingroup
        \let\href=\href@Orig
        \let\texttt=\textttOrig
        \let\protect\@unexpandable@protect
        \def\thanks{\protect\thanks}\def\footnote{\protect\footnote}%
        \@temptokena=\expandafter{\AB@authors}%
        {\def\\{\protect\\\protect\Affilfont}\xdef\AB@temp{#2}}%
         \xdef\AB@authors{\the\@temptokena\AB@las\AB@au@str
         \protect\\[\affilsep]\protect\Affilfont\AB@temp}%
         \gdef\AB@las{}\gdef\AB@au@str{}%
        {\def\\{, \ignorespaces}\xdef\AB@temp{#2}}%
        \@temptokena=\expandafter{\AB@affillist}%
        \xdef\AB@affillist{\the\@temptokena \AB@affilsep
          \AB@affilnote{\AB@note}\protect\Affilfont\AB@temp}%
      \endgroup
       \let\AB@affilsep\AB@affilsepx
}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{fontspec}

  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}

\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={starry\_process: Interpretable Gaussian processes for stellar light curves},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

% --- We redefined \texttt, but in sections and captions we want the
% --- old definition
\let\addcontentslineOrig=\addcontentsline
\def\addcontentsline#1#2#3{\bgroup
  \let\texttt=\textttOrig\addcontentslineOrig{#1}{#2}{#3}\egroup}
\let\markbothOrig\markboth
\def\markboth#1#2{\bgroup
  \let\texttt=\textttOrig\markbothOrig{#1}{#2}\egroup}
\let\markrightOrig\markright
\def\markright#1{\bgroup
  \let\texttt=\textttOrig\markrightOrig{#1}\egroup}

\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\title{starry\_process: Interpretable Gaussian processes for stellar
light curves}

        \author[1, 2]{Rodrigo Luger}
          \author[1]{Daniel Foreman-Mackey}
          \author[3, 4]{Christina Hedges}

      \affil[1]{Center for Computational Astrophysics, Flatiron
Institute, New York, NY}
      \affil[2]{Virtual Planetary Laboratory, University of Washington,
Seattle, WA}
      \affil[3]{Bay Area Environmental Research Institute, P.O. Box 25,
Moffett Field, CA 94035, USA}
      \affil[4]{NASA Ames Research Center, Moffett Field, CA}
  \date{\vspace{-7ex}}

\begin{document}
\maketitle

\marginpar{

  \begin{flushleft}
  %\hrule
  \sffamily\small

  {\bfseries DOI:} \href{https://doi.org/DOI unavailable}{\color{linky}{DOI unavailable}}

  \vspace{2mm}

  {\bfseries Software}
  \begin{itemize}
    \setlength\itemsep{0em}
    \item \href{N/A}{\color{linky}{Review}} \ExternalLink
    \item \href{NO_REPOSITORY}{\color{linky}{Repository}} \ExternalLink
    \item \href{DOI unavailable}{\color{linky}{Archive}} \ExternalLink
  \end{itemize}

  \vspace{2mm}

  \par\noindent\hrulefill\par

  \vspace{2mm}

  {\bfseries Editor:} \href{https://example.com}{Pending
Editor} \ExternalLink \\
  \vspace{1mm}
    {\bfseries Reviewers:}
  \begin{itemize}
  \setlength\itemsep{0em}
    \item \href{https://github.com/Pending Reviewers}{@Pending
Reviewers}
    \end{itemize}
    \vspace{2mm}

  {\bfseries Submitted:} N/A\\
  {\bfseries Published:} N/A

  \vspace{2mm}
  {\bfseries License}\\
  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{http://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC BY 4.0}}).

  \end{flushleft}
}

\hypertarget{section}{%
\section{}\label{section}}

\hypertarget{summary}{%
\section{Summary}\label{summary}}

The \texttt{starry\_process} code implements an interpretable Gaussian
process (GP) for modeling variability in stellar light curves. As dark
starspots rotate in and out of view, the total flux received from a
distant star will change over time. Unresolved flux time series
therefore encode information about the spatial structure of features on
the stellar surface. The \texttt{starry\_process} software package
allows one to easily model the flux variability due to starspots,
whether one is interested in understanding the properties of these spots
or marginalizing over the stellar variability when it is treated as a
nuisance signal. The main difference between the GP implemented here and
typical GPs used to model stellar variability is the explicit dependence
of our GP on physical properties of the star, such as its period,
inclination, and limb darkening coefficients, and on properties of the
spots, such as their radius and latitude distributions (see Figure
\ref{fig:samples}). This code is the \texttt{Python} implementation of
the interpretable GP algorithm developed in Luger, Foreman-Mackey, \&
Hedges (2021).

\begin{figure}
\centering
\includegraphics{figures/samples.pdf}
\caption{Five random samples from our GP (columns) conditioned on two
different hyperparameter vectors \(\pmb{\theta}_\bullet\) (rows). The
samples are shown on the surface of the star in a Mollweide projection
alongside the corresponding light curves viewed at several different
inclinations. (a) Samples from a GP describing a star with small
mid-latitude spots. (b) Samples from a GP describing a star with larger
high-latitude spots. \label{fig:samples}}
\end{figure}

\hypertarget{section-1}{%
\section{}\label{section-1}}

\hypertarget{statement-of-need}{%
\section{Statement of need}\label{statement-of-need}}

Mapping the surfaces of stars using time series measurements is a
fundamental problem in modern time-domain stellar astrophysics. This
inverse problem is ill-posed and computationally intractable, but in the
associated AAS Journals publication submitted in parallel to this paper
(Luger, Foreman-Mackey, \& Hedges, 2021), we derive an interpretable
effective Gaussian Process (GP) model for this problem that enables
robust probabilistic characterization of stellar surfaces using
photometric time series observations. Our model builds on previous work
by Perger et al. (2020) on semi-interpretable Gaussian processes for
stellar timeseries data and by Morris (2020) on approximate inference
for large ensembles of stellar light curves. Implementation of our model
requires the efficient evaluation of a set of special functions and
recursion relations that are not readily available in existing
probabilistic programming frameworks. The \texttt{starry\_process}
package provides the necessary elements to perform this analysis with
existing and forthcoming astronomical datasets.

\hypertarget{implementation}{%
\section{Implementation}\label{implementation}}

We implement our interpretable GP in the user-friendly \texttt{Python}
package \texttt{starry\_process}, which can be installed via
\texttt{pip} or from source on
\href{https://github.com/rodluger/starry_process}{GitHub}. The code is
thoroughly
\href{https://github.com/rodluger/starry_process/tree/master/tests}{unit-tested}
and \href{https://starry_process.readthedocs.io}{well documented}, with
\href{https://starry-process.readthedocs.io/en/latest/examples}{examples}
on how to use the GP in custom inference problems. As discussed in the
associated AAS Journals publication (Luger, Foreman-Mackey, \& Hedges,
2021), users can choose, among other options, whether or not to
marginalize over the stellar inclination and whether or not to model a
normalized process. Users can also choose the spherical harmonic degree
of the expansion, although it is recommended to use
\(l_\mathrm{max} = 15\) (see below). Users may compute the mean vector
and covariance matrix in either the spherical harmonic basis or the flux
basis, or they may sample from it or use it to compute marginal
likelihoods. Arbitrary order limb darkening is implemented following
Agol et al. (2020).

The code was designed to maximize the speed and numerical stability of
the computation. Although the computation of the GP covariance involves
many layers of nested sums over spherical harmonic coefficients, these
may be expressed as high-dimensional tensor products, which can be
evaluated efficiently on modern hardware. Many of the expressions can
also be either pre-computed or computed recursively. To maximize the
speed of the algorithm, the code is implemented in hybrid
\texttt{C++}/\texttt{Python} using the just-in-time compilation
capability of the \texttt{Theano} package (Theano Development Team,
2016). Since all equations derived here have closed form expressions,
these can be autodifferentiated in a straightforward and numerically
stable manner, enabling the computation of backpropagated gradients
within \texttt{Theano}. As such, \texttt{starry\_process} is designed to
work out-of-the box with \texttt{Theano}-based inference tools such as
\texttt{PyMC3} for NUTS/HMC or ADVI sampling (Salvatier et al., 2016).

Figure \ref{fig:speed} shows the computational scaling of the
\texttt{Python} implementation of the algorithm for the case where we
condition the GP on a specific value of the inclination (blue) and the
case where we marginalize over inclination (orange). Both curves show
the time in seconds to compute the likelihood (averaged over many trials
to obtain a robust estimate) as a function of the number of points \(K\)
in a single light curve. For \(K \lesssim 100\), the computation time is
constant at \(10-30\) ms for both algorithms. This is the approximate
time (on a typical modern laptop) taken to compute the GP covariance
matrix given a set of hyperparameters \(\pmb{\theta}_\bullet\). For
larger values of \(K\), the cost approaches a scaling of \(K^{2.6}\),
which is dominated by the factorization of the covariance matrix and the
solve operation to compute the likelihood. The likelihood marginalized
over inclination is only slightly slower to compute, thanks to the
tricks discussed in Luger, Foreman-Mackey, \& Hedges (2021).

\begin{figure}
\centering
\includegraphics{figures/speed.pdf}
\caption{Evaluation time in seconds for a single log-likelihood
computation as a function of the number of points \(K\) in each light
curve when conditioning on a value of the inclination (blue) and when
marginalizing over the inclination (orange). At \(l_\mathrm{max} = 15\),
computation of the covariance matrix of the GP takes about 20ms on a
typical laptop. The dashed line shows the asymptotic scaling of the
algorithm, which is due to the Cholesky factorization and solve
operations.\label{fig:speed}}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/stability.pdf}
\caption{Log of the condition number of the covariance in the spherical
harmonic basis as a function of the spherical harmonic degree of the
expansion, \(l_\mathrm{max}\). Different lines correspond to different
values of \(\pmb{\theta}_\bullet\) drawn from a uniform prior (see text
for details). In the majority of the cases, the matrix becomes
ill-conditioned above \(l_\mathrm{max} = 15\).\label{fig:stability}}
\end{figure}

Many modern GP packages (e.g., Ambikasaran et al., 2015; Foreman-Mackey
et al., 2017) have significantly better asymptotic scalings, but these
are usually due to specific structure imposed on the kernel functions,
such as the assumption of stationarity. Our kernel structure is
determined by the physics (or perhaps more accurately, the geometry) of
stellar surfaces, and its nonstationarity is a consequence of the
normalization step in relative photometry (Luger, Foreman-Mackey,
Hedges, \& Hogg, 2021). Moreover, and unlike the typical kernels used
for GP regression, our kernel is a nontrivial function of the
hyperparameters \(\pmb{\theta}_\bullet\), so its computation is
necessarily more expensive. Nevertheless, the fact that our GP may be
used for likelihood evaluation in a small fraction of a second for
typical datasets (\(K \sim 1{,000}\)) makes it extremely useful for
inference.

Our algorithm is also numerically stable over nearly all of the prior
volume up to \(l_\mathrm{max} = 15\). Figure \ref{fig:stability} shows
the log of the condition number of the covariance matrix in the
spherical harmonic basis as a function of the spherical harmonic degree
of the expansion for 100 draws from a uniform prior over the domain of
the hyperparameters. The condition number is nearly constant up to
\(l_\mathrm{max} = 15\) in almost all cases; above this value, the
algorithm suddenly becomes unstable and the covariance is
ill-conditioned. The instability occurs within the computation of the
latitude and longitude moment integrals and is likely due to the large
number of operations involving linear combinations of hypergeometric and
gamma functions. While it may be possible to achieve stability at higher
values of \(l_\mathrm{max}\) via careful reparametrization of some of
those equations, we find that \(l_\mathrm{max} = 15\) is high enough for
most practical purposes.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Agol2020}{}%
Agol, E., Luger, R., \& Foreman-Mackey, D. (2020). {Analytic Planetary
Transit Light Curves and Derivatives for Stars with Polynomial Limb
Darkening}. \emph{159}(3), 123.
\url{https://doi.org/10.3847/1538-3881/ab4fee}

\leavevmode\hypertarget{ref-Ambikasaran2015}{}%
Ambikasaran, S., Foreman-Mackey, D., Greengard, L., Hogg, D. W., \&
O'Neil, M. (2015). {Fast Direct Methods for Gaussian Processes}.
\emph{IEEE Transactions on Pattern Analysis and Machine Intelligence},
\emph{38}, 252. \url{https://doi.org/10.1109/TPAMI.2015.2448083}

\leavevmode\hypertarget{ref-ForemanMackey2017}{}%
Foreman-Mackey, D., Agol, E., Ambikasaran, S., \& Angus, R. (2017).
{Fast and Scalable Gaussian Process Modeling with Applications to
Astronomical Time Series}. \emph{154}(6), 220.
\url{https://doi.org/10.3847/1538-3881/aa9332}

\leavevmode\hypertarget{ref-PaperII}{}%
Luger, R., Foreman-Mackey, D., \& Hedges, C. (2021). {Mapping stellar
surfaces II: An interpretable Gaussian process model for light curves}.
\emph{arXiv e-Prints}, arXiv:2102.01697.
\url{http://arxiv.org/abs/2102.01697}

\leavevmode\hypertarget{ref-PaperI}{}%
Luger, R., Foreman-Mackey, D., Hedges, C., \& Hogg, D. W. (2021).
{Mapping stellar surfaces I: Degeneracies in the rotational light curve
problem}. \emph{arXiv e-Prints}, arXiv:2102.00007.
\url{http://arxiv.org/abs/2102.00007}

\leavevmode\hypertarget{ref-Morris2020b}{}%
Morris, B. (2020). {fleck: Fast approximate light curves for starspot
rotational modulation}. \emph{The Journal of Open Source Software},
\emph{5}(47), 2103. \url{https://doi.org/10.21105/joss.02103}

\leavevmode\hypertarget{ref-Perger2020}{}%
Perger, M., Anglada-Escudé, G., Ribas, I., Rosich, A., Herrero, E., \&
Morales, J. C. (2020). {Auto-correlation functions of astrophysical
processes, and their relation to Gaussian processes; Application to
radial velocities of different starspot configurations}. \emph{arXiv
e-Prints}, arXiv:2012.01862. \url{http://arxiv.org/abs/2012.01862}

\leavevmode\hypertarget{ref-Salvatier2016}{}%
Salvatier, J., Wiecki, T. V., \& Fonnesbeck, C. (2016). {Probabilistic
programming in Python using PyMC3}. \emph{PeerJ Computer Science},
\emph{2}, e55.

\leavevmode\hypertarget{ref-Theano2016}{}%
Theano Development Team. (2016). {Theano: A {Python} framework for fast
computation of mathematical expressions}. \emph{arXiv e-Prints},
\emph{abs/1605.02688}. \url{http://arxiv.org/abs/1605.02688}

\end{CSLReferences}

\end{document}

### 2021-02-17 joss/figures/speed.py

@@ -1,0 +1,73 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from starry_process import StarryProcess
+import theano
+import theano.tensor as tt
+from tqdm import tqdm
+import os
+import time
+
+
+# Compile the function to compute the likelihood
+def get_loglike_function(ydeg, marginalize_over_inclination=False):
+    t = tt.dvector()
+    return theano.function(
+        [t],
+        StarryProcess(
+            ydeg=ydeg,
+            marginalize_over_inclination=marginalize_over_inclination,
+        ).log_likelihood(t, tt.ones_like(t), 1.0),
+    )
+
+
+# Settings
+ydeg = 15
+npts = np.sort(list(set(list(np.array(np.logspace(0, 4, 20), dtype=int)))))
+ncalls = 10
+
+# Compute
+tloglike = np.zeros((2, len(npts)))
+for i, marginalize_over_inclination in enumerate([False, True]):
+    get_loglike = get_loglike_function(ydeg, marginalize_over_inclination)
+    for j in range(len(npts)):
+        tstart = time.time()
+        t = np.linspace(0, 1, npts[j])
+        for k in range(ncalls):
+            get_loglike(t)
+        tloglike[i, j] = (time.time() - tstart) / ncalls
+
+# Plot
+fig, ax = plt.subplots(1)
+ax.plot(npts, tloglike[0], label="conditional")
+ax.plot(npts, tloglike[1], label="marginal")
+
+# Asymptotic
+n = np.array([2e2, 1.0e4])
+ax.plot(n, 1.2e-10 * n ** 2.6, "k--", alpha=0.5)
+ax.annotate(
+    r"$\propto K^{2.6}$",
+    fontsize=16,
+    xy=(2e3, 1e-2),
+    xycoords="data",
+    xytext=(0, 0),
+    textcoords="offset points",
+    ha="center",
+    va="center",
+    rotation=35,
+    alpha=0.5,
+)
+
+# Appearance
+ax.margins(0, 0)
+ax.legend(fontsize=12)
+ax.set_xlabel("number of points")
+ax.set_ylabel("likelihood eval time [s]")
+ax.set_yscale("log")
+ax.set_xscale("log")
+
+# We're done
+fig.savefig(
+    os.path.abspath(__file__).replace(".py", ".pdf"),
+    bbox_inches="tight",
+    dpi=300,
+)
